### 2020-05-17 Papers/01_interf_transf/Interference Transform.tex
started paper 01 presenting the interference transform
added various explorations into a closed form estimation of max frequency in the folder python/explorations/cf

\documentclass{article}

\usepackage[margin=0.5in]{geometry}
\usepackage[utf8]{inputenc}

\title{Interference Transform: Estimating the frequency and phase of low resolution samples}
\author{Carlos Tarjano}

\begin{document}

\maketitle

\begin{abstract}

\end{abstract}

{\bf Keywords:}

\section{Introduction}

\section{Bibliographic Review}

\subsection{Sound Representation}
frequency domain, one hot encoding, spectrogram

\subsection{Computer Graphics}
Bezier, splines, some raster, image rotation

\subsection{Sound synthesis}

\section{Methodology}

\section{Results}

\section{Discussion + Conclusion}

\section{References}

\end{document}

### 2020-05-17 Python/Explorations/Draw Isolines.py

@@ -1,92 +1,0 @@
-# import plotly.express as px
-import plotly.graph_objects as go
-import numpy as np
-
-
-#######################
-n = 10
-
-X = np.arange(n)
-
-f = 3
-p = np.pi / 2
-W = np.cos(p + 2 * np.pi * f * X / n)
-#######################
-
-fig = go.Figure()
-
-fig.update_xaxes(tickvals=[0, np.pi / 2, np.pi, 3 * np.pi / 2, 2 * np.pi])
-fig.update_yaxes(tickvals=[i for i in range(n)])
-
-fig.update_layout(
-    # width = 2 * np.pi * 220,
-    # height = n * 220,
-    # yaxis = dict(scaleanchor = "x", scaleratio = 1),
-    title=f"n = {n}",
-    xaxis_title="Phase",
-    yaxis_title="Frequency",
-    font=dict(
-        family="Courier New, monospace",
-        size=18,
-        color="#7f7f7f"
-    )
-)
-px = -2*np.pi*(n - 2)
-res = 20
-dp = (2 * np.pi - px) / res
-sum_d = np.zeros(res)
-for t in range(1,n):
-  # f = n
-  
-  fy = -n*p/(2*np.pi*t) + n/t
-
-  for i in range(res):
-    sum_d[i] += (-n*(px + i * dp)/(2*np.pi*t) + n/t - f)**2 * W[t]/n
-
-  fig.add_trace(
-    go.Scatter(
-      x=[2*np.pi,px], 
-      y=[0,fy],
-      hoverinfo=f"all",
-      name=f"t={t}, a={W[t]:.2f}", 
-      mode='lines',
-      line=go.scatter.Line(
-        width=max(1, W[t]**2 * 10),
-        dash="solid" if W[t]>0 else "dash"
-        )
-    )
-  )
-
-Xs = np.arange(px, 2 * np.pi, res)
-fig.add_trace(
-  go.Scatter(
-    x=Xs,
-    y=sum_d,
-    name=f"sum d",
-    mode='lines',
-      line=go.scatter.Line(
-        width=2,
-        dash="solid",
-        color="red"
-      )
-  )
-)
-
-fig.add_trace(
-  go.Scatter(
-    x=[p], 
-    y=[f],
-    name=f"max", 
-    mode='markers',
-    marker=dict(
-        size=8,
-        color="black", #set color equal to a variable
-        showscale=False
-    )
-  )
-)
-
-fig.show(config=dict({'scrollZoom': False}))
-
-
-

### 2020-05-17 Python/Explorations/cf/2Lines.py

@@ -1,0 +1,235 @@
+import plotly.graph_objects as go
+import numpy as np
+
+r = np.array([[0, -1], [1, 0]])  # rotation matrix
+
+p1 = np.array([[4], [2]])        # point in l1
+v1 = np.array([[1], [3]])        # vector perpendicular to l1
+n1 = v1 / np.sqrt(v1.T @ v1)     # unit vector perpendicular to l1
+u1 = r @ n1                      # unit vector in the direction of l1
+
+p2 = np.array([[8], [-1]])        # point in l2
+v2 = np.array([[1], [2]])        # vector perpendicular to l2
+n2 = v2 / np.sqrt(v2.T @ v2)     # unit vector perpendicular to l2
+u2 = r @ n2                      # unit vector in the direction of l2
+
+xi = -9
+xf = 9
+l1i = (xi - p1[0, 0]) / u1[0, 0]
+l1f = (xf - p1[0, 0]) / u1[0, 0]
+y1i = (p1 + l1i * u1)[1, 0]
+y1f = (p1 + l1f * u1)[1, 0]
+
+l2i = (xi - p2[0, 0]) / u2[0, 0]
+l2f = (xf - p2[0, 0]) / u2[0, 0]
+y2i = (p2 + l2i * u2)[1, 0]
+y2f = (p2 + l2f * u2)[1, 0]
+
+p = np.array([[4], [6]])
+
+d1 = np.sqrt((p - p1).T @ n1 @ n1.T @ (p - p1))
+d2 = np.sqrt((p - p2).T @ n2 @ n2.T @ (p - p2))
+
+
+min_x = min(xi, xf)
+max_x = max(xi, xf)
+min_y = int(np.floor(min([y1i, y1f, y2i, y2f])))
+max_y = int(np.ceil(max([y1i, y1f, y2i, y2f])))
+
+fig = go.Figure()
+
+fig.update_xaxes(tickvals=[i for i in range(min_x, max_x + 1, 1)])
+fig.update_yaxes(tickvals=[i for i in range(min_y, max_y + 1, 1)])
+
+##### l1
+fig.add_trace(
+  go.Scatter(
+    name="line 1",
+    x=[xi, xf], 
+    y=[y1i, y1f],
+    line = dict(color='blue'),
+    mode='lines',
+  ))
+##### d1
+fig.add_trace(
+  go.Scatter(
+    name="d 1",
[…]

### 2020-05-17 Python/Explorations/cf/min distance to parallel lines - LP obj only.py

@@ -1,0 +1,59 @@
+import pulp
+import plotly.graph_objects as go
+import numpy as np
+
+
+#######################
+n = 5
+X = np.arange(n)
+f = 1
+p = np.pi
+W = np.cos(p + 2 * np.pi * f * X / n)
+#######################
+
+prob = pulp.LpProblem("LP", pulp.LpMinimize)
+
+x = pulp.LpVariable("p")
+y = pulp.LpVariable("f")
+
+O = [] # opt
+K = []
+for t in range(n):
+  d = (n*x + 2*np.pi*t*y)/np.sqrt(n**2 + 4*np.pi**2 * t**2)
+
+  x_vec = (2 * np.pi * n**2) / (n**2 + 4 * np.pi**2 * t**2)     # p
+  y_vec = (4 * np.pi**2 * n * t) / (n**2 + 4 * np.pi**2 * t**2) # f
+  m_vec = np.sqrt(x_vec**2 + y_vec**2)
+
+  k = pulp.LpVariable(f"k_{t}", cat=pulp.LpInteger)
+  K.append(k)
+  # prob += k <= (d / m_vec) + 0.5
+  # prob += k >= (d / m_vec) - 0.5
+  r = d - m_vec * k                                 # <|<|<|<|<|<|<|<|
+  a = pulp.LpVariable(f"a_{t}")
+  if W[t] >= 0: # residue abs value must be the smallest possible
+    prob += a >= r
+    prob += a >= -r
+  else:       # W[t] < 0 -> residue abs value must be the highest possible
+    prob += a <= r
+    prob += a <= -r
+  O.append(a)
+
+prob += pulp.lpSum([O[t] * (W[t]) for t in range(n)]) # <|<|<|<|<|<|<|<|
+
+
+prob.writeLP("LP.lp")
+
+# status = prob.solve()
+status = prob.solve(pulp.GLPK(msg = 0))
+
+print(pulp.LpStatus[status])
+
+print(f"f={round(pulp.value(y), 2)} p={round(pulp.value(x), 2)}")
+
+for t, a in enumerate(O):
+  print(f"a_{t}={pulp.value(a)}")
+
+for t, k in enumerate(K):
+  print(f"k_{t}={pulp.value(k)}")
+

### 2020-05-17 Papers/01_interf_transf/Interference Transform.tex
added "guide to publish" folder
updated guide to publish

@@
 \author{Carlos Tarjano}
+\author{Valdecy Pereira}
 
@@
 
-\begin{abstract}
-
+\begin{abstract} % 250 words
+  % background
+  The Fourier Transform, naturally due to the context and time in which it was idealized, was formulated without regard to discrete waves or computational complexity.
+  Some developments, such as the Discrete Fourier Transform and the Fast Fourier Transform algorithm updated the theory to the digital, discrete era.
+  % motivation
+  Some shortcomings were introduced with those methods, such as diminishing resolution in dealing with a small number of os samples.
+  % objective
+  In this paper we propose a transform specifically designed for the discrete case, optimizing the inference of frequency and amplitude from whatever sampled points are available.
+  % methods
+  The algorithm is based on the representation of the set of samples in a frequency versus phase space and the investigation of the interference pattern thus generated.
+Each sample defines a sinusoidal planar wave in this space, with a particular amplitude, direction and frequency.
+The determination of the point in which maximum constructive interference occurs translates to the determination of the most accurate sinusoidal description of the whole set of samples.
+  % results
+  TODO: improvement of the naive algorithm. better accuracy
+  % implications
+  TODO: an alternative to wavelet and short-time Fourier transform
 \end{abstract}

### 2020-05-17 Papers/00_guide/Paper_structure.js

@@ -1,0 +1,31 @@
+
+var details = document.getElementsByTagName("details");
+console.log(details.length);
+
+var expandButtons = document.getElementsByClassName("expand");
+console.log(expandButtons.length);
+
+var collapseButtons = document.getElementsByClassName("collapse");
+console.log(collapseButtons.length);
+
+var ea = function expand_all(){
+  for (var i = 0; i < details.length; i++) {
+    details[i].open = true;
+  };
+};
+
+var ca = function collapse_all(){
+  for (var i = 0; i < details.length; i++) {
+    details[i].open = false;
+  };
+};
+
+for (var i = 0; i < expandButtons.length; i++) {
+  expandButtons[i].addEventListener("click", ea);
+};
+
+for (var i = 0; i < collapseButtons.length; i++) {
+  collapseButtons[i].addEventListener("click", ca);
+};
+
+ea();

### 2020-05-17 Papers/01_interf_transf/Interference Transform.tex
migrating image production to plotly

@@
 \title{Interference Transform: Estimating the frequency and phase of low resolution samples}
-\author{Carlos Tarjano}
-\author{Valdecy Pereira}
+\author{
+  Carlos Tarjano
+  \and 
+  Valdecy Pereira
+  }
 
@@
   % results
-  TODO: improvement of the naive algorithm. better accuracy
+  TODO: improvement of the naïve algorithm. better accuracy
   % implications
@@
 
-{\bf Keywords:}
+{\bf Keywords:} Fast Fourier Transform, Digital Signal Processing, Wavelet Transform
 
@@
 
-\section{Bibliographic Review}
-
-\subsection{Sound Representation}
-frequency domain, one hot encoding, spectrogram
-
-\subsection{Computer Graphics}
-Bezier, splines, some raster, image rotation
-
-\subsection{Sound synthesis}
+\section{Literature Review}
 
 \section{Methodology}
+
+\subsection{Theory}
 

### 2020-05-18 Papers/01_interf_transf/Interference Transform.tex
adjusting typesseting tools
added ghostscript to compress pdf

@@
 \documentclass{article}
-
+\usepackage{amsfonts} % \mathbb
 \usepackage[margin=0.5in]{geometry}
 \usepackage[utf8]{inputenc}
+\usepackage{graphicx}
 
@@
 \subsection{Theory}
+  One can represent the discrete wave as an an array $ W[i] $ of $ n-1 $ real numbers.
+Such wave is related, as illustrated in figure \ref{fig:ContinuousVsDiscrete}, to an hypothetical continuous wave $ S(t) $ of which its components are sampled in uniform time intervals at a rate of $\textrm{fps}$ frames per second.
+
+  \begin{figure}[h!]
+    \centering
+      \includegraphics[width=0.8\linewidth]{images/01ContinuousVsDiscrete}
+    \caption{Discrete samples (black points) from a continuous wave (grey line). Note that the amplitude is also truncated}
+    \label{fig:ContinuousVsDiscrete}
+  \end{figure}
+
+  Thus, we have:
+  $ W[i] \in \mathbb{R} \forall i \in \mathbb{N}, 0 \le i < n $,
+  $ S(t) \in \mathbb{R} \forall t \in \mathbb{R}, 0 \le t < t_f $ where $ t_f=n / \textrm{fps} $
 

### 2020-05-18 Papers/01_interf_transf/images/01ContinuousVsDiscrete copy.py

@@ -78,7 +78,7 @@ fig.add_trace(go.Scatter(name= "Discrete Wave",
 
 
 fig.show()
-fig.write_image("01.png")
+fig.write_image("./01ContinuousVsDiscrete.png")
 exit()
 
 # X = np.linspace(0, n, n * 2)

### 2020-05-18 Papers/01_interf_transf/images/01ContinuousVsDiscrete.py

@@ -1,0 +1,62 @@
+import numpy as np
+import random
+import plotly.graph_objects as go
+
+
+random.seed(0)
+n = 1000+1
+
+X = np.arange(n)
+Y = np.zeros(n)
+
+for _ in range(100):
+  a = random.uniform(1, 5)
+  f = random.uniform(1, 10)
+  p = random.uniform(0, 2 * np.pi)
+  Y += a * np.cos(p + 2 * np.pi * f * X / n)
+
+Y = Y / np.max(np.abs(Y))
+
+X_lowres = X[::20]
+Y_lowres = np.round(Y[::20], 1)
+
+fig = go.Figure()
+
+fig.layout.template ="plotly_white"
+
+fig.update_layout(
+    # title="Continuous x Discrete Wave",
+    xaxis_title="Time",
+    yaxis_title="Amplitude")
+
+fig.update_yaxes(tickvals=np.linspace(-1, 1, 21), zerolinewidth=2, zerolinecolor='black')
+
+fig.update_layout(legend_orientation="h")
+
+fig.add_trace(go.Scatter(name= "Continuous Wave",
+  x=X,
+  y=Y,
+  mode='lines',
+  line_shape='spline',
+  line=go.scatter.Line(color="gray", width=2)
+  ))
+
+fig.add_trace(go.Scatter(name= "Discrete Wave Stem",
+  showlegend=False,
+  x=[i for x in X_lowres for i in (x, x, None)],
+  y=[i for y in Y_lowres for i in (0, y, None)],
+  mode='lines',
+  line=go.scatter.Line(color="black", width=1)
+  ))
+
+fig.add_trace(go.Scatter(name= "Discrete Wave",
+  x=X_lowres,
+  y=Y_lowres,
+  mode='markers',
+  marker=dict(
+      size=5,
+      color="black",
+  )))
+
+# fig.show()
+fig.write_image("./01ContinuousVsDiscrete.eps", width=800, height=400, scale=10)

### 2020-05-18 Papers/01_interf_transf/Interference Transform.tex
fixed formatting issues with plotly

@@
     \centering
-      \includegraphics[width=0.8\linewidth]{images/01ContinuousVsDiscrete}
+      \includegraphics[width=0.8\linewidth]{images/01ContinuousVsDiscrete.eps}
     \caption{Discrete samples (black points) from a continuous wave (grey line). Note that the amplitude is also truncated}
@@
   Thus, we have:
-  $ W[i] \in \mathbb{R} \forall i \in \mathbb{N}, 0 \le i < n $,
-  $ S(t) \in \mathbb{R} \forall t \in \mathbb{R}, 0 \le t < t_f $ where $ t_f=n / \textrm{fps} $
+  $ W[i] \in \mathbb{R} \  \forall \  i \in \mathbb{N}, 0 \le i < n $
+  $ S(t) \in \mathbb{R} \  \forall \  t \in \mathbb{R}, 0 \le t < t_f $ where $ t_f=\frac{n}{\textrm{fps}} $.
+
+  The relationship is better illustrated comparing two equivalent simple sinusoids, being one continuous and its discrete counterpart. Those can be described, respectively as 
+  $$ S(t) = a \cos \left( \phi + 2 \pi \mathcal{F} t \right) $$
+  and
+  $$ W[i] = a \cos \left( \phi + 2 \pi f \frac{i}{n} \right) $$
+  where:
+
+  $ \mathcal{F} \in \mathbb{R}^+ $, absolute frequency of the continuous wave in Hz;
+
+  $ f \in \mathbb{R}^+ $, relative frequency of the discrete wave in cycles per $n$ samples;
+
+  $ f = \frac{n}{\textrm{fps}} \mathcal{F} $;
+
+  $ i = t \  \textrm{fps}, i \in \mathbb{N} $, frame number;
+
+  $ t \in \mathbb{R} $, time in seconds;
+
+  $ n \in \mathbb{N} $, number os discrete samples;
+
+  $ 0 \le \frac{i}{n} < 1 $;
+
+  $ a \in \mathbb{R}^+ $, amplitude;
+
+  $ \phi \in \mathbb{R}^+, 0 \le \phi < 2 \pi $, phase;
+
+  \begin{figure}[h!]
+    \centering
+      \includegraphics[width=0.8\linewidth]{images/01ContinuousVsDiscreteSinusoid}
+    \caption{Continuous sinusoid (grey line) and the correspondent Discrete sinusoid (black points) illustrating the correspondence between frames i and time t}
+    \label{fig:ContinuousVsDiscreteSinusoid}
+  \end{figure}
+
+  \begin{figure}[h!]
+    \centering
+      \includegraphics[width=0.8\linewidth]{images/03OptSinusoids}
+    \caption{Sinusoids with maximum coinciding with sampled points i=3 and i=28 }
+    \label{fig:OptSinusoids}
+  \end{figure}
 

### 2020-05-18 Papers/01_interf_transf/images/01ContinuousVsDiscrete copy.py

@@ -1,104 +1,0 @@
-import numpy as np
-import random
-import plotly.graph_objects as go
-
-
-random.seed(0)
-n = 1000+1
-
-X = np.arange(n)
-Y = np.zeros(n)
-
-for _ in range(100):
-  a = random.uniform(1, 5)
-  f = random.uniform(1, 10)
-  p = random.uniform(0, 2 * np.pi)
-  Y += a * np.cos(p + 2 * np.pi * f * X / n)
-
-Y = Y / np.max(np.abs(Y))
-
-X_lowres = X[::20]
-Y_lowres = np.round(Y[::20], 1)
-
-fig = go.Figure()
-
-fig.layout.template ="plotly_white"
-
-fig.update_layout(
-    title="Continuous x Discrete Wave",
-    xaxis_title="Time",
-    yaxis_title="Amplitude",
-    # font=dict(
-    #     family="Courier New, monospace",
-    #     size=18,
-    #     color="#7f7f7f"
-    # )
-)
-
-fig.update_yaxes(tickvals=np.linspace(-1, 1, 21), zerolinewidth=2, zerolinecolor='black')
-
-fig.update_layout(legend_orientation="h")
-
-# fig.update_xaxes()
-
-# fig.update_layout(
-#     yaxis = dict(
-#         # tickmode = 'linear',
-#         # tick0 = 0.5,
-#         dtick = 0.1
-#     )
-# )
-
-fig.add_trace(go.Scatter(name= "Original Wave",
-  x=X,
-  y=Y,
-  mode='lines',
-  line_shape='spline',
-  line=go.scatter.Line(color="blue", width=2)
-  ))
-
-fig.add_trace(go.Scatter(name= "Discrete Wave Stem",
-  # legendgroup="Discrete Wave",
-  showlegend=False,
-  x=[i for x in X_lowres for i in (x, x, None)],
-  y=[i for y in Y_lowres for i in (0, y, None)],
-  mode='lines',
-  line=go.scatter.Line(color="grey")
-  ))
-
-fig.add_trace(go.Scatter(name= "Discrete Wave",
-  # legendgroup="Discrete Wave",
-  x=X_lowres,
-  y=Y_lowres,
-  mode='markers',
-  marker=dict(
-      size=5,
-      color="black",
-  )))
-
-
-fig.show()
[…]

### 2020-05-18 Papers/01_interf_transf/Interference Transform.tex
changed all images to .eps
changed images to pdf
some progress with the methodology section from first paper

@@
     \centering
-      \includegraphics[width=0.8\linewidth]{images/01ContinuousVsDiscrete.eps}
+      \includegraphics[width=0.8\linewidth]{images/01ContinuousVsDiscrete.pdf}
     \caption{Discrete samples (black points) from a continuous wave (grey line). Note that the amplitude is also truncated}
@@
 
-  The relationship is better illustrated comparing two equivalent simple sinusoids, being one continuous and its discrete counterpart. Those can be described, respectively as 
+  The relationship is better illustrated comparing two equivalent simple sinusoids; a continuous one and its discrete counterpart. Those can be described, respectively as 
   $$ S(t) = a \cos \left( \phi + 2 \pi \mathcal{F} t \right) $$
@@
 
-  $ \phi \in \mathbb{R}^+, 0 \le \phi < 2 \pi $, phase;
+  $ \phi \in \mathbb{R}^+, 0 \le \phi < 2 \pi $, phase; \\
+
+  Figure \ref{fig:ContinuousVsDiscreteSinusoid} illustrates the relationship between the two.
+The values for the amplitude of both coincide, in contrast with the sampled example, as the case here is conceptually different: both sinusoids were constructed from identical parameters.
 
@@
     \centering
-      \includegraphics[width=0.8\linewidth]{images/01ContinuousVsDiscreteSinusoid}
+      \includegraphics[width=0.8\linewidth]{images/02ContinuousVsDiscreteSinusoid.pdf}
     \caption{Continuous sinusoid (grey line) and the correspondent Discrete sinusoid (black points) illustrating the correspondence between frames i and time t}
@@
 
+  So it's easy to see for each frame of a given discrete wave $W[i]$, there exists an infinite family of continuous sinusoids that pass exactly through the point defined by $i$ and $W[i]$.
+In fact, almost any sinusoid of the form $a \cos \left( \phi + 2 \pi f \frac{i}{n} \right)$ can be forced to contain the point $(i, W[i])$, given a suitable choice of $a$, with the exception of the particular case where $W[i] \ne 0$ and $\cos \left( \phi + 2 \pi \mathcal{F} \frac{i}{\textrm{fps}} \right)$.
+Of interest to this work are those sinusoids, as shown in Figure \ref{fig:OptSinusoids}, that have one of their infinite extrema, either maximum or minimum, coinciding with the abscissa $i$ or, setting $a=W[t]$, coincident with the point $(i, W[i])$.
+
   \begin{figure}[h!]
     \centering
-      \includegraphics[width=0.8\linewidth]{images/03OptSinusoids}
+      \includegraphics[width=0.8\linewidth]{images/03OptSinusoids.pdf}
     \caption{Sinusoids with maximum coinciding with sampled points i=3 and i=28 }
@@
   \end{figure}
+
+  One can define, for each frame, those waves mathematically as the family of continuous sinusoids that satisfy the equality $\cos \left( \phi + 2 \pi \mathcal{F} t \right) = \textrm{sgn}(W[t])$ or, in terms of discrete sinusoids, $\cos \left( \phi + 2 \pi f \frac{i}{n} \right) = \textrm{sgn}(W[t])$ where the sign function is defined as
+  $ \textrm{sgn}(x) = \{ -1 \  \textrm{if} \ x < 0, \  0  \  \textrm{if} \ x = 0, \  1  \  \textrm{if} \ x > 0\}$.
 

### 2020-05-18 Papers/01_interf_transf/images/02ContinuousVsDiscreteSinusoid.py

@@ -23,19 +23,23 @@ fig = go.Figure()
 fig.layout.template ="plotly_white"
 
 fig.update_layout(
-    xaxis_title="Frame number, Time (seconds)",
-    yaxis_title="Amplitude",
-    legend=dict(orientation='h', yanchor='top', xanchor='left', y=1.1),
-    margin=dict(l=5, r=5, b=5, t=5)
-)
+  xaxis_title="Frame number, Time (seconds)",
+  yaxis_title="Amplitude",
+  legend=dict(orientation='h', yanchor='top', xanchor='left', y=1.1),
+  margin=dict(l=5, r=5, b=5, t=5),
+  font=dict(family="Computer Modern", color="black", size=18)
+  )
 
 stp = 5
+
 fig.update_xaxes(
-    ticktext=[f"i={i}, t={round(i/fps, 3)}" for i in X[::stp]],
-    tickvals=X[::stp] / fps
+  gridwidth=1, 
+  gridcolor='silver',
+  ticktext=[f"i={i}, t={round(i/fps, 3)}" for i in X[::stp]],
+  tickvals=X[::stp] / fps
 )
 
-fig.update_yaxes(zerolinewidth=2, zerolinecolor='black', automargin=True)
+fig.update_yaxes(zerolinewidth=2, zerolinecolor='silver', automargin=True)
 
 fig.add_trace(go.Scatter(name=f"Continuous Wave",
   x=T,
@@ -50,20 +54,8 @@ fig.add_trace(go.Scatter(name=f"Discrete Wave",
   y=W,
   mode='lines+markers',
   # line_shape='spline',
-  line=go.scatter.Line(color="black", width=2)
+  line=go.scatter.Line(color="black", width=1)
   ))
 
-# fig.add_trace(go.Scatter(name= "Discrete Wave",
-#   # legendgroup="Discrete Wave",
-#   x=X_lowres,
-#   y=Y_lowres,
-#   mode='markers',
-#   marker=dict(
-#       size=5,
-#       color="black",
-#   )))
-
-
 # fig.show()
-fig.write_image("./02ContinuousVsDiscreteSinusoid.eps", width=800, height=400, scale=10)
-fig.write_image("./02ContinuousVsDiscreteSinusoid.png", width=800, height=400, scale=10)
+fig.write_image("./02ContinuousVsDiscreteSinusoid.pdf", width=800, height=400, scale=1)

### 2020-05-18 Papers/01_interf_transf/images/03OptSinusoids.py

@@ -2,11 +2,10 @@ import numpy as np
 import random
 import plotly.graph_objects as go
 
-
-random.seed(0)
+random.seed(1)
 
 fps = 44100
-n = 35+1
+n = 20
 
 X = np.arange(n)
 W = np.zeros(n)
@@ -20,61 +19,63 @@ for _ in range(100):
 W = W / np.max(np.abs(W))
 
 
-
 fig = go.Figure()
 
 fig.layout.template ="plotly_white"
- 
+
 fig.update_layout(
     xaxis_title="Frame",
     yaxis_title="Amplitude",
     legend=dict(orientation='h', yanchor='top', xanchor='left', y=1.1),
-    margin=dict(l=5, r=5, b=5, t=5))
+    margin=dict(l=5, r=5, b=5, t=5),
+    font=dict(family="Computer Modern", color="black", size=18)
+    )
 
+fig.update_xaxes(gridwidth=1, gridcolor='silver')
 
-fig.update_yaxes(zerolinewidth=2, zerolinecolor='black')
+fig.update_yaxes(zerolinewidth=2, zerolinecolor='silver')
 
-
+### family 01
 t = 3
 a = W[t]
 P = np.linspace(0, 2 * np.pi, 25)
 F = n * 1 / t - P * n / (2 * np.pi * t)
 
 for i in range(P.shape[0]):
-  T = np.linspace(0, 12.5, 100)
+  T = np.linspace(0, n, 100)
   S = a * np.cos(2 * np.pi * F[i] * T / n + P[i])
 
   fig.add_trace(go.Scatter(name=f"t={t}, f={F[i]}, p={P[i]}",
+  opacity=0.5,
   showlegend=False,
   x=T,
   y=S,
   mode='lines',
   line_shape='spline',
-  line=go.scatter.Line(color="gray", width=1)
+  line=go.scatter.Line(color="blue", width=1)
   ))
 
-
-
-t = 28
+### family 02
+t = 15
 a = W[t]
 P = np.linspace(0, 2 * np.pi, 25)
 F = n * 1 / t - P * n / (2 * np.pi * t)
 
 for i in range(P.shape[0]):
-  T = np.linspace(18.5, n, 100)
+  T = np.linspace(0, n, 100)
   S = a * np.cos(2 * np.pi * F[i] * T / n + P[i])
 
   fig.add_trace(go.Scatter(name=f"t={t}, f={F[i]}, p={P[i]}",
+  opacity=0.5,
   showlegend=False,
   x=T,
   y=S,
[…]

### 2020-05-23 Papers/01_interf_transf/Interference Transform.tex
web technologies version implemented

@@
 
-  So it's easy to see for each frame of a given discrete wave $W[i]$, there exists an infinite family of continuous sinusoids that pass exactly through the point defined by $i$ and $W[i]$.
+  So it's easy to see that, for each frame of a given discrete wave $W[i]$, there exists an infinite family of continuous sinusoids that pass exactly through the point defined by $i$ and $W[i]$.
 In fact, almost any sinusoid of the form $a \cos \left( \phi + 2 \pi f \frac{i}{n} \right)$ can be forced to contain the point $(i, W[i])$, given a suitable choice of $a$, with the exception of the particular case where $W[i] \ne 0$ and $\cos \left( \phi + 2 \pi \mathcal{F} \frac{i}{\textrm{fps}} \right)$.

### 2020-05-23 Web/InterferenceTransform.js

@@ -1,9 +1,3 @@
-document.getElementById("myBtn").onclick = function() {Update()};
-
-
-var RandomWaveDiv = document.getElementById("RandomWave");
-var IsolinesDiv = document.getElementById("Isolines");
-
 var n = 10;
 var X = [...Array(n).keys()];
 var W = Array(n);
@@ -11,8 +5,6 @@ var Wvis = Array(n).fill(true);
 var RandomWaveData;
 
 var Colors = Plotly.d3.scale.linear().domain([0, n-1]).range(["blue", "red"]);
-
-
 
 function RandomWave() {
   W.fill(0.0);
@@ -49,19 +41,174 @@ var RandomWaveLayout = {
     family: "Computer Modern",
     size: 18,
     color: "black"
+  },
+  yaxis:{
+    range:[-1.1, 1.1]
   }
 };
  
-Plotly.plot(RandomWaveDiv, RandomWaveData, RandomWaveLayout);
+Plotly.plot("RandomWave", RandomWaveData, RandomWaveLayout);
+
+
+
+
+// Isolines
+var LinesData = []
+for (var t = 0; t < n; t++) {
+  var vx = (2 * Math.PI * n ** 2) / (n ** 2 + 4 * Math.PI ** 2 * t ** 2)
+  var vy = (4 * Math.PI ** 2 * n * t) / (n ** 2 + 4 * Math.PI ** 2 * t ** 2)
+  var Xl = []
+  var Yl = []
+  Xl.push(0)
+  Xl.push(vx)
+  Xl.push(NaN)
+  Yl.push(0)
+  Yl.push(vy)
+  Yl.push(NaN)
+  if (t == 0) {
+    Xl.push(0)
+    Xl.push(0)
+    Xl.push(NaN)
+    Yl.push(0)
+    Yl.push(n)
+    Yl.push(NaN)
+    Xl.push(2 * Math.PI)
+    Xl.push(2 * Math.PI)
+    Xl.push(NaN)
+    Yl.push(0)
+    Yl.push(n)
+    Yl.push(NaN)
+  } else {
+    for (var k = 0; k <= t; k++) {
+      Xl.push(0)
+      Xl.push(2 * Math.PI)
+      Xl.push(NaN)
+      Yl.push(k * n / t)
+      Yl.push((k - 1) * n / t)
+      Yl.push(NaN)
+    }
+  }
+  LinesData.push(
+    {
+      name: "",
+      x: Xl,
[…]

### 2020-05-31 Papers/01_interf_transf/Interference Transform.tex
updated paper and guide

@@
   % motivation
-  Some shortcomings were introduced with those methods, such as diminishing resolution in dealing with a small number of os samples.
+  Some shortcomings were introduced with those methods, such as diminishing resolution in dealing with increasingly small number of os samples.
   % objective
-  In this paper we propose a transform specifically designed for the discrete case, optimizing the inference of frequency and amplitude from whatever sampled points are available.
+  In this paper we propose a formal framework based on discrete waves, optimizing the inference of frequency and amplitude from whatever sampled points are available.
+Although arising from a intuitively different approach, the framework here presented can be viewed as a generalization of the Discrete Fourier Transform.
   % methods
@@
 
-\section{Literature Review}
+Despite the ubiquitous presence of the DFT algorithm, some of its shortcomings, specially with respect to low resolution waves, are yet to be addressed.
+Nevertheless, areas such as audio compression would greatly benefit from improved analysis of the characteristics of low resolution samples, enabling, for example, the extraction of parameters to be used in machine learning algorithms or sound restoration.
+
+This neglect is felt in the virtual inexistence of research in this specific topic.
+
+\subsection{Literature Review}
 
@@
   Figure \ref{fig:ContinuousVsDiscreteSinusoid} illustrates the relationship between the two.
-The values for the amplitude of both coincide, in contrast with the sampled example, as the case here is conceptually different: both sinusoids were constructed from identical parameters.
 
@@
 In fact, almost any sinusoid of the form $a \cos \left( \phi + 2 \pi f \frac{i}{n} \right)$ can be forced to contain the point $(i, W[i])$, given a suitable choice of $a$, with the exception of the particular case where $W[i] \ne 0$ and $\cos \left( \phi + 2 \pi \mathcal{F} \frac{i}{\textrm{fps}} \right)$.
+
   Of interest to this work are those sinusoids, as shown in Figure \ref{fig:OptSinusoids}, that have one of their infinite extrema, either maximum or minimum, coinciding with the abscissa $i$ or, setting $a=W[t]$, coincident with the point $(i, W[i])$.
@@
       \includegraphics[width=0.8\linewidth]{images/03OptSinusoids.pdf}
-    \caption{Sinusoids with maximum coinciding with sampled points i=3 and i=28 }
+    \caption{Sinusoids with maximum coinciding with frames i=3 and i=28 }
     \label{fig:OptSinusoids}

### 2020-06-29 Papers/01_Cycles/Interference Transform.tex
started new paper

--- 2020-05-31 Papers/01_interf_transf/Interference Transform.tex
@@
   % motivation
-  Some shortcomings were introduced with those methods, such as diminishing resolution in dealing with increasingly small number of os samples.
+  A meaningful,
   % objective
@@
   % results
-  TODO: improvement of the naïve algorithm. better accuracy
+  % TODO: improvement of the naïve algorithm. better accuracy
   % implications
-  TODO: an alternative to wavelet and short-time Fourier transform
+  % TODO: an alternative to wavelet and short-time Fourier transform
 \end{abstract}
 
-{\bf Keywords:} Fast Fourier Transform, Digital Signal Processing, Wavelet Transform
+{\bf Keywords:} Signal Representation, Audio Compression,
 
 \section{Introduction}
-
-Despite the ubiquitous presence of the DFT algorithm, some of its shortcomings, specially with respect to low resolution waves, are yet to be addressed.
-Nevertheless, areas such as audio compression would greatly benefit from improved analysis of the characteristics of low resolution samples, enabling, for example, the extraction of parameters to be used in machine learning algorithms or sound restoration.
-
-This neglect is felt in the virtual inexistence of research in this specific topic.
 
@@
 \subsection{Theory}
-  One can represent the discrete wave as an an array $ W[i] $ of $ n-1 $ real numbers.
-Such wave is related, as illustrated in figure \ref{fig:ContinuousVsDiscrete}, to an hypothetical continuous wave $ S(t) $ of which its components are sampled in uniform time intervals at a rate of $\textrm{fps}$ frames per second.
 
@@
 
-  Thus, we have:
-  $ W[i] \in \mathbb{R} \  \forall \  i \in \mathbb{N}, 0 \le i < n $
-  $ S(t) \in \mathbb{R} \  \forall \  t \in \mathbb{R}, 0 \le t < t_f $ where $ t_f=\frac{n}{\textrm{fps}} $.
-
-  The relationship is better illustrated comparing two equivalent simple sinusoids; a continuous one and its discrete counterpart. Those can be described, respectively as 
-  $$ S(t) = a \cos \left( \phi + 2 \pi \mathcal{F} t \right) $$
-  and
-  $$ W[i] = a \cos \left( \phi + 2 \pi f \frac{i}{n} \right) $$
-  where:
-
-  $ \mathcal{F} \in \mathbb{R}^+ $, absolute frequency of the continuous wave in Hz;
-
-  $ f \in \mathbb{R}^+ $, relative frequency of the discrete wave in cycles per $n$ samples;
-
-  $ f = \frac{n}{\textrm{fps}} \mathcal{F} $;
-
-  $ i = t \  \textrm{fps}, i \in \mathbb{N} $, frame number;
-
-  $ t \in \mathbb{R} $, time in seconds;
-
-  $ n \in \mathbb{N} $, number os discrete samples;
-
-  $ 0 \le \frac{i}{n} < 1 $;
-
-  $ a \in \mathbb{R}^+ $, amplitude;
-
-  $ \phi \in \mathbb{R}^+, 0 \le \phi < 2 \pi $, phase; \\
-
-  Figure \ref{fig:ContinuousVsDiscreteSinusoid} illustrates the relationship between the two.
-
-  \begin{figure}[h!]
-    \centering
-      \includegraphics[width=0.8\linewidth]{images/02ContinuousVsDiscreteSinusoid.pdf}
-    \caption{Continuous sinusoid (grey line) and the correspondent Discrete sinusoid (black points) illustrating the correspondence between frames i and time t}
-    \label{fig:ContinuousVsDiscreteSinusoid}
-  \end{figure}
-
-  So it's easy to see that, for each frame of a given discrete wave $W[i]$, there exists an infinite family of continuous sinusoids that pass exactly through the point defined by $i$ and $W[i]$.
-In fact, almost any sinusoid of the form $a \cos \left( \phi + 2 \pi f \frac{i}{n} \right)$ can be forced to contain the point $(i, W[i])$, given a suitable choice of $a$, with the exception of the particular case where $W[i] \ne 0$ and $\cos \left( \phi + 2 \pi \mathcal{F} \frac{i}{\textrm{fps}} \right)$.
-
-  Of interest to this work are those sinusoids, as shown in Figure \ref{fig:OptSinusoids}, that have one of their infinite extrema, either maximum or minimum, coinciding with the abscissa $i$ or, setting $a=W[t]$, coincident with the point $(i, W[i])$.
-
-  \begin{figure}[h!]
-    \centering
-      \includegraphics[width=0.8\linewidth]{images/03OptSinusoids.pdf}
-    \caption{Sinusoids with maximum coinciding with frames i=3 and i=28 }
-    \label{fig:OptSinusoids}
-  \end{figure}
-
-  One can define, for each frame, those waves mathematically as the family of continuous sinusoids that satisfy the equality $\cos \left( \phi + 2 \pi \mathcal{F} t \right) = \textrm{sgn}(W[t])$ or, in terms of discrete sinusoids, $\cos \left( \phi + 2 \pi f \frac{i}{n} \right) = \textrm{sgn}(W[t])$ where the sign function is defined as
-  $ \textrm{sgn}(x) = \{ -1 \  \textrm{if} \ x < 0, \  0  \  \textrm{if} \ x = 0, \  1  \  \textrm{if} \ x > 0\}$.
-
 \section{Results}

### 2020-06-29 Papers/01_Cycles/images/00 delete me.py

@@ -1,0 +1,290 @@
+# import plotly.express as px
+import plotly.graph_objects as go
+import numpy as np
+
+
+#######################
+n = 7
+t = 3
+res_f = 10
+res_p = 10
+min_f = 0
+max_f = n
+#######################
+
+
+
+F = np.linspace(0, max_f, res_f)
+P = np.linspace(0, 2 * np.pi, res_p)
+
+FP = np.zeros((res_f, res_p))
+
+for i in range(res_f):
+  for j in range(res_p):
+    FP[i,j] += np.cos(P[j] + 2*np.pi*F[i]*t/n)
+
+
+X, Y, Z = [], [], []
+for i in range(res_f):
+  for j in range(res_p):
+    X.append(P[j])
+    Y.append(F[i])
+    Z.append(FP[i,j])
+    
+    
+fig = go.Figure(data=go.Scatter(
+    x= X,
+    y= Y,
+    name= "Amplitudes",
+    legendgroup="Normal",
+    customdata=Z,
+    hovertemplate='phase:%{x:.3f}<br>frequency:%{y:.3f}<br>amplitude: %{customdata:.3f} ',
+    mode='markers',
+    marker=dict(
+        size=18,
+        color=Z, #set color equal to a variable
+        colorscale='Bluered', # one of plotly colorscales
+        showscale=False
+    )
+))
+
+X_lines=[]
+Y_lines=[]
+for k in range(1, t + 1):
+  f0 = n * k / t #- n * p / (2 * np.pi * t)
+  f2pi = n * k / t - n * 2 * np.pi / (2 * np.pi * t)
+  X_lines.append(0)
+  X_lines.append(2 * np.pi)
+  X_lines.append(np.nan)
+  Y_lines.append(f0)
+  Y_lines.append(f2pi)
+  Y_lines.append(np.nan)
+
+fig.add_trace(go.Scatter(x=X_lines, y=Y_lines,name=f"Amplitude=1", mode='lines',line=go.scatter.Line(color="red")))
+
+p_vec = (2 * np.pi * n**2) / (n**2 + 4 * np.pi**2 * t**2)
+f_vec = (4 * np.pi**2 * n * t) / (n**2 + 4 * np.pi**2 * t**2)
+
+fig.add_annotation(
+  ax=0, 
+  ay=0, 
+  x=p_vec, 
+  y=f_vec, 
+  showarrow=True, 
+  text = "",
+  xref="x",
[…]

### 2020-07-16 Papers/01_paused/02_MILP/Interference Transform.tex
moved old papers to "paused"

--- 2020-05-31 Papers/01_interf_transf/Interference Transform.tex
@@
   % background
-  The Fourier Transform, naturally due to the context and time in which it was idealized, was formulated without regard to discrete waves or computational complexity.
-  Some developments, such as the Discrete Fourier Transform and the Fast Fourier Transform algorithm updated the theory to the digital, discrete era.
+  TODO low res samples, interest in best sinusoidal approximation
   % motivation
-  Some shortcomings were introduced with those methods, such as diminishing resolution in dealing with increasingly small number of os samples.
+  The Discrete Fourier Transform decomposes an periodic wave in a finite set of predefined sinusoids, commonly with frequencies coinciding with the natural numbers, offering no guarantees that one of those frequencies is the best one, in the sense of the frequency that generates the closest sinusoid to the original wave.
   % objective
-  In this paper we propose a formal framework based on discrete waves, optimizing the inference of frequency and amplitude from whatever sampled points are available.
-Although arising from a intuitively different approach, the framework here presented can be viewed as a generalization of the Discrete Fourier Transform.
+  Starting from the observation of the intersection of plane waves in the frequency versus phase space we propose a mixed-integer programming formulation of the problem.
   % methods
-  The algorithm is based on the representation of the set of samples in a frequency versus phase space and the investigation of the interference pattern thus generated.
-Each sample defines a sinusoidal planar wave in this space, with a particular amplitude, direction and frequency.
-The determination of the point in which maximum constructive interference occurs translates to the determination of the most accurate sinusoidal description of the whole set of samples.
+  The algorithm is based on the representation of the set of samples in a frequency versus phase space and the observation of the interference pattern thus generated.
+Each sample defines a sinusoidal plane wave in this space, with a particular amplitude, direction and frequency.
+The determination of the point in which maximum constructive interference occurs translates to the determination of the closest point to the lines that represent the maxima of each plane wave.
   % results
-  TODO: improvement of the naïve algorithm. better accuracy
+  TODO: better accuracy than FFT, more time consuming
   % implications
@@
 
-  Of interest to this work are those sinusoids, as shown in Figure \ref{fig:OptSinusoids}, that have one of their infinite extrema, either maximum or minimum, coinciding with the abscissa $i$ or, setting $a=W[t]$, coincident with the point $(i, W[i])$.
+  Of interest to this work are those sinusoids, as shown in Figure \ref{fig:OptSinusoids}, that have one of their infinite extrema, either maximum or minimum, coinciding with the abscissa $i$ or, setting $a=W[t]$, coincident with the point $(i, W[i])$; those sinusoids represent the combinations of phase an frequency that would maximize the value of $W[t] S[t]$ for frame $t$.
 
@@
   $ \textrm{sgn}(x) = \{ -1 \  \textrm{if} \ x < 0, \  0  \  \textrm{if} \ x = 0, \  1  \  \textrm{if} \ x > 0\}$.
+  TODO We can plot
 

### 2020-07-16 Papers/01_interf_transf/images/01ContinuousVsDiscrete.py

@@ -1,73 +1,0 @@
-import numpy as np
-import random
-import plotly.graph_objects as go
-
-
-random.seed(0)
-
-fps = 44100
-n = 1000+1
-
-X = np.arange(n) / fps
-Y = np.zeros(n)
-
-for _ in range(100):
-  a = random.uniform(1, 5)
-  f = random.uniform(1, 10)
-  p = random.uniform(0, 2 * np.pi)
-  Y += a * np.cos(p + 2 * np.pi * f * X * fps / n)
-
-Y = Y / np.max(np.abs(Y))
-
-X_lowres = X[::20]
-Y_lowres = np.round(Y[::20], 1)
-
-fig = go.Figure()
-
-fig.layout.template ="plotly_white"
- 
-fig.update_layout(
-    xaxis_title="Time (seconds)",
-    yaxis_title="Amplitude",
-    legend=dict(orientation='h', yanchor='top', xanchor='left', y=1.1),
-    margin=dict(l=5, r=5, b=5, t=5),
-    font=dict(
-    family="Computer Modern",
-    color="black",
-    size=18
-    ))
-
-fig.update_xaxes(
-  gridwidth=1, 
-  gridcolor='silver'
-)
-
-fig.update_yaxes(tickvals=np.linspace(-1, 1, 11), zerolinewidth=2, zerolinecolor='gray')
-
-fig.add_trace(go.Scatter(name= "Continuous Wave",
-  x=X,
-  y=Y,
-  mode='lines',
-  line_shape='spline',
-  line=go.scatter.Line(color="gray", width=2)
-  ))
-
-fig.add_trace(go.Scatter(name= "Discrete Wave Stem",
-  showlegend=False,
-  x=[i for x in X_lowres for i in (x, x, None)],
-  y=[i for y in Y_lowres for i in (0, y, None)],
-  mode='lines',
-  line=go.scatter.Line(color="black", width=1)
-  ))
-
-fig.add_trace(go.Scatter(name= "Discrete Wave",
-  x=X_lowres,
-  y=Y_lowres,
-  mode='markers',
-  marker=dict(
-      size=5,
-      color="black",
-  )))
-
-fig.show()
-# fig.write_image("./01ContinuousVsDiscrete.pdf", width=800, height=400, scale=1)

### 2020-07-31 Papers/02_Pulses_Envelope/PulsesEnvelope.tex
renamed paper

\documentclass{article}
\usepackage{amsfonts} % \mathbb
\usepackage[margin=0.5in]{geometry}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}

\usepackage{amsmath}
\DeclareMathOperator{\sign}{sign}

\title{Interference Transform: Estimating the frequency and phase of low resolution samples}
\author{
  Carlos Tarjano
  \and 
  Valdecy Pereira
  }

\begin{document}

\maketitle

\begin{abstract} % 250 words
  % background estimate parameters from the signal itself
  Despite being an elusive concept, only well defined mathematically in artificially generated waves, the envelope of a signal is essential for its complete characterization, being the primary information carrying medium in telecommunications, for example.
  the concept of an envelope is not uniformly defined across the literature and can be understood intuitively as a smooth function of the same variable as the principal wave that modulates the signal, responsible for his outer shape. It is implied in this definition that the envelope modulates the instantaneous amplitude of the underlying signal, being generally non periodic and thus better addressed by specific techniques diverse from those used in the wave it encompasses.
  % motivation
  Envelope detection techniques have applications in areas like health, sound classification and synthesis, seismology, speech recognition.
  % objective
  In this paper we propose a framework that uses information about the curvature of a signal to identify its boundary, eliminating the necessity of parameter tuning.
  % methods
  The approach here described draws inspiration from geometric algorithms, most notably the concept of alpha-shapes and concave hulls, to identify the frontier of an arbitrary signal. By defining a pulse as the atomic entity of a signal we greatly reduce the dimensionality and randomness of the problem, enabling to asses estimates about the curvature of a signal to inform in the envelope identification.
  % results

  % implications

\end{abstract}

{\bf Keywords:} DSP, alpha-shapes, envelope detection

\section{Introduction}

\subsection{Literature Review}

\section{Methodology} % 1000

We start by defining a pulse as a series of consecutive samples in a discrete wave with the same signal. More formally, let $ W[i] \in \mathbb{R} \forall i \in \mathbb{N}_0, i < n \in \mathbb{N}_0 $ be a real, discrete and finite signal indexed, without loss of generality, over a subset of the natural numbers. This definition relates closely to the concept of an array or vector in programming languages, and is used in the interest of simplicity. A pulse $ P[j], j \in \mathbb{N}_0 $ in $ W $ can then be defined as a sequence of samples indexed by $ i $ such that $ \sign(W[a-1]) \ne \sign(W[a]) = \sign(W[a+1]) = \dots = \sign(W[b-1]) = \sign(W[b]) \ne \sign(W[b+1]) \forall a \le i \le b | a,b \in \mathbb{N}_0 $; that is to say, a cluster of samples between the change of sign. For a continuous function, that would correspond to the pieces between roots, and would be equal to half the cycle of a sinusoid.

It is useful now to introduce the concept of a frontier as the set of points $ (i, W[i]) $ that lie on the smooth, continuous envelope of the signal forming, in this way, a discrete version of the envelope itself. Because we are interested in a smooth function over the whole domain of the wave, we can assume that, in the context of a pulse, this function will be very close to a straight line and, thus, won't touch any point inside the convex hull defined by the pulse. By the same token, as the change in amplitude from one pulse to the next is motivated by changes in the envelope, one can safely assume that $ \max(P[j]) \approx \max(P[j]+1) $ and approximate the convex hull of a pulse with the triangle defined by the points where it crosses the horizontal axis and it's interior point of max amplitude. In practice, this will allow a considerable dimension reduction.

To obtain an estimate of the behaviour of the envelope we observe the evolution of it's instantaneous curvature. For a continuous, twice differentiable function $ y = f(x); x, y \in \mathbb{R} $, the curvature as a function of the independent variable is defined as $c(x) = \frac{f''(x)}{(f'(x)^2 + 1)^{\frac{3}{2}}} $. As we are dealing with a discrete variable, we may use a finite differences scheme with modifications to account for the fact that, in general, the distance from pulse to pulse is variable.

\subsection{Theory}

  % \begin{figure}[h!]
  %   \centering
  %     \includegraphics[width=0.8\linewidth]{images/01ContinuousVsDiscrete.pdf}
  %   \caption{Discrete samples (black points) from a continuous wave (grey line). Note that the amplitude is also truncated}
  %   \label{fig:ContinuousVsDiscrete}
  % \end{figure}

\section{Results}

\section{Discussion + Conclusion}

\section{References}

\end{document}

### 2020-08-03 Papers/02_Pulses_Envelope/PulsesEnvelope.tex
progress in the paper

@@
 \usepackage{graphicx}
-
+\usepackage[backend=biber, style=apa]{biblatex}
 \usepackage{amsmath}
 \DeclareMathOperator{\sign}{sign}
+\bibliography{bibli}
 
@@
 \begin{abstract} % 250 words
-  % background estimate parameters from the signal itself
-  Despite being an elusive concept, only well defined mathematically in artificially generated waves, the envelope of a signal is essential for its complete characterization, being the primary information carrying medium in telecommunications, for example.
-  the concept of an envelope is not uniformly defined across the literature and can be understood intuitively as a smooth function of the same variable as the principal wave that modulates the signal, responsible for his outer shape.
-It is implied in this definition that the envelope modulates the instantaneous amplitude of the underlying signal, being generally non periodic and thus better addressed by specific techniques diverse from those used in the wave it encompasses.
+
+% estimate parameters from the signal itself
+% Dominant frequency as by-product
+% we define several entities in the process, introducing a new measure of discrete curve curvature, hoping to consolidate the terminology in the area
+
+  % background  
+  Despite being an elusive concept, only mathematically well defined in artificially generated waves, the temporal amplitude envelope of a signal is essential for its complete characterization, being the primary information carrying medium in spoken voice and telecommunications, for example.
+  intuitively, the temporal envelope can be understood as a smooth function on the same variable as the principal wave, or temporal fine structure, that modulates the signal, responsible for its outer shape.
+It is implied in this definition that the envelope modulates the instantaneous amplitude of the underlying signal, being non periodic in general and thus better addressed by specific techniques diverse from those used to analyse the wave it encompasses.
   % motivation
-  Envelope detection techniques have applications in areas like health, sound classification and synthesis, seismology, speech recognition.
+  Envelope detection techniques have applications in areas like health, sound classification and synthesis, seismology and speech recognition.
   % objective
-  In this paper we propose a framework that uses information about the curvature of a signal to identify its boundary, eliminating the necessity of parameter tuning.
+  In this paper we propose a framework that uses intrinsic characteristics of a signal to estimate its envelope, eliminating the necessity of parameter tuning.
   % methods
   The approach here described draws inspiration from geometric algorithms, most notably the concept of alpha-shapes and concave hulls, to identify the frontier of an arbitrary signal.
-By defining a pulse as the atomic entity of a signal we greatly reduce the dimensionality and randomness of the problem, enabling to asses estimates about the curvature of a signal to inform in the envelope identification.
+By defining a pulse as the atomic entity of a signal we greatly reduce the dimensionality and randomness of the problem.
   % results
@@
 
+% Opener sentence
+% Literature review | Brief Context of Prior Research
+In many applications the temporal amplitude envelope of a signal plays a prominent role in the characteristics exhibited: According to Qi et al (\citedate{2017QiRelative}), for example, the envelope is at least as important as the fine structure of the soundwave in the context of the intelligibility of mandarin tones.
+That is also the case for English, according to \cite{1995ShannonSpeech}, where even envelopes modulating mostly noise were capable of conveying meaning.
+Still in the context of speech, the envelope helps to convey emotion and identity \parencite{2018ZhuContributions}.
+
+Also, the envelope is shown to add complexity to the spectral representation of a wave \parencite{2019TarjanoNeuro}, and a good description of the envelope would be helpful for a subsequent spectral analysis.
+
+% Restate Your Question as Something Not Known or Fully Understood by Prior Research
+% State the Significance of Your Question
+% State Your Claim | Objective | Hypothesis
+% brief outlook on the structure of the paper
+
 \subsection{Literature Review}
 
-\section{Methodology} % 1000
+\section{Methodology} % 1000 words
 
+% Essential background information
 We start by defining a pulse as a series of consecutive samples in a discrete wave with the same signal.
@@
 
-It is useful now to introduce the concept of a frontier as the set of points $ (i, W[i]) $ that lie on the smooth, continuous envelope of the signal forming, in this way, a discrete version of the envelope itself.
-Because we are interested in a smooth function over the whole domain of the wave, we can assume that, in the context of a pulse, this function will be very close to a straight line and, thus, won't touch any point inside the convex hull defined by the pulse.
-By the same token, as the change in amplitude from one pulse to the next is motivated by changes in the envelope, one can safely assume that $ \max(P[j]) \approx \max(P[j]+1) $ and approximate the convex hull of a pulse with the triangle defined by the points where it crosses the horizontal axis and it's interior point of max amplitude.
-In practice, this will allow a considerable dimension reduction.
+As we are seeking a geometric treatment of the subject, it is necessary to make abscissa and ordinate units consistent.
+this can be accomplished by multiplying each sample of the signal by the average length of the pulses, after having divided by the average of signal amplitudes, putting both axes in the same time related units.
 
-To obtain an estimate of the behaviour of the envelope we observe the evolution of it's instantaneous curvature.
-For a continuous, twice differentiable function $ y = f(x); x, y \in \mathbb{R} $, the curvature as a function of the independent variable is defined as $c(x) = \frac{f''(x)}{(f'(x)^2 + 1)^{\frac{3}{2}}} $.
-As we are dealing with a discrete variable, we may use a finite differences scheme with modifications to account for the fact that, in general, the distance from pulse to pulse is variable.
+It is useful now to introduce the concept of a frontier as the set of points $ (i, W[i]) $ that bound the smooth, continuous envelope of the signal.
+Because we are interested in a smooth function over the whole domain of the wave, we can assume that, in the scale of a pulse, this function will be very close to a straight line and, thus, won't touch any point inside the convex hull defined by the pulse.
+By the same token, as the change in amplitude from one pulse to the next is motivated by changes in the envelope, one can assume that $ \max(P[j]) \approx \max(P[j+1]) $ and approximate the convex hull of a pulse with the triangle defined by the points where it crosses the horizontal axis and it's point of max amplitude; this will allow a considerable dimension reduction as this triangular pulse becomes the atomic unit for the rest of the method.
+
+For a continuous, twice differentiable function $ y = f(x); x, y \in \mathbb{R} $, the curvature as a function of the independent variable is given by $c(x) = \frac{f''(x)}{(f'(x)^2 + 1)^{\frac{3}{2}}} $ which is also the inverse of the radius of the osculating circle tangent to the curve at $ x $.
+This definition provides a welcome link between algebra and geometry, but also poses two difficulties to our particular case that need to be addressed: the fact that our signal isn't even continuous, and thus neither smooth, and that we don't have a solid unit for the amplitude.
 
@@
 \section{References}
+\printbibliography[heading=none]
 

### 2020-08-03 Papers/02_Pulses_Envelope/PulsesEnvelope.tex
some progress

@@
 
-\title{Interference Transform: Estimating the frequency and phase of low resolution samples}
+\title{Robust Envelope Estimation Via Geometric Properties of a Signal}
 \author{
@@
 
-\section{Introduction}
-
+\section{Introduction} % 500 words
 % Opener sentence
 % Literature review | Brief Context of Prior Research
-In many applications the temporal amplitude envelope of a signal plays a prominent role in the characteristics exhibited: According to Qi et al (\citedate{2017QiRelative}), for example, the envelope is at least as important as the fine structure of the soundwave in the context of the intelligibility of mandarin tones.
-That is also the case for English, according to \cite{1995ShannonSpeech}, where even envelopes modulating mostly noise were capable of conveying meaning.
-Still in the context of speech, the envelope helps to convey emotion and identity \parencite{2018ZhuContributions}.
+In many applications the temporal amplitude envelope of a signal plays a prominent role in the characteristics exhibited: According to \textcite{2017QiRelative}, for example, the envelope is at least as important as the fine structure of the soundwave in the context of the intelligibility of mandarin tones.
+That is also the case for English \parencite{1995ShannonSpeech}, where even envelopes modulating mostly noise were still capable of conveying meaning.
+The envelope helps to convey emotion and identity to the human voice \parencite{2018ZhuContributions}, and envelope preserving characteristics of concert halls are associated with their pleasantness \parencite{2011LokkiEngaging}.
 
-Also, the envelope is shown to add complexity to the spectral representation of a wave \parencite{2019TarjanoNeuro}, and a good description of the envelope would be helpful for a subsequent spectral analysis.
+% enumerate some uses in diverse areas
+
+Also, the envelope is shown to add complexity to the spectral representation of a wave \parencite{2019TarjanoNeuro}, and a accurate description of the envelope would be useful for a subsequent spectral analysis.
+In the context of sound synthesis, for example, one would be able to apply different methods for the recreation of the envelope, more in line with its smooth, slow varying and non periodic nature and an appropriate approach to the periodic and relatively fast changes of the temporal fine structure.
 
@@
 % brief outlook on the structure of the paper
-
-\subsection{Literature Review}
 

### 2020-08-04 Papers/02_Pulses_Envelope/PulsesEnvelope.tex
some progress in the paper

@@
 
-\title{Robust Envelope Estimation Via Geometric Properties of a Signal}
+\title{Robust Digital Envelope Estimation Via Geometric Properties of a Arbitrary Signal}
 \author{
@@
 % estimate parameters from the signal itself
+% broadband
 % Dominant frequency as by-product
@@
   % background  
-  Despite being an elusive concept, only mathematically well defined in artificially generated waves, the temporal amplitude envelope of a signal is essential for its complete characterization, being the primary information carrying medium in spoken voice and telecommunications, for example.
+  Despite being an elusive concept, mathematically well defined only for artificially generated waves, the temporal amplitude envelope of a signal is essential for its complete characterization, being the primary information carrying medium in spoken voice and telecommunications, for example.
   intuitively, the temporal envelope can be understood as a smooth function on the same variable as the principal wave, or temporal fine structure, that modulates the signal, responsible for its outer shape.
-It is implied in this definition that the envelope modulates the instantaneous amplitude of the underlying signal, being non periodic in general and thus better addressed by specific techniques diverse from those used to analyse the wave it encompasses.
+It is implied in this definition that the envelope is non periodic in general and is thus better addressed by specific techniques diverse from those used in the analysis of the periodic, rapidly changing wave it encompasses.
   % motivation
   Envelope detection techniques have applications in areas like health, sound classification and synthesis, seismology and speech recognition.
+Despite of that, a general approach to envelope detection of signals with rich spectral content seems to be lacking, and most methods involve manual intervention, like filter design, based on prior knowledge about the wave under investigation.
   % objective
@@
   % methods
-  The approach here described draws inspiration from geometric algorithms, most notably the concept of alpha-shapes and concave hulls, to identify the frontier of an arbitrary signal.
-By defining a pulse as the atomic entity of a signal we greatly reduce the dimensionality and randomness of the problem.
+  The approach here described draws inspiration from general and differential geometry to isolate the frontier of an arbitrary signal.
+The concepts of alpha-shapes, concave hulls, arc-length parametrization and discrete curve curvature are adapted.
+We also define a pulse in the context of a arbitrary digital signal as a means to reduce dimensionality and subsequently, the complexity of the algorithm.
   % results
@@
 % Opener sentence
+Despite being ubiquitous in digital signal processing, the literature about envelope detection is very fragmented \parencite{2017LyonsDigital}.
+Besides, most envelope detection techniques are designed to account for very specific settings, like pure sinusoids with moderate noise contents, in line with the most common usages of those algorithms with artificial signals in the context of analog telecommunications.
 % Literature review | Brief Context of Prior Research
-In many applications the temporal amplitude envelope of a signal plays a prominent role in the characteristics exhibited: According to \textcite{2017QiRelative}, for example, the envelope is at least as important as the fine structure of the soundwave in the context of the intelligibility of mandarin tones.
+In many natural signals, however, the temporal amplitude envelope of a signal plays a prominent role in the characteristics exhibited: According to \textcite{2017QiRelative}, for example, the envelope is at least as important as the fine structure of the soundwave in the context of the intelligibility of mandarin tones.
 That is also the case for English \parencite{1995ShannonSpeech}, where even envelopes modulating mostly noise were still capable of conveying meaning.
@@
 Also, the envelope is shown to add complexity to the spectral representation of a wave \parencite{2019TarjanoNeuro}, and a accurate description of the envelope would be useful for a subsequent spectral analysis.
-In the context of sound synthesis, for example, one would be able to apply different methods for the recreation of the envelope, more in line with its smooth, slow varying and non periodic nature and an appropriate approach to the periodic and relatively fast changes of the temporal fine structure.
+In the context of sound synthesis, for example, one would be able to apply different methods for the recreation of the envelope, more in line with its smooth, slow varying and non periodic nature and an different approach to the periodic and relatively fast changes of the temporal fine structure.
+
+% Cite current algorithms
 
 % Restate Your Question as Something Not Known or Fully Understood by Prior Research
+Thus, in this paper we formulate a general approach to envelope detection, examining the intrinsic characteristics of a generic, spectrally complex wave in order to avoid parameter tuning.
 % State the Significance of Your Question
+While the robustness of the proposed approach allows it to be used as a plug in replacement for many methods encountered in the literature, we feel that it would be particularly useful for sound synthesis.
 % State Your Claim | Objective | Hypothesis

### 2020-08-06 Papers/02_Pulses_Envelope/PulsesEnvelope.tex
Advances

@@
 
-\title{Robust Digital Envelope Estimation Via Geometric Properties of a Arbitrary Signal}
+\title{Robust Digital Envelope Estimation Via Geometric Properties of an Arbitrary Signal}
 \author{
@@
   % methods
-  The approach here described draws inspiration from general and differential geometry to isolate the frontier of an arbitrary signal.
+  The approach here described draws inspiration from general, computational and differential geometry to isolate the frontier of an arbitrary signal.
 The concepts of alpha-shapes, concave hulls, arc-length parametrization and discrete curve curvature are adapted.
-We also define a pulse in the context of a arbitrary digital signal as a means to reduce dimensionality and subsequently, the complexity of the algorithm.
+We also define a pulse in the context of a arbitrary digital signal as a means to reduce dimensionality and the complexity of the proposed algorithm.
   % results
@@
 The envelope helps to convey emotion and identity to the human voice \parencite{2018ZhuContributions}, and envelope preserving characteristics of concert halls are associated with their pleasantness \parencite{2011LokkiEngaging}.
+% Cite current algorithms
+When dealing with broadband signals, approaches tailored to specific applications are prevalent, such as the one presented by \textcite{2014YangFast} for the distributed monitoring of fibre optic or the one formulated by \textcite{2018AssefModeling} in the context of medical ultrasound imaging.
 
-% enumerate some uses in diverse areas
+If one adresses the different units in the horizontal and vertical axes, one can transform the DSP problem of envelope detection in the geometric problem of defining the shape of a set of points in $ \mathbb{R}^2 $.
+For that purpose \textcite{1983Edelsbrunnershape} introduced the concept of alpha-shapes, a mathematically well defined extension to the convex hull of a finite set of points, related to the Delaunay triangulation and Voronoi diagrams of those points.
 
-Also, the envelope is shown to add complexity to the spectral representation of a wave \parencite{2019TarjanoNeuro}, and a accurate description of the envelope would be useful for a subsequent spectral analysis.
-In the context of sound synthesis, for example, one would be able to apply different methods for the recreation of the envelope, more in line with its smooth, slow varying and non periodic nature and an different approach to the periodic and relatively fast changes of the temporal fine structure.
-
-% Cite current algorithms
+This approach is used in areas such as the detection of features in images \parencite{2016VarytimidisAlpha}, reconstruction of surfaces from a cloud of points \parencite{2015WuAutomated} and Spectroscopy \parencite{2019XuModeling}, with the last work, that involves the estimation and removal of the Blaze function -a kind of envelope- of a echelle spectrograph, being particularly align with what is intended here.
 
 % Restate Your Question as Something Not Known or Fully Understood by Prior Research
-Thus, in this paper we formulate a general approach to envelope detection, examining the intrinsic characteristics of a generic, spectrally complex wave in order to avoid parameter tuning.
+Thus, in this paper we formulate a general approach to envelope detection, exploiting the intrinsic characteristics of a generic, spectrally complex, wave in order to avoid the need to manual intervention or parameter tuning.
 % State the Significance of Your Question
 While the robustness of the proposed approach allows it to be used as a plug in replacement for many methods encountered in the literature, we feel that it would be particularly useful for sound synthesis.
+The envelope is shown to add complexity to the spectral representation of a wave \parencite{2019TarjanoNeuro}, and a accurate description of the envelope would be useful for a cleaner spectral analysis.
+Moreover, the algorithm naturally divides a signal in pseudo-cycles, that could serve pragmatic building blocks for the reconstruction of the fine structure of the wave.
 % State Your Claim | Objective | Hypothesis
+In the context of sound synthesis, one would be able to apply specific methods for the recreation of the envelope, more in line with its smooth, slow varying and non periodic nature and a different approach to the periodic and relatively fast changes of the temporal fine structure.
 % brief outlook on the structure of the paper
@@
 
+% pulse, simplified pulse, pseudo-cycle image
+
 As we are seeking a geometric treatment of the subject, it is necessary to make abscissa and ordinate units consistent.
 this can be accomplished by multiplying each sample of the signal by the average length of the pulses, after having divided by the average of signal amplitudes, putting both axes in the same time related units.
+One can call this approach arc length regularization, as it translates the widespread approach of arc length parametrization to a non parametric setting.
 
@@
 
-For a continuous, twice differentiable function $ y = f(x); x, y \in \mathbb{R} $, the curvature as a function of the independent variable is given by $c(x) = \frac{f''(x)}{(f'(x)^2 + 1)^{\frac{3}{2}}} $ which is also the inverse of the radius of the osculating circle tangent to the curve at $ x $.
-This definition provides a welcome link between algebra and geometry, but also poses two difficulties to our particular case that need to be addressed: the fact that our signal isn't even continuous, and thus neither smooth, and that we don't have a solid unit for the amplitude.
+Besides, those definitions will allow us to derive the frontier of a wave using an alpha-shapes inspired algorithm without the need to compute the Delaunay triangulation first.
+Translating the intuitive explanation of alpha-shapes in \textcite{1994EdelsbrunnerThree}, we can say that the points in the frontier are those touched by a circle outside the signal that is not allowed to contain any point of the signal; one can picture a circle being rolled above (or below, in the case of the negative envelope) the signal, and marking the points it touches as frontier points.
 
-\subsection{Theory}
+% snowball image
+
+It is now necessary to infer the appropriate radius of such a circle and, to that end, a measure of the instantaneous curvature of a discrete function is needed.
+Discrete curvature estimation is an important task in image processing \parencite{2010FleischmannNovel} for which no default definition exists; The two possible approaches are the derivation of direct methods, or the calculation of the curvature of a fitted curve \parencite{2001CoeurjollyDiscrete}.
+As we are interested in the average curvature, the second approach is more natural, since it enables us to calculate the average curvature via the integral of the curvature in the fitted curve.
+
+% osculating circle image, fitted line
+
+For a continuous, twice differentiable function $ y = f(x); x, y \in \mathbb{R} $, the curvature as a function of the independent variable is given by $ \kappa(x) = \frac{f''(x)}{(f'(x)^2 + 1)^{\frac{3}{2}}} $ which is also the inverse of the radius of the osculating circle tangent to the curve at $ x $.
+Since we are assuming smoothness of envelope in the neighbourhood of a pulse, it makes sense to approximate it with a parabola $ y(x)=ax^2+bx+c $, since a higher order polynomial would provide unnecessary complexity.
+Besides some characteristics of the parabola derived from its relation with Bezier curves will be useful later.
+
+Therefore, the definition of the curvature becomes $ \kappa(x) = \frac{2 a}{\big( (b + 2 a x)^2 + 1 \big)^\frac{3}{2}} $ with the average curvature between points $ x_0 $ and $ x_1 $ given by equation \ref{eq:1}.
+
+\begin{equation} \label{eq:1}
+  \overline{\kappa}(x) = 
+  \frac{
+      \left( \dfrac{2 a x_1 + b}{\sqrt{\big(2 a x_1 + b \big)^2+1}} - \dfrac{2 a x_0 + b}{\sqrt{\big(2 a x_0 + b \big)^2+1}} \right) 
+  }{
+    (x_1 - x_0)
+  }
+\end{equation}
+
+% \subsection{Theory}
 

### 2020-08-07 Papers/02_Pulses_Envelope/PulsesEnvelope.tex
some progress in the paper

@@
 For a continuous function, that would correspond to the pieces between roots, and would be equal to half the cycle of a sinusoid.
+Thus, in our analogy, two consecutive, opposed pulses have the potential do delimitate a pseudo-cycle, as illustrated in figure \ref{fig:pulses}.
+For that to be true, however, the pulses need to make part of the frontier, as will be defined later.
 
-% pulse, simplified pulse, pseudo-cycle image
+  \begin{figure}[h!]
+    \centering
+      \includegraphics[width=0.8\linewidth]{01Pulses.pdf}
+    \caption{The grey areas encompass the simplified convex hulls of the pulses. The dashed lines mark the frontier between pseudo-cycles}
+    \label{fig:pulses}
+  \end{figure}
 
@@
 
-  % \begin{figure}[h!]
-  %   \centering
-  %     \includegraphics[width=0.8\linewidth]{images/01ContinuousVsDiscrete.pdf}
-  %   \caption{Discrete samples (black points) from a continuous wave (grey line). Note that the amplitude is also truncated}
-  %   \label{fig:ContinuousVsDiscrete}
-  % \end{figure}
-
 \section{Results}

### 2020-08-07 Papers/02_Pulses_Envelope/images/02Curvature.py

@@ -1,0 +1,169 @@
+# import random
+# random.seed(2)
+import sys
+sys.path.append("c:/Users/tesse/Desktop/Files/Dropbox/0_Thesis/Python")
+# import site
+# site.addsitedir("c:/Users/tesse/Desktop/Files/Dropbox/0_Thesis/Python")
+import numpy as np
+
+import plotly.graph_objects as go
+
+from Helper import signal_to_pulses, get_pulses_area, split_pulses
+
+def get_parabola_coefficients(X, Y):
+  assert len(X) == len(Y) and len(X) == 3
+  x0, x1, x2 = X[0], X[1], X[2]
+  y0, y1, y2 = Y[0], Y[1], Y[2]
+
+  den = (-x0 + x2) * (x0 - x1) * (x1 - x2)
+
+  a = ((x0 - x1) * (y1 - y2) - (x1 - x2) * (y0 - y1)) / (2 * den)
+  b = (-x0 * (x0 - x1) * (y1 - y2) + x2 * (x1 - x2) * (y0 - y1)) / den
+  c = -a * x0**2 - b * x0 + y0
+  return a, b, c
+
+def get_approximation(X, Y):
+  Y = np.abs(Y)
+  final_X = []
+  final_Y = []
+  for i in range(1, len(X) - 1):
+    x0, x1, x2 = (X[i - 1] + X[i]) / 2, X[i], (X[i] + X[i + 1]) / 2
+    y0, y1, y2 = (Y[i - 1] + Y[i]) / 2, Y[i], (Y[i] + Y[i + 1]) / 2
+    a, b, c = get_parabola_coefficients([x0, x1, x2], [y0, y1, y2])
+    XX = np.linspace(x0, x2, 10)
+    YY = a * XX**2 + b * XX + c
+    for x, y in zip(XX, YY):
+      final_X.append(x), final_Y.append(y)
+    final_X.append(None), final_Y.append(None)
+  return final_X, final_Y
+
+'''Generating Wave'''
+np.random.seed(1)
+# fps = 10
+n = 200+1
+
+X = np.arange(n)# / fps
+W = np.zeros(n)
+
+afp = [
+  [1, 8, np.pi],
+  # [1, 2, 1.2 * np.pi],
+  # [1, 2.3, 1.4 * np.pi]
+  ]
+for a, f, p in afp:
[…]

### 2020-08-17 Papers/02_Pulses_Envelope/PulsesEnvelope.tex
pulse from triangle to box
Some progress with the paper

@@
 % Essential background information
-We start by defining a pulse as a series of consecutive samples in a discrete wave with the same signal.
+We start by defining a pulse as a series of consecutive samples in a discrete wave with the same sign.
 More formally, let $ W[i] \in \mathbb{R} \forall i \in \mathbb{N}_0, i < n \in \mathbb{N}_0 $ be a real, discrete and finite signal indexed, without loss of generality, over a subset of the natural numbers.
 This definition relates closely to the concept of an array or vector in programming languages, and is used in the interest of simplicity.
-A pulse $ P[j], j \in \mathbb{N}_0 $ in $ W $ can then be defined as a sequence of samples indexed by $ i $ such that $ \sign(W[a-1]) \ne \sign(W[a]) = \sign(W[a+1]) = \dots = \sign(W[b-1]) = \sign(W[b]) \ne \sign(W[b+1]) \forall a \le i \le b | a,b \in \mathbb{N}_0 $; that is to say, a cluster of samples between the change of sign.
-For a continuous function, that would correspond to the pieces between roots, and would be equal to half the cycle of a sinusoid.
+A pulse $ P[j], j \in \mathbb{N}_0 $ in $ W $ can then be defined as a sequence of samples indexed by $ i $ such that $ \sign(W[a-1]) \ne \sign(W[a]) = \sign(W[a+1]) = \dots = \sign(W[b-1]) = \sign(W[b]) \ne \sign(W[b+1]) \forall a \le i \le b | a,b \in \mathbb{N}_0 $.
+In other words a pulse is a cluster of samples between the changes of sign in the original wave.
+For a continuous function, that would correspond to the pieces between it's roots, and would be equal to half the cycle of a sinusoid.
 Thus, in our analogy, two consecutive, opposed pulses have the potential do delimitate a pseudo-cycle, as illustrated in figure \ref{fig:pulses}.
@@
 
-As we are seeking a geometric treatment of the subject, it is necessary to make abscissa and ordinate units consistent.
-this can be accomplished by multiplying each sample of the signal by the average length of the pulses, after having divided by the average of signal amplitudes, putting both axes in the same time related units.
+As we are seeking a geometric treatment of the subject, it is first necessary to make abscissa and ordinate units consistent.
+this can be accomplished by multiplying each sample of the signal by the average length of the pulses, after having divided by the average of signal amplitudes, putting both axes in the same frame related unit $i$, that relates to time by the equation $i = t \ \mathit{fps}$.
 One can call this approach arc length regularization, as it translates the widespread approach of arc length parametrization to a non parametric setting.
 
-It is useful now to introduce the concept of a frontier as the set of points $ (i, W[i]) $ that bound the smooth, continuous envelope of the signal.
-Because we are interested in a smooth function over the whole domain of the wave, we can assume that, in the scale of a pulse, this function will be very close to a straight line and, thus, won't touch any point inside the convex hull defined by the pulse.
-By the same token, as the change in amplitude from one pulse to the next is motivated by changes in the envelope, one can assume that $ \max(P[j]) \approx \max(P[j+1]) $ and approximate the convex hull of a pulse with the triangle defined by the points where it crosses the horizontal axis and it's point of max amplitude; this will allow a considerable dimension reduction as this triangular pulse becomes the atomic unit for the rest of the method.
+% Define envelope
+
+It will be useful to introduce the concept of a frontier as the set of points $ (i, W[i]) $ that are closest to the unknown smooth, continuous envelope of the signal.
+Because we are interested in a function that is smooth over the whole domain of the wave, we can assume that, in the scale of a pulse, this function will be very close to a horizontal straight line and, thus, won't touch any point inside the convex hull defined by the pulse.
+By the same token, as the change in amplitude from one pulse to the next is motivated by changes in the temporal envelope, one can assume that $ \max(P[j]) \approx \max(P[j+1]) $ and approximate the convex hull of a pulse with the rectangle whose length is defined by the points where the pulse crosses the horizontal axis and whose height is equal to the point of max amplitude; this will allow a considerable dimension reduction as this rectangular pulse becomes the atomic unit for the rest of the method.
 
@@
 It is now necessary to infer the appropriate radius of such a circle and, to that end, a measure of the instantaneous curvature of a discrete function is needed.
-Discrete curvature estimation is an important task in image processing \parencite{2010FleischmannNovel} for which no default definition exists; The two possible approaches are the derivation of direct methods, or the calculation of the curvature of a fitted curve \parencite{2001CoeurjollyDiscrete}.
-As we are interested in the average curvature, the second approach is more natural, since it enables us to calculate the average curvature via the integral of the curvature in the fitted curve.
+Discrete curvature estimation is an important task in image processing \parencite{2010FleischmannNovel} for which no default definition exists; The two possible approaches are the derivation of direct methods that use characteristics of the discrete wave to calculate its curvature at each point, or the calculation of the curvature of a curve fitted to the discrete wave \parencite{2001CoeurjollyDiscrete}.
+To fit a polynomial to the a generic wave two approaches are readily available: the least mean squares approach, that seeks to minimize the dependant variable errors and the total least squares problem, that treat both variables symmetrically \parencite{1980GolubAnalysis}.
+The geometric, symmetrical nature of the problem excludes the more computationally economic least mean squares approach however, leaving us with the total least squares, for which no general closed form solutions are available \parencite{2007MarkovskyOverview}; this leads us to turn our attention to direct approaches of curvature estimation.
+
+\subsection{Discrete Curvature Estimation - The Equivalent Circle Approach}
+
+As previously stated, no default discrete curvature definition exists
 
@@
 
-For a continuous, twice differentiable function $ y = f(x); x, y \in \mathbb{R} $, the curvature as a function of the independent variable is given by $ \kappa(x) = \frac{f''(x)}{(f'(x)^2 + 1)^{\frac{3}{2}}} $ which is also the inverse of the radius of the osculating circle tangent to the curve at $ x $.
+% For a continuous, twice differentiable function $ y = f(x); x, y \in \mathbb{R} $, the curvature as a function of the independent variable is given by $ \kappa(x) = \frac{f''(x)}{(f'(x)^2 + 1)^{\frac{3}{2}}} $ which is also the inverse of the radius of the osculating circle tangent to the curve at $ x $.
 Since we are assuming smoothness of envelope in the neighbourhood of a pulse, it makes sense to approximate it with a parabola $ y(x)=ax^2+bx+c $, since a higher order polynomial would provide unnecessary complexity.
@@
 
-Therefore, the definition of the curvature becomes $ \kappa(x) = \frac{2 a}{\big( (b + 2 a x)^2 + 1 \big)^\frac{3}{2}} $ with the average curvature between points $ x_0 $ and $ x_1 $ given by equation \ref{eq:1}.
+% Therefore, the definition of the curvature becomes $ \kappa(x) = \frac{2 a}{\big( (b + 2 a x)^2 + 1 \big)^\frac{3}{2}} $ with the average curvature between points $ x_0 $ and $ x_1 $ given by equation \ref{eq:1}.
 
-\begin{equation} \label{eq:1}
-  \overline{\kappa}(x) = 
-  \frac{
-      \left( \dfrac{2 a x_1 + b}{\sqrt{\big(2 a x_1 + b \big)^2+1}} - \dfrac{2 a x_0 + b}{\sqrt{\big(2 a x_0 + b \big)^2+1}} \right) 
-  }{
-    (x_1 - x_0)
-  }
-\end{equation}
+% \begin{equation} \label{eq:1}
+%   \overline{\kappa}(x) = 
+%   \frac{
+%       \left( \dfrac{2 a x_1 + b}{\sqrt{\big(2 a x_1 + b \big)^2+1}} - \dfrac{2 a x_0 + b}{\sqrt{\big(2 a x_0 + b \big)^2+1}} \right) 
+%   }{
+%     (x_1 - x_0)
+%   }
+% \end{equation}
 

### 2020-08-18 Papers/02_Pulses_Envelope/PulsesEnvelope.tex
progress in the paper

@@
 More formally, let $ W[i] \in \mathbb{R} \forall i \in \mathbb{N}_0, i < n \in \mathbb{N}_0 $ be a real, discrete and finite signal indexed, without loss of generality, over a subset of the natural numbers.
-This definition relates closely to the concept of an array or vector in programming languages, and is used in the interest of simplicity.
-A pulse $ P[j], j \in \mathbb{N}_0 $ in $ W $ can then be defined as a sequence of samples indexed by $ i $ such that $ \sign(W[a-1]) \ne \sign(W[a]) = \sign(W[a+1]) = \dots = \sign(W[b-1]) = \sign(W[b]) \ne \sign(W[b+1]) \forall a \le i \le b | a,b \in \mathbb{N}_0 $.
-In other words a pulse is a cluster of samples between the changes of sign in the original wave.
-For a continuous function, that would correspond to the pieces between it's roots, and would be equal to half the cycle of a sinusoid.
-Thus, in our analogy, two consecutive, opposed pulses have the potential do delimitate a pseudo-cycle, as illustrated in figure \ref{fig:pulses}.
+This definition relates closely to the concept of an array or vector in most programming languages, and is used in the interest of simplicity.
+A pulse $ P[j], j \in \mathbb{N}_0, j < q $ in $ W $ can then be defined as a sequence of samples indexed by $ i $ such that $ \sign(W[a-1]) \ne \sign(W[a]) = \sign(W[a+1]) = \dots = \sign(W[b-1]) = \sign(W[b]) \ne \sign(W[b+1]) \forall i, a \le i \le b | a,b \in \mathbb{N}_0 $.
+In other words, each pulse is a unique subset of $ W $ with its members being the maximum sequence of samples in $ W $ with the same signal and $ P[0] \cup P[0] \cup \cdots \cup P[q-1] = W $.
+Naturally, each pulse can be classified as positive or negative, based on the signal of its samples, which must be the same.
+For a continuous function, then, a pulse would correspond to the pieces between it's roots, and would be equal to half the cycle of a sinusoid.
+Thus, in our analogy, two consecutive pulses, naturally with opposing signs, have the potential do delimitate a pseudo-cycle, as illustrated in figure \ref{fig:pulses}.
 For that to be true, however, the pulses need to make part of the frontier, as will be defined later.
@@
 
-As we are seeking a geometric treatment of the subject, it is first necessary to make abscissa and ordinate units consistent.
-this can be accomplished by multiplying each sample of the signal by the average length of the pulses, after having divided by the average of signal amplitudes, putting both axes in the same frame related unit $i$, that relates to time by the equation $i = t \ \mathit{fps}$.
+% TODO Define envelope, Hilbert, etc
+
+In practice, however, is not uncommon for a discrete wave, specially in the case of sound, to present somewhat different positive and negative contours; As our method will eventually have to be applied separately to both sides, superior and inferior, of the discrete wave, it is only natural to introduce such separation as soon as possible.
+To do so we will define the positive (upper) frontier and negative (lower) frontier as the set of points $ (i, |W[i]|) $ that are closest to the unknown smooth, continuous envelope of the signal.
+The set of all these points such that $ W[i] > 0 $ will be called the positive, piecewise linear frontier denoted by $ F_+ $; similarly, $ F_- $ denotes the negative frontier of $ W $, defined by the subset of points $ (i, W[i]) $ where $ W[i] < 0 $.
+To identify such points, we need to address the different units in which the amplitude and horizontal distance of our wave are expressed.
+
+% TODO envelope, positive and negative frontiers image
+
+% TODO clarify this part
+To make abscissa and ordinate units consistent for the computation of the frontier, and proceed with a geometric treatment of the subject, we consider the set of positive and negative pulses separately.
+For each set, we can multiply all pulses by the average length of the pulses in the set, after having divided by the average of pulses amplitudes, putting both axes in the same frame related unit $i$, that relates to time by the equation $i = t \ \mathit{fps}$.
 One can call this approach arc length regularization, as it translates the widespread approach of arc length parametrization to a non parametric setting.
 
-% Define envelope
+Before proceeding, some simplifications based on the smoothness of the temporal envelope will unburden the future development of the method.
+Because we are interested in approximating a function that is smooth over the whole domain of the wave, we can assume that, in the scale of a pulse, this function will be very close to a horizontal straight line and, thus, won't touch any point inside the convex hull defined by the pulse.
+By the same token, as the change in amplitude from one pulse to the next pulse with the same sign is motivated by changes in this same envelope, one can reasonably suppose that $ \max(P[j]) \approx \max(P[j+2]) $ and approximate the convex hull of a pulse with the triangle defined by the points where the pulse crosses the horizontal the point of max amplitude; this will allow a considerable dimension reduction as this triangular pulse can become the atomic unit for the rest of the method.
+This definition also makes clear that only one point of an arbitrary pulse can be in the frontier, namely, its point of max amplitude.
 
-It will be useful to introduce the concept of a frontier as the set of points $ (i, W[i]) $ that are closest to the unknown smooth, continuous envelope of the signal.
-Because we are interested in a function that is smooth over the whole domain of the wave, we can assume that, in the scale of a pulse, this function will be very close to a horizontal straight line and, thus, won't touch any point inside the convex hull defined by the pulse.
-By the same token, as the change in amplitude from one pulse to the next is motivated by changes in the temporal envelope, one can assume that $ \max(P[j]) \approx \max(P[j+1]) $ and approximate the convex hull of a pulse with the rectangle whose length is defined by the points where the pulse crosses the horizontal axis and whose height is equal to the point of max amplitude; this will allow a considerable dimension reduction as this rectangular pulse becomes the atomic unit for the rest of the method.
+Those definitions will allow us to derive the frontiers of a wave using an alpha-shapes inspired algorithm without the need to compute the Delaunay triangulation first.
+Translating the intuitive explanation of alpha-shapes in \textcite{1994EdelsbrunnerThree}, we can say that the points in the frontier are those touched by a circle outside the signal that is not allowed to contain any point of the signal; The pulses that contain said points can be seen as frontier pulses.
+Intuitively, one can picture a circle being rolled above (or below, in the case of the negative envelope) the signal, and marking the points it touches as frontier points.
+We can therefore revisit the concept of a pseudo-cycle, only hinted at before, defining it as the samples of the original wave that belong to two successive pulses that belong to their respective frontiers, including eventual pulses between them.
 
-Besides, those definitions will allow us to derive the frontier of a wave using an alpha-shapes inspired algorithm without the need to compute the Delaunay triangulation first.
-Translating the intuitive explanation of alpha-shapes in \textcite{1994EdelsbrunnerThree}, we can say that the points in the frontier are those touched by a circle outside the signal that is not allowed to contain any point of the signal; one can picture a circle being rolled above (or below, in the case of the negative envelope) the signal, and marking the points it touches as frontier points.
+% TODO pseudo cycles image
 
-% snowball image
+% TODO snowball image
 
-It is now necessary to infer the appropriate radius of such a circle and, to that end, a measure of the instantaneous curvature of a discrete function is needed.
+It is now necessary to infer the appropriate radius of such a circle and, to that end, a measure of the curvature of a discrete function is needed.
 Discrete curvature estimation is an important task in image processing \parencite{2010FleischmannNovel} for which no default definition exists; The two possible approaches are the derivation of direct methods that use characteristics of the discrete wave to calculate its curvature at each point, or the calculation of the curvature of a curve fitted to the discrete wave \parencite{2001CoeurjollyDiscrete}.
@@
 
-As previously stated, no default discrete curvature definition exists
+As previously stated, no default discrete curvature definition exists; \textcite{2014CarrollSurvey}, for example, derives 3 such different definitions based on the approximation of a circle by inscribed, centred e circumscribed polygon.
+
+The approach here presented differs from most others found in the literature in that it is defined over the edges linking the points that forma the discrete wave.
+The idea is to find the equivalent circle whose tangent presents the same change in direction, in the same horizontal distance, as the edge of interest.
+Formally, we define $ n-1 $ vectors
+
+As an edge defines only one direction, there is a need for a referential initial direction, which wav chosen in this work as the direction of the average vector.
 
@@
 
+\subsection{Fundamental Frequency as a Welcome Side-effect}
+
 \section{Discussion + Conclusion}

### 2020-08-20 Papers/02_Pulses_Envelope/PulsesEnvelope.tex
advancements in the paper

@@
 
-\title{Robust Digital Envelope Estimation Via Geometric Properties of an Arbitrary Signal}
+\title{Robust Digital Envelope Estimation Via Geometric Properties of an Arbitrary Real Signal}
 \author{
@@
 Besides, most envelope detection techniques are designed to account for very specific settings, like pure sinusoids with moderate noise contents, in line with the most common usages of those algorithms with artificial signals in the context of analog telecommunications.
+
 % Literature review | Brief Context of Prior Research
@@
 The envelope helps to convey emotion and identity to the human voice \parencite{2018ZhuContributions}, and envelope preserving characteristics of concert halls are associated with their pleasantness \parencite{2011LokkiEngaging}.
+
 % Cite current algorithms
@@
 Thus, in this paper we formulate a general approach to envelope detection, exploiting the intrinsic characteristics of a generic, spectrally complex, wave in order to avoid the need to manual intervention or parameter tuning.
+
 % State the Significance of Your Question
@@
 Moreover, the algorithm naturally divides a signal in pseudo-cycles, that could serve pragmatic building blocks for the reconstruction of the fine structure of the wave.
+
 % State Your Claim | Objective | Hypothesis
@@
 This definition relates closely to the concept of an array or vector in most programming languages, and is used in the interest of simplicity.
+
 A pulse $ P[j], j \in \mathbb{N}_0, j < q $ in $ W $ can then be defined as a sequence of samples indexed by $ i $ such that $ \sign(W[a-1]) \ne \sign(W[a]) = \sign(W[a+1]) = \dots = \sign(W[b-1]) = \sign(W[b]) \ne \sign(W[b+1]) \forall i, a \le i \le b | a,b \in \mathbb{N}_0 $.
 In other words, each pulse is a unique subset of $ W $ with its members being the maximum sequence of samples in $ W $ with the same signal and $ P[0] \cup P[0] \cup \cdots \cup P[q-1] = W $.
+
 Naturally, each pulse can be classified as positive or negative, based on the signal of its samples, which must be the same.
@@
 Thus, in our analogy, two consecutive pulses, naturally with opposing signs, have the potential do delimitate a pseudo-cycle, as illustrated in figure \ref{fig:pulses}.
-For that to be true, however, the pulses need to make part of the frontier, as will be defined later.
+For that to be true, however, the pulses need to make part of the frontiers of the signal, as will be defined later.
 
@@
 
-% TODO Define envelope, Hilbert, etc
+The most mathematically sound definition of an envelope involves its representation as an analytic signal, introduced by Gabor in 1946 \parencite{2007HahnHistory}.
+In his work, Gabor applies the then relatively new mathematical machinery of the quantum mechanics to unify time and frequency representations of a wave, showing how the Hilbert transform could be applied to a real signal in order to obtain a complex signal that would become known as the analytic signal.
 
-In practice, however, is not uncommon for a discrete wave, specially in the case of sound, to present somewhat different positive and negative contours; As our method will eventually have to be applied separately to both sides, superior and inferior, of the discrete wave, it is only natural to introduce such separation as soon as possible.
+This analytic signal is of the form $ a(t) = s(t) + j \mathbb{H}(s(t)) $ \parencite{2016HePraat} where $ s(t) $ is the original real signal, that becomes the real part of the analytic signal.
+$ \mathbb{H}(s(t)) $, the Hilbert transform of the original signal becomes, then, the imaginary part of the analytic signal; the envelope of a signal thus represented can be straightforwardly obtained by the computation of its complex modulus.
+
+\begin{figure}[h!]
+  \centering
+    \includegraphics[width=0.8\linewidth]{02AnalyticSignal.pdf}
+  \caption{Envelope of a pure sinusoid modulated by a polynomial of degree 3, as obtained by the Hilbert Transform.
+In the first half the sinusoid is free of noise, while in the second half white gaussian noise with a standard deviation of 1/10 of the max amplitude of the wave was added to the base signal}
+  \label{fig:AnalyticSignal}
+\end{figure}
+
+Figure \ref{fig:AnalyticSignal} shows the envelope obtained via the Hilbert transform, for a pure sinusoid of local frequency 20, modulated by a polynomial of the third degree, in the presence and absence of gaussian white noise with standard deviation of $ 1/10 $ of the wave's maximum amplitude.
+It is readily noticeable that, as soon as noise is introduced in the original signal, it is reflected in the envelope, breaking our assumption of smoothness.
+Other than that, this definition yields a unique, positive envelope.
+
+In practice, however, is not uncommon for a discrete wave, specially in the case of sound, to present somewhat different positive and negative contours; as our method will forcibly have to be applied separately to both sides, superior and inferior, of the discrete wave, it is only natural to introduce such separation as soon as possible.
+
 To do so we will define the positive (upper) frontier and negative (lower) frontier as the set of points $ (i, |W[i]|) $ that are closest to the unknown smooth, continuous envelope of the signal.
-The set of all these points such that $ W[i] > 0 $ will be called the positive, piecewise linear frontier denoted by $ F_+ $; similarly, $ F_- $ denotes the negative frontier of $ W $, defined by the subset of points $ (i, W[i]) $ where $ W[i] < 0 $.
+The set of all these points such that $ W[i] > 0 $ will be called the positive, piecewise linear frontier and will be denoted by $ F_+ $; similarly, $ F_- $ denotes the negative frontier of $ W $, defined by the subset of points $ (i, W[i]) $ where $ W[i] < 0 $.
 To identify such points, we need to address the different units in which the amplitude and horizontal distance of our wave are expressed.

### 2020-08-20 Papers/02_Pulses_Envelope/images/02AnalyticSignal.py

@@ -1,0 +1,96 @@
+from scipy.signal import hilbert
+import numpy as np
+import plotly.graph_objects as go
+import numpy.polynomial.polynomial as poly
+
+'''Generating Wave'''
+
+np.random.seed(1)
+# fps = 10
+n = 1000+1
+
+X = np.arange(n)# / fps
+W = np.zeros(n)
+
+fp = [
+  [20, np.pi],
+  # [2, 1.2 * np.pi],
+  # [2.3, 1.4 * np.pi]
+  ]
+
+coefs = poly.polyfit([0, n//4, 3 * n//4, n], [0.6, 0.1, 0.9, .2], 3)
+A = poly.polyval(X, coefs)
+
+noise = np.random.normal(0, 0.1, n)
+noise[0 : n//2] = 0
+
+for f, p in fp:
+  W += np.cos(p + 2 * np.pi * f * X / n) + noise
+
+W = A * W / np.max(np.abs(W))
+
+'''Analytic_Signal'''
+analytic_signal = hilbert(W)
+hilbert_envelope = np.abs(analytic_signal)
+
+
+'''Plotting'''
+fig = go.Figure()
+fig.layout.template ="plotly_white" 
+fig.update_layout(
+  xaxis_title="$i$",
+  yaxis_title="Amplitude",
+  legend=dict(orientation='h', yanchor='top', xanchor='left', y=1.1),
+  margin=dict(l=5, r=5, b=5, t=5),
+  font=dict(
+  family="Computer Modern",
+  color="black",
+  size=18
+  )
+)
+fig.update_xaxes(showline=False, showgrid=False, zeroline=False)
+fig.update_yaxes(showline=False, showgrid=False, zerolinewidth=2, zerolinecolor="silver")
+
+'''Plotting Signal'''
+fig.add_trace(
+  go.Scatter(
+    name="Signal",
+    # x=X,
+    y=W,
+    mode="lines",
+    line=dict(
+        width=.8,
+        color="gray",
+    )
+  )
+)
+
+fig.add_trace(
+  go.Scatter(
+    name="Hilbert Envelope",
+    # x=X,
+    y=hilbert_envelope,
+    mode="lines",
+    line=dict(
+        width=.8,
+        color="black",
+    )
+  )
+)
+
+fig.add_trace(
+  go.Scatter(
+    name="$ i = n/2 $",
+    x=[n//2, n//2],
+    y=[-1, 1],
+    mode="lines",
+    line=dict(
[…]

### 2020-08-20 Papers/02_Pulses_Envelope/images/ContinuousVsDiscreteSinusoid.py

@@ -1,61 +1,0 @@
-import numpy as np
-import random
-import plotly.graph_objects as go
-
-a = 1
-F = 2
-p = 1
-fps=31
-
-''' CONTINUOUS '''
-t_f = 2 # seconds
-T = np.linspace(0, t_f, 1000)
-S = a * np.cos(p + 2 * np.pi * F * T)
-
-''' DISCRETE '''
-n = fps * t_f
-X = np.arange(n)
-f = n * F / fps
-W = a * np.cos(p + 2 * np.pi * f * X / n)
-
-fig = go.Figure()
-
-fig.layout.template ="plotly_white"
-
-fig.update_layout(
-  xaxis_title="Frame number, Time (seconds)",
-  yaxis_title="Amplitude",
-  legend=dict(orientation='h', yanchor='top', xanchor='left', y=1.1),
-  margin=dict(l=5, r=5, b=5, t=5),
-  font=dict(family="Computer Modern", color="black", size=18)
-  )
-
-stp = 5
-
-fig.update_xaxes(
-  gridwidth=1, 
-  gridcolor='silver',
-  ticktext=[f"i={i}, t={round(i/fps, 3)}" for i in X[::stp]],
-  tickvals=X[::stp] / fps
-)
-
-fig.update_yaxes(zerolinewidth=2, zerolinecolor='silver', automargin=True)
-
-fig.add_trace(go.Scatter(name=f"Continuous Wave",
-  x=T,
-  y=S,
-  mode='lines',
-  line_shape='spline',
-  line=go.scatter.Line(color="gray", width=12)
-  ))
-
-fig.add_trace(go.Scatter(name=f"Discrete Wave",
-  x=X/fps,
-  y=W,
-  mode='lines+markers',
-  # line_shape='spline',
-  line=go.scatter.Line(color="black", width=1)
-  ))
-
-# fig.show()
-fig.write_image("./02ContinuousVsDiscreteSinusoid.pdf", width=800, height=400, scale=1)

### 2020-08-22 Papers/02_Pulses_Envelope/PulsesEnvelope.tex
progress in the paper

@@
 
-This approach is used in areas such as the detection of features in images \parencite{2016VarytimidisAlpha}, reconstruction of surfaces from a cloud of points \parencite{2015WuAutomated} and Spectroscopy \parencite{2019XuModeling}, with the last work, that involves the estimation and removal of the Blaze function -a kind of envelope- of a echelle spectrograph, being particularly align with what is intended here.
+This approach is used in areas such as the detection of features in images \parencite{2016VarytimidisAlpha}, reconstruction of surfaces from a cloud of points \parencite{2015WuAutomated} and Spectroscopy \parencite{2019XuModeling}, with the last work, that involves the estimation and removal of the Blaze function (a kind of envelope) of a echelle spectrograph, being particularly align with what is intended here.
 
@@
 In the context of sound synthesis, one would be able to apply specific methods for the recreation of the envelope, more in line with its smooth, slow varying and non periodic nature and a different approach to the periodic and relatively fast changes of the temporal fine structure.
-% brief outlook on the structure of the paper
+% TODO brief outlook on the structure of the paper
 
@@
 A pulse $ P[j], j \in \mathbb{N}_0, j < q $ in $ W $ can then be defined as a sequence of samples indexed by $ i $ such that $ \sign(W[a-1]) \ne \sign(W[a]) = \sign(W[a+1]) = \dots = \sign(W[b-1]) = \sign(W[b]) \ne \sign(W[b+1]) \forall i, a \le i \le b | a,b \in \mathbb{N}_0 $.
+
 In other words, each pulse is a unique subset of $ W $ with its members being the maximum sequence of samples in $ W $ with the same signal and $ P[0] \cup P[0] \cup \cdots \cup P[q-1] = W $.
+It is convenient to attribute to each pulse a pair of coordinates: those will be the coordinates of the sample with the maximum amplitude in each pulse.
 
@@
 For a continuous function, then, a pulse would correspond to the pieces between it's roots, and would be equal to half the cycle of a sinusoid.
-Thus, in our analogy, two consecutive pulses, naturally with opposing signs, have the potential do delimitate a pseudo-cycle, as illustrated in figure \ref{fig:pulses}.
+Thus, in our analogy, two consecutive pulses, naturally with opposing signs, have the potential do delimitate a pseudo-cycle, as illustrated in figure \ref{fig:Pulses}.
 For that to be true, however, the pulses need to make part of the frontiers of the signal, as will be defined later.
@@
       \includegraphics[width=0.8\linewidth]{01Pulses.pdf}
-    \caption{The grey areas encompass the simplified convex hulls of the pulses. The dashed lines mark the frontier between pseudo-cycles}
-    \label{fig:pulses}
+    \caption{The grey areas encompass the simplified convex hulls of the pulses. The dashed lines mark the frontier between pseudo-cycles.}
+    \label{fig:Pulses}
   \end{figure}
@@
 The most mathematically sound definition of an envelope involves its representation as an analytic signal, introduced by Gabor in 1946 \parencite{2007HahnHistory}.
-In his work, Gabor applies the then relatively new mathematical machinery of the quantum mechanics to unify time and frequency representations of a wave, showing how the Hilbert transform could be applied to a real signal in order to obtain a complex signal that would become known as the analytic signal.
+In his work, \textcite{1946GaborTheory} applies the then relatively new mathematical machinery of the quantum mechanics to unify time and frequency representations of a wave, showing how the Hilbert transform could be applied to a real signal in order to obtain a complex signal that would become known as the analytic signal.
 
@@
   \caption{Envelope of a pure sinusoid modulated by a polynomial of degree 3, as obtained by the Hilbert Transform.
-In the first half the sinusoid is free of noise, while in the second half white gaussian noise with a standard deviation of 1/10 of the max amplitude of the wave was added to the base signal}
+In the first half the sinusoid is free of noise, while in the second half white gaussian noise with a standard deviation of $ 1/10 $ of the maximum amplitude of the wave was added to the base signal.}
   \label{fig:AnalyticSignal}
@@
 
-To do so we will define the positive (upper) frontier and negative (lower) frontier as the set of points $ (i, |W[i]|) $ that are closest to the unknown smooth, continuous envelope of the signal.
-The set of all these points such that $ W[i] > 0 $ will be called the positive, piecewise linear frontier and will be denoted by $ F_+ $; similarly, $ F_- $ denotes the negative frontier of $ W $, defined by the subset of points $ (i, W[i]) $ where $ W[i] < 0 $.
+To do so we will define the positive (upper) frontier and negative (lower) frontier as the set of points $ (i, |W[i]|) $ that are, necessarily but not sufficiently, the point of maximum amplitude of their respective pulses.
+The set of all these points such that $ W[i] > 0 $ will be called the positive, piecewise linear frontier and will be denoted by $ F_+ $; similarly, $ F_- $ will denote the negative frontier of $ W $, defined by the subset of points $ (i, W[i]) $ where $ W[i] < 0 $, as is shown in figure \ref{fig:Frontiers}.
 To identify such points, we need to address the different units in which the amplitude and horizontal distance of our wave are expressed.
 
-% TODO envelope, positive and negative frontiers image
+\begin{figure}[h!]
+  \centering
+    \includegraphics[width=0.8\linewidth]{03Frontiers.pdf}
+  \caption{Positive and negative frontiers, envelope and pseudo-cycles for a fragment of piano sound.
+While we have only one (positive) envelope, here mirrored in the inferior space, we have two frontiers, the superior and inferior.}
+  \label{fig:Frontiers}
+\end{figure}
 
-% TODO clarify this part
 To make abscissa and ordinate units consistent for the computation of the frontier, and proceed with a geometric treatment of the subject, we consider the set of positive and negative pulses separately.
-For each set, we can multiply all pulses by the average length of the pulses in the set, after having divided by the average of pulses amplitudes, putting both axes in the same frame related unit $i$, that relates to time by the equation $i = t \ \mathit{fps}$.
-One can call this approach arc length regularization, as it translates the widespread approach of arc length parametrization to a non parametric setting.
+For each set, we multiply all pulses by the average length of the pulses in the set, after having divided by the average of the pulses' amplitudes, putting both axes in the same frame related unit $i$, related to time by the equation $i = t \ \mathit{fps}$.
 
-Before proceeding, some simplifications based on the smoothness of the temporal envelope will unburden the future development of the method.
 Because we are interested in approximating a function that is smooth over the whole domain of the wave, we can assume that, in the scale of a pulse, this function will be very close to a horizontal straight line and, thus, won't touch any point inside the convex hull defined by the pulse.
-By the same token, as the change in amplitude from one pulse to the next pulse with the same sign is motivated by changes in this same envelope, one can reasonably suppose that $ \max(P[j]) \approx \max(P[j+2]) $ and approximate the convex hull of a pulse with the triangle defined by the points where the pulse crosses the horizontal the point of max amplitude; this will allow a considerable dimension reduction as this triangular pulse can become the atomic unit for the rest of the method.
-This definition also makes clear that only one point of an arbitrary pulse can be in the frontier, namely, its point of max amplitude.
 
-Those definitions will allow us to derive the frontiers of a wave using an alpha-shapes inspired algorithm without the need to compute the Delaunay triangulation first.
-Translating the intuitive explanation of alpha-shapes in \textcite{1994EdelsbrunnerThree}, we can say that the points in the frontier are those touched by a circle outside the signal that is not allowed to contain any point of the signal; The pulses that contain said points can be seen as frontier pulses.
-Intuitively, one can picture a circle being rolled above (or below, in the case of the negative envelope) the signal, and marking the points it touches as frontier points.
-We can therefore revisit the concept of a pseudo-cycle, only hinted at before, defining it as the samples of the original wave that belong to two successive pulses that belong to their respective frontiers, including eventual pulses between them.
+By the same token, as the change in amplitude from one pulse to the next pulse with the same sign is motivated by changes in this same envelope, one can reasonably suppose that $ \max(P[j]) \approx \max(P[j+2]) $.
+That justifies the approximation of the convex hull of a pulse by the triangle defined by the points where the pulse crosses the horizontal the point of maximum amplitude; this will allow a considerable dimension reduction as this triangular pulse can become the atomic unit for the rest of the method.
 
-% TODO pseudo cycles image
+This definition also makes clear that only one point of an arbitrary pulse can be in the frontier, namely, its point of maximum amplitude; those definitions will allow us to derive the frontiers of a wave using an alpha-shapes inspired algorithm without the need to compute the Delaunay triangulation first.
+
+Translating the intuitive explanation of alpha-shapes in \textcite{1994EdelsbrunnerThree} to our context, we can say that the points in the frontier are those touched by a circle outside the signal that is not allowed to contain any point of the signal; The pulses that contain said points can be seen as frontier pulses.
+Intuitively, one can picture a circle being rolled above (or below, in the case of the negative frontier) the signal, and marking the points it touches as frontier points.
+
+We can therefore revisit the concept of a pseudo-cycle, only hinted at before, defining it as the samples of the original wave that belong to two successive pulses that belong to their respective frontiers, including eventual pulses between them, as can be seen in figure \ref{fig:Frontiers}.
 
@@
 Discrete curvature estimation is an important task in image processing \parencite{2010FleischmannNovel} for which no default definition exists; The two possible approaches are the derivation of direct methods that use characteristics of the discrete wave to calculate its curvature at each point, or the calculation of the curvature of a curve fitted to the discrete wave \parencite{2001CoeurjollyDiscrete}.
-To fit a polynomial to the a generic wave two approaches are readily available: the least mean squares approach, that seeks to minimize the dependant variable errors and the total least squares problem, that treat both variables symmetrically \parencite{1980GolubAnalysis}.
+
+To fit a polynomial to the a generic wave two approaches are readily available: the least mean squares approach, that seeks to minimize the dependant variable errors and the total least squares problem, that treats both variables symmetrically \parencite{1980GolubAnalysis}.
+
 The geometric, symmetrical nature of the problem excludes the more computationally economic least mean squares approach however, leaving us with the total least squares, for which no general closed form solutions are available \parencite{2007MarkovskyOverview}; this leads us to turn our attention to direct approaches of curvature estimation.
@@
 
-As previously stated, no default discrete curvature definition exists; \textcite{2014CarrollSurvey}, for example, derives 3 such different definitions based on the approximation of a circle by inscribed, centred e circumscribed polygon.
+As previously stated, no default discrete curvature definition exists; \textcite{2014CarrollSurvey}, for example, derives 3 such different definitions based on the approximation of a circle by an inscribed, centred e circumscribed polygon.
+In the context of 3d meshes, \textcite{2016VasaMultivariate} evaluates multiple estimators from a multivariate point o view.
 
-The approach here presented differs from most others found in the literature in that it is defined over the edges linking the points that forma the discrete wave.
+For our needs, however, it is more convenient to define our own discrete curvature estimator; the approach here presented differs from most others found in the literature in that the curvature is defined over the edges linking the points that form the discrete wave, instead of the vertices.
+
+To that end we are going to apply the definition of smooth curvature as the rate of change of the unit tangent to a curve, noting that this is equivalente to that of the osculating circle \parencite{2016VasaMesh}.
 The idea is to find the equivalent circle whose tangent presents the same change in direction, in the same horizontal distance, as the edge of interest.
-Formally, we define $ n-1 $ vectors
 
-As an edge defines only one direction, there is a need for a referential initial direction, which wav chosen in this work as the direction of the average vector.
+Formally, for each frontier, positive or negative, we define $ m - 1 $ vectors, $ m $ being the number of positive or negative pulses respectively, as the coordinates of the posterior pulse with the relevant signal minus the preceding pulse: $\vec{v}[l]^\pm = (P[l+1]_x^\pm - P[l]_x^\pm, P[l+1]_y^\pm - P[l]_y^\pm), \forall 0 \le l < m - 1 $, recalling that the $x$ coordinate of a pulse was defined as the $x$ coordinate of its sample with maximum amplitude, while the $y$ coordinate of a pulse is this amplitude itself.
 
-% osculating circle image, fitted line
+Since an edge defines only one direction, there remains the need for a reference.
+As we are interested in the relative curvature, it is natural to use the average direction of all vectors, $ \vec{\overline{v}} = \frac{1}{m-1} \sum_{l=0}^{m-1} \vec{v}[l] $, so as to measure the relative change of direction of each edge.
+
+% TODO osculating circle image, average vector
 

### 2020-08-22 Papers/02_Pulses_Envelope/images/01Pulses.py

@@ -1,13 +1,7 @@
-# import random
-# random.seed(2)
 import sys
 sys.path.append("c:/Users/tesse/Desktop/Files/Dropbox/0_Thesis/Python")
-# import site
-# site.addsitedir("c:/Users/tesse/Desktop/Files/Dropbox/0_Thesis/Python")
 import numpy as np
-
 import plotly.graph_objects as go
-
 from Helper import signal_to_pulses, get_pulses_area
 
 '''Generating Wave'''
@@ -36,9 +30,7 @@ pulses_X, pulses_Y = get_pulses_area(pulses)
 
 '''Plotting'''
 fig = go.Figure()
-
-fig.layout.template ="plotly_white"
- 
+fig.layout.template ="plotly_white" 
 fig.update_layout(
   xaxis_title="$i$",
   yaxis_title="Amplitude",
@@ -50,7 +42,6 @@ fig.update_layout(
   size=18
   )
 )
-
 fig.update_xaxes(showline=False, showgrid=False, zeroline=False)
 fig.update_yaxes(showline=False, showgrid=False, zerolinewidth=2, zerolinecolor='gray')
 

### 2020-08-22 Papers/02_Pulses_Envelope/images/03Frontiers.py

@@ -1,46 +1,77 @@
-import numpy.polynomial.polynomial as poly
 import sys
 sys.path.append("c:/Users/tesse/Desktop/Files/Dropbox/0_Thesis/Python")
-import numpy as np
-from scipy import signal
-
+import os
+from pathlib import Path
 import plotly.graph_objects as go
+import numpy as np
+from scipy.interpolate.interpolate import interp1d
+from Helper import read_wav, get_pulses_area, split_pulses, signal_to_pulses, get_frontier#, save_wav, get_frontier
+from scipy.signal import savgol_filter
 
-from Helper import signal_to_pulses, get_pulses_area
+# os.path.dirname(os.path.dirname(path))
 
-'''Generating Wave'''
-
-np.random.seed(1)
-# fps = 10
-n = 1000+1
-
-X = np.linspace(0, 1, n)
-W = np.zeros(n)
-
-afp = [
-  [3, 10, np.pi],
-  [2, 2, 1.2 * np.pi],
-  [1, 2.3, 1.4 * np.pi]
-  ]
-for a, f, p in afp:
-  W += a * np.cos(p + 2 * np.pi * f * X / n)# + np.random.normal(0, a / 10, n)
-
-# coefs = poly.polyfit([0, n//4, 3 * n//4, n], [0.0, 0.1, 2.0, 2.0], 3)
-# A = poly.polyval(X, coefs)
-# W += A
-
-W = signal.sawtooth(2 * np.pi * 5 * X)
-
-'''Wave to pulses'''
-pulses = signal_to_pulses(W)
-pulses_X, pulses_Y = get_pulses_area(pulses)
-
-
-'''Plotting'''
+'''==============='''
+''' Read wav file '''
+'''==============='''
 fig = go.Figure()
 
-fig.layout.template ="plotly_white"
- 
+parent_path = str(Path(os.path.abspath('./')).parents[1])
+print(parent_path)
+name = "piano33"
+W, fps = read_wav(f"{parent_path}/Python/Samples/{name}.wav")
+
+W = W[100103 : 100775]
+
+W = W - np.average(W)
+amplitude = np.max(np.abs(W))
+W = W / amplitude
+n = W.size
+print(n)
+X = np.arange(n)
+
+
+pulses = signal_to_pulses(W)
+
+areas_X, areas_Y = get_pulses_area(pulses)
+
+pos_pulses, neg_pulses, pos_noises, neg_noises = split_pulses(pulses)
+
[…]

### 2020-08-24 Papers/02_Pulses_Envelope/PulsesEnvelope.tex
created curvature related images for the paper

@@
 \documentclass{article}
+\usepackage{lmodern}
+\usepackage[T1]{fontenc}
 \usepackage{amsfonts} % \mathbb
@@
 
-  \begin{figure}[h!]
+  \begin{figure}[ht!]
     \centering
@@
 
-\begin{figure}[h!]
+\begin{figure}[ht!]
   \centering
@@
 As previously stated, no default discrete curvature definition exists; \textcite{2014CarrollSurvey}, for example, derives 3 such different definitions based on the approximation of a circle by an inscribed, centred e circumscribed polygon.
-In the context of 3d meshes, \textcite{2016VasaMultivariate} evaluates multiple estimators from a multivariate point o view.
+In the context of 3d meshes, \textcite{2016VasaMultivariate} evaluate a range of estimators from a multivariate point o view.
 
@@
 
-To that end we are going to apply the definition of smooth curvature as the rate of change of the unit tangent to a curve, noting that this is equivalente to that of the osculating circle \parencite{2016VasaMesh}.
+To that end we are going to apply the definition of smooth curvature as the rate of change of the unit tangent to a curve, noting that this is equivalent to that of the osculating circle \parencite{2016VasaMesh}.
 The idea is to find the equivalent circle whose tangent presents the same change in direction, in the same horizontal distance, as the edge of interest.
+This approach presents the convenience of estimating the radius directly.
+
+\begin{figure}[ht!]
+  \centering
+    \includegraphics[width=0.5\linewidth]{04DiscreteCurvature.pdf}
+  \caption{Change of direction from horizontal, in $x_0$, until an angle of $\theta$, in $x_1$, that is equivalent to the angle between the average direction and the direction of the vector $\vec{v}[l]$ whose radius of curvature is being calculated.
+$x_0$ and $x_1$} correspond to the the horizontal coordinates of the tale and tip of $\vec{v}[l]$, respectively.
+  \label{fig:DiscreteCurvature}
+\end{figure}
 
@@
 Since an edge defines only one direction, there remains the need for a reference.
-As we are interested in the relative curvature, it is natural to use the average direction of all vectors, $ \vec{\overline{v}} = \frac{1}{m-1} \sum_{l=0}^{m-1} \vec{v}[l] $, so as to measure the relative change of direction of each edge.
+As we are interested in the relative curvature, it is natural to use the average direction of all vectors, $ \vec{\overline{v}}^\pm = \frac{1}{m-1} \sum_{l=0}^{m-1} \vec{v}[l]^\pm $, so as to measure the relative change of direction of each edge.
+This procedure, as well as the circles generated by it, are illustrated in figure \ref{fig:Curvatures}
 
-% TODO osculating circle image, average vector
+\begin{figure}[ht!]
+  \centering
+    \includegraphics[width=0.8\linewidth]{05Curvatures.pdf}
+  \caption{The equivalent circles obtained via the procedure described for the positive frontier.
+$\theta$, the angle between $ \vec{\overline{v}}^+$ and each of the $\vec{v}[l]^+$ is used, as is the horizontal component of each $\vec{v}[l]^+$ to obtain the radii of the circles.}
+  \label{fig:Curvatures}
+\end{figure}
 
-% For a continuous, twice differentiable function $ y = f(x); x, y \in \mathbb{R} $, the curvature as a function of the independent variable is given by $ \kappa(x) = \frac{f''(x)}{(f'(x)^2 + 1)^{\frac{3}{2}}} $ which is also the inverse of the radius of the osculating circle tangent to the curve at $ x $.
-Since we are assuming smoothness of envelope in the neighbourhood of a pulse, it makes sense to approximate it with a parabola $ y(x)=ax^2+bx+c $, since a higher order polynomial would provide unnecessary complexity.
-Besides some characteristics of the parabola derived from its relation with Bezier curves will be useful later.
+After this procedure one obtains a set of points in $ \mathbb{R}^2 $ % TODO
+
+\begin{figure}[ht!]
+  \centering
+    \includegraphics[width=0.8\linewidth]{06AvgCircle.pdf}
+  \caption{The positive frontier as obtained by the use of the circle representing the average curvature of the discrete wave illustrated.}
+  \label{fig:AvgCircle}
+\end{figure}
 
@@
 
+\subsection{Robustness}
+
 \subsection{Fundamental Frequency as a Welcome Side-effect}

### 2020-08-26 Papers/02_Pulses_Envelope/PulsesEnvelope.tex
updated discrete curvature svg image

@@
   % methods
-  The approach here described draws inspiration from general, computational and differential geometry to isolate the frontier of an arbitrary signal.
-The concepts of alpha-shapes, concave hulls, arc-length parametrization and discrete curve curvature are adapted.
+  The approach here described draws inspiration from general, computational and differential geometry to isolate the frontiers and thus estimate the temporal envelope of an arbitrary signal; to that end The concepts of alpha-shapes, concave hulls, and discrete curve curvature are explored.
 We also define a pulse in the context of a arbitrary digital signal as a means to reduce dimensionality and the complexity of the proposed algorithm.
@@
 
-This analytic signal is of the form $ a(t) = s(t) + j \mathbb{H}(s(t)) $ \parencite{2016HePraat} where $ s(t) $ is the original real signal, that becomes the real part of the analytic signal.
+This analytic signal is of the form $ a(t) = s(t) + \mathbb{H}(s(t)) \ \text{i} $ \parencite{2016HePraat} where $ s(t) $ is the original real signal, that becomes the real part of the analytic signal.
 $ \mathbb{H}(s(t)) $, the Hilbert transform of the original signal becomes, then, the imaginary part of the analytic signal; the envelope of a signal thus represented can be straightforwardly obtained by the computation of its complex modulus.
@@
 
-% TODO snowball image
-
 It is now necessary to infer the appropriate radius of such a circle and, to that end, a measure of the curvature of a discrete function is needed.
@@
   \centering
-    \includegraphics[width=0.5\linewidth]{04DiscreteCurvature.pdf}
-  \caption{Change of direction from horizontal, in $x_0$, until an angle of $\theta$, in $x_1$, that is equivalent to the angle between the average direction and the direction of the vector $\vec{v}[l]$ whose radius of curvature is being calculated.
-$x_0$ and $x_1$} correspond to the the horizontal coordinates of the tale and tip of $\vec{v}[l]$, respectively.
+    \includegraphics[width=0.3\linewidth]{04DiscreteCurvature.pdf}
+  \caption{Change of direction from horizontal, in $ \hat{u}_0 $, until an angle of $ \theta $, in $ \hat{u}_1 $, that is equivalent to the angle between the average direction and the direction of the vector $ \vec{v}_l $ whose radius of curvature is being calculated, shown in the picture in grey.}
   \label{fig:DiscreteCurvature}
@@
 
-Formally, for each frontier, positive or negative, we define $ m - 1 $ vectors, $ m $ being the number of positive or negative pulses respectively, as the coordinates of the posterior pulse with the relevant signal minus the preceding pulse: $\vec{v}[l]^\pm = (P[l+1]_x^\pm - P[l]_x^\pm, P[l+1]_y^\pm - P[l]_y^\pm), \forall 0 \le l < m - 1 $, recalling that the $x$ coordinate of a pulse was defined as the $x$ coordinate of its sample with maximum amplitude, while the $y$ coordinate of a pulse is this amplitude itself.
+Formally, for each frontier, positive or negative, we define $ m-1 $ vectors, $ m $ being the number of positive or negative pulses respectively, as the coordinates of the posterior pulse with the relevant signal minus the preceding pulse: $\vec{v}_l^\pm = (P[l+1]_x^\pm - P[l]_x^\pm, P[l+1]_y^\pm - P[l]_y^\pm), \forall 0 \ \le l < m - 1 $, recalling that the $x$ coordinate of a pulse was defined as the $x$ coordinate of its sample with maximum amplitude, while the $y$ coordinate of a pulse is this amplitude itself.
 
 Since an edge defines only one direction, there remains the need for a reference.
-As we are interested in the relative curvature, it is natural to use the average direction of all vectors, $ \vec{\overline{v}}^\pm = \frac{1}{m-1} \sum_{l=0}^{m-1} \vec{v}[l]^\pm $, so as to measure the relative change of direction of each edge.
-This procedure, as well as the circles generated by it, are illustrated in figure \ref{fig:Curvatures}
+As we are interested in the relative curvature, it is natural to use the average direction of all vectors, $ \vec{\overline{v}}^\pm = \frac{1}{m-1} \sum_{l=0}^{m-1} \vec{v}_l^\pm $, so as to measure the relative change of direction of each edge.
+This procedure, as well as the circles generated by it, are illustrated in figure \ref{fig:Curvatures}.
+
+The angle $ \theta $ between the average direction and each of the vectors can be obtained via the relation of the slope of the respective lines, as in the expression $ \theta = \arctan( (m - \overline{m}) / (1 + m * \overline{m}) ) $.
+The radii of the equivalent circles are thus given by $ r_l = x_l / \sin (\theta_l) $ with the curvature given by $\kappa_l =1/r_l=\sin (\theta_l) / x_l$.
 
@@
   \caption{The equivalent circles obtained via the procedure described for the positive frontier.
-$\theta$, the angle between $ \vec{\overline{v}}^+$ and each of the $\vec{v}[l]^+$ is used, as is the horizontal component of each $\vec{v}[l]^+$ to obtain the radii of the circles.}
+$\theta$, the angle between $ \vec{\overline{v}}^+$ and each of the $\vec{v}_l^+$ is used, as is the horizontal component of each $\vec{v}_l^+$ to obtain the radii of the circles.}
   \label{fig:Curvatures}
@@
 
-After this procedure one obtains a set of points in $ \mathbb{R}^2 $ % TODO
+After this procedure one obtains a set of points in $ \mathbb{R}^2 $ where the ordinate represents the instantaneous radius, with its localization in the space of the wave given by the ordinates.
+Given the robustness of the method, for an average wave of a couple of seconds at an ordinary rate of 44100 $ \mathit{fps} $ the average of the radius values is sufficient to assure the identification of the frontier.
+However, a curve fitting method can be used in the case of longer waves.
+An example of the use of the average equivalent circle can be seen in figure \ref{fig:AvgCircle}.
 
@@
 
+Applying the method to a diverse set of sound waves, we can have an idea of its generality.
+% Revisit purpose and methods
+% Overview of results
+% Key results in details
+
 \subsection{Robustness}
@@
 \subsection{Fundamental Frequency as a Welcome Side-effect}
+% Comparison with results in other works
+% Comparison with predictions
+% Problems with results
+% Possible implications
 

### 2020-08-29 Papers/02_Pulses_Envelope/PulsesEnvelope.tex
improvements in the images for the paper

@@
 \documentclass{article}
+% \usepackage{svg}
 \usepackage{lmodern}
@@
 \section{Results}
-
-Applying the method to a diverse set of sound waves, we can have an idea of its generality.
+% TODO
+Figure \ref{fig:FullFrontiers} illustrates the frontiers obtained by the application of the method to 6 diverse discrete sound waves.
 % Revisit purpose and methods
@@
 % Key results in details
+
+\begin{figure}[ht!]
+  \centering
+    \includegraphics[width=0.8\linewidth]{07FullFrontiers.pdf}
+  \caption{Positive and negative frontiers of 6 digital waves.
+For each wave, the region highlighted in black is shown in detail.
+For each wave the horizontal axis is the sample number $i$, while the vertical axis is the normalized amplitude.
+All 6 waves were recorded at 44100 $ \mathit{fps} $.}
+  \label{fig:FullFrontiers}
+\end{figure}
 

### 2020-08-31 Papers/02_Pulses_Envelope/PulsesEnvelope.tex
added manual recipe to build pdf from tex

@@
 \begin{abstract} % 250 words
-
-% estimate parameters from the signal itself
-% broadband
-% Dominant frequency as by-product
-% we define several entities in the process, introducing a new measure of discrete curve curvature, hoping to consolidate the terminology in the area
-
 % background  
-  Despite being an elusive concept, mathematically well defined only for artificially generated waves, the temporal amplitude envelope of a signal is essential for its complete characterization, being the primary information carrying medium in spoken voice and telecommunications, for example.
-  intuitively, the temporal envelope can be understood as a smooth function on the same variable as the principal wave, or temporal fine structure, that modulates the signal, responsible for its outer shape.
-It is implied in this definition that the envelope is non periodic in general and is thus better addressed by specific techniques diverse from those used in the analysis of the periodic, rapidly changing wave it encompasses.
+Despite being an elusive concept, the temporal amplitude envelope of a signal is essential for its complete characterization, being the primary information carrying medium in spoken voice and telecommunications, for example.
+intuitively, the temporal envelope can be understood as a smooth function, that modulates the signal, being responsible for its outer shape.
 % motivation
 Envelope detection techniques have applications in areas like health, sound classification and synthesis, seismology and speech recognition.
-Despite of that, a general approach to envelope detection of signals with rich spectral content seems to be lacking, and most methods involve manual intervention, like filter design, based on prior knowledge about the wave under investigation.
+Despite of that, a general approach to envelope detection of signals with rich spectral content seems to be lacking, and most methods involve manual intervention, like filter design or smoothing, based on prior knowledge about the wave under investigation.
 % objective
-  In this paper we propose a framework that uses intrinsic characteristics of a signal to estimate its envelope, eliminating the necessity of parameter tuning.
+In this paper we propose a framework that uses intrinsic characteristics of a signal to estimate its envelope, completely eliminating the necessity of parameter tuning.
 % methods
-  The approach here described draws inspiration from general, computational and differential geometry to isolate the frontiers and thus estimate the temporal envelope of an arbitrary signal; to that end The concepts of alpha-shapes, concave hulls, and discrete curve curvature are explored.
+The approach here described draws inspiration from general, computational and differential geometry to isolate the frontiers and thus estimate the temporal envelope of an arbitrary signal; to that end The concepts of alpha-shapes, concave hulls, and discrete curvature are explored.
 We also define a pulse in the context of a arbitrary digital signal as a means to reduce dimensionality and the complexity of the proposed algorithm.
 % results
-
-  % implications
-
+We find that the algorithm accurate in the identification of the frontiers of a range of digital sound signal with very diverse characteristics, while providing the dominant frequency of the wave as a by-product.
+The algorithm also compares favourably with a common detection technique based on the Hilbert Transform.
+While the results are satisfactory per se, we define several geometric entities in this work, including a new measure of discrete curvature; we hope to consolidate theory in the area, while providing a solid basis for the development of new techniques, specially in sound synthesis.
 \end{abstract}
@@
 % Opener sentence
-Despite being ubiquitous in digital signal processing, the literature about envelope detection is very fragmented \parencite{2017LyonsDigital}.
-Besides, most envelope detection techniques are designed to account for very specific settings, like pure sinusoids with moderate noise contents, in line with the most common usages of those algorithms with artificial signals in the context of analog telecommunications.
+Despite being ubiquitous in both analog and digital signal processing, the literature about envelope detection is very fragmented \parencite{2017LyonsDigital}.
+Besides, most envelope detection techniques are designed to account for very specific settings, like pure sinusoids with moderate noise contents, in line with the most common usages of those algorithms with artificial signals in the context of analog telecommunications; that limitation arises in part by the lack of a mathematical definition for a temporal envelope \parencite{2013Mengempirical}.
 
@@
 % TODO
-Figure \ref{fig:FullFrontiers} illustrates the frontiers obtained by the application of the method to 6 diverse discrete sound waves.
 % Revisit purpose and methods
 % Overview of results
+Figure \ref{fig:FullFrontiers} illustrates the frontiers obtained to six diverse discrete sound waves, as well as a detail view of a shorter segment for each wave.
+All waves are records of physical sounds, chosen to represent the applicability of the algorithm in real world scenarios.
+
+We can see that the frontier is satisfactorily detected in harmonic an inharmonic sounds, and presents robustness in relation to the number of samples and the frequencies of the waves.
+
 % Key results in details
@@
   \label{fig:FullFrontiers}
+\end{figure}
+
+The impact of the normalization of the wave with the frontiers obtained can be seen in figure \ref{fig:Fourier}, which shows the original and normalized wave in the case of a recording of the key 33 of a piano both in the time and frequency domains.
+
+\begin{figure}[ht!]
+  \centering
+    \includegraphics[width=0.8\linewidth]{08Fourier.pdf}
+  \caption{Original and normalized views of a recording of key 33 (F3, 175 Hz) from a grand piano both in time and frequency domains.}
+  \label{fig:Fourier}
 \end{figure}

### 2020-09-02 Papers/02_Pulses_Envelope/PulsesEnvelope.tex
added robustness proof and images to the paper

@@
 \documentclass{article}
-% \usepackage{svg}
-\usepackage{lmodern}
+\usepackage{lmodern} % Latin Modern Font
 \usepackage[T1]{fontenc}
@@
 \usepackage{amsmath}
+\usepackage{amssymb} % blackbox symbol
+\usepackage{hyperref}
+\usepackage{hypcap}
 \DeclareMathOperator{\sign}{sign}
 \bibliography{bibli}
-
+\hypersetup{
+  colorlinks=true, 
+  allcolors=black,
+  pdfauthor={Carlos Tarjano},
+  pdftitle={Robust Digital Envelope Estimation Via Geometric Properties of an Arbitrary Real Signal}
+}
 \title{Robust Digital Envelope Estimation Via Geometric Properties of an Arbitrary Real Signal}
-\author{
-  Carlos Tarjano
-  \and 
-  Valdecy Pereira
-  }
+\author{Carlos Tarjano \and Valdecy Pereira}
 
@@
 
-The angle $ \theta $ between the average direction and each of the vectors can be obtained via the relation of the slope of the respective lines, as in the expression $ \theta = \arctan( (m - \overline{m}) / (1 + m * \overline{m}) ) $.
+The angle $ \theta $ between the average direction and each of the vectors can be obtained via the relation of the slope of the respective lines, as in the expression $ \theta = \arctan( (m - \overline{m}) / (1 + m \overline{m}) ) $.
 The radii of the equivalent circles are thus given by $ r_l = x_l / \sin (\theta_l) $ with the curvature given by $\kappa_l =1/r_l=\sin (\theta_l) / x_l$.
+
+\begin{equation} \label{eq:1}
+  r_l = \left\lvert \frac{x_l \sqrt{\overline{m}^2 + 1} \ \sqrt{x_l^2 + y^2}}{\overline{m} x_l - y} \right\rvert
+\end{equation}
 
@@
 \end{figure}
-
-% Therefore, the definition of the curvature becomes $ \kappa(x) = \frac{2 a}{\big( (b + 2 a x)^2 + 1 \big)^\frac{3}{2}} $ with the average curvature between points $ x_0 $ and $ x_1 $ given by equation \ref{eq:1}.
-
-% \begin{equation} \label{eq:1}
-%   \overline{\kappa}(x) = 
-%   \frac{
-%       \left( \dfrac{2 a x_1 + b}{\sqrt{\big(2 a x_1 + b \big)^2+1}} - \dfrac{2 a x_0 + b}{\sqrt{\big(2 a x_0 + b \big)^2+1}} \right) 
-%   }{
-%     (x_1 - x_0)
-%   }
-% \end{equation}
-
-% \subsection{Theory}
 
@@
 
+While a theoretical basis for the normalization of the amplitude is given, and is justified in practice by the computational stability issues that might arise with arbitrary values, the algorithm is, from a pure mathematical point of view, impervious to amplitude scaling, as we now prove.
+
+let $ s $ be a positive real number that multiplies every sample of a discrete wave $ W $.
+The subdivision in pulses of such a wave is unchanged by the vertical scaling operation, and we can write the slope of a line segment from a pulse to the next pulse with the same sign as $ m_l = (s y_{l + 1} - s y_l) / (x_{l + 1} - x_l) $.
+
+Assuming, for the sake of simplicity, $ \overline{m} = 0 $, the radius of the equivalent circle of such a segment, as per equation \ref{eq:1}, is
+
+$$ r_l =\left\lvert \frac{\sqrt{\left(s {y}_{l + 1} - s {y}_{l}\right)^{2} + \left({x}_{l + 1} - {x}_{l}\right)^{2}} \left({x}_{l + 1} - {x}_{l}\right)}{s {y}_{l + 1} - s {y}_{l}} \right\rvert $$
+
+And the square of the distance $ d_l $ from the center of a circle $ C $ with that radius passing through the points defined by the coordinates of both pulses to the median point $\overline{p} = \big((x_l + x_{l + 1}) / 2, (y_l + y_{l + 1}) / 2\big) $ of the line segment is
+
+$$ d_l^2 = - \frac{\left(s {y}_{l + 1} - s {y}_{l} - 2 {x}_{l + 1} + 2 {x}_{l}\right) \left(s {y}_{l + 1} - s {y}_{l} + 2 {x}_{l + 1} - 2 {x}_{l}\right) \left(s^{2} {y}_{l + 1}^{2} - 2 s^{2} {y}_{l + 1} {y}_{l} + s^{2} {y}_{l}^{2} + {x}_{l + 1}^{2} - 2 {x}_{l + 1} {x}_{l} + {x}_{l}^{2}\right)}{4 s^{2} \left({y}_{l + 1} - {y}_{l}\right)^{2}} $$
+
+Assuming invariance to vertical scaling, $ \overline{p} $ lying inside or outside the circle must not depend on $ s $; $ |d_l| \le |r_l| \iff d_l^2 / r_l^2 \le 1 \iff \overline{p} \in C $.
+
+$$ \frac{d_l^2}{r_l^2} = \frac{- \frac{s^{2} \left({y}_{i + 1}^{2} + {y}_{i}^{2}\right)}{4} + \frac{s^{2} {y}_{i + 1} {y}_{i}}{2} + ({x}_{i + 1} - {x}_{i})^{2}}{({x}_{i + 1} - {x}_{i})^{2}} \le 1$$
+
+Reorganizing:
+
+$$ y_{i + 1} y_i \le \frac{y_{i + 1}^2 + y_i^2}{2} \iff \overline{p} \in C $$
+
+Thus, the circle $ C $ containing the point $\overline{p}$ doesn't depend on $ s $ nor on the horizontal parametrization of $ W \ quad \blacksquare $
+
+\begin{figure}[ht!]
+  \centering
+    \includegraphics[width=0.8\linewidth]{10xRobustness.pdf}
+  \caption{Minimum and maximum scales of the abscissa before computational stability issues arose, for a recording of a drumkit tom.}
+  \label{fig:xRobustness}
+\end{figure}
+
+\begin{figure}[ht!]
+  \centering
+    \includegraphics[width=0.8\linewidth]{09yRobustness.pdf}
+  \caption{Minimum and maximum values of the amplitude before computational stability issues arose, for a recording of a drumkit tom.}
+  \label{fig:yRobustness}
+\end{figure}
+
+Figures \ref{fig:xRobustness} and \ref{fig:yRobustness} illustrate the computational resilience of the algorithm for scaling in both the abscissa and ordinate axes, respectively.
+Although mathematically independent of such parametrizations, in practice care must be taken, for different reasons, in the choice of values for both axes.
+% TODO lack of standard and great number of samples
+
 \subsection{Fundamental Frequency as a Welcome Side-effect}
@@
 \section{References}
+
 \printbibliography[heading=none]

### 2020-09-02 Papers/02_Pulses_Envelope/images/09yRobustness.py

@@ -1,0 +1,252 @@
+import plotly.graph_objects as go
+import numpy as np
+from scipy.interpolate import interp1d
+import numpy.polynomial.polynomial as poly
+import wave
+from scipy.signal import savgol_filter
+import os
+from pathlib import Path
+import sys
+import argparse
+from plotly.subplots import make_subplots
+
+def read_wav(path): 
+  """returns signal & fps"""
+  wav = wave.open(path , 'r')
+  signal = np.frombuffer(wav.readframes(-1) , np.int16)
+  fps = wav.getframerate()
+  return signal, fps
+
+
+def get_circle(x0, y0, x1, y1, r):
+  '''returns center of circle that passes through two points'''
+  
+  radsq = r * r
+  q = np.sqrt((x1 - x0)**2 + (y1 - y0)**2)
+
+  # assert q <= 2 * r, f"shit"
+
+  x3 = (x0 + x1) / 2
+  y3 = (y0 + y1) / 2
+
+  if y0 + y1 >= 0:
+    xc = x3 + np.sqrt(radsq - (q / 2)**2) * ((y0 - y1) / q)
+    yc = y3 + np.sqrt(radsq - (q / 2)**2) * ((x1 - x0) / q)
+  else:
+    xc = x3 - np.sqrt(radsq - (q / 2)**2) * ((y0 - y1) / q)
+    yc = y3 - np.sqrt(radsq - (q / 2)**2) * ((x1 - x0) / q)
+
+  return xc, yc
+
+
+def get_radius_function(X, Y):
+  m0 = np.average(Y[1:] - Y[:-1]) / np.average(X[1:] - X[:-1])
+  curvatures_X = []
+  curvatures_Y = []
+  for i in range(len(X) - 1):
+    m1 = (Y[i + 1] - Y[i]) / (X[i + 1] - X[i])
+
+    theta = np.arctan( (m1 - m0) / (1 + m1 * m0) )
+    k = np.sin(theta) / (X[i + 1] - X[i])
+
+    # if k >= 0:
+    curvatures_X.append((X[i + 1] + X[i]) / 2)
+    curvatures_Y.append(k)
+    
+  curvatures_Y = np.array(curvatures_Y)
+  coefs = poly.polyfit(curvatures_X, curvatures_Y, 0)
+  smooth_curvatures_Y = poly.polyval(curvatures_X, coefs)
+
[…]

### 2020-09-02 Papers/02_Pulses_Envelope/images/09yRobustnessTest.py

@@ -1,0 +1,263 @@
+import plotly.graph_objects as go
+import numpy as np
+from scipy.interpolate import interp1d
+import numpy.polynomial.polynomial as poly
+import wave
+from scipy.signal import savgol_filter
+import os
+from pathlib import Path
+import sys
+import argparse
+from plotly.subplots import make_subplots
+
+def read_wav(path): 
+  """returns signal & fps"""
+  wav = wave.open(path , 'r')
+  signal = np.frombuffer(wav.readframes(-1) , np.int16)
+  fps = wav.getframerate()
+  return signal, fps
+
+
+def get_circle(x0, y0, x1, y1, r):
+  '''returns center of circle that passes through two points'''
+  
+  radsq = r * r
+  q = np.sqrt((x1 - x0)**2 + (y1 - y0)**2)
+
+  # assert q <= 2 * r, f"shit"
+
+  x3 = (x0 + x1) / 2
+  y3 = (y0 + y1) / 2
+
+  if y0 + y1 >= 0:
+    xc = x3 + np.sqrt(radsq - (q / 2)**2) * ((y0 - y1) / q)
+    yc = y3 + np.sqrt(radsq - (q / 2)**2) * ((x1 - x0) / q)
+  else:
+    xc = x3 - np.sqrt(radsq - (q / 2)**2) * ((y0 - y1) / q)
+    yc = y3 - np.sqrt(radsq - (q / 2)**2) * ((x1 - x0) / q)
+
+  return xc, yc
+
+
+def get_radius_function(X, Y):
+  m0 = np.average(Y[1:] - Y[:-1]) / np.average(X[1:] - X[:-1])
+  curvatures_X = []
+  curvatures_Y = []
+  for i in range(len(X) - 1):
+    m1 = (Y[i + 1] - Y[i]) / (X[i + 1] - X[i])
+
+    theta = np.arctan( (m1 - m0) / (1 + m1 * m0) )
+    k = np.sin(theta) / (X[i + 1] - X[i])
+
+    # if k >= 0:
+    curvatures_X.append((X[i + 1] + X[i]) / 2)
+    curvatures_Y.append(k)
+    
+  curvatures_Y = np.array(curvatures_Y)
+  coefs = poly.polyfit(curvatures_X, curvatures_Y, 0)
+  smooth_curvatures_Y = poly.polyval(curvatures_X, coefs)
+
[…]

### 2020-09-02 Papers/02_Pulses_Envelope/images/10xRobustnessTest.py

@@ -1,0 +1,267 @@
+import plotly.graph_objects as go
+import numpy as np
+from scipy.interpolate import interp1d
+import numpy.polynomial.polynomial as poly
+import wave
+from scipy.signal import savgol_filter
+import os
+from pathlib import Path
+import sys
+import argparse
+from plotly.subplots import make_subplots
+
+def read_wav(path): 
+  """returns signal & fps"""
+  wav = wave.open(path , 'r')
+  signal = np.frombuffer(wav.readframes(-1) , np.int16)
+  fps = wav.getframerate()
+  return signal, fps
+
+
+def get_circle(x0, y0, x1, y1, r):
+  '''returns center of circle that passes through two points'''
+  
+  radsq = r * r
+  q = np.sqrt((x1 - x0)**2 + (y1 - y0)**2)
+
+  # assert q <= 2 * r, f"shit"
+
+  x3 = (x0 + x1) / 2
+  y3 = (y0 + y1) / 2
+
+  if y0 + y1 >= 0:
+    xc = x3 + np.sqrt(radsq - (q / 2)**2) * ((y0 - y1) / q)
+    yc = y3 + np.sqrt(radsq - (q / 2)**2) * ((x1 - x0) / q)
+  else:
+    xc = x3 - np.sqrt(radsq - (q / 2)**2) * ((y0 - y1) / q)
+    yc = y3 - np.sqrt(radsq - (q / 2)**2) * ((x1 - x0) / q)
+
+  return xc, yc
+
+
+def get_radius_function(X, Y):
+  m0 = np.average(Y[1:] - Y[:-1]) / np.average(X[1:] - X[:-1])
+  curvatures_X = []
+  curvatures_Y = []
+  for i in range(len(X) - 1):
+    m1 = (Y[i + 1] - Y[i]) / (X[i + 1] - X[i])
+
+    theta = np.arctan( (m1 - m0) / (1 + m1 * m0) )
+    k = np.sin(theta) / (X[i + 1] - X[i])
+
+    # if k >= 0:
+    curvatures_X.append((X[i + 1] + X[i]) / 2)
+    curvatures_Y.append(k)
+    
+  curvatures_Y = np.array(curvatures_Y)
+  coefs = poly.polyfit(curvatures_X, curvatures_Y, 0)
+  smooth_curvatures_Y = poly.polyval(curvatures_X, coefs)
+
[…]

### 2020-09-04 Papers/02_Pulses_Envelope/PulsesEnvelope.tex
some progress with the paper

--- 2020-08-31 Papers/02_Pulses_Envelope/PulsesEnvelope.tex
@@
 % results
-We find that the algorithm accurate in the identification of the frontiers of a range of digital sound signal with very diverse characteristics, while providing the dominant frequency of the wave as a by-product.
+We find the algorithm accurate in the identification of the frontiers of a wide range of digital sound signal with very diverse characteristics, while providing the dominant frequency of the wave as a by-product.
 The algorithm also compares favourably with a common detection technique based on the Hilbert Transform.
-While the results are satisfactory per se, we define several geometric entities in this work, including a new measure of discrete curvature; we hope to consolidate theory in the area, while providing a solid basis for the development of new techniques, specially in sound synthesis.
+While the results are satisfactory per se, we define several geometric entities in this work, including a new measure of discrete curvature; hoping to consolidate theory in the area, while providing a solid basis for the development of new techniques, specially in sound synthesis and classification.
 \end{abstract}
@@
 % Opener sentence
-Despite being ubiquitous in both analog and digital signal processing, the literature about envelope detection is very fragmented \parencite{2017LyonsDigital}.
+Despite being ubiquitous in both analog and digital signal processing \parencite{2011CaetanoImproved}, the literature about envelope detection is very fragmented \parencite{2017LyonsDigital}.
 Besides, most envelope detection techniques are designed to account for very specific settings, like pure sinusoids with moderate noise contents, in line with the most common usages of those algorithms with artificial signals in the context of analog telecommunications; that limitation arises in part by the lack of a mathematical definition for a temporal envelope \parencite{2013Mengempirical}.
@@
 That is also the case for English \parencite{1995ShannonSpeech}, where even envelopes modulating mostly noise were still capable of conveying meaning.
+
 The envelope helps to convey emotion and identity to the human voice \parencite{2018ZhuContributions}, and envelope preserving characteristics of concert halls are associated with their pleasantness \parencite{2011LokkiEngaging}.
@@
 
+Interesting steps in the direction of translating geometric algorithms to the context of envelope detection were made by \textcite{2015YangSkeleton} with an approach based on the construction of a skeleton underlying the wave, and with the direct translation of computer vision methods to the task \parencite{2015YangRepresenting}.
+
 % Restate Your Question as Something Not Known or Fully Understood by Prior Research
@@
 % State the Significance of Your Question
-While the robustness of the proposed approach allows it to be used as a plug in replacement for many methods encountered in the literature, we feel that it would be particularly useful for sound synthesis.
-The envelope is shown to add complexity to the spectral representation of a wave \parencite{2019TarjanoNeuro}, and a accurate description of the envelope would be useful for a cleaner spectral analysis.
+While the robustness of the proposed approach allows it to be used as a plug in replacement for many methods encountered in the literature, we feel that it would be particularly useful for sound synthesis; the envelope is shown to add complexity to the spectral representation of a wave \parencite{2019TarjanoNeuro}, and a accurate description of the envelope would be useful for a cleaner spectral analysis.
 Moreover, the algorithm naturally divides a signal in pseudo-cycles, that could serve pragmatic building blocks for the reconstruction of the fine structure of the wave.
@@
 % State Your Claim | Objective | Hypothesis
-In the context of sound synthesis, one would be able to apply specific methods for the recreation of the envelope, more in line with its smooth, slow varying and non periodic nature and a different approach to the periodic and relatively fast changes of the temporal fine structure.
-% TODO brief outlook on the structure of the paper
+In the context of sound synthesis, for example, one would be able to apply specific methods for the recreation of the envelope, more in line with its smooth, slow varying and non periodic nature and a different approach to the periodic and relatively fast changes of the temporal fine structure.
+
+% brief outlook on the structure of the paper
+After explaining and defining the various entities used later, of which the concept of frontiers is arguably the most important, in the methodology section, we dedicate a section to present a new discrete curvature estimation approach.
+We proceed to illustrate the results of the application of the algorithm in a set of 6 diverse sound waves, ranging from voice to unpitched instruments.
+
+We then discuss the future implications of the work, emphasizing the focus on the construction of new interpretations of the relation between time domain and frequency domain sound descriptions, and our vision of the impact of this methodology in the field of sound synthesis.
+A range of potential future developments is also presented.
 
@@
 % Essential background information
+In this work it is assumed, without loss of generality, that one is interested in retrieving the mathematically ill defined \parencite{2013Mengempirical} temporal envelope of a arbitrary physical sound wave, of which only a recording is available.
+
+Naturally, due to the necessary truncations both in time and in amplitude, as well as measure errors introduced by the instruments used to record the signal, it is not possible to characterise exactly the original, continuous wave, from its digital counterpart.
+
+The best representation one can hope to get of the original wave is, then, the original recording.
+It is sensible, thus, to characterise the envelope of that wave in terms of a subset of the exact samples that comprise the digital representation.
+
+To that end, this work will emphasize the concept of frontier, as the points of the digital wave that mark the wave's upper and lower boundaries; from those points an envelope, conforming to the vague definitions of uniqueness and smoothness, can be trivially constructed.
+
 We start by defining a pulse as a series of consecutive samples in a discrete wave with the same sign.
-More formally, let $ W[i] \in \mathbb{R} \forall i \in \mathbb{N}_0, i < n \in \mathbb{N}_0 $ be a real, discrete and finite signal indexed, without loss of generality, over a subset of the natural numbers.
+More formally, let $ W[i] \in \mathbb{R} \ \forall \ i \in \mathbb{N}_0, \ i < n \in \mathbb{N}_0 $ be a real, discrete and finite signal indexed, without loss of generality, over a adjacent subset of the natural numbers starting at zero.
 This definition relates closely to the concept of an array or vector in most programming languages, and is used in the interest of simplicity.
@@
 
-In other words, each pulse is a unique subset of $ W $ with its members being the maximum sequence of samples in $ W $ with the same signal and $ P[0] \cup P[0] \cup \cdots \cup P[q-1] = W $.
-It is convenient to attribute to each pulse a pair of coordinates: those will be the coordinates of the sample with the maximum amplitude in each pulse.
+In other words, each pulse is a unique subset of $ W $ with its members being the maximum sequence of samples in $ W $ with the same signal and $ P[0] \cup P[1] \cup \cdots \cup P[q-1] = W $.
+It is convenient to attribute to each pulse a pair of coordinates: those will be the coordinates of the sample with the maximum absolute amplitude in each pulse, being denoted by $ P[j]_x $ and $ P[j]_y $; additionally, the length of a pulse will be denoted by $ P[j]_l $.
 
-Naturally, each pulse can be classified as positive or negative, based on the signal of its samples, which must be the same.
-For a continuous function, then, a pulse would correspond to the pieces between it's roots, and would be equal to half the cycle of a sinusoid.
+Naturally, each pulse can be classified as positive or negative, based on the sign of its samples, which must be the same.
+For a continuous function, then, a pulse would correspond to the pieces between its roots, and would be equal to half the cycle of a sinusoid.
 Thus, in our analogy, two consecutive pulses, naturally with opposing signs, have the potential do delimitate a pseudo-cycle, as illustrated in figure \ref{fig:Pulses}.
@@
 
-To do so we will define the positive (upper) frontier and negative (lower) frontier as the set of points $ (i, |W[i]|) $ that are, necessarily but not sufficiently, the point of maximum amplitude of their respective pulses.
+To do so we will define the positive (upper) frontier and negative (lower) frontier as the set of points $ (i, |W[i]|) $ that are, necessarily but not sufficiently, the points of maximum absolute amplitude of their respective pulses.
 The set of all these points such that $ W[i] > 0 $ will be called the positive, piecewise linear frontier and will be denoted by $ F_+ $; similarly, $ F_- $ will denote the negative frontier of $ W $, defined by the subset of points $ (i, W[i]) $ where $ W[i] < 0 $, as is shown in figure \ref{fig:Frontiers}.
@@
 To make abscissa and ordinate units consistent for the computation of the frontier, and proceed with a geometric treatment of the subject, we consider the set of positive and negative pulses separately.
-For each set, we multiply all pulses by the average length of the pulses in the set, after having divided by the average of the pulses' amplitudes, putting both axes in the same frame related unit $i$, related to time by the equation $i = t \ \mathit{fps}$.
+For each set, we multiply all pulses lengths by the average length of the pulses in the set, after having divided by the average of the pulses' amplitudes, putting both axes in the same unit $i$, related to time by the equation $i = t \ \mathit{fps}$.
+
+Formally, supposing we have $ m $ pulses with the same sign, we scale each pulse's vertical coordinate by $s$, where:
+
+\begin{equation} \label{eq:scale}
+  s = \frac{\sum_{i=1}^{m} P[i]_l}{\sum_{i=1}^{m} P[i]_y}
+\end{equation}
 
@@
 
-Since an edge defines only one direction, there remains the need for a reference.
-As we are interested in the relative curvature, it is natural to use the average direction of all vectors, $ \vec{\overline{v}}^\pm = \frac{1}{m-1} \sum_{l=0}^{m-1} \vec{v}_l^\pm $, so as to measure the relative change of direction of each edge.
+Since an edge defines only one direction, there remains the need for a direction of reference.
+As we are interested in the relative curvature, it is natural to use the average direction of all positive or negative vectors, $ \vec{\overline{v}}^\pm = \frac{1}{m-1} \sum_{l=0}^{m-1} \vec{v}_l^\pm $, so as to measure the relative change of direction of each edge.
 This procedure, as well as the circles generated by it, are illustrated in figure \ref{fig:Curvatures}.
 
-The angle $ \theta $ between the average direction and each of the vectors can be obtained via the relation of the slope of the respective lines, as in the expression $ \theta = \arctan( (m - \overline{m}) / (1 + m * \overline{m}) ) $.
+The angle $ \theta $ between the average direction and each of the vectors can be obtained via the relation between the slopes of the respective lines, as in the expression $ \theta = \arctan( (m - \overline{m}) / (1 + m \overline{m}) ) $.
 The radii of the equivalent circles are thus given by $ r_l = x_l / \sin (\theta_l) $ with the curvature given by $\kappa_l =1/r_l=\sin (\theta_l) / x_l$.
+After some algebra, one can write:
+
+\begin{equation} \label{eq:radius}
+  r_l = \left\lvert \frac{x_l \sqrt{\overline{m}^2 + 1} \ \sqrt{x_l^2 + y^2}}{\overline{m} x_l - y} \right\rvert
+\end{equation}
 
@@
 
-% Therefore, the definition of the curvature becomes $ \kappa(x) = \frac{2 a}{\big( (b + 2 a x)^2 + 1 \big)^\frac{3}{2}} $ with the average curvature between points $ x_0 $ and $ x_1 $ given by equation \ref{eq:1}.
+\section{Results}
 
-% \begin{equation} \label{eq:1}
-%   \overline{\kappa}(x) = 
-%   \frac{
-%       \left( \dfrac{2 a x_1 + b}{\sqrt{\big(2 a x_1 + b \big)^2+1}} - \dfrac{2 a x_0 + b}{\sqrt{\big(2 a x_0 + b \big)^2+1}} \right) 
-%   }{
-%     (x_1 - x_0)
-%   }
-% \end{equation}
-
-% \subsection{Theory}
-
-\section{Results}
-% TODO
-% Revisit purpose and methods
-% Overview of results
 Figure \ref{fig:FullFrontiers} illustrates the frontiers obtained to six diverse discrete sound waves, as well as a detail view of a shorter segment for each wave.
@@
     \includegraphics[width=0.8\linewidth]{07FullFrontiers.pdf}
-  \caption{Positive and negative frontiers of 6 digital waves.
-For each wave, the region highlighted in black is shown in detail.
+  \caption{Positive and negative frontiers of 6 digital waves, as extracted by the algorithm here presented.
+For each wave, the region highlighted in black is shown in detail besides the whole wave.
 For each wave the horizontal axis is the sample number $i$, while the vertical axis is the normalized amplitude.
@@
 
-\subsection{Fundamental Frequency as a Welcome Side-effect}
+Figures \ref{fig:xRobustness} and \ref{fig:yRobustness} illustrate the computational resilience of the algorithm for scaling in both the abscissa and ordinate axes, respectively.
+This is due to the normalization scheme presented in equation \ref{eq:scale}.
+
+Although important from a theoretical standpoint, this step has a special computational relevance in the case of digital sound waves, as they tend to be composed of a high number of samples in moderate amplitudes, generally normalized between -1 and 1, leading to a numerical disparity between axes.
+
+\begin{figure}[ht!]
+  \centering
+    \includegraphics[width=0.8\linewidth]{10xRobustness.pdf}
+  \caption{Minimum and maximum scales of the abscissa before computational stability issues arose, for a recording of a drumkit tom.  $ \text{B} = 10^6 $.}
+  \label{fig:xRobustness}
+\end{figure}
+
+\begin{figure}[ht!]
+  \centering
+    \includegraphics[width=0.8\linewidth]{09yRobustness.pdf}
+  \caption{Minimum and maximum values of the amplitude before computational stability issues arose, for a recording of a drumkit tom.}
+  \label{fig:yRobustness}
+\end{figure}
+
 % Comparison with results in other works
-% Comparison with predictions
-% Problems with results
-% Possible implications
+Direct comparison with other algorithms is made difficult by the unavailability of digital implementations of such works, many designed to process analog signals \parencite[e.g.,][]{2018AssefModeling}.
 
-\section{Discussion + Conclusion}
+\section{Discussion}
+
+This work hopes to fill a gap identified by the authors in the context of sound synthesis, where the lack of procedures for the accurate identification of the temporal envelope of arbitrary waves consists of an obstacle to the complete description, and eventual manipulation of signals.
+
+While relevant in its own accord, the procedure here presented isolates the individual cycles of a wave, pinpointing them in the time domain.
+Due to that, a rich set of theoretical advancements, of whose an example was given in the preceding section, can be built upon this initial building block.
+
+With the knowledge of the exact position and characteristics of the cycles of a wave, one can investigate its evolution in time, in terms of frequency content for example, free from the potencial interference of a modulating wave; in other words, one can have a very satisfactory view of the instantaneous frequencies and phase of an arbitrary wave and its evolution in time, providing a way to bridge the gap between time domain and frequency domain interpretations.
 
 \section{References}
+
 \printbibliography[heading=none]

### 2020-09-05 Papers/02_Pulses_Envelope/PulsesEnvelope.tex
finished paper

@@
 To do so we will define the positive (upper) frontier and negative (lower) frontier as the set of points $ (i, |W[i]|) $ that are, necessarily but not sufficiently, the points of maximum absolute amplitude of their respective pulses.
-The set of all these points such that $ W[i] > 0 $ will be called the positive, piecewise linear frontier and will be denoted by $ F_+ $; similarly, $ F_- $ will denote the negative frontier of $ W $, defined by the subset of points $ (i, W[i]) $ where $ W[i] < 0 $, as is shown in figure \ref{fig:Frontiers}.
+The set of all these points such that $ W[i] > 0 $ will be called the positive, piecewise linear frontier and will be denoted by $ F^+ $; similarly, $ F^- $ will denote the negative frontier of $ W $, defined by the subset of points $ (i, W[i]) $ where $ W[i] < 0 $, as is shown in figure \ref{fig:Frontiers}.
 To identify such points, we need to address the different units in which the amplitude and horizontal distance of our wave are expressed.
@@
 
-% Key results in details
+Besides, the algorithm presents a definition of the position of the cycles of a wave, with which on can, for example, derive its fundamental frequency and phase: using one of the two frontiers, we have $ f = (\#F^\pm - 1) n / (F^\pm[-1] - F^\pm[0])$, where $ \#F^\pm $ denotes the cardinality of the frontier, while $ F^\pm[0] $ and $ F^\pm[-1] $ stand for the frontier's first and last item, respectively.
+
+Similarly, each item of each of the frontiers can be used to estimate the local phase at that location.
+For the positive frontier, it is necessary to note that, for each item, the following equation holds: $ \cos(\phi + 2 \pi f i / n) = 1 $, and thus $ \phi = k 2 \pi - 2 \pi f i / n $, with $ k \in \mathbb{N } $, while for the negative frontier $ \cos(\phi + 2 \pi f i / n) = -1 $ and $ \phi = k 3 \pi - 2 \pi f i / n) $.
 
@@
 
-% Comparison with results in other works
-Direct comparison with other algorithms is made difficult by the unavailability of digital implementations of such works, many designed to process analog signals \parencite[e.g.,][]{2018AssefModeling}.
+\subsection{Comparison with traditional algorithms}
+
+Direct comparison with many of the more recent algorithms is made difficult by the unavailability of digital implementations of such works, many designed to process analog signals \parencite[e.g.,][]{2018AssefModeling}.
+
+Nevertheless, insight can be gained from a comparison of the results of the method here proposed with some of the most common envelope extraction algorithms, as cane be seen in figure \ref{fig:Comparison}.
+
+\begin{figure}[ht!]
+  \centering
+    \includegraphics[width=0.8\linewidth]{12Comparison.pdf}
+  \caption{Comparison of the algorithm present in this work with the most common methods of digital envelope identification.}
+  \label{fig:Comparison}
+\end{figure}
+
+For a recording of a guitar bend, we contrast the proposed algorithm with four common approaches of varied complexity: simple smoothing of the rectified original wave, lowpass filtering followed by peak identification and subsequent smoothing, and the extraction of the envelope via Hilbert transform, as presented earlier in the paper, with a previous smoothing of the original wave to avoid the artefacts shown in figure \ref{fig:AnalyticSignal}.
+
+All those methods demanded careful parameter tuning: in the case of filtering, a local cut-off frequency of approximately 88 Hz was used, consisting of one additional parameter to be tuned.
+For all the other traditional approaches, the Savitzky-Golay smoothing algorithm was used, and the window size was the parameter to be defined.
+
+In the case of the proposed algorithm, the smooth temporal envelope was obtained merging both positive and negative frontiers, that were also smoothed with the same procedure described.
+Figure \ref{fig:FrontiersToEnvelope} illustrates this process.
+
+\begin{figure}[ht!]
+  \centering
+    \includegraphics[width=0.8\linewidth]{13FrontiersToEnvelope.pdf}
+  \caption{Original positive and negative frontiers in dark gray and their mirrored versions. The temporal envelope, in black, is obtained by merging and approximating both frontiers.}
+  \label{fig:FrontiersToEnvelope}
+\end{figure}
+
+Table \ref{table:Comparison} presents some computational results.
+The times are for custom implementations of those methods in the Python programming language.
+The least mean squares was calculated halving the envelopes presented in figure \ref{fig:Comparison} and comparing then with the rectified absolute wave, so as to avoid the potential influence of an estimated frequency.
+We can see that, while the algorithm here presented is the most accurate, it's also the most computationally intensive.
+The connection between alpha-shapes and Delaunay triangulation can be explored in future works as a mean to improve the efficiency of the algorithm.
+
+\begin{table}[ht!]
+\centering
+\begin{tabular}{ l c c }
+\hline
+Method & LMS & time(s) \\
+\hline
+Presented Algorithm & 0.0088 & 4.1507 \\ 
+Smoothing           & 0.0123 & 0.1911 \\ 
+Lowpass Filter      & 0.0258 & 0.5601 \\ 
+Hilbert Transform   & 0.0141 & 0.2599 \\ 
+\hline
+\end{tabular}
+\caption{Comparison of the algorithm present in this work with the most common methods of digital envelope identification.}
+\label{table:Comparison}
+\end{table}
 

### 2020-09-05 Papers/02_Pulses_Envelope/images/12Comparison.py

@@ -1,0 +1,238 @@
+import plotly.graph_objects as go
+import numpy as np
+from scipy.interpolate import interp1d
+import numpy.polynomial.polynomial as poly
+import wave
+from scipy.signal import savgol_filter, hilbert
+import os
+from pathlib import Path
+import sys
+sys.path.append("c:/Users/tesse/Desktop/Files/Dropbox/0_Thesis/Python")
+from Wheel import frontiers
+from plotly.subplots import make_subplots
+from timeit import default_timer as timer
+
+
+
+def read_wav(path): 
+  """returns signal & fps"""
+  wav = wave.open(path , 'r')
+  signal = np.frombuffer(wav.readframes(-1) , np.int16)
+  fps = wav.getframerate()
+  return signal, fps
+
+
+
+
+'''==============='''
+''' Read wav file '''
+'''==============='''
+name = "bend"
+parent_path = str(Path(os.path.abspath('./')).parents[1])
+path = f"{parent_path}/Python/Samples/{name}.wav"
+print(path)
+
+W, fps = read_wav(path)
+
+W = W - np.average(W)
+amplitude = np.max(np.abs(W))
+W = W / amplitude
+n = W.size
+print(f"n={n}")
+X = np.arange(n)
+
+# '''==============='''
+# '''    Sinusoid   '''
+# '''==============='''
+# FT = np.fft.rfft(W)# * 2 / n
+# PW = np.abs(FT)
+# f = np.argmax(PW)
+# p = np.angle(FT[f])
+
+# S = np.cos(p + 2 * np.pi * f * X / n)
+
+'''==============='''
+'''    Snowball   '''
+'''==============='''
+start = timer()
+pos_frontierX, pos_frontierY, neg_frontierX, neg_frontierY = frontiers(W)
+
+frontierX = np.concatenate([pos_frontierX, neg_frontierX])
+frontierY = np.concatenate([pos_frontierY, np.abs(neg_frontierY)])
+
+idxs = np.argsort(frontierX)
+frontierX = frontierX[idxs]
+frontierY = frontierY[idxs]
+
+f = interp1d(frontierX, frontierY, kind="linear", fill_value="extrapolate", assume_sorted=False)
+
+smooth_frontierY = savgol_filter(f(X), 5000 + 1, 2)
+
+lms = np.average((np.abs(W) - smooth_frontierY * 0.5)**2)
+
[…]

### 2020-09-05 Papers/02_Pulses_Envelope/images/13FrontiersToEnvelope.py

@@ -1,0 +1,172 @@
+import plotly.graph_objects as go
+import numpy as np
+from scipy.interpolate import interp1d
+import numpy.polynomial.polynomial as poly
+import wave
+from scipy.signal import savgol_filter, hilbert
+import os
+from pathlib import Path
+import sys
+sys.path.append("c:/Users/tesse/Desktop/Files/Dropbox/0_Thesis/Python")
+from Wheel import frontiers
+from plotly.subplots import make_subplots
+from timeit import default_timer as timer
+
+
+
+def read_wav(path): 
+  """returns signal & fps"""
+  wav = wave.open(path , 'r')
+  signal = np.frombuffer(wav.readframes(-1) , np.int16)
+  fps = wav.getframerate()
+  return signal, fps
+
+
+
+
+'''==============='''
+''' Read wav file '''
+'''==============='''
+name = "bend"
+parent_path = str(Path(os.path.abspath('./')).parents[1])
+path = f"{parent_path}/Python/Samples/{name}.wav"
+print(path)
+
+W, fps = read_wav(path)
+
+W = W - np.average(W)
+amplitude = np.max(np.abs(W))
+W = W / amplitude
+n = W.size
+print(f"n={n}")
+X = np.arange(n)
+
+# '''==============='''
+# '''    Sinusoid   '''
+# '''==============='''
+# FT = np.fft.rfft(W)# * 2 / n
+# PW = np.abs(FT)
+# f = np.argmax(PW)
+# p = np.angle(FT[f])
+
+# S = np.cos(p + 2 * np.pi * f * X / n)
+
+'''==============='''
+'''    Snowball   '''
+'''==============='''
+start = timer()
+pos_frontierX, pos_frontierY, neg_frontierX, neg_frontierY = frontiers(W)
+
+frontierX = np.concatenate([pos_frontierX, neg_frontierX])
+frontierY = np.concatenate([pos_frontierY, np.abs(neg_frontierY)])
+
+idxs = np.argsort(frontierX)
+frontierX = frontierX[idxs]
+frontierY = frontierY[idxs]
+
+f = interp1d(frontierX, frontierY, kind="linear", fill_value="extrapolate", assume_sorted=False)
+
+smooth_frontierY = savgol_filter(f(X), 5000 + 1, 2)
+
+lms = np.average((np.abs(W) - smooth_frontierY * 0.5)**2)
+
+envX = [x for x in X]
[…]

### 2020-09-06 Papers/02_Pulses_Envelope/PulsesEnvelope.tex
Paper version 1.0
first full revision

--- 2020-09-04 Papers/02_Pulses_Envelope/PulsesEnvelope.tex
@@
 \usepackage{amsmath}
-\usepackage{amssymb} % blackbox symbol
 \usepackage{hyperref}
@@
 
-\begin{abstract} % 250 words
-% background  
-Despite being an elusive concept, the temporal amplitude envelope of a signal is essential for its complete characterization, being the primary information carrying medium in spoken voice and telecommunications, for example.
-intuitively, the temporal envelope can be understood as a smooth function, that modulates the signal, being responsible for its outer shape.
-% motivation
+\begin{abstract}
+Despite being an elusive concept, the temporal amplitude envelope of a signal is essential for its complete characterization, being the primary information-carrying medium in spoken voice and telecommunications, for example.
+Intuitively, the temporal envelope can be understood as a smooth function that multiplies the signal, being responsible for its outer shape.
 Envelope detection techniques have applications in areas like health, sound classification and synthesis, seismology and speech recognition.
-Despite of that, a general approach to envelope detection of signals with rich spectral content seems to be lacking, and most methods involve manual intervention, like filter design or smoothing, based on prior knowledge about the wave under investigation.
-% objective
-In this paper we propose a framework that uses intrinsic characteristics of a signal to estimate its envelope, completely eliminating the necessity of parameter tuning.
-% methods
-The approach here described draws inspiration from general, computational and differential geometry to isolate the frontiers and thus estimate the temporal envelope of an arbitrary signal; to that end The concepts of alpha-shapes, concave hulls, and discrete curvature are explored.
-We also define a pulse in the context of a arbitrary digital signal as a means to reduce dimensionality and the complexity of the proposed algorithm.
-% results
-We find the algorithm accurate in the identification of the frontiers of a wide range of digital sound signal with very diverse characteristics, while providing the dominant frequency of the wave as a by-product.
-The algorithm also compares favourably with a common detection technique based on the Hilbert Transform.
-While the results are satisfactory per se, we define several geometric entities in this work, including a new measure of discrete curvature; hoping to consolidate theory in the area, while providing a solid basis for the development of new techniques, specially in sound synthesis and classification.
+Nevertheless, a general method to digital envelope detection of signals with rich spectral content doesn't exist, as most methods involve manual intervention, in the form of filter design, smoothing, as well as other specific design choices, based on a priori knowledge about the nature of the specific waves under investigation.
+To address this problem, we propose a framework that uses intrinsic characteristics of a signal to estimate its envelope, completely eliminating the necessity of parameter tuning.
+The approach here described draws inspiration from geometric concepts to isolate the frontiers and thus estimate the temporal envelope of an arbitrary signal; to that end, alpha-shapes, concave hulls, and discrete curvature are explored.
+We also define entities, such as a pulse and frontiers, in the context of an arbitrary digital signal, as a means to reduce dimensionality and the complexity of the proposed algorithm.
+Specifically, a new measure of discrete curvature is used to obtain the average radius of curvature of a discrete wave, serving as a threshold to identify the wave's frontier points.
+We find the algorithm accurate in the identification of the frontiers of a wide range of digital sound waves with very diverse characteristics, while localizing each pseudo-cycle of the wave in the time domain.
+The algorithm also compares favourably with classic envelope detection techniques based on filtering and the Hilbert Transform.
+Besides the most direct applications of this work to audio classification and synthesis, we foresee impact in compression techniques and machine learning approaches to audio.
+The discrete curvature definition presented could also be extended to three-dimensional settings, to improve shape detection algorithms based on alpha-shapes.
 \end{abstract}
 
-{\bf Keywords:} DSP, alpha-shapes, envelope detection
+{\bf Keywords:} DSP, alpha-shapes, envelope detection, discrete curvature
 
-\section{Introduction} % 500 words
-% Opener sentence
-Despite being ubiquitous in both analog and digital signal processing \parencite{2011CaetanoImproved}, the literature about envelope detection is very fragmented \parencite{2017LyonsDigital}.
-Besides, most envelope detection techniques are designed to account for very specific settings, like pure sinusoids with moderate noise contents, in line with the most common usages of those algorithms with artificial signals in the context of analog telecommunications; that limitation arises in part by the lack of a mathematical definition for a temporal envelope \parencite{2013Mengempirical}.
+\section{Introduction}
+Envelope detection is ubiquitous in both analogue and digital signal processing \parencite{2011CaetanoImproved}.
+Nevertheless, the literature in this area is very fragmented \parencite{2017LyonsDigital}.
+Besides, most envelope detection techniques are designed to account for very specific settings, like pure sinusoids with moderate noise content, a limitation that excludes most physical signals, as is the case of recorded sound, for examples; that limitation arises in part by the lack of a strict mathematical definition for a temporal envelope \parencite{2013Mengempirical}.
 
-% Literature review | Brief Context of Prior Research
-In many natural signals, however, the temporal amplitude envelope of a signal plays a prominent role in the characteristics exhibited: According to \textcite{2017QiRelative}, for example, the envelope is at least as important as the fine structure of the soundwave in the context of the intelligibility of mandarin tones.
+In many natural signals, however, the temporal amplitude envelope of a signal plays a prominent role in the characteristics exhibited: according to \textcite{2017QiRelative}, for example, the envelope is at least as important as the fine structure of a soundwave in the context of the intelligibility of mandarin tones.
 That is also the case for English \parencite{1995ShannonSpeech}, where even envelopes modulating mostly noise were still capable of conveying meaning.
 
-The envelope helps to convey emotion and identity to the human voice \parencite{2018ZhuContributions}, and envelope preserving characteristics of concert halls are associated with their pleasantness \parencite{2011LokkiEngaging}.
+The envelope helps to impart emotion and identity to the human voice \parencite{2018ZhuContributions}, and envelope preserving characteristics in concert halls are associated with their pleasantness \parencite{2011LokkiEngaging}.
 
-% Cite current algorithms
 When dealing with broadband signals, approaches tailored to specific applications are prevalent, such as the one presented by \textcite{2014YangFast} for the distributed monitoring of fibre optic or the one formulated by \textcite{2018AssefModeling} in the context of medical ultrasound imaging.
 
-If one adresses the different units in the horizontal and vertical axes, one can transform the DSP problem of envelope detection in the geometric problem of defining the shape of a set of points in $ \mathbb{R}^2 $.
-For that purpose \textcite{1983Edelsbrunnershape} introduced the concept of alpha-shapes, a mathematically well defined extension to the convex hull of a finite set of points, related to the Delaunay triangulation and Voronoi diagrams of those points.
+Addressing the different units in the horizontal and vertical axes, one can transform the DSP problem of envelope detection in the geometric problem of defining the shape of a set of points in $ \mathbb{R}^2 $.
+For that purpose \textcite{1983Edelsbrunnershape} introduced the concept of alpha-shapes, a mathematically well-defined extension to the convex hull of a finite set of points, closely related to the Delaunay triangulation and Voronoi diagrams of those points.
 
-This approach is used in areas such as the detection of features in images \parencite{2016VarytimidisAlpha}, reconstruction of surfaces from a cloud of points \parencite{2015WuAutomated} and Spectroscopy \parencite{2019XuModeling}, with the last work, that involves the estimation and removal of the Blaze function (a kind of envelope) of a echelle spectrograph, being particularly align with what is intended here.
+This approach is used in areas such as the detection of features in images \parencite{2016VarytimidisAlpha}, reconstruction of surfaces from a cloud of points \parencite{2015WuAutomated} and Spectroscopy \parencite{2019XuModeling}, with the last work, which involves the estimation and removal of the Blaze function (a kind of envelope) of an echelle spectrograph, being particularly illustrative of the potential of cross-fertilization.
 
-Interesting steps in the direction of translating geometric algorithms to the context of envelope detection were made by \textcite{2015YangSkeleton} with an approach based on the construction of a skeleton underlying the wave, and with the direct translation of computer vision methods to the task \parencite{2015YangRepresenting}.
+Steps in the direction of translating geometric algorithms to the context of envelope detection were also made by \textcite{2015YangSkeleton} with an approach based on the construction of a skeleton underlying the wave, and with the direct translation of computer vision methods to the task \parencite{2015YangRepresenting}.
 
-% Restate Your Question as Something Not Known or Fully Understood by Prior Research
-Thus, in this paper we formulate a general approach to envelope detection, exploiting the intrinsic characteristics of a generic, spectrally complex, wave in order to avoid the need to manual intervention or parameter tuning.
+Following this path, we present a general approach to envelope detection, exploiting the intrinsic characteristics of a generic, spectrally complex wave in order to avoid the need for manual intervention or parameter tuning.
 
-% State the Significance of Your Question
-While the robustness of the proposed approach allows it to be used as a plug in replacement for many methods encountered in the literature, we feel that it would be particularly useful for sound synthesis; the envelope is shown to add complexity to the spectral representation of a wave \parencite{2019TarjanoNeuro}, and a accurate description of the envelope would be useful for a cleaner spectral analysis.
-Moreover, the algorithm naturally divides a signal in pseudo-cycles, that could serve pragmatic building blocks for the reconstruction of the fine structure of the wave.
+While the robustness of the proposed approach allows it to be used as a plug-in replacement for many methods encountered in the literature, we feel that it would be particularly useful for sound synthesis; the envelope is shown to add complexity to the spectral representation of a wave \parencite{2019TarjanoNeuro}, and an accurate description of the envelope, while solving the problem of the evolution of the instantaneous amplitude of a signal in time, would also greatly simplify further spectral analysis.
+Moreover, the algorithm naturally divides a signal in its pseudo-cycles, pinpointing them in the time domain, providing the building blocks for the reconstruction of the fine structure of the wave.
 
-% State Your Claim | Objective | Hypothesis
-In the context of sound synthesis, for example, one would be able to apply specific methods for the recreation of the envelope, more in line with its smooth, slow varying and non periodic nature and a different approach to the periodic and relatively fast changes of the temporal fine structure.
+The rest of this paper is developed as follows: After explaining and defining the various entities used subsequently, of which the concept of frontiers is arguably the most important, in the methodology section, we dedicate a section to present a new discrete curvature estimation approach.
 
-% brief outlook on the structure of the paper
-After explaining and defining the various entities used later, of which the concept of frontiers is arguably the most important, in the methodology section, we dedicate a section to present a new discrete curvature estimation approach.
-We proceed to illustrate the results of the application of the algorithm in a set of 6 diverse sound waves, ranging from voice to unpitched instruments.
+We then proceed to illustrate the results of the application of the algorithm in a set of six diverse sound waves, ranging from voice to unpitched instruments comparing, afterwards, its performance with that of some classic digital envelope estimation algorithms, both from a visual and numerical standpoint.
 
-We then discuss the future implications of the work, emphasizing the focus on the construction of new interpretations of the relation between time domain and frequency domain sound descriptions, and our vision of the impact of this methodology in the field of sound synthesis.
-A range of potential future developments is also presented.
+We then discuss the implications of the work, emphasizing the focus on the construction of new interpretations of the relationship between time domain and frequency domain sound descriptions, and our vision of the impact of this methodology in the field of sound synthesis.
 
-\section{Methodology} % 1000 words
+\section{Methodology}
 
-% Essential background information
-In this work it is assumed, without loss of generality, that one is interested in retrieving the mathematically ill defined \parencite{2013Mengempirical} temporal envelope of a arbitrary physical sound wave, of which only a recording is available.
+In this work it is assumed, without loss of generality, that one is interested in retrieving the mathematically ill-defined \parencite{2013Mengempirical} temporal envelope of an arbitrary physical sound wave, of which only a digital representation is available.
 
-Naturally, due to the necessary truncations both in time and in amplitude, as well as measure errors introduced by the instruments used to record the signal, it is not possible to characterise exactly the original, continuous wave, from its digital counterpart.
+Naturally, due to the necessary truncations both in time and amplitude, as well as measurement errors introduced by the instruments used to record the signal, it is not possible to exactly recover the original, continuous wave, from its digital counterpart.
 
-The best representation one can hope to get of the original wave is, then, the original recording.
+The best representation one can hope to get of the original wave is, then, the recording.
 It is sensible, thus, to characterise the envelope of that wave in terms of a subset of the exact samples that comprise the digital representation.
@@
 We start by defining a pulse as a series of consecutive samples in a discrete wave with the same sign.
-More formally, let $ W[i] \in \mathbb{R} \ \forall \ i \in \mathbb{N}_0, \ i < n \in \mathbb{N}_0 $ be a real, discrete and finite signal indexed, without loss of generality, over a adjacent subset of the natural numbers starting at zero.
+More formally, let $ W[i] \in \mathbb{R} \ \forall \ i \in \mathbb{N}_0, \ i < n \in \mathbb{N}_0 $ be a real, discrete and finite signal indexed, without loss of generality, over an adjacent subset of the natural numbers starting at zero.
 This definition relates closely to the concept of an array or vector in most programming languages, and is used in the interest of simplicity.
@@
 In other words, each pulse is a unique subset of $ W $ with its members being the maximum sequence of samples in $ W $ with the same signal and $ P[0] \cup P[1] \cup \cdots \cup P[q-1] = W $.
-It is convenient to attribute to each pulse a pair of coordinates: those will be the coordinates of the sample with the maximum absolute amplitude in each pulse, being denoted by $ P[j]_x $ and $ P[j]_y $; additionally, the length of a pulse will be denoted by $ P[j]_l $.
+It is convenient to attribute to each pulse a pair of coordinates: those will be the coordinates of the sample with the maximum absolute amplitude in each pulse, being denoted by $ P[j]_x $ and $ P[j]_y $; additionally, the length, or cardinality, of a pulse will be denoted by $ \#P[j] $.
 
-Naturally, each pulse can be classified as positive or negative, based on the sign of its samples, which must be the same.
-For a continuous function, then, a pulse would correspond to the pieces between its roots, and would be equal to half the cycle of a sinusoid.
-Thus, in our analogy, two consecutive pulses, naturally with opposing signs, have the potential do delimitate a pseudo-cycle, as illustrated in figure \ref{fig:Pulses}.
-For that to be true, however, the pulses need to make part of the frontiers of the signal, as will be defined later.
+Naturally, each pulse can be classified as positive ($ P[j]^+ $) or negative($ P[j]^- $), based on the sign of its samples, which must be the same.
+For an arbitrary continuous function, then, a pulse would correspond to the pieces between its roots, and would be equal to half the cycle of a sinusoid.
+Thus, in our definition, two consecutive pulses, naturally with opposing signs, have the potential to delimitate a pseudo-cycle, as illustrated in figure \ref{fig:Pulses}.
+For that to be true, however, the pulses need to make part of the frontiers of the signal, as yet to be defined.
 
@@
 The most mathematically sound definition of an envelope involves its representation as an analytic signal, introduced by Gabor in 1946 \parencite{2007HahnHistory}.
-In his work, \textcite{1946GaborTheory} applies the then relatively new mathematical machinery of the quantum mechanics to unify time and frequency representations of a wave, showing how the Hilbert transform could be applied to a real signal in order to obtain a complex signal that would become known as the analytic signal.
+In his work, \textcite{1946GaborTheory} applies the then relatively new mathematical machinery of the quantum mechanics to unify time and frequency representations of a wave, showing how the Hilbert transform could be applied to a real signal in order to obtain an equivalent complex signal, later known as the analytic signal.
 
@@
 $ \mathbb{H}(s(t)) $, the Hilbert transform of the original signal becomes, then, the imaginary part of the analytic signal; the envelope of a signal thus represented can be straightforwardly obtained by the computation of its complex modulus.
+
+Figure \ref{fig:AnalyticSignal} shows the envelope obtained via the Hilbert transform, for a pure sinusoid with a local frequency of 20 cycles, modulated by a polynomial of the third degree, in the presence and absence of gaussian white noise with standard deviation of $ 1/10 $ of the wave's maximum amplitude.
 
@@
   \caption{Envelope of a pure sinusoid modulated by a polynomial of degree 3, as obtained by the Hilbert Transform.
-In the first half the sinusoid is free of noise, while in the second half white gaussian noise with a standard deviation of $ 1/10 $ of the maximum amplitude of the wave was added to the base signal.}
+In the first half, the sinusoid is free of noise, while in the second half white gaussian noise with a standard deviation of $ 1/10 $ of the maximum amplitude of the wave was added to the base signal.}
   \label{fig:AnalyticSignal}
@@
 
-Figure \ref{fig:AnalyticSignal} shows the envelope obtained via the Hilbert transform, for a pure sinusoid of local frequency 20, modulated by a polynomial of the third degree, in the presence and absence of gaussian white noise with standard deviation of $ 1/10 $ of the wave's maximum amplitude.
-It is readily noticeable that, as soon as noise is introduced in the original signal, it is reflected in the envelope, breaking our assumption of smoothness.
+It is readily noticeable from figure \ref{fig:AnalyticSignal} that, as soon as noise is introduced in the original signal, it is reflected in the envelope, breaking our assumption of smoothness.
 Other than that, this definition yields a unique, positive envelope.
+In practice, however, is not uncommon for a discrete wave, especially in the case of sound, to present somewhat different positive and negative contours; as our method will forcibly have to be applied separately to both sides, superior and inferior, of the discrete wave, it is only natural to introduce such separation as soon as possible.
 
-In practice, however, is not uncommon for a discrete wave, specially in the case of sound, to present somewhat different positive and negative contours; as our method will forcibly have to be applied separately to both sides, superior and inferior, of the discrete wave, it is only natural to introduce such separation as soon as possible.
+To do so we will define the positive (upper) frontier and negative (lower) frontiers of a discrete wave as the set of points $ (i, |W[i]|) $ that are, necessarily but not sufficiently, the points of maximum absolute amplitude of their respective pulses.
 
-To do so we will define the positive (upper) frontier and negative (lower) frontier as the set of points $ (i, |W[i]|) $ that are, necessarily but not sufficiently, the points of maximum absolute amplitude of their respective pulses.
-The set of all these points such that $ W[i] > 0 $ will be called the positive, piecewise linear frontier and will be denoted by $ F_+ $; similarly, $ F_- $ will denote the negative frontier of $ W $, defined by the subset of points $ (i, W[i]) $ where $ W[i] < 0 $, as is shown in figure \ref{fig:Frontiers}.
+The set of all these points such that $ W[i] > 0 $ will be called the positive, piecewise linear frontier and will be denoted by $ F^+ $; similarly, $ F^- $ will denote the negative frontier of $ W $, defined by the subset of points $ (i, W[i]) $ where $ W[i] < 0 $, as is shown in figure \ref{fig:Frontiers}.
+
 To identify such points, we need to address the different units in which the amplitude and horizontal distance of our wave are expressed.
@@
 
-To make abscissa and ordinate units consistent for the computation of the frontier, and proceed with a geometric treatment of the subject, we consider the set of positive and negative pulses separately.
+To make abscissa and ordinate units consistent for the computation of the frontiers, and proceed with a geometric treatment of the subject, we consider the set of positive and negative pulses separately.
+
 For each set, we multiply all pulses lengths by the average length of the pulses in the set, after having divided by the average of the pulses' amplitudes, putting both axes in the same unit $i$, related to time by the equation $i = t \ \mathit{fps}$.
 
-Formally, supposing we have $ m $ pulses with the same sign, we scale each pulse's vertical coordinate by $s$, where:
+Formally, supposing we have $ m $ pulses with the same sign, we scale each pulse's vertical coordinate $ P[j]^\pm_y $ by $ s^\pm $, where:
 
 \begin{equation} \label{eq:scale}
-  s = \frac{\sum_{i=1}^{m} P[i]_l}{\sum_{i=1}^{m} P[i]_y}
+  s^\pm = \frac{\sum_{j=1}^{m} \#P[j]^\pm}{\sum_{j=1}^{m} P[j]^\pm_y}
 \end{equation}
@@
 By the same token, as the change in amplitude from one pulse to the next pulse with the same sign is motivated by changes in this same envelope, one can reasonably suppose that $ \max(P[j]) \approx \max(P[j+2]) $.
-That justifies the approximation of the convex hull of a pulse by the triangle defined by the points where the pulse crosses the horizontal the point of maximum amplitude; this will allow a considerable dimension reduction as this triangular pulse can become the atomic unit for the rest of the method.
+
+That justifies the approximation of the convex hull of a pulse by the triangle defined by the points where the pulse crosses the horizontal axis and the point of maximum amplitude; this will allow a considerable dimension reduction as this triangular pulse can become the atomic unit for the rest of the method.
 
@@
 Translating the intuitive explanation of alpha-shapes in \textcite{1994EdelsbrunnerThree} to our context, we can say that the points in the frontier are those touched by a circle outside the signal that is not allowed to contain any point of the signal; The pulses that contain said points can be seen as frontier pulses.
+
 Intuitively, one can picture a circle being rolled above (or below, in the case of the negative frontier) the signal, and marking the points it touches as frontier points.
 
-We can therefore revisit the concept of a pseudo-cycle, only hinted at before, defining it as the samples of the original wave that belong to two successive pulses that belong to their respective frontiers, including eventual pulses between them, as can be seen in figure \ref{fig:Frontiers}.
+We can therefore revisit the concept of a pseudo-cycle, only hinted at before, defining it as the samples of the original wave that belong to two successive pulses that in turn belong to their respective frontiers, including eventual pulses between them, as can be seen in figure \ref{fig:Frontiers}.
 
 It is now necessary to infer the appropriate radius of such a circle and, to that end, a measure of the curvature of a discrete function is needed.
-Discrete curvature estimation is an important task in image processing \parencite{2010FleischmannNovel} for which no default definition exists; The two possible approaches are the derivation of direct methods that use characteristics of the discrete wave to calculate its curvature at each point, or the calculation of the curvature of a curve fitted to the discrete wave \parencite{2001CoeurjollyDiscrete}.
+Discrete curvature estimation is an important task in image processing \parencite{2010FleischmannNovel} for which no default definition exists.
 
-To fit a polynomial to the a generic wave two approaches are readily available: the least mean squares approach, that seeks to minimize the dependant variable errors and the total least squares problem, that treats both variables symmetrically \parencite{1980GolubAnalysis}.
+The two possible general approaches are the derivation of direct methods that use characteristics of the discrete wave to calculate its curvature at each point, or the calculation of the curvature of a curve fitted to the discrete wave \parencite{2001CoeurjollyDiscrete}.
 
-The geometric, symmetrical nature of the problem excludes the more computationally economic least mean squares approach however, leaving us with the total least squares, for which no general closed form solutions are available \parencite{2007MarkovskyOverview}; this leads us to turn our attention to direct approaches of curvature estimation.
+To fit a polynomial to a generic wave two approaches are readily available: the least mean squares approach, that seeks to minimize the dependent variable errors and the total least squares problem, that treats both variables symmetrically \parencite{1980GolubAnalysis}.
+
+The geometric, symmetrical nature of the problem excludes the more computationally economic least mean squares approach, however, leaving us with the total least squares, for which no general closed-form solutions are available \parencite{2007MarkovskyOverview}; this leads us to turn our attention to direct approaches of curvature estimation.
 
@@
 
-As previously stated, no default discrete curvature definition exists; \textcite{2014CarrollSurvey}, for example, derives 3 such different definitions based on the approximation of a circle by an inscribed, centred e circumscribed polygon.
-In the context of 3d meshes, \textcite{2016VasaMultivariate} evaluate a range of estimators from a multivariate point o view.
+As previously noted, no default discrete curvature definition exists; \textcite{2014CarrollSurvey}, for example, derives three such different definitions based on the approximation of a circle by an inscribed, centred and circumscribed polygon.
+In the context of 3d meshes, \textcite{2016VasaMultivariate} evaluate a range of existing estimators from a multivariate point of view.
 
@@
 To that end we are going to apply the definition of smooth curvature as the rate of change of the unit tangent to a curve, noting that this is equivalent to that of the osculating circle \parencite{2016VasaMesh}.
-The idea is to find the equivalent circle whose tangent presents the same change in direction, in the same horizontal distance, as the edge of interest.
-This approach presents the convenience of estimating the radius directly.
+
+The idea is to find the equivalent circle whose tangent presents the same change in direction, in the same horizontal distance, as the edge of interest, as shown in \ref{fig:DiscreteCurvature}.
+This approach has the convenience of estimating the radius directly.
 
@@
 
-Formally, for each frontier, positive or negative, we define $ m-1 $ vectors, $ m $ being the number of positive or negative pulses respectively, as the coordinates of the posterior pulse with the relevant signal minus the preceding pulse: $\vec{v}_l^\pm = (P[l+1]_x^\pm - P[l]_x^\pm, P[l+1]_y^\pm - P[l]_y^\pm), \forall 0 \ \le l < m - 1 $, recalling that the $x$ coordinate of a pulse was defined as the $x$ coordinate of its sample with maximum amplitude, while the $y$ coordinate of a pulse is this amplitude itself.
+For each frontier, the angle $ \theta $ between the average direction and each of the vectors can be obtained via the relation between the slopes of the respective lines, as in the expression $ \theta = \arctan( (m - \overline{m}) / (1 + m \overline{m}) ) $, with $\overline{m} = \sum_{l=1}^{m-1} (P[l+1]_y - P[l]_y) / (P[l+1]_x - P[l]_x) $ and $ m_l = (P[l+1]_y - P[l]_y)/ (P[l+1]_x - P[l]_x) $
 
-Since an edge defines only one direction, there remains the need for a direction of reference.
-As we are interested in the relative curvature, it is natural to use the average direction of all positive or negative vectors, $ \vec{\overline{v}}^\pm = \frac{1}{m-1} \sum_{l=0}^{m-1} \vec{v}_l^\pm $, so as to measure the relative change of direction of each edge.
-This procedure, as well as the circles generated by it, are illustrated in figure \ref{fig:Curvatures}.
-
-The angle $ \theta $ between the average direction and each of the vectors can be obtained via the relation between the slopes of the respective lines, as in the expression $ \theta = \arctan( (m - \overline{m}) / (1 + m \overline{m}) ) $.
-The radii of the equivalent circles are thus given by $ r_l = x_l / \sin (\theta_l) $ with the curvature given by $\kappa_l =1/r_l=\sin (\theta_l) / x_l$.
-After some algebra, one can write:
+The radii of the equivalent circles are thus given by $ r_l = x_l / \sin (\theta_l) $ with the curvature given by $\kappa_l = 1/r_l=\sin (\theta_l) / x_l$, where $ x_l = P[l+1]_x - P[l]_x $ and $ y_l = P[l+1]_y - P[l]_y $.
+Joining both equations, one can write:
 
 \begin{equation} \label{eq:radius}
-  r_l = \left\lvert \frac{x_l \sqrt{\overline{m}^2 + 1} \ \sqrt{x_l^2 + y^2}}{\overline{m} x_l - y} \right\rvert
+  r_l = \left\lvert \frac{x_l \sqrt{\overline{m}^2 + 1} \ \sqrt{x_l^2 + y_l^2}}{\overline{m} x_l - y_l} \right\rvert
 \end{equation}
+
+After this procedure, illustrated in figure \ref{fig:Curvatures}, two sets of points in $ \mathbb{R}^2 $, corresponding to the positive and negative frontiers, are obtained, where the ordinates represent the instantaneous radius, with its localization in the time domain given by the abscissas.
 
@@
 
-After this procedure one obtains a set of points in $ \mathbb{R}^2 $ where the ordinate represents the instantaneous radius, with its localization in the space of the wave given by the ordinates.
-Given the robustness of the method, for an average wave of a couple of seconds at an ordinary rate of 44100 $ \mathit{fps} $ the average of the radius values is sufficient to assure the identification of the frontier.
-However, a curve fitting method can be used in the case of longer waves.
-An example of the use of the average equivalent circle can be seen in figure \ref{fig:AvgCircle}.
+Given the robustness of the method, for an average wave of a few seconds at an ordinary rate of 44100 $ \mathit{fps} $ the average of the radius values is sufficient to assure the accurate identification of the frontier.
+
+However, a curve fitting method can be used in the case of longer or ill-behaved waves. An example of the use of the average equivalent circle can be seen in figure \ref{fig:AvgCircle}.
 
@@
 
-Figure \ref{fig:FullFrontiers} illustrates the frontiers obtained to six diverse discrete sound waves, as well as a detail view of a shorter segment for each wave.
-All waves are records of physical sounds, chosen to represent the applicability of the algorithm in real world scenarios.
-
-We can see that the frontier is satisfactorily detected in harmonic an inharmonic sounds, and presents robustness in relation to the number of samples and the frequencies of the waves.
-
-% Key results in details
+Figure \ref{fig:FullFrontiers} illustrates the frontiers of six diverse discrete sound waves, as well as an in-detail view of a shorter segment for each signal.
+All waves used are records of physical sounds, chosen to represent the applicability of the algorithm in real-world scenarios.
 
@@
     \includegraphics[width=0.8\linewidth]{07FullFrontiers.pdf}
-  \caption{Positive and negative frontiers of 6 digital waves, as extracted by the algorithm here presented.
+  \caption{Positive and negative frontiers of six digital waves, as extracted by the algorithm here presented.
 For each wave, the region highlighted in black is shown in detail besides the whole wave.
@@
 
-The impact of the normalization of the wave with the frontiers obtained can be seen in figure \ref{fig:Fourier}, which shows the original and normalized wave in the case of a recording of the key 33 of a piano both in the time and frequency domains.
+We can see that the frontier is satisfactorily detected in the case of harmonic and inharmonic sounds, and presents robustness concerning the number of samples and the frequencies of the waves.
+
+Besides, the algorithm provides the position of the pseudo-cycles of a wave, with which one can, for example, derive its fundamental frequency and phase: using one of the two frontiers, we have $ f = (\#F^\pm - 1) n / (F^\pm[-1] - F^\pm[0])$, where $ \#F^\pm $ denotes the cardinality of the frontier, while $ F^\pm[0] $ and $ F^\pm[-1] $ stand for the frontier's first and last item, respectively.
+
+Similarly, each item of each of the frontiers can be used to estimate the local phase at that location.
+For the positive frontier, it is necessary to note that, for each item, the following equation holds: $ \cos(\phi + 2 \pi f i / n) = 1 $, and thus $ \phi = k \ 2 \pi - 2 \pi f i / n $, with $ k \in \mathbb{N } $, while for the negative frontier $ \cos(\phi + 2 \pi f i / n) = -1 $ and $ \phi = k \ 3 \pi - 2 \pi f i / n) $.
+
+The impact of the normalization of a wave using the frontiers obtained can be seen in figure \ref{fig:Fourier}, which shows the original and normalized wave in the case of a recording of the key 33 of a piano both in the time and frequency domains.
+We can see that the Fourier power spectra is greatly simplified in the case of the normalized wave.
 
@@
 
-Although important from a theoretical standpoint, this step has a special computational relevance in the case of digital sound waves, as they tend to be composed of a high number of samples in moderate amplitudes, generally normalized between -1 and 1, leading to a numerical disparity between axes.
+Although important from a theoretical standpoint, this step has an especial computational relevance in the case of digital sound waves, as they tend to be composed of a high number of samples with relatively low amplitudes, generally normalized between -1 and 1, leading to a numerical disparity between axes.
 
@@
     \includegraphics[width=0.8\linewidth]{10xRobustness.pdf}
-  \caption{Minimum and maximum scales of the abscissa before computational stability issues arose, for a recording of a drumkit tom.  $ \text{B} = 10^6 $.}
+  \caption{Minimum and maximum scales of the abscissa before computational stability issues arose, for a recording of a drum kit tom.  $ \text{B} = 10^6 $.}
   \label{fig:xRobustness}
@@
     \includegraphics[width=0.8\linewidth]{09yRobustness.pdf}
-  \caption{Minimum and maximum values of the amplitude before computational stability issues arose, for a recording of a drumkit tom.}
+  \caption{Minimum and maximum values of the amplitude before computational stability issues arose, for a recording of a drum kit tom.}
   \label{fig:yRobustness}
@@
 
-% Comparison with results in other works
-Direct comparison with other algorithms is made difficult by the unavailability of digital implementations of such works, many designed to process analog signals \parencite[e.g.,][]{2018AssefModeling}.
+\subsection{Comparison with traditional algorithms}
+
+Direct comparison with many of the more recent algorithms is made difficult by the unavailability of digital implementations of such works, many designed to process analogue signals \parencite[e.g.,][]{2018AssefModeling}.
+
+Nevertheless, insight can be gained from a comparison of the results of the method here proposed with some of the most common envelope extraction algorithms, as can be seen in figure \ref{fig:Comparison}.
+
+\begin{figure}[ht!]
+  \centering
+    \includegraphics[width=0.8\linewidth]{12Comparison.pdf}
+  \caption{Comparison of the algorithm present in this work with the most common methods of digital envelope identification.}
+  \label{fig:Comparison}
+\end{figure}
+
+For a recording of a guitar bend, we contrast the proposed algorithm with four common approaches of varying complexity: simple smoothing of the rectified original wave, low pass filtering followed by rectification, peak identification and subsequent smoothing, and the extraction of the envelope via Hilbert transform, as presented earlier in the paper, with a previous smoothing of the original wave to avoid the artefacts shown in figure \ref{fig:AnalyticSignal}.
+
+All those methods demanded careful choice of parameters: in the case of filtering, a cut-off frequency of approximately 88 Hz was used, consisting of one additional parameter to be tuned.
+For all the other traditional approaches, the Savitzky-Golay smoothing algorithm was used, and the window size was the parameter to be defined.
+
+In the case of the proposed algorithm, the smooth temporal envelope was obtained merging both positive and negative frontiers, that were also smoothed with the same procedure described.
+Figure \ref{fig:FrontiersToEnvelope} illustrates this process.
+
+\begin{figure}[ht!]
+  \centering
+    \includegraphics[width=0.8\linewidth]{13FrontiersToEnvelope.pdf}
+  \caption{Original positive and negative frontiers in dark grey and their mirrored versions. The temporal envelope, in black, is obtained by merging and approximating both frontiers.}
+  \label{fig:FrontiersToEnvelope}
+\end{figure}
+
+Table \ref{table:Comparison} presents a computational comparison.
+The times are for custom implementations of those methods in the Python programming language.
+The least mean squares result was calculated by halving the envelopes presented in figure \ref{fig:Comparison} and comparing them with the rectified absolute wave, so as to avoid the potential influence of an estimated frequency.
+
+We can see that, while the algorithm here presented is the most accurate, it's also the most computationally intensive.
+The connection between alpha-shapes and Delaunay triangulation can be explored in future works as a means to improve the efficiency of the algorithm.
+
+\begin{table}[ht!]
+\centering
+\begin{tabular}{ l c c }
+\hline
+Method & LMS & time(s) \\
+\hline
+Presented Algorithm & 0.0088 & 4.1507 \\ 
+Smoothing           & 0.0123 & 0.1911 \\ 
+Low pass Filter      & 0.0258 & 0.5601 \\ 
+Hilbert Transform   & 0.0141 & 0.2599 \\ 
+\hline
+\end{tabular}
+\caption{Comparison of the algorithm present in this work with the most common methods of digital envelope identification.}
+\label{table:Comparison}
+\end{table}
 
@@
 
-This work hopes to fill a gap identified by the authors in the context of sound synthesis, where the lack of procedures for the accurate identification of the temporal envelope of arbitrary waves consists of an obstacle to the complete description, and eventual manipulation of signals.
+This work hopes to fill a gap identified by the authors in the context of sound synthesis, where the lack of procedures for the accurate identification of the temporal envelope of arbitrary waves consists of an obstacle to the complete description and eventual manipulation of signals.
 
-While relevant in its own accord, the procedure here presented isolates the individual cycles of a wave, pinpointing them in the time domain.
-Due to that, a rich set of theoretical advancements, of whose an example was given in the preceding section, can be built upon this initial building block.
+While relevant in its own accord, the procedure here presented isolates the individual pseudo-cycles of a wave, pinpointing them in the time domain.
+Due to that, a rich set of theoretical advancements, of whose an example was given in the preceding section, can be built upon this initial algorithm.
 
-With the knowledge of the exact position and characteristics of the cycles of a wave, one can investigate its evolution in time, in terms of frequency content for example, free from the potencial interference of a modulating wave; in other words, one can have a very satisfactory view of the instantaneous frequencies and phase of an arbitrary wave and its evolution in time, providing a way to bridge the gap between time domain and frequency domain interpretations.
+With the knowledge of the exact position and characteristics of the pseudo-cycles of a wave, one can investigate its evolution in time, in terms of waveform, for example, free from the potential interference of a modulating wave; in other words, one can have a very satisfactory view of the instantaneous shape and phase of an arbitrary wave and its evolution in time, providing a way to bridge the gap between time domain and frequency domain interpretations.
+
+In the context of sound synthesis, for example, one would be able to apply specific methods for the recreation of the envelope, more in line with its smooth, slow varying and non-periodic nature and a different approach to the periodic and relatively fast changes of the temporal fine structure.
+
+\subsection{Conclusion}
+
+It is our opinion that such theoretical advances can be used in the formulation of better representations of discrete sounds, that in turn can improve sound compression techniques, while providing more adequate representation for the application of machine learning techniques for both sound classification and synthesis.
 

### 2020-09-06 Papers/02_Pulses_Envelope/PulsesEnvelope.tex
checked phase calculation in the paper

@@
 Nevertheless, the literature in this area is very fragmented \parencite{2017LyonsDigital}.
-Besides, most envelope detection techniques are designed to account for very specific settings, like pure sinusoids with moderate noise content, a limitation that excludes most physical signals, as is the case of recorded sound, for examples; that limitation arises in part by the lack of a strict mathematical definition for a temporal envelope \parencite{2013Mengempirical}.
+Besides, most envelope detection techniques are designed to account for very specific settings, like pure sinusoids with moderate noise content, a limitation that excludes most physical signals, as is the case of recorded sound, for example; that limitation arises in part due to the lack of a strict mathematical definition of a temporal envelope \parencite{2013Mengempirical}.
 
 In many natural signals, however, the temporal amplitude envelope of a signal plays a prominent role in the characteristics exhibited: according to \textcite{2017QiRelative}, for example, the envelope is at least as important as the fine structure of a soundwave in the context of the intelligibility of mandarin tones.
-That is also the case for English \parencite{1995ShannonSpeech}, where even envelopes modulating mostly noise were still capable of conveying meaning.
+This is also the case for English \parencite{1995ShannonSpeech}, where even envelopes modulating mostly noise were still capable of conveying meaning.
 
@@
 
-Addressing the different units in the horizontal and vertical axes, one can transform the DSP problem of envelope detection in the geometric problem of defining the shape of a set of points in $ \mathbb{R}^2 $.
-For that purpose \textcite{1983Edelsbrunnershape} introduced the concept of alpha-shapes, a mathematically well-defined extension to the convex hull of a finite set of points, closely related to the Delaunay triangulation and Voronoi diagrams of those points.
+Addressing the different units in the horizontal and vertical axes, one can transform the DSP problem of envelope detection into the geometric problem of defining the shape of a set of points in $ \mathbb{R}^2 $.
+For that purpose, \textcite{1983Edelsbrunnershape} introduced the concept of alpha-shapes, a mathematically well-defined extension to the convex hull of a finite set of points, closely related to the Delaunay triangulation and Voronoi diagrams of those points.
 
@@
 
-Steps in the direction of translating geometric algorithms to the context of envelope detection were also made by \textcite{2015YangSkeleton} with an approach based on the construction of a skeleton underlying the wave, and with the direct translation of computer vision methods to the task \parencite{2015YangRepresenting}.
+Steps in the direction of translating geometric algorithms to the context of envelope detection were also made by \textcite{2015YangSkeleton} via an algorithm based on the construction of a skeleton underlying the wave, and also via the direct translation of computer vision methods to the task \parencite{2015YangRepresenting}.
 
-Following this path, we present a general approach to envelope detection, exploiting the intrinsic characteristics of a generic, spectrally complex wave in order to avoid the need for manual intervention or parameter tuning.
+Following this path, we present a general approach to envelope detection, exploiting the intrinsic characteristics of a generic, spectrally complex wave, in order to avoid the need for manual intervention or parameter tuning.
 
-While the robustness of the proposed approach allows it to be used as a plug-in replacement for many methods encountered in the literature, we feel that it would be particularly useful for sound synthesis; the envelope is shown to add complexity to the spectral representation of a wave \parencite{2019TarjanoNeuro}, and an accurate description of the envelope, while solving the problem of the evolution of the instantaneous amplitude of a signal in time, would also greatly simplify further spectral analysis.
-Moreover, the algorithm naturally divides a signal in its pseudo-cycles, pinpointing them in the time domain, providing the building blocks for the reconstruction of the fine structure of the wave.
+While the robustness of the proposed approach allows it to be used as a plug-in replacement for many methods encountered in the literature, we feel that it would be particularly useful for sound synthesis.
 
-The rest of this paper is developed as follows: After explaining and defining the various entities used subsequently, of which the concept of frontiers is arguably the most important, in the methodology section, we dedicate a section to present a new discrete curvature estimation approach.
+The envelope is shown to add complexity to the spectral representation of a wave \parencite{2019TarjanoNeuro}, and an accurate description of the envelope, while describing the evolution of the instantaneous amplitude of a signal in time, would also greatly simplify further spectral analysis.
+Moreover, the algorithm developed in this work naturally divides a signal in its pseudo-cycles, pinpointing them in the time domain, providing the building blocks for the reconstruction of the fine structure of the wave.
 
-We then proceed to illustrate the results of the application of the algorithm in a set of six diverse sound waves, ranging from voice to unpitched instruments comparing, afterwards, its performance with that of some classic digital envelope estimation algorithms, both from a visual and numerical standpoint.
+The rest of this paper is divided as follows: After explaining and defining the various entities to be used subsequently, of which the concept of frontiers is arguably the most important, in the methodology section, we dedicate a section to present a new discrete curvature estimation approach, and how it was used to identify the frontiers of a discrete wave.
 
-We then discuss the implications of the work, emphasizing the focus on the construction of new interpretations of the relationship between time domain and frequency domain sound descriptions, and our vision of the impact of this methodology in the field of sound synthesis.
+We then proceed to illustrate the results of the application of the algorithm in a set of six diverse sound waves, ranging from voice to unpitched instruments comparing, afterwards, its performance with that of classic digital envelope estimation algorithms, both from a visual and numerical standpoint.
+
+We then discuss the implications of the work, emphasizing the focus on the construction of new interpretations of the relationship between time domain and frequency domain sound descriptions, and our vision of the impact of this methodology in the field of sound synthesis, providing directions for future developments.
 
@@
 
-The best representation one can hope to get of the original wave is, then, the recording.
-It is sensible, thus, to characterise the envelope of that wave in terms of a subset of the exact samples that comprise the digital representation.
+The best representation one can hope to get of the original wave is, then, the recording itself.
+It is sensible, thus, to characterise the envelope of the wave in terms of a subset of the exact samples that comprise its digital representation.
 
-To that end, this work will emphasize the concept of frontier, as the points of the digital wave that mark the wave's upper and lower boundaries; from those points an envelope, conforming to the vague definitions of uniqueness and smoothness, can be trivially constructed.
+To that end, this work will emphasize the concept of frontier, as the points of the digital wave that mark the wave's upper and lower boundaries; from those points an envelope, conforming to the vague definitions of uniqueness and smoothness, can be trivially constructed, as will be shown.
 
-We start by defining a pulse as a series of consecutive samples in a discrete wave with the same sign.
-More formally, let $ W[i] \in \mathbb{R} \ \forall \ i \in \mathbb{N}_0, \ i < n \in \mathbb{N}_0 $ be a real, discrete and finite signal indexed, without loss of generality, over an adjacent subset of the natural numbers starting at zero.
+We start by defining a pulse as a series of consecutive samples of a discrete wave that have the same sign.
+More formally, let $ W[i] \in \mathbb{R} \ \forall \ i \in \mathbb{N}_0; \ i < n \in \mathbb{N}_0 $ be a real, discrete and finite signal indexed, without loss of generality, over an adjacent subset of the natural numbers starting at zero.
 This definition relates closely to the concept of an array or vector in most programming languages, and is used in the interest of simplicity.
 
-A pulse $ P[j], j \in \mathbb{N}_0, j < q $ in $ W $ can then be defined as a sequence of samples indexed by $ i $ such that $ \sign(W[a-1]) \ne \sign(W[a]) = \sign(W[a+1]) = \dots = \sign(W[b-1]) = \sign(W[b]) \ne \sign(W[b+1]) \forall i, a \le i \le b | a,b \in \mathbb{N}_0 $.
+A pulse $ P[j]; \ \forall \ j \in \mathbb{N}_0; \ j < q \in \mathbb{N}_0; q \le n $ is, then, a subset of $ W $ with cardinality $ \#P[j] = b - a $ such that $ \sign(W[a-1]) \ne \sign(W[a]) = \sign(W[a+1]) = \dots = \sign(W[b-1]) = \sign(W[b]) \ne \sign(W[b+1]) \ \forall \ i; \ a \le i \le b; \ a, b \in \mathbb{N}_0 $.
 
-In other words, each pulse is a unique subset of $ W $ with its members being the maximum sequence of samples in $ W $ with the same signal and $ P[0] \cup P[1] \cup \cdots \cup P[q-1] = W $.
-It is convenient to attribute to each pulse a pair of coordinates: those will be the coordinates of the sample with the maximum absolute amplitude in each pulse, being denoted by $ P[j]_x $ and $ P[j]_y $; additionally, the length, or cardinality, of a pulse will be denoted by $ \#P[j] $.
+In other words, each pulse is a unique subset of $ W $ with its members being the maximum sequence of samples of $ W $ having the same signal.
+Thus $ P[0] \cup P[1] \cup \cdots \cup P[q-1] = W $ and $ P[i_1] \cap P[i_2] = \emptyset \ \forall \ i_1, i_2 \in \mathbb{N}_0;\ i_1 \ne i_2; \ i_1, i_2 < q $.
 
-Naturally, each pulse can be classified as positive ($ P[j]^+ $) or negative($ P[j]^- $), based on the sign of its samples, which must be the same.
-For an arbitrary continuous function, then, a pulse would correspond to the pieces between its roots, and would be equal to half the cycle of a sinusoid.
-Thus, in our definition, two consecutive pulses, naturally with opposing signs, have the potential to delimitate a pseudo-cycle, as illustrated in figure \ref{fig:Pulses}.
+Naturally, each pulse can be classified as positive ($ P[j]^+ $) or negative ($ P[j]^- $), based on the sign of its samples, which must be the same.
+It is also convenient to attribute to each pulse a pair of coordinates: those will be the coordinates of the sample with the maximum absolute amplitude in each pulse, and will be denoted by $ P[j]^\pm_x $ and $ P[j]^\pm_y $.
+Additionally, the length, or cardinality, of a pulse will be denoted by $ \#P[j]^\pm $.
+
+For an arbitrary continuous function, then, a pulse would correspond to the regions between its roots; for a sinusoid, a pulse would be equal to half of its cycle.
+In analogy, two consecutive pulses, naturally with opposing signs, have the potential to delimitate a pseudo-cycle, as illustrated in figure \ref{fig:Pulses}.
 For that to be true, however, the pulses need to make part of the frontiers of the signal, as yet to be defined.
@@
       \includegraphics[width=0.8\linewidth]{01Pulses.pdf}
-    \caption{The grey areas encompass the simplified convex hulls of the pulses. The dashed lines mark the frontier between pseudo-cycles.}
+    \caption{Example of a discrete wave.
+The black stems represent the individual samples of the signal.
+The grey areas encompass the simplified convex hulls of the pulses.
+The dashed lines mark the frontiers between pseudo-cycles.}
     \label{fig:Pulses}
@@
 
-The most mathematically sound definition of an envelope involves its representation as an analytic signal, introduced by Gabor in 1946 \parencite{2007HahnHistory}.
-In his work, \textcite{1946GaborTheory} applies the then relatively new mathematical machinery of the quantum mechanics to unify time and frequency representations of a wave, showing how the Hilbert transform could be applied to a real signal in order to obtain an equivalent complex signal, later known as the analytic signal.
+The most mathematically sound definition of an envelope involves the representation of its underlying wave as an analytic signal, introduced by Gabor in 1946 \parencite{2007HahnHistory}.
+In his work, \textcite{1946GaborTheory} applies the then relatively new mathematical machinery of the quantum mechanics to unify the time and frequency representations of a wave, showing how the Hilbert transform could be applied to a real signal in order to obtain an equivalent complex signal, later known as the analytic signal.
 
-This analytic signal is of the form $ a(t) = s(t) + \mathbb{H}(s(t)) \ \text{i} $ \parencite{2016HePraat} where $ s(t) $ is the original real signal, that becomes the real part of the analytic signal.
-$ \mathbb{H}(s(t)) $, the Hilbert transform of the original signal becomes, then, the imaginary part of the analytic signal; the envelope of a signal thus represented can be straightforwardly obtained by the computation of its complex modulus.
+This analytic signal has the form $ a(t) = s(t) + \mathbb{H}(s(t)) \ \text{i} $ \parencite{2016HePraat} where $ s(t) $ is the original real signal.
+$ \mathbb{H}(s(t)) $, the Hilbert transform of the original signal, becomes the imaginary part of the analytic signal; the envelope of a signal thus represented can be straightforwardly obtained by the computation of its complex modulus.
 
-Figure \ref{fig:AnalyticSignal} shows the envelope obtained via the Hilbert transform, for a pure sinusoid with a local frequency of 20 cycles, modulated by a polynomial of the third degree, in the presence and absence of gaussian white noise with standard deviation of $ 1/10 $ of the wave's maximum amplitude.
+Figure \ref{fig:AnalyticSignal} shows the envelope obtained via the analytic signal, for a pure sinusoid with a local frequency of 20 cycles, modulated by a polynomial of the third degree, illustrating the difference in the shape of the envelope in the presence and absence of gaussian white noise with standard deviation of $ 1/10 $ of the wave's maximum amplitude.
 
@@
 Other than that, this definition yields a unique, positive envelope.
-In practice, however, is not uncommon for a discrete wave, especially in the case of sound, to present somewhat different positive and negative contours; as our method will forcibly have to be applied separately to both sides, superior and inferior, of the discrete wave, it is only natural to introduce such separation as soon as possible.
 
-To do so we will define the positive (upper) frontier and negative (lower) frontiers of a discrete wave as the set of points $ (i, |W[i]|) $ that are, necessarily but not sufficiently, the points of maximum absolute amplitude of their respective pulses.
+In practice, however, is not uncommon for a discrete wave, especially in the case of sound, to present somewhat different positive and negative contours; to account for this discrepancy, we will define a positive (upper) and a negative (lower) frontier of a discrete wave.
+They will be composed by the points $ (i, W[i]) $ that are, necessarily but not sufficiently, the points of respectively maximum and minimum amplitude of their pulses, i.e.
+$ (P[j]^\pm_x, P[j]^\pm_y) $.
 
-The set of all these points such that $ W[i] > 0 $ will be called the positive, piecewise linear frontier and will be denoted by $ F^+ $; similarly, $ F^- $ will denote the negative frontier of $ W $, defined by the subset of points $ (i, W[i]) $ where $ W[i] < 0 $, as is shown in figure \ref{fig:Frontiers}.
-
-To identify such points, we need to address the different units in which the amplitude and horizontal distance of our wave are expressed.
+The set of all these points such that $ W[i] > 0 $ will be called the positive, piecewise linear frontier and will be denoted by $ F^+ $.
+Similarly, $ F^- $ will denote the negative frontier of $ W $, defined by the subset of points $ (i, W[i]) $ where $ W[i] < 0 $, as is shown in figure \ref{fig:Frontiers}.
 
@@
   \caption{Positive and negative frontiers, envelope and pseudo-cycles for a fragment of piano sound.
-While we have only one (positive) envelope, here mirrored in the inferior space, we have two frontiers, the superior and inferior.}
+While we have only one (positive) envelope, here mirrored in the inferior space, we have two frontiers, the superior and inferior, that are obtained independently.}
   \label{fig:Frontiers}
@@
 
+To identify such points, we need first to address the different units in which the amplitude and horizontal distance of our wave are expressed.
 To make abscissa and ordinate units consistent for the computation of the frontiers, and proceed with a geometric treatment of the subject, we consider the set of positive and negative pulses separately.
 
-For each set, we multiply all pulses lengths by the average length of the pulses in the set, after having divided by the average of the pulses' amplitudes, putting both axes in the same unit $i$, related to time by the equation $i = t \ \mathit{fps}$.
+For each set, we multiply all pulses vertical coordinates by the average length of the pulses in the set, after having divided by the average of the pulses' amplitudes, putting both axes in the same unit $i$, related to time by the equation $i = t \ \mathit{fps}$.
 
-Formally, supposing we have $ m $ pulses with the same sign, we scale each pulse's vertical coordinate $ P[j]^\pm_y $ by $ s^\pm $, where:
+Formally, supposing we have $ m $ pulses of the same sign, we scale each pulse's vertical coordinate $ P[j]^\pm_y $ by $ s^\pm $, where:
 
@@
 
-That justifies the approximation of the convex hull of a pulse by the triangle defined by the points where the pulse crosses the horizontal axis and the point of maximum amplitude; this will allow a considerable dimension reduction as this triangular pulse can become the atomic unit for the rest of the method.
+This makes clear that only one point of an arbitrary pulse can be in the frontier, namely, its point of maximum amplitude; those definitions will allow us to derive the frontiers of a wave using an alpha-shapes inspired algorithm without the need to compute the Delaunay triangulation first.
 
-This definition also makes clear that only one point of an arbitrary pulse can be in the frontier, namely, its point of maximum amplitude; those definitions will allow us to derive the frontiers of a wave using an alpha-shapes inspired algorithm without the need to compute the Delaunay triangulation first.
+That also justifies the approximation of the convex hull of a pulse by the triangle defined by the points where the pulse crosses the horizontal axis and the point of maximum amplitude $ (P[j]^\pm_x, P[j]^\pm_y) $; this triangular pulse and, more specifically, its point of maximum, will become the atomic unit for the rest of the method.
 
-Translating the intuitive explanation of alpha-shapes in \textcite{1994EdelsbrunnerThree} to our context, we can say that the points in the frontier are those touched by a circle outside the signal that is not allowed to contain any point of the signal; The pulses that contain said points can be seen as frontier pulses.
+Translating the intuitive explanation of alpha-shapes in \textcite{1994EdelsbrunnerThree} to our context, we can say that the points in the frontier are those touched by a circle of a given radius $ r $ outside the signal that is not allowed to contain any point of the signal; the pulses that contain said points can be seen as frontier pulses.
 
@@
 
-We can therefore revisit the concept of a pseudo-cycle, only hinted at before, defining it as the samples of the original wave that belong to two successive pulses that in turn belong to their respective frontiers, including eventual pulses between them, as can be seen in figure \ref{fig:Frontiers}.
+We can now revisit the concept of a pseudo-cycle, only hinted at before, defining it as the samples of the original wave that belong to two successive pulses that, in turn, belong to their respective frontiers, including eventual pulses between them, as can be seen in figure \ref{fig:Frontiers}.
 
-It is now necessary to infer the appropriate radius of such a circle and, to that end, a measure of the curvature of a discrete function is needed.
+It is necessary to infer the appropriate radius $ r $ of such a circle and, to that end, a measure of the curvature of a discrete function is needed.
 Discrete curvature estimation is an important task in image processing \parencite{2010FleischmannNovel} for which no default definition exists.
 
-The two possible general approaches are the derivation of direct methods that use characteristics of the discrete wave to calculate its curvature at each point, or the calculation of the curvature of a curve fitted to the discrete wave \parencite{2001CoeurjollyDiscrete}.
+The two possible general approaches are the derivation of direct methods that use characteristics of the discrete wave to calculate the curvature, or the use of the smooth curvature of a curve fitted to the discrete wave \parencite{2001CoeurjollyDiscrete}.
 
-To fit a polynomial to a generic wave two approaches are readily available: the least mean squares approach, that seeks to minimize the dependent variable errors and the total least squares problem, that treats both variables symmetrically \parencite{1980GolubAnalysis}.
+Concerning the last method, two approaches to fit a polynomial to a generic wave are readily available: the least mean squares approach, that seeks to minimize the dependent variable errors and the total least squares problem, that treats both variables symmetrically \parencite{1980GolubAnalysis}.
 
@@
 
-As previously noted, no default discrete curvature definition exists; \textcite{2014CarrollSurvey}, for example, derives three such different definitions based on the approximation of a circle by an inscribed, centred and circumscribed polygon.
+As previously noted, no default discrete curvature definition exists; \textcite{2014CarrollSurvey}, for example, derive three such different definitions based on the approximation of a circle by an inscribed, centred and circumscribed polygon.
 In the context of 3d meshes, \textcite{2016VasaMultivariate} evaluate a range of existing estimators from a multivariate point of view.
 
-For our needs, however, it is more convenient to define our own discrete curvature estimator; the approach here presented differs from most others found in the literature in that the curvature is defined over the edges linking the points that form the discrete wave, instead of the vertices.
-
+None of those approaches, especially for being defined over the vertices of a discrete wave, fit seamlessly into our algorithm; we thus proceed to define a discrete curvature measure over the edges of a discrete wave.
 To that end we are going to apply the definition of smooth curvature as the rate of change of the unit tangent to a curve, noting that this is equivalent to that of the osculating circle \parencite{2016VasaMesh}.
 
-The idea is to find the equivalent circle whose tangent presents the same change in direction, in the same horizontal distance, as the edge of interest, as shown in \ref{fig:DiscreteCurvature}.
-This approach has the convenience of estimating the radius directly.
+The idea is to find the equivalent circle whose tangent presents the same change in direction, in the same horizontal distance, as the edge of interest, as shown in figure \ref{fig:DiscreteCurvature}.
 
@@
   \centering
-    \includegraphics[width=0.3\linewidth]{04DiscreteCurvature.pdf}
-  \caption{Change of direction from horizontal, in $ \hat{u}_0 $, until an angle of $ \theta $, in $ \hat{u}_1 $, that is equivalent to the angle between the average direction and the direction of the vector $ \vec{v}_l $ whose radius of curvature is being calculated, shown in the picture in grey.}
+    \includegraphics[width=0.2\linewidth]{04DiscreteCurvature.pdf}
+  \caption{The tangent unit vector of the circle changes from the horizontal direction in $ \hat{u}_0 $ to an inclination of $ \theta $ in $ \hat{u}_1 $.
+$ \theta $ is also the angle between the average direction $ \vec{\overline{v}}_l $ and the vector $ \vec{v}_l $ that represents the edge of interest, shown in grey.}
   \label{fig:DiscreteCurvature}
@@
 
-For each frontier, the angle $ \theta $ between the average direction and each of the vectors can be obtained via the relation between the slopes of the respective lines, as in the expression $ \theta = \arctan( (m - \overline{m}) / (1 + m \overline{m}) ) $, with $\overline{m} = \sum_{l=1}^{m-1} (P[l+1]_y - P[l]_y) / (P[l+1]_x - P[l]_x) $ and $ m_l = (P[l+1]_y - P[l]_y)/ (P[l+1]_x - P[l]_x) $
+For each frontier, the angle $ \theta $ between the average direction and each of the vectors can be obtained via the relation between the slopes $ d $ of the respective lines, as in the expression $ \theta = \arctan( (d - \overline{d}) / (1 + d \overline{d}) ) $, with $\overline{d} = \sum_{l=1}^{m-1} (P[l+1]_y - P[l]_y) / (P[l+1]_x - P[l]_x) $ and $ d_l = (P[l+1]_y - P[l]_y)/ (P[l+1]_x - P[l]_x) $
 
@@
 \begin{equation} \label{eq:radius}
-  r_l = \left\lvert \frac{x_l \sqrt{\overline{m}^2 + 1} \ \sqrt{x_l^2 + y_l^2}}{\overline{m} x_l - y_l} \right\rvert
+  r_l = \left\lvert \frac{x_l \sqrt{\overline{d}^2 + 1} \ \sqrt{x_l^2 + y_l^2}}{\overline{d} x_l - y_l} \right\rvert
 \end{equation}
@@
 Given the robustness of the method, for an average wave of a few seconds at an ordinary rate of 44100 $ \mathit{fps} $ the average of the radius values is sufficient to assure the accurate identification of the frontier.
-
-However, a curve fitting method can be used in the case of longer or ill-behaved waves. An example of the use of the average equivalent circle can be seen in figure \ref{fig:AvgCircle}.
+However, a curve fitting method can be used in the case of longer or ill-behaved waves.
+An example of the use of the average equivalent circle can be seen in figure \ref{fig:AvgCircle}.
 
@@
 
-Figure \ref{fig:FullFrontiers} illustrates the frontiers of six diverse discrete sound waves, as well as an in-detail view of a shorter segment for each signal.
-All waves used are records of physical sounds, chosen to represent the applicability of the algorithm in real-world scenarios.
+Figure \ref{fig:FullFrontiers} illustrates the frontiers of six diverse discrete sound waves, as well as an in-detail view of the highlighted segment for each signal.
+All waves are records of physical sounds, chosen to represent the applicability of the algorithm in real-world scenarios.
 
@@
 
-We can see that the frontier is satisfactorily detected in the case of harmonic and inharmonic sounds, and presents robustness concerning the number of samples and the frequencies of the waves.
-
-Besides, the algorithm provides the position of the pseudo-cycles of a wave, with which one can, for example, derive its fundamental frequency and phase: using one of the two frontiers, we have $ f = (\#F^\pm - 1) n / (F^\pm[-1] - F^\pm[0])$, where $ \#F^\pm $ denotes the cardinality of the frontier, while $ F^\pm[0] $ and $ F^\pm[-1] $ stand for the frontier's first and last item, respectively.
-
-Similarly, each item of each of the frontiers can be used to estimate the local phase at that location.
-For the positive frontier, it is necessary to note that, for each item, the following equation holds: $ \cos(\phi + 2 \pi f i / n) = 1 $, and thus $ \phi = k \ 2 \pi - 2 \pi f i / n $, with $ k \in \mathbb{N } $, while for the negative frontier $ \cos(\phi + 2 \pi f i / n) = -1 $ and $ \phi = k \ 3 \pi - 2 \pi f i / n) $.
+We can see that the frontier is satisfactorily detected in the case of harmonic and inharmonic sounds, and is robust in relation to the number of samples and the frequencies of the waves.
 
@@
 
-For a recording of a guitar bend, we contrast the proposed algorithm with four common approaches of varying complexity: simple smoothing of the rectified original wave, low pass filtering followed by rectification, peak identification and subsequent smoothing, and the extraction of the envelope via Hilbert transform, as presented earlier in the paper, with a previous smoothing of the original wave to avoid the artefacts shown in figure \ref{fig:AnalyticSignal}.
+For a recording of a guitar sound, in which a bend is performed, we contrast the proposed algorithm with four common approaches of varying complexity: simple smoothing of the rectified original wave, low pass filtering followed by rectification, peak identification and subsequent smoothing, and the extraction of the envelope via Hilbert transform, as presented earlier in the paper, with a previous smoothing of the original wave to avoid the artefacts shown in figure \ref{fig:AnalyticSignal}.
 
@@
 
-In the case of the proposed algorithm, the smooth temporal envelope was obtained merging both positive and negative frontiers, that were also smoothed with the same procedure described.
-Figure \ref{fig:FrontiersToEnvelope} illustrates this process.
+In the case of the proposed algorithm, the smooth temporal envelope was obtained merging both positive and negative frontiers, the result being smoothed with the same procedure described.
+Figure \ref{fig:FrontiersToEnvelope} illustrates the results.
 
@@
 
-Table \ref{table:Comparison} presents a computational comparison.
+Table \ref{table:Comparison} presents a computational comparison between the techniques.
 The times are for custom implementations of those methods in the Python programming language.
@@
 
+\subsection{Retrieving information about the underlying wave}
+
+The algorithm provides the position of the pseudo-cycles of a wave, with which one can, for example, derive its fundamental frequency and phase: using one of the two frontiers, we have $ f = (\#F^\pm - 1) n / (F^\pm[-1] - F^\pm[0])$, where $ \#F^\pm $ denotes the cardinality of the frontier, while $ F^\pm[0] $ and $ F^\pm[-1] $ stand for the frontier's first and last item, respectively.
+
+Similarly, each point of each of the frontiers can be used to estimate the local phase at that location.
+For the positive frontier, it is necessary to note that, for each point, the following equation holds: $ \cos(\phi + 2 \pi f i / n) = 1 $, and thus $ \phi = 2 \pi k - 2 \pi f i / n $, with $ k \in \mathbb{N} $, while for the negative frontier $ \cos(\phi + 2 \pi f i / n) = -1 $ and $ \phi = 2 \pi k + \pi - 2 \pi f i / n) $.
+
 \section{Discussion}
@@
 While relevant in its own accord, the procedure here presented isolates the individual pseudo-cycles of a wave, pinpointing them in the time domain.
-Due to that, a rich set of theoretical advancements, of whose an example was given in the preceding section, can be built upon this initial algorithm.
+Due to that, a rich set of theoretical advancements, of whose an example was given in the end of the preceding section, can be built upon this initial algorithm.
 

### 2020-09-06 Papers/02_Pulses_Envelope/PulsesEnvelope.tex
revised with outwrite

@@
 
-In many natural signals, however, the temporal amplitude envelope of a signal plays a prominent role in the characteristics exhibited: according to \textcite{2017QiRelative}, for example, the envelope is at least as important as the fine structure of a soundwave in the context of the intelligibility of mandarin tones.
+In many natural signals, however, the temporal amplitude envelope of a signal plays a prominent role in the characteristics exhibited: according to \textcite{2017QiRelative}, for example, the envelope is at least as important as the fine structure of a sound wave in the context of the intelligibility of mandarin tones.
 This is also the case for English \parencite{1995ShannonSpeech}, where even envelopes modulating mostly noise were still capable of conveying meaning.
@@
 The envelope is shown to add complexity to the spectral representation of a wave \parencite{2019TarjanoNeuro}, and an accurate description of the envelope, while describing the evolution of the instantaneous amplitude of a signal in time, would also greatly simplify further spectral analysis.
-Moreover, the algorithm developed in this work naturally divides a signal in its pseudo-cycles, pinpointing them in the time domain, providing the building blocks for the reconstruction of the fine structure of the wave.
+Moreover, the algorithm developed in this work naturally divides a signal into its pseudo-cycles, pinpointing them in the time domain, providing the building blocks for the reconstruction of the fine structure of the wave.
 
@@
 The best representation one can hope to get of the original wave is, then, the recording itself.
-It is sensible, thus, to characterise the envelope of the wave in terms of a subset of the exact samples that comprise its digital representation.
+It is sensible, thus, to characterize the envelope of the wave in terms of a subset of the exact samples that comprise its digital representation.
 
@@
 
-For a recording of a guitar sound, in which a bend is performed, we contrast the proposed algorithm with four common approaches of varying complexity: simple smoothing of the rectified original wave, low pass filtering followed by rectification, peak identification and subsequent smoothing, and the extraction of the envelope via Hilbert transform, as presented earlier in the paper, with a previous smoothing of the original wave to avoid the artefacts shown in figure \ref{fig:AnalyticSignal}.
+For a recording of a guitar sound, in which a bend is performed, we contrast the proposed algorithm with four common approaches of varying complexity: simple smoothing of the rectified original wave, low pass filtering followed by rectification, peak identification and subsequent smoothing, and the extraction of the envelope via Hilbert transform, as presented earlier in the paper, with a previous smoothing of the original wave to avoid the artifacts shown in figure \ref{fig:AnalyticSignal}.
 

### 2020-09-07 Papers/02_Pulses_Envelope/PulsesEnvelope.tex
version submitted to Elsevier's Signal Processing

@@
 \usepackage[T1]{fontenc}
+\usepackage{authblk}
 \usepackage{amsfonts} % \mathbb
@@
   allcolors=black,
-  pdfauthor={Carlos Tarjano},
+  pdfauthor={Carlos Henrique Tarjano Santos},
   pdftitle={Robust Digital Envelope Estimation Via Geometric Properties of an Arbitrary Real Signal}
@@
 \title{Robust Digital Envelope Estimation Via Geometric Properties of an Arbitrary Real Signal}
-\author{Carlos Tarjano \and Valdecy Pereira}
+\author[1]{Carlos Henrique Tarjano Santos}
+\author[2]{Valdecy Pereira}
+\affil[1]{(corresponding author, carlostarjano@id.uff.br) Department of Production Engineering, Universidade Federal Fluminense, Rua Passo da Pátria, 156, Campus Praia Vermelha, Bloco D - sala 309, São Domingos, Niterói, RJ, Brasil, CEP: 24.210-240}
+\affil[2]{Department of Production Engineering, Universidade Federal Fluminense, Rua Passo da Pátria, 156, Campus Praia Vermelha, Bloco D - sala 309, São Domingos, Niterói, RJ, Brasil, CEP: 24.210-240}
 
@@
 
-{\bf Keywords:} DSP, alpha-shapes, envelope detection, discrete curvature
+{\bf Keywords:} DSP, alpha-shapes, envelope detection, discrete curvature estimation
 

### 2020-10-07 Text/00_Include.tex
organized text division

\documentclass[12pt]{article}
\usepackage[a4paper, top=3cm, left=3cm, bottom=2cm, right=2cm]{geometry}
\usepackage[utf8]{inputenc}
% \usepackage{fontspec}
% \setmainfont{Arial}
\title{Sound Synthesis}
\author{Carlos Tarjano}

### 2020-10-07 Text/03_Introduction_en.tex
organized text division

\input{00_Include.tex}
\usepackage[portuguese]{babel}

\begin{document}

\section{Introduction}
% Tell the reader the problem you are tackling in this project.

Humans are a creative bunch. Since the beginning of computing, even when mainframes were the only available medium, restricted to organizations with huge budgets, people found ways to develop artistic applications, like proto video games, in their (and the machine's) spare time. % TODO find citation

Music, on the other hand, is another staple of humanity's creativity. % TODO first musical instruments
 Is only natural that, as soon as this technology became reasonably stable, musical applications started to emerge.

% State clearly how you aim to deal with this problem. 
% Limit the scope of your study. 
% Sketch out how the thesis is structured to achieve your aim.

% objective
Ambitious as it may seem, it is the goal of this monograph to motivate a change of approach in domains that, despite their modest intersections, share more attributes than is apparent at a superficial glance: Both machine learning and musical acoustics are areas that, despite their current relevance, are fresh out of their respective infancies, and a scrutiny of both promptly reveals the difficulties that arise from this state of affairs. A lack of their own terminology, with frequent borrows from other, more established areas, can be readily cited, as the pronounced prominence of few individual contributions.

Those shortcomings have, however, a bright side to them, in that they expose those areas to reinterpretation, rendering them somewhat open to the influence of outside ideas.

Musical acoustics borrows heavily from the digital signal processing terminology, the last having its roots in the analogic world. Many algorithms are conceived in terms of filters, circuits and other similar legacy constructs, a fact that burdens the conceptual framework of the area. Besides the introduction of a noveau envelope extraction method, we find that his formulation in terms of (differential) geometric concepts is, perhaps, as great a contribution as the algorithm per se.

Perhaps less daring, and owing more to the advanced age of the author than to his research prowess, is the presentation of this monograph in a format more akin to a book; In addition to turning this somewhat dense tome into a more palatable experience. Besides, he whom maketh a monograph wishes not to repeat the experience, save if a new challenge presents itself.
scope
structure
socioeconomical relevance

\end{document}

### 2020-10-07 Text/old/00Preambulo/Thesis.tex
organized text division

--- 2020-10-07 Text/03_Introduction_en.tex
@@
-\input{00_Include.tex}
-\usepackage[portuguese]{babel}
+\documentclass{article}
+\usepackage[margin=0.5in]{geometry}
+\usepackage[utf8]{inputenc}
+\title{Sound Synthesis}
+\author{Carlos Tarjano}
 
@@
 
+\maketitle
+
+\begin{abstract}
+
+\end{abstract}
+
+{\bf Keywords:}
+
 \section{Introduction}
 % Tell the reader the problem you are tackling in this project. 
-
-Humans are a creative bunch.
-Since the beginning of computing, even when mainframes were the only available medium, restricted to organizations with huge budgets, people found ways to develop artistic applications, like proto video games, in their (and the machine's) spare time.
-% TODO find citation
-
-Music, on the other hand, is another staple of humanity's creativity. % TODO first musical instruments
- Is only natural that, as soon as this technology became reasonably stable, musical applications started to emerge.
-
 % State clearly how you aim to deal with this problem. 
@@
 
+\section{Bibliographic Review}
+
+\subsection{Sound Representation}
+frequency domain, one hot encoding, spectrogram
+
+\subsection{Computer Graphics}
+Bezier, splines, some raster, image rotation
+
+\subsection{Sound synthesis}
+
+\section{Methodology}
+
+\section{Results}
+
+\section{Discussion + Conclusion}
+
+\section{References}
+
 \end{document}

### 2020-10-08 Text/03_Introduction_pt.tex
structured thesis text in separated .tex files

--- 2020-10-07 Text/03_Introduction_en.tex
@@
 % Tell the reader the problem you are tackling in this project.
-
-Humans are a creative bunch.
-Since the beginning of computing, even when mainframes were the only available medium, restricted to organizations with huge budgets, people found ways to develop artistic applications, like proto video games, in their (and the machine's) spare time.
-% TODO find citation
-
-Music, on the other hand, is another staple of humanity's creativity. % TODO first musical instruments
- Is only natural that, as soon as this technology became reasonably stable, musical applications started to emerge.
 

### 2020-10-08 Text/06_SignalEnvelope.tex
structured thesis text in separated .tex files

--- 2020-09-06 Papers/02_Pulses_Envelope/PulsesEnvelope.tex
@@
-\documentclass{article}
-\usepackage{lmodern} % Latin Modern Font
-\usepackage[T1]{fontenc}
-\usepackage{amsfonts} % \mathbb
-\usepackage[margin=0.5in]{geometry}
-\usepackage[utf8]{inputenc}
-\usepackage{graphicx}
-\usepackage[backend=biber, style=apa]{biblatex}
-\usepackage{amsmath}
-\usepackage{hyperref}
-\usepackage{hypcap}
-\DeclareMathOperator{\sign}{sign}
-\bibliography{bibli}
-\hypersetup{
-  colorlinks=true, 
-  allcolors=black,
-  pdfauthor={Carlos Tarjano},
-  pdftitle={Robust Digital Envelope Estimation Via Geometric Properties of an Arbitrary Real Signal}
-}
-\title{Robust Digital Envelope Estimation Via Geometric Properties of an Arbitrary Real Signal}
-\author{Carlos Tarjano \and Valdecy Pereira}
-
-\begin{document}
-
-\maketitle
-
-\begin{abstract}
-Despite being an elusive concept, the temporal amplitude envelope of a signal is essential for its complete characterization, being the primary information-carrying medium in spoken voice and telecommunications, for example.
-Intuitively, the temporal envelope can be understood as a smooth function that multiplies the signal, being responsible for its outer shape.
-Envelope detection techniques have applications in areas like health, sound classification and synthesis, seismology and speech recognition.
-Nevertheless, a general method to digital envelope detection of signals with rich spectral content doesn't exist, as most methods involve manual intervention, in the form of filter design, smoothing, as well as other specific design choices, based on a priori knowledge about the nature of the specific waves under investigation.
-To address this problem, we propose a framework that uses intrinsic characteristics of a signal to estimate its envelope, completely eliminating the necessity of parameter tuning.
-The approach here described draws inspiration from geometric concepts to isolate the frontiers and thus estimate the temporal envelope of an arbitrary signal; to that end, alpha-shapes, concave hulls, and discrete curvature are explored.
-We also define entities, such as a pulse and frontiers, in the context of an arbitrary digital signal, as a means to reduce dimensionality and the complexity of the proposed algorithm.
-Specifically, a new measure of discrete curvature is used to obtain the average radius of curvature of a discrete wave, serving as a threshold to identify the wave's frontier points.
-We find the algorithm accurate in the identification of the frontiers of a wide range of digital sound waves with very diverse characteristics, while localizing each pseudo-cycle of the wave in the time domain.
-The algorithm also compares favourably with classic envelope detection techniques based on filtering and the Hilbert Transform.
-Besides the most direct applications of this work to audio classification and synthesis, we foresee impact in compression techniques and machine learning approaches to audio.
-The discrete curvature definition presented could also be extended to three-dimensional settings, to improve shape detection algorithms based on alpha-shapes.
-\end{abstract}
-
-{\bf Keywords:} DSP, alpha-shapes, envelope detection, discrete curvature
-
-\section{Introduction}
-Envelope detection is ubiquitous in both analogue and digital signal processing \parencite{2011CaetanoImproved}.
-Nevertheless, the literature in this area is very fragmented \parencite{2017LyonsDigital}.
-Besides, most envelope detection techniques are designed to account for very specific settings, like pure sinusoids with moderate noise content, a limitation that excludes most physical signals, as is the case of recorded sound, for example; that limitation arises in part due to the lack of a strict mathematical definition of a temporal envelope \parencite{2013Mengempirical}.
-
-In many natural signals, however, the temporal amplitude envelope of a signal plays a prominent role in the characteristics exhibited: according to \textcite{2017QiRelative}, for example, the envelope is at least as important as the fine structure of a sound wave in the context of the intelligibility of mandarin tones.
-This is also the case for English \parencite{1995ShannonSpeech}, where even envelopes modulating mostly noise were still capable of conveying meaning.
-
-The envelope helps to impart emotion and identity to the human voice \parencite{2018ZhuContributions}, and envelope preserving characteristics in concert halls are associated with their pleasantness \parencite{2011LokkiEngaging}.
-
-When dealing with broadband signals, approaches tailored to specific applications are prevalent, such as the one presented by \textcite{2014YangFast} for the distributed monitoring of fibre optic or the one formulated by \textcite{2018AssefModeling} in the context of medical ultrasound imaging.
-
-Addressing the different units in the horizontal and vertical axes, one can transform the DSP problem of envelope detection into the geometric problem of defining the shape of a set of points in $ \mathbb{R}^2 $.
-For that purpose, \textcite{1983Edelsbrunnershape} introduced the concept of alpha-shapes, a mathematically well-defined extension to the convex hull of a finite set of points, closely related to the Delaunay triangulation and Voronoi diagrams of those points.
-
-This approach is used in areas such as the detection of features in images \parencite{2016VarytimidisAlpha}, reconstruction of surfaces from a cloud of points \parencite{2015WuAutomated} and Spectroscopy \parencite{2019XuModeling}, with the last work, which involves the estimation and removal of the Blaze function (a kind of envelope) of an echelle spectrograph, being particularly illustrative of the potential of cross-fertilization.
-
-Steps in the direction of translating geometric algorithms to the context of envelope detection were also made by \textcite{2015YangSkeleton} via an algorithm based on the construction of a skeleton underlying the wave, and also via the direct translation of computer vision methods to the task \parencite{2015YangRepresenting}.
-
-Following this path, we present a general approach to envelope detection, exploiting the intrinsic characteristics of a generic, spectrally complex wave, in order to avoid the need for manual intervention or parameter tuning.
-
-While the robustness of the proposed approach allows it to be used as a plug-in replacement for many methods encountered in the literature, we feel that it would be particularly useful for sound synthesis.
-
-The envelope is shown to add complexity to the spectral representation of a wave \parencite{2019TarjanoNeuro}, and an accurate description of the envelope, while describing the evolution of the instantaneous amplitude of a signal in time, would also greatly simplify further spectral analysis.
-Moreover, the algorithm developed in this work naturally divides a signal into its pseudo-cycles, pinpointing them in the time domain, providing the building blocks for the reconstruction of the fine structure of the wave.
-
-The rest of this paper is divided as follows: After explaining and defining the various entities to be used subsequently, of which the concept of frontiers is arguably the most important, in the methodology section, we dedicate a section to present a new discrete curvature estimation approach, and how it was used to identify the frontiers of a discrete wave.
-
-We then proceed to illustrate the results of the application of the algorithm in a set of six diverse sound waves, ranging from voice to unpitched instruments comparing, afterwards, its performance with that of classic digital envelope estimation algorithms, both from a visual and numerical standpoint.
-
-We then discuss the implications of the work, emphasizing the focus on the construction of new interpretations of the relationship between time domain and frequency domain sound descriptions, and our vision of the impact of this methodology in the field of sound synthesis, providing directions for future developments.
-
-\section{Methodology}
+\section{Using geometric properties of a signal to estimate the envelope }
 
@@
     \centering
-      \includegraphics[width=0.8\linewidth]{01Pulses.pdf}
+      \includegraphics[width=0.8\linewidth]{./imgs/envelope/01Pulses.pdf}
     \caption{Example of a discrete wave.
@@
   \centering
-    \includegraphics[width=0.8\linewidth]{02AnalyticSignal.pdf}
+    \includegraphics[width=0.8\linewidth]{./imgs/envelope/02AnalyticSignal.pdf}
   \caption{Envelope of a pure sinusoid modulated by a polynomial of degree 3, as obtained by the Hilbert Transform.
@@
   \centering
-    \includegraphics[width=0.8\linewidth]{03Frontiers.pdf}
+    \includegraphics[width=0.8\linewidth]{./imgs/envelope/03Frontiers.pdf}
   \caption{Positive and negative frontiers, envelope and pseudo-cycles for a fragment of piano sound.
@@
   \centering
-    \includegraphics[width=0.2\linewidth]{04DiscreteCurvature.pdf}
+    \includegraphics[width=0.2\linewidth]{./imgs/envelope/04DiscreteCurvature.pdf}
   \caption{The tangent unit vector of the circle changes from the horizontal direction in $ \hat{u}_0 $ to an inclination of $ \theta $ in $ \hat{u}_1 $.
@@
   \centering
-    \includegraphics[width=0.8\linewidth]{05Curvatures.pdf}
+    \includegraphics[width=0.8\linewidth]{./imgs/envelope/05Curvatures.pdf}
   \caption{The equivalent circles obtained via the procedure described for the positive frontier.
@@
   \centering
-    \includegraphics[width=0.8\linewidth]{06AvgCircle.pdf}
+    \includegraphics[width=0.8\linewidth]{./imgs/envelope/06AvgCircle.pdf}
   \caption{The positive frontier as obtained by the use of the circle representing the average curvature of the discrete wave illustrated.}
@@
 
-\section{Results}
+\subsection{Results}
 
@@
   \centering
-    \includegraphics[width=0.8\linewidth]{07FullFrontiers.pdf}
+    \includegraphics[width=0.8\linewidth]{./imgs/envelope/07FullFrontiers.pdf}
   \caption{Positive and negative frontiers of six digital waves, as extracted by the algorithm here presented.
@@
   \centering
-    \includegraphics[width=0.8\linewidth]{08Fourier.pdf}
+    \includegraphics[width=0.8\linewidth]{./imgs/envelope/08Fourier.pdf}
   \caption{Original and normalized views of a recording of key 33 (F3, 175 Hz) from a grand piano both in time and frequency domains.}
@@
   \centering
-    \includegraphics[width=0.8\linewidth]{10xRobustness.pdf}
+    \includegraphics[width=0.8\linewidth]{./imgs/envelope/10xRobustness.pdf}
   \caption{Minimum and maximum scales of the abscissa before computational stability issues arose, for a recording of a drum kit tom.  $ \text{B} = 10^6 $.}
@@
   \centering
-    \includegraphics[width=0.8\linewidth]{09yRobustness.pdf}
+    \includegraphics[width=0.8\linewidth]{./imgs/envelope/09yRobustness.pdf}
   \caption{Minimum and maximum values of the amplitude before computational stability issues arose, for a recording of a drum kit tom.}
@@
   \centering
-    \includegraphics[width=0.8\linewidth]{12Comparison.pdf}
+    \includegraphics[width=0.8\linewidth]{./imgs/envelope/12Comparison.pdf}
   \caption{Comparison of the algorithm present in this work with the most common methods of digital envelope identification.}
@@
   \centering
-    \includegraphics[width=0.8\linewidth]{13FrontiersToEnvelope.pdf}
+    \includegraphics[width=0.8\linewidth]{./imgs/envelope/13FrontiersToEnvelope.pdf}
   \caption{Original positive and negative frontiers in dark grey and their mirrored versions. The temporal envelope, in black, is obtained by merging and approximating both frontiers.}
@@
 
-\section{Discussion}
+\subsection{Discussion}
 
@@
 It is our opinion that such theoretical advances can be used in the formulation of better representations of discrete sounds, that in turn can improve sound compression techniques, while providing more adequate representation for the application of machine learning techniques for both sound classification and synthesis.
-
-\section{References}
-
-\printbibliography[heading=none]
-
-\end{document}

### 2020-10-08 Text/03_Introduction.tex
structured thesis text in separated .tex files
custom recipe to use with glossaries

--- 2020-10-07 Text/03_Introduction_en.tex
@@
-\input{00_Include.tex}
-\usepackage[portuguese]{babel}
-
-\begin{document}
+\newacronym{ENIAC}{ENIAC}{Electronic Numeric Integrator and Calculator}
 
@@
 Humans are a creative bunch.
-Since the beginning of computing, even when mainframes were the only available medium, restricted to organizations with huge budgets, people found ways to develop artistic applications, like proto video games, in their (and the machine's) spare time.
-% TODO find citation
+Since the beginning of computing, even when mainframes were the only available medium, restricted to organizations with considerable budgets, people found ways and motivation to overcome the technical limitations and develop artistic applications, like video games, in their (and the machine's) spare time.
+
+On a parallel note, also relevant to this thesis, those first machines, like the \gls{ENIAC} in 1946, brought with then the promise of the artificial intelligence as their ultimate goal \parencite{2010DonovanReplay}.
 
@@
 % Sketch out how the thesis is structured to achieve your aim. 
-
 % objective
-Ambitious as it may seem, it is the goal of this monograph to motivate a change of approach in domains that, despite their modest intersections, share more attributes than is apparent at a superficial glance: Both machine learning and musical acoustics are areas that, despite their current relevance, are fresh out of their respective infancies, and a scrutiny of both promptly reveals the difficulties that arise from this state of affairs.
+Ambitious as it may seem, it is the goal of this monograph to motivate a change of approach in domains that, despite their modest intersections, share more attributes than is apparent at a superficial glance: Both machine learning and digital musical acoustics are areas that, despite their current relevance, are fresh out of their respective infancies, and a scrutiny of both promptly reveals the difficulties that arise from this state of affairs.
 A lack of their own terminology, with frequent borrows from other, more established areas, can be readily cited, as the pronounced prominence of few individual contributions.
 
-Those shortcomings have, however, a bright side to them, in that they expose those areas to reinterpretation, rendering them somewhat open to the influence of outside ideas.
+Those shortcomings have, however, a bright side to them, in that they still expose those areas to reinterpretation, rendering them somewhat open to the influence of outside ideas.
 
-Musical acoustics borrows heavily from the digital signal processing terminology, the last having its roots in the analogic world.
-Many algorithms are conceived in terms of filters, circuits and other similar legacy constructs, a fact that burdens the conceptual framework of the area.
+Digital musical acoustics borrows heavily from the digital signal processing terminology, the last having its roots in the analogic world.
+Many algorithms are conceived in terms of filters, circuits and other similar legacy constructs, a fact that burdens the conceptual, theoretical framework of the area.
 Besides the introduction of a noveau envelope extraction method, we find that his formulation in terms of (differential) geometric concepts is, perhaps, as great a contribution as the algorithm per se.
@@
 Perhaps less daring, and owing more to the advanced age of the author than to his research prowess, is the presentation of this monograph in a format more akin to a book; In addition to turning this somewhat dense tome into a more palatable experience.
-Besides, he whom maketh a monograph wishes not to repeat the experience, save if a new challenge presents itself.
-scope
-structure
-socioeconomical relevance
-
-\end{document}

### 2020-10-09 Text/00_Include.tex
structured thesis text in separated .tex files
custom recipe to use with glossaries
English set-up ok

@@
 \documentclass[12pt]{article}
+
+% \usepackage[utf8]{inputenc}
+% \usepackage[T1]{fontenc}
+% \usepackage{authblk}
+
+\usepackage[backend=biber]{biblatex}
+%\usepackage{biblatex}
+
 \usepackage[a4paper, top=3cm, left=3cm, bottom=2cm, right=2cm]{geometry}
-\usepackage[utf8]{inputenc}
-% \usepackage{fontspec}
-% \setmainfont{Arial}
+\usepackage{indentfirst}
+\setlength{\parindent}{1.25cm}
+
+\usepackage[no-math]{fontspec}
+\defaultfontfeatures{Ligatures=TeX}
+\setmainfont{Tex Gyre Termes}
+
+\usepackage[nodisplayskipstretch]{setspace}
+\setstretch{1.427465}
+
+\usepackage[explicit]{titlesec}
+\titleformat{\section}[hang]{\normalfont\bfseries}{\thesection}{1em}{\MakeUppercase{#1}}
+\titleformat{\subsection}[hang]{\normalfont}{\thesubsection}{1em}{\MakeUppercase{#1}}
+\titleformat{\subsubsection}[hang]{\normalfont\bfseries}{\thesubsubsection}{1em}{#1}%{\MakeLowercase{#1}}
+
+\usepackage{titletoc}
+\titlecontents{section}[1em]{\vspace{12pt}}{\bfseries\normalsize\contentslabel[\thecontentslabel]{3em}\uppercase}{\hspace*{-2em}\uppercase}{\titlerule*[.75em]{.}\contentspage}
+\titlecontents{subsection}[1em]{\vspace{12pt}}{\normalfont\normalsize\contentslabel[\thecontentslabel]{3em}\uppercase}{\hspace*{-2em}\uppercase}{\titlerule*[.75em]{.}\contentspage}
+\titlecontents{subsubsection}[1em]{\vspace{12pt}}{\bfseries\normalsize\contentslabel[\thecontentslabel]{3em}}{\hspace*{-2em}}{\titlerule*[.75em]{.}\contentspage}
+
+\usepackage{amsmath}  % enables declaremathoperator
+\DeclareMathOperator{\sign}{sign}
+
+\usepackage{amsfonts} % \mathbb
+\usepackage{graphicx} % manipulate imported images
+
+\usepackage{fancyhdr} % \pagestyle{fancy}
+\fancyhf{}
+\fancyhead[R]{\thepage}
+\renewcommand{\headrulewidth}{0pt}
+
+\usepackage[toc,section=section]{glossaries}
+\makeglossaries
+
+\usepackage{hyperref}
+\usepackage{hypcap}
+\hypersetup{
+  colorlinks=true, 
+  allcolors=black,
+  pdfauthor={Carlos Henrique Tarjano Santos},
+  pdftitle={Robust Digital Envelope Estimation Via Geometric Properties of an Arbitrary Real Signal}
+}
+
+\bibliography{00_BIBLI}
+
 \title{Sound Synthesis}
 \author{Carlos Tarjano}
+
+% \newglossaryentry{VST}{name=VST, description={Virtual Studio Technology}}

### 2020-10-14 Text/04_LMS.tex
some cleanup; added lms section

\section{Theory}

\subsection{Least Mean Squares}
The linear least mean squares, also known as linear regression \parencite{2018StephenBoydIntroduction} is a data fitting technique. % TODO

We present here the theory for real numbers, noting that it can be easily extended for complex numbers \parencite{1995LawsonSolving}, should the necessity arise. That, however, will not be the case in this work.

Given two vectors $ X $ and $ Y $, such that:
$$ X = [x_0, x_1, \cdots ,x_{m-1}]^T, \ x_i \ \in \ \mathbb{R} \ \forall \ i, \ 0 \le i < m \in \mathbb{N} $$
$$ Y = [y_0, y_1, \cdots ,y_{m-1}]^T, \ y_i \ \in \ \mathbb{R} \ \forall \ i, \ 0 \le i < m \in \mathbb{N} $$
We ideally want to find the vector of coefficients $ A = [a_0, a_1, \cdots ,a_k]^T, \ a_j \ \in \ \mathbb{R} \ \forall \ j, \ 0 \le j \le k \in \mathbb{N} $ that satisfies the equality $ x^0_i a_0 + x^1_i a_1 + \cdots + x^{k-1}_i a_{k-1} + x^k_i a_k = y_i \quad \forall \quad i \in \{0, \cdots , m-1\} $.

Put in that way, what we have is a system of linear equations that can be rewritten in matrix form as seen in equation \ref{eq:lms}.

\begin{equation}
  \label{eq:lms}
\stackrel{Q_{m \times (k+1)}}
{
\begin{bmatrix}
1 & x_0 & \cdots & x_0^k \\
1 & x_1 & \cdots & x_1^k \\
\vdots & \vdots & \ddots & \vdots  \\
1 & x_{m-1} & \cdots & x_{m-1} \\
\end{bmatrix}
}
\stackrel{A_{(k+1) \times 1}}
{
\begin{bmatrix}
a_0 \\
a_1 \\
\vdots  \\
a_k \\
\end{bmatrix}
}
=
\stackrel{Y_{m \times 1}}
{
\begin{bmatrix}
y_0 \\
y_1 \\
\vdots  \\
y_{m-1} \\
\end{bmatrix}
}
\end{equation}

More compactly, we have that $ Q A = Y $. This system will generally be over-determined, only having an exact solution in the case where $ Y $ is a linear combination of the columns in $ Q $ \parencite{2018StephenBoydIntroduction}.

Since an exact solution is often beyond reach, we can at least choose $ A $ in a way that makes $ Q A $ and $ Y $ close. A way to do that is to minimize the absolute errors (or their square) between those two quantities; $ \underset{A}{\min} ||Q A - Y||^2_2 $

Using calculus, for example, on can arrive at the equation $ \hat{A} = (Q^T Q)^{-1} Q^T Y $, where $ \hat{A} $ is the vector of coefficients that minimizes the squared errors. A detailed derivation is available at \cite{2018StephenBoydIntroduction}.

\subsection{Constrained Least Mean Squares}

In the constrained version of the least squares problem we are still interested in finding the coefficient vector $ \hat{A} $ that minimizes $ ||Q A - Y||^2_2 $, but respecting a restriction on this vector of the general form of $ V A = B $.

We can then define the Lagrangian and set its derivatives to zero to obtain, as described in \textcite{2013SelesnickLeast}, the closed form solution

$$ \hat{A} = \left(Q^{T} Q\right)^{-1} \left(Q^{T} Y - V^{T} \left(V \left(Q^{T} Q\right)^{-1} V^{T}\right)^{-1} V \left(Q^{T} Q\right)^{-1} Q^{T} Y - B \right) $$

### 2020-10-14 Text/05_Characterization.tex
some cleanup; added lms section

--- 2020-09-06 Papers/02_Pulses_Envelope/PulsesEnvelope.tex
@@
-\documentclass{article}
-\usepackage{lmodern} % Latin Modern Font
-\usepackage[T1]{fontenc}
-\usepackage{amsfonts} % \mathbb
-\usepackage[margin=0.5in]{geometry}
-\usepackage[utf8]{inputenc}
-\usepackage{graphicx}
-\usepackage[backend=biber, style=apa]{biblatex}
-\usepackage{amsmath}
-\usepackage{hyperref}
-\usepackage{hypcap}
-\DeclareMathOperator{\sign}{sign}
-\bibliography{bibli}
-\hypersetup{
-  colorlinks=true, 
-  allcolors=black,
-  pdfauthor={Carlos Tarjano},
-  pdftitle={Robust Digital Envelope Estimation Via Geometric Properties of an Arbitrary Real Signal}
-}
-\title{Robust Digital Envelope Estimation Via Geometric Properties of an Arbitrary Real Signal}
-\author{Carlos Tarjano \and Valdecy Pereira}
+\section{Characterization of a digital Signal}
+% TODO continuous to digital
 
-\begin{document}
+\subsection{Envelope of a digital signal}
 
-\maketitle
-
-\begin{abstract}
 Despite being an elusive concept, the temporal amplitude envelope of a signal is essential for its complete characterization, being the primary information-carrying medium in spoken voice and telecommunications, for example.
@@
 The discrete curvature definition presented could also be extended to three-dimensional settings, to improve shape detection algorithms based on alpha-shapes.
-\end{abstract}
 
-{\bf Keywords:} DSP, alpha-shapes, envelope detection, discrete curvature
-
-\section{Introduction}
+\subsubsection{Envelope Detection}
 Envelope detection is ubiquitous in both analogue and digital signal processing \parencite{2011CaetanoImproved}.
@@
 
-The rest of this paper is divided as follows: After explaining and defining the various entities to be used subsequently, of which the concept of frontiers is arguably the most important, in the methodology section, we dedicate a section to present a new discrete curvature estimation approach, and how it was used to identify the frontiers of a discrete wave.
+% The rest of this paper is divided as follows: After explaining and defining the various entities to be used subsequently, of which the concept of frontiers is arguably the most important, in the methodology section, we dedicate a section to present a new discrete curvature estimation approach, and how it was used to identify the frontiers of a discrete wave.
 
@@
 
-\section{Methodology}
+\subsection{Frequency content}
 
-In this work it is assumed, without loss of generality, that one is interested in retrieving the mathematically ill-defined \parencite{2013Mengempirical} temporal envelope of an arbitrary physical sound wave, of which only a digital representation is available.
+\subsection{Representations}
 
-Naturally, due to the necessary truncations both in time and amplitude, as well as measurement errors introduced by the instruments used to record the signal, it is not possible to exactly recover the original, continuous wave, from its digital counterpart.
-
-The best representation one can hope to get of the original wave is, then, the recording itself.
-It is sensible, thus, to characterize the envelope of the wave in terms of a subset of the exact samples that comprise its digital representation.
-
-To that end, this work will emphasize the concept of frontier, as the points of the digital wave that mark the wave's upper and lower boundaries; from those points an envelope, conforming to the vague definitions of uniqueness and smoothness, can be trivially constructed, as will be shown.
-
-We start by defining a pulse as a series of consecutive samples of a discrete wave that have the same sign.
-More formally, let $ W[i] \in \mathbb{R} \ \forall \ i \in \mathbb{N}_0; \ i < n \in \mathbb{N}_0 $ be a real, discrete and finite signal indexed, without loss of generality, over an adjacent subset of the natural numbers starting at zero.
-This definition relates closely to the concept of an array or vector in most programming languages, and is used in the interest of simplicity.
-
-A pulse $ P[j]; \ \forall \ j \in \mathbb{N}_0; \ j < q \in \mathbb{N}_0; q \le n $ is, then, a subset of $ W $ with cardinality $ \#P[j] = b - a $ such that $ \sign(W[a-1]) \ne \sign(W[a]) = \sign(W[a+1]) = \dots = \sign(W[b-1]) = \sign(W[b]) \ne \sign(W[b+1]) \ \forall \ i; \ a \le i \le b; \ a, b \in \mathbb{N}_0 $.
-
-In other words, each pulse is a unique subset of $ W $ with its members being the maximum sequence of samples of $ W $ having the same signal.
-Thus $ P[0] \cup P[1] \cup \cdots \cup P[q-1] = W $ and $ P[i_1] \cap P[i_2] = \emptyset \ \forall \ i_1, i_2 \in \mathbb{N}_0;\ i_1 \ne i_2; \ i_1, i_2 < q $.
-
-Naturally, each pulse can be classified as positive ($ P[j]^+ $) or negative ($ P[j]^- $), based on the sign of its samples, which must be the same.
-It is also convenient to attribute to each pulse a pair of coordinates: those will be the coordinates of the sample with the maximum absolute amplitude in each pulse, and will be denoted by $ P[j]^\pm_x $ and $ P[j]^\pm_y $.
-Additionally, the length, or cardinality, of a pulse will be denoted by $ \#P[j]^\pm $.
-
-For an arbitrary continuous function, then, a pulse would correspond to the regions between its roots; for a sinusoid, a pulse would be equal to half of its cycle.
-In analogy, two consecutive pulses, naturally with opposing signs, have the potential to delimitate a pseudo-cycle, as illustrated in figure \ref{fig:Pulses}.
-For that to be true, however, the pulses need to make part of the frontiers of the signal, as yet to be defined.
-
-  \begin{figure}[ht!]
-    \centering
-      \includegraphics[width=0.8\linewidth]{01Pulses.pdf}
-    \caption{Example of a discrete wave.
-The black stems represent the individual samples of the signal.
-The grey areas encompass the simplified convex hulls of the pulses.
-The dashed lines mark the frontiers between pseudo-cycles.}
-    \label{fig:Pulses}
-  \end{figure}
-
-The most mathematically sound definition of an envelope involves the representation of its underlying wave as an analytic signal, introduced by Gabor in 1946 \parencite{2007HahnHistory}.
-In his work, \textcite{1946GaborTheory} applies the then relatively new mathematical machinery of the quantum mechanics to unify the time and frequency representations of a wave, showing how the Hilbert transform could be applied to a real signal in order to obtain an equivalent complex signal, later known as the analytic signal.
-
-This analytic signal has the form $ a(t) = s(t) + \mathbb{H}(s(t)) \ \text{i} $ \parencite{2016HePraat} where $ s(t) $ is the original real signal.
-$ \mathbb{H}(s(t)) $, the Hilbert transform of the original signal, becomes the imaginary part of the analytic signal; the envelope of a signal thus represented can be straightforwardly obtained by the computation of its complex modulus.
-
-Figure \ref{fig:AnalyticSignal} shows the envelope obtained via the analytic signal, for a pure sinusoid with a local frequency of 20 cycles, modulated by a polynomial of the third degree, illustrating the difference in the shape of the envelope in the presence and absence of gaussian white noise with standard deviation of $ 1/10 $ of the wave's maximum amplitude.
-
-\begin{figure}[ht!]
-  \centering
-    \includegraphics[width=0.8\linewidth]{02AnalyticSignal.pdf}
-  \caption{Envelope of a pure sinusoid modulated by a polynomial of degree 3, as obtained by the Hilbert Transform.
-In the first half, the sinusoid is free of noise, while in the second half white gaussian noise with a standard deviation of $ 1/10 $ of the maximum amplitude of the wave was added to the base signal.}
-  \label{fig:AnalyticSignal}
-\end{figure}
-
-It is readily noticeable from figure \ref{fig:AnalyticSignal} that, as soon as noise is introduced in the original signal, it is reflected in the envelope, breaking our assumption of smoothness.
-Other than that, this definition yields a unique, positive envelope.
-
-In practice, however, is not uncommon for a discrete wave, especially in the case of sound, to present somewhat different positive and negative contours; to account for this discrepancy, we will define a positive (upper) and a negative (lower) frontier of a discrete wave.
-They will be composed by the points $ (i, W[i]) $ that are, necessarily but not sufficiently, the points of respectively maximum and minimum amplitude of their pulses, i.e.
-$ (P[j]^\pm_x, P[j]^\pm_y) $.
-
-The set of all these points such that $ W[i] > 0 $ will be called the positive, piecewise linear frontier and will be denoted by $ F^+ $.
-Similarly, $ F^- $ will denote the negative frontier of $ W $, defined by the subset of points $ (i, W[i]) $ where $ W[i] < 0 $, as is shown in figure \ref{fig:Frontiers}.
-
-\begin{figure}[ht!]
-  \centering
-    \includegraphics[width=0.8\linewidth]{03Frontiers.pdf}
-  \caption{Positive and negative frontiers, envelope and pseudo-cycles for a fragment of piano sound.
-While we have only one (positive) envelope, here mirrored in the inferior space, we have two frontiers, the superior and inferior, that are obtained independently.}
-  \label{fig:Frontiers}
-\end{figure}
-
-To identify such points, we need first to address the different units in which the amplitude and horizontal distance of our wave are expressed.
-To make abscissa and ordinate units consistent for the computation of the frontiers, and proceed with a geometric treatment of the subject, we consider the set of positive and negative pulses separately.
-
-For each set, we multiply all pulses vertical coordinates by the average length of the pulses in the set, after having divided by the average of the pulses' amplitudes, putting both axes in the same unit $i$, related to time by the equation $i = t \ \mathit{fps}$.
-
-Formally, supposing we have $ m $ pulses of the same sign, we scale each pulse's vertical coordinate $ P[j]^\pm_y $ by $ s^\pm $, where:
-
-\begin{equation} \label{eq:scale}
-  s^\pm = \frac{\sum_{j=1}^{m} \#P[j]^\pm}{\sum_{j=1}^{m} P[j]^\pm_y}
-\end{equation}
-
-Because we are interested in approximating a function that is smooth over the whole domain of the wave, we can assume that, in the scale of a pulse, this function will be very close to a horizontal straight line and, thus, won't touch any point inside the convex hull defined by the pulse.
-
-By the same token, as the change in amplitude from one pulse to the next pulse with the same sign is motivated by changes in this same envelope, one can reasonably suppose that $ \max(P[j]) \approx \max(P[j+2]) $.
-
-This makes clear that only one point of an arbitrary pulse can be in the frontier, namely, its point of maximum amplitude; those definitions will allow us to derive the frontiers of a wave using an alpha-shapes inspired algorithm without the need to compute the Delaunay triangulation first.
-
-That also justifies the approximation of the convex hull of a pulse by the triangle defined by the points where the pulse crosses the horizontal axis and the point of maximum amplitude $ (P[j]^\pm_x, P[j]^\pm_y) $; this triangular pulse and, more specifically, its point of maximum, will become the atomic unit for the rest of the method.
-
-Translating the intuitive explanation of alpha-shapes in \textcite{1994EdelsbrunnerThree} to our context, we can say that the points in the frontier are those touched by a circle of a given radius $ r $ outside the signal that is not allowed to contain any point of the signal; the pulses that contain said points can be seen as frontier pulses.
-
-Intuitively, one can picture a circle being rolled above (or below, in the case of the negative frontier) the signal, and marking the points it touches as frontier points.
-
-We can now revisit the concept of a pseudo-cycle, only hinted at before, defining it as the samples of the original wave that belong to two successive pulses that, in turn, belong to their respective frontiers, including eventual pulses between them, as can be seen in figure \ref{fig:Frontiers}.
-
-It is necessary to infer the appropriate radius $ r $ of such a circle and, to that end, a measure of the curvature of a discrete function is needed.
-Discrete curvature estimation is an important task in image processing \parencite{2010FleischmannNovel} for which no default definition exists.
-
-The two possible general approaches are the derivation of direct methods that use characteristics of the discrete wave to calculate the curvature, or the use of the smooth curvature of a curve fitted to the discrete wave \parencite{2001CoeurjollyDiscrete}.
-
-Concerning the last method, two approaches to fit a polynomial to a generic wave are readily available: the least mean squares approach, that seeks to minimize the dependent variable errors and the total least squares problem, that treats both variables symmetrically \parencite{1980GolubAnalysis}.
-
-The geometric, symmetrical nature of the problem excludes the more computationally economic least mean squares approach, however, leaving us with the total least squares, for which no general closed-form solutions are available \parencite{2007MarkovskyOverview}; this leads us to turn our attention to direct approaches of curvature estimation.
-
-\subsection{Discrete Curvature Estimation - The Equivalent Circle Approach}
-
-As previously noted, no default discrete curvature definition exists; \textcite{2014CarrollSurvey}, for example, derive three such different definitions based on the approximation of a circle by an inscribed, centred and circumscribed polygon.
-In the context of 3d meshes, \textcite{2016VasaMultivariate} evaluate a range of existing estimators from a multivariate point of view.
-
-None of those approaches, especially for being defined over the vertices of a discrete wave, fit seamlessly into our algorithm; we thus proceed to define a discrete curvature measure over the edges of a discrete wave.
-To that end we are going to apply the definition of smooth curvature as the rate of change of the unit tangent to a curve, noting that this is equivalent to that of the osculating circle \parencite{2016VasaMesh}.
-
-The idea is to find the equivalent circle whose tangent presents the same change in direction, in the same horizontal distance, as the edge of interest, as shown in figure \ref{fig:DiscreteCurvature}.
-
-\begin{figure}[ht!]
-  \centering
-    \includegraphics[width=0.2\linewidth]{04DiscreteCurvature.pdf}
-  \caption{The tangent unit vector of the circle changes from the horizontal direction in $ \hat{u}_0 $ to an inclination of $ \theta $ in $ \hat{u}_1 $.
-$ \theta $ is also the angle between the average direction $ \vec{\overline{v}}_l $ and the vector $ \vec{v}_l $ that represents the edge of interest, shown in grey.}
-  \label{fig:DiscreteCurvature}
-\end{figure}
-
-For each frontier, the angle $ \theta $ between the average direction and each of the vectors can be obtained via the relation between the slopes $ d $ of the respective lines, as in the expression $ \theta = \arctan( (d - \overline{d}) / (1 + d \overline{d}) ) $, with $\overline{d} = \sum_{l=1}^{m-1} (P[l+1]_y - P[l]_y) / (P[l+1]_x - P[l]_x) $ and $ d_l = (P[l+1]_y - P[l]_y)/ (P[l+1]_x - P[l]_x) $
-
-The radii of the equivalent circles are thus given by $ r_l = x_l / \sin (\theta_l) $ with the curvature given by $\kappa_l = 1/r_l=\sin (\theta_l) / x_l$, where $ x_l = P[l+1]_x - P[l]_x $ and $ y_l = P[l+1]_y - P[l]_y $.
-Joining both equations, one can write:
-
-\begin{equation} \label{eq:radius}
-  r_l = \left\lvert \frac{x_l \sqrt{\overline{d}^2 + 1} \ \sqrt{x_l^2 + y_l^2}}{\overline{d} x_l - y_l} \right\rvert
-\end{equation}
-
-After this procedure, illustrated in figure \ref{fig:Curvatures}, two sets of points in $ \mathbb{R}^2 $, corresponding to the positive and negative frontiers, are obtained, where the ordinates represent the instantaneous radius, with its localization in the time domain given by the abscissas.
-
-\begin{figure}[ht!]
-  \centering
-    \includegraphics[width=0.8\linewidth]{05Curvatures.pdf}
-  \caption{The equivalent circles obtained via the procedure described for the positive frontier.
-$\theta$, the angle between $ \vec{\overline{v}}^+$ and each of the $\vec{v}_l^+$ is used, as is the horizontal component of each $\vec{v}_l^+$ to obtain the radii of the circles.}
-  \label{fig:Curvatures}
-\end{figure}
-
-Given the robustness of the method, for an average wave of a few seconds at an ordinary rate of 44100 $ \mathit{fps} $ the average of the radius values is sufficient to assure the accurate identification of the frontier.
-However, a curve fitting method can be used in the case of longer or ill-behaved waves.
-An example of the use of the average equivalent circle can be seen in figure \ref{fig:AvgCircle}.
-
-\begin{figure}[ht!]
-  \centering
-    \includegraphics[width=0.8\linewidth]{06AvgCircle.pdf}
-  \caption{The positive frontier as obtained by the use of the circle representing the average curvature of the discrete wave illustrated.}
-  \label{fig:AvgCircle}
-\end{figure}
-
-\section{Results}
-
-Figure \ref{fig:FullFrontiers} illustrates the frontiers of six diverse discrete sound waves, as well as an in-detail view of the highlighted segment for each signal.
-All waves are records of physical sounds, chosen to represent the applicability of the algorithm in real-world scenarios.
-
-\begin{figure}[ht!]
-  \centering
-    \includegraphics[width=0.8\linewidth]{07FullFrontiers.pdf}
-  \caption{Positive and negative frontiers of six digital waves, as extracted by the algorithm here presented.
-For each wave, the region highlighted in black is shown in detail besides the whole wave.
-For each wave the horizontal axis is the sample number $i$, while the vertical axis is the normalized amplitude.
-All 6 waves were recorded at 44100 $ \mathit{fps} $.}
-  \label{fig:FullFrontiers}
-\end{figure}
-
-We can see that the frontier is satisfactorily detected in the case of harmonic and inharmonic sounds, and is robust in relation to the number of samples and the frequencies of the waves.
-
-The impact of the normalization of a wave using the frontiers obtained can be seen in figure \ref{fig:Fourier}, which shows the original and normalized wave in the case of a recording of the key 33 of a piano both in the time and frequency domains.
-We can see that the Fourier power spectra is greatly simplified in the case of the normalized wave.
-
-\begin{figure}[ht!]
-  \centering
-    \includegraphics[width=0.8\linewidth]{08Fourier.pdf}
-  \caption{Original and normalized views of a recording of key 33 (F3, 175 Hz) from a grand piano both in time and frequency domains.}
-  \label{fig:Fourier}
-\end{figure}
-
-\subsection{Robustness}
-
-Figures \ref{fig:xRobustness} and \ref{fig:yRobustness} illustrate the computational resilience of the algorithm for scaling in both the abscissa and ordinate axes, respectively.
-This is due to the normalization scheme presented in equation \ref{eq:scale}.
-
-Although important from a theoretical standpoint, this step has an especial computational relevance in the case of digital sound waves, as they tend to be composed of a high number of samples with relatively low amplitudes, generally normalized between -1 and 1, leading to a numerical disparity between axes.
-
-\begin{figure}[ht!]
-  \centering
-    \includegraphics[width=0.8\linewidth]{10xRobustness.pdf}
-  \caption{Minimum and maximum scales of the abscissa before computational stability issues arose, for a recording of a drum kit tom.  $ \text{B} = 10^6 $.}
-  \label{fig:xRobustness}
-\end{figure}
-
-\begin{figure}[ht!]
-  \centering
-    \includegraphics[width=0.8\linewidth]{09yRobustness.pdf}
-  \caption{Minimum and maximum values of the amplitude before computational stability issues arose, for a recording of a drum kit tom.}
-  \label{fig:yRobustness}
-\end{figure}
-
-\subsection{Comparison with traditional algorithms}
-
-Direct comparison with many of the more recent algorithms is made difficult by the unavailability of digital implementations of such works, many designed to process analogue signals \parencite[e.g.,][]{2018AssefModeling}.
-
-Nevertheless, insight can be gained from a comparison of the results of the method here proposed with some of the most common envelope extraction algorithms, as can be seen in figure \ref{fig:Comparison}.
-
-\begin{figure}[ht!]
-  \centering
-    \includegraphics[width=0.8\linewidth]{12Comparison.pdf}
-  \caption{Comparison of the algorithm present in this work with the most common methods of digital envelope identification.}
-  \label{fig:Comparison}
-\end{figure}
-
-For a recording of a guitar sound, in which a bend is performed, we contrast the proposed algorithm with four common approaches of varying complexity: simple smoothing of the rectified original wave, low pass filtering followed by rectification, peak identification and subsequent smoothing, and the extraction of the envelope via Hilbert transform, as presented earlier in the paper, with a previous smoothing of the original wave to avoid the artifacts shown in figure \ref{fig:AnalyticSignal}.
-
-All those methods demanded careful choice of parameters: in the case of filtering, a cut-off frequency of approximately 88 Hz was used, consisting of one additional parameter to be tuned.
-For all the other traditional approaches, the Savitzky-Golay smoothing algorithm was used, and the window size was the parameter to be defined.
-
-In the case of the proposed algorithm, the smooth temporal envelope was obtained merging both positive and negative frontiers, the result being smoothed with the same procedure described.
-Figure \ref{fig:FrontiersToEnvelope} illustrates the results.
-
-\begin{figure}[ht!]
-  \centering
-    \includegraphics[width=0.8\linewidth]{13FrontiersToEnvelope.pdf}
-  \caption{Original positive and negative frontiers in dark grey and their mirrored versions. The temporal envelope, in black, is obtained by merging and approximating both frontiers.}
-  \label{fig:FrontiersToEnvelope}
-\end{figure}
-
-Table \ref{table:Comparison} presents a computational comparison between the techniques.
-The times are for custom implementations of those methods in the Python programming language.
-The least mean squares result was calculated by halving the envelopes presented in figure \ref{fig:Comparison} and comparing them with the rectified absolute wave, so as to avoid the potential influence of an estimated frequency.
-
-We can see that, while the algorithm here presented is the most accurate, it's also the most computationally intensive.
-The connection between alpha-shapes and Delaunay triangulation can be explored in future works as a means to improve the efficiency of the algorithm.
-
-\begin{table}[ht!]
-\centering
-\begin{tabular}{ l c c }
-\hline
-Method & LMS & time(s) \\
-\hline
-Presented Algorithm & 0.0088 & 4.1507 \\ 
-Smoothing           & 0.0123 & 0.1911 \\ 
-Low pass Filter      & 0.0258 & 0.5601 \\ 
-Hilbert Transform   & 0.0141 & 0.2599 \\ 
-\hline
-\end{tabular}
-\caption{Comparison of the algorithm present in this work with the most common methods of digital envelope identification.}
-\label{table:Comparison}
-\end{table}
-
-\subsection{Retrieving information about the underlying wave}
-
-The algorithm provides the position of the pseudo-cycles of a wave, with which one can, for example, derive its fundamental frequency and phase: using one of the two frontiers, we have $ f = (\#F^\pm - 1) n / (F^\pm[-1] - F^\pm[0])$, where $ \#F^\pm $ denotes the cardinality of the frontier, while $ F^\pm[0] $ and $ F^\pm[-1] $ stand for the frontier's first and last item, respectively.
-
-Similarly, each point of each of the frontiers can be used to estimate the local phase at that location.
-For the positive frontier, it is necessary to note that, for each point, the following equation holds: $ \cos(\phi + 2 \pi f i / n) = 1 $, and thus $ \phi = 2 \pi k - 2 \pi f i / n $, with $ k \in \mathbb{N} $, while for the negative frontier $ \cos(\phi + 2 \pi f i / n) = -1 $ and $ \phi = 2 \pi k + \pi - 2 \pi f i / n) $.
-
-\section{Discussion}
-
-This work hopes to fill a gap identified by the authors in the context of sound synthesis, where the lack of procedures for the accurate identification of the temporal envelope of arbitrary waves consists of an obstacle to the complete description and eventual manipulation of signals.
-
-While relevant in its own accord, the procedure here presented isolates the individual pseudo-cycles of a wave, pinpointing them in the time domain.
-Due to that, a rich set of theoretical advancements, of whose an example was given in the end of the preceding section, can be built upon this initial algorithm.
-
-With the knowledge of the exact position and characteristics of the pseudo-cycles of a wave, one can investigate its evolution in time, in terms of waveform, for example, free from the potential interference of a modulating wave; in other words, one can have a very satisfactory view of the instantaneous shape and phase of an arbitrary wave and its evolution in time, providing a way to bridge the gap between time domain and frequency domain interpretations.
-
-In the context of sound synthesis, for example, one would be able to apply specific methods for the recreation of the envelope, more in line with its smooth, slow varying and non-periodic nature and a different approach to the periodic and relatively fast changes of the temporal fine structure.
-
-\subsection{Conclusion}
-
-It is our opinion that such theoretical advances can be used in the formulation of better representations of discrete sounds, that in turn can improve sound compression techniques, while providing more adequate representation for the application of machine learning techniques for both sound classification and synthesis.
-
-\section{References}
-
-\printbibliography[heading=none]
-
-\end{document}
+% one hot

### 2020-10-14 Papers/02_Sound_Representation/SoundRepresentation.tex
moved
progress with the paper

--- 2020-09-07 Papers/02_Pulses_Envelope/PulsesEnvelope.tex
@@
 \usepackage[T1]{fontenc}
-\usepackage{authblk}
 \usepackage{amsfonts} % \mathbb
@@
 \usepackage{graphicx}
-\usepackage[backend=biber, style=apa]{biblatex}
+% \graphicspath{{images/}}
+\usepackage[backend=biber]{biblatex}
 \usepackage{amsmath}
-\usepackage{hyperref}
-\usepackage{hypcap}
-\DeclareMathOperator{\sign}{sign}
+
+% \usepackage{auto-pst-pdf}
+
+% \usepackage{hyperref}
+% \usepackage{hypcap}
+% \DeclareMathOperator{\sign}{sign}
+% \hypersetup{
+%   colorlinks=true, 
+%   allcolors=black,
+%   pdfauthor={Carlos Henrique Tarjano Santos},
+%   pdftitle={Robust Digital Envelope Estimation Via Geometric Properties of an Arbitrary Real Signal}
+% }
+
 \bibliography{bibli}
-\hypersetup{
-  colorlinks=true, 
-  allcolors=black,
-  pdfauthor={Carlos Henrique Tarjano Santos},
-  pdftitle={Robust Digital Envelope Estimation Via Geometric Properties of an Arbitrary Real Signal}
-}
+
+\usepackage{authblk}
 \title{Robust Digital Envelope Estimation Via Geometric Properties of an Arbitrary Real Signal}
@@
 \begin{abstract}
-Despite being an elusive concept, the temporal amplitude envelope of a signal is essential for its complete characterization, being the primary information-carrying medium in spoken voice and telecommunications, for example.
-Intuitively, the temporal envelope can be understood as a smooth function that multiplies the signal, being responsible for its outer shape.
-Envelope detection techniques have applications in areas like health, sound classification and synthesis, seismology and speech recognition.
-Nevertheless, a general method to digital envelope detection of signals with rich spectral content doesn't exist, as most methods involve manual intervention, in the form of filter design, smoothing, as well as other specific design choices, based on a priori knowledge about the nature of the specific waves under investigation.
-To address this problem, we propose a framework that uses intrinsic characteristics of a signal to estimate its envelope, completely eliminating the necessity of parameter tuning.
-The approach here described draws inspiration from geometric concepts to isolate the frontiers and thus estimate the temporal envelope of an arbitrary signal; to that end, alpha-shapes, concave hulls, and discrete curvature are explored.
-We also define entities, such as a pulse and frontiers, in the context of an arbitrary digital signal, as a means to reduce dimensionality and the complexity of the proposed algorithm.
-Specifically, a new measure of discrete curvature is used to obtain the average radius of curvature of a discrete wave, serving as a threshold to identify the wave's frontier points.
-We find the algorithm accurate in the identification of the frontiers of a wide range of digital sound waves with very diverse characteristics, while localizing each pseudo-cycle of the wave in the time domain.
-The algorithm also compares favourably with classic envelope detection techniques based on filtering and the Hilbert Transform.
-Besides the most direct applications of this work to audio classification and synthesis, we foresee impact in compression techniques and machine learning approaches to audio.
-The discrete curvature definition presented could also be extended to three-dimensional settings, to improve shape detection algorithms based on alpha-shapes.
+
 \end{abstract}
@@
 \section{Introduction}
-Envelope detection is ubiquitous in both analogue and digital signal processing \parencite{2011CaetanoImproved}.
-Nevertheless, the literature in this area is very fragmented \parencite{2017LyonsDigital}.
-Besides, most envelope detection techniques are designed to account for very specific settings, like pure sinusoids with moderate noise content, a limitation that excludes most physical signals, as is the case of recorded sound, for example; that limitation arises in part due to the lack of a strict mathematical definition of a temporal envelope \parencite{2013Mengempirical}.
-
-In many natural signals, however, the temporal amplitude envelope of a signal plays a prominent role in the characteristics exhibited: according to \textcite{2017QiRelative}, for example, the envelope is at least as important as the fine structure of a sound wave in the context of the intelligibility of mandarin tones.
-This is also the case for English \parencite{1995ShannonSpeech}, where even envelopes modulating mostly noise were still capable of conveying meaning.
-
-The envelope helps to impart emotion and identity to the human voice \parencite{2018ZhuContributions}, and envelope preserving characteristics in concert halls are associated with their pleasantness \parencite{2011LokkiEngaging}.
-
-When dealing with broadband signals, approaches tailored to specific applications are prevalent, such as the one presented by \textcite{2014YangFast} for the distributed monitoring of fibre optic or the one formulated by \textcite{2018AssefModeling} in the context of medical ultrasound imaging.
-
-Addressing the different units in the horizontal and vertical axes, one can transform the DSP problem of envelope detection into the geometric problem of defining the shape of a set of points in $ \mathbb{R}^2 $.
-For that purpose, \textcite{1983Edelsbrunnershape} introduced the concept of alpha-shapes, a mathematically well-defined extension to the convex hull of a finite set of points, closely related to the Delaunay triangulation and Voronoi diagrams of those points.
-
-This approach is used in areas such as the detection of features in images \parencite{2016VarytimidisAlpha}, reconstruction of surfaces from a cloud of points \parencite{2015WuAutomated} and Spectroscopy \parencite{2019XuModeling}, with the last work, which involves the estimation and removal of the Blaze function (a kind of envelope) of an echelle spectrograph, being particularly illustrative of the potential of cross-fertilization.
-
-Steps in the direction of translating geometric algorithms to the context of envelope detection were also made by \textcite{2015YangSkeleton} via an algorithm based on the construction of a skeleton underlying the wave, and also via the direct translation of computer vision methods to the task \parencite{2015YangRepresenting}.
-
-Following this path, we present a general approach to envelope detection, exploiting the intrinsic characteristics of a generic, spectrally complex wave, in order to avoid the need for manual intervention or parameter tuning.
-
-While the robustness of the proposed approach allows it to be used as a plug-in replacement for many methods encountered in the literature, we feel that it would be particularly useful for sound synthesis.
-
-The envelope is shown to add complexity to the spectral representation of a wave \parencite{2019TarjanoNeuro}, and an accurate description of the envelope, while describing the evolution of the instantaneous amplitude of a signal in time, would also greatly simplify further spectral analysis.
-Moreover, the algorithm developed in this work naturally divides a signal into its pseudo-cycles, pinpointing them in the time domain, providing the building blocks for the reconstruction of the fine structure of the wave.
-
-The rest of this paper is divided as follows: After explaining and defining the various entities to be used subsequently, of which the concept of frontiers is arguably the most important, in the methodology section, we dedicate a section to present a new discrete curvature estimation approach, and how it was used to identify the frontiers of a discrete wave.
-
-We then proceed to illustrate the results of the application of the algorithm in a set of six diverse sound waves, ranging from voice to unpitched instruments comparing, afterwards, its performance with that of classic digital envelope estimation algorithms, both from a visual and numerical standpoint.
-
-We then discuss the implications of the work, emphasizing the focus on the construction of new interpretations of the relationship between time domain and frequency domain sound descriptions, and our vision of the impact of this methodology in the field of sound synthesis, providing directions for future developments.
 
@@
 
-In this work it is assumed, without loss of generality, that one is interested in retrieving the mathematically ill-defined \parencite{2013Mengempirical} temporal envelope of an arbitrary physical sound wave, of which only a digital representation is available.
+Using the algorithm presented in \textcite{2020TarjanoRobust} one can identify the frontiers, both positive and negative, of a digital signal: Those are, ideally, respectively composed of the maxima and minima of each pseudo-cycle of a discrete signal.
 
-Naturally, due to the necessary truncations both in time and amplitude, as well as measurement errors introduced by the instruments used to record the signal, it is not possible to exactly recover the original, continuous wave, from its digital counterpart.
+If we disconsider inaccuracies we could use either one of the frontiers to infer the exact position, wavelength and amplitude of each pseudo-cycle and to, provided that the shape of the wave is also known, reconstruct the original wave with those pieces of information.
+Consider, for example, the digital signal presented in figure \ref{fig:signalenvelope}
+and the extracted envelopes; the signal is an example of the voice of an alto singer, with its frontiers in black.
 
-The best representation one can hope to get of the original wave is, then, the recording itself.
+\begin{figure}[ht!]
+  \centering
+    %  \includegraphics[width=0.9\linewidth]{01signalenvelope}
+     \includegraphics{01signalenvelope}
+    % \def\svgwidth{\linewidth}
+    %\input{01signalenvelope.pdf_tex}
+  \caption{A discrete wave of the singing voice of an alto singer.
+The black lines are the frontiers; in the detail view, it can be seen that each region between two diamonds comprises a pseudo-cycle, as defined by the points belonging to the upper frontier.
+Similarly, two adjacent squares delimitate a pseudo-cycle, from the standpoint of the lower frontier.}
+  \label{fig:signalenvelope}
+\end{figure}
+
+It can be seen that both frontiers provide approximately the same information about the location of each pseudo-cycle, with a $ \pi $ phase difference.
+
+\begin{figure}[ht!]
+  \centering
+    %  \includegraphics[width=0.9\linewidth]{01signalenvelope}
+     \includegraphics{02pcsraw-alto}
+    % \def\svgwidth{\linewidth}
+    %\input{01signalenvelope.pdf_tex}
+  \caption{Raw pseudo-cycles superimposed as defined by the positive and negative frontiers, for the alto singer signal.}
+  \label{fig:pcrawalto}
+\end{figure}
+
+We can normalize the amplitude the pseudo-cycles between zero and one, by dividing each of them by their maximum.
+The lengths of each pseudo-cycle can be normalized with the use of the Fourier transform; to avoid any loss of data, we can use the maximum wavelength as the standard.
+
+\begin{figure}[ht!]
+  \centering
+    %  \includegraphics[width=0.9\linewidth]{01signalenvelope}
+     \includegraphics{03pcs-alto}
+    % \def\svgwidth{\linewidth}
+    %\input{01signalenvelope.pdf_tex}
+  \caption{Normalized pseudo-cycles superimposed as defined by the positive and negative frontiers, for the alto singer signal, and the average of all normalized pseudo-cycles.}
+  \label{fig:pcalto}
+\end{figure}
+
+Comparing figures \ref{fig:pcrawalto} and \ref{fig:pcalto} we note the effect of the normalization, as a means to reduce the variance of the pseudo-cycles, specially in the regions near the discontinuities.
+This will enable a more accurate approximation of the average wave.
+
+\begin{figure}[ht!]
+  \centering
+    %  \includegraphics[width=0.9\linewidth]{01signalenvelope}
+     \includegraphics{04averagepc-alto}
+    % \def\svgwidth{\linewidth}
+    %\input{01signalenvelope.pdf_tex}
+  \caption{Average pseudo-cycles, superimposed, after a change of phase of $-\pi$ on the average pseudo-cycle defined by the negative frontier.}
+  \label{fig:averagepc-alto}
+\end{figure}
+
+Figure \ref{fig:averagepc-alto} makes clear that the average pseudo-cycles, as estimated by both frontiers, are in reasonable agreement, despite the evident vertical asymmetries of the underlying wave.
+
+This, however, is not always the case, as
+
+% In this work it is assumed, without loss of generality, that one is interested in retrieving the mathematically ill-defined \parencite{2013Mengempirical} temporal envelope of an arbitrary physical sound wave, of which only a digital representation is available.
+
+% Naturally, due to the necessary truncations both in time and amplitude, as well as measurement errors introduced by the instruments used to record the signal, it is not possible to exactly recover the original, continuous wave, from its digital counterpart.
+
+% The best representation one can hope to get of the original wave is, then, the recording itself.
 It is sensible, thus, to characterize the envelope of the wave in terms of a subset of the exact samples that comprise its digital representation.
 
-To that end, this work will emphasize the concept of frontier, as the points of the digital wave that mark the wave's upper and lower boundaries; from those points an envelope, conforming to the vague definitions of uniqueness and smoothness, can be trivially constructed, as will be shown.
+% To that end, this work will emphasize the concept of frontier, as the points of the digital wave that mark the wave's upper and lower boundaries; from those points an envelope, conforming to the vague definitions of uniqueness and smoothness, can be trivially constructed, as will be shown.
 
-We start by defining a pulse as a series of consecutive samples of a discrete wave that have the same sign.
+% We start by defining a pulse as a series of consecutive samples of a discrete wave that have the same sign.
 More formally, let $ W[i] \in \mathbb{R} \ \forall \ i \in \mathbb{N}_0; \ i < n \in \mathbb{N}_0 $ be a real, discrete and finite signal indexed, without loss of generality, over an adjacent subset of the natural numbers starting at zero.
@@
 
-A pulse $ P[j]; \ \forall \ j \in \mathbb{N}_0; \ j < q \in \mathbb{N}_0; q \le n $ is, then, a subset of $ W $ with cardinality $ \#P[j] = b - a $ such that $ \sign(W[a-1]) \ne \sign(W[a]) = \sign(W[a+1]) = \dots = \sign(W[b-1]) = \sign(W[b]) \ne \sign(W[b+1]) \ \forall \ i; \ a \le i \le b; \ a, b \in \mathbb{N}_0 $.
+% A pulse $ P[j]; \ \forall \ j \in \mathbb{N}_0; \ j < q \in \mathbb{N}_0; q \le n $ is, then, a subset of $ W $ with cardinality $ \#P[j] = b - a $ such that $ \sign(W[a-1]) \ne \sign(W[a]) = \sign(W[a+1]) = \dots = \sign(W[b-1]) = \sign(W[b]) \ne \sign(W[b+1]) \ \forall \ i; \ a \le i \le b; \ a, b \in \mathbb{N}_0 $.
 
-In other words, each pulse is a unique subset of $ W $ with its members being the maximum sequence of samples of $ W $ having the same signal.
+% In other words, each pulse is a unique subset of $ W $ with its members being the maximum sequence of samples of $ W $ having the same signal.
 Thus $ P[0] \cup P[1] \cup \cdots \cup P[q-1] = W $ and $ P[i_1] \cap P[i_2] = \emptyset \ \forall \ i_1, i_2 \in \mathbb{N}_0;\ i_1 \ne i_2; \ i_1, i_2 < q $.
 
-Naturally, each pulse can be classified as positive ($ P[j]^+ $) or negative ($ P[j]^- $), based on the sign of its samples, which must be the same.
+% Naturally, each pulse can be classified as positive ($ P[j]^+ $) or negative ($ P[j]^- $), based on the sign of its samples, which must be the same.
 It is also convenient to attribute to each pulse a pair of coordinates: those will be the coordinates of the sample with the maximum absolute amplitude in each pulse, and will be denoted by $ P[j]^\pm_x $ and $ P[j]^\pm_y $.
@@
 
-For an arbitrary continuous function, then, a pulse would correspond to the regions between its roots; for a sinusoid, a pulse would be equal to half of its cycle.
+% For an arbitrary continuous function, then, a pulse would correspond to the regions between its roots; for a sinusoid, a pulse would be equal to half of its cycle.
 In analogy, two consecutive pulses, naturally with opposing signs, have the potential to delimitate a pseudo-cycle, as illustrated in figure \ref{fig:Pulses}.
@@
 
-  \begin{figure}[ht!]
-    \centering
-      \includegraphics[width=0.8\linewidth]{01Pulses.pdf}
-    \caption{Example of a discrete wave.
-The black stems represent the individual samples of the signal.
-The grey areas encompass the simplified convex hulls of the pulses.
-The dashed lines mark the frontiers between pseudo-cycles.}
-    \label{fig:Pulses}
-  \end{figure}
-
-The most mathematically sound definition of an envelope involves the representation of its underlying wave as an analytic signal, introduced by Gabor in 1946 \parencite{2007HahnHistory}.
+% The most mathematically sound definition of an envelope involves the representation of its underlying wave as an analytic signal, introduced by Gabor in 1946 \parencite{2007HahnHistory}.
 In his work, \textcite{1946GaborTheory} applies the then relatively new mathematical machinery of the quantum mechanics to unify the time and frequency representations of a wave, showing how the Hilbert transform could be applied to a real signal in order to obtain an equivalent complex signal, later known as the analytic signal.
 
-This analytic signal has the form $ a(t) = s(t) + \mathbb{H}(s(t)) \ \text{i} $ \parencite{2016HePraat} where $ s(t) $ is the original real signal.
+% This analytic signal has the form $ a(t) = s(t) + \mathbb{H}(s(t)) \ \text{i} $ \parencite{2016HePraat} where $ s(t) $ is the original real signal.
 $ \mathbb{H}(s(t)) $, the Hilbert transform of the original signal, becomes the imaginary part of the analytic signal; the envelope of a signal thus represented can be straightforwardly obtained by the computation of its complex modulus.
 
-Figure \ref{fig:AnalyticSignal} shows the envelope obtained via the analytic signal, for a pure sinusoid with a local frequency of 20 cycles, modulated by a polynomial of the third degree, illustrating the difference in the shape of the envelope in the presence and absence of gaussian white noise with standard deviation of $ 1/10 $ of the wave's maximum amplitude.
+% Figure \ref{fig:AnalyticSignal} shows the envelope obtained via the analytic signal, for a pure sinusoid with a local frequency of 20 cycles, modulated by a polynomial of the third degree, illustrating the difference in the shape of the envelope in the presence and absence of gaussian white noise with standard deviation of $ 1/10 $ of the wave's maximum amplitude.
 
-\begin{figure}[ht!]
-  \centering
-    \includegraphics[width=0.8\linewidth]{02AnalyticSignal.pdf}
-  \caption{Envelope of a pure sinusoid modulated by a polynomial of degree 3, as obtained by the Hilbert Transform.
+% \begin{figure}[ht!]
+%   \centering
+%     \includegraphics[width=0.8\linewidth]{02AnalyticSignal.pdf}
+%   \caption{Envelope of a pure sinusoid modulated by a polynomial of degree 3, as obtained by the Hilbert Transform.
 In the first half, the sinusoid is free of noise, while in the second half white gaussian noise with a standard deviation of $ 1/10 $ of the maximum amplitude of the wave was added to the base signal.}
-  \label{fig:AnalyticSignal}
-\end{figure}
+%   \label{fig:AnalyticSignal}
+% \end{figure}
 
-It is readily noticeable from figure \ref{fig:AnalyticSignal} that, as soon as noise is introduced in the original signal, it is reflected in the envelope, breaking our assumption of smoothness.
+% It is readily noticeable from figure \ref{fig:AnalyticSignal} that, as soon as noise is introduced in the original signal, it is reflected in the envelope, breaking our assumption of smoothness.
 Other than that, this definition yields a unique, positive envelope.
 
-In practice, however, is not uncommon for a discrete wave, especially in the case of sound, to present somewhat different positive and negative contours; to account for this discrepancy, we will define a positive (upper) and a negative (lower) frontier of a discrete wave.
+% In practice, however, is not uncommon for a discrete wave, especially in the case of sound, to present somewhat different positive and negative contours; to account for this discrepancy, we will define a positive (upper) and a negative (lower) frontier of a discrete wave.
 They will be composed by the points $ (i, W[i]) $ that are, necessarily but not sufficiently, the points of respectively maximum and minimum amplitude of their pulses, i.e.
@@
 
-The set of all these points such that $ W[i] > 0 $ will be called the positive, piecewise linear frontier and will be denoted by $ F^+ $.
+% The set of all these points such that $ W[i] > 0 $ will be called the positive, piecewise linear frontier and will be denoted by $ F^+ $.
 Similarly, $ F^- $ will denote the negative frontier of $ W $, defined by the subset of points $ (i, W[i]) $ where $ W[i] < 0 $, as is shown in figure \ref{fig:Frontiers}.
 
-\begin{figure}[ht!]
-  \centering
-    \includegraphics[width=0.8\linewidth]{03Frontiers.pdf}
-  \caption{Positive and negative frontiers, envelope and pseudo-cycles for a fragment of piano sound.
+% \begin{figure}[ht!]
+%   \centering
+%     \includegraphics[width=0.8\linewidth]{03Frontiers.pdf}
+%   \caption{Positive and negative frontiers, envelope and pseudo-cycles for a fragment of piano sound.
 While we have only one (positive) envelope, here mirrored in the inferior space, we have two frontiers, the superior and inferior, that are obtained independently.}
-  \label{fig:Frontiers}
-\end{figure}
+%   \label{fig:Frontiers}
+% \end{figure}
 
-To identify such points, we need first to address the different units in which the amplitude and horizontal distance of our wave are expressed.
+% To identify such points, we need first to address the different units in which the amplitude and horizontal distance of our wave are expressed.
 To make abscissa and ordinate units consistent for the computation of the frontiers, and proceed with a geometric treatment of the subject, we consider the set of positive and negative pulses separately.
 
-For each set, we multiply all pulses vertical coordinates by the average length of the pulses in the set, after having divided by the average of the pulses' amplitudes, putting both axes in the same unit $i$, related to time by the equation $i = t \ \mathit{fps}$.
+% For each set, we multiply all pulses vertical coordinates by the average length of the pulses in the set, after having divided by the average of the pulses' amplitudes, putting both axes in the same unit $i$, related to time by the equation $i = t \ \mathit{fps}$.
 
-Formally, supposing we have $ m $ pulses of the same sign, we scale each pulse's vertical coordinate $ P[j]^\pm_y $ by $ s^\pm $, where:
+% Formally, supposing we have $ m $ pulses of the same sign, we scale each pulse's vertical coordinate $ P[j]^\pm_y $ by $ s^\pm $, where:
 
-\begin{equation} \label{eq:scale}
-  s^\pm = \frac{\sum_{j=1}^{m} \#P[j]^\pm}{\sum_{j=1}^{m} P[j]^\pm_y}
-\end{equation}
+% \begin{equation} \label{eq:scale}
+%   s^\pm = \frac{\sum_{j=1}^{m} \#P[j]^\pm}{\sum_{j=1}^{m} P[j]^\pm_y}
+% \end{equation}
 
-Because we are interested in approximating a function that is smooth over the whole domain of the wave, we can assume that, in the scale of a pulse, this function will be very close to a horizontal straight line and, thus, won't touch any point inside the convex hull defined by the pulse.
+% Because we are interested in approximating a function that is smooth over the whole domain of the wave, we can assume that, in the scale of a pulse, this function will be very close to a horizontal straight line and, thus, won't touch any point inside the convex hull defined by the pulse.
 
-By the same token, as the change in amplitude from one pulse to the next pulse with the same sign is motivated by changes in this same envelope, one can reasonably suppose that $ \max(P[j]) \approx \max(P[j+2]) $.
+% By the same token, as the change in amplitude from one pulse to the next pulse with the same sign is motivated by changes in this same envelope, one can reasonably suppose that $ \max(P[j]) \approx \max(P[j+2]) $.
 
-This makes clear that only one point of an arbitrary pulse can be in the frontier, namely, its point of maximum amplitude; those definitions will allow us to derive the frontiers of a wave using an alpha-shapes inspired algorithm without the need to compute the Delaunay triangulation first.
+% This makes clear that only one point of an arbitrary pulse can be in the frontier, namely, its point of maximum amplitude; those definitions will allow us to derive the frontiers of a wave using an alpha-shapes inspired algorithm without the need to compute the Delaunay triangulation first.
 
-That also justifies the approximation of the convex hull of a pulse by the triangle defined by the points where the pulse crosses the horizontal axis and the point of maximum amplitude $ (P[j]^\pm_x, P[j]^\pm_y) $; this triangular pulse and, more specifically, its point of maximum, will become the atomic unit for the rest of the method.
+% That also justifies the approximation of the convex hull of a pulse by the triangle defined by the points where the pulse crosses the horizontal axis and the point of maximum amplitude $ (P[j]^\pm_x, P[j]^\pm_y) $; this triangular pulse and, more specifically, its point of maximum, will become the atomic unit for the rest of the method.
 
-Translating the intuitive explanation of alpha-shapes in \textcite{1994EdelsbrunnerThree} to our context, we can say that the points in the frontier are those touched by a circle of a given radius $ r $ outside the signal that is not allowed to contain any point of the signal; the pulses that contain said points can be seen as frontier pulses.
+% Translating the intuitive explanation of alpha-shapes in \textcite{1994EdelsbrunnerThree} to our context, we can say that the points in the frontier are those touched by a circle of a given radius $ r $ outside the signal that is not allowed to contain any point of the signal; the pulses that contain said points can be seen as frontier pulses.
 
-Intuitively, one can picture a circle being rolled above (or below, in the case of the negative frontier) the signal, and marking the points it touches as frontier points.
+% Intuitively, one can picture a circle being rolled above (or below, in the case of the negative frontier) the signal, and marking the points it touches as frontier points.
 
-We can now revisit the concept of a pseudo-cycle, only hinted at before, defining it as the samples of the original wave that belong to two successive pulses that, in turn, belong to their respective frontiers, including eventual pulses between them, as can be seen in figure \ref{fig:Frontiers}.
+% We can now revisit the concept of a pseudo-cycle, only hinted at before, defining it as the samples of the original wave that belong to two successive pulses that, in turn, belong to their respective frontiers, including eventual pulses between them, as can be seen in figure \ref{fig:Frontiers}.
 
-It is necessary to infer the appropriate radius $ r $ of such a circle and, to that end, a measure of the curvature of a discrete function is needed.
+% It is necessary to infer the appropriate radius $ r $ of such a circle and, to that end, a measure of the curvature of a discrete function is needed.
 Discrete curvature estimation is an important task in image processing \parencite{2010FleischmannNovel} for which no default definition exists.
 
-The two possible general approaches are the derivation of direct methods that use characteristics of the discrete wave to calculate the curvature, or the use of the smooth curvature of a curve fitted to the discrete wave \parencite{2001CoeurjollyDiscrete}.
+% The two possible general approaches are the derivation of direct methods that use characteristics of the discrete wave to calculate the curvature, or the use of the smooth curvature of a curve fitted to the discrete wave \parencite{2001CoeurjollyDiscrete}.
 
-Concerning the last method, two approaches to fit a polynomial to a generic wave are readily available: the least mean squares approach, that seeks to minimize the dependent variable errors and the total least squares problem, that treats both variables symmetrically \parencite{1980GolubAnalysis}.
+% Concerning the last method, two approaches to fit a polynomial to a generic wave are readily available: the least mean squares approach, that seeks to minimize the dependent variable errors and the total least squares problem, that treats both variables symmetrically \parencite{1980GolubAnalysis}.
 
-The geometric, symmetrical nature of the problem excludes the more computationally economic least mean squares approach, however, leaving us with the total least squares, for which no general closed-form solutions are available \parencite{2007MarkovskyOverview}; this leads us to turn our attention to direct approaches of curvature estimation.
+% The geometric, symmetrical nature of the problem excludes the more computationally economic least mean squares approach, however, leaving us with the total least squares, for which no general closed-form solutions are available \parencite{2007MarkovskyOverview}; this leads us to turn our attention to direct approaches of curvature estimation.
 
-\subsection{Discrete Curvature Estimation - The Equivalent Circle Approach}
+% \subsection{Discrete Curvature Estimation - The Equivalent Circle Approach}
 
-As previously noted, no default discrete curvature definition exists; \textcite{2014CarrollSurvey}, for example, derive three such different definitions based on the approximation of a circle by an inscribed, centred and circumscribed polygon.
+% As previously noted, no default discrete curvature definition exists; \textcite{2014CarrollSurvey}, for example, derive three such different definitions based on the approximation of a circle by an inscribed, centred and circumscribed polygon.
 In the context of 3d meshes, \textcite{2016VasaMultivariate} evaluate a range of existing estimators from a multivariate point of view.
 
-None of those approaches, especially for being defined over the vertices of a discrete wave, fit seamlessly into our algorithm; we thus proceed to define a discrete curvature measure over the edges of a discrete wave.
+% None of those approaches, especially for being defined over the vertices of a discrete wave, fit seamlessly into our algorithm; we thus proceed to define a discrete curvature measure over the edges of a discrete wave.
 To that end we are going to apply the definition of smooth curvature as the rate of change of the unit tangent to a curve, noting that this is equivalent to that of the osculating circle \parencite{2016VasaMesh}.
 
-The idea is to find the equivalent circle whose tangent presents the same change in direction, in the same horizontal distance, as the edge of interest, as shown in figure \ref{fig:DiscreteCurvature}.
+% The idea is to find the equivalent circle whose tangent presents the same change in direction, in the same horizontal distance, as the edge of interest, as shown in figure \ref{fig:DiscreteCurvature}.
 
-\begin{figure}[ht!]
-  \centering
-    \includegraphics[width=0.2\linewidth]{04DiscreteCurvature.pdf}
-  \caption{The tangent unit vector of the circle changes from the horizontal direction in $ \hat{u}_0 $ to an inclination of $ \theta $ in $ \hat{u}_1 $.
+% \begin{figure}[ht!]
+%   \centering
+%     \includegraphics[width=0.2\linewidth]{04DiscreteCurvature.pdf}
+%   \caption{The tangent unit vector of the circle changes from the horizontal direction in $ \hat{u}_0 $ to an inclination of $ \theta $ in $ \hat{u}_1 $.
 $ \theta $ is also the angle between the average direction $ \vec{\overline{v}}_l $ and the vector $ \vec{v}_l $ that represents the edge of interest, shown in grey.}
-  \label{fig:DiscreteCurvature}
-\end{figure}
+%   \label{fig:DiscreteCurvature}
+% \end{figure}
 
-For each frontier, the angle $ \theta $ between the average direction and each of the vectors can be obtained via the relation between the slopes $ d $ of the respective lines, as in the expression $ \theta = \arctan( (d - \overline{d}) / (1 + d \overline{d}) ) $, with $\overline{d} = \sum_{l=1}^{m-1} (P[l+1]_y - P[l]_y) / (P[l+1]_x - P[l]_x) $ and $ d_l = (P[l+1]_y - P[l]_y)/ (P[l+1]_x - P[l]_x) $
+% For each frontier, the angle $ \theta $ between the average direction and each of the vectors can be obtained via the relation between the slopes $ d $ of the respective lines, as in the expression $ \theta = \arctan( (d - \overline{d}) / (1 + d \overline{d}) ) $, with $\overline{d} = \sum_{l=1}^{m-1} (P[l+1]_y - P[l]_y) / (P[l+1]_x - P[l]_x) $ and $ d_l = (P[l+1]_y - P[l]_y)/ (P[l+1]_x - P[l]_x) $
 
-The radii of the equivalent circles are thus given by $ r_l = x_l / \sin (\theta_l) $ with the curvature given by $\kappa_l = 1/r_l=\sin (\theta_l) / x_l$, where $ x_l = P[l+1]_x - P[l]_x $ and $ y_l = P[l+1]_y - P[l]_y $.
+% The radii of the equivalent circles are thus given by $ r_l = x_l / \sin (\theta_l) $ with the curvature given by $\kappa_l = 1/r_l=\sin (\theta_l) / x_l$, where $ x_l = P[l+1]_x - P[l]_x $ and $ y_l = P[l+1]_y - P[l]_y $.
 Joining both equations, one can write:
 
-\begin{equation} \label{eq:radius}
-  r_l = \left\lvert \frac{x_l \sqrt{\overline{d}^2 + 1} \ \sqrt{x_l^2 + y_l^2}}{\overline{d} x_l - y_l} \right\rvert
-\end{equation}
+% \begin{equation} \label{eq:radius}
+%   r_l = \left\lvert \frac{x_l \sqrt{\overline{d}^2 + 1} \ \sqrt{x_l^2 + y_l^2}}{\overline{d} x_l - y_l} \right\rvert
+% \end{equation}
 
-After this procedure, illustrated in figure \ref{fig:Curvatures}, two sets of points in $ \mathbb{R}^2 $, corresponding to the positive and negative frontiers, are obtained, where the ordinates represent the instantaneous radius, with its localization in the time domain given by the abscissas.
+% After this procedure, illustrated in figure \ref{fig:Curvatures}, two sets of points in $ \mathbb{R}^2 $, corresponding to the positive and negative frontiers, are obtained, where the ordinates represent the instantaneous radius, with its localization in the time domain given by the abscissas.
 
-\begin{figure}[ht!]
-  \centering
-    \includegraphics[width=0.8\linewidth]{05Curvatures.pdf}
-  \caption{The equivalent circles obtained via the procedure described for the positive frontier.
+% \begin{figure}[ht!]
+%   \centering
+%     \includegraphics[width=0.8\linewidth]{05Curvatures.pdf}
+%   \caption{The equivalent circles obtained via the procedure described for the positive frontier.
 $\theta$, the angle between $ \vec{\overline{v}}^+$ and each of the $\vec{v}_l^+$ is used, as is the horizontal component of each $\vec{v}_l^+$ to obtain the radii of the circles.}
-  \label{fig:Curvatures}
-\end{figure}
+%   \label{fig:Curvatures}
+% \end{figure}
 
-Given the robustness of the method, for an average wave of a few seconds at an ordinary rate of 44100 $ \mathit{fps} $ the average of the radius values is sufficient to assure the accurate identification of the frontier.
+% Given the robustness of the method, for an average wave of a few seconds at an ordinary rate of 44100 $ \mathit{fps} $ the average of the radius values is sufficient to assure the accurate identification of the frontier.
 However, a curve fitting method can be used in the case of longer or ill-behaved waves.
-An example of the use of the average equivalent circle can be seen in figure \ref{fig:AvgCircle}.
+% An example of the use of the average equivalent circle can be seen in figure \ref{fig:AvgCircle}.
 
-\begin{figure}[ht!]
-  \centering
-    \includegraphics[width=0.8\linewidth]{06AvgCircle.pdf}
-  \caption{The positive frontier as obtained by the use of the circle representing the average curvature of the discrete wave illustrated.}
-  \label{fig:AvgCircle}
-\end{figure}
+% \begin{figure}[ht!]
+%   \centering
+%     \includegraphics[width=0.8\linewidth]{06AvgCircle.pdf}
+%   \caption{The positive frontier as obtained by the use of the circle representing the average curvature of the discrete wave illustrated.}
+%   \label{fig:AvgCircle}
+% \end{figure}
 
-\section{Results}
+% \section{Results}
 
-Figure \ref{fig:FullFrontiers} illustrates the frontiers of six diverse discrete sound waves, as well as an in-detail view of the highlighted segment for each signal.
+% Figure \ref{fig:FullFrontiers} illustrates the frontiers of six diverse discrete sound waves, as well as an in-detail view of the highlighted segment for each signal.
 All waves are records of physical sounds, chosen to represent the applicability of the algorithm in real-world scenarios.
 
-\begin{figure}[ht!]
-  \centering
-    \includegraphics[width=0.8\linewidth]{07FullFrontiers.pdf}
-  \caption{Positive and negative frontiers of six digital waves, as extracted by the algorithm here presented.
+% \begin{figure}[ht!]
+%   \centering
+%     \includegraphics[width=0.8\linewidth]{07FullFrontiers.pdf}
+%   \caption{Positive and negative frontiers of six digital waves, as extracted by the algorithm here presented.
 For each wave, the region highlighted in black is shown in detail besides the whole wave.
@@
 All 6 waves were recorded at 44100 $ \mathit{fps} $.}
-  \label{fig:FullFrontiers}
-\end{figure}
+%   \label{fig:FullFrontiers}
+% \end{figure}
 
-We can see that the frontier is satisfactorily detected in the case of harmonic and inharmonic sounds, and is robust in relation to the number of samples and the frequencies of the waves.
+% We can see that the frontier is satisfactorily detected in the case of harmonic and inharmonic sounds, and is robust in relation to the number of samples and the frequencies of the waves.
 
-The impact of the normalization of a wave using the frontiers obtained can be seen in figure \ref{fig:Fourier}, which shows the original and normalized wave in the case of a recording of the key 33 of a piano both in the time and frequency domains.
+% The impact of the normalization of a wave using the frontiers obtained can be seen in figure \ref{fig:Fourier}, which shows the original and normalized wave in the case of a recording of the key 33 of a piano both in the time and frequency domains.
 We can see that the Fourier power spectra is greatly simplified in the case of the normalized wave.
 
-\begin{figure}[ht!]
-  \centering
-    \includegraphics[width=0.8\linewidth]{08Fourier.pdf}
-  \caption{Original and normalized views of a recording of key 33 (F3, 175 Hz) from a grand piano both in time and frequency domains.}
-  \label{fig:Fourier}
-\end{figure}
+% \begin{figure}[ht!]
+%   \centering
+%     \includegraphics[width=0.8\linewidth]{08Fourier.pdf}
+%   \caption{Original and normalized views of a recording of key 33 (F3, 175 Hz) from a grand piano both in time and frequency domains.}
+%   \label{fig:Fourier}
+% \end{figure}
 
-\subsection{Robustness}
+% \subsection{Robustness}
 
-Figures \ref{fig:xRobustness} and \ref{fig:yRobustness} illustrate the computational resilience of the algorithm for scaling in both the abscissa and ordinate axes, respectively.
+% Figures \ref{fig:xRobustness} and \ref{fig:yRobustness} illustrate the computational resilience of the algorithm for scaling in both the abscissa and ordinate axes, respectively.
 This is due to the normalization scheme presented in equation \ref{eq:scale}.
 
-Although important from a theoretical standpoint, this step has an especial computational relevance in the case of digital sound waves, as they tend to be composed of a high number of samples with relatively low amplitudes, generally normalized between -1 and 1, leading to a numerical disparity between axes.
+% Although important from a theoretical standpoint, this step has an especial computational relevance in the case of digital sound waves, as they tend to be composed of a high number of samples with relatively low amplitudes, generally normalized between -1 and 1, leading to a numerical disparity between axes.
 
-\begin{figure}[ht!]
-  \centering
-    \includegraphics[width=0.8\linewidth]{10xRobustness.pdf}
-  \caption{Minimum and maximum scales of the abscissa before computational stability issues arose, for a recording of a drum kit tom.  $ \text{B} = 10^6 $.}
-  \label{fig:xRobustness}
-\end{figure}
+% \begin{figure}[ht!]
+%   \centering
+%     \includegraphics[width=0.8\linewidth]{10xRobustness.pdf}
+%   \caption{Minimum and maximum scales of the abscissa before computational stability issues arose, for a recording of a drum kit tom.  $ \text{B} = 10^6 $.}
+%   \label{fig:xRobustness}
+% \end{figure}
 
-\begin{figure}[ht!]
-  \centering
-    \includegraphics[width=0.8\linewidth]{09yRobustness.pdf}
-  \caption{Minimum and maximum values of the amplitude before computational stability issues arose, for a recording of a drum kit tom.}
-  \label{fig:yRobustness}
-\end{figure}
+% \begin{figure}[ht!]
+%   \centering
+%     \includegraphics[width=0.8\linewidth]{09yRobustness.pdf}
+%   \caption{Minimum and maximum values of the amplitude before computational stability issues arose, for a recording of a drum kit tom.}
+%   \label{fig:yRobustness}
+% \end{figure}
 
-\subsection{Comparison with traditional algorithms}
+% \subsection{Comparison with traditional algorithms}
 
-Direct comparison with many of the more recent algorithms is made difficult by the unavailability of digital implementations of such works, many designed to process analogue signals \parencite[e.g.,][]{2018AssefModeling}.
+% Direct comparison with many of the more recent algorithms is made difficult by the unavailability of digital implementations of such works, many designed to process analogue signals \parencite[e.g.,][]{2018AssefModeling}.
 
-Nevertheless, insight can be gained from a comparison of the results of the method here proposed with some of the most common envelope extraction algorithms, as can be seen in figure \ref{fig:Comparison}.
+% Nevertheless, insight can be gained from a comparison of the results of the method here proposed with some of the most common envelope extraction algorithms, as can be seen in figure \ref{fig:Comparison}.
 
-\begin{figure}[ht!]
-  \centering
-    \includegraphics[width=0.8\linewidth]{12Comparison.pdf}
-  \caption{Comparison of the algorithm present in this work with the most common methods of digital envelope identification.}
-  \label{fig:Comparison}
-\end{figure}
+% \begin{figure}[ht!]
+%   \centering
+%     \includegraphics[width=0.8\linewidth]{12Comparison.pdf}
+%   \caption{Comparison of the algorithm present in this work with the most common methods of digital envelope identification.}
+%   \label{fig:Comparison}
+% \end{figure}
 
-For a recording of a guitar sound, in which a bend is performed, we contrast the proposed algorithm with four common approaches of varying complexity: simple smoothing of the rectified original wave, low pass filtering followed by rectification, peak identification and subsequent smoothing, and the extraction of the envelope via Hilbert transform, as presented earlier in the paper, with a previous smoothing of the original wave to avoid the artifacts shown in figure \ref{fig:AnalyticSignal}.
+% For a recording of a guitar sound, in which a bend is performed, we contrast the proposed algorithm with four common approaches of varying complexity: simple smoothing of the rectified original wave, low pass filtering followed by rectification, peak identification and subsequent smoothing, and the extraction of the envelope via Hilbert transform, as presented earlier in the paper, with a previous smoothing of the original wave to avoid the artifacts shown in figure \ref{fig:AnalyticSignal}.
 
-All those methods demanded careful choice of parameters: in the case of filtering, a cut-off frequency of approximately 88 Hz was used, consisting of one additional parameter to be tuned.
+% All those methods demanded careful choice of parameters: in the case of filtering, a cut-off frequency of approximately 88 Hz was used, consisting of one additional parameter to be tuned.
 For all the other traditional approaches, the Savitzky-Golay smoothing algorithm was used, and the window size was the parameter to be defined.
 
-In the case of the proposed algorithm, the smooth temporal envelope was obtained merging both positive and negative frontiers, the result being smoothed with the same procedure described.
+% In the case of the proposed algorithm, the smooth temporal envelope was obtained merging both positive and negative frontiers, the result being smoothed with the same procedure described.
 Figure \ref{fig:FrontiersToEnvelope} illustrates the results.
 
-\begin{figure}[ht!]
-  \centering
-    \includegraphics[width=0.8\linewidth]{13FrontiersToEnvelope.pdf}
-  \caption{Original positive and negative frontiers in dark grey and their mirrored versions. The temporal envelope, in black, is obtained by merging and approximating both frontiers.}
-  \label{fig:FrontiersToEnvelope}
-\end{figure}
+% \begin{figure}[ht!]
+%   \centering
+%     \includegraphics[width=0.8\linewidth]{13FrontiersToEnvelope.pdf}
+%   \caption{Original positive and negative frontiers in dark grey and their mirrored versions. The temporal envelope, in black, is obtained by merging and approximating both frontiers.}
+%   \label{fig:FrontiersToEnvelope}
+% \end{figure}
 
-Table \ref{table:Comparison} presents a computational comparison between the techniques.
+% Table \ref{table:Comparison} presents a computational comparison between the techniques.
 The times are for custom implementations of those methods in the Python programming language.
@@
 
-We can see that, while the algorithm here presented is the most accurate, it's also the most computationally intensive.
+% We can see that, while the algorithm here presented is the most accurate, it's also the most computationally intensive.
 The connection between alpha-shapes and Delaunay triangulation can be explored in future works as a means to improve the efficiency of the algorithm.
 
-\begin{table}[ht!]
-\centering
-\begin{tabular}{ l c c }
-\hline
-Method & LMS & time(s) \\
-\hline
-Presented Algorithm & 0.0088 & 4.1507 \\ 
-Smoothing           & 0.0123 & 0.1911 \\ 
-Low pass Filter      & 0.0258 & 0.5601 \\ 
-Hilbert Transform   & 0.0141 & 0.2599 \\ 
-\hline
-\end{tabular}
-\caption{Comparison of the algorithm present in this work with the most common methods of digital envelope identification.}
-\label{table:Comparison}
-\end{table}
+% \begin{table}[ht!]
+% \centering
+% \begin{tabular}{ l c c }
+% \hline
+% Method & LMS & time(s) \\
+% \hline
+% Presented Algorithm & 0.0088 & 4.1507 \\ 
+% Smoothing           & 0.0123 & 0.1911 \\ 
+% Low pass Filter      & 0.0258 & 0.5601 \\ 
+% Hilbert Transform   & 0.0141 & 0.2599 \\ 
+% \hline
+% \end{tabular}
+% \caption{Comparison of the algorithm present in this work with the most common methods of digital envelope identification.}
+% \label{table:Comparison}
+% \end{table}
 
-\subsection{Retrieving information about the underlying wave}
+% \subsection{Retrieving information about the underlying wave}
 
-The algorithm provides the position of the pseudo-cycles of a wave, with which one can, for example, derive its fundamental frequency and phase: using one of the two frontiers, we have $ f = (\#F^\pm - 1) n / (F^\pm[-1] - F^\pm[0])$, where $ \#F^\pm $ denotes the cardinality of the frontier, while $ F^\pm[0] $ and $ F^\pm[-1] $ stand for the frontier's first and last item, respectively.
+% The algorithm provides the position of the pseudo-cycles of a wave, with which one can, for example, derive its fundamental frequency and phase: using one of the two frontiers, we have $ f = (\#F^\pm - 1) n / (F^\pm[-1] - F^\pm[0])$, where $ \#F^\pm $ denotes the cardinality of the frontier, while $ F^\pm[0] $ and $ F^\pm[-1] $ stand for the frontier's first and last item, respectively.
 
-Similarly, each point of each of the frontiers can be used to estimate the local phase at that location.
+% Similarly, each point of each of the frontiers can be used to estimate the local phase at that location.
 For the positive frontier, it is necessary to note that, for each point, the following equation holds: $ \cos(\phi + 2 \pi f i / n) = 1 $, and thus $ \phi = 2 \pi k - 2 \pi f i / n $, with $ k \in \mathbb{N} $, while for the negative frontier $ \cos(\phi + 2 \pi f i / n) = -1 $ and $ \phi = 2 \pi k + \pi - 2 \pi f i / n) $.
 
-\section{Discussion}
+% \section{Discussion}
 
-This work hopes to fill a gap identified by the authors in the context of sound synthesis, where the lack of procedures for the accurate identification of the temporal envelope of arbitrary waves consists of an obstacle to the complete description and eventual manipulation of signals.
+% This work hopes to fill a gap identified by the authors in the context of sound synthesis, where the lack of procedures for the accurate identification of the temporal envelope of arbitrary waves consists of an obstacle to the complete description and eventual manipulation of signals.
 
-While relevant in its own accord, the procedure here presented isolates the individual pseudo-cycles of a wave, pinpointing them in the time domain.
+% While relevant in its own accord, the procedure here presented isolates the individual pseudo-cycles of a wave, pinpointing them in the time domain.
 Due to that, a rich set of theoretical advancements, of whose an example was given in the end of the preceding section, can be built upon this initial algorithm.
 
-With the knowledge of the exact position and characteristics of the pseudo-cycles of a wave, one can investigate its evolution in time, in terms of waveform, for example, free from the potential interference of a modulating wave; in other words, one can have a very satisfactory view of the instantaneous shape and phase of an arbitrary wave and its evolution in time, providing a way to bridge the gap between time domain and frequency domain interpretations.
+% With the knowledge of the exact position and characteristics of the pseudo-cycles of a wave, one can investigate its evolution in time, in terms of waveform, for example, free from the potential interference of a modulating wave; in other words, one can have a very satisfactory view of the instantaneous shape and phase of an arbitrary wave and its evolution in time, providing a way to bridge the gap between time domain and frequency domain interpretations.
 
-In the context of sound synthesis, for example, one would be able to apply specific methods for the recreation of the envelope, more in line with its smooth, slow varying and non-periodic nature and a different approach to the periodic and relatively fast changes of the temporal fine structure.
+% In the context of sound synthesis, for example, one would be able to apply specific methods for the recreation of the envelope, more in line with its smooth, slow varying and non-periodic nature and a different approach to the periodic and relatively fast changes of the temporal fine structure.
 
-\subsection{Conclusion}
+% \subsection{Conclusion}
 
-It is our opinion that such theoretical advances can be used in the formulation of better representations of discrete sounds, that in turn can improve sound compression techniques, while providing more adequate representation for the application of machine learning techniques for both sound classification and synthesis.
+% It is our opinion that such theoretical advances can be used in the formulation of better representations of discrete sounds, that in turn can improve sound compression techniques, while providing more adequate representation for the application of machine learning techniques for both sound classification and synthesis.
 
-\section{References}
+% \section{References}
 
-\printbibliography[heading=none]
+\printbibliography
 

### 2020-10-14 Papers/02_Sound_Representation/images/01signalenvelope/Frontier.py

@@ -1,0 +1,359 @@
+import plotly.graph_objects as go
+import numpy as np
+from scipy.interpolate import interp1d
+import numpy.polynomial.polynomial as poly
+import wave
+from scipy.signal import savgol_filter
+import os
+from pathlib import Path
+import sys
+import argparse
+
+def read_wav(path): 
+  """returns signal & fps"""
+  wav = wave.open(path , 'r')
+  signal = np.frombuffer(wav.readframes(-1) , np.int16)
+  fps = wav.getframerate()
+  return signal, fps
+
+
+def get_circle(x0, y0, x1, y1, r):
+  '''returns center of circle that passes through two points'''
+  
+  radsq = r * r
+  q = np.sqrt((x1 - x0)**2 + (y1 - y0)**2)
+
+  # assert q <= 2 * r, f"shit"
+
+  x3 = (x0 + x1) / 2
+  y3 = (y0 + y1) / 2
+
+  if y0 + y1 >= 0:
+    xc = x3 + np.sqrt(radsq - (q / 2)**2) * ((y0 - y1) / q)
+    yc = y3 + np.sqrt(radsq - (q / 2)**2) * ((x1 - x0) / q)
+  else:
+    xc = x3 - np.sqrt(radsq - (q / 2)**2) * ((y0 - y1) / q)
+    yc = y3 - np.sqrt(radsq - (q / 2)**2) * ((x1 - x0) / q)
+
+  return xc, yc
+
+
+def get_radius_function(X, Y):
+  m0 = np.average(Y[1:] - Y[:-1]) / np.average(X[1:] - X[:-1])
+  curvatures_X = []
+  curvatures_Y = []
+  for i in range(len(X) - 1):
+    m1 = (Y[i + 1] - Y[i]) / (X[i + 1] - X[i])
+
+    theta = np.arctan( (m1 - m0) / (1 + m1 * m0) )
+    k = np.sin(theta) / (X[i + 1] - X[i])
+
+    # if k >= 0:
+    curvatures_X.append((X[i + 1] + X[i]) / 2)
+    curvatures_Y.append(k)
+    
+  curvatures_Y = np.array(curvatures_Y)
+  coefs = poly.polyfit(curvatures_X, curvatures_Y, 0)
+  smooth_curvatures_Y = poly.polyval(curvatures_X, coefs)
+
+  r_of_x = interp1d(curvatures_X, 1 / np.abs(smooth_curvatures_Y), fill_value="extrapolate")
[…]

### 2020-10-14 Papers/02_Sound_Representation/images/03pcs.py

@@ -1,0 +1,175 @@
+from ctypes import resize
+import plotly.graph_objects as go
+import numpy as np
+import signal_envelope as se
+import sys
+import os
+os.chdir('../..')
+print (os.path.abspath(os.curdir))
+from plotly.subplots import make_subplots
+
+'''==============='''
+''' Read wav file '''
+'''==============='''
+
+name = "alto"
+W, fps = se.read_wav(f"./Python/Samples/{name}.wav")
+W = W - np.average(W)
+n = W.size
+X = np.arange(n)
+
+save_path = os.path.abspath(os.curdir) + '''\\Papers\\02_Sound_Representation\\images\\''' + sys.argv[0].split('/')[-1].replace(".py", "") + "-" + name + ".svg"
+print(save_path)
+
+# exit()
+
+Xpos, Xneg = se.get_frontiers(W)
+
+maxL = []
+for i in range(1, Xpos.size):
+  maxL.append(Xpos[i] - Xpos[i - 1])
+for i in range(1, Xneg.size):
+  maxL.append(Xneg[i] - Xneg[i - 1])
+maxL = np.max(np.array(maxL))
+print(f"Max L = {maxL}")
+
+pospseudoCyclesX = []
+pospseudoCyclesY = []
+for i in range(1, Xpos.size):
+  x0 = Xpos[i - 1]
+  x1 = Xpos[i]
+  # a = np.max(np.abs(W[x0 : x1]))
+  ft = np.fft.rfft(W[x0 : x1])
+  npulse = np.fft.irfft(ft, maxL)
+  pospseudoCyclesY.append(npulse / np.max(np.abs(npulse)))
+  pospseudoCyclesX.append(np.arange(maxL))
+pospseudoCyclesX = np.array(pospseudoCyclesX)
+pospseudoCyclesY = np.array(pospseudoCyclesY)
+
+negpseudoCyclesX = []
+negpseudoCyclesY = []
+for i in range(1, Xneg.size):
+  x0 = Xneg[i - 1]
+  x1 = Xneg[i]
+  # a = np.max(np.abs(W[x0 : x1]))
+  ft = np.fft.rfft(W[x0 : x1])
+  npulse = np.fft.irfft(ft, maxL)
+  negpseudoCyclesY.append(npulse / np.max(np.abs(npulse)))
+  negpseudoCyclesX.append(np.arange(maxL))
[…]

### 2020-10-14 Papers/02_Sound_Representation/images/04averagepc.py

@@ -1,0 +1,174 @@
+from ctypes import resize
+import plotly.graph_objects as go
+import numpy as np
+import signal_envelope as se
+import sys
+import os
+os.chdir('../..')
+print (os.path.abspath(os.curdir))
+
+'''==============='''
+''' Read wav file '''
+'''==============='''
+
+name = "alto"
+W, fps = se.read_wav(f"./Python/Samples/{name}.wav")
+W = W - np.average(W)
+n = W.size
+X = np.arange(n)
+
+save_path = os.path.abspath(os.curdir) + '''\\Papers\\02_Sound_Representation\\images\\''' + sys.argv[0].split('/')[-1].replace(".py", "") + "-" + name + ".svg"
+print(save_path)
+
+# exit()
+
+Xpos, Xneg = se.get_frontiers(W)
+
+maxL = []
+for i in range(1, Xpos.size):
+  maxL.append(Xpos[i] - Xpos[i - 1])
+for i in range(1, Xneg.size):
+  maxL.append(Xneg[i] - Xneg[i - 1])
+maxL = np.max(np.array(maxL))
+print(f"Max L = {maxL}")
+
+pospseudoCyclesX = []
+pospseudoCyclesY = []
+for i in range(1, Xpos.size):
+  x0 = Xpos[i - 1]
+  x1 = Xpos[i]
+  # a = np.max(np.abs(W[x0 : x1]))
+  ft = np.fft.rfft(W[x0 : x1])
+  npulse = np.fft.irfft(ft, maxL)
+  pospseudoCyclesY.append(npulse / np.max(np.abs(npulse)))
+  pospseudoCyclesX.append(np.arange(maxL))
+pospseudoCyclesX = np.array(pospseudoCyclesX)
+pospseudoCyclesY = np.array(pospseudoCyclesY)
+
+negpseudoCyclesX = []
+negpseudoCyclesY = []
+for i in range(1, Xneg.size):
+  x0 = Xneg[i - 1]
+  x1 = Xneg[i]
+  # a = np.max(np.abs(W[x0 : x1]))
+  ft = np.fft.rfft(W[x0 : x1])
+  npulse = np.fft.irfft(ft, maxL)
+  negpseudoCyclesY.append(npulse / np.max(np.abs(npulse)))
+  negpseudoCyclesX.append(np.arange(maxL))
+negpseudoCyclesX = np.array(negpseudoCyclesX)
+negpseudoCyclesY = np.array(negpseudoCyclesY)
+
+
+
+'''============================================================================'''
[…]

### 2020-10-21 Text/04_LMS.tex
added LMS chapter

@@
 \subsection{Least Mean Squares}
-The linear least mean squares, also known as linear regression \parencite{2018StephenBoydIntroduction} is a data fitting technique. % TODO
+
+\newacronym{LMS}{LMS}{Least Mean Squares}
+
+The linear \gls{LMS}, also known as linear regression \parencite{2018StephenBoydIntroduction} is a data fitting technique. % TODO
 
@@
 
-In the constrained version of the least squares problem we are still interested in finding the coefficient vector $ \hat{A} $ that minimizes $ ||Q A - Y||^2_2 $, but respecting a restriction on this vector of the general form of $ V A = B $.
+\newacronym{CLMS}{CLMS}{Constrained Least Mean Squares}
+
+In the constrained version of the \gls{LMS}, known as the \gls{CLMS} problem we are still interested in finding the coefficient vector $ \hat{A} $ that minimizes $ ||Q A - Y||^2_2 $, but respecting a restriction on this vector of the general form of $ V A = B $.
 

### 2020-10-21 Papers/02_Sound_Representation/SoundRepresentation.tex
some progress with the paper

@@
 \usepackage{amsmath}
+\usepackage[makeroom]{cancel}
 
@@
 
+We aim to exploit redundancies in the constitution of a discrete sound wave to formulate an alternative, more compact and thus easier to analyse, representation.
+Implicit in this affirmation is the somewhat loose assumption that the wave considered is reasonably well behaved, in the sense that redundancies exist to be identified.
+In contrast with \textcite{2019TarjanoNeuro}, we don't require the signal to be strictly harmonic.
+
 \section{Methodology}
@@
     %  \includegraphics[width=0.9\linewidth]{01signalenvelope}
-     \includegraphics{01signalenvelope}
+     \includegraphics{01signalenvelope.pdf}
     % \def\svgwidth{\linewidth}
@@
     %  \includegraphics[width=0.9\linewidth]{01signalenvelope}
-     \includegraphics{02pcsraw-alto}
+     \includegraphics{02pcsraw-alto.pdf}
     % \def\svgwidth{\linewidth}
@@
     %  \includegraphics[width=0.9\linewidth]{01signalenvelope}
-     \includegraphics{03pcs-alto}
+     \includegraphics{03pcs-alto.pdf}
     % \def\svgwidth{\linewidth}
     %\input{01signalenvelope.pdf_tex}
-  \caption{Normalized pseudo-cycles superimposed as defined by the positive and negative frontiers, for the alto singer signal, and the average of all normalized pseudo-cycles.}
+  \caption{Normalized pseudo-cycles superimposed as defined by the positive and negative frontiers, for an alto singer signal, and the average of all normalized pseudo-cycles.}
   \label{fig:pcalto}
@@
     %  \includegraphics[width=0.9\linewidth]{01signalenvelope}
-     \includegraphics{04averagepc-alto}
+     \includegraphics{04averagepc-alto.pdf}
     % \def\svgwidth{\linewidth}
     %\input{01signalenvelope.pdf_tex}
-  \caption{Average pseudo-cycles, superimposed, after a change of phase of $-\pi$ on the average pseudo-cycle defined by the negative frontier.}
+  \caption{Average pseudo-cycles, superimposed, after a change of phase of $-\pi$ on the average pseudo-cycle defined by the negative frontier, for an alto singer digital signal.}
   \label{fig:averagepc-alto}
@@
 
-Figure \ref{fig:averagepc-alto} makes clear that the average pseudo-cycles, as estimated by both frontiers, are in reasonable agreement, despite the evident vertical asymmetries of the underlying wave.
+Figure \ref{fig:averagepc-alto} makes clear that the average pseudo-cycles for the alto singer signal, as estimated by both frontiers, are in reasonable agreement, despite the evident vertical asymmetries of the underlying wave.
 
-This, however, is not always the case, as
+That is not always the case, however, as figure \ref{fig:averagepc-piano33} illustrates.
+For a discrete signal representing the recording o the key 33 of a grand piano, we can see that the average pseudo-cycle inferred from the positive frontier is substantially different from the one obtained with the negative frontier.
 
-% In this work it is assumed, without loss of generality, that one is interested in retrieving the mathematically ill-defined \parencite{2013Mengempirical} temporal envelope of an arbitrary physical sound wave, of which only a digital representation is available.
+The horizontal lines indicate the maximum of each pseudo cycle, being the point of adjacency between two subsequent pseudo cycles.
+For the average defined by the negative frontier we see that there is a local maxima near that middle that is almost as high as the global maxima, which is not the case as the average defined by the positive one.
 
-% Naturally, due to the necessary truncations both in time and amplitude, as well as measurement errors introduced by the instruments used to record the signal, it is not possible to exactly recover the original, continuous wave, from its digital counterpart.
+\begin{figure}[ht!]
+  \centering
+    %  \includegraphics[width=0.9\linewidth]{01signalenvelope}
+     \includegraphics{04averagepc-piano33.pdf}
+    % \def\svgwidth{\linewidth}
+    %\input{01signalenvelope.pdf_tex}
+  \caption{Average pseudo-cycles, superimposed, after a change of phase of $-\pi$ on the average pseudo-cycle defined by the negative frontier, for a grand piano digital signal.
+The horizontal lines in the figure mark the points of maximum of each of the average pseudo cycles.}
+  \label{fig:averagepc-piano33}
+\end{figure}
 
-% The best representation one can hope to get of the original wave is, then, the recording itself.
-It is sensible, thus, to characterize the envelope of the wave in terms of a subset of the exact samples that comprise its digital representation.
+We could take the average of those two representations, but that would dilute the parametric representation of the pseudo-cycle, besides introducing unnecessary complication in the stage of recreating the wave, as we would have to deal with redundant information about the location of each pseudo-cycle.
 
-% To that end, this work will emphasize the concept of frontier, as the points of the digital wave that mark the wave's upper and lower boundaries; from those points an envelope, conforming to the vague definitions of uniqueness and smoothness, can be trivially constructed, as will be shown.
+Instead, a measure of the statistical variance of the lengths of the pseudo cycles will be used in order to select the most appropriate frontier as guide for the construction of the average pseudo cycle waveform.
 
-% We start by defining a pulse as a series of consecutive samples of a discrete wave that have the same sign.
-More formally, let $ W[i] \in \mathbb{R} \ \forall \ i \in \mathbb{N}_0; \ i < n \in \mathbb{N}_0 $ be a real, discrete and finite signal indexed, without loss of generality, over an adjacent subset of the natural numbers starting at zero.
-This definition relates closely to the concept of an array or vector in most programming languages, and is used in the interest of simplicity.
+Figure \ref{fig:std-piano33} shows the distribution of the lengths of the pseudo cycles of the grand piano digital signal.
+We can use the standard deviation as this measure, and use the pseudo cycles as defined by the frontier with the least standard deviation.
 
-% A pulse $ P[j]; \ \forall \ j \in \mathbb{N}_0; \ j < q \in \mathbb{N}_0; q \le n $ is, then, a subset of $ W $ with cardinality $ \#P[j] = b - a $ such that $ \sign(W[a-1]) \ne \sign(W[a]) = \sign(W[a+1]) = \dots = \sign(W[b-1]) = \sign(W[b]) \ne \sign(W[b+1]) \ \forall \ i; \ a \le i \le b; \ a, b \in \mathbb{N}_0 $.
+\begin{figure}[ht!]
+  \centering
+    %  \includegraphics[width=0.9\linewidth]{01signalenvelope}
+     \includegraphics{05std-piano33.pdf}
+    % \def\svgwidth{\linewidth}
+    %\input{01signalenvelope.pdf_tex}
+  \caption{Histogram of the distribution of lengths of the pseudo cycles, as defined by the positive and negative frontiers, and the accompanying standard deviation, for a digital signal of the key 33 of a grand piano.}
+  \label{fig:std-piano33}
+\end{figure}
 
-% In other words, each pulse is a unique subset of $ W $ with its members being the maximum sequence of samples of $ W $ having the same signal.
-Thus $ P[0] \cup P[1] \cup \cdots \cup P[q-1] = W $ and $ P[i_1] \cap P[i_2] = \emptyset \ \forall \ i_1, i_2 \in \mathbb{N}_0;\ i_1 \ne i_2; \ i_1, i_2 < q $.
+Thus, a tillable parametric representation of the average pseudo-cycle can be obtained via the least mean squares method; we are interested in the polynomial that best approximates the average pseudo cycle, respecting the conditions that the beginning and the end of this polynomial must coincide, as must the first derivative of that polynomial in those places, if we are to obtain a smooth transition between pseudo cycles.
 
-% Naturally, each pulse can be classified as positive ($ P[j]^+ $) or negative ($ P[j]^- $), based on the sign of its samples, which must be the same.
-It is also convenient to attribute to each pulse a pair of coordinates: those will be the coordinates of the sample with the maximum absolute amplitude in each pulse, and will be denoted by $ P[j]^\pm_x $ and $ P[j]^\pm_y $.
-Additionally, the length, or cardinality, of a pulse will be denoted by $ \#P[j]^\pm $.
+From \textcite{2013SelesnickLeast} we know that a constrained least mean squares problem of the form $ \underset{A}{\min} ||Q A - Y||^2_2 \quad \text{subject to} \quad V A = B $ has a closed form approximate solution $ \hat{A} = \left(Q^{T} Q\right)^{-1} \left(Q^{T} Y - V^{T} \left(V \left(Q^{T} Q\right)^{-1} V^{T}\right)^{-1} V \left(Q^{T} Q\right)^{-1} Q^{T} Y - B \right) $, where $ Y $ is the vector we are interested in approximating.
 
-% For an arbitrary continuous function, then, a pulse would correspond to the regions between its roots; for a sinusoid, a pulse would be equal to half of its cycle.
-In analogy, two consecutive pulses, naturally with opposing signs, have the potential to delimitate a pseudo-cycle, as illustrated in figure \ref{fig:Pulses}.
-For that to be true, however, the pulses need to make part of the frontiers of the signal, as yet to be defined.
+We can proceed to define the real vector $ U = \{ u_0, u_1, \cdots, u_{m-1}\} $ as the average of all normalized pseudo-cycles, with $ m \in \mathbb{N} $ equal to the number of elements and the maximum length of the pseudo-cycles identified by the envelope detection algorithm.
 
-% The most mathematically sound definition of an envelope involves the representation of its underlying wave as an analytic signal, introduced by Gabor in 1946 \parencite{2007HahnHistory}.
-In his work, \textcite{1946GaborTheory} applies the then relatively new mathematical machinery of the quantum mechanics to unify the time and frequency representations of a wave, showing how the Hilbert transform could be applied to a real signal in order to obtain an equivalent complex signal, later known as the analytic signal.
+As we are interested in maintaining $ C^0 $ and $ C^1 $ continuity between two adjacent pseudo-pulses, it is convenient to define the vector $ Y $ as two vectors $ E $ stacked, that is $ Y = \{ u_0, u_1, \cdots, u_{m-1}, u_0, u_1, \cdots, u_{m-1}\} $.
 
-% This analytic signal has the form $ a(t) = s(t) + \mathbb{H}(s(t)) \ \text{i} $ \parencite{2016HePraat} where $ s(t) $ is the original real signal.
-$ \mathbb{H}(s(t)) $, the Hilbert transform of the original signal, becomes the imaginary part of the analytic signal; the envelope of a signal thus represented can be straightforwardly obtained by the computation of its complex modulus.
+$ A $, the vector of coefficients we are interested in estimating, will thus be composed of $ 2 (k + 1) $ items, where $ k $ is the order of the polynomial to be used in the approximation.
+Half those values are redundant, however, and $ A $ can thus be defined as $ A = \{ a_0, a_1, \cdots, a_k, a_0, a_1, \cdots, a_k \} $.
+$ Q $ is formed as usual, with the difference that
+it is duplicated horizontally, as seen in equation \ref{eq:matrix}.
 
-% Figure \ref{fig:AnalyticSignal} shows the envelope obtained via the analytic signal, for a pure sinusoid with a local frequency of 20 cycles, modulated by a polynomial of the third degree, illustrating the difference in the shape of the envelope in the presence and absence of gaussian white noise with standard deviation of $ 1/10 $ of the wave's maximum amplitude.
+Anticipating that we will eventually evaluate this polynomial in various lengths, it is convenient to normalize $ X $ between $ 0 $ and $ 1 $, defining $ X = \left\{ \cancelto{0}{\frac{0}{m-1}}, \frac{1}{m-1}, \frac{2}{m-1}, \cdots, \cancelto{1}{\frac{m-1}{m-1}} \right\} $
 
-% \begin{figure}[ht!]
-%   \centering
-%     \includegraphics[width=0.8\linewidth]{02AnalyticSignal.pdf}
-%   \caption{Envelope of a pure sinusoid modulated by a polynomial of degree 3, as obtained by the Hilbert Transform.
-In the first half, the sinusoid is free of noise, while in the second half white gaussian noise with a standard deviation of $ 1/10 $ of the maximum amplitude of the wave was added to the base signal.}
-%   \label{fig:AnalyticSignal}
-% \end{figure}
+\begin{equation} \label{eq:matrix}
+\stackrel{Q_{m \times 2 (k + 1)}} {
+\begin{bmatrix}
+  1      & x_0     & \cdots & x_0^k   & 1      & x_0     & \cdots & x_0^k   \\
+  1      & x_1     & \cdots & x_1^k   & 1      & x_1     & \cdots & x_1^k   \\
+  \vdots & \vdots  & \ddots & \vdots  & \vdots & \vdots  & \ddots & \vdots  \\
+  1      & x_{m-1} & \cdots & x_{m-1} & 1      & x_{m-1} & \cdots & x_{m-1} \\
+\end{bmatrix}
+}
+\stackrel{A_{2 (k + 1) \times 1}}
+{
+\begin{bmatrix}
+a_0    \\
+a_1    \\
+\vdots \\
+a_k    \\
+a_0    \\
+a_1    \\
+\vdots \\
+a_k    \\
+\end{bmatrix}
+}
+=
+\stackrel{Y_{2 m \times 1}}
+{
+\begin{bmatrix}
+y_0     \\
+y_1     \\
+\vdots  \\
+y_{m-1} \\
+y_0     \\
+y_1     \\
+\vdots  \\
+y_{m-1} \\
+\end{bmatrix}
+}
+\end{equation}
 
-% It is readily noticeable from figure \ref{fig:AnalyticSignal} that, as soon as noise is introduced in the original signal, it is reflected in the envelope, breaking our assumption of smoothness.
-Other than that, this definition yields a unique, positive envelope.
+To maintain the continuity and smoothness, as well as the identity between the first and second half of the coefficients, $ V $ can be defined as in equation \ref{eq:constraint}.
 
-% In practice, however, is not uncommon for a discrete wave, especially in the case of sound, to present somewhat different positive and negative contours; to account for this discrepancy, we will define a positive (upper) and a negative (lower) frontier of a discrete wave.
-They will be composed by the points $ (i, W[i]) $ that are, necessarily but not sufficiently, the points of respectively maximum and minimum amplitude of their pulses, i.e.
-$ (P[j]^\pm_x, P[j]^\pm_y) $.
+\begin{equation} \label{eq:constraint}
+\stackrel{V_{k + 3 \times 2 (k + 1)}} {
+\begin{bmatrix}
+  1      & 0       & \cdots & 0           & -1     & 0        & \cdots & 0                \\
+  0      & 1       & \cdots & 0           & 0      & -1       & \cdots & 0                \\
+  \vdots & \vdots  & \ddots & \vdots      & \vdots & \vdots   & \ddots & \vdots           \\
+  0      & 0       & \cdots & 1           & 0      & 0        & \cdots & -1               \\
+  1      & x_0     & \cdots & x_0^k       & -1     & -x_{m-1} & \cdots & -x_{m-1}^k       \\
+  0      & 1       & \cdots & k x_0^{k-1} & 0      & -1       & \cdots & -k x_{m-1}^{k-1} \\
+\end{bmatrix}
+}
+\stackrel{A_{2 (k + 1) \times 1}}
+{
+\begin{bmatrix}
+a_0    \\
+a_1    \\
+\vdots \\
+a_k    \\
+a_0    \\
+a_1    \\
+\vdots \\
+a_k    \\
+\end{bmatrix}
+}
+=
+\stackrel{B_{k + 2 \times 1}}
+{
+\begin{bmatrix}
+0      \\
+\vdots \\
+0      \\
+\end{bmatrix}
+}
+\end{equation}
 
-% The set of all these points such that $ W[i] > 0 $ will be called the positive, piecewise linear frontier and will be denoted by $ F^+ $.
-Similarly, $ F^- $ will denote the negative frontier of $ W $, defined by the subset of points $ (i, W[i]) $ where $ W[i] < 0 $, as is shown in figure \ref{fig:Frontiers}.
+The initial $k+1$ lines of $V$ assure that the first half of the coefficients of $A$ are equal to the second half, while the penultimate line ensures $C^0$ continuity and the last line ensures $C^1$ continuity, guaranteeing that the derivatives at the beginning and the end of the parametric pseudo cycles are equal.
 
-% \begin{figure}[ht!]
-%   \centering
-%     \includegraphics[width=0.8\linewidth]{03Frontiers.pdf}
-%   \caption{Positive and negative frontiers, envelope and pseudo-cycles for a fragment of piano sound.
-While we have only one (positive) envelope, here mirrored in the inferior space, we have two frontiers, the superior and inferior, that are obtained independently.}
-%   \label{fig:Frontiers}
-% \end{figure}
-
-% To identify such points, we need first to address the different units in which the amplitude and horizontal distance of our wave are expressed.
-To make abscissa and ordinate units consistent for the computation of the frontiers, and proceed with a geometric treatment of the subject, we consider the set of positive and negative pulses separately.
-
-% For each set, we multiply all pulses vertical coordinates by the average length of the pulses in the set, after having divided by the average of the pulses' amplitudes, putting both axes in the same unit $i$, related to time by the equation $i = t \ \mathit{fps}$.
-
-% Formally, supposing we have $ m $ pulses of the same sign, we scale each pulse's vertical coordinate $ P[j]^\pm_y $ by $ s^\pm $, where:
-
-% \begin{equation} \label{eq:scale}
-%   s^\pm = \frac{\sum_{j=1}^{m} \#P[j]^\pm}{\sum_{j=1}^{m} P[j]^\pm_y}
-% \end{equation}
-
-% Because we are interested in approximating a function that is smooth over the whole domain of the wave, we can assume that, in the scale of a pulse, this function will be very close to a horizontal straight line and, thus, won't touch any point inside the convex hull defined by the pulse.
-
-% By the same token, as the change in amplitude from one pulse to the next pulse with the same sign is motivated by changes in this same envelope, one can reasonably suppose that $ \max(P[j]) \approx \max(P[j+2]) $.
-
-% This makes clear that only one point of an arbitrary pulse can be in the frontier, namely, its point of maximum amplitude; those definitions will allow us to derive the frontiers of a wave using an alpha-shapes inspired algorithm without the need to compute the Delaunay triangulation first.
-
-% That also justifies the approximation of the convex hull of a pulse by the triangle defined by the points where the pulse crosses the horizontal axis and the point of maximum amplitude $ (P[j]^\pm_x, P[j]^\pm_y) $; this triangular pulse and, more specifically, its point of maximum, will become the atomic unit for the rest of the method.
-
-% Translating the intuitive explanation of alpha-shapes in \textcite{1994EdelsbrunnerThree} to our context, we can say that the points in the frontier are those touched by a circle of a given radius $ r $ outside the signal that is not allowed to contain any point of the signal; the pulses that contain said points can be seen as frontier pulses.
-
-% Intuitively, one can picture a circle being rolled above (or below, in the case of the negative frontier) the signal, and marking the points it touches as frontier points.
-
-% We can now revisit the concept of a pseudo-cycle, only hinted at before, defining it as the samples of the original wave that belong to two successive pulses that, in turn, belong to their respective frontiers, including eventual pulses between them, as can be seen in figure \ref{fig:Frontiers}.
-
-% It is necessary to infer the appropriate radius $ r $ of such a circle and, to that end, a measure of the curvature of a discrete function is needed.
-Discrete curvature estimation is an important task in image processing \parencite{2010FleischmannNovel} for which no default definition exists.
-
-% The two possible general approaches are the derivation of direct methods that use characteristics of the discrete wave to calculate the curvature, or the use of the smooth curvature of a curve fitted to the discrete wave \parencite{2001CoeurjollyDiscrete}.
-
-% Concerning the last method, two approaches to fit a polynomial to a generic wave are readily available: the least mean squares approach, that seeks to minimize the dependent variable errors and the total least squares problem, that treats both variables symmetrically \parencite{1980GolubAnalysis}.
-
-% The geometric, symmetrical nature of the problem excludes the more computationally economic least mean squares approach, however, leaving us with the total least squares, for which no general closed-form solutions are available \parencite{2007MarkovskyOverview}; this leads us to turn our attention to direct approaches of curvature estimation.
-
-% \subsection{Discrete Curvature Estimation - The Equivalent Circle Approach}
-
-% As previously noted, no default discrete curvature definition exists; \textcite{2014CarrollSurvey}, for example, derive three such different definitions based on the approximation of a circle by an inscribed, centred and circumscribed polygon.
-In the context of 3d meshes, \textcite{2016VasaMultivariate} evaluate a range of existing estimators from a multivariate point of view.
-
-% None of those approaches, especially for being defined over the vertices of a discrete wave, fit seamlessly into our algorithm; we thus proceed to define a discrete curvature measure over the edges of a discrete wave.
-To that end we are going to apply the definition of smooth curvature as the rate of change of the unit tangent to a curve, noting that this is equivalent to that of the osculating circle \parencite{2016VasaMesh}.
-
-% The idea is to find the equivalent circle whose tangent presents the same change in direction, in the same horizontal distance, as the edge of interest, as shown in figure \ref{fig:DiscreteCurvature}.
-
-% \begin{figure}[ht!]
-%   \centering
-%     \includegraphics[width=0.2\linewidth]{04DiscreteCurvature.pdf}
-%   \caption{The tangent unit vector of the circle changes from the horizontal direction in $ \hat{u}_0 $ to an inclination of $ \theta $ in $ \hat{u}_1 $.
-$ \theta $ is also the angle between the average direction $ \vec{\overline{v}}_l $ and the vector $ \vec{v}_l $ that represents the edge of interest, shown in grey.}
-%   \label{fig:DiscreteCurvature}
-% \end{figure}
-
-% For each frontier, the angle $ \theta $ between the average direction and each of the vectors can be obtained via the relation between the slopes $ d $ of the respective lines, as in the expression $ \theta = \arctan( (d - \overline{d}) / (1 + d \overline{d}) ) $, with $\overline{d} = \sum_{l=1}^{m-1} (P[l+1]_y - P[l]_y) / (P[l+1]_x - P[l]_x) $ and $ d_l = (P[l+1]_y - P[l]_y)/ (P[l+1]_x - P[l]_x) $
-
-% The radii of the equivalent circles are thus given by $ r_l = x_l / \sin (\theta_l) $ with the curvature given by $\kappa_l = 1/r_l=\sin (\theta_l) / x_l$, where $ x_l = P[l+1]_x - P[l]_x $ and $ y_l = P[l+1]_y - P[l]_y $.
-Joining both equations, one can write:
-
-% \begin{equation} \label{eq:radius}
-%   r_l = \left\lvert \frac{x_l \sqrt{\overline{d}^2 + 1} \ \sqrt{x_l^2 + y_l^2}}{\overline{d} x_l - y_l} \right\rvert
-% \end{equation}
-
-% After this procedure, illustrated in figure \ref{fig:Curvatures}, two sets of points in $ \mathbb{R}^2 $, corresponding to the positive and negative frontiers, are obtained, where the ordinates represent the instantaneous radius, with its localization in the time domain given by the abscissas.
-
-% \begin{figure}[ht!]
-%   \centering
-%     \includegraphics[width=0.8\linewidth]{05Curvatures.pdf}
-%   \caption{The equivalent circles obtained via the procedure described for the positive frontier.
-$\theta$, the angle between $ \vec{\overline{v}}^+$ and each of the $\vec{v}_l^+$ is used, as is the horizontal component of each $\vec{v}_l^+$ to obtain the radii of the circles.}
-%   \label{fig:Curvatures}
-% \end{figure}
-
-% Given the robustness of the method, for an average wave of a few seconds at an ordinary rate of 44100 $ \mathit{fps} $ the average of the radius values is sufficient to assure the accurate identification of the frontier.
-However, a curve fitting method can be used in the case of longer or ill-behaved waves.
-% An example of the use of the average equivalent circle can be seen in figure \ref{fig:AvgCircle}.
-
-% \begin{figure}[ht!]
-%   \centering
-%     \includegraphics[width=0.8\linewidth]{06AvgCircle.pdf}
-%   \caption{The positive frontier as obtained by the use of the circle representing the average curvature of the discrete wave illustrated.}
-%   \label{fig:AvgCircle}
-% \end{figure}
-
-% \section{Results}
-
-% Figure \ref{fig:FullFrontiers} illustrates the frontiers of six diverse discrete sound waves, as well as an in-detail view of the highlighted segment for each signal.
-All waves are records of physical sounds, chosen to represent the applicability of the algorithm in real-world scenarios.
-
-% \begin{figure}[ht!]
-%   \centering
-%     \includegraphics[width=0.8\linewidth]{07FullFrontiers.pdf}
-%   \caption{Positive and negative frontiers of six digital waves, as extracted by the algorithm here presented.
-For each wave, the region highlighted in black is shown in detail besides the whole wave.
-For each wave the horizontal axis is the sample number $i$, while the vertical axis is the normalized amplitude.
-All 6 waves were recorded at 44100 $ \mathit{fps} $.}
-%   \label{fig:FullFrontiers}
-% \end{figure}
-
-% We can see that the frontier is satisfactorily detected in the case of harmonic and inharmonic sounds, and is robust in relation to the number of samples and the frequencies of the waves.
-
-% The impact of the normalization of a wave using the frontiers obtained can be seen in figure \ref{fig:Fourier}, which shows the original and normalized wave in the case of a recording of the key 33 of a piano both in the time and frequency domains.
-We can see that the Fourier power spectra is greatly simplified in the case of the normalized wave.
-
-% \begin{figure}[ht!]
-%   \centering
-%     \includegraphics[width=0.8\linewidth]{08Fourier.pdf}
-%   \caption{Original and normalized views of a recording of key 33 (F3, 175 Hz) from a grand piano both in time and frequency domains.}
-%   \label{fig:Fourier}
-% \end{figure}
-
-% \subsection{Robustness}
-
-% Figures \ref{fig:xRobustness} and \ref{fig:yRobustness} illustrate the computational resilience of the algorithm for scaling in both the abscissa and ordinate axes, respectively.
-This is due to the normalization scheme presented in equation \ref{eq:scale}.
-
-% Although important from a theoretical standpoint, this step has an especial computational relevance in the case of digital sound waves, as they tend to be composed of a high number of samples with relatively low amplitudes, generally normalized between -1 and 1, leading to a numerical disparity between axes.
-
-% \begin{figure}[ht!]
-%   \centering
-%     \includegraphics[width=0.8\linewidth]{10xRobustness.pdf}
-%   \caption{Minimum and maximum scales of the abscissa before computational stability issues arose, for a recording of a drum kit tom.  $ \text{B} = 10^6 $.}
-%   \label{fig:xRobustness}
-% \end{figure}
-
-% \begin{figure}[ht!]
-%   \centering
-%     \includegraphics[width=0.8\linewidth]{09yRobustness.pdf}
-%   \caption{Minimum and maximum values of the amplitude before computational stability issues arose, for a recording of a drum kit tom.}
-%   \label{fig:yRobustness}
-% \end{figure}
-
-% \subsection{Comparison with traditional algorithms}
-
-% Direct comparison with many of the more recent algorithms is made difficult by the unavailability of digital implementations of such works, many designed to process analogue signals \parencite[e.g.,][]{2018AssefModeling}.
-
-% Nevertheless, insight can be gained from a comparison of the results of the method here proposed with some of the most common envelope extraction algorithms, as can be seen in figure \ref{fig:Comparison}.
-
-% \begin{figure}[ht!]
-%   \centering
-%     \includegraphics[width=0.8\linewidth]{12Comparison.pdf}
-%   \caption{Comparison of the algorithm present in this work with the most common methods of digital envelope identification.}
-%   \label{fig:Comparison}
-% \end{figure}
-
-% For a recording of a guitar sound, in which a bend is performed, we contrast the proposed algorithm with four common approaches of varying complexity: simple smoothing of the rectified original wave, low pass filtering followed by rectification, peak identification and subsequent smoothing, and the extraction of the envelope via Hilbert transform, as presented earlier in the paper, with a previous smoothing of the original wave to avoid the artifacts shown in figure \ref{fig:AnalyticSignal}.
-
-% All those methods demanded careful choice of parameters: in the case of filtering, a cut-off frequency of approximately 88 Hz was used, consisting of one additional parameter to be tuned.
-For all the other traditional approaches, the Savitzky-Golay smoothing algorithm was used, and the window size was the parameter to be defined.
-
-% In the case of the proposed algorithm, the smooth temporal envelope was obtained merging both positive and negative frontiers, the result being smoothed with the same procedure described.
-Figure \ref{fig:FrontiersToEnvelope} illustrates the results.
-
-% \begin{figure}[ht!]
-%   \centering
-%     \includegraphics[width=0.8\linewidth]{13FrontiersToEnvelope.pdf}
-%   \caption{Original positive and negative frontiers in dark grey and their mirrored versions. The temporal envelope, in black, is obtained by merging and approximating both frontiers.}
-%   \label{fig:FrontiersToEnvelope}
-% \end{figure}
-
-% Table \ref{table:Comparison} presents a computational comparison between the techniques.
-The times are for custom implementations of those methods in the Python programming language.
-The least mean squares result was calculated by halving the envelopes presented in figure \ref{fig:Comparison} and comparing them with the rectified absolute wave, so as to avoid the potential influence of an estimated frequency.
-
-% We can see that, while the algorithm here presented is the most accurate, it's also the most computationally intensive.
-The connection between alpha-shapes and Delaunay triangulation can be explored in future works as a means to improve the efficiency of the algorithm.
+\begin{figure}[ht!]
+  \centering
+    %  \includegraphics[width=0.9\linewidth]{01signalenvelope}
+     \includegraphics{06approximation-alto}
+    % \def\svgwidth{\linewidth}
+    %\input{01signalenvelope.pdf_tex}
+  \caption{Two adjacent pseudo-cycles. The detail view illustrates that $C^0$ and $C^1$ smoothness are maintained in the parametric representation.}
+  \label{fig:05approximation-alto}
+\end{figure}
 
@@
 
-% \subsection{Retrieving information about the underlying wave}
-
-% The algorithm provides the position of the pseudo-cycles of a wave, with which one can, for example, derive its fundamental frequency and phase: using one of the two frontiers, we have $ f = (\#F^\pm - 1) n / (F^\pm[-1] - F^\pm[0])$, where $ \#F^\pm $ denotes the cardinality of the frontier, while $ F^\pm[0] $ and $ F^\pm[-1] $ stand for the frontier's first and last item, respectively.
-
-% Similarly, each point of each of the frontiers can be used to estimate the local phase at that location.
-For the positive frontier, it is necessary to note that, for each point, the following equation holds: $ \cos(\phi + 2 \pi f i / n) = 1 $, and thus $ \phi = 2 \pi k - 2 \pi f i / n $, with $ k \in \mathbb{N} $, while for the negative frontier $ \cos(\phi + 2 \pi f i / n) = -1 $ and $ \phi = 2 \pi k + \pi - 2 \pi f i / n) $.
-
-% \section{Discussion}
-
-% This work hopes to fill a gap identified by the authors in the context of sound synthesis, where the lack of procedures for the accurate identification of the temporal envelope of arbitrary waves consists of an obstacle to the complete description and eventual manipulation of signals.
-
-% While relevant in its own accord, the procedure here presented isolates the individual pseudo-cycles of a wave, pinpointing them in the time domain.
-Due to that, a rich set of theoretical advancements, of whose an example was given in the end of the preceding section, can be built upon this initial algorithm.
-
-% With the knowledge of the exact position and characteristics of the pseudo-cycles of a wave, one can investigate its evolution in time, in terms of waveform, for example, free from the potential interference of a modulating wave; in other words, one can have a very satisfactory view of the instantaneous shape and phase of an arbitrary wave and its evolution in time, providing a way to bridge the gap between time domain and frequency domain interpretations.
-
-% In the context of sound synthesis, for example, one would be able to apply specific methods for the recreation of the envelope, more in line with its smooth, slow varying and non-periodic nature and a different approach to the periodic and relatively fast changes of the temporal fine structure.
-
-% \subsection{Conclusion}
-
-% It is our opinion that such theoretical advances can be used in the formulation of better representations of discrete sounds, that in turn can improve sound compression techniques, while providing more adequate representation for the application of machine learning techniques for both sound classification and synthesis.
-
-% \section{References}
-
 \printbibliography

### 2020-10-21 Papers/02_Sound_Representation/images/06approximation.py

@@ -1,0 +1,238 @@
+import plotly.graph_objects as go
+import numpy as np
+import signal_envelope as se
+from scipy.linalg import block_diag
+import numpy.polynomial.polynomial as poly
+import os, sys
+os.chdir('../..')
+print (os.path.abspath(os.curdir))
+
+'''==============='''
+''' Read wav file '''
+'''==============='''
+
+name = "alto"
+W, fps = se.read_wav(f"./Python/Samples/{name}.wav")
+W = W - np.average(W)
+n = W.size
+X = np.arange(n)
+
+save_path = os.path.abspath(os.curdir) + '''\\Papers\\02_Sound_Representation\\images\\''' + sys.argv[0].split('/')[-1].replace(".py", "") + "-" + name + ".svg"
+print(save_path)
+
+Xpos, Xneg = se.get_frontiers(W)
+
+posL = []
+for i in range(1, Xpos.size):
+  posL.append(Xpos[i] - Xpos[i - 1])
+
+negL = []
+for i in range(1, Xneg.size):
+  negL.append(Xneg[i] - Xneg[i - 1])
+
+if np.std(posL) < np.std(negL):
+  print("using positive frontier")
+  maxL = np.max(np.array(posL))
+  pseudoCyclesX = []
+  pseudoCyclesY = []
+  for i in range(1, Xpos.size):
+    x0 = Xpos[i - 1]
+    x1 = Xpos[i]
+    # a = np.max(np.abs(W[x0 : x1]))
+    ft = np.fft.rfft(W[x0 : x1])
+    npulse = np.fft.irfft(ft, maxL)
+    pseudoCyclesY.append(npulse / np.max(np.abs(npulse)))
+    pseudoCyclesX.append(np.arange(maxL))
+  pseudoCyclesX = np.array(pseudoCyclesX)
+  pseudoCyclesY = np.array(pseudoCyclesY)
+else:
+  print("using negative frontier")
+  maxL = np.max(np.array(negL))
+  pseudoCyclesX = []
+  pseudoCyclesY = []
+  for i in range(1, Xpos.size):
+    x0 = Xpos[i - 1]
+    x1 = Xpos[i]
+    # a = np.max(np.abs(W[x0 : x1]))
+    ft = np.fft.rfft(W[x0 : x1])
+    npulse = np.fft.irfft(ft, maxL)
+    pseudoCyclesY.append(npulse / np.max(np.abs(npulse)))
[…]

### 2020-10-25 Papers/02_Sound_Representation/SoundRepresentation.tex
progress with the paper
added some brief intro

@@
 In contrast with \textcite{2019TarjanoNeuro}, we don't require the signal to be strictly harmonic.
+
+We aim in providing a method that, with minimal manual interference, is able to reduce a digital wave to a more compact and meaninful representation.
 
@@
     %\input{01signalenvelope.pdf_tex}
-  \caption{Two adjacent pseudo-cycles. The detail view illustrates that $C^0$ and $C^1$ smoothness are maintained in the parametric representation.}
+  \caption{Two adjacent pseudo-cycles. The detail view illustrates that $C^0$ and $C^1$ smoothness are maintained in the parametric representation when two reconstructions are stacked horizontally.}
   \label{fig:05approximation-alto}

### 2020-11-01 Text/04_Theory.tex
Reorganized, renamed files and extended text a bit

--- 2020-10-21 Text/04_LMS.tex
@@
 
+\subsection{Waves}
+
+It can be insightful to analyse the behaviour of the crytical points and the zeroes of an arbitrary, periodic wave over time.
+We will assume, for the sake of simplicity, that each cycle of said wave has a unique maximum and minimum, and crosses the horizontal axis one single time.
+Those assumption dont compromise the generality of the analysis, as in the case of multiple maxima, minima and/or zeroes, one of thosee can be choosen randomly to represent the global one, as long as consistency between cycles is maintained.
+
+To keep our roots in the physical case, we`ll also assume that a wave with a zero phase has an instantaneous amplitude of zero at time zero, as is the case of the sine wave, for example.
+
+For a continuous wave, if we were to plot the time at which any of those values ocurred as a function of their ordinality, the result would be parallel straight lines, their slope somehow related to the frequency of the underlying wave.
+We will illustrated the point in the case of discrete waves, as this case will be of more immediate use in this work.
+
+\subsubsection{Continuous and discrete waves}
+
+First, it is necessary to stablish a correspondence between a continuous and a discrete wave, as the latter will be the object of most of this work.
+To do that we will define a simple continuous sinusoidal wave of the form $ S(t) = a \cos \left( \phi + 2 \pi \mathcal{F} t \right) $, where $ a \in \mathbb{R}^+ $ is the amplitude; $ \phi \in \mathbb{R}^+, 0 \le \phi < 2 \pi $ is the phase; $ \mathcal{F} \in \mathbb{R}^+ $ is the absolute frequency of the wave, in Hz and $ t \in \mathbb{R} $ is time, in seconds.
+
+The cosine function is used for a twofold reason: it is historical background, and its use helps to underline the periodic nature of the wave in consideration.
+It is important to keep in mind, however, that this discussion applyes any arbitrary function of the form $ f(x) = f(x + P) \forall x \in D(f) $, where $ D $ is the domain of the function $ f $.
+% TODO This will be further illustrated in the discrete part below.
+
+If we were to sample $ S(t) $ at regular intervals, that is, to measure the value of the function at a constant rate of $ h $ \gls{FPS}, neglecting errors of measurement, we would have a discrete representation of the continuous wave of the form $ W[i] = a \cos \left( \phi + 2 \pi f \frac{i}{n} \right) $ where, alongside the quantities introduced before, we would have $ i \in \mathbb{N} $ as the frame number; $ n \in \mathbb{N} $ as the number of discrete samples and $ f \in \mathbb{R}^+ $ being the relative frequency of our discrete wave in cycles per $ n $ samples.
+
+The relative, local frequency $ f $ of the discrete representation is related to the absolute frequency $ \mathcal{F} $ via the equation $ f = \frac{n}{\text{fps}} \mathcal{F} $ while the absolute time, in the continuous case can be obtained from the frame number $ i $ of the discrete signal via $ i = t \ \text{fps} $.
+
+\subsubsection{Points of interest of periodic waves}
+
 \subsection{Least Mean Squares}
-
-\newacronym{LMS}{LMS}{Least Mean Squares}
 

### 2020-11-01 Text/06_SignalEnvelope.tex
Reorganized, renamed files and extended text a bit

@@
     \centering
-      \includegraphics[width=0.8\linewidth]{./imgs/envelope/01Pulses.pdf}
+      \includegraphics{./imgs/01Pulses.pdf}
     \caption{Example of a discrete wave.
@@
   \centering
-    \includegraphics[width=0.8\linewidth]{./imgs/envelope/02AnalyticSignal.pdf}
+    \includegraphics{./imgs/02AnalyticSignal.pdf}
   \caption{Envelope of a pure sinusoid modulated by a polynomial of degree 3, as obtained by the Hilbert Transform.
@@
   \centering
-    \includegraphics[width=0.8\linewidth]{./imgs/envelope/03Frontiers.pdf}
+    \includegraphics{./imgs/03Frontiers.pdf}
   \caption{Positive and negative frontiers, envelope and pseudo-cycles for a fragment of piano sound.
@@
 
-For each set, we multiply all pulses vertical coordinates by the average length of the pulses in the set, after having divided by the average of the pulses' amplitudes, putting both axes in the same unit $i$, related to time by the equation $i = t \ \mathit{fps}$.
+For each set, we multiply all pulses vertical coordinates by the average length of the pulses in the set, after having divided by the average of the pulses' amplitudes, putting both axes in the same unit $i$, related to time by the equation $i = t \ \text{FPS}$.
 
@@
 
-\subsection{Discrete Curvature Estimation - The Equivalent Circle Approach}
+\subsection{Discrete Curvature Estimation - Equivalent Circle Approach}
 
@@
   \centering
-    \includegraphics[width=0.2\linewidth]{./imgs/envelope/04DiscreteCurvature.pdf}
+    \includegraphics{./imgs/04DiscreteCurvature.pdf}
   \caption{The tangent unit vector of the circle changes from the horizontal direction in $ \hat{u}_0 $ to an inclination of $ \theta $ in $ \hat{u}_1 $.
@@
   \centering
-    \includegraphics[width=0.8\linewidth]{./imgs/envelope/05Curvatures.pdf}
+    \includegraphics{./imgs/05Curvatures.pdf}
   \caption{The equivalent circles obtained via the procedure described for the positive frontier.
@@
 
-Given the robustness of the method, for an average wave of a few seconds at an ordinary rate of 44100 $ \mathit{fps} $ the average of the radius values is sufficient to assure the accurate identification of the frontier.
+Given the robustness of the method, for an average wave of a few seconds at an ordinary rate of 44100 \gls{FPS} the average of the radius values is sufficient to assure the accurate identification of the frontier.
 However, a curve fitting method can be used in the case of longer or ill-behaved waves.
@@
   \centering
-    \includegraphics[width=0.8\linewidth]{./imgs/envelope/06AvgCircle.pdf}
+    \includegraphics{./imgs/06AvgCircle.pdf}
   \caption{The positive frontier as obtained by the use of the circle representing the average curvature of the discrete wave illustrated.}
@@
   \centering
-    \includegraphics[width=0.8\linewidth]{./imgs/envelope/07FullFrontiers.pdf}
+    \includegraphics{./imgs/07FullFrontiers.pdf}
   \caption{Positive and negative frontiers of six digital waves, as extracted by the algorithm here presented.
@@
 For each wave the horizontal axis is the sample number $i$, while the vertical axis is the normalized amplitude.
-All 6 waves were recorded at 44100 $ \mathit{fps} $.}
+All 6 waves were recorded at 44100 \gls{FPS}.}
   \label{fig:FullFrontiers}
@@
   \centering
-    \includegraphics[width=0.8\linewidth]{./imgs/envelope/08Fourier.pdf}
+    \includegraphics{./imgs/08Fourier.pdf}
   \caption{Original and normalized views of a recording of key 33 (F3, 175 Hz) from a grand piano both in time and frequency domains.}
@@
   \centering
-    \includegraphics[width=0.8\linewidth]{./imgs/envelope/10xRobustness.pdf}
+    \includegraphics{./imgs/10xRobustness.pdf}
   \caption{Minimum and maximum scales of the abscissa before computational stability issues arose, for a recording of a drum kit tom.  $ \text{B} = 10^6 $.}
@@
   \centering
-    \includegraphics[width=0.8\linewidth]{./imgs/envelope/09yRobustness.pdf}
+    \includegraphics{./imgs/09yRobustness.pdf}
   \caption{Minimum and maximum values of the amplitude before computational stability issues arose, for a recording of a drum kit tom.}
@@
   \centering
-    \includegraphics[width=0.8\linewidth]{./imgs/envelope/12Comparison.pdf}
+    \includegraphics{./imgs/12Comparison.pdf}
   \caption{Comparison of the algorithm present in this work with the most common methods of digital envelope identification.}
@@
   \centering
-    \includegraphics[width=0.8\linewidth]{./imgs/envelope/13FrontiersToEnvelope.pdf}
+    \includegraphics{./imgs/13FrontiersToEnvelope.pdf}
   \caption{Original positive and negative frontiers in dark grey and their mirrored versions. The temporal envelope, in black, is obtained by merging and approximating both frontiers.}

### 2020-11-04 Text/04_Theory.tex
some progress

@@
 
-To keep our roots in the physical case, we`ll also assume that a wave with a zero phase has an instantaneous amplitude of zero at time zero, as is the case of the sine wave, for example.
+We are going to also assume that a wave with a zero phase has an maximum instantaneous amplitude at time zero, as is the case of the cosine wave, for example.
+Altough that assumption somewhat departs from the physical intuition, it is useful in making our description consistent with the theory behind the Fourier Transform.
 
@@
 
-First, it is necessary to stablish a correspondence between a continuous and a discrete wave, as the latter will be the object of most of this work.
+First, it is necessary to establish a correspondence between a continuous and a discrete wave, as the latter will be the object of most of this work.
 To do that we will define a simple continuous sinusoidal wave of the form $ S(t) = a \cos \left( \phi + 2 \pi \mathcal{F} t \right) $, where $ a \in \mathbb{R}^+ $ is the amplitude; $ \phi \in \mathbb{R}^+, 0 \le \phi < 2 \pi $ is the phase; $ \mathcal{F} \in \mathbb{R}^+ $ is the absolute frequency of the wave, in Hz and $ t \in \mathbb{R} $ is time, in seconds.
 
-The cosine function is used for a twofold reason: it is historical background, and its use helps to underline the periodic nature of the wave in consideration.
-It is important to keep in mind, however, that this discussion applyes any arbitrary function of the form $ f(x) = f(x + P) \forall x \in D(f) $, where $ D $ is the domain of the function $ f $.
+The cosine function is used for a twofold reason: its use helps to underline the periodic nature of the wave in consideration and is the parameter of the phase when the FT is used.
+It is important to keep in mind, however, that this discussion applies any arbitrary function of the form $ f(x) = f(x + P) \forall x \in D(f) $, where $ D $ is the domain of the function $ f $.
 % TODO This will be further illustrated in the discrete part below.

### 2020-11-05 Papers/02_Sound_Representation/SoundRepresentation.tex
reorganized images

@@
 
-We aim in providing a method that, with minimal manual interference, is able to reduce a digital wave to a more compact and meaninful representation.
+We aim in providing a method that, with minimal manual interference, is able to reduce a digital wave to a more compact and meaningful representation.
 
 \section{Methodology}
+Conceptually, one can think of the procedure here presented as a way to represent discrete, pitched, signal with a set of 3 tuples.
+Those tuples are composed, to describe them loosely, of a polynomial description of the envelope of the signal, that dictates the outer shape of the wave, or the volume of the sound as a function of time; another polynomial description, but this time for the (pseudo)-cycles of the wave, meant to represent the shape of each pseudo cycle of the signal, and a description of the beginning of each said pseudo cycle, and thus the end of the preceding one.
 
-Using the algorithm presented in \textcite{2020TarjanoRobust} one can identify the frontiers, both positive and negative, of a digital signal: Those are, ideally, respectively composed of the maxima and minima of each pseudo-cycle of a discrete signal.
+This aids in putting in perspective the broad strokes of the algorithm.
+
+\subsection{Pseudo Cycles}
+
+This work proposes the pseudo cycle as the building block of a discrete wave, assuming its parametric shape to be constant throughout the space of the digital signal; in order words, the adjective pseudo is used because the two other descriptions of the wave, namely, the envelope and the length of the pseudo cycles act into the pseudo cycles shape itself, modifying it in function of the time.
+
+This is a simplification, as many factors are known to influence the shape pseudo cycles of an arbitrary wave throughout time, as is the case of the decaying high frequencies in a rigid string, for example.
+% TODO Those factors will be eventually accounted for
+
+Using the envelope detection algorithm presented in \textcite{2020TarjanoRobust} one can identify the frontiers, both positive and negative, of a digital signal: Those are, ideally, respectively composed of the maxima and minima of each pseudo-cycle of a discrete signal.
 
@@
 Consider, for example, the digital signal presented in figure \ref{fig:signalenvelope}
-and the extracted envelopes; the signal is an example of the voice of an alto singer, with its frontiers in black.
+and the extracted frontiers; the signal is an example of the voice of an alto singer, with its frontiers in black.
 
@@
   \centering
-    %  \includegraphics[width=0.9\linewidth]{01signalenvelope}
-     \includegraphics{01signalenvelope.pdf}
-    % \def\svgwidth{\linewidth}
-    %\input{01signalenvelope.pdf_tex}
+     \includegraphics{images/01signalenvelope.pdf}
   \caption{A discrete wave of the singing voice of an alto singer.
@@
 
-It can be seen that both frontiers provide approximately the same information about the location of each pseudo-cycle, with a $ \pi $ phase difference.
+It can be seen that both frontiers provide approximately the same information about the location of each pseudo-cycle, albeit with a phase difference between them.
+As the original algorithm proposed in \textcite{2020TarjanoRobust} is meant for envelope detection, maxima or minima that lies close to the already established frontiers are sometimes neglected, without significant impact in the envelope extracted.
+
+This can be seen in the discrete representation of the sound of the key 33 of a grand piano in figure \ref{fig:frontiergaps}, as the black squares and diamonds: intuitively, one can see that some local maxima and minima are missing, compromising the identification of all pseudo cycles.
+For our purposes, however, it is beneficial to accurately identify as many local maxima and minima, and hence pseudo cycles, as possible, as the impact on the shape of a pseudo cycle is significant.
+
+We are going to assume that all the local extremes identified by the envelope detection algorithm are indeed correct; albeit not being strictily accurate, this approximation is very reasonable and simplifies further steps immensely, as we now need only to idenfity pseudo cycles with periods $ T $, that is, the difference between two adjacent maxima or minima, that are significantly bigger than the average observed in the discrete wave, and identify their local maximum or minimum, respectively.
+
+The result of this procedure can be seen ad the gray squares and diamonds in figure \ref{fig:frontiergaps}.
 
@@
   \centering
-    %  \includegraphics[width=0.9\linewidth]{01signalenvelope}
+     \includegraphics{images/02frontiergaps.pdf}
+  \caption{Close-up view of a discrete signal representing the sound of the key 33 of a grand piano and its original positive and negative frontiers, in black, as well as the additional extrema identified during the refining process.}
+  \label{fig:frontiergaps}
+\end{figure}
+
+% Consider, for example, the periodic discrete signal shown in the figure \ref{fig:periodicwave} below, consisting of the repetition of a cycle constructed from a simple sinusoid with gaussian noise added, where the maxima and the minima were highlighted.
+
+% \begin{figure}[ht!]
+%   \centering
+%      \includegraphics{07periodicwave.pdf}
+%   \caption{Three Cycles, divided by the vertical lines, of a discrete periodic wave, with the maxima and minima highlighted}
+%   \label{fig:periodicwave}
+% \end{figure}
+
+% The first thing we can observe from \ref{fig:periodicwave} is that there are three types of well defined points of interest in the cycle of an arbitrary periodic discrete wave of the form $ f(x) = f(x + P) \forall x \in D(f) $, where $ D $ is the domain of the function $ f $: The extreme points, that is, the maxima and minima, and the zeroes.
+
+% It is easy to see that if we choose to plot the frame of ocurrence of those points as a function of their cardinality, breaking ties arbitrarily and consistently, we will invariably have a straight line of the form $ y = a x + b $, in the case of strictily periodic waves, with the slope related to the dominant frequency, and the intercept related to the concept of phase.
+
+% Also, we can argue that using the maxima, instead of the zeroes, brings some simplifications to the process, as the maxima are generally less abundant, and more directly well defined, both in the horizontal and vertical axes.
+The vertical difference of one of such points and the immediately adjacent one is the period $ T $ of the periodic wave.
+
+% Two things about this straight line become, then, immeadiately apparent: the first is that unit in the vertical axis is $ i $, that is, the indices of the frames of the discrete wave, and the second is that $ x \in \mathbb{N} $.
+Also, we have that the slope of the line equation is $ a = \frac{y_{k+1} - y_{k}}{x_{k+1} - x_{k}} = T $.
+
+% We can then rewrite the line equation as $ y_k = T x_k + b $
+
+% % $ b = \frac{x_{k+1} y_{k} - x_{k} y_{k+1}}{x_{k+1} - x_{k}} $
+
+% If we define the local frequency $ f $ of a discrete, periodic wave as the ratio between the number cycles of said wave and the number $ n $ of frames of that wave, we can, recalling that the period of a wave is the inverse of its frequency, obtain the dominant frequency of the wave by using the equation $ f = \frac{n}{T} = \frac{n}{a} $.
+Alternatively, the relation $ f = (\#F^\pm - 1) n / (F^\pm[-1] - F^\pm[0])$, where $ \#F^\pm $ denotes the cardinality of the frontier, while $ F^\pm[0] $ and $ F^\pm[-1] $ stand for the frontier's first and last item, respectively, could also be used.
+
+% Similarly, it is simpler to obtain information about the phase of a periodic wave from the extremes.
+Keeping in mind that a phase shift of $ 2 \pi $ would shift the wave by one whole period and also that a wave with zero phase would have its first maximum at $ i = 0 $, we need only to multiply the ratio between the frame where the first maximum ocurred and the period by
+% $ 2 \pi $, as in the formula $ \phi = \frac{2 \pi F^+[0]}{T}$
+
+% About the result above it is important to keep in mind that congruency with the results of the Fourier Transform is not guarateed, as the shape of an arbitrary wave and, consequently, the relative position of its maximum can depart considerably from the position of an ordinary sinusoid.
+
+% By this token, it is more convenient to define the phase obtained from the negative frontier in terms of the one obtained with the use of the positive one; To achieve that, it is sufficient to infer the relative position of the maximum in relation to the minimum, and use that in the phase formula presented above.
+
+\begin{figure}[ht!]
+  \centering
      \includegraphics{02pcsraw-alto.pdf}
-    % \def\svgwidth{\linewidth}
-    %\input{01signalenvelope.pdf_tex}
   \caption{Raw pseudo-cycles superimposed as defined by the positive and negative frontiers, for the alto singer signal.}

### 2020-11-05 Papers/02_Sound_Representation/images/02frontiergaps.py

@@ -1,0 +1,146 @@
+import plotly.graph_objects as go
+import numpy as np
+import signal_envelope as se
+import sys
+
+def refine_frontier(Xp, W):
+  "find additional frontier points, and return an array with then, if found"
+  if W[Xp[0]] >= 0:
+    e = np.argmax
+  else:
+    e = np.argmin
+  L = Xp[1:] - Xp[:-1]
+  avgL = np.average(L)
+  stdL = np.std(L)
+  Xnew = []
+  for i in range(1, Xp.size):
+    x0 = int(Xp[i - 1])
+    x1 = int(Xp[i])
+    if x1 - x0 > avgL + 2 * stdL:
+      Xzeroes = []
+      currsign = np.sign(W[x0])
+      for i in range(x0 + 1, x1):
+        if currsign != np.sign(W[i]):
+          Xzeroes.append(i)
+          currsign = np.sign(W[i])
+      if len(Xzeroes) > 1:
+        Xnew.append(Xzeroes[0] + e(W[Xzeroes[0] : Xzeroes[-1]]))
+  return np.array(Xnew, dtype=np.int)
+
+
+'''==============='''
+''' Read wav file '''
+'''==============='''
+
+name = "piano33"
+
+W, fps = se.read_wav(f"C:/Users/tesse/Desktop/Files/Dropbox/0_Thesis/Python/Samples/{name}.wav")
+W = W - np.average(W)
+n = W.size
+X = np.arange(n)
+
+Xpos, Xneg = se.get_frontiers(W)
+
+Xposnew = refine_frontier(Xpos, W)
+Xnegnew = refine_frontier(Xneg, W)
+
+
+
+'''============================================================================'''
+'''                                    PLOT                                    '''
+'''============================================================================'''
+FONT = dict(
+    family="Latin Modern Roman",
+    color="black",
+    size=13.3333
+  )
+
+'''Plotting'''
+fig = go.Figure()
+fig.layout.template ="plotly_white" 
+fig.update_layout(
+  xaxis_title="<b><i>i</i></b>",
+  yaxis_title="<b>Amplitude</b>",
+  legend=dict(orientation='h', yanchor='top', xanchor='left', y=1.1),
+  margin=dict(l=0, r=0, b=0, t=0),
+  font=FONT,
+  titlefont=FONT
+)
+fig.layout.xaxis.title.font=FONT
[…]

### 2020-11-12 Papers/02_Sound_Representation/SoundRepresentation.tex
minor changes
progress with the paper

@@
 \usepackage[makeroom]{cancel}
-
 % \usepackage{auto-pst-pdf}
@@
 \begin{abstract}
+We propose representing a discrete wave based on the description of the pseudo cycle as a function of time.
 
@@
 
-\section{Introduction}
+\section{Introduction} % 500 words
+% Opener sentence -> Question | Assertion | Intriguing Revelations
+The idea, based on the Fourier transform theory, that a complex wave can be represented by a sum of sinusoids is ubiquituous in sound processing, owing considerably to the reversebility of the fourier transform.
 
+Despite its many uses, for discrete real waves, frequency and time domain representations are equivalent, as the later represents the discrete with the same number of floats.
+
+Additionally, eventual variations in the envelope of the wave have to be encoded as additional sinusoids \textcite{2019TarjanoNeuro}.
+
+% Literature review | Brief Context of Prior Research
+For these reasons, mixed representations were introduced, specially in the context of lossy and lossless sound compression
+
+% Restate Your Question as Something Not Fully Understood by Prior Research -> identify the gap in the literature use but, however, etc…
+
+However, those representations lack a more physical connection, and generally dont lend themselves well as basis representatios for further tasks, such as machine learning synthesis.
+
+% State the Significance of Your Question -> make readers understand why they should read your report
+The aim of this work is, then, to fill this gap, introducing a compact representation
+
+% What I did - State Your Claim | Objective | Hypothesis -> should clearly relate to the information gap
 We aim to exploit redundancies in the constitution of a discrete sound wave to formulate an alternative, more compact and thus easier to analyse, representation.
 Implicit in this affirmation is the somewhat loose assumption that the wave considered is reasonably well behaved, in the sense that redundancies exist to be identified.
-In contrast with \textcite{2019TarjanoNeuro}, we don't require the signal to be strictly harmonic.
+
+To that end, we use the cycle as the basic building block of a signal and describe its changes both in the vertical axis, assumed to be due to the varying amplitude of the signal, and in the local horizontal axis, reflecting global frequency variations.
 
 We aim in providing a method that, with minimal manual interference, is able to reduce a digital wave to a more compact and meaningful representation.
+% brief outlook on the structure of the paper
 
@@
 
-\subsection{Pseudo Cycles}
+\subsection{Identifying the Pseudo Cycles}
 
-This work proposes the pseudo cycle as the building block of a discrete wave, assuming its parametric shape to be constant throughout the space of the digital signal; in order words, the adjective pseudo is used because the two other descriptions of the wave, namely, the envelope and the length of the pseudo cycles act into the pseudo cycles shape itself, modifying it in function of the time.
+This work proposes the (pseudo) cycle as the building block of a discrete wave, assuming its waveform to be constant throughout the space of the digital signal; in other words, the adjective pseudo is used because the two other descriptions of the wave, namely, the envelope and the length of the pseudo cycles act into the pseudo cycles shape itself, modifying it in function of the time, during the lifetime of the signal.
 
@@
 
-Using the envelope detection algorithm presented in \textcite{2020TarjanoRobust} one can identify the frontiers, both positive and negative, of a digital signal: Those are, ideally, respectively composed of the maxima and minima of each pseudo-cycle of a discrete signal.
+Using the envelope detection algorithm presented in \textcite{2020TarjanoRobust} one can identify the frontiers, both positive and negative, of a digital signal: Those are respectively composed of the maxima and minima of each pseudo-cycle of a discrete signal.
 
-If we disconsider inaccuracies we could use either one of the frontiers to infer the exact position, wavelength and amplitude of each pseudo-cycle and to, provided that the shape of the wave is also known, reconstruct the original wave with those pieces of information.
+If we disconsider inaccuracies we could use either one of the frontiers indifferently to infer the exact position, waveform and amplitude of each pseudo-cycle and to reconstruct the original wave with those information.
 Consider, for example, the digital signal presented in figure \ref{fig:signalenvelope}
@@
 It can be seen that both frontiers provide approximately the same information about the location of each pseudo-cycle, albeit with a phase difference between them.
-As the original algorithm proposed in \textcite{2020TarjanoRobust} is meant for envelope detection, maxima or minima that lies close to the already established frontiers are sometimes neglected, without significant impact in the envelope extracted.
 
-This can be seen in the discrete representation of the sound of the key 33 of a grand piano in figure \ref{fig:frontiergaps}, as the black squares and diamonds: intuitively, one can see that some local maxima and minima are missing, compromising the identification of all pseudo cycles.
-For our purposes, however, it is beneficial to accurately identify as many local maxima and minima, and hence pseudo cycles, as possible, as the impact on the shape of a pseudo cycle is significant.
+However, as the algorithm proposed in \textcite{2020TarjanoRobust} is originally meant for envelope detection, maxima or minima that lies close to the already established frontiers are sometimes neglected, without significant impact in the shape of the envelope extracted.
 
-We are going to assume that all the local extremes identified by the envelope detection algorithm are indeed correct; albeit not being strictily accurate, this approximation is very reasonable and simplifies further steps immensely, as we now need only to idenfity pseudo cycles with periods $ T $, that is, the difference between two adjacent maxima or minima, that are significantly bigger than the average observed in the discrete wave, and identify their local maximum or minimum, respectively.
+This can be seen in the discrete representation of the sound of the key 33 of a grand piano in figure \ref{fig:frontiergaps}, as the black squares and diamonds: intuitively, one can see that some local maxima and minima were not detected, compromising the identification of all pseudo cycles.
 
-The result of this procedure can be seen ad the gray squares and diamonds in figure \ref{fig:frontiergaps}.
+For our purposes, however, it is beneficial to accurately identify as many local maxima and minima, and the resulting pseudo cycles, as possible, as the impact on the shape of a pseudo cycle is significant.
+
+We are going to assume that all the local extremes identified by the envelope detection algorithm are indeed correct; albeit not being strictily accurate, this approximation is very reasonable and simplifies further steps immensely, as we now need only to idenfity pseudo cycles with periods $ T $, that is, the difference between two adjacent maxima or minima, that are significantly bigger than the majority of the ones observed in the discrete wave, and identify their local maximum or minimum, respectively.
+
+To that end we are going to use the mode of the periods as the centrality measure, instead of the average.
+Similarly, we will use the mean absolute deviation in relation to the mode as a measure of the dispersion of the periods.
+
+In figure \ref{fig:modevsaverage} an except of the plot of the frequency of the errors is shown for both measures, for the case of a discrete wave representing the sound of the key 33 of a grand piano.
+
+The errors are the frequency of absolute difference between the centrality of the distribution of all combined periods, as defined by both the positive and negative frontiers, and their values.
+
+In addition to the smaller mean absolute deviation, the errors when the mode is used are centered around zero, which is not the case for when the centrality measure used is the average.
 
@@
   \centering
-     \includegraphics{images/02frontiergaps.pdf}
+     \includegraphics{images/02modevsaverage.pdf}
+  \caption{Frequency of errors for the periods of a digital signal representing the sound of the key 33 of a grand piano when two different centrality measures are used: the mode of the periods and their average.}
+  \label{fig:modevsaverage}
+\end{figure}
+
+The result of this procedure can be seen as the additional squares and diamonds in figure \ref{fig:frontiergaps}, representing new maxima and minima, respectively, found by reanalising pulses whose period is superior to 2 mean absolute deviations from the mode of the periods distribution, both for the positive and the negative frontiers.
+
+A new mode is then calculated from the new pulse distribution, as well as the mean absolute deviation; the procedure is repeated until no more new pseudo pulses are found or the mean absolute deviation is increased by the addition of the newfound extrema.
+
+\begin{figure}[ht!]
+  \centering
+     \includegraphics{images/03frontiergaps.pdf}
   \caption{Close-up view of a discrete signal representing the sound of the key 33 of a grand piano and its original positive and negative frontiers, in black, as well as the additional extrema identified during the refining process.}
@@
 \end{figure}
+
+figure \ref{fig:plotXps} illustrates the behaviour of the periods as inferred by both frontiers, before and after the refinement.
+
+\begin{figure}[ht!]
+  \centering
+     \includegraphics{images/04plotXps.pdf}
+  \caption{Frame of occurrence of the maxima and minima of a discrete signal representing the sound of the key 33 of a grand piano before and after the refinement process.}
+  \label{fig:plotXps}
+\end{figure}
+
+\subsection{Pseudo Cycles evolution over time}
 
@@
 
-\begin{figure}[ht!]
-  \centering
-     \includegraphics{02pcsraw-alto.pdf}
-  \caption{Raw pseudo-cycles superimposed as defined by the positive and negative frontiers, for the alto singer signal.}
-  \label{fig:pcrawalto}
-\end{figure}
+% \begin{figure}[ht!]
+%   \centering
+%      \includegraphics{02pcsraw-alto.pdf}
+%   \caption{Raw pseudo-cycles superimposed as defined by the positive and negative frontiers, for the alto singer signal.}
+%   \label{fig:pcrawalto}
+% \end{figure}
 
-We can normalize the amplitude the pseudo-cycles between zero and one, by dividing each of them by their maximum.
+% We can normalize the amplitude the pseudo-cycles between zero and one, by dividing each of them by their maximum.
 The lengths of each pseudo-cycle can be normalized with the use of the Fourier transform; to avoid any loss of data, we can use the maximum wavelength as the standard.
 
-\begin{figure}[ht!]
-  \centering
-    %  \includegraphics[width=0.9\linewidth]{01signalenvelope}
-     \includegraphics{03pcs-alto.pdf}
-    % \def\svgwidth{\linewidth}
-    %\input{01signalenvelope.pdf_tex}
-  \caption{Normalized pseudo-cycles superimposed as defined by the positive and negative frontiers, for an alto singer signal, and the average of all normalized pseudo-cycles.}
-  \label{fig:pcalto}
-\end{figure}
+% \begin{figure}[ht!]
+%   \centering
+%     %  \includegraphics[width=0.9\linewidth]{01signalenvelope}
+%      \includegraphics{03pcs-alto.pdf}
+%     % \def\svgwidth{\linewidth}
+%     %\input{01signalenvelope.pdf_tex}
+%   \caption{Normalized pseudo-cycles superimposed as defined by the positive and negative frontiers, for an alto singer signal, and the average of all normalized pseudo-cycles.}
+%   \label{fig:pcalto}
+% \end{figure}
 
-Comparing figures \ref{fig:pcrawalto} and \ref{fig:pcalto} we note the effect of the normalization, as a means to reduce the variance of the pseudo-cycles, specially in the regions near the discontinuities.
+% Comparing figures \ref{fig:pcrawalto} and \ref{fig:pcalto} we note the effect of the normalization, as a means to reduce the variance of the pseudo-cycles, specially in the regions near the discontinuities.
 This will enable a more accurate approximation of the average wave.
 
-\begin{figure}[ht!]
-  \centering
-    %  \includegraphics[width=0.9\linewidth]{01signalenvelope}
-     \includegraphics{04averagepc-alto.pdf}
-    % \def\svgwidth{\linewidth}
-    %\input{01signalenvelope.pdf_tex}
-  \caption{Average pseudo-cycles, superimposed, after a change of phase of $-\pi$ on the average pseudo-cycle defined by the negative frontier, for an alto singer digital signal.}
-  \label{fig:averagepc-alto}
-\end{figure}
+% \begin{figure}[ht!]
+%   \centering
+%     %  \includegraphics[width=0.9\linewidth]{01signalenvelope}
+%      \includegraphics{04averagepc-alto.pdf}
+%     % \def\svgwidth{\linewidth}
+%     %\input{01signalenvelope.pdf_tex}
+%   \caption{Average pseudo-cycles, superimposed, after a change of phase of $-\pi$ on the average pseudo-cycle defined by the negative frontier, for an alto singer digital signal.}
+%   \label{fig:averagepc-alto}
+% \end{figure}
 
-Figure \ref{fig:averagepc-alto} makes clear that the average pseudo-cycles for the alto singer signal, as estimated by both frontiers, are in reasonable agreement, despite the evident vertical asymmetries of the underlying wave.
+% Figure \ref{fig:averagepc-alto} makes clear that the average pseudo-cycles for the alto singer signal, as estimated by both frontiers, are in reasonable agreement, despite the evident vertical asymmetries of the underlying wave.
 
-That is not always the case, however, as figure \ref{fig:averagepc-piano33} illustrates.
+% That is not always the case, however, as figure \ref{fig:averagepc-piano33} illustrates.
 For a discrete signal representing the recording o the key 33 of a grand piano, we can see that the average pseudo-cycle inferred from the positive frontier is substantially different from the one obtained with the negative frontier.
 
-The horizontal lines indicate the maximum of each pseudo cycle, being the point of adjacency between two subsequent pseudo cycles.
+% The horizontal lines indicate the maximum of each pseudo cycle, being the point of adjacency between two subsequent pseudo cycles.
 For the average defined by the negative frontier we see that there is a local maxima near that middle that is almost as high as the global maxima, which is not the case as the average defined by the positive one.
 
-\begin{figure}[ht!]
-  \centering
-    %  \includegraphics[width=0.9\linewidth]{01signalenvelope}
-     \includegraphics{04averagepc-piano33.pdf}
-    % \def\svgwidth{\linewidth}
-    %\input{01signalenvelope.pdf_tex}
-  \caption{Average pseudo-cycles, superimposed, after a change of phase of $-\pi$ on the average pseudo-cycle defined by the negative frontier, for a grand piano digital signal.
+% \begin{figure}[ht!]
+%   \centering
+%     %  \includegraphics[width=0.9\linewidth]{01signalenvelope}
+%      \includegraphics{04averagepc-piano33.pdf}
+%     % \def\svgwidth{\linewidth}
+%     %\input{01signalenvelope.pdf_tex}
+%   \caption{Average pseudo-cycles, superimposed, after a change of phase of $-\pi$ on the average pseudo-cycle defined by the negative frontier, for a grand piano digital signal.
 The horizontal lines in the figure mark the points of maximum of each of the average pseudo cycles.}
-  \label{fig:averagepc-piano33}
-\end{figure}
+%   \label{fig:averagepc-piano33}
+% \end{figure}
 
-We could take the average of those two representations, but that would dilute the parametric representation of the pseudo-cycle, besides introducing unnecessary complication in the stage of recreating the wave, as we would have to deal with redundant information about the location of each pseudo-cycle.
+% We could take the average of those two representations, but that would dilute the parametric representation of the pseudo-cycle, besides introducing unnecessary complication in the stage of recreating the wave, as we would have to deal with redundant information about the location of each pseudo-cycle.
 
-Instead, a measure of the statistical variance of the lengths of the pseudo cycles will be used in order to select the most appropriate frontier as guide for the construction of the average pseudo cycle waveform.
+% Instead, a measure of the statistical variance of the lengths of the pseudo cycles will be used in order to select the most appropriate frontier as guide for the construction of the average pseudo cycle waveform.
 
-Figure \ref{fig:std-piano33} shows the distribution of the lengths of the pseudo cycles of the grand piano digital signal.
+% Figure \ref{fig:std-piano33} shows the distribution of the lengths of the pseudo cycles of the grand piano digital signal.
 We can use the standard deviation as this measure, and use the pseudo cycles as defined by the frontier with the least standard deviation.
 
-\begin{figure}[ht!]
-  \centering
-    %  \includegraphics[width=0.9\linewidth]{01signalenvelope}
-     \includegraphics{05std-piano33.pdf}
-    % \def\svgwidth{\linewidth}
-    %\input{01signalenvelope.pdf_tex}
-  \caption{Histogram of the distribution of lengths of the pseudo cycles, as defined by the positive and negative frontiers, and the accompanying standard deviation, for a digital signal of the key 33 of a grand piano.}
-  \label{fig:std-piano33}
-\end{figure}
+% \begin{figure}[ht!]
+%   \centering
+%     %  \includegraphics[width=0.9\linewidth]{01signalenvelope}
+%      \includegraphics{05std-piano33.pdf}
+%     % \def\svgwidth{\linewidth}
+%     %\input{01signalenvelope.pdf_tex}
+%   \caption{Histogram of the distribution of lengths of the pseudo cycles, as defined by the positive and negative frontiers, and the accompanying standard deviation, for a digital signal of the key 33 of a grand piano.}
+%   \label{fig:std-piano33}
+% \end{figure}
 
-Thus, a tillable parametric representation of the average pseudo-cycle can be obtained via the least mean squares method; we are interested in the polynomial that best approximates the average pseudo cycle, respecting the conditions that the beginning and the end of this polynomial must coincide, as must the first derivative of that polynomial in those places, if we are to obtain a smooth transition between pseudo cycles.
+% Thus, a tillable parametric representation of the average pseudo-cycle can be obtained via the least mean squares method; we are interested in the polynomial that best approximates the average pseudo cycle, respecting the conditions that the beginning and the end of this polynomial must coincide, as must the first derivative of that polynomial in those places, if we are to obtain a smooth transition between pseudo cycles.
 
-From \textcite{2013SelesnickLeast} we know that a constrained least mean squares problem of the form $ \underset{A}{\min} ||Q A - Y||^2_2 \quad \text{subject to} \quad V A = B $ has a closed form approximate solution $ \hat{A} = \left(Q^{T} Q\right)^{-1} \left(Q^{T} Y - V^{T} \left(V \left(Q^{T} Q\right)^{-1} V^{T}\right)^{-1} V \left(Q^{T} Q\right)^{-1} Q^{T} Y - B \right) $, where $ Y $ is the vector we are interested in approximating.
+% From \textcite{2013SelesnickLeast} we know that a constrained least mean squares problem of the form $ \underset{A}{\min} ||Q A - Y||^2_2 \quad \text{subject to} \quad V A = B $ has a closed form approximate solution $ \hat{A} = \left(Q^{T} Q\right)^{-1} \left(Q^{T} Y - V^{T} \left(V \left(Q^{T} Q\right)^{-1} V^{T}\right)^{-1} V \left(Q^{T} Q\right)^{-1} Q^{T} Y - B \right) $, where $ Y $ is the vector we are interested in approximating.
 
-We can proceed to define the real vector $ U = \{ u_0, u_1, \cdots, u_{m-1}\} $ as the average of all normalized pseudo-cycles, with $ m \in \mathbb{N} $ equal to the number of elements and the maximum length of the pseudo-cycles identified by the envelope detection algorithm.
+% We can proceed to define the real vector $ U = \{ u_0, u_1, \cdots, u_{m-1}\} $ as the average of all normalized pseudo-cycles, with $ m \in \mathbb{N} $ equal to the number of elements and the maximum length of the pseudo-cycles identified by the envelope detection algorithm.
 
-As we are interested in maintaining $ C^0 $ and $ C^1 $ continuity between two adjacent pseudo-pulses, it is convenient to define the vector $ Y $ as two vectors $ E $ stacked, that is $ Y = \{ u_0, u_1, \cdots, u_{m-1}, u_0, u_1, \cdots, u_{m-1}\} $.
+% As we are interested in maintaining $ C^0 $ and $ C^1 $ continuity between two adjacent pseudo-pulses, it is convenient to define the vector $ Y $ as two vectors $ E $ stacked, that is $ Y = \{ u_0, u_1, \cdots, u_{m-1}, u_0, u_1, \cdots, u_{m-1}\} $.
 
-$ A $, the vector of coefficients we are interested in estimating, will thus be composed of $ 2 (k + 1) $ items, where $ k $ is the order of the polynomial to be used in the approximation.
+% $ A $, the vector of coefficients we are interested in estimating, will thus be composed of $ 2 (k + 1) $ items, where $ k $ is the order of the polynomial to be used in the approximation.
 Half those values are redundant, however, and $ A $ can thus be defined as $ A = \{ a_0, a_1, \cdots, a_k, a_0, a_1, \cdots, a_k \} $.
 $ Q $ is formed as usual, with the difference that
-it is duplicated horizontally, as seen in equation \ref{eq:matrix}.
+% it is duplicated horizontally, as seen in equation \ref{eq:matrix}.
 
-Anticipating that we will eventually evaluate this polynomial in various lengths, it is convenient to normalize $ X $ between $ 0 $ and $ 1 $, defining $ X = \left\{ \cancelto{0}{\frac{0}{m-1}}, \frac{1}{m-1}, \frac{2}{m-1}, \cdots, \cancelto{1}{\frac{m-1}{m-1}} \right\} $
+% Anticipating that we will eventually evaluate this polynomial in various lengths, it is convenient to normalize $ X $ between $ 0 $ and $ 1 $, defining $ X = \left\{ \cancelto{0}{\frac{0}{m-1}}, \frac{1}{m-1}, \frac{2}{m-1}, \cdots, \cancelto{1}{\frac{m-1}{m-1}} \right\} $
 
-\begin{equation} \label{eq:matrix}
-\stackrel{Q_{m \times 2 (k + 1)}} {
-\begin{bmatrix}
-  1      & x_0     & \cdots & x_0^k   & 1      & x_0     & \cdots & x_0^k   \\
-  1      & x_1     & \cdots & x_1^k   & 1      & x_1     & \cdots & x_1^k   \\
-  \vdots & \vdots  & \ddots & \vdots  & \vdots & \vdots  & \ddots & \vdots  \\
-  1      & x_{m-1} & \cdots & x_{m-1} & 1      & x_{m-1} & \cdots & x_{m-1} \\
-\end{bmatrix}
-}
-\stackrel{A_{2 (k + 1) \times 1}}
-{
-\begin{bmatrix}
-a_0    \\
-a_1    \\
-\vdots \\
-a_k    \\
-a_0    \\
-a_1    \\
-\vdots \\
-a_k    \\
-\end{bmatrix}
-}
-=
-\stackrel{Y_{2 m \times 1}}
-{
-\begin{bmatrix}
-y_0     \\
-y_1     \\
-\vdots  \\
-y_{m-1} \\
-y_0     \\
-y_1     \\
-\vdots  \\
-y_{m-1} \\
-\end{bmatrix}
-}
-\end{equation}
+% \begin{equation} \label{eq:matrix}
+% \stackrel{Q_{m \times 2 (k + 1)}} {
+% \begin{bmatrix}
+%   1      & x_0     & \cdots & x_0^k   & 1      & x_0     & \cdots & x_0^k   \\
+%   1      & x_1     & \cdots & x_1^k   & 1      & x_1     & \cdots & x_1^k   \\
+%   \vdots & \vdots  & \ddots & \vdots  & \vdots & \vdots  & \ddots & \vdots  \\
+%   1      & x_{m-1} & \cdots & x_{m-1} & 1      & x_{m-1} & \cdots & x_{m-1} \\
+% \end{bmatrix}
+% }
+% \stackrel{A_{2 (k + 1) \times 1}}
+% {
+% \begin{bmatrix}
+% a_0    \\
+% a_1    \\
+% \vdots \\
+% a_k    \\
+% a_0    \\
+% a_1    \\
+% \vdots \\
+% a_k    \\
+% \end{bmatrix}
+% }
+% =
+% \stackrel{Y_{2 m \times 1}}
+% {
+% \begin{bmatrix}
+% y_0     \\
+% y_1     \\
+% \vdots  \\
+% y_{m-1} \\
+% y_0     \\
+% y_1     \\
+% \vdots  \\
+% y_{m-1} \\
+% \end{bmatrix}
+% }
+% \end{equation}
 
-To maintain the continuity and smoothness, as well as the identity between the first and second half of the coefficients, $ V $ can be defined as in equation \ref{eq:constraint}.
+% To maintain the continuity and smoothness, as well as the identity between the first and second half of the coefficients, $ V $ can be defined as in equation \ref{eq:constraint}.
 
-\begin{equation} \label{eq:constraint}
-\stackrel{V_{k + 3 \times 2 (k + 1)}} {
-\begin{bmatrix}
-  1      & 0       & \cdots & 0           & -1     & 0        & \cdots & 0                \\
-  0      & 1       & \cdots & 0           & 0      & -1       & \cdots & 0                \\
-  \vdots & \vdots  & \ddots & \vdots      & \vdots & \vdots   & \ddots & \vdots           \\
-  0      & 0       & \cdots & 1           & 0      & 0        & \cdots & -1               \\
-  1      & x_0     & \cdots & x_0^k       & -1     & -x_{m-1} & \cdots & -x_{m-1}^k       \\
-  0      & 1       & \cdots & k x_0^{k-1} & 0      & -1       & \cdots & -k x_{m-1}^{k-1} \\
-\end{bmatrix}
-}
-\stackrel{A_{2 (k + 1) \times 1}}
-{
-\begin{bmatrix}
-a_0    \\
-a_1    \\
-\vdots \\
-a_k    \\
-a_0    \\
-a_1    \\
-\vdots \\
-a_k    \\
-\end{bmatrix}
-}
-=
-\stackrel{B_{k + 2 \times 1}}
-{
-\begin{bmatrix}
-0      \\
-\vdots \\
-0      \\
-\end{bmatrix}
-}
-\end{equation}
+% \begin{equation} \label{eq:constraint}
+% \stackrel{V_{k + 3 \times 2 (k + 1)}} {
+% \begin{bmatrix}
+%   1      & 0       & \cdots & 0           & -1     & 0        & \cdots & 0                \\
+%   0      & 1       & \cdots & 0           & 0      & -1       & \cdots & 0                \\
+%   \vdots & \vdots  & \ddots & \vdots      & \vdots & \vdots   & \ddots & \vdots           \\
+%   0      & 0       & \cdots & 1           & 0      & 0        & \cdots & -1               \\
+%   1      & x_0     & \cdots & x_0^k       & -1     & -x_{m-1} & \cdots & -x_{m-1}^k       \\
+%   0      & 1       & \cdots & k x_0^{k-1} & 0      & -1       & \cdots & -k x_{m-1}^{k-1} \\
+% \end{bmatrix}
+% }
+% \stackrel{A_{2 (k + 1) \times 1}}
+% {
+% \begin{bmatrix}
+% a_0    \\
+% a_1    \\
+% \vdots \\
+% a_k    \\
+% a_0    \\
+% a_1    \\
+% \vdots \\
+% a_k    \\
+% \end{bmatrix}
+% }
+% =
+% \stackrel{B_{k + 2 \times 1}}
+% {
+% \begin{bmatrix}
+% 0      \\
+% \vdots \\
+% 0      \\
+% \end{bmatrix}
+% }
+% \end{equation}
 
-The initial $k+1$ lines of $V$ assure that the first half of the coefficients of $A$ are equal to the second half, while the penultimate line ensures $C^0$ continuity and the last line ensures $C^1$ continuity, guaranteeing that the derivatives at the beginning and the end of the parametric pseudo cycles are equal.
+% The initial $k+1$ lines of $V$ assure that the first half of the coefficients of $A$ are equal to the second half, while the penultimate line ensures $C^0$ continuity and the last line ensures $C^1$ continuity, guaranteeing that the derivatives at the beginning and the end of the parametric pseudo cycles are equal.
 
-\begin{figure}[ht!]
-  \centering
-    %  \includegraphics[width=0.9\linewidth]{01signalenvelope}
-     \includegraphics{06approximation-alto}
-    % \def\svgwidth{\linewidth}
-    %\input{01signalenvelope.pdf_tex}
-  \caption{Two adjacent pseudo-cycles. The detail view illustrates that $C^0$ and $C^1$ smoothness are maintained in the parametric representation when two reconstructions are stacked horizontally.}
-  \label{fig:05approximation-alto}
-\end{figure}
+% \begin{figure}[ht!]
+%   \centering
+%     %  \includegraphics[width=0.9\linewidth]{01signalenvelope}
+%      \includegraphics{06approximation-alto}
+%     % \def\svgwidth{\linewidth}
+%     %\input{01signalenvelope.pdf_tex}
+%   \caption{Two adjacent pseudo-cycles. The detail view illustrates that $C^0$ and $C^1$ smoothness are maintained in the parametric representation when two reconstructions are stacked horizontally.}
+%   \label{fig:05approximation-alto}
+% \end{figure}
 

### 2020-11-12 Papers/02_Sound_Representation/images/02modevsaverage.py

@@ -1,0 +1,107 @@
+import plotly.graph_objects as go
+import numpy as np
+import signal_envelope as se
+from statistics import mode
+import sys
+sys.path.insert(0,"C:/Users/tesse/Desktop/Files/Dropbox/0_Thesis/Python")
+import hp as hp
+
+'''==============='''
+''' Read wav file '''
+'''==============='''
+
+# def stats(L):
+#   avg = np.average(L)
+#   mde = mode(L)
+#   std_avg = np.average(np.abs(L - avg))
+#   std_mde = np.average(np.abs(L - mde))
+#   print("std:", std_avg, std_mde, np.min(L), mde, np.max(L))
+#   return avg, std_avg, mde, std_mde, np.std(L)
+
+name = "piano33"
+W, fps = se.read_wav(f"C:/Users/tesse/Desktop/Files/Dropbox/0_Thesis/Python/Samples/{name}.wav")
+W = W - np.average(W)
+amp = np.max(np.abs(W))
+n = W.size
+X = np.arange(n)
+
+Xpos, Xneg = se.get_frontiers(W)
+Xpos, Xneg = Xpos.astype(np.int), Xneg.astype(np.int)
+posT = Xpos[1:] - Xpos[:-1]
+negT = Xneg[1:] - Xneg[:-1]
+T = np.hstack([posT, negT])
+
+# Xpos = hp.refine_frontier_iter(Xpos, W)
+# Xneg = hp.refine_frontier_iter(Xneg, W)
+# posT_ref = Xpos[1:] - Xpos[:-1]
+# negT_ref = Xneg[1:] - Xneg[:-1]
+# T_ref = np.hstack([posT_ref, negT_ref])
+
+mode_error = T - mode(T)
+modeX, modeY = np.unique(mode_error, return_counts=True)
+
+avg_error = T - np.average(T)
+avgX, avgY = np.unique(avg_error, return_counts=True)
+
+'''============================================================================'''
+'''                              PLOT LINES                                    '''
+'''============================================================================'''
+FONT = dict(
+    family="Latin Modern Roman",
+    color="black",
+    size=13.3333
+  )
+
+'''Plotting'''
+fig = go.Figure()
+fig.layout.template ="plotly_white" 
+fig.update_layout(
+  xaxis_title="<b><i>T</i></b>",
+  yaxis_title="<b>Number of Ocurrences</b>",
[…]

### 2020-11-12 Papers/02_Sound_Representation/images/03frontiergaps.py

@@ -1,31 +1,34 @@
 import plotly.graph_objects as go
 import numpy as np
 import signal_envelope as se
+from statistics import mode
 import sys
+sys.path.insert(0,"C:/Users/tesse/Desktop/Files/Dropbox/0_Thesis/Python")
+import hp as hp
 
-def refine_frontier(Xp, W):
-  "find additional frontier points, and return an array with then, if found"
-  if W[Xp[0]] >= 0:
-    e = np.argmax
-  else:
-    e = np.argmin
-  L = Xp[1:] - Xp[:-1]
-  avgL = np.average(L)
-  stdL = np.std(L)
-  Xnew = []
-  for i in range(1, Xp.size):
-    x0 = int(Xp[i - 1])
-    x1 = int(Xp[i])
-    if x1 - x0 > avgL + 2 * stdL:
-      Xzeroes = []
-      currsign = np.sign(W[x0])
-      for i in range(x0 + 1, x1):
-        if currsign != np.sign(W[i]):
-          Xzeroes.append(i)
-          currsign = np.sign(W[i])
-      if len(Xzeroes) > 1:
-        Xnew.append(Xzeroes[0] + e(W[Xzeroes[0] : Xzeroes[-1]]))
-  return np.array(Xnew, dtype=np.int)
+# def refine_frontier(Xp, W):
+#   "find additional frontier points, and return an array with then, if found"
+#   if W[Xp[0]] >= 0:
+#     e = np.argmax
+#   else:
+#     e = np.argmin
+#   L = Xp[1:] - Xp[:-1]
+#   avgL = mode(L)
+#   stdL = np.average(np.abs(L - avgL))
+#   Xnew = []
+#   for i in range(1, Xp.size):
+#     x0 = int(Xp[i - 1])
+#     x1 = int(Xp[i])
+#     if x1 - x0 > avgL + 2 * stdL:
+#       Xzeroes = []
+#       currsign = np.sign(W[x0])
+#       for i in range(x0 + 1, x1):
+#         if currsign != np.sign(W[i]):
+#           Xzeroes.append(i)
+#           currsign = np.sign(W[i])
+#       if len(Xzeroes) > 1:
+#         Xnew.append(Xzeroes[0] + e(W[Xzeroes[0] : Xzeroes[-1]]))
+#   return np.array(Xnew, dtype=np.int)
 
 
 '''==============='''
@@ -41,8 +44,10 @@ X = np.arange(n)
 
[…]

### 2020-11-12 Papers/02_Sound_Representation/images/04plotXps.py

@@ -1,0 +1,123 @@
+import plotly.graph_objects as go
+import numpy as np
+import signal_envelope as se
+import sys
+sys.path.insert(0,"C:/Users/tesse/Desktop/Files/Dropbox/0_Thesis/Python")
+import hp as hp
+
+
+'''==============='''
+''' Read wav file '''
+'''==============='''
+
+
+name = "piano33"
+W, fps = se.read_wav(f"C:/Users/tesse/Desktop/Files/Dropbox/0_Thesis/Python/Samples/{name}.wav")
+W = W - np.average(W)
+amp = np.max(np.abs(W))
+n = W.size
+X = np.arange(n)
+
+Xpos_orig, Xneg_orig = se.get_frontiers(W)
+
+Xpos = hp.refine_frontier_iter(Xpos_orig, W)
+Xneg = hp.refine_frontier_iter(Xneg_orig, W)
+
+
+
+'''============================================================================'''
+'''                              PLOT LINES                                    '''
+'''============================================================================'''
+FONT = dict(
+    family="Latin Modern Roman",
+    color="black",
+    size=13.3333
+  )
+
+'''Plotting'''
+fig = go.Figure()
+fig.layout.template ="plotly_white" 
+fig.update_layout(
+  xaxis_title="<b>Ordinality</b>",
+  yaxis_title="<b><i>i</i></b>",
+  legend=dict(orientation='h', yanchor='top', xanchor='left', y=1.1),
+  margin=dict(l=0, r=0, b=0, t=0),
+  font=FONT,
+  titlefont=FONT
+)
+fig.layout.xaxis.title.font=FONT
+fig.layout.yaxis.title.font=FONT
+fig.update_xaxes(showline=False, showgrid=False, zerolinewidth=1, zerolinecolor='silver')
+fig.update_yaxes(showline=False, showgrid=False, zerolinewidth=1, zerolinecolor='silver')
+
+fig.add_trace(
+  go.Scatter(
+    name="Maxima", # <|<|<|<|<|<|<|<|<|<|<|<|
+    # x=Xpos,
+    y=Xpos_orig,
+    # fill="toself",
+    mode="lines",
+    line=dict(
+        width=2,
+        color="black",
+        dash="dash"
+        # showscale=False
+    ),
+    # visible = "legendonly"
+  )
+)
+
+fig.add_trace(
+  go.Scatter(
[…]

### 2020-11-18 Papers/02_Sound_Representation/SoundRepresentation.tex
minor changes

@@
 
-The result of this procedure can be seen as the additional squares and diamonds in figure \ref{fig:frontiergaps}, representing new maxima and minima, respectively, found by reanalising pulses whose period is superior to 2 mean absolute deviations from the mode of the periods distribution, both for the positive and the negative frontiers.
+The result of this procedure can be seen as the additional squares and diamonds in figure \ref{fig:frontiergaps}, representing new maxima and minima, respectively, found by iteratively reanalising and eventually splitting pulses whose period is superior to 2 mean absolute deviations from the mode of the periods distribution, both for the positive and the negative frontiers.
 
@@
 
-figure \ref{fig:plotXps} illustrates the behaviour of the periods as inferred by both frontiers, before and after the refinement.
+Figure \ref{fig:plotXps} illustrates the behaviour of the extrema as inferred by both frontiers, before and after the refinement.
+It can be seen, particularly at the end of the wave that each inference is a little biased, both before and after the refinement process.
+To solve that, we can calculate an intermediate measure of both inferred values, by averaging compatible values of both frontiers.
+
+In this way we have a balanced, unbiased estimative of the position of the pseudo pulses along the time that can be used to calculate the average waveform of the pseudo pulses in the signal of interest.
+
+We won't, in general, have an accurate estimative - unless methods such as the FFT are used - of the relative phase of this signal, since the points marked by the average frontier wont, in general, concide with points of interest of the waveform.
+
+In special cases such as sinusoids, this would be the case, as the average frontier would mark the zeros of each pseudo cycle.
+Nevertheless, our interest here is in periodicity, since the envetual extracted average waveform will have to be made tileable, rendering ineffective such a discussion.
+It will be shown below that the average waveform identified as described is indeed very close to tileable, and minimal further processing is required.
 
@@
      \includegraphics{images/04plotXps.pdf}
-  \caption{Frame of occurrence of the maxima and minima of a discrete signal representing the sound of the key 33 of a grand piano before and after the refinement process.}
+  \caption{Frame of occurrence of the maxima and minima of a discrete signal representing the sound of the key 33 of a grand piano before and after the refinement process, as well as the average of both refined extrema.}
   \label{fig:plotXps}
 \end{figure}
+
+We could use a polinomial approximation, fitting an arbitrary direct polynomial to the values.
+Given the range of number, that procedure would bring numerical instabillity to our model, however.
+Instead, we chose to divide the approximation into a straight line and a polynomial approximation of the deviation from this line.
+In that way we are able to enforce the overall pitch of the wave, while avoiding computational instability.
 

### 2020-11-18 Papers/02_Sound_Representation/images/04plotXps.py

@@ -23,7 +23,8 @@ Xpos_orig, Xneg_orig = se.get_frontiers(W)
 Xpos = hp.refine_frontier_iter(Xpos_orig, W)
 Xneg = hp.refine_frontier_iter(Xneg_orig, W)
 
-
+nn = min(Xpos.size, Xneg.size)
+Average_ref_X = np.round((Xpos[:nn] + Xneg[:nn]) / 2)#.astype(np.int)
 
 '''============================================================================'''
 '''                              PLOT LINES                                    '''
@@ -58,7 +59,7 @@ fig.add_trace(
     # fill="toself",
     mode="lines",
     line=dict(
-        width=2,
+        width=1,
         color="black",
         dash="dash"
         # showscale=False
@@ -75,7 +76,7 @@ fig.add_trace(
     # fill="toself",
     mode="lines",
     line=dict(
-        width=2,
+        width=1,
         color="gray",
         dash="dash"
         # showscale=False
@@ -92,7 +93,7 @@ fig.add_trace(
     # fill="toself",
     mode="lines",
     line=dict(
-        width=2,
+        width=1,
         color="black",
         # showscale=False
     ),
@@ -108,7 +109,7 @@ fig.add_trace(
     # fill="toself",
     mode="lines",
     line=dict(
-        width=2,
+        width=1,
         color="gray",
         # dash="dash"
         # showscale=False
@@ -117,7 +118,24 @@ fig.add_trace(
   )
 )
 
-fig.show(config=dict({'scrollZoom': True}))
+fig.add_trace(
+  go.Scatter(
+    name="Average Pseudo Pulses Position", # <|<|<|<|<|<|<|<|<|<|<|<|
+    # x=Xpos,
+    y=Average_ref_X,
+    # fill="toself",
+    mode="lines",
+    line=dict(
+        width=1,
+        color="silver",
+        # dash="dash"
+        # showscale=False
+    ),
+    # visible = "legendonly"
+  )
+)
+
+# fig.show(config=dict({'scrollZoom': True}))
 save_name = "./images/" + sys.argv[0].split('/')[-1].replace(".py", ".svg")
 fig.write_image(save_name, width=650, height=300, engine="kaleido", format="svg")
 print("saved:", save_name)

### 2020-11-23 Papers/01_Pulses_Envelope/PulsesEnvelope.tex
moved files to make rooms for new developments
revising "pulses envelope"

--- 2020-09-07 Papers/02_Pulses_Envelope/PulsesEnvelope.tex
@@
 \usepackage{graphicx}
-\usepackage[backend=biber, style=apa]{biblatex}
+\usepackage[backend=biber]{biblatex}
 \usepackage{amsmath}
-\usepackage{hyperref}
-\usepackage{hypcap}
+% \usepackage{hyperref}
+% \usepackage{hypcap}
 \DeclareMathOperator{\sign}{sign}
 \bibliography{bibli}
-\hypersetup{
-  colorlinks=true, 
-  allcolors=black,
-  pdfauthor={Carlos Henrique Tarjano Santos},
-  pdftitle={Robust Digital Envelope Estimation Via Geometric Properties of an Arbitrary Real Signal}
-}
+% \hypersetup{
+%   colorlinks=true, 
+%   allcolors=black,
+%   pdfauthor={Carlos Henrique Tarjano Santos},
+%   pdftitle={Robust Digital Envelope Estimation Via Geometric Properties of an Arbitrary Real Signal}
+% }
 \title{Robust Digital Envelope Estimation Via Geometric Properties of an Arbitrary Real Signal}
@@
 \section{Introduction}
-Envelope detection is ubiquitous in both analogue and digital signal processing \parencite{2011CaetanoImproved}.
+Envelope detection, also known as demodulation, is ubiquitous in both analogue and digital signal processing \parencite{2011CaetanoImproved}.
 Nevertheless, the literature in this area is very fragmented \parencite{2017LyonsDigital}.
@@
 
-In many natural signals, however, the temporal amplitude envelope of a signal plays a prominent role in the characteristics exhibited: according to \textcite{2017QiRelative}, for example, the envelope is at least as important as the fine structure of a sound wave in the context of the intelligibility of mandarin tones.
-This is also the case for English \parencite{1995ShannonSpeech}, where even envelopes modulating mostly noise were still capable of conveying meaning.
+In many contexts, however, the temporal amplitude envelope of a signal plays a prominent role: according to \textcite{2017QiRelative}, for example, the envelope is at least as important as the fine structure of a sound wave in the context of the intelligibility of Mandarin tones.
+This is also the case for the English language \parencite{1995ShannonSpeech}, where even envelopes modulating mostly noise were still capable of conveying meaning.
 
-The envelope helps to impart emotion and identity to the human voice \parencite{2018ZhuContributions}, and envelope preserving characteristics in concert halls are associated with their pleasantness \parencite{2011LokkiEngaging}.
+The envelope also helps to impart emotion and identity to the human voice \parencite{2018ZhuContributions}, and envelope preserving characteristics in concert halls are associated with their pleasantness \parencite{2011LokkiEngaging}.
 
@@
 
-Addressing the different units in the horizontal and vertical axes, one can transform the DSP problem of envelope detection into the geometric problem of defining the shape of a set of points in $ \mathbb{R}^2 $.
-For that purpose, \textcite{1983Edelsbrunnershape} introduced the concept of alpha-shapes, a mathematically well-defined extension to the convex hull of a finite set of points, closely related to the Delaunay triangulation and Voronoi diagrams of those points.
+The problem of envelope detection, normally associated with the field of digital signal processing, can be made equivalent to the geometric problem of defining the (outer) shape of a set of points in $ \mathbb{R}^2 $, provided a method for addressing the difference of units in the horizontal, time related and vertical, amplitude related, axes of a digital signal is available, rendering the original DSP problem readily approachable via geometric techniques.
 
-This approach is used in areas such as the detection of features in images \parencite{2016VarytimidisAlpha}, reconstruction of surfaces from a cloud of points \parencite{2015WuAutomated} and Spectroscopy \parencite{2019XuModeling}, with the last work, which involves the estimation and removal of the Blaze function (a kind of envelope) of an echelle spectrograph, being particularly illustrative of the potential of cross-fertilization.
+One direct way of defining the shape of a set of points in $ \mathbb{R}^2 $ is using the convex hull: as defined in \textcite{2008BergComputational}, the convex hull of a set of points $ S $ in two dimensional space is the smallest convex subset of the plane that contains all the points in $ S $.
+Convex hulls are unique, in the sense that a set of points define one and only one convex hull.
 
-Steps in the direction of translating geometric algorithms to the context of envelope detection were also made by \textcite{2015YangSkeleton} via an algorithm based on the construction of a skeleton underlying the wave, and also via the direct translation of computer vision methods to the task \parencite{2015YangRepresenting}.
+Generalizing that definition, \textcite{1983Edelsbrunnershape} introduced the concept of alpha-shapes, that can be seen as a Concave Hull; a mathematically well-defined generalization of the Convex Hull of a finite set of points, closely related to the Delaunay triangulation and Voronoi diagrams of those points.
 
-Following this path, we present a general approach to envelope detection, exploiting the intrinsic characteristics of a generic, spectrally complex wave, in order to avoid the need for manual intervention or parameter tuning.
+The intuition behind the alpha shapes algorithm as that only points that can be touched by a circle of radius $ \alpha $ coming from an infinite distance towards the set of points of interest are part of the frontier, that is, of the Concave Hull of this set.
+
+Once an $ \alpha $ parameter is defined alpha shapes conserve the uniqueness property, besides reducing to the Convex Hull when $ \alpha = \infty $.
+Consider, for example, the artificially generate discrete wave in figure \ref{fig:Hulls}
+below, where the links between consecutive samples were deemphasised to highlight the "set of points" nature of a signal: The Convex Hull and a Concave Hull, in the form of an alpha shape with a particular value of $ \alpha $ are illustrated.
+
+\begin{figure}[ht!]
+  \centering
+    \includegraphics{images/00ConvexHull.pdf}
+  \caption{A discrete signal interpreted as a set of points in the Cartesian plane, and the Convex Hull and Concave Hull at $ \alpha=300 $ defined by it}
+  \label{fig:Hulls}
+\end{figure}
+
+The alpha shapes approach is used in areas such as the detection of features in images \parencite{2016VarytimidisAlpha}, reconstruction of surfaces from a cloud of points \parencite{2015WuAutomated} and Spectroscopy \parencite{2019XuModeling}, with the last work, which involves the estimation and removal of the Blaze function (a kind of envelope) of an echelle spectrograph, being particularly illustrative of the potential of synergy between geometric and DSP approaches.
+
+Other steps in the direction of translating geometric algorithms to the context of envelope detection were also made by \textcite{2015YangSkeleton} via an algorithm based on the construction of a skeleton underlying the digital wave of interest, and also via the direct translation of computer vision methods to the task of envelope detection \parencite{2015YangRepresenting}.
+
+Following this path, we present a general approach to envelope detection, exploiting the intrinsic characteristics of a generic, spectrally complex wave, in order to avoid as much as possible the need for manual intervention or parameter tuning, such as the choice of $ \alpha $.
+
+% <|<|<|<|<| TODO
 
@@
 
+\section{Characterization of the temporal envelope}
+
+The interest in the theory of envelope detection arose in the analogue domain, with the widespread adoption of radio communications, where the problem is typically restricted to simple sinusoids, often with well defined frequencies \parencite{2011TurnerDemodulation}.
+
+In this context, the most mathematically sound definition of an envelope involves the representation of its underlying wave as an analytic signal, introduced by Gabor in 1946 \parencite{2007HahnHistory}.
+In his work, \textcite{1946GaborTheory} applies the then relatively new mathematical machinery of the quantum mechanics to unify the time and frequency domain representations of a wave, showing how the Hilbert transform could be applied to a real signal in order to obtain an equivalent complex signal, later known as the analytic signal.
+
+This analytic signal has the form $ a(t) = s(t) + \mathbb{H}(s(t)) \ \text{i} $ \parencite{2016HePraat} where $ s(t) $ is the original real signal.
+$ \mathbb{H}(s(t)) $, the Hilbert transform of the original signal, becomes the imaginary part of the analytic signal; the envelope of a signal thus represented can be straightforwardly obtained by the computation of its complex modulus.
+
+Despite its widespread use, and effectiveness in the context of narrowband signals, envelope detection techniques based on the Hilbert transform don't behave well for broadband signals \parencite{2012DauSpeech}, being even physically
+paradoxical in some cases \parencite{1996Loughlinamplitude}.
+
+Figure \ref{fig:AnalyticSignal} exemplifies that, showing the envelope obtained via the analytic signal, for a pure sinusoid with a local frequency of 20 cycles, modulated by a polynomial of the third degree, illustrating the difference in the shape of the envelope in the presence and absence of Gaussian white noise with standard deviation of $ 1/10 $ of the wave's maximum amplitude.
+
+It is readily noticeable from figure \ref{fig:AnalyticSignal} that, as soon as noise is introduced in the original signal, it is reflected in the envelope.
+
+\begin{figure}[ht!]
+  \centering
+    \includegraphics[width=0.8\linewidth]{02AnalyticSignal.pdf}
+  \caption{Envelope of a pure sinusoid modulated by a polynomial of degree 3, as obtained by the Hilbert Transform approach.
+In the first half, the sinusoid is free of noise, while in the second half white gaussian noise with a standard deviation of $ 1/10 $ of the maximum amplitude of the wave was added to the base signal.}
+  \label{fig:AnalyticSignal}
+\end{figure}
+
+In general, the problem of envelope detection, or demodulation, can be interpreted as the task of, given a continuous wave $ W(t) $, decomposing it in two components $ E(t) $ and $ C(t) $, such that $ W(t) = E(t) C(t) $ \parencite{2011TurnerDemodulation}.
+$ E(t) $ represents the slow varying part of the wave, also known as the (temporal) envelope, modulator component, or amplitude modulation (AM) of $ W $, while $ C(t) $ models its fast varying part, called throughout the literature as its (temporal) fine structure, carrier component, or frequency modulation (FM).
+
+In the case of broadband signals, the problem of envelope detection of an arbitrary wave is ill posed, in the sense that an infinite number of pairs of $ E(t) $ and $ C(t) $ can result in the same specific $ W(t) $ \parencite{2011TurnerDemodulation, 2013Mengempirical, 1996Loughlinamplitude}.
+
+We address this problem by assuming $ C(t) $ normalized between $ \{-1, 1\}$.
+This assumption doesn't cause any loss of generality since, given an arbitrary $ C(t) $, we can obtain its normalization by dividing the function by its absolute global maximum, that is $ \hat{C}(t) = C(t) / \max(|C(t)|) $ provided that $ \max(|C(t)|) != 0 $.
+
+Also, in this work we are concerned with the discrete version of this problem: given a finite digital wave represented by the vector $ W \in \mathbb{R}^n $, obtaining the temporal envelope of $ W $.
+The preceding definitions can be translated to a discrete setting assuming that the discrete quantities arise from observing the continuous ones at regular time intervals.
+In that case, the equality $ i = t \ \mathit{fps} $ can be used, where $ i $ is the index of each observation, $ t $ stands for the time in seconds, and $ \mathit{fps} $ is the frame rate, or the number of observations made in the period of a second.
+
+% <|<|<|<|<| TODO We can thus define:
+
 \section{Methodology}
+% Overview of methods
 
-In this work it is assumed, without loss of generality, that one is interested in retrieving the mathematically ill-defined \parencite{2013Mengempirical} temporal envelope of an arbitrary physical sound wave, of which only a digital representation is available.
+ Since $ W $ is a wave, we expect it to be quasi-periodic, as well as uniformly sampled.
 
-Naturally, due to the necessary truncations both in time and amplitude, as well as measurement errors introduced by the instruments used to record the signal, it is not possible to exactly recover the original, continuous wave, from its digital counterpart.
+Instrumental to the method here presented is the definition of a pulse $ P \subset W $: each time $ W $ crosses the horizontal axis, the beginning of a pulse is defined, with the next crossing defining its end.
 
-The best representation one can hope to get of the original wave is, then, the recording itself.
-It is sensible, thus, to characterize the envelope of the wave in terms of a subset of the exact samples that comprise its digital representation.
+This definition is in line with the one presented in \textcite{national1996telecommunications}, where a pulse is defined as
+a rapid change in the amplitude of a signal, followed by a fast return to the baseline value.
 
-To that end, this work will emphasize the concept of frontier, as the points of the digital wave that mark the wave's upper and lower boundaries; from those points an envelope, conforming to the vague definitions of uniqueness and smoothness, can be trivially constructed, as will be shown.
+With a pulse thus defined, we can interpret a wave as composed of a succession of alternating positive and negative pulses.
+
+As such, some structures can be defined, as illustrated in figure \ref{fig:Pulses}.
+
+$ W = \{w_0, w_1, \cdots, w_{n-1}\} $
+
+% <|<|<|<|<|
+
+Half of a pseudo cycle, then, can be defined as the maximum set of adjacent samples with the same sign.
+Two adjacent half pseudo cycles form a full pseudo cycle, with the whole wave being formed by groups of adjacent pseudo cycles.
+The wave in figure \ref{fig:Pulses} if formed by 100 samples, that can be divided in 7 half pseudo cycles, of which two - the leftmost and the rightmost - are malformed, since they are not completely observable in the data provided by $ W $.
+
+With the remaining half pseudo cycles, two full pseudo cycles can be formed.
+Given the quasi-periodicity of the digital wave $ W $, the order in which two adjacent half pseudo cycles are joined to form a full pseudo cycle can be chosen arbitrarily, provided consistency is maintained; in figure \ref{fig:Pulses}, a positive half pseudo cycle is followed by the adjacent negative one, forming a pseudo cycle.
+
+\begin{figure}[ht!]
+  \centering
+    \includegraphics{images/01Pulses.pdf}
+  \caption{Example of a discrete wave.
+The black circles are the individual samples of the signal, representing the intensity of the wave at regular time intervals.
+The grey areas are the half pseudo cycles.
+The dashed lines mark the divisions between adjacent pseudo cycles.}
+  \label{fig:Pulses}
+\end{figure}
+
+From figure \ref{fig:Pulses} and keeping in mind the fact that the variation of the temporal envelope is much slower than the temporal fine structure variation, it is reasonable to suppose that the envelope would ideally pass through the point of maximum amplitude of each half pseudo cycle, following the least sinuous trajectory possible, while still tangencing each half pseudo cycle.
+
+% <|<|<|<|<|
 
@@
 
-  \begin{figure}[ht!]
-    \centering
-      \includegraphics[width=0.8\linewidth]{01Pulses.pdf}
-    \caption{Example of a discrete wave.
-The black stems represent the individual samples of the signal.
-The grey areas encompass the simplified convex hulls of the pulses.
-The dashed lines mark the frontiers between pseudo-cycles.}
-    \label{fig:Pulses}
-  \end{figure}
+Naturally, due to the necessary truncations both in time and amplitude, as well as measurement errors introduced by the instruments used to record the signal, it is not possible to exactly recover the original, continuous wave, from its digital counterpart.
 
-The most mathematically sound definition of an envelope involves the representation of its underlying wave as an analytic signal, introduced by Gabor in 1946 \parencite{2007HahnHistory}.
-In his work, \textcite{1946GaborTheory} applies the then relatively new mathematical machinery of the quantum mechanics to unify the time and frequency representations of a wave, showing how the Hilbert transform could be applied to a real signal in order to obtain an equivalent complex signal, later known as the analytic signal.
+The best representation one can hope to get of the original wave is, then, the digital wave itself.
+It is sensible, thus, to characterize the envelope of the wave in terms of a subset of the exact samples that comprise its digital representation.
 
-This analytic signal has the form $ a(t) = s(t) + \mathbb{H}(s(t)) \ \text{i} $ \parencite{2016HePraat} where $ s(t) $ is the original real signal.
-$ \mathbb{H}(s(t)) $, the Hilbert transform of the original signal, becomes the imaginary part of the analytic signal; the envelope of a signal thus represented can be straightforwardly obtained by the computation of its complex modulus.
+To that end, this work will emphasize the concept of frontier, as the points of the digital wave that mark the wave's upper and lower boundaries; from those points an envelope, conforming to the vague definitions of uniqueness and smoothness, can be trivially constructed, as will be shown.
 
-Figure \ref{fig:AnalyticSignal} shows the envelope obtained via the analytic signal, for a pure sinusoid with a local frequency of 20 cycles, modulated by a polynomial of the third degree, illustrating the difference in the shape of the envelope in the presence and absence of gaussian white noise with standard deviation of $ 1/10 $ of the wave's maximum amplitude.
-
-\begin{figure}[ht!]
-  \centering
-    \includegraphics[width=0.8\linewidth]{02AnalyticSignal.pdf}
-  \caption{Envelope of a pure sinusoid modulated by a polynomial of degree 3, as obtained by the Hilbert Transform.
-In the first half, the sinusoid is free of noise, while in the second half white gaussian noise with a standard deviation of $ 1/10 $ of the maximum amplitude of the wave was added to the base signal.}
-  \label{fig:AnalyticSignal}
-\end{figure}
-
-It is readily noticeable from figure \ref{fig:AnalyticSignal} that, as soon as noise is introduced in the original signal, it is reflected in the envelope, breaking our assumption of smoothness.
-Other than that, this definition yields a unique, positive envelope.
+%
 

### 2020-11-23 Papers/01_Pulses_Envelope/images/00ConvexHull.py

@@ -1,0 +1,118 @@
+import sys
+print(sys.version)
+import plotly.graph_objects as go
+sys.path.append("c:/Users/tesse/Desktop/Files/Dropbox/0_Thesis/Python/01_Envelope")
+import numpy as np
+from scipy.spatial import ConvexHull
+from Helper import signal_to_pulses, get_pulses_area, alpha_shape
+from plotly.subplots import make_subplots
+'''Generating Wave'''
+
+np.random.seed(1)
+# fps = 10
+n = 200+1
+
+X = np.arange(n)# / fps
+W = np.zeros(n)
+
+afp = [
+  [1, 19, np.pi],
+  [1, 29, 1.2 * np.pi],
+  [.6, 9.3, 1.4 * np.pi]
+  ]
+for a, f, p in afp:
+  W += a * np.cos(p + 2 * np.pi * f * X / n) + np.random.normal(0, a / 20, n)
+
+W = W / np.max(np.abs(W))
+
+
+
+points = np.array([[x, w] for x, w in zip(X, W)])
+hull = ConvexHull(points)
+alpha = 300
+alpha_edges = alpha_shape(points, alpha)
+
+
+XX = []
+YY = []
+for e in alpha_edges:
+  idx0 = e[0]
+  idx1 = e[1]
+
+  XX.append(points[idx0][0]),XX.append(points[idx0][0]), XX.append(points[idx1][0]), XX.append(points[idx1][0]), XX.append(None)
+  YY.append(0),              YY.append(points[idx0][1]), YY.append(points[idx1][1]), YY.append(0), YY.append(None)
+
+
+'''============================================================================'''
+'''                              PLOT LINES                                    '''
+'''============================================================================'''
+FONT = dict(
+    family="Latin Modern Roman",
+    color="black",
+    size=13.3333
+  )
+
+'''Plotting'''
+fig = go.Figure()
+fig.layout.template ="plotly_white" 
+fig.update_layout(
+  xaxis_title="<b><i>x</i></b>",
+  yaxis_title="<b><i>y</i></b>",
+  legend=dict(orientation='h', yanchor='top', xanchor='left', y=1.1),
+  margin=dict(l=0, r=0, b=0, t=0),
+  font=FONT,
+  titlefont=FONT
+)
+fig.layout.xaxis.title.font=FONT
+fig.layout.yaxis.title.font=FONT
+
[…]

### 2020-11-30 Papers/01_Pulses_Envelope/PulsesEnvelope.tex
extensive review of the paper

@@
 \usepackage{amsmath}
+
+\usepackage{algorithm}
+\usepackage{algorithmic}
+
+% \usepackage{bm} % bold & upright inside math env
 % \usepackage{hyperref}
@@
 % }
-\title{Robust Digital Envelope Estimation Via Geometric Properties of an Arbitrary Real Signal}
+\title{Envelope Estimation using Geometric Properties of a Discrete Real Signal}
 \author[1]{Carlos Henrique Tarjano Santos}
@@
 \begin{abstract}
+% Background
 Despite being an elusive concept, the temporal amplitude envelope of a signal is essential for its complete characterization, being the primary information-carrying medium in spoken voice and telecommunications, for example.
-Intuitively, the temporal envelope can be understood as a smooth function that multiplies the signal, being responsible for its outer shape.
+Intuitively, the temporal envelope can be understood as a slow varying function that multiplies the signal, being responsible for its outer shape.
 Envelope detection techniques have applications in areas like health, sound classification and synthesis, seismology and speech recognition. 
-Nevertheless, a general method to digital envelope detection of signals with rich spectral content doesn't exist, as most methods involve manual intervention, in the form of filter design, smoothing, as well as other specific design choices, based on a priori knowledge about the nature of the specific waves under investigation.
-To address this problem, we propose a framework that uses intrinsic characteristics of a signal to estimate its envelope, completely eliminating the necessity of parameter tuning.
-The approach here described draws inspiration from geometric concepts to isolate the frontiers and thus estimate the temporal envelope of an arbitrary signal; to that end, alpha-shapes, concave hulls, and discrete curvature are explored.
-We also define entities, such as a pulse and frontiers, in the context of an arbitrary digital signal, as a means to reduce dimensionality and the complexity of the proposed algorithm.
-Specifically, a new measure of discrete curvature is used to obtain the average radius of curvature of a discrete wave, serving as a threshold to identify the wave's frontier points.
-We find the algorithm accurate in the identification of the frontiers of a wide range of digital sound waves with very diverse characteristics, while localizing each pseudo-cycle of the wave in the time domain.
-The algorithm also compares favourably with classic envelope detection techniques based on filtering and the Hilbert Transform.
+% Problem | Objectives
+Nevertheless, a general approach to digital envelope detection of signals with rich spectral content doesn't exist, as most methods involve manual intervention, in the form of filter design, smoothing, as well as other specific design choices, based on a priori knowledge of the nature of the specific waves under investigation.
+To address this problem, we propose a framework that uses intrinsic characteristics of a signal to estimate its envelope, eliminating the necessity of parameter tuning.
+% Methods | Procedure | Approach
+The approach here described draws inspiration from geometric concepts to estimate the temporal envelope of an arbitrary signal; specifically, a new measure of discrete curvature is used to obtain the average radius of curvature of a discrete wave, that will as a threshold to identify the wave's samples that are part of the envelope, in a manner similar to the alpha-shapes approach to the identification of the concave hull of a set of points.
+% Results | Findings | Product
+We provide visualizations of the envelope extracted via the algorithm for various real world signals, with very diverse characteristics, such as voice, spoken and sang, and pitched and non pitched musical instruments, and discuss some approaches to assess the quality of the obtained envelopes.
+The algorithm compares favourably with classic envelope detection techniques based on filtering and the Hilbert Transform.
+A Python module, installable via the Python Package Index (PyPI), with a pure Python and a specialized C++ implementation for 64 bit Windows systems, and interactive visualizations of envelopes for a diverse range of digital waves, as well as the source code for the Python implementation, are available online.
+% Conclusions | Implications
 Besides the most direct applications of this work to audio classification and synthesis, we foresee impact in compression techniques and machine learning approaches to audio.
@@
 
-{\bf Keywords:} DSP, alpha-shapes, envelope detection, discrete curvature estimation
+{\bf Keywords:} DSP, alpha-shapes, envelope detection, discrete curvature estimation, demodulation
 
@@
 
-The problem of envelope detection, normally associated with the field of digital signal processing, can be made equivalent to the geometric problem of defining the (outer) shape of a set of points in $ \mathbb{R}^2 $, provided a method for addressing the difference of units in the horizontal, time related and vertical, amplitude related, axes of a digital signal is available, rendering the original DSP problem readily approachable via geometric techniques.
+The problem of envelope detection, normally associated with the field of digital signal processing, can be made equivalent to the geometric problem of defining the (outer) shape of a set of points in $ \mathbb{R}^2 $, provided a method for addressing the difference of units in the horizontal, time related, and vertical, amplitude related, axes of a digital signal is available, rendering the original DSP problem readily approachable via geometric techniques.
 
@@
 
-% <|<|<|<|<| TODO
+% TODO <|<|<|<|<|
 
@@
 
-\section{Characterization of the temporal envelope}
+\section{Methodology}
+% Overview of methods
+
+\subsection{Characterization of a temporal envelope}
 
@@
 
-This analytic signal has the form $ a(t) = s(t) + \mathbb{H}(s(t)) \ \text{i} $ \parencite{2016HePraat} where $ s(t) $ is the original real signal.
-$ \mathbb{H}(s(t)) $, the Hilbert transform of the original signal, becomes the imaginary part of the analytic signal; the envelope of a signal thus represented can be straightforwardly obtained by the computation of its complex modulus.
+This analytic signal has the form $ A(t) = S(t) + \mathbb{H}(S(t)) \ \text{i} $ \parencite{2016HePraat} where $ S(t) $ is the original real signal.
+$ \mathbb{H}(S(t)) $, the Hilbert transform of the original signal, becomes the imaginary part of the analytic signal; the envelope of a signal thus represented can be straightforwardly obtained by the computation of the complex modulus of $ \mathbb{H}(S(t)) $.
 
@@
   \centering
-    \includegraphics[width=0.8\linewidth]{02AnalyticSignal.pdf}
+    \includegraphics{images/01AnalyticSignal.pdf}
   \caption{Envelope of a pure sinusoid modulated by a polynomial of degree 3, as obtained by the Hilbert Transform approach.
-In the first half, the sinusoid is free of noise, while in the second half white gaussian noise with a standard deviation of $ 1/10 $ of the maximum amplitude of the wave was added to the base signal.}
+In the first half, the sinusoid is free of noise, while in the second half Gaussian noise with a standard deviation of $ 1/10 $ of the maximum amplitude of the wave was added to the base signal.}
   \label{fig:AnalyticSignal}
@@
 We address this problem by assuming $ C(t) $ normalized between $ \{-1, 1\}$.
-This assumption doesn't cause any loss of generality since, given an arbitrary $ C(t) $, we can obtain its normalization by dividing the function by its absolute global maximum, that is $ \hat{C}(t) = C(t) / \max(|C(t)|) $ provided that $ \max(|C(t)|) != 0 $.
+This assumption doesn't cause any loss of generality since, given an arbitrary $ C(t) $, we can obtain its normalization by dividing the function by its absolute global maximum, that is, $ \hat{C}(t) = C(t) / \max(|C(t)|) $, provided that $ \max(|C(t)|) \ne 0 $.
 
-Also, in this work we are concerned with the discrete version of this problem: given a finite digital wave represented by the vector $ W \in \mathbb{R}^n $, obtaining the temporal envelope of $ W $.
-The preceding definitions can be translated to a discrete setting assuming that the discrete quantities arise from observing the continuous ones at regular time intervals.
-In that case, the equality $ i = t \ \mathit{fps} $ can be used, where $ i $ is the index of each observation, $ t $ stands for the time in seconds, and $ \mathit{fps} $ is the frame rate, or the number of observations made in the period of a second.
+Also, in this work we are concerned with the discrete version of this problem: given a finite digital wave represented by the vector $ \textbf{w} \in \mathbb{R}^n $, instead of the continuous function $ W(t) $, obtaining the temporal envelope of $ \textbf{w} $.
 
-% <|<|<|<|<| TODO We can thus define:
+The preceding definitions can be translated to this discrete scenario assuming that the discrete quantities arise from observing the continuous ones at regular time intervals.
+In that case, the equality $ i = t \ \text{fps} $ can be used to link both settings, where $ i $ is the index of each observation, $ t $ stands for the time in seconds, and fps is the frame rate, or the number of observations made in the period of a second.
+We can thus define:
 
-\section{Methodology}
-% Overview of methods
+\begin{equation} \label{eq:Envelope}
+  \begin{align}
+    & \textbf{w}, \textbf{e}, \textbf{c} \in \mathbb{R}^n \\
+    & \textbf{w} = (w_0, w_1, \cdots, w_{n-1}) \\
+    & \textbf{e} = (e_0, e_1, \cdots, e_{n-1}) \\
+    & \textbf{c} = (c_0, c_1, \cdots, c_{n-1}) \\
+    & \textbf{w} = \textbf{e} \odot \textbf{c} \\
+    & w_i = e_i c_i \quad \forall \quad i, 0 \le i \le n-1
+  \end{align}
+\end{equation}
 
- Since $ W $ is a wave, we expect it to be quasi-periodic, as well as uniformly sampled.
+The $ \odot $ operator in \ref{eq:Envelope} stands for the Hadamard product, denoting element wise multiplication of two vectors.
+Figure \ref{fig:Envelope} provides an example of the vectors just defined: \textbf{w} is obtained by modulating the normalized carrier wave \textbf{c} using the smooth envelope \textbf{e}.
 
-Instrumental to the method here presented is the definition of a pulse $ P \subset W $: each time $ W $ crosses the horizontal axis, the beginning of a pulse is defined, with the next crossing defining its end.
+Note from Figure \ref{fig:Envelope} that, at some local maxima, the envelope is defined, being in fact equal to the value of the wave \textbf{w}, since $ c_i = 1 $ at those points, rendering $ e_i = 1 w_i $.
+The carrier wave \textbf{c} is periodic, and normalized between $ \{-1, 1\}$.
+Each cycle of the wave, then, reaches the value 1 only at its maximum making $ w_i$ equal to the value of the envelope at that point.
 
-This definition is in line with the one presented in \textcite{national1996telecommunications}, where a pulse is defined as
-a rapid change in the amplitude of a signal, followed by a fast return to the baseline value.
+% TODO <|<|<|<|<| we will assume that this assumption holds for quasi periodic signals: that is, the absolute maximum of each pseudo cycle is equal to the value of the envelope at that point, the other local extrema in that same cycle are the ones that should be filtered out.
 
-With a pulse thus defined, we can interpret a wave as composed of a succession of alternating positive and negative pulses.
+% TODO <|<|<|<|<| We lack a indicator to evaluate the quality of the envelope, once one is identified, but figure suggests that, once the envelope is removed from w, the wave should then be confined between the interval $ \{-1, 1\}$.
 
-As such, some structures can be defined, as illustrated in figure \ref{fig:Pulses}.
+Image \ref{fig:Envelope}, along with the Concave Hull illustrated in figure \ref{fig:Hulls}, helps forming the intuition behind the method developed in this work: the idea is to identify the local extrema that touch the envelope (or, to put in another way, the local extrema in relation to the whole wave that are global extrema inside their own pseudo cycles), filtering out those who don't; in this way, we're defining the envelope itself at those points.
 
-$ W = \{w_0, w_1, \cdots, w_{n-1}\} $
+In this work, this is accomplished with the use of a circle of a defined radius, large enough to prevent it from reaching the smallest local extrema when "rolled" around the (geometric) representation of the wave: only the extrema touched by the circle will be marked as part of the envelope.
 
-% <|<|<|<|<|
+In the next section such a geometric representation will be formulated, along with the necessary tool to estimate the average curvature of the wave and, consequently, the appropriate radius of the circle.
 
-Half of a pseudo cycle, then, can be defined as the maximum set of adjacent samples with the same sign.
-Two adjacent half pseudo cycles form a full pseudo cycle, with the whole wave being formed by groups of adjacent pseudo cycles.
-The wave in figure \ref{fig:Pulses} if formed by 100 samples, that can be divided in 7 half pseudo cycles, of which two - the leftmost and the rightmost - are malformed, since they are not completely observable in the data provided by $ W $.
-
-With the remaining half pseudo cycles, two full pseudo cycles can be formed.
-Given the quasi-periodicity of the digital wave $ W $, the order in which two adjacent half pseudo cycles are joined to form a full pseudo cycle can be chosen arbitrarily, provided consistency is maintained; in figure \ref{fig:Pulses}, a positive half pseudo cycle is followed by the adjacent negative one, forming a pseudo cycle.
+% TODO <|<|<|<|<|
 
@@
   \centering
-    \includegraphics{images/01Pulses.pdf}
-  \caption{Example of a discrete wave.
-The black circles are the individual samples of the signal, representing the intensity of the wave at regular time intervals.
-The grey areas are the half pseudo cycles.
-The dashed lines mark the divisions between adjacent pseudo cycles.}
+    \includegraphics{images/02Envelope.pdf}
+  \caption{Example of a discrete wave \textbf{w} composed by the element wise multiplication of an envelope \textbf{e} and a carrier \textbf{c}.
+The local extrema of \textbf{w} are highlighted with a circle.}
+  \label{fig:Envelope}
+\end{figure}
+
+\subsection{Simplified Representation}
+
+Instrumental to the method here presented is the definition of a pulse: each time the value of \textbf{w} changes sign, that is, every time the discrete wave \textbf{w} represents crosses the horizontal axis, the beginning of a pulse is defined, with the next crossing defining its end.
+This definition is in line with the one presented in \textcite{national1996telecommunications}, where a pulse is defined as a rapid change in the amplitude of a signal, followed by a fast return to the baseline value; zero, in our case.
+
+From \ref{fig:Envelope} we can see that only the maximum point of a positive pulse or the minimum point of a negative one can potentially be equal to the envelope and are, thus, our only points of interest.
+We can then proceed with the rest of the method considering only those points, to great computational advantage.
+
+For this reason we define P as the set of the points $ \{P_0, P_1, \cdots, P_{m-1}\} $ where $ P_j = (i_j, \lvert w_j \lvert) $, that is, $ \lvert w_j \lvert $, the absolute value of the extrema of each pulse, becomes the ordinate of each point and $ i_j $, its original index, becomes the abscissa.
+
+This relation is illustrated in figure \ref{fig:Pulses}. We call it P to emphasize the connection with the pulses of \textbf{w}, noting that P is, in fact, a set of points.
+
+\begin{figure}[ht!]
+  \centering
+    \includegraphics{images/03Pulses.pdf}
+  \caption{Example of a discrete wave \textbf{w} divided into pulses, of which the extrema are highlighted with a circle.
+P is the set of the points $ P_j = (i_j, \lvert w_j \lvert) $ representing the absolute value of those extrema.}
   \label{fig:Pulses}
@@
 
-From figure \ref{fig:Pulses} and keeping in mind the fact that the variation of the temporal envelope is much slower than the temporal fine structure variation, it is reasonable to suppose that the envelope would ideally pass through the point of maximum amplitude of each half pseudo cycle, following the least sinuous trajectory possible, while still tangencing each half pseudo cycle.
+P is not fit for geometric interpretation yet, since the abscissa and ordinate of his orthogonal coordinate system have different units: in the vertical axis we have a unit related to the instantaneous intensity of the wave, while in the horizontal axis we have the index at which each extremum occurs, ultimately a time related unit.
 
-% <|<|<|<|<|
+Specifically, this coordinate system has a basis vector $ \mathbf{b} = \{ \mathbf{b}_x, \mathbf{b}_y \}$ with $ \mathbf{b}_x = (a, 0), \mathbf{b}_y = (0, i) $ where $ a $ is an amplitude unit such as decibel, a measure of the instantaneous sound pressure, volt or even a fraction of a fixed maximum value.
+$ i $ represents the index of the sample, and is linked to time by the aforementioned relation $ i = t \ \text{fps} $.
 
-We start by defining a pulse as a series of consecutive samples of a discrete wave that have the same sign.
-More formally, let $ W[i] \in \mathbb{R} \ \forall \ i \in \mathbb{N}_0; \ i < n \in \mathbb{N}_0 $ be a real, discrete and finite signal indexed, without loss of generality, over an adjacent subset of the natural numbers starting at zero.
-This definition relates closely to the concept of an array or vector in most programming languages, and is used in the interest of simplicity.
+We are interested in finding a coordinate system where a pulse's amplitude $ \lvert w_j \lvert $ and length would define, in average, a square.
 
-A pulse $ P[j]; \ \forall \ j \in \mathbb{N}_0; \ j < q \in \mathbb{N}_0; q \le n $ is, then, a subset of $ W $ with cardinality $ \#P[j] = b - a $ such that $ \sign(W[a-1]) \ne \sign(W[a]) = \sign(W[a+1]) = \dots = \sign(W[b-1]) = \sign(W[b]) \ne \sign(W[b+1]) \ \forall \ i; \ a \le i \le b; \ a, b \in \mathbb{N}_0 $.
+To achieve that, we divide the basis vector of the original vertical axis by the average of all ordinates, effectively cancelling the unit of the vertical components of the points.
+We could divide the basis vector of the horizontal axis by the average of the difference between the abscissa of a point and the abscissa of the immediately posterior, likewise eliminating the axis' unit.
 
-In other words, each pulse is a unique subset of $ W $ with its members being the maximum sequence of samples of $ W $ having the same signal.
-Thus $ P[0] \cup P[1] \cup \cdots \cup P[q-1] = W $ and $ P[i_1] \cap P[i_2] = \emptyset \ \forall \ i_1, i_2 \in \mathbb{N}_0;\ i_1 \ne i_2; \ i_1, i_2 < q $.
+The drawback is that we would lose the direct link between P and \textbf{w} that arises from the fact that the abscissas of the points in P are indices $ i $ of \textbf{w}: we choose instead to leave the abscissa intact and multiply the now unitless vertical basis axis by the average of the difference between the abscissa of a point and the abscissa of the immediately posterior.
+In this way, both axes will have unit $ i $, and the relation is preserved, making it easier to recover the envelope points later.
 
-Naturally, each pulse can be classified as positive ($ P[j]^+ $) or negative ($ P[j]^- $), based on the sign of its samples, which must be the same.
-It is also convenient to attribute to each pulse a pair of coordinates: those will be the coordinates of the sample with the maximum absolute amplitude in each pulse, and will be denoted by $ P[j]^\pm_x $ and $ P[j]^\pm_y $.
-Additionally, the length, or cardinality, of a pulse will be denoted by $ \#P[j]^\pm $.
+In this new coordinate system we have a basis vector $ \mathbf{b}^\prime = \{ \mathbf{b}^\prime_x, \mathbf{b}^\prime_y \}$ where $ \mathbf{b}^\prime_x = (i, 0) $ and $\mathbf{b}^\prime_y = (0, i) $, related to the original basis vector by:
 
-For an arbitrary continuous function, then, a pulse would correspond to the regions between its roots; for a sinusoid, a pulse would be equal to half of its cycle.
-In analogy, two consecutive pulses, naturally with opposing signs, have the potential to delimitate a pseudo-cycle, as illustrated in figure \ref{fig:Pulses}.
-For that to be true, however, the pulses need to make part of the frontiers of the signal, as yet to be defined.
+\begin{equation} \label{eq:Basis}
+\begin{align}
+\mathbf{b}^\prime_x &= \mathbf{b}_x  \\
+\mathbf{b}^\prime_y &= \frac{\left( \frac{i_{m-1} - i_0}{m-1} \right)}{\left( \frac{\sum_{j=0}^{m-1} \lvert w_j \lvert}{m} \right)} \mathbf{b}_y 
+\end{align}
+\end{equation}
 
-Naturally, due to the necessary truncations both in time and amplitude, as well as measurement errors introduced by the instruments used to record the signal, it is not possible to exactly recover the original, continuous wave, from its digital counterpart.
+The effect of this normalization can be seen in the points shown in figure \ref{fig:Vectors}, shown in scale in both coordinate systems.
+The set V of the vectors from each point of P to the next is also shown in the picture, as it will be used to estimate the average curvature of P, we define this set below.
 
-The best representation one can hope to get of the original wave is, then, the digital wave itself.
-It is sensible, thus, to characterize the envelope of the wave in terms of a subset of the exact samples that comprise its digital representation.
-
-To that end, this work will emphasize the concept of frontier, as the points of the digital wave that mark the wave's upper and lower boundaries; from those points an envelope, conforming to the vague definitions of uniqueness and smoothness, can be trivially constructed, as will be shown.
-
-%
-
-In practice, however, is not uncommon for a discrete wave, especially in the case of sound, to present somewhat different positive and negative contours; to account for this discrepancy, we will define a positive (upper) and a negative (lower) frontier of a discrete wave.
-They will be composed by the points $ (i, W[i]) $ that are, necessarily but not sufficiently, the points of respectively maximum and minimum amplitude of their pulses, i.e.
-$ (P[j]^\pm_x, P[j]^\pm_y) $.
-
-The set of all these points such that $ W[i] > 0 $ will be called the positive, piecewise linear frontier and will be denoted by $ F^+ $.
-Similarly, $ F^- $ will denote the negative frontier of $ W $, defined by the subset of points $ (i, W[i]) $ where $ W[i] < 0 $, as is shown in figure \ref{fig:Frontiers}.
+\begin{equation} \label{eq:Vectors}
+\begin{align}
+\text{V} = \{ v_0, v_1, v_{m-2} \} \quad \text{where} \quad v_k = P_{j+1} - P_j \quad \forall \quad 0 \le k \le m-2
+\end{align}
+\end{equation}
 
@@
   \centering
-    \includegraphics[width=0.8\linewidth]{03Frontiers.pdf}
-  \caption{Positive and negative frontiers, envelope and pseudo-cycles for a fragment of piano sound.
-While we have only one (positive) envelope, here mirrored in the inferior space, we have two frontiers, the superior and inferior, that are obtained independently.}
-  \label{fig:Frontiers}
+    \includegraphics{images/04Vectors.pdf}
+  \caption{The set of points P and the set V of vectors between two adjacent points of P, in both the original and normalized coordinate systems, shown in scale.}
+  \label{fig:Vectors}
 \end{figure}
 
-To identify such points, we need first to address the different units in which the amplitude and horizontal distance of our wave are expressed.
-To make abscissa and ordinate units consistent for the computation of the frontiers, and proceed with a geometric treatment of the subject, we consider the set of positive and negative pulses separately.
+\subsection{Discrete Curvature Estimation}
 
-For each set, we multiply all pulses vertical coordinates by the average length of the pulses in the set, after having divided by the average of the pulses' amplitudes, putting both axes in the same unit $i$, related to time by the equation $i = t \ \mathit{fps}$.
+Translating the explanation of alpha-shapes in \textcite{1994EdelsbrunnerThree} to our context, we can say that the points that are part of the envelope are those touched by a circle of a given radius $ r $, coming from $ y \to \infty $ towards the signal, that is not allowed to contain any point of the signal.
+% TODO refine
 
-Formally, supposing we have $ m $ pulses of the same sign, we scale each pulse's vertical coordinate $ P[j]^\pm_y $ by $ s^\pm $, where:
+Intuitively, one can picture a circle being rolled above the signal, and marking the points it touches as envelope points.
 
-\begin{equation} \label{eq:scale}
-  s^\pm = \frac{\sum_{j=1}^{m} \#P[j]^\pm}{\sum_{j=1}^{m} P[j]^\pm_y}
-\end{equation}
-
-Because we are interested in approximating a function that is smooth over the whole domain of the wave, we can assume that, in the scale of a pulse, this function will be very close to a horizontal straight line and, thus, won't touch any point inside the convex hull defined by the pulse.
-
-By the same token, as the change in amplitude from one pulse to the next pulse with the same sign is motivated by changes in this same envelope, one can reasonably suppose that $ \max(P[j]) \approx \max(P[j+2]) $.
-
-This makes clear that only one point of an arbitrary pulse can be in the frontier, namely, its point of maximum amplitude; those definitions will allow us to derive the frontiers of a wave using an alpha-shapes inspired algorithm without the need to compute the Delaunay triangulation first.
-
-That also justifies the approximation of the convex hull of a pulse by the triangle defined by the points where the pulse crosses the horizontal axis and the point of maximum amplitude $ (P[j]^\pm_x, P[j]^\pm_y) $; this triangular pulse and, more specifically, its point of maximum, will become the atomic unit for the rest of the method.
-
-Translating the intuitive explanation of alpha-shapes in \textcite{1994EdelsbrunnerThree} to our context, we can say that the points in the frontier are those touched by a circle of a given radius $ r $ outside the signal that is not allowed to contain any point of the signal; the pulses that contain said points can be seen as frontier pulses.
-
-Intuitively, one can picture a circle being rolled above (or below, in the case of the negative frontier) the signal, and marking the points it touches as frontier points.
-
-We can now revisit the concept of a pseudo-cycle, only hinted at before, defining it as the samples of the original wave that belong to two successive pulses that, in turn, belong to their respective frontiers, including eventual pulses between them, as can be seen in figure \ref{fig:Frontiers}.
-
-It is necessary to infer the appropriate radius $ r $ of such a circle and, to that end, a measure of the curvature of a discrete function is needed.
+To infer the appropriate radius $ r $ of such circle a measure of discrete curvature is needed.
 Discrete curvature estimation is an important task in image processing \parencite{2010FleischmannNovel} for which no default definition exists.
 
-The two possible general approaches are the derivation of direct methods that use characteristics of the discrete wave to calculate the curvature, or the use of the smooth curvature of a curve fitted to the discrete wave \parencite{2001CoeurjollyDiscrete}.
+The two possible general approaches are the derivation of direct methods, that use characteristics of the discrete wave to calculate the curvature, or the use of the curvature of a smooth, continuous curve fitted to the discrete wave \parencite{2001CoeurjollyDiscrete}.
 
-Concerning the last method, two approaches to fit a polynomial to a generic wave are readily available: the least mean squares approach, that seeks to minimize the dependent variable errors and the total least squares problem, that treats both variables symmetrically \parencite{1980GolubAnalysis}.
+% Concerning the last method, two approaches to fit a polynomial to a generic wave are readily available: the least mean squares approach, that seeks to minimize the dependent variable errors and the total least squares problem, that treats both variables symmetrically \parencite{1980GolubAnalysis}.
 
-The geometric, symmetrical nature of the problem excludes the more computationally economic least mean squares approach, however, leaving us with the total least squares, for which no general closed-form solutions are available \parencite{2007MarkovskyOverview}; this leads us to turn our attention to direct approaches of curvature estimation.
+% The geometric, symmetrical nature of the problem excludes the more computationally economic least mean squares approach, however, leaving us with the total least squares, for which no general closed-form solutions are available \parencite{2007MarkovskyOverview}; this leads us to turn our attention to direct approaches of curvature estimation.
 
-\subsection{Discrete Curvature Estimation - The Equivalent Circle Approach}
+Concerning direct methods \textcite{2014CarrollSurvey}, for example, derive three such definitions based on the approximation of a circle by an inscribed, centred and circumscribed polygon.
+In the context of tridimentional meshes, \textcite{2016VasaMultivariate} evaluate a range of existing estimators from a multivariate point of view.
 
-As previously noted, no default discrete curvature definition exists; \textcite{2014CarrollSurvey}, for example, derive three such different definitions based on the approximation of a circle by an inscribed, centred and circumscribed polygon.
-In the context of 3d meshes, \textcite{2016VasaMultivariate} evaluate a range of existing estimators from a multivariate point of view.
+Those approaches define the discrete curvature in relation to the vertices of a wave.
+In our particular case, it wouldn't make sense to talk about the curvature of a single pulse, since it is what our points ultimately represent.
+Instead, we are interested in the change of direction, what the term curvature ultimately means, between two adjacent pulses, as the vectors in the set V suggests.
 
-None of those approaches, especially for being defined over the vertices of a discrete wave, fit seamlessly into our algorithm; we thus proceed to define a discrete curvature measure over the edges of a discrete wave.
+We thus proceed to define a discrete curvature measure over the edges of a discrete wave.
 To that end we are going to apply the definition of smooth curvature as the rate of change of the unit tangent to a curve, noting that this is equivalent to that of the osculating circle \parencite{2016VasaMesh}.
 
-The idea is to find the equivalent circle whose tangent presents the same change in direction, in the same horizontal distance, as the edge of interest, as shown in figure \ref{fig:DiscreteCurvature}.
+\subsubsection{The Equivalent Circle Approach}
+
+The rationale is to find, for each vector $ v_k \in \text{V} $, the radius $ r_k $ of the equivalent circle whose tangent has the same change in direction, in the same horizontal distance, as the vector of interest, as shown in figure \ref{fig:DiscreteCurvature}.
+The average curvature of the P will then be obtained by the average of all radii $ r_k $.
 
@@
   \centering
-    \includegraphics[width=0.2\linewidth]{04DiscreteCurvature.pdf}
-  \caption{The tangent unit vector of the circle changes from the horizontal direction in $ \hat{u}_0 $ to an inclination of $ \theta $ in $ \hat{u}_1 $.
-$ \theta $ is also the angle between the average direction $ \vec{\overline{v}}_l $ and the vector $ \vec{v}_l $ that represents the edge of interest, shown in grey.}
+    \includegraphics{images/05DiscreteCurvature.pdf}
+  \caption{The tangent unit vector of the circle changes from the horizontal direction in $ \hat{u}_0 $ to an inclination of $ \theta $ in $ \hat{u}_1 $, $ \theta $ being the angle that the vector $ v_k $ makes with the horizontal direction.}
   \label{fig:DiscreteCurvature}
@@
 
-For each frontier, the angle $ \theta $ between the average direction and each of the vectors can be obtained via the relation between the slopes $ d $ of the respective lines, as in the expression $ \theta = \arctan( (d - \overline{d}) / (1 + d \overline{d}) ) $, with $\overline{d} = \sum_{l=1}^{m-1} (P[l+1]_y - P[l]_y) / (P[l+1]_x - P[l]_x) $ and $ d_l = (P[l+1]_y - P[l]_y)/ (P[l+1]_x - P[l]_x) $
+From figure \ref{fig:DiscreteCurvature} is easy to see that $ r_k = v_{k,x} \sin(\theta_k)$ where $ v_{k,x} $ is the component of $ v_k $ in the horizontal direction; $ \theta_k $ itself can be obtained using the slope $ m_0 = 0 $ of any line in the horizontal direction and the slope $ m_k = v_{k,y} / v_{k,x} $ defined by $ v_k $ via the equality $ \theta_j = \arctan \big( (m_k - m_0) / (1 + m_k m_0) \big) $.
 
-The radii of the equivalent circles are thus given by $ r_l = x_l / \sin (\theta_l) $ with the curvature given by $\kappa_l = 1/r_l=\sin (\theta_l) / x_l$, where $ x_l = P[l+1]_x - P[l]_x $ and $ y_l = P[l+1]_y - P[l]_y $.
-Joining both equations, one can write:
+The radius of the circle that represents the average curvature of V, and will be used to obtain the envelope of the discrete wave \textbf{w} can be obtained via equation \ref{eq:Radius}:
 
-\begin{equation} \label{eq:radius}
-  r_l = \left\lvert \frac{x_l \sqrt{\overline{d}^2 + 1} \ \sqrt{x_l^2 + y_l^2}}{\overline{d} x_l - y_l} \right\rvert
+\begin{equation} \label{eq:Radius}
+% \begin{gather}
+r = \frac{\sum\limits_{k=0}^{m-2} \left( \frac{v_{k,x} \sqrt{v_{k,x}^2 + v_{k,y}^2}}{ v_{k,y}} \right)}{m-2}
+% \end{gather}
 \end{equation}
 
-After this procedure, illustrated in figure \ref{fig:Curvatures}, two sets of points in $ \mathbb{R}^2 $, corresponding to the positive and negative frontiers, are obtained, where the ordinates represent the instantaneous radius, with its localization in the time domain given by the abscissas.
+We need now to construct an algorithm, using the radius just obtained, to identify the points that belong to the envelope.
+Generally, the algorithm for the alpha-shapes approach resorts to the construction of the Delaunay triangulation of the set of points, that is latter filtered.
+We can adopt a more straightforward method, as our points have a defined structure:
+
+Let $ x_o, y_o \in \mathbb{R}, \quad 0 \le y_o < +\infty $ be the coordinates of the center of a circle with radius $ r $, that can be placed anywhere in the upper plane of the coordinate system;
+
+Let $ P_a, P_b, P_c \in \text{P}, \quad P_a \ne P_b \ne P_c $ be three different points of the set P;
+
+For all circles of center $ x_o, y_o $ and radius $ r $ that can be constructed to pass through any two different points of P, that is, $ \forall \quad P_b, P_c $ such that $ (P_{b,x} - x_o)^2 + (P_{b,y} - y_o)^2 = (P_{c,x} - x_o)^2 + (P_{c,y} - y_o)^2 = r^2 $, $ P_a $ will be part of the set $ E $ of the points belonging to the envelope if and only if none of those circles contain $ P_a $, or $ P_a \in \text{E} \leftrightarrow (P_{b,x} - x_o)^2 + (P_{b,y} - y_o)^2 > r^2 $.
+Formalizing, we have:
+
+\begin{equation} \label{eq:InEnvelope}
+\begin{align}
+& P_a \in \text{E} \leftrightarrow (P_{b,x} - x_o)^2 + (P_{b,y} - y_o)^2 > r^2, \\
+& \forall \ P_b, P_c \mid (P_{b,x} - x_o)^2 + (P_{b,y} - y_o)^2 = (P_{c,x} - x_o)^2 + (P_{c,y} - y_o)^2 = r^2, \\
+& P_a, P_b, P_c \in \text{P}, \quad P_a \ne P_b \ne P_c \quad \text{and} \quad x_o, y_o \in \mathbb{R}, \quad 0 \le y_o < +\infty
+\end{align}
+\end{equation}
+
+The algorithm follows directly from the definition in \ref{eq:InEnvelope} after noting that the first and last members of P will always be part of the envelope, since they are reachable a circle of any radius $ r $.
+From $ P_0 $, our first pivot point, we construct circles passing through $ P_1, P_2, \cdots, P_d $ until a circle that doesn't contain any point in P is found; $ P_d $ then becomes the new pivot point and is included in E, and this procedure continues until $ P_{m-1} $ is reached.
+
+The procedure is formalized in algorithm \ref{FindEnvelope}. We don't provide similar algorithmic descriptions of the preceding steps since the Python implementation can easily serve as pseudocode.
+
+\begin{algorithm}[H]
+\caption{Retrieve Envelope} \label{FindEnvelope}
+\begin{algorithmic}
+\STATE \textbf{Given} $ \text{P} = \{P_0, P_1, \cdots, P_{m-1}\}, r $, \textbf{circle}, a function that returns the center of a circle of a given radius passing through two points and \textbf{distance}, a function that returns the Euclidean distance between two points,
+\STATE \textbf{Let} $ \text{E} = \{ P_0 \}, \text{id1} = 0, \text{id2} = 1 $
+\STATE \textbf{Define} $ P_c = \text{point}, \text{empty} = \text{boolean} $
+\WHILE {$ \text{id2} < m $}
+  \STATE $ P_c \leftarrow \textbf{circle}(r, P_\text{id1}, P_\text{id2}) $
+  \STATE empty $ \leftarrow $ \TRUE 
+  \FOR{$ i = \text{id2} + 1 $ \TO $ m $}
+    \IF {$ \textbf{distance}(P_c, P_i) < r $}
+      \STATE empty $ \leftarrow $ \FALSE
+      \STATE id2 $ \leftarrow \text{id2} + 1 $
+      \STATE \textbf{break}
+    \ENDIF
+  \ENDFOR
+  \IF {empty}
+    \STATE E $ \leftarrow $ E $ \cup \ \{ P_\text{id2} \} $
+    \STATE id1 $ \leftarrow $ id2
+    \STATE id2 $ \leftarrow \text{id2} + 1 $
+  \ENDIF
+\ENDWHILE
+\end{algorithmic}
+\end{algorithm}
+
+The set E obtained via algorithm \ref{FindEnvelope} is a set of points.
+The most straightforward way to transform those points into the vector $ \textbf{e} \in \mathbb{R}^n $ is via linear interpolation, the approach used in this work.
+is worthy noting, however, that many other procedures are available, both for the interpolation and the smooth approximation of points.
+We deliberately choose linear interpolation to avoid blurrying the results of the method.
+
+\section{Results}
+% Revisit purpose and methods
+% Overview of results
+
+\subsection{Quality of an envelope}
+
+There are no agreed upon objective metrics to assess the quality of an envelope.
+Recalling figure \ref{fig:Envelope}, however, and the accompanying discussion, an intuitive way is to consider the carrier wave \textbf{c}, and how well it seems to be bounded in amplitude in the interval $ \{-1, 1\} $.
+To that end, with provide the follow figures of original discrete waves \textbf{w}, their envelope \textbf{e}, and the inferred carrier, obtained after dividing, element wise, the original wave by its envelope.
 
@@
   \centering
-    \includegraphics[width=0.8\linewidth]{05Curvatures.pdf}
-  \caption{The equivalent circles obtained via the procedure described for the positive frontier.
-$\theta$, the angle between $ \vec{\overline{v}}^+$ and each of the $\vec{v}_l^+$ is used, as is the horizontal component of each $\vec{v}_l^+$ to obtain the radii of the circles.}
-  \label{fig:Curvatures}
+    \includegraphics{images/06Graph_bend.pdf}
+  \caption{Signal, envelope and carrier for a record of a bend performed on an electric guitar.}
+  \label{fig:Bend}
 \end{figure}
 
-Given the robustness of the method, for an average wave of a few seconds at an ordinary rate of 44100 $ \mathit{fps} $ the average of the radius values is sufficient to assure the accurate identification of the frontier.
-However, a curve fitting method can be used in the case of longer or ill-behaved waves.
-An example of the use of the average equivalent circle can be seen in figure \ref{fig:AvgCircle}.
+Due to an assortment of factors, such as the unconventional approach presented in this work and the intrinsic dificult in visualizing discrete waves given their high dimensionality, the authors deemed helpful to provide a companion website with interactive visualizations.
 
-\begin{figure}[ht!]
-  \centering
-    \includegraphics[width=0.8\linewidth]{06AvgCircle.pdf}
-  \caption{The positive frontier as obtained by the use of the circle representing the average curvature of the discrete wave illustrated.}
-  \label{fig:AvgCircle}
-\end{figure}
+An implementation of the method in pure Python, fully functional, can also be found at the repository dedicated to this work as well as a Python module available at the Python Package Index (PyPI) and conveniently instalabbe via the command `pip install signal-envelope`.
+Besides, those working under Windows 64 bit can benefit from a specialized dll coded in C++ that is automatically used within this module.
 
-\section{Results}
+In an effort to formalize the assessment of the quality of temporal envelopes, \textcite{1996Loughlinamplitude} proposed some conditions necessary, but not sufficient, to ensure the physical plausibility of an envelope.
+We comment briefly below that the algorithm here presented satisfies the four conditions presented.
 
-Figure \ref{fig:FullFrontiers} illustrates the frontiers of six diverse discrete sound waves, as well as an in-detail view of the highlighted segment for each signal.
-All waves are records of physical sounds, chosen to represent the applicability of the algorithm in real-world scenarios.
+The first one states that if a signal is bounded in magnitude, then its envelope should be, as well.
+Our algorithm complies with this requisite, as the envelope is composed of selected samples from the original signal itself;
 
-\begin{figure}[ht!]
-  \centering
-    \includegraphics[width=0.8\linewidth]{07FullFrontiers.pdf}
-  \caption{Positive and negative frontiers of six digital waves, as extracted by the algorithm here presented.
-For each wave, the region highlighted in black is shown in detail besides the whole wave.
-For each wave the horizontal axis is the sample number $i$, while the vertical axis is the normalized amplitude.
-All 6 waves were recorded at 44100 $ \mathit{fps} $.}
-  \label{fig:FullFrontiers}
-\end{figure}
+The second statement is that if the signal has a finite frequency range, that frequency range must not be exceeded by the envelope.
+This is only directly applicable in the continuous case, where this condition was proposed.
 
-We can see that the frontier is satisfactorily detected in the case of harmonic and inharmonic sounds, and is robust in relation to the number of samples and the frequencies of the waves.
+Nevertheless, if we picture both the wave and the envelope as samples from continuous functions, is easy to see that this t is the case for our algorithm; since the envelope is composed by some samples of the digital wave, in the limit where all the samples of the wave are also part of the envelope, we have that the maximum frequency for both is the same.
 
-The impact of the normalization of a wave using the frontiers obtained can be seen in figure \ref{fig:Fourier}, which shows the original and normalized wave in the case of a recording of the key 33 of a piano both in the time and frequency domains.
-We can see that the Fourier power spectra is greatly simplified in the case of the normalized wave.
+In all other cases, the maximum frequency of the envelope will be smaller than that of the original wave.
+The range of frequencies of the envelope, then, will always lie between zero and the maximum frequency of the original wave.
 
-\begin{figure}[ht!]
-  \centering
-    \includegraphics[width=0.8\linewidth]{08Fourier.pdf}
-  \caption{Original and normalized views of a recording of key 33 (F3, 175 Hz) from a grand piano both in time and frequency domains.}
-  \label{fig:Fourier}
-\end{figure}
+The third condition states what, for a periodic signal, the envelope should be a straight line with the intercept equal to the amplitude of this signal.
+Since all the local maxima have the same amplitude, our envelope will indeed be a straight line, with the same amplitude as the original wave.
 
-\subsection{Robustness}
-
-Figures \ref{fig:xRobustness} and \ref{fig:yRobustness} illustrate the computational resilience of the algorithm for scaling in both the abscissa and ordinate axes, respectively.
-This is due to the normalization scheme presented in equation \ref{eq:scale}.
-
-Although important from a theoretical standpoint, this step has an especial computational relevance in the case of digital sound waves, as they tend to be composed of a high number of samples with relatively low amplitudes, generally normalized between -1 and 1, leading to a numerical disparity between axes.
-
-\begin{figure}[ht!]
-  \centering
-    \includegraphics[width=0.8\linewidth]{10xRobustness.pdf}
-  \caption{Minimum and maximum scales of the abscissa before computational stability issues arose, for a recording of a drum kit tom.  $ \text{B} = 10^6 $.}
-  \label{fig:xRobustness}
-\end{figure}
-
-\begin{figure}[ht!]
-  \centering
-    \includegraphics[width=0.8\linewidth]{09yRobustness.pdf}
-  \caption{Minimum and maximum values of the amplitude before computational stability issues arose, for a recording of a drum kit tom.}
-  \label{fig:yRobustness}
-\end{figure}
+The fourth and last condition states that if a wave is multiplied by a constant, the envelope should also be multiplied by the same constant.
+That is also the case, again because the points that form the envelope are points of the original wave itself.
 
@@
 
-\subsection{Retrieving information about the underlying wave}
+% \subsection{Retrieving information about the underlying wave}
 
-The algorithm provides the position of the pseudo-cycles of a wave, with which one can, for example, derive its fundamental frequency and phase: using one of the two frontiers, we have $ f = (\#F^\pm - 1) n / (F^\pm[-1] - F^\pm[0])$, where $ \#F^\pm $ denotes the cardinality of the frontier, while $ F^\pm[0] $ and $ F^\pm[-1] $ stand for the frontier's first and last item, respectively.
+% The algorithm provides the position of the pseudo-cycles of a wave, with which one can, for example, derive its fundamental frequency and phase: using one of the two frontiers, we have $ f = (\#F^\pm - 1) n / (F^\pm[-1] - F^\pm[0])$, where $ \#F^\pm $ denotes the cardinality of the frontier, while $ F^\pm[0] $ and $ F^\pm[-1] $ stand for the frontier's first and last item, respectively.
 
-Similarly, each point of each of the frontiers can be used to estimate the local phase at that location.
+% Similarly, each point of each of the frontiers can be used to estimate the local phase at that location.
 For the positive frontier, it is necessary to note that, for each point, the following equation holds: $ \cos(\phi + 2 \pi f i / n) = 1 $, and thus $ \phi = 2 \pi k - 2 \pi f i / n $, with $ k \in \mathbb{N} $, while for the negative frontier $ \cos(\phi + 2 \pi f i / n) = -1 $ and $ \phi = 2 \pi k + \pi - 2 \pi f i / n) $.
 
+\subsection{Frontiers}
+
+In practice, is not uncommon for a discrete wave, especially in the case of sound, to present somewhat different positive and negative contours; in those cases, the algorithm here presented can be used independently in the positive and negative pulses of the wave to define two "envelopes", that we are going to call superior and inferior frontiers.
+
+Figure \ref{fig:FullFrontiers} illustrates the frontiers of six diverse discrete sound waves, as well as an in-detail view of the highlighted segment for each signal.
+All waves are records of physical sounds, chosen to represent the applicability of the algorithm in real-world scenarios.
+
+\begin{figure}[ht!]
+  \centering
+    \includegraphics{images/07FullFrontiers.pdf}
+  \caption{Positive and negative frontiers of six digital waves, as extracted by the algorithm here presented.
+For each wave, the region highlighted in black is shown in detail besides the whole wave.
+For each wave the horizontal axis is the sample number $i$, while the vertical axis is the normalized amplitude.
+All 6 waves were recorded at 44100 fps.}
+  \label{fig:FullFrontiers}
+\end{figure}
+
+The frontiers are satisfactorily detected in the case of harmonic and inharmonic sounds, and is robust in relation to the number of samples and the frequencies of the waves.
+
+% The impact of the normalization of a wave using the frontiers obtained can be seen in figure \ref{fig:Fourier}, which shows the original and normalized wave in the case of a recording of the key 33 of a piano both in the time and frequency domains.
+We can see that the Fourier power spectra is greatly simplified in the case of the normalized wave.
+
+% \begin{figure}[ht!]
+%   \centering
+%     \includegraphics[width=0.8\linewidth]{08Fourier.pdf}
+%   \caption{Original and normalized views of a recording of key 33 (F3, 175 Hz) from a grand piano both in time and frequency domains.}
+%   \label{fig:Fourier}
+% \end{figure}
+
+% \subsection{Robustness}
+
+% Figures \ref{fig:xRobustness} and \ref{fig:yRobustness} illustrate the computational resilience of the algorithm for scaling in both the abscissa and ordinate axes, respectively.
+This is due to the normalization scheme presented in equation \ref{eq:scale}.
+
+% Although important from a theoretical standpoint, this step has an especial computational relevance in the case of digital sound waves, as they tend to be composed of a high number of samples with relatively low amplitudes, generally normalized between -1 and 1, leading to a numerical disparity between axes.
+
+% \begin{figure}[ht!]
+%   \centering
+%     \includegraphics[width=0.8\linewidth]{10xRobustness.pdf}
+%   \caption{Minimum and maximum scales of the abscissa before computational stability issues arose, for a recording of a drum kit tom.  $ \text{B} = 10^6 $.}
+%   \label{fig:xRobustness}
+% \end{figure}
+
+% \begin{figure}[ht!]
+%   \centering
+%     \includegraphics[width=0.8\linewidth]{09yRobustness.pdf}
+%   \caption{Minimum and maximum values of the amplitude before computational stability issues arose, for a recording of a drum kit tom.}
+%   \label{fig:yRobustness}
+% \end{figure}
+
 \section{Discussion}
+% Put results into a broader context.
+
+% What does it all mean?
+
+% what principles have been established or reinforced
 

### 2020-11-30 Papers/01_Pulses_Envelope/images/00ConvexHull.py

@@ -72,7 +72,7 @@ fig.update_yaxes(showline=False, showgrid=False, zerolinewidth=1, zerolinecolor=
 '''Samples'''
 fig.add_trace(
   go.Scatter(
-    name="Samples",
+    name="Samples      ",
     # showlegend=False,
     x=X,
     y=W,
@@ -85,7 +85,7 @@ fig.add_trace(
 I = [v for v in hull.vertices] + [hull.vertices[0]]
 fig.add_trace(
   go.Scatter(
-    name="Convex Hull",
+    name="Convex Hull      ",
     # showlegend=False,
     x=X[I],
     y=W[I],
@@ -100,7 +100,7 @@ fig.add_trace(
 
 fig.add_trace(
   go.Scatter(
-    name=f"Concave Hull (Alpha-Shape, \u03B1={alpha})    ",
+    name=f"Concave Hull (Alpha-Shape, \u03B1={alpha})      ",
     # showlegend=False,
     x=XX,
     y=YY,
@@ -112,7 +112,7 @@ fig.add_trace(
   )
 )
 
-fig.show(config=dict({'scrollZoom': True}))
+# fig.show(config=dict({'scrollZoom': True}))
 save_name = "./images/" + sys.argv[0].split('/')[-1].replace(".py", ".svg")
 fig.write_image(save_name, width=650, height=300, engine="kaleido", format="svg")
 print("saved:", save_name)

### 2020-11-30 Papers/01_Pulses_Envelope/images/01AnalyticSignal.py

@@ -35,27 +35,35 @@ analytic_signal = hilbert(W)
 hilbert_envelope = np.abs(analytic_signal)
 
 
+'''============================================================================'''
+'''                              PLOT LINES                                    '''
+'''============================================================================'''
+FONT = dict(
+    family="Latin Modern Roman",
+    color="black",
+    size=13.3333
+  )
+
 '''Plotting'''
 fig = go.Figure()
 fig.layout.template ="plotly_white" 
 fig.update_layout(
-  xaxis_title="$i$",
-  yaxis_title="Amplitude",
+  xaxis_title="<b><i>i</i></b>",
+  yaxis_title="<b>Amplitude</b>",
   legend=dict(orientation='h', yanchor='top', xanchor='left', y=1.1),
-  margin=dict(l=5, r=5, b=5, t=5),
-  font=dict(
-  family="Latin Modern Roman",
-  color="black",
-  size=18
-  )
+  margin=dict(l=0, r=0, b=0, t=0),
+  font=FONT,
+  titlefont=FONT
 )
+fig.layout.xaxis.title.font=FONT
+fig.layout.yaxis.title.font=FONT
 fig.update_xaxes(showline=False, showgrid=False, zeroline=False)
 fig.update_yaxes(showline=False, showgrid=False, zerolinewidth=2, zerolinecolor="silver")
 
 '''Plotting Signal'''
 fig.add_trace(
   go.Scatter(
-    name="Signal",
+    name="Signal      ",
     # x=X,
     y=W,
     mode="lines",
@@ -68,7 +76,7 @@ fig.add_trace(
 
 fig.add_trace(
   go.Scatter(
-    name="Hilbert Envelope",
+    name="Hilbert Envelope      ",
     # x=X,
     y=hilbert_envelope,
     mode="lines",
@@ -81,7 +89,7 @@ fig.add_trace(
 
 fig.add_trace(
   go.Scatter(
-    name="$ i = n/2 $",
+    name="<i>i = n/2 </i>      ",
     x=[n//2, n//2],
     y=[-1, 1],
     mode="lines",
@@ -93,7 +101,7 @@ fig.add_trace(
   )
 )
 
-# fig.show()
-save_name = "./" + sys.argv[0].split('/')[-1].replace(".py", ".pdf")
-fig.write_image(save_name, width=800, height=400, scale=1, engine="kaleido")
+# fig.show(config=dict({'scrollZoom': True}))
[…]

### 2020-11-30 Papers/01_Pulses_Envelope/images/01Pulses.py

@@ -1,146 +1,0 @@
-import sys
-print(sys.version)
-import plotly.graph_objects as go
-sys.path.append("c:/Users/tesse/Desktop/Files/Dropbox/0_Thesis/Python/01_Envelope")
-import numpy as np
-
-from Helper import signal_to_pulses, get_pulses_area
-
-'''Generating Wave'''
-
-np.random.seed(1)
-# fps = 10
-n = 100+1
-
-X = np.arange(n)# / fps
-W = np.zeros(n)
-
-afp = [
-  [2, 3, np.pi],
-  # [1, 2, 1.2 * np.pi],
-  # [1, 2.3, 1.4 * np.pi]
-  ]
-for a, f, p in afp:
-  W += a * np.cos(p + 2 * np.pi * f * X / n) + np.random.normal(0, a / 5, n)
-
-W = W / np.max(np.abs(W))
-
-'''Wave to pulses'''
-pulses = signal_to_pulses(W)
-pulses_X, pulses_Y = get_pulses_area(pulses)
-
-'''============================================================================'''
-'''                              PLOT LINES                                    '''
-'''============================================================================'''
-FONT = dict(
-    family="Latin Modern Roman",
-    color="black",
-    size=13.3333
-  )
-
-'''Plotting'''
-fig = go.Figure()
-fig.layout.template ="plotly_white" 
-fig.update_layout(
-  xaxis_title="<b><i>i</i></b>",
-  yaxis_title="<b>Normalized Amplitude</b>",
-  legend=dict(orientation='h', yanchor='top', xanchor='left', y=1.1),
-  margin=dict(l=0, r=0, b=0, t=0),
-  font=FONT,
-  titlefont=FONT
-)
-fig.layout.xaxis.title.font=FONT
-fig.layout.yaxis.title.font=FONT
-
-fig.update_xaxes(showline=False, showgrid=False, zeroline=False)
-fig.update_yaxes(showline=False, showgrid=False, zerolinewidth=2, zerolinecolor='black')
-
-'''Samples'''
-fig.add_trace(
-  go.Scatter(
-    name="Samples",
-    showlegend=False,
-    x=X,
-    y=W,
-    mode='markers',
-    fill="tozeroy",
-    fillcolor="gray",
-    marker=dict(
-        size=5,
-        color="black",
-    )
-  )
-)
-fig.add_trace(
-  go.Scatter(
-    name= "Samples Stem",
[…]

### 2020-11-30 Papers/01_Pulses_Envelope/images/06Graph.py

@@ -1,0 +1,114 @@
+'''To be used AFTER signal_envelope is installed via pip'''
+import sys
+# sys.path.append("signal_envelope/")
+import numpy as np
+import signal_envelope as se
+from scipy import interpolate
+import plotly.graph_objects as go
+
+name = "bend"
+W, _ = se.read_wav(f"C:/Users/tesse/Desktop/Files/Dropbox/0_Science/reps/envelope/test_samples/{name}.wav")
+amp = np.max(np.abs(W))
+W = 4 * W / amp
+X = np.arange(W.size)
+
+Xpos, Xneg = se.get_frontiers(W, 0)
+E = se.get_frontiers(W, 1)
+
+f = interpolate.interp1d(E, np.abs(W[E]), kind="linear", fill_value="extrapolate")
+E = f(X)
+
+for i in range(E.size):
+  if E[i] < 0.1:
+    E[i] = 0.1
+
+C = W / E
+
+'''============================================================================'''
+'''                                    PLOT                                    '''
+'''============================================================================'''
+
+FONT = dict(
+    family="Latin Modern Roman",
+    color="black",
+    size=13.3333
+  )
+
+'''Plotting'''
+fig = go.Figure()
+fig.layout.template ="plotly_white" 
+fig.update_layout(
+  xaxis_title="<b><i>i</i></b>",
+  yaxis_title="<b>Amplitude</b>",
+  legend=dict(orientation='h', yanchor='top', xanchor='left', y=1.1),
+  margin=dict(l=0, r=0, b=0, t=0),
+  font=FONT,
+  titlefont=FONT
+)
+fig.layout.xaxis.title.font=FONT
+fig.layout.yaxis.title.font=FONT
+
+fig.update_xaxes(showline=False, showgrid=False, zeroline=False)
+fig.update_yaxes(showline=False, zeroline=False, showgrid=False, gridwidth=1, gridcolor='gray', tickvals=[-1, 0, 1])
+
+
+fig.add_trace(
+  go.Scatter(
+    name="Signal (<b>w = e</b> ⊙ <b>c</b>)      ",
+    # x=np.arange(W.size),
+    y=W,
+    mode="lines",
+    # fill="tozeroy",
+    # fillcolor="rgba(0,0,0,0.6)",
+    line=dict(width=3, color="gray",),
+  )
+)
+
+fig.add_trace(
+  go.Scatter(
[…]

### 2020-11-30 Papers/01_Pulses_Envelope/images/07Comparison.py

@@ -7,19 +7,11 @@ from scipy.signal import savgol_filter, hilbert
 import os
 from pathlib import Path
 import sys
-sys.path.append("c:/Users/tesse/Desktop/Files/Dropbox/0_Thesis/Python")
-from Wheel import frontiers
+# sys.path.append("c:/Users/tesse/Desktop/Files/Dropbox/0_Thesis/Python")
+# from Wheel import frontiers
 from plotly.subplots import make_subplots
 from timeit import default_timer as timer
-
-
-
-def read_wav(path): 
-  """returns signal & fps"""
-  wav = wave.open(path , 'r')
-  signal = np.frombuffer(wav.readframes(-1) , np.int16)
-  fps = wav.getframerate()
-  return signal, fps
+import signal_envelope as se
 
 
 
@@ -28,12 +20,7 @@ def read_wav(path):
 ''' Read wav file '''
 '''==============='''
 name = "bend"
-parent_path = str(Path(os.path.abspath('./')).parents[1])
-path = f"{parent_path}/Python/Samples/{name}.wav"
-print(path)
-
-W, fps = read_wav(path)
-
+W, fps = se.read_wav(f"C:/Users/tesse/Desktop/Files/Dropbox/0_Science/reps/envelope/test_samples/{name}.wav")
 W = W - np.average(W)
 amplitude = np.max(np.abs(W))
 W = W / amplitude
@@ -41,39 +28,18 @@ n = W.size
 print(f"n={n}")
 X = np.arange(n)
 
-# '''==============='''
-# '''    Sinusoid   '''
-# '''==============='''
-# FT = np.fft.rfft(W)# * 2 / n
-# PW = np.abs(FT)
-# f = np.argmax(PW)
-# p = np.angle(FT[f])
-
-# S = np.cos(p + 2 * np.pi * f * X / n)
-
 '''==============='''
 '''    Snowball   '''
 '''==============='''
 start = timer()
-pos_frontierX, pos_frontierY, neg_frontierX, neg_frontierY = frontiers(W)
+Ex = se.get_frontiers(W, 1)
 
-frontierX = np.concatenate([pos_frontierX, neg_frontierX])
-frontierY = np.concatenate([pos_frontierY, np.abs(neg_frontierY)])
 
-idxs = np.argsort(frontierX)
-frontierX = frontierX[idxs]
-frontierY = frontierY[idxs]
[…]

### 2020-11-30 Papers/01_Pulses_Envelope/images/old/10xRobustness.py

@@ -177,8 +177,7 @@ fig.add_trace(
     mode="lines",
     line=dict(color="silver"),
     showlegend=True,
-  ),
-  row=1, col=2
+  ), row=1, col=2
 )
 
 fig.add_trace(

### 2020-12-05 Papers/01_Pulses_Envelope/PulsesEnvelope.tex
grammar checked

@@
 
-% \usepackage{bm} % bold & upright inside math env
 % \usepackage{hyperref}
 % \usepackage{hypcap}
-\DeclareMathOperator{\sign}{sign}
-\bibliography{bibli}
 % \hypersetup{
@@
 % }
+
+\DeclareMathOperator{\sign}{sign}
+\bibliography{bibli}
+
 \title{Envelope Estimation using Geometric Properties of a Discrete Real Signal}
@@
 Despite being an elusive concept, the temporal amplitude envelope of a signal is essential for its complete characterization, being the primary information-carrying medium in spoken voice and telecommunications, for example.
-Intuitively, the temporal envelope can be understood as a slow varying function that multiplies the signal, being responsible for its outer shape.
+Intuitively, the temporal envelope can be understood as a slowly varying function that multiplies the signal, being responsible for its outer shape.
 Envelope detection techniques have applications in areas like health, sound classification and synthesis, seismology and speech recognition. 
 % Problem | Objectives
-Nevertheless, a general approach to digital envelope detection of signals with rich spectral content doesn't exist, as most methods involve manual intervention, in the form of filter design, smoothing, as well as other specific design choices, based on a priori knowledge of the nature of the specific waves under investigation.
+Nevertheless, a general approach to digital envelope detection of signals with rich spectral content doesn't exist, as most methods involve manual intervention, in the form of filter design, smoothing, as well as other specific design choices, based on prior knowledge of the nature of the specific waves under investigation.
 To address this problem, we propose a framework that uses intrinsic characteristics of a signal to estimate its envelope, eliminating the necessity of parameter tuning.
 % Methods | Procedure | Approach
-The approach here described draws inspiration from geometric concepts to estimate the temporal envelope of an arbitrary signal; specifically, a new measure of discrete curvature is used to obtain the average radius of curvature of a discrete wave, that will as a threshold to identify the wave's samples that are part of the envelope, in a manner similar to the alpha-shapes approach to the identification of the concave hull of a set of points.
+The approach here described draws inspiration from geometric concepts to estimate the temporal envelope of an arbitrary signal; specifically, a new measure of discrete curvature is used to obtain the average radius of curvature of a discrete wave, that will as a threshold to identify the wave's samples that are part of the envelope, in a manner similar to the alpha shapes approach to the identification of the concave hull of a set of points.
 % Results | Findings | Product
-We provide visualizations of the envelope extracted via the algorithm for various real world signals, with very diverse characteristics, such as voice, spoken and sang, and pitched and non pitched musical instruments, and discuss some approaches to assess the quality of the obtained envelopes.
-The algorithm compares favourably with classic envelope detection techniques based on filtering and the Hilbert Transform.
-A Python module, installable via the Python Package Index (PyPI), with a pure Python and a specialized C++ implementation for 64 bit Windows systems, and interactive visualizations of envelopes for a diverse range of digital waves, as well as the source code for the Python implementation, are available online.
+The algorithm compares favourably with classic envelope detection techniques based on smoothing, filtering and the Hilbert Transform, besides being physically plausible.
+We provide visualizations of the envelope extracted via the algorithm for various real-world signals, with very diverse characteristics, such as voice, spoken and sang, and pitched and non-pitched musical instruments, and discuss some approaches to assess the quality of the obtained envelopes.
+A Python module, installable via the Python Package Index (PyPI), with a pure Python and a specialized C++ implementation for 64-bit Windows systems, interactive visualizations of envelopes for a diverse range of digital waves, as well as the source code for the Python implementation, are available online at \textcite{2020TarjanoEnvelope}.
 % Conclusions | Implications
 Besides the most direct applications of this work to audio classification and synthesis, we foresee impact in compression techniques and machine learning approaches to audio.
-The discrete curvature definition presented could also be extended to three-dimensional settings, to improve shape detection algorithms based on alpha-shapes.
+The discrete curvature definition presented could also be extended to three-dimensional settings, to improve shape detection algorithms based on alpha shapes.
 \end{abstract}
 
-{\bf Keywords:} DSP, alpha-shapes, envelope detection, discrete curvature estimation, demodulation
+{\bf Keywords:} DSP, alpha shapes, envelope detection, discrete curvature estimation, demodulation
 
@@
 Envelope detection, also known as demodulation, is ubiquitous in both analogue and digital signal processing \parencite{2011CaetanoImproved}.
-Nevertheless, the literature in this area is very fragmented \parencite{2017LyonsDigital}.
+Nevertheless, the literature in this area is fragmented \parencite{2017LyonsDigital}.
 Besides, most envelope detection techniques are designed to account for very specific settings, like pure sinusoids with moderate noise content, a limitation that excludes most physical signals, as is the case of recorded sound, for example; that limitation arises in part due to the lack of a strict mathematical definition of a temporal envelope \parencite{2013Mengempirical}.
@@
 
-One direct way of defining the shape of a set of points in $ \mathbb{R}^2 $ is using the convex hull: as defined in \textcite{2008BergComputational}, the convex hull of a set of points $ S $ in two dimensional space is the smallest convex subset of the plane that contains all the points in $ S $.
+One direct way of defining the shape of a set of points in $ \mathbb{R}^2 $ is using the convex hull: as defined in \textcite{2008BergComputational}, the convex hull of a set of points $ S $ in two-dimensional space is the smallest convex subset of the plane that contains all the points in $ S $.
 Convex hulls are unique, in the sense that a set of points define one and only one convex hull.
 
-Generalizing that definition, \textcite{1983Edelsbrunnershape} introduced the concept of alpha-shapes, that can be seen as a Concave Hull; a mathematically well-defined generalization of the Convex Hull of a finite set of points, closely related to the Delaunay triangulation and Voronoi diagrams of those points.
+Generalizing that definition, \textcite{1983Edelsbrunnershape} introduced the concept of alpha shapes, that can be seen as a Concave Hull; a mathematically well-defined generalization of the Convex Hull of a finite set of points, closely related to the Delaunay triangulation and Voronoi diagrams of those points.
 
-The intuition behind the alpha shapes algorithm as that only points that can be touched by a circle of radius $ \alpha $ coming from an infinite distance towards the set of points of interest are part of the frontier, that is, of the Concave Hull of this set.
+The intuition behind the alpha shapes algorithm is that only points that can be touched by a circle of radius $ \alpha $ coming from an infinite distance towards the set of points of interest are part of the frontier, that is, of the Concave Hull of this set.
 
 Once an $ \alpha $ parameter is defined alpha shapes conserve the uniqueness property, besides reducing to the Convex Hull when $ \alpha = \infty $.
-Consider, for example, the artificially generate discrete wave in figure \ref{fig:Hulls}
-below, where the links between consecutive samples were deemphasised to highlight the "set of points" nature of a signal: The Convex Hull and a Concave Hull, in the form of an alpha shape with a particular value of $ \alpha $ are illustrated.
+Consider, for example, the artificially generated discrete wave in figure \ref{fig:Hulls}
+below, where the links between consecutive samples were deemphasized to highlight the “set of points” nature of a signal: The Convex Hull and a Concave Hull, in the form of an alpha shape with a particular value of $ \alpha $ are illustrated.
 
@@
     \includegraphics{images/00ConvexHull.pdf}
-  \caption{A discrete signal interpreted as a set of points in the Cartesian plane, and the Convex Hull and Concave Hull at $ \alpha=300 $ defined by it}
+  \caption{A discrete signal interpreted as a set of points in the Cartesian plane, and the Convex Hull and Concave Hull at $ \alpha=300 $ defined by it.}
   \label{fig:Hulls}
@@
 
-Following this path, we present a general approach to envelope detection, exploiting the intrinsic characteristics of a generic, spectrally complex wave, in order to avoid as much as possible the need for manual intervention or parameter tuning, such as the choice of $ \alpha $.
+Following this path, we present a general approach to envelope detection, exploiting the intrinsic characteristics of a generic, spectrally complex wave, in order to completely avoid the need for manual intervention or parameter tuning.
 
-% TODO <|<|<|<|<|
+The rest of this paper is divided as follows: 
+We start by trying to define as specifically as possible what a temporal envelope is, especially in the context of this work, providing a mathematical definition, in terms of the Hilbert transform, for the case of a narrowband signal.
 
-While the robustness of the proposed approach allows it to be used as a plug-in replacement for many methods encountered in the literature, we feel that it would be particularly useful for sound synthesis.
+We proceed to present the representation of a discrete wave used in this work, largely based on the concept of dividing the wave into its constituent pulses, and how it can simplify the algorithm, reducing the high dimensionality generally present by digital representations of waves.
 
-The envelope is shown to add complexity to the spectral representation of a wave \parencite{2019TarjanoNeuro}, and an accurate description of the envelope, while describing the evolution of the instantaneous amplitude of a signal in time, would also greatly simplify further spectral analysis.
-Moreover, the algorithm developed in this work naturally divides a signal into its pseudo-cycles, pinpointing them in the time domain, providing the building blocks for the reconstruction of the fine structure of the wave.
+Next, we review methods for the estimation of the curvature of a discrete wave, after which we present our own; this method will then be used to estimate the equivalent circle, that will be used to define the envelope of the wave.
 
-The rest of this paper is divided as follows: After explaining and defining the various entities to be used subsequently, of which the concept of frontiers is arguably the most important, in the methodology section, we dedicate a section to present a new discrete curvature estimation approach, and how it was used to identify the frontiers of a discrete wave.
+In the results section, we present some illustrations of the envelopes obtained with our method, comparing both accuracy and running time with the most common algorithms encountered in the literature.
+We also suggest an alternative use of the algorithm to identify the negative and positive frontiers of a wave, a kind of partial envelope considering only the positive and negative pulses of a wave, respectively.
 
-We then proceed to illustrate the results of the application of the algorithm in a set of six diverse sound waves, ranging from voice to unpitched instruments comparing, afterwards, its performance with that of classic digital envelope estimation algorithms, both from a visual and numerical standpoint.
-
-We then discuss the implications of the work, emphasizing the focus on the construction of new interpretations of the relationship between time domain and frequency domain sound descriptions, and our vision of the impact of this methodology in the field of sound synthesis, providing directions for future developments.
+We follow with some suggestions about how the approach here presented can be useful beyond envelope detection, exemplifying with some applications, such as the extractions of the average waveform of a digital wave, concluding with some commentary on potential future developments of the research here presented.
 
 \section{Methodology}
-% Overview of methods
+After defining what will be considered an envelope in the context of this work, in this section we explore some simplifications that can be applied to a discrete wave, defining the representation that will be used in the subsequent part of the work.
+
+We next explore the concept of discrete curvature, putting forth our own method, tailored to the definition of the equivalent circle of a wave; we end this section by showing how this circle can be used to identify the envelope of a discrete wave, in a similar approach to that of the alpha shapes theory, but taking advantage of some unique structure present in discrete waves.
 
@@
 
-The interest in the theory of envelope detection arose in the analogue domain, with the widespread adoption of radio communications, where the problem is typically restricted to simple sinusoids, often with well defined frequencies \parencite{2011TurnerDemodulation}.
+The interest in the theory of envelope detection arose in the analogue domain, with the widespread adoption of radio communications, where the problem is typically restricted to simple sinusoids, often with well-defined frequencies \parencite{2011TurnerDemodulation}.
 
 In this context, the most mathematically sound definition of an envelope involves the representation of its underlying wave as an analytic signal, introduced by Gabor in 1946 \parencite{2007HahnHistory}.
-In his work, \textcite{1946GaborTheory} applies the then relatively new mathematical machinery of the quantum mechanics to unify the time and frequency domain representations of a wave, showing how the Hilbert transform could be applied to a real signal in order to obtain an equivalent complex signal, later known as the analytic signal.
+In his work, \textcite{1946GaborTheory} applies the then relatively new mathematical machinery of the quantum mechanics to unify the time and frequency-domain representations of a wave, showing how the Hilbert transform could be applied to a real signal in order to obtain an equivalent complex signal, later known as the analytic signal.
 
@@
 
+%%%%%%%%%%%%%%%%%
+
 \begin{figure}[ht!]
@@
 In general, the problem of envelope detection, or demodulation, can be interpreted as the task of, given a continuous wave $ W(t) $, decomposing it in two components $ E(t) $ and $ C(t) $, such that $ W(t) = E(t) C(t) $ \parencite{2011TurnerDemodulation}.
+
 $ E(t) $ represents the slow varying part of the wave, also known as the (temporal) envelope, modulator component, or amplitude modulation (AM) of $ W $, while $ C(t) $ models its fast varying part, called throughout the literature as its (temporal) fine structure, carrier component, or frequency modulation (FM).
 
-In the case of broadband signals, the problem of envelope detection of an arbitrary wave is ill posed, in the sense that an infinite number of pairs of $ E(t) $ and $ C(t) $ can result in the same specific $ W(t) $ \parencite{2011TurnerDemodulation, 2013Mengempirical, 1996Loughlinamplitude}.
+In the case of broadband signals, the problem of envelope detection of an arbitrary wave is ill-posed, in the sense that an infinite number of pairs of $ E(t) $ and $ C(t) $ can result in the same specific $ W(t) $ \parencite{2011TurnerDemodulation, 2013Mengempirical, 1996Loughlinamplitude}.
 
@@
 
-\begin{equation} \label{eq:Envelope}
-  \begin{align}
+\begin{align} \label{eq:Envelope}
   & \textbf{w}, \textbf{e}, \textbf{c} \in \mathbb{R}^n \\
@@
 \end{align}
-\end{equation}
 
@@
 
-% TODO <|<|<|<|<| we will assume that this assumption holds for quasi periodic signals: that is, the absolute maximum of each pseudo cycle is equal to the value of the envelope at that point, the other local extrema in that same cycle are the ones that should be filtered out.
-
-% TODO <|<|<|<|<| We lack a indicator to evaluate the quality of the envelope, once one is identified, but figure suggests that, once the envelope is removed from w, the wave should then be confined between the interval $ \{-1, 1\}$.
-
-Image \ref{fig:Envelope}, along with the Concave Hull illustrated in figure \ref{fig:Hulls}, helps forming the intuition behind the method developed in this work: the idea is to identify the local extrema that touch the envelope (or, to put in another way, the local extrema in relation to the whole wave that are global extrema inside their own pseudo cycles), filtering out those who don't; in this way, we're defining the envelope itself at those points.
+Image \ref{fig:Envelope}, along with the Concave Hull illustrated in figure \ref{fig:Hulls}, helps forming the intuition behind the method developed in this work: the idea is to identify the local extrema that touch the envelope (or, to put in another way, the local extrema in relation to the whole wave that are global extrema inside their own pseudo-cycles), filtering out those who don't; in this way, we're defining the envelope itself at those points.
 
@@
 
-In the next section such a geometric representation will be formulated, along with the necessary tool to estimate the average curvature of the wave and, consequently, the appropriate radius of the circle.
-
-% TODO <|<|<|<|<|
+In the next section, such a geometric representation will be formulated, along with the necessary tool to estimate the average curvature of the wave and, consequently, the appropriate radius of the circle.
 
@@
 
-P is not fit for geometric interpretation yet, since the abscissa and ordinate of his orthogonal coordinate system have different units: in the vertical axis we have a unit related to the instantaneous intensity of the wave, while in the horizontal axis we have the index at which each extremum occurs, ultimately a time related unit.
+P is not fit for geometric interpretation yet, since the abscissa and ordinate of his orthogonal coordinate system have different units: in the vertical axis, we have a unit related to the instantaneous intensity of the wave, while in the horizontal axis we have the index at which each extremum occurs, ultimately a time-related unit.
 
@@
 
-We are interested in finding a coordinate system where a pulse's amplitude $ \lvert w_j \lvert $ and length would define, in average, a square.
+We are interested in finding a coordinate system where a pulse's amplitude $ \lvert w_j \lvert $ and length would define, on average, a square.
 
@@
 The drawback is that we would lose the direct link between P and \textbf{w} that arises from the fact that the abscissas of the points in P are indices $ i $ of \textbf{w}: we choose instead to leave the abscissa intact and multiply the now unitless vertical basis axis by the average of the difference between the abscissa of a point and the abscissa of the immediately posterior.
-In this way, both axes will have unit $ i $, and the relation is preserved, making it easier to recover the envelope points later.
+In this way, both axes will have unit $ i $, and the relationship is preserved, making it easier to recover the envelope points later.
 
@@
 
-\begin{equation} \label{eq:Basis}
-\begin{align}
+\begin{align} \label{eq:Basis}
 \mathbf{b}^\prime_x &= \mathbf{b}_x  \\
@@
 \begin{equation} \label{eq:Vectors}
-\begin{align}
 \text{V} = \{ v_0, v_1, v_{m-2} \} \quad \text{where} \quad v_k = P_{j+1} - P_j \quad \forall \quad 0 \le k \le m-2
-\end{align}
 \end{equation}
@@
 
-Translating the explanation of alpha-shapes in \textcite{1994EdelsbrunnerThree} to our context, we can say that the points that are part of the envelope are those touched by a circle of a given radius $ r $, coming from $ y \to \infty $ towards the signal, that is not allowed to contain any point of the signal.
-% TODO refine
-
-Intuitively, one can picture a circle being rolled above the signal, and marking the points it touches as envelope points.
+Translating the explanation of alpha shapes in \textcite{1994EdelsbrunnerThree} to our context, we can say that the points that are part of the envelope are those touched by a circle of a given radius $ r $, coming from $ y \to \infty $ towards the signal, that is not allowed to contain any point of the signal; Intuitively, one can picture a circle being rolled above the signal, and marking the points it touches as envelope points.
 
@@
 
-% Concerning the last method, two approaches to fit a polynomial to a generic wave are readily available: the least mean squares approach, that seeks to minimize the dependent variable errors and the total least squares problem, that treats both variables symmetrically \parencite{1980GolubAnalysis}.
-
-% The geometric, symmetrical nature of the problem excludes the more computationally economic least mean squares approach, however, leaving us with the total least squares, for which no general closed-form solutions are available \parencite{2007MarkovskyOverview}; this leads us to turn our attention to direct approaches of curvature estimation.
-
-Concerning direct methods \textcite{2014CarrollSurvey}, for example, derive three such definitions based on the approximation of a circle by an inscribed, centred and circumscribed polygon.
-In the context of tridimentional meshes, \textcite{2016VasaMultivariate} evaluate a range of existing estimators from a multivariate point of view.
+Concerning direct methods \textcite{2014CarrollSurvey}, for example, derive three such definitions based on the approximation of a circle by an inscribed, centered and circumscribed polygon.
+In the context of tridimensional meshes, \textcite{2016VasaMultivariate} evaluate a range of existing estimators from a multivariate point of view.
 
@@
 \begin{equation} \label{eq:Radius}
-% \begin{gather}
 r = \frac{\sum\limits_{k=0}^{m-2} \left( \frac{v_{k,x} \sqrt{v_{k,x}^2 + v_{k,y}^2}}{ v_{k,y}} \right)}{m-2}
-% \end{gather}
 \end{equation}
@@
 We need now to construct an algorithm, using the radius just obtained, to identify the points that belong to the envelope.
-Generally, the algorithm for the alpha-shapes approach resorts to the construction of the Delaunay triangulation of the set of points, that is latter filtered.
+Generally, the algorithm for the alpha shapes approach resorts to the construction of the Delaunay triangulation of the set of points, that is latter filtered.
 We can adopt a more straightforward method, as our points have a defined structure:
@@
 
-\begin{equation} \label{eq:InEnvelope}
-\begin{align}
+\begin{align} \label{eq:InEnvelope}
 & P_a \in \text{E} \leftrightarrow (P_{b,x} - x_o)^2 + (P_{b,y} - y_o)^2 > r^2, \\
@@
 The most straightforward way to transform those points into the vector $ \textbf{e} \in \mathbb{R}^n $ is via linear interpolation, the approach used in this work.
-is worthy noting, however, that many other procedures are available, both for the interpolation and the smooth approximation of points.
-We deliberately choose linear interpolation to avoid blurrying the results of the method.
+It is worth noting, however, that many other procedures are available, both for the interpolation and the smooth approximation of points.
+We deliberately choose linear interpolation to avoid blurring the results of the method.
 
 \section{Results}
-% Revisit purpose and methods
-% Overview of results
+Given the lack of a precise definition of what an envelope should be, we begin this section suggesting an empirical metric, based on the behaviour of the carrier wave \textbf{c}.
+We also comment on some metrics presented in the literature, and compare our algorithm with traditional demodulation approaches, both in relation to execution time and accuracy; to this last end, a numerical indicator is introduced.
+
+We also illustrate how the method here presented can be used to isolate not only a single envelope of the wave, but the superior and inferior envelopes as well, that we call frontiers, ending the section with a suggestion on how the proposed algorithm can be useful beyond envelope detection, by identifying the approximate locations of the pseudo-cycles in a quasi-periodic wave.
+
+Due to an assortment of factors, such as the unconventional approach presented in this work and the intrinsic difficulty in visualizing discrete waves given their high dimensionality, the authors deemed helpful to provide a companion website with interactive visualizations \parencite{2020TarjanoEnvelope}.
+
+An implementation of the method in pure Python, fully functional, can also be found at the repository dedicated to this work as well as a Python module available at the Python Package Index (PyPI) and conveniently installable via the command `pip install signal-envelope`.
+Besides, those working under Windows 64-bit can benefit from a specialized dynamic library coded in C++ that is automatically used within this module.
 
@@
 
-There are no agreed upon objective metrics to assess the quality of an envelope.
+There are no agreed-upon objective metrics to assess the quality of an envelope.
 Recalling figure \ref{fig:Envelope}, however, and the accompanying discussion, an intuitive way is to consider the carrier wave \textbf{c}, and how well it seems to be bounded in amplitude in the interval $ \{-1, 1\} $.
-To that end, with provide the follow figures of original discrete waves \textbf{w}, their envelope \textbf{e}, and the inferred carrier, obtained after dividing, element wise, the original wave by its envelope.
+To that end, we provide the following figures of original discrete waves \textbf{w}, their envelope \textbf{e}, and the inferred carrier, obtained after dividing, element-wise, the original wave by its envelope.
+
+From figure \ref{fig:spoken_voice} it can be seen that the algorithm performs well under the carrier metric.
+This is especially remarkable for the case of spoken voice, a notorious complex sound with fast changes in pitch and high noise content.
+
+These characteristics are in stark contrast with the sang voice in figure \ref{fig:singer_voice}, where a recording of an alto singer is shown; there, the wave is almost periodic, with a stable waveform.
 
@@
 
-Due to an assortment of factors, such as the unconventional approach presented in this work and the intrinsic dificult in visualizing discrete waves given their high dimensionality, the authors deemed helpful to provide a companion website with interactive visualizations.
+\begin{figure}[ht!]
+  \centering
+    \includegraphics{images/06Graph_spoken_voice.pdf}
+  \caption{Signal, envelope and carrier for a record of male voice uttering the word "amazing".}
+  \label{fig:spoken_voice}
+\end{figure}
 
-An implementation of the method in pure Python, fully functional, can also be found at the repository dedicated to this work as well as a Python module available at the Python Package Index (PyPI) and conveniently instalabbe via the command `pip install signal-envelope`.
-Besides, those working under Windows 64 bit can benefit from a specialized dll coded in C++ that is automatically used within this module.
+\begin{figure}[ht!]
+  \centering
+    \includegraphics{images/06Graph_alto.pdf}
+  \caption{Signal, envelope and carrier for a record of an alto singer.}
+  \label{fig:singer_voice}
+\end{figure}
 
@@
 The first one states that if a signal is bounded in magnitude, then its envelope should be, as well.
-Our algorithm complies with this requisite, as the envelope is composed of selected samples from the original signal itself;
+Our algorithm complies with this requisite, as the envelope is composed of selected samples from the original signal itself.
 
@@
 
-Nevertheless, insight can be gained from a comparison of the results of the method here proposed with some of the most common envelope extraction algorithms, as can be seen in figure \ref{fig:Comparison}.
+Nevertheless, insight can be gained from a comparison of the results of the method here proposed with some of the most common envelope extraction algorithms.
+Specifically, we compare the present method with the following approaches:
+
+\begin{enumerate}
+\item Smoothing: This is a relatively simple procedure that, applied to the absolute values of a signal, can provide an approximation of the variation of its amplitude in time, by smoothing the movement of the carrier.
+We use the Savitzky-Golay algorithm \parencite{1964SavitzkySmoothing} with a window width of 3001 samples and a cubic polynomial.
+\item Filtering: The absolute value of the wave is filtered using a digital implementation of the Butterworth \parencite{butterworth1930theory} low pass filter.
+We use an order 2 filter with a cut off frequency of 10 Hz.
+\item Hilbert Transform: We apply the Hilbert transform to the absolute values of the wave after filtering the original signal with a Butterworth filter with a cut off frequency of 100 Hz, using the absolute value of the analytic signal to obtain the envelope.
+\end{enumerate}
+
+In \ref{fig:Comparison_sinusoid} we illustrate the envelope extracted by the conventional algorithms, contrasting it with the result of the method presented in this work, for a simple sinusoid.
+Note that none of the conventional algorithms complies with the third condition proposed in \textcite{1996Loughlinamplitude} that the envelope of a periodic wave should be a straight line.
 
@@
   \centering
-    \includegraphics[width=0.8\linewidth]{12Comparison.pdf}
-  \caption{Comparison of the algorithm present in this work with the most common methods of digital envelope identification.}
-  \label{fig:Comparison}
+    \includegraphics{images/07Comparison_sinusoid.pdf}
+  \caption{Comparison of envelope detection algorithms for a simple sinusoid}
+  \label{fig:Comparison_sinusoid}
 \end{figure}
 
-For a recording of a guitar sound, in which a bend is performed, we contrast the proposed algorithm with four common approaches of varying complexity: simple smoothing of the rectified original wave, low pass filtering followed by rectification, peak identification and subsequent smoothing, and the extraction of the envelope via Hilbert transform, as presented earlier in the paper, with a previous smoothing of the original wave to avoid the artifacts shown in figure \ref{fig:AnalyticSignal}.
-
-All those methods demanded careful choice of parameters: in the case of filtering, a cut-off frequency of approximately 88 Hz was used, consisting of one additional parameter to be tuned.
-For all the other traditional approaches, the Savitzky-Golay smoothing algorithm was used, and the window size was the parameter to be defined.
-
-In the case of the proposed algorithm, the smooth temporal envelope was obtained merging both positive and negative frontiers, the result being smoothed with the same procedure described.
-Figure \ref{fig:FrontiersToEnvelope} illustrates the results.
+For a recording of the key 33 of a grand piano, where parts of the sound are highly percussive and noisy, the Hilbert approach presents considerable oscillations, as can be seen in figure \ref{fig:Comparison_piano}.
 
@@
   \centering
-    \includegraphics[width=0.8\linewidth]{13FrontiersToEnvelope.pdf}
-  \caption{Original positive and negative frontiers in dark grey and their mirrored versions. The temporal envelope, in black, is obtained by merging and approximating both frontiers.}
-  \label{fig:FrontiersToEnvelope}
+    \includegraphics{images/07Comparison_piano.pdf}
+  \caption{Comparison of envelope detection algorithms for the recording of the key 33 of a grand piano}
+  \label{fig:Comparison_piano}
 \end{figure}
 
-Table \ref{table:Comparison} presents a computational comparison between the techniques.
-The times are for custom implementations of those methods in the Python programming language.
-The least mean squares result was calculated by halving the envelopes presented in figure \ref{fig:Comparison} and comparing them with the rectified absolute wave, so as to avoid the potential influence of an estimated frequency.
+In figure \ref{fig:Comparison_soprano} all envelopes exhibit a similar movement along the samples, but the ones generated by traditional algorithms are undershooting the original wave.
 
-We can see that, while the algorithm here presented is the most accurate, it's also the most computationally intensive.
-The connection between alpha-shapes and Delaunay triangulation can be explored in future works as a means to improve the efficiency of the algorithm.
+\begin{figure}[ht!]
+  \centering
+    \includegraphics{images/07Comparison_soprano.pdf}
+  \caption{Comparison of envelope detection algorithms for the recording of the vocalization of a soprano singer}
+  \label{fig:Comparison_soprano}
+\end{figure}
+
+All the traditional methods demanded careful choice of parameters.
+In the case of the Hilbert transform, pre-processing in the form of filtering was applied.
+That is in contrast with our method, that organically tunes itself, in the form of the automatic choice of the radius of the circle representing the average discrete curvature of the discrete signal.
+
+Table \ref{table:Error_comparison} presents a comparison of the average errors of the techniques.
+The average error was calculated by the average of the squared distance of the points in the envelope divided by two and the absolute value of the wave.
+The rationale is that the values of the envelope divide by two should pass near the average of the absolute values of the wave.
+
+We can see from the table \ref{table:Error_comparison} that the method presented here consistently outperforms the traditional approaches for all waves tested, yielding an error 33\% smaller, on average, than the other methods.
+It is also worth noting that the errors of the traditional methods are very similar between them.
 
@@
 \centering
-\begin{tabular}{ l c c }
+\begin{tabular}{lcccc}
 \hline
-Method & LMS & time(s) \\
+\multicolumn{5}{c}{\textbf{Average error per frame}} \\
+              & Present Work & Smoothing & Low pass & Hilbert \\ 
 \hline
-Presented Algorithm & 0.0088 & 4.1507 \\ 
-Smoothing           & 0.0123 & 0.1911 \\ 
-Low pass Filter      & 0.0258 & 0.5601 \\ 
-Hilbert Transform   & 0.0141 & 0.2599 \\ 
+alto          & 0.073        & 0.083     & 0.083   & 0.082   \\
+bend          & 0.009        & 0.015     & 0.015   & 0.014   \\
+piano         & 0.005        & 0.006     & 0.006   & 0.005   \\
+brass         & 0.024        & 0.035     & 0.035   & 0.033   \\
+soprano       & 0.042        & 0.065     & 0.065   & 0.065   \\
+spoken voice  & 0.024        & 0.037     & 0.037   & 0.034   \\
+tom           & 0.003        & 0.005     & 0.005   & 0.005   \\
+sinusoid      & 0.113        & 0.193     & 0.195   & 0.191   \\ 
 \hline
+Average       & 0.037        & 0.055     & 0.055   & 0.054  
 \end{tabular}
-\caption{Comparison of the algorithm present in this work with the most common methods of digital envelope identification.}
-\label{table:Comparison}
+\caption{Comparison of the average errors of the algorithm presented in this work with the most common methods of digital envelope identification.
+The discrete waves used were normalized between -1 and 1.
+The average is between the absolute value of those waves and the envelope divided by two.}
+\label{table:Error_comparison}
 \end{table}
 
-% \subsection{Retrieving information about the underlying wave}
+The times taken for the algorithms compared here to process each wave are shown in \ref{table:Time_comparison}.
+We used specialized implementations available in libraries of the Python programming language.
+The presented algorithm is consistently the second-fastest.
 
-% The algorithm provides the position of the pseudo-cycles of a wave, with which one can, for example, derive its fundamental frequency and phase: using one of the two frontiers, we have $ f = (\#F^\pm - 1) n / (F^\pm[-1] - F^\pm[0])$, where $ \#F^\pm $ denotes the cardinality of the frontier, while $ F^\pm[0] $ and $ F^\pm[-1] $ stand for the frontier's first and last item, respectively.
-
-% Similarly, each point of each of the frontiers can be used to estimate the local phase at that location.
-For the positive frontier, it is necessary to note that, for each point, the following equation holds: $ \cos(\phi + 2 \pi f i / n) = 1 $, and thus $ \phi = 2 \pi k - 2 \pi f i / n $, with $ k \in \mathbb{N} $, while for the negative frontier $ \cos(\phi + 2 \pi f i / n) = -1 $ and $ \phi = 2 \pi k + \pi - 2 \pi f i / n) $.
+\begin{table}[ht!]
+\centering
+\begin{tabular}{lcccc}
+\hline
+\multicolumn{5}{c}{\textbf{Time in seconds}} \\
+              & Present Work & Smoothing & Low pass & Hilbert \\ 
+\hline
+alto          & 0.008        & 0.101     & 0.007   & 0.011   \\
+bend          & 0.030        & 0.226     & 0.012   & 0.131   \\
+piano         & 0.034        & 0.374     & 0.014   & 0.042   \\
+brass         & 0.032        & 0.297     & 0.011   & 0.045   \\
+soprano       & 0.177        & 0.735     & 0.027   & 0.566   \\
+spoken voice  & 0.014        & 0.105     & 0.003   & 0.038   \\
+tom           & 0.007        & 0.084     & 0.018   & 0.026   \\
+sinusoid      & 0.007        & 0.091     & 0.005   & 0.011   \\ 
+\hline
+Average       & 0.039        & 0.252     & 0.012   & 0.109  
+\end{tabular}
+\caption{Comparison of the processing time of the algorithm presented in this work with the most common methods of digital envelope identification.}
+\label{table:Time_comparison}
+\end{table}
 
@@
 
-In practice, is not uncommon for a discrete wave, especially in the case of sound, to present somewhat different positive and negative contours; in those cases, the algorithm here presented can be used independently in the positive and negative pulses of the wave to define two "envelopes", that we are going to call superior and inferior frontiers.
+In practice, is not uncommon for a discrete wave, especially in the case of sound, to present somewhat different positive and negative contours; in those cases, the algorithm here presented can be used independently in the positive and negative pulses of the wave to define two “envelopes”, that we are going to call superior and inferior frontiers.
 
@@
   \centering
-    \includegraphics{images/07FullFrontiers.pdf}
+    \includegraphics{images/08FullFrontiers.pdf}
   \caption{Positive and negative frontiers of six digital waves, as extracted by the algorithm here presented.
 For each wave, the region highlighted in black is shown in detail besides the whole wave.
-For each wave the horizontal axis is the sample number $i$, while the vertical axis is the normalized amplitude.
+For each wave, the horizontal axis is the sample number $i$, while the vertical axis is the normalized amplitude.
 All 6 waves were recorded at 44100 fps.}
@@
 
-The frontiers are satisfactorily detected in the case of harmonic and inharmonic sounds, and is robust in relation to the number of samples and the frequencies of the waves.
+The frontiers are satisfactorily detected in the case of harmonic and inharmonic sounds, and are robust in relation to the number of samples and the frequencies of the waves.
+That is a peculiarity of our algorithm: the traditional algorithms are unable to generate two distinct envelopes for the same wave.
 
-% The impact of the normalization of a wave using the frontiers obtained can be seen in figure \ref{fig:Fourier}, which shows the original and normalized wave in the case of a recording of the key 33 of a piano both in the time and frequency domains.
-We can see that the Fourier power spectra is greatly simplified in the case of the normalized wave.
+\section{Approximated location of pseudo-cycles}
 
-% \begin{figure}[ht!]
-%   \centering
-%     \includegraphics[width=0.8\linewidth]{08Fourier.pdf}
-%   \caption{Original and normalized views of a recording of key 33 (F3, 175 Hz) from a grand piano both in time and frequency domains.}
-%   \label{fig:Fourier}
-% \end{figure}
+The envelope is shown to add complexity to the spectral representation of a wave \parencite{2019TarjanoNeuro}, and an accurate description of the envelope, while describing the evolution of the instantaneous amplitude of a signal in time, would also simplify further spectral analysis.
+Figure \ref{fig:Fourier} shows the frequency-domain power spectrum for the wave and the carrier presented in \ref{fig:Envelope}.
+For the carrier, the power spectrum presents two nonzero values, while the power spectrum for the wave composed of the carrier modulated by the envelope is more complex.
 
-% \subsection{Robustness}
+\begin{figure}[ht!]
+  \centering
+    \includegraphics{images/11Fourier.pdf}
+  \caption{Fourier power spectrum for the wave and carrier shown in \ref{fig:Envelope}.}
+  \label{fig:Fourier}
+\end{figure}
 
-% Figures \ref{fig:xRobustness} and \ref{fig:yRobustness} illustrate the computational resilience of the algorithm for scaling in both the abscissa and ordinate axes, respectively.
-This is due to the normalization scheme presented in equation \ref{eq:scale}.
+Moreover, the algorithm developed in this work naturally divides a signal into its pseudo-cycles, pinpointing them in the time-domain, providing the building blocks for the reconstruction of the fine structure of the wave.
 
-% Although important from a theoretical standpoint, this step has an especial computational relevance in the case of digital sound waves, as they tend to be composed of a high number of samples with relatively low amplitudes, generally normalized between -1 and 1, leading to a numerical disparity between axes.
+\begin{figure}[ht!]
+  \centering
+    \includegraphics{images/09PCs.pdf}
+  \caption{Positive and negative frontiers of six digital waves, as extracted by the algorithm here presented.
+For each wave, the region highlighted in black is shown in detail besides the whole wave.
+For each wave, the horizontal axis is the sample number $i$, while the vertical axis is the normalized amplitude.
+All 6 waves were recorded at 44100 fps.}
+  \label{fig:PCs}
+\end{figure}
 
-% \begin{figure}[ht!]
-%   \centering
-%     \includegraphics[width=0.8\linewidth]{10xRobustness.pdf}
-%   \caption{Minimum and maximum scales of the abscissa before computational stability issues arose, for a recording of a drum kit tom.  $ \text{B} = 10^6 $.}
-%   \label{fig:xRobustness}
-% \end{figure}
+One can see in figure \ref{fig:PCs} a section of the plot of where the local maxima and minima occurs in a recording of an alto singer.
+The average between the two is also shown.
+With this information, one can split a discrete wave into its pseudo-cycles, as is illustrated in Figure \ref{fig:AvgPc}: It shows the superposition of this extracted pseudo-cycles, after being normalized by length, and their average.
+This same average is shown shifted, to exemplify that approximate $C^0$ and $C^1$ continuity is achieved.
 
-% \begin{figure}[ht!]
-%   \centering
-%     \includegraphics[width=0.8\linewidth]{09yRobustness.pdf}
-%   \caption{Minimum and maximum values of the amplitude before computational stability issues arose, for a recording of a drum kit tom.}
-%   \label{fig:yRobustness}
-% \end{figure}
+\begin{figure}[ht!]
+  \centering
+    \includegraphics{images/10AvgPc.pdf}
+  \caption{Positive and negative frontiers of six digital waves, as extracted by the algorithm here presented.
+For each wave, the region highlighted in black is shown in detail besides the whole wave.
+For each wave, the horizontal axis is the sample number $i$, while the vertical axis is the normalized amplitude.
+All 6 waves were recorded at 44100 fps.}
+  \label{fig:AvgPc}
+\end{figure}
 
-\section{Discussion}
-% Put results into a broader context.
+\section{Conclusion}
+This work fills a gap identified in the context of signal processing, where the lack of general procedures for the accurate identification of the temporal envelope of arbitrary waves consists of an obstacle to the complete description and eventual manipulation of signals.
 
-% What does it all mean?
+By presenting a general method for envelope detection that obtains information for his only variable parameter from the wave itself, thus completely eliminating the need for manual parameter tuning, and illustrating the efficiency of the algorithm on a wide range of real-world signals, we hope to facilitate future research in the many areas that rely on envelope detection techniques.
 
-% what principles have been established or reinforced
+With the suggestion of a new metric to evaluate the quality of an envelope, we hope to consolidate a path to a more solid theory of envelopes.
 
-This work hopes to fill a gap identified by the authors in the context of sound synthesis, where the lack of procedures for the accurate identification of the temporal envelope of arbitrary waves consists of an obstacle to the complete description and eventual manipulation of signals.
-
-While relevant in its own accord, the procedure here presented isolates the individual pseudo-cycles of a wave, pinpointing them in the time domain.
-Due to that, a rich set of theoretical advancements, of whose an example was given in the end of the preceding section, can be built upon this initial algorithm.
-
-With the knowledge of the exact position and characteristics of the pseudo-cycles of a wave, one can investigate its evolution in time, in terms of waveform, for example, free from the potential interference of a modulating wave; in other words, one can have a very satisfactory view of the instantaneous shape and phase of an arbitrary wave and its evolution in time, providing a way to bridge the gap between time domain and frequency domain interpretations.
-
-In the context of sound synthesis, for example, one would be able to apply specific methods for the recreation of the envelope, more in line with its smooth, slow varying and non-periodic nature and a different approach to the periodic and relatively fast changes of the temporal fine structure.
-
-\subsection{Conclusion}
-
-It is our opinion that such theoretical advances can be used in the formulation of better representations of discrete sounds, that in turn can improve sound compression techniques, while providing more adequate representation for the application of machine learning techniques for both sound classification and synthesis.
+While relevant in its own accord, the procedure here presented approximately isolates the individual pseudo-cycles of a wave, pinpointing them in the time-domain.
+Due to that, a rich set of theoretical advancements, especially in the areas of sound compression and synthesis, can be built upon this initial algorithm.
 

### 2020-12-05 Papers/01_Pulses_Envelope/Filter.py

@@ -1,0 +1,41 @@
+import plotly.graph_objects as go
+import signal_envelope as se
+import numpy as np
+from scipy.signal import butter, filtfilt
+
+def butter_lowpass_filter(data, fps, cutoff = 2000, order = 2):
+  nyq = 0.5 * fps
+  normal_cutoff = cutoff / nyq
+  # Get the filter coefficients 
+  b, a = butter(order, normal_cutoff, btype='low', analog=False)
+  y = filtfilt(b, a, data)
+  return y
+
+'''==============='''
+''' Read wav file '''
+'''==============='''
+name = "alto"
+W, fps = se.read_wav(f"C:/Users/tesse/Desktop/Files/Dropbox/0_Science/reps/envelope/test_samples/{name}.wav")
+W = W - np.average(W)
+amplitude = np.max(np.abs(W))
+W = W / amplitude
+n = W.size
+print(f"n={n}")
+X = np.arange(n)
+
+
+
+y = butter_lowpass_filter(W, fps)
+
+fig = go.Figure()
+fig.add_trace(go.Scatter(
+            y = W,
+            # line =  dict(shape =  'spline' ),
+            name = 'signal with noise'
+            ))
+fig.add_trace(go.Scatter(
+            y = y,
+            # line =  dict(shape =  'spline' ),
+            name = 'filtered signal'
+            ))
+fig.show()

### 2020-12-06 Papers/01_Pulses_Envelope/PulsesEnvelope.tex
revised

@@
 Intuitively, the temporal envelope can be understood as a slowly varying function that multiplies the signal, being responsible for its outer shape.
-Envelope detection techniques have applications in areas like health, sound classification and synthesis, seismology and speech recognition. 
+Envelope detection techniques have applications in areas like medicine, sound classification and synthesis, seismology and speech recognition.
 % Problem | Objectives
-Nevertheless, a general approach to digital envelope detection of signals with rich spectral content doesn't exist, as most methods involve manual intervention, in the form of filter design, smoothing, as well as other specific design choices, based on prior knowledge of the nature of the specific waves under investigation.
-To address this problem, we propose a framework that uses intrinsic characteristics of a signal to estimate its envelope, eliminating the necessity of parameter tuning.
+Nevertheless, a general approach to digital envelope detection of signals with rich spectral content doesn't exist, as most methods involve manual intervention, in the form of filter design, smoothing, and other specific design choices, based on prior knowledge of the signals under investigation.
+To address this problem, we propose an algorithm that uses intrinsic characteristics of a signal to estimate its envelope, eliminating the necessity of parameter tuning.
 % Methods | Procedure | Approach
-The approach here described draws inspiration from geometric concepts to estimate the temporal envelope of an arbitrary signal; specifically, a new measure of discrete curvature is used to obtain the average radius of curvature of a discrete wave, that will as a threshold to identify the wave's samples that are part of the envelope, in a manner similar to the alpha shapes approach to the identification of the concave hull of a set of points.
+The approach here described draws inspiration from geometric concepts to estimate the temporal envelope of an arbitrary signal; specifically, a new measure of discrete curvature is used to obtain the average radius of curvature of a discrete wave, that will serve as a threshold to identify the wave's samples that are part of the envelope.
 % Results | Findings | Product
@@
 We provide visualizations of the envelope extracted via the algorithm for various real-world signals, with very diverse characteristics, such as voice, spoken and sang, and pitched and non-pitched musical instruments, and discuss some approaches to assess the quality of the obtained envelopes.
-A Python module, installable via the Python Package Index (PyPI), with a pure Python and a specialized C++ implementation for 64-bit Windows systems, interactive visualizations of envelopes for a diverse range of digital waves, as well as the source code for the Python implementation, are available online at \textcite{2020TarjanoEnvelope}.
+A Python module implementing the algorithm was made available via the Python Package Index (PyPI); interactive visualizations of envelopes for a diverse range of digital waves, as well as the source code for the Python implementation, are available online at \textcite{2020TarjanoEnvelope}.
 % Conclusions | Implications
 Besides the most direct applications of this work to audio classification and synthesis, we foresee impact in compression techniques and machine learning approaches to audio.
+We briefly discuss some potential paths in this direction.
 The discrete curvature definition presented could also be extended to three-dimensional settings, to improve shape detection algorithms based on alpha shapes.
@@
 In many contexts, however, the temporal amplitude envelope of a signal plays a prominent role: according to \textcite{2017QiRelative}, for example, the envelope is at least as important as the fine structure of a sound wave in the context of the intelligibility of Mandarin tones.
-This is also the case for the English language \parencite{1995ShannonSpeech}, where even envelopes modulating mostly noise were still capable of conveying meaning.
+This is also the case for the English language \parencite{1995ShannonSpeech}, where even envelopes modulating mostly noise are still capable of conveying meaning.
 
-The envelope also helps to impart emotion and identity to the human voice \parencite{2018ZhuContributions}, and envelope preserving characteristics in concert halls are associated with their pleasantness \parencite{2011LokkiEngaging}.
+The envelope also helps to impart emotion and identity to the human voice \parencite{2018ZhuContributions} and music: Concert halls with envelope preserving characteristics are deemed more pleasant \parencite{2011LokkiEngaging}, for example.
 
@@
 
-One direct way of defining the shape of a set of points in $ \mathbb{R}^2 $ is using the convex hull: as defined in \textcite{2008BergComputational}, the convex hull of a set of points $ S $ in two-dimensional space is the smallest convex subset of the plane that contains all the points in $ S $.
+One direct way of defining the shape of a set of points in $ \mathbb{R}^2 $ is using the concept of a convex hull: as defined in \textcite{2008BergComputational}, the convex hull of a set of points $ S $ in two-dimensional space is the smallest convex subset of the plane that contains all the points in $ S $.
 Convex hulls are unique, in the sense that a set of points define one and only one convex hull.
 
-Generalizing that definition, \textcite{1983Edelsbrunnershape} introduced the concept of alpha shapes, that can be seen as a Concave Hull; a mathematically well-defined generalization of the Convex Hull of a finite set of points, closely related to the Delaunay triangulation and Voronoi diagrams of those points.
+Generalizing that definition, \textcite{1983Edelsbrunnershape} introduced the concept of alpha shapes, that can be seen as a concave hull; a mathematically well-defined generalization of the convex hull of a finite set of points, closely related to the Delaunay triangulation and Voronoi diagrams of those points.
 
-The intuition behind the alpha shapes algorithm is that only points that can be touched by a circle of radius $ \alpha $ coming from an infinite distance towards the set of points of interest are part of the frontier, that is, of the Concave Hull of this set.
+The intuition behind the alpha shapes algorithm is that only points that can be touched by a circle of radius $ \alpha $ coming from an infinite distance towards the set of points of interest are part of the frontier, that is, of the concave hull of this set.
 
-Once an $ \alpha $ parameter is defined alpha shapes conserve the uniqueness property, besides reducing to the Convex Hull when $ \alpha = \infty $.
-Consider, for example, the artificially generated discrete wave in figure \ref{fig:Hulls}
-below, where the links between consecutive samples were deemphasized to highlight the “set of points” nature of a signal: The Convex Hull and a Concave Hull, in the form of an alpha shape with a particular value of $ \alpha $ are illustrated.
+Once a radius $ \alpha $ is defined, alpha shapes conserve the uniqueness property, in fact tending to the convex hull when $ \alpha \to \infty $.
+Consider, for example, the artificially generated discrete wave in figure \ref{fig:Hulls} below, where the links between consecutive samples were deemphasized to highlight the “set of points” nature of a signal: The convex hull and a concave hull, in the form of an alpha shape with a particular value of $ \alpha $ are illustrated.
 
@@
     \includegraphics{images/00ConvexHull.pdf}
-  \caption{A discrete signal interpreted as a set of points in the Cartesian plane, and the Convex Hull and Concave Hull at $ \alpha=300 $ defined by it.}
+  \caption{A discrete signal interpreted as a set of points in the Cartesian plane, and the convex and concave hull at $ \alpha=300 $ defined by it.}
   \label{fig:Hulls}
@@
 
-The alpha shapes approach is used in areas such as the detection of features in images \parencite{2016VarytimidisAlpha}, reconstruction of surfaces from a cloud of points \parencite{2015WuAutomated} and Spectroscopy \parencite{2019XuModeling}, with the last work, which involves the estimation and removal of the Blaze function (a kind of envelope) of an echelle spectrograph, being particularly illustrative of the potential of synergy between geometric and DSP approaches.
+The alpha shapes approach is used in areas such as detection of features in images \parencite{2016VarytimidisAlpha}, reconstruction of surfaces from a cloud of points \parencite{2015WuAutomated} and spectroscopy \parencite{2019XuModeling}.
 
-Other steps in the direction of translating geometric algorithms to the context of envelope detection were also made by \textcite{2015YangSkeleton} via an algorithm based on the construction of a skeleton underlying the digital wave of interest, and also via the direct translation of computer vision methods to the task of envelope detection \parencite{2015YangRepresenting}.
+This last work, which involves the estimation and removal of the Blaze function (a kind of envelope) of an echelle spectrograph, being particularly illustrative of the potential synergy between geometric and DSP approaches.
+
+Other steps in the direction of translating geometric algorithms to the context of envelope detection were made by \textcite{2015YangSkeleton} via an algorithm based on the construction of a skeleton underlying the digital wave of interest, and also via the direct translation of computer vision methods to the task of envelope detection \parencite{2015YangRepresenting}.
 
@@
 
-The rest of this paper is divided as follows: 
-We start by trying to define as specifically as possible what a temporal envelope is, especially in the context of this work, providing a mathematical definition, in terms of the Hilbert transform, for the case of a narrowband signal.
+The rest of this paper is divided as follows: We start by defining as specifically as possible what a temporal envelope is, especially in the context of this work and providing a mathematical definition, in terms of the Hilbert transform, for the case of a narrowband signal.
 
-We proceed to present the representation of a discrete wave used in this work, largely based on the concept of dividing the wave into its constituent pulses, and how it can simplify the algorithm, reducing the high dimensionality generally present by digital representations of waves.
+We proceed to illustrate the representation of a discrete wave used in this work, largely based on the concept of dividing the wave into its constituent pulses, and how it can simplify the algorithm, reducing the high dimensionality generally present in digital representations of waves.
 
-Next, we review methods for the estimation of the curvature of a discrete wave, after which we present our own; this method will then be used to estimate the equivalent circle, that will be used to define the envelope of the wave.
+Next, we review methods for the estimation of the curvature of a discrete wave, after which we present our own; this method will then be used to estimate the equivalent circle, that will be used to define the envelope of the wave, via the procedure described in sequence.
 
 In the results section, we present some illustrations of the envelopes obtained with our method, comparing both accuracy and running time with the most common algorithms encountered in the literature.
-We also suggest an alternative use of the algorithm to identify the negative and positive frontiers of a wave, a kind of partial envelope considering only the positive and negative pulses of a wave, respectively.
+We also suggest an alternative use of the algorithm to identify the negative and positive frontiers of a wave, that can be seen as partial envelopes considering only the positive and negative pulses of a wave, respectively.
 
-We follow with some suggestions about how the approach here presented can be useful beyond envelope detection, exemplifying with some applications, such as the extractions of the average waveform of a digital wave, concluding with some commentary on potential future developments of the research here presented.
+We follow with suggestions about how the approach here presented can be useful beyond envelope detection, exemplifying with some applications, such as the extractions of the average waveform of a digital wave, and concluding with commentary on potential future developments of the method.
 
@@
 
-The interest in the theory of envelope detection arose in the analogue domain, with the widespread adoption of radio communications, where the problem is typically restricted to simple sinusoids, often with well-defined frequencies \parencite{2011TurnerDemodulation}.
+The interest in the theory of envelope detection arose in the analog domain, with the widespread adoption of radio communications, where the problem is typically restricted to simple sinusoids, often with well-defined frequencies \parencite{2011TurnerDemodulation}.
 
-In this context, the most mathematically sound definition of an envelope involves the representation of its underlying wave as an analytic signal, introduced by Gabor in 1946 \parencite{2007HahnHistory}.
+In this context, the most mathematically sound definition of an envelope involves the representation of its underlying wave as an analytic signal, as introduced by Gabor in 1946 \parencite{2007HahnHistory}.
 In his work, \textcite{1946GaborTheory} applies the then relatively new mathematical machinery of the quantum mechanics to unify the time and frequency-domain representations of a wave, showing how the Hilbert transform could be applied to a real signal in order to obtain an equivalent complex signal, later known as the analytic signal.
-
 This analytic signal has the form $ A(t) = S(t) + \mathbb{H}(S(t)) \ \text{i} $ \parencite{2016HePraat} where $ S(t) $ is the original real signal.
@@
 
-Despite its widespread use, and effectiveness in the context of narrowband signals, envelope detection techniques based on the Hilbert transform don't behave well for broadband signals \parencite{2012DauSpeech}, being even physically
-paradoxical in some cases \parencite{1996Loughlinamplitude}.
+Despite its widespread use, and effectiveness in the context of narrowband signals, envelope detection techniques based on the Hilbert transform don't behave well for broadband signals \parencite{2012DauSpeech}, being even physically paradoxical in some cases \parencite{1996Loughlinamplitude}.
 
-Figure \ref{fig:AnalyticSignal} exemplifies that, showing the envelope obtained via the analytic signal, for a pure sinusoid with a local frequency of 20 cycles, modulated by a polynomial of the third degree, illustrating the difference in the shape of the envelope in the presence and absence of Gaussian white noise with standard deviation of $ 1/10 $ of the wave's maximum amplitude.
-
-It is readily noticeable from figure \ref{fig:AnalyticSignal} that, as soon as noise is introduced in the original signal, it is reflected in the envelope.
-
-%%%%%%%%%%%%%%%%%
+Figure \ref{fig:AnalyticSignal} exemplifies the Hilbert envelope, for a pure sinusoid modulated by a polynomial, illustrating the difference in the shape of the envelope in the presence and absence of Gaussian white noise.
+It is readily noticeable from the figure that, as soon as noise is introduced in the original signal, it is reflected in the envelope.
 
@@
     \includegraphics{images/01AnalyticSignal.pdf}
-  \caption{Envelope of a pure sinusoid modulated by a polynomial of degree 3, as obtained by the Hilbert Transform approach.
+  \caption{Envelope of a pure sinusoid with a local frequency of 20 cycles modulated by a polynomial of degree 3, as obtained by the Hilbert Transform approach.
 In the first half, the sinusoid is free of noise, while in the second half Gaussian noise with a standard deviation of $ 1/10 $ of the maximum amplitude of the wave was added to the base signal.}
@@
 
-Also, in this work we are concerned with the discrete version of this problem: given a finite digital wave represented by the vector $ \textbf{w} \in \mathbb{R}^n $, instead of the continuous function $ W(t) $, obtaining the temporal envelope of $ \textbf{w} $.
+In this work we are concerned with the discrete version of this problem: given a finite digital wave represented by the vector $ \textbf{w} \in \mathbb{R}^n $, instead of the continuous function $ W(t) $, obtaining the temporal envelope of $ \textbf{w} $.
 
@@
   & \textbf{c} = (c_0, c_1, \cdots, c_{n-1}) \\
-  & \textbf{w} = \textbf{e} \odot \textbf{c} \\
-  & w_i = e_i c_i \quad \forall \quad i, 0 \le i \le n-1
+  & \textbf{w} = \textbf{e} \odot \textbf{c} \quad \therefore \quad w_i = e_i c_i \quad \forall \quad i, 0 \le i \le n-1
 \end{align}
 
-The $ \odot $ operator in \ref{eq:Envelope} stands for the Hadamard product, denoting element wise multiplication of two vectors.
-Figure \ref{fig:Envelope} provides an example of the vectors just defined: \textbf{w} is obtained by modulating the normalized carrier wave \textbf{c} using the smooth envelope \textbf{e}.
+The $ \odot $ operator in \ref{eq:Envelope} stands for the Hadamard product, denoting element-wise multiplication of two vectors.
+Figure \ref{fig:Envelope} provides an example of the vectors just defined: \textbf{w} is obtained by modulating the normalized carrier wave \textbf{c} using the envelope \textbf{e}.
 
-Note from Figure \ref{fig:Envelope} that, at some local maxima, the envelope is defined, being in fact equal to the value of the wave \textbf{w}, since $ c_i = 1 $ at those points, rendering $ e_i = 1 w_i $.
-The carrier wave \textbf{c} is periodic, and normalized between $ \{-1, 1\}$.
-Each cycle of the wave, then, reaches the value 1 only at its maximum making $ w_i$ equal to the value of the envelope at that point.
+Note from Figure \ref{fig:Envelope} that, at some local maxima, the envelope is well-defined, being in fact equal to the value of the wave $ w_i $, since $ c_i = 1 $ at those points, rendering $ e_i = 1 w_i $.
 
-Image \ref{fig:Envelope}, along with the Concave Hull illustrated in figure \ref{fig:Hulls}, helps forming the intuition behind the method developed in this work: the idea is to identify the local extrema that touch the envelope (or, to put in another way, the local extrema in relation to the whole wave that are global extrema inside their own pseudo-cycles), filtering out those who don't; in this way, we're defining the envelope itself at those points.
+The carrier wave \textbf{c} in the figure is periodic, and normalized between $ \{-1, 1\}$.
+Each cycle of the wave, then, reaches the value 1 only at its maximum making $ w_i $ equal to the value of the envelope at one point in each cycle.
 
-In this work, this is accomplished with the use of a circle of a defined radius, large enough to prevent it from reaching the smallest local extrema when "rolled" around the (geometric) representation of the wave: only the extrema touched by the circle will be marked as part of the envelope.
+Image \ref{fig:Envelope}, along with the concave hull illustrated in figure \ref{fig:Hulls}, helps forming the intuition behind the method developed in this work: the idea is to identify the local extrema that touch the envelope, those being also the extrema of each (pseudo-)cycle.
 
-In the next section, such a geometric representation will be formulated, along with the necessary tool to estimate the average curvature of the wave and, consequently, the appropriate radius of the circle.
+This is accomplished with the use of a circle with a defined radius, large enough to prevent it from reaching the smallest local extrema when "rolled" around the (geometric) representation of the wave.
+Only the extrema touched by the circle will be marked as part of the envelope.
+
+In the next section, a compatible geometric representation will be formulated, along with the necessary tools to estimate the average curvature of the wave and, consequently, the appropriate radius of the circle.
 
@@
     \includegraphics{images/02Envelope.pdf}
-  \caption{Example of a discrete wave \textbf{w} composed by the element wise multiplication of an envelope \textbf{e} and a carrier \textbf{c}.
+  \caption{Example of a discrete wave \textbf{w} composed by the element-wise multiplication of an envelope \textbf{e} and a carrier \textbf{c}.
 The local extrema of \textbf{w} are highlighted with a circle.}
@@
 
-Instrumental to the method here presented is the definition of a pulse: each time the value of \textbf{w} changes sign, that is, every time the discrete wave \textbf{w} represents crosses the horizontal axis, the beginning of a pulse is defined, with the next crossing defining its end.
+Instrumental to the method here presented is the definition of a pulse: each time the value of \textbf{w} changes sign, that is, every time the discrete wave \textbf{w} crosses the horizontal axis, the beginning of a pulse is defined, with the next crossing defining its end.
 This definition is in line with the one presented in \textcite{national1996telecommunications}, where a pulse is defined as a rapid change in the amplitude of a signal, followed by a fast return to the baseline value; zero, in our case.
 
-From \ref{fig:Envelope} we can see that only the maximum point of a positive pulse or the minimum point of a negative one can potentially be equal to the envelope and are, thus, our only points of interest.
-We can then proceed with the rest of the method considering only those points, to great computational advantage.
+From \ref{fig:Envelope} and the discussion in the preceding chapter, we can see that only the maximum point of a positive pulse or the minimum point of a negative one can potentially be equal to the envelope and are, thus, our only points of interest.
+We can then proceed, for the rest of the method, considering only those points, to great computational economy.
 
@@
 
-This relation is illustrated in figure \ref{fig:Pulses}. We call it P to emphasize the connection with the pulses of \textbf{w}, noting that P is, in fact, a set of points.
+This relation is illustrated in figure \ref{fig:Pulses}. We call it P to emphasize the connection with the pulses of \textbf{w}.
 
@@
 
-P is not fit for geometric interpretation yet, since the abscissa and ordinate of his orthogonal coordinate system have different units: in the vertical axis, we have a unit related to the instantaneous intensity of the wave, while in the horizontal axis we have the index at which each extremum occurs, ultimately a time-related unit.
+The points in P are not fit for geometric interpretation yet, since the abscissa and ordinate of their orthogonal coordinate system have different units: in the vertical axis, we have a unit related to the instantaneous intensity of the wave, while in the horizontal axis we have the index at which each extremum occurs, ultimately a time-related unit.
 
@@
 
-In this new coordinate system we have a basis vector $ \mathbf{b}^\prime = \{ \mathbf{b}^\prime_x, \mathbf{b}^\prime_y \}$ where $ \mathbf{b}^\prime_x = (i, 0) $ and $\mathbf{b}^\prime_y = (0, i) $, related to the original basis vector by:
+In this new coordinate system we have a basis vector $ \mathbf{b}^\prime = \{ \mathbf{b}^\prime_x, \mathbf{b}^\prime_y \}$ where $ \mathbf{b}^\prime_x = (i, 0) $ and $\mathbf{b}^\prime_y = (0, i) $, related to the original basis vector by equation \ref{eq:Basis}:
 
@@
 The effect of this normalization can be seen in the points shown in figure \ref{fig:Vectors}, shown in scale in both coordinate systems.
-The set V of the vectors from each point of P to the next is also shown in the picture, as it will be used to estimate the average curvature of P, we define this set below.
+The set V of the vectors from each point of P to the next is also shown in the picture, and it will be used to estimate the average curvature of P.
+We define this set in equation \ref{eq:Vectors} below.
 
@@
 Concerning direct methods \textcite{2014CarrollSurvey}, for example, derive three such definitions based on the approximation of a circle by an inscribed, centered and circumscribed polygon.
-In the context of tridimensional meshes, \textcite{2016VasaMultivariate} evaluate a range of existing estimators from a multivariate point of view.
+In the context of three-dimensional meshes, \textcite{2016VasaMultivariate} evaluate a range of existing estimators from a multivariate point of view.
 
-Those approaches define the discrete curvature in relation to the vertices of a wave.
+Those approaches define the discrete curvature in relation to the vertices of a wave (or mesh, in the three-dimensional case).
 In our particular case, it wouldn't make sense to talk about the curvature of a single pulse, since it is what our points ultimately represent.
@@
 The rationale is to find, for each vector $ v_k \in \text{V} $, the radius $ r_k $ of the equivalent circle whose tangent has the same change in direction, in the same horizontal distance, as the vector of interest, as shown in figure \ref{fig:DiscreteCurvature}.
-The average curvature of the P will then be obtained by the average of all radii $ r_k $.
+The average curvature of P will then be obtained by the average of all radii $ r_k $.
+
+From figure \ref{fig:DiscreteCurvature} is easy to see that $ r_k = v_{k,x} \sin(\theta_k)$ where $ v_{k,x} $ is the component of $ v_k $ in the horizontal direction; $ \theta_k $ itself can be obtained using the slope $ m_0 = 0 $ of a line in the horizontal direction and the slope $ m_k = v_{k,y} / v_{k,x} $ defined by $ v_k $ via the equality $ \theta_k = \arctan \big( (m_k - m_0) / (1 + m_k m_0) \big) $.
 
@@
     \includegraphics{images/05DiscreteCurvature.pdf}
-  \caption{The tangent unit vector of the circle changes from the horizontal direction in $ \hat{u}_0 $ to an inclination of $ \theta $ in $ \hat{u}_1 $, $ \theta $ being the angle that the vector $ v_k $ makes with the horizontal direction.}
+  \caption{The tangent unit vector of the circle changes from the horizontal direction in $ \hat{u}_0 $ to an inclination of $ \theta_k $ in $ \hat{u}_1 $, $ \theta_k $ being the angle that the vector $ v_k $ makes with the horizontal direction.}
   \label{fig:DiscreteCurvature}
@@
 
-From figure \ref{fig:DiscreteCurvature} is easy to see that $ r_k = v_{k,x} \sin(\theta_k)$ where $ v_{k,x} $ is the component of $ v_k $ in the horizontal direction; $ \theta_k $ itself can be obtained using the slope $ m_0 = 0 $ of any line in the horizontal direction and the slope $ m_k = v_{k,y} / v_{k,x} $ defined by $ v_k $ via the equality $ \theta_j = \arctan \big( (m_k - m_0) / (1 + m_k m_0) \big) $.
-
-The radius of the circle that represents the average curvature of V, and will be used to obtain the envelope of the discrete wave \textbf{w} can be obtained via equation \ref{eq:Radius}:
+The radius of the circle that represents the average curvature of V (and P), and will be used to obtain the envelope of the discrete wave \textbf{w} can be obtained via equation \ref{eq:Radius}:
 
@@
 We need now to construct an algorithm, using the radius just obtained, to identify the points that belong to the envelope.
-Generally, the algorithm for the alpha shapes approach resorts to the construction of the Delaunay triangulation of the set of points, that is latter filtered.
+Generally, the algorithm for the alpha shapes approach resorts to the construction of the Delaunay triangulation of the set of points, that is latter filtered to contain only the outer edges.
 We can adopt a more straightforward method, as our points have a defined structure:
@@
 For all circles of center $ x_o, y_o $ and radius $ r $ that can be constructed to pass through any two different points of P, that is, $ \forall \quad P_b, P_c $ such that $ (P_{b,x} - x_o)^2 + (P_{b,y} - y_o)^2 = (P_{c,x} - x_o)^2 + (P_{c,y} - y_o)^2 = r^2 $, $ P_a $ will be part of the set $ E $ of the points belonging to the envelope if and only if none of those circles contain $ P_a $, or $ P_a \in \text{E} \leftrightarrow (P_{b,x} - x_o)^2 + (P_{b,y} - y_o)^2 > r^2 $.
-Formalizing, we have:
+Summarizing this, we have equation \ref{eq:InEnvelope}:
 
@@
 
-The algorithm follows directly from the definition in \ref{eq:InEnvelope} after noting that the first and last members of P will always be part of the envelope, since they are reachable a circle of any radius $ r $.
+The algorithm follows directly from the definition in \ref{eq:InEnvelope} after noting that the first and last members of P will always be part of the envelope, since they are reachable by a circle of any radius $ r $.
 From $ P_0 $, our first pivot point, we construct circles passing through $ P_1, P_2, \cdots, P_d $ until a circle that doesn't contain any point in P is found; $ P_d $ then becomes the new pivot point and is included in E, and this procedure continues until $ P_{m-1} $ is reached.
 
-The procedure is formalized in algorithm \ref{FindEnvelope}. We don't provide similar algorithmic descriptions of the preceding steps since the Python implementation can easily serve as pseudocode.
+The procedure is formalized in the algorithm \ref{FindEnvelope}.
+We don't provide similar algorithmic descriptions of the preceding steps since the Python implementation can easily serve as pseudocode.
 
@@
 It is worth noting, however, that many other procedures are available, both for the interpolation and the smooth approximation of points.
-We deliberately choose linear interpolation to avoid blurring the results of the method.
 
@@
 Given the lack of a precise definition of what an envelope should be, we begin this section suggesting an empirical metric, based on the behaviour of the carrier wave \textbf{c}.
-We also comment on some metrics presented in the literature, and compare our algorithm with traditional demodulation approaches, both in relation to execution time and accuracy; to this last end, a numerical indicator is introduced.
+We also comment on some metrics presented in the literature, and compare our algorithm with traditional demodulation approaches, both in relation to execution time and accuracy; to this last end, a numerical indicator is also introduced.
 
@@
 There are no agreed-upon objective metrics to assess the quality of an envelope.
-Recalling figure \ref{fig:Envelope}, however, and the accompanying discussion, an intuitive way is to consider the carrier wave \textbf{c}, and how well it seems to be bounded in amplitude in the interval $ \{-1, 1\} $.
-To that end, we provide the following figures of original discrete waves \textbf{w}, their envelope \textbf{e}, and the inferred carrier, obtained after dividing, element-wise, the original wave by its envelope.
+Recalling figure \ref{fig:Envelope}, however, and the accompanying discussion, an intuitive way is to consider the carrier wave \textbf{c}, and how well it seems to be bounded by the interval $ \{-1, 1\} $.
+
+To illustrate how the algorithm performs on such metric, we provide the following figures of original discrete waves \textbf{w}, their extracted envelope \textbf{e}, and the inferred carrier, obtained after dividing, element-wise, the original wave by its envelope.
 
 From figure \ref{fig:spoken_voice} it can be seen that the algorithm performs well under the carrier metric.
-This is especially remarkable for the case of spoken voice, a notorious complex sound with fast changes in pitch and high noise content.
+This is especially remarkable for the case of spoken voice, a notorious complex wave with fast changes in pitch and high noise content.
 
-These characteristics are in stark contrast with the sang voice in figure \ref{fig:singer_voice}, where a recording of an alto singer is shown; there, the wave is almost periodic, with a stable waveform.
+These characteristics are in stark contrast with those of the sang voice in figure \ref{fig:singer_voice}, where a recording of an alto singer is shown; there, the wave is almost periodic, with a stable waveform.
 
@@
     \includegraphics{images/06Graph_spoken_voice.pdf}
-  \caption{Signal, envelope and carrier for a record of male voice uttering the word "amazing".}
+  \caption{Signal, envelope and carrier for a record of male voice uttering the word “amazing”.}
   \label{fig:spoken_voice}
@@
 
-The first one states that if a signal is bounded in magnitude, then its envelope should be, as well.
+The first one states that, if a signal is bounded in magnitude, then its envelope should be, as well.
 Our algorithm complies with this requisite, as the envelope is composed of selected samples from the original signal itself.
@@
 
-Nevertheless, if we picture both the wave and the envelope as samples from continuous functions, is easy to see that this t is the case for our algorithm; since the envelope is composed by some samples of the digital wave, in the limit where all the samples of the wave are also part of the envelope, we have that the maximum frequency for both is the same.
+Nevertheless, if we picture both the wave and the envelope as samples from continuous functions, is easy to see that this is the case for our algorithm; since the envelope is composed by some samples of the digital wave, in the limit where all the samples of the wave are also part of the envelope, we have that the maximum frequency for both is the same.
 
@@
 The third condition states what, for a periodic signal, the envelope should be a straight line with the intercept equal to the amplitude of this signal.
-Since all the local maxima have the same amplitude, our envelope will indeed be a straight line, with the same amplitude as the original wave.
+Since all the local maxima have the same amplitude in the case of a periodic wave, our envelope will indeed be a straight line, with the same amplitude as the original wave.
+An example of such behaviour can be observed in figure \ref{fig:Comparison_sinusoid}.
 
@@
 
-Direct comparison with many of the more recent algorithms is made difficult by the unavailability of digital implementations of such works, many designed to process analogue signals \parencite[e.g.,][]{2018AssefModeling}.
+Direct comparison with many of the more recent algorithms is made difficult by the unavailability of digital implementations of such works, many designed to process analog signals \parencite[e.g.,][]{2018AssefModeling}.
 
@@
 \begin{enumerate}
-\item Smoothing: This is a relatively simple procedure that, applied to the absolute values of a signal, can provide an approximation of the variation of its amplitude in time, by smoothing the movement of the carrier.
+\item Smoothing: This is a relatively simple procedure that, applied to the absolute values of a discrete signal, can provide an approximation of the variation of its amplitude in time.
 We use the Savitzky-Golay algorithm \parencite{1964SavitzkySmoothing} with a window width of 3001 samples and a cubic polynomial.
@@
 In the case of the Hilbert transform, pre-processing in the form of filtering was applied.
-That is in contrast with our method, that organically tunes itself, in the form of the automatic choice of the radius of the circle representing the average discrete curvature of the discrete signal.
+That is in contrast with our method, that organically tunes itself via the automatic choice of the radius of the circle representing the average discrete curvature of the signal.
 
@@
 The average error was calculated by the average of the squared distance of the points in the envelope divided by two and the absolute value of the wave.
-The rationale is that the values of the envelope divide by two should pass near the average of the absolute values of the wave.
+The rationale is that the values of the envelope divided by two should pass near the average of the absolute values of the wave.
+This metric is only valid for reasonably smooth envelopes, however.
 
@@
 We used specialized implementations available in libraries of the Python programming language.
-The presented algorithm is consistently the second-fastest.
+The presented algorithm is in average the second-fastest.
 
@@
 For each wave, the region highlighted in black is shown in detail besides the whole wave.
-For each wave, the horizontal axis is the sample number $i$, while the vertical axis is the normalized amplitude.
+For each wave, the horizontal axis is the sample number $ i $, while the vertical axis is the normalized amplitude.
 All 6 waves were recorded at 44100 fps.}
@@
 
-Moreover, the algorithm developed in this work naturally divides a signal into its pseudo-cycles, pinpointing them in the time-domain, providing the building blocks for the reconstruction of the fine structure of the wave.
+Moreover, the algorithm developed in this work naturally divides a signal into its pseudo-cycles, pinpointing them in the time-domain, thus providing the building blocks for the reconstruction of the fine structure of the wave.
 
@@
     \includegraphics{images/09PCs.pdf}
-  \caption{Positive and negative frontiers of six digital waves, as extracted by the algorithm here presented.
-For each wave, the region highlighted in black is shown in detail besides the whole wave.
-For each wave, the horizontal axis is the sample number $i$, while the vertical axis is the normalized amplitude.
-All 6 waves were recorded at 44100 fps.}
+  \caption{Section of the lines formed by the positions of the positive and negative frontiers of the digital wave illustrated in \ref{fig:singer_voice}, as extracted by the algorithm presented, and their average.}
   \label{fig:PCs}
@@
 
-One can see in figure \ref{fig:PCs} a section of the plot of where the local maxima and minima occurs in a recording of an alto singer.
+One can see in figure \ref{fig:PCs} a section of the plot of where the local maxima and minima occurs in the recording of an alto singer show in \ref{fig:singer_voice}.
 The average between the two is also shown.
@@
     \includegraphics{images/10AvgPc.pdf}
-  \caption{Positive and negative frontiers of six digital waves, as extracted by the algorithm here presented.
-For each wave, the region highlighted in black is shown in detail besides the whole wave.
-For each wave, the horizontal axis is the sample number $i$, while the vertical axis is the normalized amplitude.
-All 6 waves were recorded at 44100 fps.}
+  \caption{Pseudo-cycles of the wave illustrated in \ref{fig:singer_voice}, segmented and superposed, and their average.
+The average is also shown shifted by one average period, to illustrate the average continuity.}
   \label{fig:AvgPc}
@@
 \section{Conclusion}
-This work fills a gap identified in the context of signal processing, where the lack of general procedures for the accurate identification of the temporal envelope of arbitrary waves consists of an obstacle to the complete description and eventual manipulation of signals.
+This work fills a gap identified in the theory of digital signal processing, where the lack of general procedures for the accurate identification of the temporal envelope of arbitrary waves poses an obstacle to the complete description and eventual manipulation of signals.
 
-By presenting a general method for envelope detection that obtains information for his only variable parameter from the wave itself, thus completely eliminating the need for manual parameter tuning, and illustrating the efficiency of the algorithm on a wide range of real-world signals, we hope to facilitate future research in the many areas that rely on envelope detection techniques.
+By presenting a general approach for envelope detection, and eliminating the need for parameter tuning, we expect to facilitate further research at the many areas that rely on envelope detection techniques, and improve already existing algorithm that have envelope detection as an intermediary step.
 
-With the suggestion of a new metric to evaluate the quality of an envelope, we hope to consolidate a path to a more solid theory of envelopes.
+Illustrating the efficiency of an algorithm inspired by geometric aspects of a discrete signal on a wide range of real-world signals, we also hope to encourage research in this direction.
+
+The present work also highlights some gaps in our current understanding and definition of temporal envelope and, in so doing, endeavors to motivate further research in this topic, that can ultimately give rise to a more mathematically sound envelope theory, especially concerning broadband waves; suggestions in this direction are also presented.
 
 While relevant in its own accord, the procedure here presented approximately isolates the individual pseudo-cycles of a wave, pinpointing them in the time-domain.
-Due to that, a rich set of theoretical advancements, especially in the areas of sound compression and synthesis, can be built upon this initial algorithm.
+
+Similarly, by decomposing a wave into its envelope and carrier, and simplifying the representation of such carrier in the frequency-domain, a rich set of advancements, especially in the areas of sound compression and synthesis in the time and frequency-domains, can be built upon this initial algorithm.
+
+By illustrating how this information can be used to obtain the average waveform of a discrete wave, we hope to encourage research in this area, ultimately leading to alternative representations of discrete waves.
 

### 2020-12-06 Papers/01_Pulses_Envelope/PulsesEnvelope.tex
last adjustments

@@
 
-% \usepackage{hyperref}
-% \usepackage{hypcap}
-% \hypersetup{
-%   colorlinks=true, 
-%   allcolors=black,
-%   pdfauthor={Carlos Henrique Tarjano Santos},
-%   pdftitle={Robust Digital Envelope Estimation Via Geometric Properties of an Arbitrary Real Signal}
-% }
+\usepackage{hyperref}
+\usepackage{hypcap}
+\hypersetup{
+  colorlinks=true, 
+  allcolors=black,
+  pdfauthor={Carlos Henrique Tarjano Santos},
+  pdftitle={Robust Digital Envelope Estimation Via Geometric Properties of an Arbitrary Real Signal}
+}
 
@@
 
-{\bf Keywords:} DSP, alpha shapes, envelope detection, discrete curvature estimation, demodulation
+{\bf Keywords:} DSP; alpha shapes; envelope detection; discrete curvature estimation; demodulation
 
@@
 
-\section{Methodology}
+\section{Methods}
 After defining what will be considered an envelope in the context of this work, in this section we explore some simplifications that can be applied to a discrete wave, defining the representation that will be used in the subsequent part of the work.
@@
 \begin{align} \label{eq:Envelope}
-  & \textbf{w}, \textbf{e}, \textbf{c} \in \mathbb{R}^n \\
-  & \textbf{w} = (w_0, w_1, \cdots, w_{n-1}) \\
-  & \textbf{e} = (e_0, e_1, \cdots, e_{n-1}) \\
-  & \textbf{c} = (c_0, c_1, \cdots, c_{n-1}) \\
-  & \textbf{w} = \textbf{e} \odot \textbf{c} \quad \therefore \quad w_i = e_i c_i \quad \forall \quad i, 0 \le i \le n-1
+  & \mathbf{w}, \mathbf{e}, \mathbf{c} \in \mathbb{R}^n \\
+  & \mathbf{w} = (w_0, w_1, \cdots, w_{n-1}) \\
+  & \mathbf{e} = (e_0, e_1, \cdots, e_{n-1}) \\
+  & \mathbf{c} = (c_0, c_1, \cdots, c_{n-1}) \\
+  & \mathbf{w} = \mathbf{e} \odot \mathbf{c} \ \therefore \ w_i = e_i c_i \ \forall \ i, \ 0 \le i \le n-1
 \end{align}
@@
 
-\section{Results}
+\section{Results and Discussion}
 Given the lack of a precise definition of what an envelope should be, we begin this section suggesting an empirical metric, based on the behaviour of the carrier wave \textbf{c}.
@@
 
-\section{Approximated location of pseudo-cycles}
+\subsection{Approximated location of pseudo-cycles}
 
@@
 
-The present work also highlights some gaps in our current understanding and definition of temporal envelope and, in so doing, endeavors to motivate further research in this topic, that can ultimately give rise to a more mathematically sound envelope theory, especially concerning broadband waves; suggestions in this direction are also presented.
+The present work also highlights some gaps in our current understanding and definition of temporal envelope and, in so doing, endeavours to motivate further research in this topic, that can ultimately give rise to a more mathematically sound envelope theory, especially concerning broadband waves; suggestions in this direction are also presented.
 

### 2020-12-06 Papers/01_Pulses_Envelope/00ElsevierPulsesEnvelope.tex
Elsevier Digital Signal Processing preparation

--- 2020-12-06 Papers/01_Pulses_Envelope/PulsesEnvelope.tex
@@
-\documentclass{article}
-\usepackage{lmodern} % Latin Modern Font
-\usepackage[T1]{fontenc}
-\usepackage{authblk}
+\documentclass{elsarticle}
+% \usepackage{lmodern} % Latin Modern Font
+% \usepackage[T1]{fontenc}
+% \usepackage{authblk}
 \usepackage{amsfonts} % \mathbb
 \usepackage[margin=0.5in]{geometry}
-\usepackage[utf8]{inputenc}
-\usepackage{graphicx}
+% \usepackage[utf8]{inputenc}
+% \usepackage{graphicx}
 \usepackage[backend=biber]{biblatex}
@@
 
-\usepackage{hyperref}
-\usepackage{hypcap}
-\hypersetup{
-  colorlinks=true, 
-  allcolors=black,
-  pdfauthor={Carlos Henrique Tarjano Santos},
-  pdftitle={Robust Digital Envelope Estimation Via Geometric Properties of an Arbitrary Real Signal}
-}
+% \usepackage{hyperref}
+% \usepackage{hypcap}
+% \hypersetup{
+%   colorlinks=true, 
+%   allcolors=black,
+%   pdfauthor={Carlos Henrique Tarjano Santos},
+%   pdftitle={Robust Digital Envelope Estimation Via Geometric Properties of an Arbitrary Real Signal}
+% }
 
@@
 \title{Envelope Estimation using Geometric Properties of a Discrete Real Signal}
-\author[1]{Carlos Henrique Tarjano Santos}
-\author[2]{Valdecy Pereira}
-\affil[1]{(corresponding author, carlostarjano@id.uff.br) Department of Production Engineering, Universidade Federal Fluminense, Rua Passo da Pátria, 156, Campus Praia Vermelha, Bloco D - sala 309, São Domingos, Niterói, RJ, Brasil, CEP: 24.210-240}
-\affil[2]{Department of Production Engineering, Universidade Federal Fluminense, Rua Passo da Pátria, 156, Campus Praia Vermelha, Bloco D - sala 309, São Domingos, Niterói, RJ, Brasil, CEP: 24.210-240}
 
-\begin{document}
+\author{Carlos Henrique Tarjano Santos\corref{cor1}}
+\ead{carlostarjano@id.uff.br}
+\address{Department of Production Engineering, Universidade Federal Fluminense, Rua Passo da Pátria, 156, Campus Praia Vermelha, Bloco D - sala 309, São Domingos, Niterói, RJ, Brasil, CEP: 24.210-240}
+\cortext[cor1]{Corresponding author, carlostarjano@id.uff.br}
 
-\maketitle
+\author{Valdecy Pereira}
+\address{Department of Production Engineering, Universidade Federal Fluminense, Rua Passo da Pátria, 156, Campus Praia Vermelha, Bloco D - sala 309, São Domingos, Niterói, RJ, Brasil, CEP: 24.210-240}
 
@@
 We provide visualizations of the envelope extracted via the algorithm for various real-world signals, with very diverse characteristics, such as voice, spoken and sang, and pitched and non-pitched musical instruments, and discuss some approaches to assess the quality of the obtained envelopes.
-A Python module implementing the algorithm was made available via the Python Package Index (PyPI); interactive visualizations of envelopes for a diverse range of digital waves, as well as the source code for the Python implementation, are available online at \textcite{2020TarjanoEnvelope}.
+A Python module implementing the algorithm was made available via the Python Package Index (PyPI); interactive visualizations of envelopes for a diverse range of digital waves, as well as the source code for the Python implementation, are available online.
 % Conclusions | Implications
@@
 
-{\bf Keywords:} DSP; alpha shapes; envelope detection; discrete curvature estimation; demodulation
+\begin{keyword}
+  DSP \sep alpha shapes \sep envelope detection \sep discrete curvature estimation \sep demodulation
+\end{keyword}
+
+\begin{document}
+
+\maketitle
 
@@
 \begin{align} \label{eq:Envelope}
+\begin{split}
   & \mathbf{w}, \mathbf{e}, \mathbf{c} \in \mathbb{R}^n \\
@@
   & \mathbf{w} = \mathbf{e} \odot \mathbf{c} \ \therefore \ w_i = e_i c_i \ \forall \ i, \ 0 \le i \le n-1
+\end{split}
 \end{align}
@@
 \begin{align} \label{eq:Basis}
+\begin{split}
 \mathbf{b}^\prime_x &= \mathbf{b}_x  \\
 \mathbf{b}^\prime_y &= \frac{\left( \frac{i_{m-1} - i_0}{m-1} \right)}{\left( \frac{\sum_{j=0}^{m-1} \lvert w_j \lvert}{m} \right)} \mathbf{b}_y 
+\end{split}
 \end{align}
@@
 \begin{equation} \label{eq:Radius}
+\begin{split}
 r = \frac{\sum\limits_{k=0}^{m-2} \left( \frac{v_{k,x} \sqrt{v_{k,x}^2 + v_{k,y}^2}}{ v_{k,y}} \right)}{m-2}
+\end{split}
 \end{equation}
@@
 \begin{align} \label{eq:InEnvelope}
+\begin{split}
 & P_a \in \text{E} \leftrightarrow (P_{b,x} - x_o)^2 + (P_{b,y} - y_o)^2 > r^2, \\
@@
 & P_a, P_b, P_c \in \text{P}, \quad P_a \ne P_b \ne P_c \quad \text{and} \quad x_o, y_o \in \mathbb{R}, \quad 0 \le y_o < +\infty
+\end{split}
 \end{align}

### 2020-12-06 Papers/01_Pulses_Envelope/images/000GraphicalAbstract.py

@@ -1,0 +1,175 @@
+import sys
+import numpy as np
+import signal_envelope as se
+from scipy import interpolate
+import plotly.graph_objects as go
+from plotly.subplots import make_subplots
+
+
+name = "piano"
+W, _ = se.read_wav(f"C:/Users/tesse/Desktop/Files/Dropbox/0_Science/reps/envelope/test_samples/{name}.wav")
+amp = np.max(np.abs(W))
+W = 4 * W / amp
+X = np.arange(W.size)
+
+Xpos, Xneg = se.get_frontiers(W, 0)
+E = se.get_frontiers(W, 1)
+
+f = interpolate.interp1d(E, np.abs(W[E]), kind="linear", fill_value="extrapolate")
+E = f(X)
+
+for i in range(E.size):
+  if E[i] < 0.1:
+    E[i] = 0.1
+
+C = W / E
+
+'''============================================================================'''
+'''                                    PLOT                                    '''
+'''============================================================================'''
+
+FONT = dict(
+    family="Arial",
+    color="black",
+    size=24
+    # size=13.3333
+  )
+
+'''Plotting'''
+fig = make_subplots(
+  rows=2, cols=1, shared_xaxes=True, vertical_spacing=0.01,
+  subplot_titles=("<b>Extracted Envelope and Carrier Wave</b> (33<sup>rd</sup> key of a grand piano)", "<b>Extracted Superior and Inferior Frontiers</b> (33<sup>rd</sup> key of a grand piano)")
+)
+fig.layout.template ="plotly_white" 
+fig.update_layout(
+  # xaxis_title="<b><i>i</i></b>",
+  # yaxis_title="<b>Amplitude</b>",
+  legend=dict(orientation='h', yanchor='top', xanchor='left', y=1.1),
+  margin=dict(l=0, r=0, b=0, t=0),
+  font=FONT,
+  # title=dict(text="<b>Teste</b>", font=FONT)
+)
+# fig.layout.xaxis.title.font=FONT
+# fig.layout.yaxis.title.font=FONT
+
+fig.update_xaxes(
+  showline=False, 
+  showgrid=False, 
+  zeroline=False, 
+  showticklabels=False, 
+  # title=dict(text="<b><i>i</i></b>",font=FONT), 
+  row=1, col=1
+)
+
[…]

### 2021-02-16 Papers/02_Sound_Compression/SoundRepresentation.tex
housekeeping

\documentclass{article}
\usepackage{lmodern} % Latin Modern Font
\usepackage[T1]{fontenc}
\usepackage{amsfonts} % \mathbb
\usepackage[margin=0.5in]{geometry}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
% \graphicspath{{images/}}
\usepackage[backend=biber]{biblatex}
\usepackage{amsmath}
\usepackage[makeroom]{cancel}
% \usepackage{auto-pst-pdf}

% \usepackage{hyperref}
% \usepackage{hypcap}
% \DeclareMathOperator{\sign}{sign}
% \hypersetup{
%   colorlinks=true, 
%   allcolors=black,
%   pdfauthor={Carlos Henrique Tarjano Santos},
%   pdftitle={Robust Digital Envelope Estimation Via Geometric Properties of an Arbitrary Real Signal}
% }

\bibliography{bibli}

\usepackage{authblk}
\title{Title}
\author[1]{Carlos Henrique Tarjano Santos}
\author[2]{Valdecy Pereira}
\affil[1]{(corresponding author, carlostarjano@id.uff.br) Department of Production Engineering, Universidade Federal Fluminense, Rua Passo da Pátria, 156, Campus Praia Vermelha, Bloco D - sala 309, São Domingos, Niterói, RJ, Brasil, CEP: 24.210-240}
\affil[2]{Department of Production Engineering, Universidade Federal Fluminense, Rua Passo da Pátria, 156, Campus Praia Vermelha, Bloco D - sala 309, São Domingos, Niterói, RJ, Brasil, CEP: 24.210-240}

\begin{document}

\maketitle

\begin{abstract} % ≅ 250 words  
% Clear and simple miniature version of the paper
% Self-contained
% Past tense for the current study
% Present tense to refer to stablished facts

% Background (What is already known about the subject)
sample based virtual instruments are the de facto standard for the emulation of real instruments. nevertheless little attention is given to the encoding of the samples used in those instruments; in fact, lossless containers like wav or proprietary formats are common, pushing the on disk size of those libraries to sizes exceeding the gigabytes.
% Motivation (Why do we care about the problem and the results?) 
This renders the most realistic digital instruments inacessible in many  but the most high end hardware, controllers, like electronic drumkits and controller keyboards. Even less robust personal computers might encounter problems.
% Problem | Objectives (What problem is the paper trying to solve?) (Scope)
We propose a lossy compression algorithm taylored specifically to highly uniform sounds, as those present in this instruments, due to the nature of the sounds, highly redudant, substantial improvements can be obtained, with minimal quality loss.
The method, being physically informed, also serves as base 
% Methods | Procedure | Approach (What did you actually do to get your results?)

% Results | Findings | Product (what did you learn | invent | create?)

% Conclusions | Implications (What are the larger implications of your findings)

\end{abstract}

{\bf Keywords:}

\section{Introduction} % ≅ 500 words
% ≅ 10\% of the text
% broad to specific
% show that the research question is clear, concise, and worthy of study.
% allow the reader to understand the rest of the paper without referring to previous publications on the topic
% What did you do?
% conclusions

% Opener sentence (Question | Assertion | Intriguing Revelations)
Many of the proposed benefits of an electronic drumkit, like portability and practicity, are compromised when, to obtain a realistic sound on par with the acoustic counterpart, the output of the drumkit has to be processed in another piece of hardware, such as a personal computer.

This transforms the instrument in a mere controller, robbing it of characteristics that provide the personality of the conventional instruments, such as timbre.

% Literature review | Brief Context of Prior Research
  % put your research in the context of other research
  % Nature and scope of the problem investigated
  % give them a framework for understanding it
  % What is the problem to be solved?
  % Are there any existing solutions?
  % Which is the best?
  % What is its main limitation?
  % What do you hope to achieve?

% Restate Your Question as Something Not Known or Fully Understood by Prior Research
  % identify the gap in the literature
  % use but, however, etc…

% State the Significance of Your Question
  % make readers understand why they should read your report

% State Your Claim | Objective | Hypothesis
  % should clearly relate to the information gap

% brief outlook on the structure of the paper

\section{Methodology} % ≅ 1000 words
  % How did you do it?
  % Justify choices
  % Describe the work so as people can reproduce it

% Overview of methods

% Restate the purpose of the work

% Essential background information

% Relate to other studies

% Indicate where problems occurred

\section{Results} % ≅ 3000 words
  % What did you find?
  % Use figures, tables, etc.
  % Clearly and simply stated.

% Revisit purpose and methods

% Overview of results

% Key results in details

% Comparison with results in other works

% Comparison with predictions

% Problems with results

% Possible implications

\section{Discussion} % ≅ 2000 words

% Put results into a broader context.

% specific to broad

% What does it all mean?

% an interpretation of how each of the experiments supported your central hypothesis

% add in supporting findings from the literature

% what principles have been established or reinforced

\section{Conclusion} % ≅ 250 words
  % leave readers with a clear idea of your claim
  % make readers understand its importance
  % suggest further research
  % how the work advances the field?
  % significance and implications of the results
  % implications for broader application
  % research opportunities

% Restate Your Claim

% Point Out: New Significance | Practical Application | New Research

% \begin{figure}[ht!]
%   \centering
%      \includegraphics{images/01signalenvelope.pdf}
%   \caption{A discrete wave of the singing voice of an alto singer. The black lines are the frontiers; in the detail view, it can be seen that each region between two diamonds comprises a pseudo-cycle, as defined by the points belonging to the upper frontier. Similarly, two adjacent squares delimitate a pseudo-cycle, from the standpoint of the lower frontier.}
%   \label{fig:signalenvelope}
% \end{figure}

\printbibliography % 20 ⇔ 60 itens

\end{document}

### 2021-02-16 Cpp/Header.h

@@ -11,6 +11,7 @@
 #include "mkl_dfti.h" // Intel MKL
 #include <tuple>
 #include <boost/math/interpolators/cardinal_cubic_b_spline.hpp>
+#include <filesystem> // to list all files in a directory
 
 //typedef int pos_integer;
 typedef unsigned long long pint;

### 2021-02-16 Cpp/Main.cpp
write and read binary to disk implemented

@@ -10,13 +10,15 @@ public:
 	v_real Envelope;
 	v_real Waveform;
 	pint n;
+	pint fps;
 
 
-	layer(v_inte X, v_real E, v_real W, pint n_) {
+	layer(v_inte X, v_real E, v_real W, pint n_, pint fps_) {
 		X_PCs = X;
 		Envelope = E;
 		Waveform = W;
 		n = n_;
+		fps = fps_;
 	}
 	pint size() {
 		return X_PCs.size() + Envelope.size() + Waveform.size();
@@ -116,7 +118,7 @@ public:
 	}
 };
 
-layer compress(const v_real& W, inte max_seconds = 100, real max_iterations = std::numeric_limits<double>::infinity(), inte mult = 3) {
+Compressed compress(const v_real& W, inte max_seconds = 100, real max_iterations = std::numeric_limits<double>::infinity(), inte mult = 3) {
 	v_pint posX, negX;
 	get_pulses(W, posX, negX);
 
@@ -177,12 +179,13 @@ layer compress(const v_real& W, inte max_seconds = 100, real max_iterations = st
 			best_Xpcs = Xpcs;
 			best_avg = avg;
 
+			pint cn = 2 * (best_Xpcs.size() + best_avg.size()) + 1;
 			std::cout
 				<< "\n"
 				<< "Iteration " << ctr << "\n"
 				<< "Number of samples in the original signal:" << W.size() << "\n"
 				//<< ", Xpcs[-1]:" << best_Xpcs.back()
-				<< "Appr. compressed size:" << best_Xpcs.size() + best_avg.size() << " (" <<  float(best_Xpcs.size() + best_avg.size()) / float(W.size()) << ")\n"
+				<< "Appr. compressed size:" << cn << " (" <<  float(cn) / float(W.size()) << ")\n"
 				<< "Waveform length mode: " << ma.mode << "\n"
 				<< "Waveform length abdm: " << ma.abdm << "\n"
 				<< "Min waveform length:  " << min_size << "\n"
@@ -193,29 +196,7 @@ layer compress(const v_real& W, inte max_seconds = 100, real max_iterations = st
 		duration = std::chrono::duration_cast<std::chrono::seconds>(std::chrono::high_resolution_clock::now() - start).count();
 	}
 
-	Wave_rep = Compressed::raw(best_Xpcs, best_avg, W);
[…]

### 2021-02-17 Papers/02_Sound_Compression/SoundRepresentation.tex
write and read binary to disk implemented
working, generally. still some bugs with some wav files

@@
 % Problem | Objectives (What problem is the paper trying to solve?) (Scope)
-We propose a lossy compression algorithm taylored specifically to highly uniform sounds, as those present in this instruments, due to the nature of the sounds, highly redudant, substantial improvements can be obtained, with minimal quality loss.
-The method, being physically informed, also serves as base 
+We propose a lossy compression algorithm tailored specifically to highly uniform sounds, as those present in this instruments, due to the nature of the sounds, highly redudant, substantial improvements can be obtained, with minimal quality loss.
+
 % Methods | Procedure | Approach (What did you actually do to get your results?)
@@
 % Conclusions | Implications (What are the larger implications of your findings)
+The method, being physically informed, also serves as base 
+lends itself well to machine learning, as the sound can be represented by an evolving waveform over time (pseudo-cycles)
 

### 2021-02-17 Cpp/Header.h

@@ -13,6 +13,8 @@
 #include <boost/math/interpolators/cardinal_cubic_b_spline.hpp>
 #include <filesystem> // to list all files in a directory
 
+//#define v // for verbose
+
 typedef unsigned long long pint;
 typedef long long          inte;
 typedef double             real;
@@ -108,7 +110,7 @@ public:
 	Wav(v_real W_, pint fps_=44100) {
 		fps = fps_;
 		W = W_;
-		std::cout << " n: " << W.size() << " fps: " << fps << "\n";
+		std::cout << "New wave with n: " << W.size() << " fps: " << fps << "\n";
 	}
 	void write(std::string path = "test.wav") {
 		if (W.size() == 0) {
@@ -124,7 +126,7 @@ public:
 		SNDFILE* outfile = sf_open(fname, SFM_WRITE, &sfinfo);
 		sf_write_double(outfile, &W[0], W.size());
 		sf_close(outfile);
-		std::cout << "file written. path=" << path << ", fps=" << fps << "\n";
+		std::cout << "Wave file written. path=" << path << ", fps=" << fps << "\n";
 	}
 };
 
@@ -320,7 +322,9 @@ public:
 	}
 
 	static Compressed raw(const v_inte Xp, const v_real& W, const v_real& S, pint f = 44100) {
-
+#ifdef v
+		std::cout << "inside Compressed::raw\n";
+#endif
 		boost::math::interpolators::cardinal_cubic_b_spline<real> spline(W.begin(), W.end(), 0.0, 1.0 / real(W.size()));
 		v_real W_reconstructed(S.size(), 0.0);
 		v_real Envelope;
@@ -421,8 +425,8 @@ Wav read_wav(std::string path) {
 		exit(1);
 	}
 
-	int fps = file.samplerate();
-	static int n = file.frames();
+	const int fps = file.samplerate();
+	const int n = file.frames();
 	std::cout << "Successfully opened file at:" << path << "\n";
 
 	v_real W(n);
@@ -516,11 +520,9 @@ void get_pulses(const v_real& W, v_pint& posX, v_pint& negX) {
 				im = argabsmax(W, i0, i);
 				i0 = i;
[…]

### 2021-02-25 Papers/02_Sound_Compression/SoundRepresentation.tex
some progress on the paper

@@
 % Background (What is already known about the subject)
-sample based virtual instruments are the de facto standard for the emulation of real instruments.
-nevertheless little attention is given to the encoding of the samples used in those instruments; in fact, lossless containers like wav or proprietary formats are common, pushing the on disk size of those libraries to sizes exceeding the gigabytes.
+Sample based virtual instruments are the de facto standard for the emulation of real word acoustic instruments.
+Nevertheless little attention is given to the encoding of the samples used in those instruments; in fact, lossless containers like wav or proprietary formats are common, pushing the on disk size of those libraries to sizes exceeding the gigabytes.
 % Motivation (Why do we care about the problem and the results?) 
-This renders the most realistic digital instruments inacessible in many  but the most high end hardware, controllers, like electronic drumkits and controller keyboards.
-Even less robust personal computers might encounter problems.
+This renders the most realistic digital instruments inaccessible in many but the most high-end hardware; controllers, like electronic drum kits and controller keyboards and even less robust personal computers might encounter problems implementing those software.
 % Problem | Objectives (What problem is the paper trying to solve?) (Scope)
-We propose a lossy compression algorithm tailored specifically to highly uniform sounds, as those present in this instruments, due to the nature of the sounds, highly redudant, substantial improvements can be obtained, with minimal quality loss.
+We propose a lossy compression algorithm tailored specifically to highly uniform sounds, as those present in these instruments, due to the nature of the sounds, highly redundant, substantial improvements can be obtained, with minimal quality loss.
 
@@
 
-This transforms the instrument in a mere controller, robbing it of characteristics that provide the personality of the conventional instruments, such as timbre.
+The same is true for digital pianos.
+
+This transforms the instrument in a mere controller, robbing it of characteristics that provide the personality of the conventional instruments, such as a particular timbre.
 
@@
 % Overview of methods
+We can interpret a continuous or discrete signal, as long as it is at least pseudo periodic, as succession of waveforms, each one representing a (pseudo) cycle of the wave.
+
+We start with an estimate of the waveform of the signal being processed.
 

### 2021-02-25 Cpp/Main.cpp

@@ -320,6 +320,7 @@ layer read_bin(std::string path) {
 int main(int argc, char** argv) {
 	pint max_secs = 20;
 	pint max_iters = std::numeric_limits<pint>::max();
+	bool save_csv = false;
 	std::vector<std::string> wav_paths;
 	std::vector<std::string> cmp_paths;
 	std::string append = "reconstructed";
@@ -328,10 +329,11 @@ int main(int argc, char** argv) {
 			std::cout
 				<< " For more info about the author: carlos-tarjano.web.app\n"
 				<< " Usage: \n"
-				<< " [-t x] [-i y] -[a z] [path/to/file1.wav]...[path/to/filen.wav]  [path/to/file1.cmp]...[path/to/filem.cmp]\n"
+				<< " [-t x] [-i y] -[a z] [-csv] [path/to/file1.wav]...[path/to/filen.wav]  [path/to/file1.cmp]...[path/to/filem.cmp]\n"
 				<< " -t or --time: (default " << max_secs <<" s) maximum time in seconds allowed for each compression task\n"
 				<< " -i or --iterations: maximum number of iterations allowed for each compression task\n"
 				<< " -a or --append: (default \"" << append << "\") string to be appended to each reconstructed file name\n"
+				<< " -csv: saves a csv file with the beginning of each pseudo-cycle\n"
 				<< " If no path is given the root folder will be scanned for .wav and .cmp files, and those will be processed accordingly\n";
 			return 0;
 		}
@@ -348,6 +350,11 @@ int main(int argc, char** argv) {
 		else if (std::strcmp(argv[i], "-a") == 0 || std::strcmp(argv[i], "--append") == 0) {
 			append = argv[i + 1];
 			std::cout << "Append \"" << append << "\" to the name of reconstructed files\n";
+			i++;
+		}
+		else if (std::strcmp(argv[i], "-csv") == 0) {
+			save_csv = true;
+			std::cout << "pseudo-cycle info will be saved as csv\n";
 			i++;
 		}
 		else if (ends_with(argv[i], ".wav")) {
@@ -377,9 +384,14 @@ int main(int argc, char** argv) {
 		//auto fps = WV.fps;
 
 		try {
-		auto C = compress(WV.W, max_secs, max_iters);
-		path.replace(path.end() - 4, path.end(), ".cmp");
[…]

### 2021-02-25 Papers/02_Sound_Compression/images/01signalenvelope/Frontier.py

@@ -1,359 +1,0 @@
-import plotly.graph_objects as go
-import numpy as np
-from scipy.interpolate import interp1d
-import numpy.polynomial.polynomial as poly
-import wave
-from scipy.signal import savgol_filter
-import os
-from pathlib import Path
-import sys
-import argparse
-
-def read_wav(path): 
-  """returns signal & fps"""
-  wav = wave.open(path , 'r')
-  signal = np.frombuffer(wav.readframes(-1) , np.int16)
-  fps = wav.getframerate()
-  return signal, fps
-
-
-def get_circle(x0, y0, x1, y1, r):
-  '''returns center of circle that passes through two points'''
-  
-  radsq = r * r
-  q = np.sqrt((x1 - x0)**2 + (y1 - y0)**2)
-
-  # assert q <= 2 * r, f"shit"
-
-  x3 = (x0 + x1) / 2
-  y3 = (y0 + y1) / 2
-
-  if y0 + y1 >= 0:
-    xc = x3 + np.sqrt(radsq - (q / 2)**2) * ((y0 - y1) / q)
-    yc = y3 + np.sqrt(radsq - (q / 2)**2) * ((x1 - x0) / q)
-  else:
-    xc = x3 - np.sqrt(radsq - (q / 2)**2) * ((y0 - y1) / q)
-    yc = y3 - np.sqrt(radsq - (q / 2)**2) * ((x1 - x0) / q)
-
-  return xc, yc
-
-
-def get_radius_function(X, Y):
-  m0 = np.average(Y[1:] - Y[:-1]) / np.average(X[1:] - X[:-1])
-  curvatures_X = []
-  curvatures_Y = []
-  for i in range(len(X) - 1):
-    m1 = (Y[i + 1] - Y[i]) / (X[i + 1] - X[i])
-
-    theta = np.arctan( (m1 - m0) / (1 + m1 * m0) )
-    k = np.sin(theta) / (X[i + 1] - X[i])
-
-    # if k >= 0:
-    curvatures_X.append((X[i + 1] + X[i]) / 2)
-    curvatures_Y.append(k)
-    
-  curvatures_Y = np.array(curvatures_Y)
-  coefs = poly.polyfit(curvatures_X, curvatures_Y, 0)
-  smooth_curvatures_Y = poly.polyval(curvatures_X, coefs)
-
-  r_of_x = interp1d(curvatures_X, 1 / np.abs(smooth_curvatures_Y), fill_value="extrapolate")
[…]

### 2021-02-25 Papers/02_Sound_Compression/images/02_waveform.py

@@ -1,0 +1,133 @@
+# from os import linesep
+from statistics import mode
+import numpy as np
+import plotly.graph_objects as go
+import signal_envelope as se
+from scipy import interpolate
+import sys
+
+
+def get_pcs(Xpc, W):
+  Xpc = Xpc.astype(int)
+  # amp = np.max(np.abs(W))
+  # max_T = int(np.max(np.abs(Xpc[1:] - Xpc[:-1])))
+  Xlocal = np.linspace(0, 1, mode(Xpc[1:] - Xpc[:-1]))
+
+  orig_pcs = []
+  norm_pcs = []
+  for i in range(2, Xpc.size):
+    x0 = Xpc[i - 1]
+    x1 = Xpc[i] + 1
+    orig_pcs.append(W[x0 : x1])
+    if x1 - x0 >= 4:
+      yx = interpolate.interp1d(np.linspace(0, 1, x1-x0), W[x0 : x1], "cubic")
+      Ylocal = yx(Xlocal)
+      # Ylocal = Ylocal / np.max(np.abs(Ylocal)) * amp
+      norm_pcs.append(Ylocal)
+  return np.average(np.array(norm_pcs), 0), orig_pcs, norm_pcs
+
+def to_plot(Matrix):
+  X = []
+  Y = []
+  for line in Matrix:
+    for x, y in enumerate(line):
+      X.append(x)
+      Y.append(y)
+    X.append(None)
+    Y.append(None)
+  return X, Y
+
+name = "alto"
+
+'''###### Read wav file ######'''
+W, fps = se.read_wav(f"images/00resources/{name}.wav")
+W = W - np.average(W)
+amp = np.max(np.abs(W))
+W = W / amp
+n = W.size
+
+'''###### Read Pseudo-cycles info ######'''
+Xpc = np.genfromtxt(f"images/00resources/{name}.csv", delimiter=",")
+
+X_xpcs, Y_xpcs = [], []
+for x in Xpc:
+  X_xpcs.append(x)
+  X_xpcs.append(x)
+  X_xpcs.append(None)
+  Y_xpcs.append(-1)
+  Y_xpcs.append(1)
+  Y_xpcs.append(None)
+
+
+average_waveform, orig_waveforms, norm_waveforms = get_pcs(Xpc, W)
+
+# norm_waveforms = norm_waveforms - average_waveform
+
+
+
+'''============================================================================'''
+
+FONT = dict(
+    family="Latin Modern Roman",
+    color="black",
+    size=13.3333
+  )
+
+
[…]

### 2021-02-25 Papers/02_Sound_Compression/images/02modevsaverage.py

@@ -1,107 +1,0 @@
-import plotly.graph_objects as go
-import numpy as np
-import signal_envelope as se
-from statistics import mode
-import sys
-sys.path.insert(0,"C:/Users/tesse/Desktop/Files/Dropbox/0_Thesis/Python")
-import hp as hp
-
-'''==============='''
-''' Read wav file '''
-'''==============='''
-
-# def stats(L):
-#   avg = np.average(L)
-#   mde = mode(L)
-#   std_avg = np.average(np.abs(L - avg))
-#   std_mde = np.average(np.abs(L - mde))
-#   print("std:", std_avg, std_mde, np.min(L), mde, np.max(L))
-#   return avg, std_avg, mde, std_mde, np.std(L)
-
-name = "piano33"
-W, fps = se.read_wav(f"C:/Users/tesse/Desktop/Files/Dropbox/0_Thesis/Python/Samples/{name}.wav")
-W = W - np.average(W)
-amp = np.max(np.abs(W))
-n = W.size
-X = np.arange(n)
-
-Xpos, Xneg = se.get_frontiers(W)
-Xpos, Xneg = Xpos.astype(np.int), Xneg.astype(np.int)
-posT = Xpos[1:] - Xpos[:-1]
-negT = Xneg[1:] - Xneg[:-1]
-T = np.hstack([posT, negT])
-
-# Xpos = hp.refine_frontier_iter(Xpos, W)
-# Xneg = hp.refine_frontier_iter(Xneg, W)
-# posT_ref = Xpos[1:] - Xpos[:-1]
-# negT_ref = Xneg[1:] - Xneg[:-1]
-# T_ref = np.hstack([posT_ref, negT_ref])
-
-mode_error = T - mode(T)
-modeX, modeY = np.unique(mode_error, return_counts=True)
-
-avg_error = T - np.average(T)
-avgX, avgY = np.unique(avg_error, return_counts=True)
-
-'''============================================================================'''
-'''                              PLOT LINES                                    '''
-'''============================================================================'''
-FONT = dict(
-    family="Latin Modern Roman",
-    color="black",
-    size=13.3333
-  )
-
-'''Plotting'''
-fig = go.Figure()
-fig.layout.template ="plotly_white" 
-fig.update_layout(
-  xaxis_title="<b><i>T</i></b>",
-  yaxis_title="<b>Number of Ocurrences</b>",
[…]

### 2021-02-25 Papers/02_Sound_Compression/images/03frontiergaps.py

@@ -1,152 +1,0 @@
-import plotly.graph_objects as go
-import numpy as np
-import signal_envelope as se
-from statistics import mode
-import sys
-sys.path.insert(0,"C:/Users/tesse/Desktop/Files/Dropbox/0_Thesis/Python")
-import hp as hp
-
-# def refine_frontier(Xp, W):
-#   "find additional frontier points, and return an array with then, if found"
-#   if W[Xp[0]] >= 0:
-#     e = np.argmax
-#   else:
-#     e = np.argmin
-#   L = Xp[1:] - Xp[:-1]
-#   avgL = mode(L)
-#   stdL = np.average(np.abs(L - avgL))
-#   Xnew = []
-#   for i in range(1, Xp.size):
-#     x0 = int(Xp[i - 1])
-#     x1 = int(Xp[i])
-#     if x1 - x0 > avgL + 2 * stdL:
-#       Xzeroes = []
-#       currsign = np.sign(W[x0])
-#       for i in range(x0 + 1, x1):
-#         if currsign != np.sign(W[i]):
-#           Xzeroes.append(i)
-#           currsign = np.sign(W[i])
-#       if len(Xzeroes) > 1:
-#         Xnew.append(Xzeroes[0] + e(W[Xzeroes[0] : Xzeroes[-1]]))
-#   return np.array(Xnew, dtype=np.int)
-
-
-'''==============='''
-''' Read wav file '''
-'''==============='''
-
-name = "tom"
-
-W, fps = se.read_wav(f"C:/Users/tesse/Desktop/Files/Dropbox/0_Thesis/Python/Samples/{name}.wav")
-W = W - np.average(W)
-n = W.size
-X = np.arange(n)
-
-Xpos, Xneg = se.get_frontiers(W)
-
-Xposnew = hp.refine_frontier_iter(Xpos, W)
-Xnegnew = hp.refine_frontier_iter(Xneg, W)
-
-
-
-
-
-'''============================================================================'''
-'''                                    PLOT                                    '''
-'''============================================================================'''
-FONT = dict(
-    family="Latin Modern Roman",
-    color="black",
-    size=13.3333
-  )
-
-'''Plotting'''
-fig = go.Figure()
-fig.layout.template ="plotly_white" 
-fig.update_layout(
-  xaxis_title="<b><i>i</i></b>",
[…]

### 2021-02-25 Papers/02_Sound_Compression/images/04plotXps.py

@@ -1,141 +1,0 @@
-import plotly.graph_objects as go
-import numpy as np
-import signal_envelope as se
-import sys
-sys.path.insert(0,"C:/Users/tesse/Desktop/Files/Dropbox/0_Thesis/Python")
-import hp as hp
-
-
-'''==============='''
-''' Read wav file '''
-'''==============='''
-
-
-name = "piano33"
-W, fps = se.read_wav(f"C:/Users/tesse/Desktop/Files/Dropbox/0_Thesis/Python/Samples/{name}.wav")
-W = W - np.average(W)
-amp = np.max(np.abs(W))
-n = W.size
-X = np.arange(n)
-
-Xpos_orig, Xneg_orig = se.get_frontiers(W)
-
-Xpos = hp.refine_frontier_iter(Xpos_orig, W)
-Xneg = hp.refine_frontier_iter(Xneg_orig, W)
-
-nn = min(Xpos.size, Xneg.size)
-Average_ref_X = np.round((Xpos[:nn] + Xneg[:nn]) / 2)#.astype(np.int)
-
-'''============================================================================'''
-'''                              PLOT LINES                                    '''
-'''============================================================================'''
-FONT = dict(
-    family="Latin Modern Roman",
-    color="black",
-    size=13.3333
-  )
-
-'''Plotting'''
-fig = go.Figure()
-fig.layout.template ="plotly_white" 
-fig.update_layout(
-  xaxis_title="<b>Ordinality</b>",
-  yaxis_title="<b><i>i</i></b>",
-  legend=dict(orientation='h', yanchor='top', xanchor='left', y=1.1),
-  margin=dict(l=0, r=0, b=0, t=0),
-  font=FONT,
-  titlefont=FONT
-)
-fig.layout.xaxis.title.font=FONT
-fig.layout.yaxis.title.font=FONT
-fig.update_xaxes(showline=False, showgrid=False, zerolinewidth=1, zerolinecolor='silver')
-fig.update_yaxes(showline=False, showgrid=False, zerolinewidth=1, zerolinecolor='silver')
-
-fig.add_trace(
-  go.Scatter(
-    name="Maxima", # <|<|<|<|<|<|<|<|<|<|<|<|
-    # x=Xpos,
-    y=Xpos_orig,
-    # fill="toself",
-    mode="lines",
-    line=dict(
-        width=1,
-        color="black",
-        dash="dash"
[…]

### 2021-02-25 Papers/02_Sound_Compression/images/0ld/02pcsraw.py

@@ -1,175 +1,0 @@
-from ctypes import resize
-import plotly.graph_objects as go
-import numpy as np
-import signal_envelope as se
-import sys
-import os
-os.chdir('../..')
-print (os.path.abspath(os.curdir))
-from plotly.subplots import make_subplots
-
-'''==============='''
-''' Read wav file '''
-'''==============='''
-
-name = "alto"
-W, fps = se.read_wav(f"./Python/Samples/{name}.wav")
-W = W - np.average(W)
-n = W.size
-X = np.arange(n)
-
-save_path = os.path.abspath(os.curdir) + '''\\Papers\\02_Sound_Representation\\images\\''' + sys.argv[0].split('/')[-1].replace(".py", "") + "-" + name + ".svg"
-print(save_path)
-
-# exit()
-
-Xpos, Xneg = se.get_frontiers(W)
-
-maxL = []
-for i in range(1, Xpos.size):
-  maxL.append(Xpos[i] - Xpos[i - 1])
-for i in range(1, Xneg.size):
-  maxL.append(Xneg[i] - Xneg[i - 1])
-maxL = np.max(np.array(maxL))
-print(f"Max L = {maxL}")
-
-pospseudoCyclesX = []
-pospseudoCyclesY = []
-for i in range(1, Xpos.size):
-  x0 = Xpos[i - 1]
-  x1 = Xpos[i]
-  # a = np.max(np.abs(W[x0 : x1]))
-  # ft = np.fft.rfft(W[x0 : x1])
-  npulse = W[x0 : x1]
-  pospseudoCyclesY.append(npulse)
-  pospseudoCyclesX.append(np.arange(npulse.size))
-pospseudoCyclesX = np.array(pospseudoCyclesX)
-pospseudoCyclesY = np.array(pospseudoCyclesY)
-
-negpseudoCyclesX = []
-negpseudoCyclesY = []
-for i in range(1, Xneg.size):
-  x0 = Xneg[i - 1]
-  x1 = Xneg[i]
-  # a = np.max(np.abs(W[x0 : x1]))
-  # ft = np.fft.rfft(W[x0 : x1])
-  npulse = W[x0 : x1]
-  negpseudoCyclesY.append(npulse)
-  negpseudoCyclesX.append(np.arange(npulse.size))
-negpseudoCyclesX = np.array(negpseudoCyclesX)
-negpseudoCyclesY = np.array(negpseudoCyclesY)
-
-
-
-'''============================================================================'''
-'''                                    PLOT                                    '''
[…]

### 2021-02-25 Papers/02_Sound_Compression/images/0ld/03pcs.py

@@ -1,175 +1,0 @@
-from ctypes import resize
-import plotly.graph_objects as go
-import numpy as np
-import signal_envelope as se
-import sys
-import os
-os.chdir('../..')
-print (os.path.abspath(os.curdir))
-from plotly.subplots import make_subplots
-
-'''==============='''
-''' Read wav file '''
-'''==============='''
-
-name = "alto"
-W, fps = se.read_wav(f"./Python/Samples/{name}.wav")
-W = W - np.average(W)
-n = W.size
-X = np.arange(n)
-
-save_path = os.path.abspath(os.curdir) + '''\\Papers\\02_Sound_Representation\\images\\''' + sys.argv[0].split('/')[-1].replace(".py", "") + "-" + name + ".svg"
-print(save_path)
-
-# exit()
-
-Xpos, Xneg = se.get_frontiers(W)
-
-maxL = []
-for i in range(1, Xpos.size):
-  maxL.append(Xpos[i] - Xpos[i - 1])
-for i in range(1, Xneg.size):
-  maxL.append(Xneg[i] - Xneg[i - 1])
-maxL = np.max(np.array(maxL))
-print(f"Max L = {maxL}")
-
-pospseudoCyclesX = []
-pospseudoCyclesY = []
-for i in range(1, Xpos.size):
-  x0 = Xpos[i - 1]
-  x1 = Xpos[i]
-  # a = np.max(np.abs(W[x0 : x1]))
-  ft = np.fft.rfft(W[x0 : x1])
-  npulse = np.fft.irfft(ft, maxL)
-  pospseudoCyclesY.append(npulse / np.max(np.abs(npulse)))
-  pospseudoCyclesX.append(np.arange(maxL))
-pospseudoCyclesX = np.array(pospseudoCyclesX)
-pospseudoCyclesY = np.array(pospseudoCyclesY)
-
-negpseudoCyclesX = []
-negpseudoCyclesY = []
-for i in range(1, Xneg.size):
-  x0 = Xneg[i - 1]
-  x1 = Xneg[i]
-  # a = np.max(np.abs(W[x0 : x1]))
-  ft = np.fft.rfft(W[x0 : x1])
-  npulse = np.fft.irfft(ft, maxL)
-  negpseudoCyclesY.append(npulse / np.max(np.abs(npulse)))
-  negpseudoCyclesX.append(np.arange(maxL))
[…]

### 2021-02-25 Papers/02_Sound_Compression/images/0ld/05std.py

@@ -1,105 +1,0 @@
-import plotly.graph_objects as go
-import numpy as np
-import signal_envelope as se
-from plotly.subplots import make_subplots
-import sys
-import os
-os.chdir("../..")
-print (os.path.abspath(os.curdir))
-
-'''==============='''
-''' Read wav file '''
-'''==============='''
-
-name = "piano33"
-W, fps = se.read_wav(f"./Python/Samples/{name}.wav")
-W = W - np.average(W)
-n = W.size
-X = np.arange(n)
-
-save_path = os.path.abspath(os.curdir) + '''\\Papers\\02_Sound_Representation\\images\\''' + sys.argv[0].split('/')[-1].replace(".py", "") + "-" + name + ".svg"
-print(save_path)
-
-
-Xpos, Xneg = se.get_frontiers(W)
-
-posL = []
-for i in range(1, Xpos.size):
-  posL.append(Xpos[i] - Xpos[i - 1])
-
-negL = []
-for i in range(1, Xneg.size):
-  negL.append(Xneg[i] - Xneg[i - 1])
-
-np.histogram(Xneg, np.arange(Xneg.size))
-
-
-'''============================================================================'''
-'''                                    PLOT                                    '''
-'''============================================================================'''
-fig = make_subplots(rows=1, cols=2, shared_yaxes=True)
-fig.layout.template ="plotly_white"
-# fig.update_yaxes(zeroline=True, zerolinewidth=2, zerolinecolor="black")
-fig.update_layout(
-  # title = name,
-  xaxis_title="Length",
-  yaxis_title="Number of Ocurrences",
-
-  # yaxis = dict(        # <|<|<|<|<|<|<|<|<|<|<|<|
-  #   scaleanchor = "x", # <|<|<|<|<|<|<|<|<|<|<|<|
-  #   scaleratio = 1,    # <|<|<|<|<|<|<|<|<|<|<|<|
-  # ),                   # <|<|<|<|<|<|<|<|<|<|<|<|
-
-  legend=dict(orientation='h', yanchor='top', xanchor='left', y=1.1),
-  margin=dict(l=5, r=5, b=5, t=5),
-  font=dict(
-    family="Latin Modern Roman",
-    color="black",
-    size=10
-  )
-)
-
-fig.add_trace(
-  go.Histogram(
-    name=f"Lengths of Positive Pseudo-Pulses (std={np.round(np.std(posL), 2)})",
[…]

### 2021-02-25 Papers/02_Sound_Compression/images/0ld/07periodicwave.py

@@ -1,122 +1,0 @@
-import sys
-# sys.path.append("c:/Users/tesse/Desktop/Files/Dropbox/0_Thesis/Python")
-import numpy as np
-import plotly.graph_objects as go
-# from Helper import signal_to_pulses, get_pulses_area
-
-'''Generating Wave'''
-
-np.random.seed(1)
-# fps = 10
-n = 40
-
-X = np.arange(n)# / fps
-W = np.zeros(n)
-
-f = 1
-p = np.pi / 3
-
-W = np.cos(p + 2 * np.pi * f * X / n) + np.random.normal(0, 1 / 2, n)
-
-FT = np.fft.rfft(W)
-frequency = np.argmax(np.abs(FT))
-phase = np.angle(FT[frequency])
-
-print(f"Ori: f={f}, p={p}")
-print(f"FFT: f={frequency}, p={phase}")
-
-W = W / np.max(np.abs(W))
-xmax = np.argmax(W)
-xmin = np.argmin(W)
-
-repeat = 3
-W = np.tile(W, repeat)
-X = np.arange(repeat * n)
-Xmax = [xmax + i * n for i in range(repeat)]
-Xmin = [xmin + i * n for i in range(repeat)]
-
-'''Plotting'''
-FONT = dict(
-    family="Latin Modern Roman",
-    color="black",
-    size=13.3333
-  )
-
-fig = go.Figure()
-fig.layout.template ="plotly_white" 
-fig.update_layout(
-  xaxis_title="<b><i>i</i></b>",
-  yaxis_title="<b>Amplitude</b>",
-  legend=dict(orientation='h', yanchor='top', xanchor='left', y=1.1),
-  margin=dict(l=0, r=0, b=0, t=0),
-  font=FONT,
-  titlefont=FONT
-)
-
-fig.layout.xaxis.title.font=FONT
-fig.layout.yaxis.title.font=FONT
-fig.update_xaxes(showline=False, showgrid=False, zeroline=False)
-fig.update_yaxes(showline=False, showgrid=False, zerolinewidth=2, zerolinecolor="gray")
-
-'''Samples'''
-fig.add_trace(
-  go.Scatter(
-    name= "Samples Stem",
-    showlegend=False,
-    x=[i for x in X for i in (x, x, None)],
-    y=[i for y in W for i in (0, y, None)],
-    mode='lines',
-    line=go.scatter.Line(color="silver", width=1)
-  )
-)
-
-fig.add_trace(
-  go.Scatter(
-    name= "Cycles",
[…]

### 2021-02-25 Papers/02_Sound_Compression/images/0ld/08line.py

@@ -1,98 +1,0 @@
-import sys
-# sys.path.append("c:/Users/tesse/Desktop/Files/Dropbox/0_Thesis/Python")
-import numpy as np
-import plotly.graph_objects as go
-# from Helper import signal_to_pulses, get_pulses_area
-
-'''Generating Wave'''
-
-np.random.seed(1)
-# fps = 10
-n = 40
-
-X = np.arange(n)# / fps
-
-f = 1
-p = np.pi / 3
-
-W = np.cos(p + 2 * np.pi * f * X / n) + np.random.normal(0, 1 / 2, n)
-W = W / np.max(np.abs(W))
-
-xmax = np.argmax(W)
-xmin = np.argmin(W)
-repeat = 3
-W = np.tile(W, repeat)
-X = np.arange(repeat * n)
-Xmax = [xmax + i * n for i in range(repeat)]
-Xmin = [xmin + i * n for i in range(repeat)]
-
-FT = np.fft.rfft(W)
-frequency = np.argmax(np.abs(FT))
-phase = np.angle(FT[frequency])
-
-print(f"Ori: f={f}, p={p}")
-print(f"FFT: f={frequency}, p={phase}")
-print(f"POS: f={n * repeat * (len(Xmax) - 1) / (Xmax[-1] - Xmax[0])}, p={2 * np.pi * Xmax[0] / n}")
-
-# exit()
-
-'''Plotting'''
-
-FONT = dict(
-    family="Latin Modern Roman",
-    color="black",
-    size=13.3333
-  )
-
-fig = go.Figure()
-fig.layout.template ="plotly_white" 
-fig.update_layout(
-  xaxis_title="<b>Ordinality</b>",
-  yaxis_title="<b><i>i</i></b>",
-  legend=dict(orientation='h', yanchor='top', xanchor='left', y=1.1),
-  margin=dict(l=0, r=0, b=0, t=0),
-  font=FONT,
-  titlefont=FONT
-)
-
-# fig.update_xaxes(
-#     # ticktext=["End of Q1", "End of Q2", "End of Q3", "End of Q4"],
-#     tickvals=[i for i in range(repeat)],
-# )
-
-fig.layout.xaxis.title.font=FONT
-fig.layout.yaxis.title.font=FONT
-fig.update_xaxes(showline=False, showgrid=True, zerolinewidth=2, zerolinecolor="gray", gridwidth=1, gridcolor='silver', tickvals=[i for i in range(repeat)])
[…]

### 2021-03-03 Papers/02_Sound_Compression/SoundRepresentation.tex
progress on the paper

@@
 
+While domain specific compression algorithms and codecs exist in many areas, to the best of the authors' knowledge, that is not the case for sample based digital musical instruments.
+Those signals, being highly uniform, offer an opportunity for compression that is not presente in most general sounds
+
 % Literature review | Brief Context of Prior Research
@@
 % Overview of methods
-We can interpret a continuous or discrete signal, as long as it is at least pseudo periodic, as succession of waveforms, each one representing a (pseudo) cycle of the wave.
+A continuous or discrete signal, as long as it is at least pseudo periodic, can be interpreted as succession of consecutive waveforms, each one representing a (pseudo) cycle of the wave.
+In the case of perfectly periodic signals, those waveforms are identical and we need only the definition of one waveform to characterise the whole signal.in pseudo periodic waves, On the other hand there are potentially variations between pseudo cycles, that must be accounted for.
+If the wave is sufficiently well behaved, as is the case with most sound arising from accoustic phenomenons, these changes can be attributed to variations in both the amplitude of the waveform and its length
 
-We start with an estimate of the waveform of the signal being processed.
+Starting with an initial estimate of the waveform of the signal being processed, one can identify the pseudo cycles of a signal by convolving the estimated waveform with the signal.
+The process can be made computationally efficient by considering some characteristics of the wave.
 
@@
 
+Since the average waveform is used to encode the specificities of the wave, signals where there are signficant variations in the sape of the waveform are naturally ill suited for compression with the presented algorithm.
+This limitation can be mitigated with a pre processing step where the waveform is segmented based on the overal shape of the waveform
+
+noise
+
 % Possible implications
@@
 % Put results into a broader context.
+While presenting an efficient algorithm to compress sufficiently well behaved signals, with immediate applications in the digital musical instrument industry, this work presents a new framework for representing pseudo periodic waves that can be extended in inumerous ways, with the help of both traditional and experimental methods.
+
+While we simply consider the average waveform when encoding the compressed wave, more sofisticated representations can be used.
+The average waveform can be further compressed via a parametric representation.
 

### 2021-03-04 Papers/02_Sound_Compression/SoundRepresentation.tex
progress with the paper

@@
 
-\bibliography{bibli}
+\bibliography{Compression}
 
@@
 % Opener sentence (Question | Assertion | Intriguing Revelations)
-Many of the proposed benefits of an electronic drumkit, like portability and practicity, are compromised when, to obtain a realistic sound on par with the acoustic counterpart, the output of the drumkit has to be processed in another piece of hardware, such as a personal computer.
+Many of the proposed benefits of an electronic drumkit, like portability and practicity for example, are compromised when, to obtain a realistic sound on par with its acoustic counterpart, the output of the drumkit has to be processed in another piece of hardware, such as a personal computer.
 
-The same is true for digital pianos.
+The same is true for digital pianos, where models providing realistic sounds are many times more expensive than entry level models, owing in great part to the additional storage needed.
 
-This transforms the instrument in a mere controller, robbing it of characteristics that provide the personality of the conventional instruments, such as a particular timbre.
+This transforms the instrument in a mere controller or practice aid, robbing it of characteristics that provide the personality of the conventional instruments, such as a particular timbre.
 
-While domain specific compression algorithms and codecs exist in many areas, to the best of the authors' knowledge, that is not the case for sample based digital musical instruments.
-Those signals, being highly uniform, offer an opportunity for compression that is not presente in most general sounds
+Domain specific compression algorithms and codecs exist in many areas.
+\textcite{2016ChenCompressing} for example, presents a scheme for compressing convolutional neural networks, while \textcite{2014CanovasLossy}, besides revising existing methods, proposes two lossy schemes for the compression of quality score data of genomic sequencing.
+
+More recently, the work of \textcite{2019CalhounExploring}, explores the benefits of specific lossy compression algorithms in the context of checkpoints of computational simulations that are stored between simulation sessions.
+
+Regarding sample based digital musical instruments, to the best of the authors' knowledge, no such specialized algorithms exist.
+Those signals, being highly uniform, offer an opportunity for compression that is not present in most general sounds and remain largely unexplored.
+
+This tendency arises in part from the marketing strategy of the vast majority of those products, where the sheer size of the library is advertised and seems to be sold as an equivalent of quality and versatility.
+An overall prejudice against compressed formats in general from the part of the music productions community supports this distorted view.
 
 % Literature review | Brief Context of Prior Research
+In that scenario, the work of \textcite{2019BlauRethinking} and, more recently, \textcite{2020O’GradyRethinking} are paramount in shedding light to the potential misconceptions permeating the opinion about lossy codecs, albeit from very different standpoints.
+
+The first authors argue, from a technical point of view that perceptual quality metrics should be favored over metrics naively based on the mathematical aspect of Shannon’s rate-distortion theory \parencite{Shannon_1948}, the most popular metrics currently used in the assessment of lossy compression codecs.
+
+The second work analyses the problem from a somewhat political viewpoint, arguing that at last some of the criticism directed to the MP3 standard and lossy codecs in general is meant to stablish a symbolic capital in the face of the crescent struggle between established recording labels and alternative channel of music diffusion, such as streaming services.
+
 % put your research in the context of other research
+Nevertheless, research in data compression in general continues with recent efforts in reviewing the advancements of the field in general presented in \textcite{2021Jayasankarsurvey}.
+Account of the research in more specific areas, such as video \parencite{2019LaudeComprehensive} and image \parencite{2014RehmanImage} were also produced recently.
+
+Concerning sound lossless audio compression research seems to be lagging when compared to lossy approaches; \textcite{2001HansLossless} suggested that a limit was about to be reached.
+
+New technologies such as 3D, immersive audio has propelled specific compression algorithms, such as the one presented in \textcite{2021HuAudio}.
+The renaissance of neural networks led to interesting results, such as the use of GANs as presented in \textcite{2018HuangBandwidth}.
+
+The ACER codec, as presented in \textcite{2014CunninghamData} offers an unconventional approach to compression, in that it identifies redundant pieces in a signal that are indexed in a dictionary.
+The results seems to be satisfactory from a perceptual standpoint, in comparison to MP3 and AAC codecs \parencite{2019CunninghamSubjective}.
+
   % Nature and scope of the problem investigated
@@
 
+We compare the results of the algorithm here presented with some of the most popular codecs.
+
+The first format we compare to is the popular MP3, the third layer of MPEG-1/2 specification, proposed in 1992 \parencite{2018BritanakAudio}.
+Despite being superseeded by the newer MPEG-2 AAC \parencite{1999BrandenburgMP3}, finalized five years later, which we also compare, its ubiquity makes the comparison crucial.
+
+besides
+
+The opus format was chosen, as, with the ACC and MP3, those are considered the most advanced \parencite{2019LaudeComprehensive} and popular \parencite{2018KwanObjective}
+
 % Revisit purpose and methods
@@
   % suggest further research
+  stereo
+  perceptual metrics
   % how the work advances the field?

### 2021-03-04 Papers/02_Sound_Compression/SoundRepresentation.tex
working char in disk implementation

@@
 The first format we compare to is the popular MP3, the third layer of MPEG-1/2 specification, proposed in 1992 \parencite{2018BritanakAudio}.
-Despite being superseeded by the newer MPEG-2 AAC \parencite{1999BrandenburgMP3}, finalized five years later, which we also compare, its ubiquity makes the comparison crucial.
+Despite being technically superseded by the newer MPEG-2 AAC \parencite{1999BrandenburgMP3}, finalized five years later, with which we also compare, its ubiquity makes the comparison crucial.
 

### 2021-03-04 Cpp/Main.cpp

@@ -265,20 +265,34 @@ inline bool ends_with(std::string const& value, std::string const& ending) {
 }
 
 void write_bin(std::string path, pint orig_n, pint fps, const v_inte& beg_of_pseudo_cycles, const v_real& waveform, const v_real& amp_of_pseudo_cycles) {
+	std::cout << "n=" << orig_n << "\n";
 
 	std::ofstream data_file;      // pay attention here! ofstream
 
-	v_pint pint_data = { orig_n, fps, amp_of_pseudo_cycles.size(), waveform.size() }; // header
+	pint max_t = 0;
+	v_pint T_pint(beg_of_pseudo_cycles.size());
+	for (size_t i = 1; i < beg_of_pseudo_cycles.size(); i++) {
+		T_pint[i] = beg_of_pseudo_cycles[i] - beg_of_pseudo_cycles[i - 1];
+		if (T_pint[i] > max_t) {
+			max_t = T_pint[i];
+		}
+	}
 
-	std::cout << "n=" << orig_n << "\n";
 
+	v_pint pint_data = { orig_n, fps, amp_of_pseudo_cycles.size(), waveform.size(), max_t }; // header
 	data_file.open(path, std::ios::out | std::ios::binary | std::fstream::trunc);
 	data_file.write((char*) &pint_data[0], pint_data.size() * sizeof(pint));
 	data_file.close();
 
-	std::vector<int> xp(beg_of_pseudo_cycles.begin(), beg_of_pseudo_cycles.end());
+	
+	std::vector<char> T(beg_of_pseudo_cycles.size());
+	T[0] = beg_of_pseudo_cycles[0];
+	for (size_t i = 1; i < beg_of_pseudo_cycles.size(); i++) {
+		T[i] = 127.0 * (real)T_pint[i] / (real)max_t;
+	}
+
 	data_file.open(path, std::ios::out | std::ios::binary | std::fstream::app);
-	data_file.write((char*)&xp[0], xp.size() * sizeof(int));
+	data_file.write((char*)&T[0], T.size() * sizeof(char));
 	data_file.close();
 
 
@@ -303,14 +317,19 @@ layer read_bin(std::string path) {
 	std::ifstream  data_file;
 	data_file.open(path, std::ios::in | std::ios::binary);
 
-	pint* header = new pint[4];
[…]

### 2021-03-06 Papers/02_Sound_Compression/Compression.tex
progress on the paper

--- 2021-03-04 Papers/02_Sound_Compression/SoundRepresentation.tex
@@
 
-\section{Introduction} % ≅ 500 words
+\section{Introduction} % ≅ 500 to 1000 words
 % ≅ 10\% of the text
@@
 \textcite{2016ChenCompressing} for example, presents a scheme for compressing convolutional neural networks, while \textcite{2014CanovasLossy}, besides revising existing methods, proposes two lossy schemes for the compression of quality score data of genomic sequencing.
+The use of lossy compression for wireless networks of sensors is studied in \textcite{2014ZordanPerformance}.
 
-More recently, the work of \textcite{2019CalhounExploring}, explores the benefits of specific lossy compression algorithms in the context of checkpoints of computational simulations that are stored between simulation sessions.
+More recently the work of \textcite{2019CalhounExploring} explores the benefits of specific lossy compression algorithms in the context of checkpoints of computational simulations that are stored between simulation sessions.
 
-Regarding sample based digital musical instruments, to the best of the authors' knowledge, no such specialized algorithms exist.
+Regarding sample based digital musical instruments, however, to the best of the authors' knowledge, no such specialized algorithms exist.
 Those signals, being highly uniform, offer an opportunity for compression that is not present in most general sounds and remain largely unexplored.
@@
 
-New technologies such as 3D, immersive audio has propelled specific compression algorithms, such as the one presented in \textcite{2021HuAudio}.
+New technologies such as 3D immersive audio has propelled specific compression algorithms, such as the one presented in \textcite{2021HuAudio}.
 The renaissance of neural networks led to interesting results, such as the use of GANs as presented in \textcite{2018HuangBandwidth}.
@@
 A continuous or discrete signal, as long as it is at least pseudo periodic, can be interpreted as succession of consecutive waveforms, each one representing a (pseudo) cycle of the wave.
-In the case of perfectly periodic signals, those waveforms are identical and we need only the definition of one waveform to characterise the whole signal.in pseudo periodic waves, On the other hand there are potentially variations between pseudo cycles, that must be accounted for.
-If the wave is sufficiently well behaved, as is the case with most sound arising from accoustic phenomenons, these changes can be attributed to variations in both the amplitude of the waveform and its length
+In the case of perfectly periodic signals, those waveforms are identical and we need only the definition of one waveform to characterise the whole signal.
+While in pseudo periodic waves, On the other hand, there are potentially variations between pseudo cycles, that must be accounted for.
+% Relate to other studies
+
+This rationale of defining a signal as a series of a modified basis function is presented more formally, in the context of wavelets, in \textcite{1993SinhaLow}; in that work, however, the bases are wavelets that are obtained via filter banks applied using a pyramid algorithm.
+A similar approach can be found in \textcite{1998SrinivasanHigh}, where a perceptual metric is used, however.
+
+In both works the wavelets coefficients are encoded in a dinamic dicinary. The present work adopts a more direct approach, obtaining a more accurate unique discrete basis.
+
+If the wave is sufficiently well behaved, as is the case with most sound arising from accoustic phenomenons, the bulk of those changes can be attributed to variations in both the amplitude of the waveform and its length; One hence needs to identify an embrionary waveform, and register how it changes during the course of the pseudo cycles that comprise the whole signal.
+
+We accomplish that via an iterative algorithm, that refines this waveform in each iteration
 
 Starting with an initial estimate of the waveform of the signal being processed, one can identify the pseudo cycles of a signal by convolving the estimated waveform with the signal.
-The process can be made computationally efficient by considering some characteristics of the wave.
+The process can be made more efficient by considering some characteristics of the wave.
+For the initial waveform, for example, the algorithm presented in \textcite{2020TarjanoDigital} is used in cases where a simple sinusoid is less accurate.
 
@@
 % Essential background information
-
-% Relate to other studies
 
@@
 
-We compare the results of the algorithm here presented with some of the most popular codecs.
+% Revisit purpose and methods
+
+% Overview of results
+We compare the results of the algorithm here presented with some of the most widespread lossy codecs, using a variety of sound signals.
+The samples chosen reflect the proposed immeadiate application of the algorithm to sample based digital musical instruments, and thus consist of excerpts from notes from various instruments, as well as singing voice.
+
+Overview of the samples 44100 etc
 
@@
 
-The opus format was chosen, as, with the ACC and MP3, those are considered the most advanced \parencite{2019LaudeComprehensive} and popular \parencite{2018KwanObjective}
-
-% Revisit purpose and methods
-
-% Overview of results
+The opus format was chosen, as, with the ACC and MP3, those are considered the most advanced \parencite{2019LaudeComprehensive} and popular \parencite{2018KwanObjective}.
 
 % Key results in details
+\subsection{Compressed file size}
 
+\begin{table}[H]
+\centering
+\begin{tabular}{lcccc}
+\hline
+\multicolumn{5}{c}{\textbf{Compressed size (bytes)}} \\
+Sample        & Present Work & MP3   & ACC   & Opus \\
+\hline
+cello         & 1159         & 3357  & 3066  & 2021 \\
+doublebass    & 1471         & 1845  & 1915  & 1339 \\
+flute         & 16678        & 11565 & 8988  & 9342 \\
+frenchhorn    & 4783         & 9189  & 7572  & 6219 \\
+mezzosoprano  & 4669         & 3717  & 3237  & 2997 \\
+saxophone     & 19260        & 20277 & 16307 & 12954 \\
+singerbass    & 1989         & 3069  & 2747  & 2302 \\
+singerfemaleA & 9125         & 8325  & 6593  & 6121 \\
+singerfemaleB & 4470         & 6021  & 4773  & 4395 \\
+singermale    & 694          & 1629  & 1777  & 977 \\
+sopranoA      & 3747         & 3069  & 2656  & 2294 \\
+sopranoB      & 2535         & 2709  & 2568  & 2021 \\
+violin        & 1575         & 2709  & 2514  & 1754 \\
+\hline
+Average       & 5550.38	     & 5960.08	&4977.92	&4210.46
+
+\end{tabular}
+\caption{Comparison of the average errors of the algorithm presented in this work with the most common methods of digital envelope identification.
+The discrete waves used were normalized between -1 and 1.
+The average is between the absolute value of those waves and the envelope divided by two.}
+\label{table:Error_comparison}
+\end{table}
+
+\subsection{Encoding and decoding time}
+
+While part of the discrepancy in encoding times can be attributed to the immaturity of the implementation when compared to well stablished ones like lame for mp3, the majority of the difference is generated by the wildly different approach used by the algorithm presented in this work, since it engages in a semi exaustive search of the best basis, or paseudo cycle waveform via a series of dilated convolutions.
 % Comparison with results in other works
@@
 
-noise
+The algorithm, seeking the most accurate basis, has a tendency of filtering out noise from the compressed signal.
+This is not always welcome in the context of sound, since, as is the case of sounds with a percussive component, noise plays an integral part in the psychoacoustics characterization of the sound.
+
+One solution would be to parametrically reintroduce the needed noise, with an approach similar to the presented in \textcite{1998LevineSinesTransientsNoise}, for example.
 

### 2021-03-14 THESIS.tex
initial

\input{00_Include.tex}

\begin{document}

%\include{01_Resumo.tex}
\pagestyle{empty}
\include{02_Abstract.tex}

% \newpage
% \listoffigures % lista de ilustracoes

% \newpage
% \listoftables % lista de tabelas

\newpage
\tableofcontents

\newpage
\pagestyle{fancy}
\include{03_Introduction.tex}
\include{04_Theory.tex}
\include{05_Characterization.tex}
\include{06_SignalEnvelope.tex}

% \section{References}
\printbibliography %[heading=none]

\newpage
\printglossaries

\end{document}

### 2021-03-14 images/Helper.py

@@ -1,0 +1,1154 @@
+import numpy as np
+import random
+import wave
+from numpy.fft import rfft, irfft
+from scipy.interpolate import interp1d, UnivariateSpline
+from scipy.linalg import block_diag
+from scipy.optimize import nnls
+from scipy.signal import savgol_filter
+import numpy.polynomial.polynomial as poly
+from scipy.signal import hilbert
+from scipy.spatial import Delaunay
+import plotly.graph_objects as go
+
+
+def read_wav(path): 
+  """returns signal & fps"""
+  wav = wave.open(path , 'r')
+  signal = np.frombuffer(wav.readframes(-1) , np.int16)
+  fps = wav.getframerate()
+  return signal, fps
+
+
+def save_wav(signal, name = 'test.wav', fps = 44100): 
+  '''save .wav file to program folder'''
+  o = wave.open(name, 'wb')
+  o.setframerate(fps)
+  o.setnchannels(1)
+  o.setsampwidth(2)
+  o.writeframes(np.int16(signal)) # Int16
+  o.close()
+
+
+class Pulse:
+  def __init__(self, x0, W, is_noise = False):
+    # self.W = W
+    self.x0 = x0 - .5
+    self.len = W.size
+    self.x1 = self.x0 + self.len
+    idx = np.argmax(np.abs(W)) #
+    self.y = W[idx]            # np.average(W)
+    self.x = x0 + idx
+    self.noise = is_noise
+
+
+def signal_to_pulses(W):
+  ''' returns a list of instances of the Pulses class'''
+  n = W.size
+  pulses = []
+  sign = np.sign(W[0])
+  x0 = 0
+  for x in range(n):
+    if sign != np.sign(W[x]):
+      p = Pulse(x0, W[x0 : x])
+      pulses.append(p)
+      x0 = x
+      assert(np.all(np.sign(W[x0 : x]) == sign))
+      sign = np.sign(W[x])
+  pulses.append(Pulse(x0, W[x0 : n]))
+  return pulses
+
+
+def split_pulses(pulses):
+  ''' Mark pulses as noise based on Nyquist | divide into positive and negative pulses and noises '''
+  pos_pulses = []
+  neg_pulses = []
+  pos_noises = []
+  neg_noises = []
+  for p in pulses:
+    if p.len <=2:
[…]

### 2021-03-14 THESIS.tex
joined thesis in one tex file

@@
-\input{00_Include.tex}
+\documentclass[12pt]{article}
+
+% \usepackage[utf8]{inputenc}
+% \usepackage[T1]{fontenc}
+% \usepackage{authblk}
+
+\usepackage[backend=biber]{biblatex}
+%\usepackage{biblatex}
+
+\usepackage[a4paper, top=3cm, left=3cm, bottom=2cm, right=2cm]{geometry}
+\usepackage{indentfirst}
+\setlength{\parindent}{1.25cm}
+
+\usepackage[no-math]{fontspec}
+\defaultfontfeatures{Ligatures=TeX}
+\setmainfont{Tex Gyre Termes}
+
+\usepackage[nodisplayskipstretch]{setspace}
+\setstretch{1.427465}
+
+\usepackage[explicit]{titlesec}
+\titleformat{\section}[hang]{\normalfont\bfseries}{\thesection}{1em}{\MakeUppercase{#1}}
+\titleformat{\subsection}[hang]{\normalfont}{\thesubsection}{1em}{\MakeUppercase{#1}}
+\titleformat{\subsubsection}[hang]{\normalfont\bfseries}{\thesubsubsection}{1em}{#1}%{\MakeLowercase{#1}}
+
+\usepackage{titletoc}
+\titlecontents{section}[1em]{\vspace{12pt}}{\bfseries\normalsize\contentslabel[\thecontentslabel]{3em}\uppercase}{\hspace*{-2em}\uppercase}{\titlerule*[.75em]{.}\contentspage}
+\titlecontents{subsection}[1em]{\vspace{12pt}}{\normalfont\normalsize\contentslabel[\thecontentslabel]{3em}\uppercase}{\hspace*{-2em}\uppercase}{\titlerule*[.75em]{.}\contentspage}
+\titlecontents{subsubsection}[1em]{\vspace{12pt}}{\bfseries\normalsize\contentslabel[\thecontentslabel]{3em}}{\hspace*{-2em}}{\titlerule*[.75em]{.}\contentspage}
+
+\usepackage{amsmath}  % enables declare math operator
+\DeclareMathOperator{\sign}{sign}
+
+\usepackage{amsfonts} % \mathbb
+\usepackage{graphicx} % manipulate imported images
+
+\usepackage{fancyhdr} % \pagestyle{fancy}
+\fancyhf{}
+\fancyhead[R]{\thepage}
+\renewcommand{\headrulewidth}{0pt}
+
+\usepackage[toc, section=section]{glossaries}
+\makeglossaries
+
+\usepackage{hyperref}
+\usepackage{hypcap}
+\hypersetup{
+  colorlinks=true, 
+  allcolors=black,
+  pdfauthor={Carlos Henrique Tarjano Santos},
+  pdftitle={Robust Digital Envelope Estimation Via Geometric Properties of an Arbitrary Real Signal}
+}
+
+\bibliography{00_BIBLI}
+
+\title{Sound Synthesis}
+\author{Carlos Tarjano}
+
+\newacronym{FPS}{FPS}{frames per second}
+\newacronym{LMS}{LMS}{Least Mean Squares}
+
+% \newglossaryentry{VST}{name=VST, description={Virtual Studio Technology}}
 
@@
 \pagestyle{empty}
-\include{02_Abstract.tex}
+
+\begin{abstract} % 150 words
+
+\end{abstract}
+
+{\bf Keywords:}
 
@@
 \pagestyle{fancy}
-\include{03_Introduction.tex}
-\include{04_Theory.tex}
-\include{05_Characterization.tex}
-\include{06_SignalEnvelope.tex}
+
+\newacronym{ENIAC}{ENIAC}{Electronic Numeric Integrator and Calculator}
+
+\section{Introduction}
+% Tell the reader the problem you are tackling in this project.
+
+Humans are a creative bunch.
+Since the beginning of computing, even when mainframes were the only available medium, restricted to organizations with considerable budgets, people found ways and motivation to overcome the technical limitations and develop artistic applications, like video games, in their (and the machine's) spare time.
+
+On a parallel note, also relevant to this thesis, those first machines, like the \gls{ENIAC} in 1946, brought with then the promise of the artificial intelligence as their ultimate goal \parencite{2010DonovanReplay}.
+
+Music, on the other hand, is another staple of humanity's creativity. % TODO first musical instruments
+ Is only natural that, as soon as this technology became reasonably stable, musical applications started to emerge.
+
+% State clearly how you aim to deal with this problem. 
+% Limit the scope of your study. 
+% Sketch out how the thesis is structured to achieve your aim. 
+% objective
+Ambitious as it may seem, it is the goal of this monograph to motivate a change of approach in domains that, despite their modest intersections, share more attributes than is apparent at a superficial glance: Both machine learning and digital musical acoustics are areas that, despite their current relevance, are fresh out of their respective infancies, and a scrutiny of both promptly reveals the difficulties that arise from this state of affairs.
+A lack of their own terminology, with frequent borrows from other, more established areas, can be readily cited, as the pronounced prominence of few individual contributions.
+
+Those shortcomings have, however, a bright side to them, in that they still expose those areas to reinterpretation, rendering them somewhat open to the influence of outside ideas.
+
+Digital musical acoustics borrows heavily from the digital signal processing terminology, the last having its roots in the analogic world.
+Many algorithms are conceived in terms of filters, circuits and other similar legacy constructs, a fact that burdens the conceptual, theoretical framework of the area.
+Besides the introduction of a noveau envelope extraction method, we find that his formulation in terms of (differential) geometric concepts is, perhaps, as great a contribution as the algorithm per se.
+
+% Perhaps less daring, and owing more to the advanced age of the author than to his research prowess, is the presentation of this monograph in a format more akin to a book; In addition to turning this somewhat dense tome into a more palatable experience.
+
+\section{Bibliographic Approach}
+
+with the advance of technology, some classic problems of bibliographic research were mitigates, while new ones were introduced.
+a modern problem encontered by the author during the development of this work is the fragmentation of data.
+In the bygone days of physical research, a paper existed in primarily  physical location.
+we adress the meta problem of organizing the bibliography introducing a piece of bibliographic software based on an novel paradigm.
+
+neural networks faced many winters since its introduction.
+The pattern seems to repeat itself: hardware improvements enable attractive new results, the interest rises, and some technical advancements are made.
+Although, in its current renaissance, the field is still blooming, there are some that are already pointing to a decrease in the research output.
+
+We feel that this tool still offers a great potential in its current state of development; a potential thtat will only be fully explored once it is realised that workflows and auxiliary theories are needed, to be used in conjunctions with neural networks.
+
+In this sense, we introduce a novel representation for pseudo periodic time series, of which sound representation are our primary interest.
+Although much of the theory is drawn from conventional areas related to digital signal processing, image processing, etc, the representation was formulated from the beginning having in mind the strengths and limitations of neural networks in general.
+
+writing a phd thesis is a peculiar endeavour in the academic world, in the sense that it is a requirement that must be fullfilled only once: while a thriving academic carreer pressuposes a constant output of papers, and possibly books, or book chapters, the closest contact an academic has with the format of a thesis is probably during the development of his masters dissertation; that one endeavour being, all in all, substantially different.
+
+It comes as no surprise, then, that bibliographic software, being in general geared towards the task of aiding in the productions of academic articles, can be subpar for a thesis, where the hability to handle a mutch larger volume of bibliographic material is needed.
+
+\section{Conventions and definitions}
+The field explored in this work, besides being extensive and fragmented, is largely interdisciplinar, with terminology strong influenced by widespread usage in the community rather than academic tradition.
+To maintain coherence thougout the text, it is convenient to stablish some definitions, as some of the terms here applyed may deviate from their usage in particular fields.
+
+Effort was made to comply with the most commom nomenclature when terms are encountered frequently in literature.
+
+\section{Theory}
+
+\subsection{Waves}
+
+It can be insightful to analyse the behaviour of the crytical points and the zeroes of an arbitrary, periodic wave over time.
+We will assume, for the sake of simplicity, that each cycle of said wave has a unique maximum and minimum, and crosses the horizontal axis one single time.
+Those assumption dont compromise the generality of the analysis, as in the case of multiple maxima, minima and/or zeroes, one of thosee can be choosen randomly to represent the global one, as long as consistency between cycles is maintained.
+
+We are going to also assume that a wave with a zero phase has an maximum instantaneous amplitude at time zero, as is the case of the cosine wave, for example.
+Altough that assumption somewhat departs from the physical intuition, it is useful in making our description consistent with the theory behind the Fourier Transform.
+
+For a continuous wave, if we were to plot the time at which any of those values ocurred as a function of their ordinality, the result would be parallel straight lines, their slope somehow related to the frequency of the underlying wave.
+We will illustrated the point in the case of discrete waves, as this case will be of more immediate use in this work.
+
+\subsubsection{Continuous and discrete waves}
+
+First, it is necessary to establish a correspondence between a continuous and a discrete wave, as the latter will be the object of most of this work.
+To do that we will define a simple continuous sinusoidal wave of the form $ S(t) = a \cos \left( \phi + 2 \pi \mathcal{F} t \right) $, where $ a \in \mathbb{R}^+ $ is the amplitude; $ \phi \in \mathbb{R}^+, 0 \le \phi < 2 \pi $ is the phase; $ \mathcal{F} \in \mathbb{R}^+ $ is the absolute frequency of the wave, in Hz and $ t \in \mathbb{R} $ is time, in seconds.
+
+The cosine function is used for a twofold reason: its use helps to underline the periodic nature of the wave in consideration and is the parameter of the phase when the FT is used.
+It is important to keep in mind, however, that this discussion applies any arbitrary function of the form $ f(x) = f(x + P) \forall x \in D(f) $, where $ D $ is the domain of the function $ f $.
+% TODO This will be further illustrated in the discrete part below.
+
+If we were to sample $ S(t) $ at regular intervals, that is, to measure the value of the function at a constant rate of $ h $ \gls{FPS}, neglecting errors of measurement, we would have a discrete representation of the continuous wave of the form $ W[i] = a \cos \left( \phi + 2 \pi f \frac{i}{n} \right) $ where, alongside the quantities introduced before, we would have $ i \in \mathbb{N} $ as the frame number; $ n \in \mathbb{N} $ as the number of discrete samples and $ f \in \mathbb{R}^+ $ being the relative frequency of our discrete wave in cycles per $ n $ samples.
+
+The relative, local frequency $ f $ of the discrete representation is related to the absolute frequency $ \mathcal{F} $ via the equation $ f = \frac{n}{\text{fps}} \mathcal{F} $ while the absolute time, in the continuous case can be obtained from the frame number $ i $ of the discrete signal via $ i = t \ \text{fps} $.
+
+\subsubsection{Points of interest of periodic waves}
+
+\subsection{Least Mean Squares}
+
+The linear \gls{LMS}, also known as linear regression \parencite{2018StephenBoydIntroduction} is a data fitting technique. % TODO
+
+We present here the theory for real numbers, noting that it can be easily extended for complex numbers \parencite{1995LawsonSolving}, should the necessity arise.
+That, however, will not be the case in this work.
+
+Given two vectors $ X $ and $ Y $, such that:
+$$ X = [x_0, x_1, \cdots ,x_{m-1}]^T, \ x_i \ \in \ \mathbb{R} \ \forall \ i, \ 0 \le i < m \in \mathbb{N} $$
+$$ Y = [y_0, y_1, \cdots ,y_{m-1}]^T, \ y_i \ \in \ \mathbb{R} \ \forall \ i, \ 0 \le i < m \in \mathbb{N} $$
+We ideally want to find the vector of coefficients $ A = [a_0, a_1, \cdots ,a_k]^T, \ a_j \ \in \ \mathbb{R} \ \forall \ j, \ 0 \le j \le k \in \mathbb{N} $ that satisfies the equality $ x^0_i a_0 + x^1_i a_1 + \cdots + x^{k-1}_i a_{k-1} + x^k_i a_k = y_i \quad \forall \quad i \in \{0, \cdots , m-1\} $.
+
+Put in that way, what we have is a system of linear equations that can be rewritten in matrix form as seen in equation \ref{eq:lms}.
+
+\begin{equation}
+  \label{eq:lms}
+\stackrel{Q_{m \times (k+1)}}
+{
+\begin{bmatrix}
+1 & x_0 & \cdots & x_0^k \\
+1 & x_1 & \cdots & x_1^k \\
+\vdots & \vdots & \ddots & \vdots  \\
+1 & x_{m-1} & \cdots & x_{m-1} \\
+\end{bmatrix}
+}
+\stackrel{A_{(k+1) \times 1}}
+{
+\begin{bmatrix}
+a_0 \\
+a_1 \\
+\vdots  \\
+a_k \\
+\end{bmatrix}
+}
+=
+\stackrel{Y_{m \times 1}}
+{
+\begin{bmatrix}
+y_0 \\
+y_1 \\
+\vdots  \\
+y_{m-1} \\
+\end{bmatrix}
+}
+\end{equation}
+
+More compactly, we have that $ Q A = Y $.
+This system will generally be over-determined, only having an exact solution in the case where $ Y $ is a linear combination of the columns in $ Q $ \parencite{2018StephenBoydIntroduction}.
+
+Since an exact solution is often beyond reach, we can at least choose $ A $ in a way that makes $ Q A $ and $ Y $ close.
+A way to do that is to minimize the absolute errors (or their square) between those two quantities; $ \underset{A}{\min} ||Q A - Y||^2_2 $
+
+Using calculus, for example, on can arrive at the equation $ \hat{A} = (Q^T Q)^{-1} Q^T Y $, where $ \hat{A} $ is the vector of coefficients that minimizes the squared errors.
+A detailed derivation is available at \cite{2018StephenBoydIntroduction}.
+
+\subsection{Constrained Least Mean Squares}
+
+\newacronym{CLMS}{CLMS}{Constrained Least Mean Squares}
+
+In the constrained version of the \gls{LMS}, known as the \gls{CLMS} problem we are still interested in finding the coefficient vector $ \hat{A} $ that minimizes $ ||Q A - Y||^2_2 $, but respecting a restriction on this vector of the general form of $ V A = B $.
+
+We can then define the Lagrangian and set its derivatives to zero to obtain, as described in \textcite{2013SelesnickLeast}, the closed form solution
+
+$$ \hat{A} = \left(Q^{T} Q\right)^{-1} \left(Q^{T} Y - V^{T} \left(V \left(Q^{T} Q\right)^{-1} V^{T}\right)^{-1} V \left(Q^{T} Q\right)^{-1} Q^{T} Y - B \right) $$
+
+\subsection{Savitzky–Golay filter}
+
+\section{Characterization of a digital Signal}
+% TODO continuous to digital
+
+\subsection{Envelope of a digital signal}
+
+Despite being an elusive concept, the temporal amplitude envelope of a signal is essential for its complete characterization, being the primary information-carrying medium in spoken voice and telecommunications, for example.
+Intuitively, the temporal envelope can be understood as a smooth function that multiplies the signal, being responsible for its outer shape.
+Envelope detection techniques have applications in areas like health, sound classification and synthesis, seismology and speech recognition.
+Nevertheless, a general method to digital envelope detection of signals with rich spectral content doesn't exist, as most methods involve manual intervention, in the form of filter design, smoothing, as well as other specific design choices, based on a priori knowledge about the nature of the specific waves under investigation.
+To address this problem, we propose a framework that uses intrinsic characteristics of a signal to estimate its envelope, completely eliminating the necessity of parameter tuning.
+The approach here described draws inspiration from geometric concepts to isolate the frontiers and thus estimate the temporal envelope of an arbitrary signal; to that end, alpha-shapes, concave hulls, and discrete curvature are explored.
+We also define entities, such as a pulse and frontiers, in the context of an arbitrary digital signal, as a means to reduce dimensionality and the complexity of the proposed algorithm.
+Specifically, a new measure of discrete curvature is used to obtain the average radius of curvature of a discrete wave, serving as a threshold to identify the wave's frontier points.
+We find the algorithm accurate in the identification of the frontiers of a wide range of digital sound waves with very diverse characteristics, while localizing each pseudo-cycle of the wave in the time domain.
+The algorithm also compares favourably with classic envelope detection techniques based on filtering and the Hilbert Transform.
+Besides the most direct applications of this work to audio classification and synthesis, we foresee impact in compression techniques and machine learning approaches to audio.
+The discrete curvature definition presented could also be extended to three-dimensional settings, to improve shape detection algorithms based on alpha-shapes.
+
+\subsubsection{Envelope Detection}
+Envelope detection is ubiquitous in both analogue and digital signal processing \parencite{2011CaetanoImproved}.
+Nevertheless, the literature in this area is very fragmented \parencite{2017LyonsDigital}.
+Besides, most envelope detection techniques are designed to account for very specific settings, like pure sinusoids with moderate noise content, a limitation that excludes most physical signals, as is the case of recorded sound, for example; that limitation arises in part due to the lack of a strict mathematical definition of a temporal envelope \parencite{2013Mengempirical}.
+
+In many natural signals, however, the temporal amplitude envelope of a signal plays a prominent role in the characteristics exhibited: according to \textcite{2017QiRelative}, for example, the envelope is at least as important as the fine structure of a sound wave in the context of the intelligibility of mandarin tones.
+This is also the case for English \parencite{1995ShannonSpeech}, where even envelopes modulating mostly noise were still capable of conveying meaning.
+
+The envelope helps to impart emotion and identity to the human voice \parencite{2018ZhuContributions}, and envelope preserving characteristics in concert halls are associated with their pleasantness \parencite{2011LokkiEngaging}.
+
+When dealing with broadband signals, approaches tailored to specific applications are prevalent, such as the one presented by \textcite{2014YangFast} for the distributed monitoring of fibre optic or the one formulated by \textcite{2018AssefModeling} in the context of medical ultrasound imaging.
+
+Addressing the different units in the horizontal and vertical axes, one can transform the DSP problem of envelope detection into the geometric problem of defining the shape of a set of points in $ \mathbb{R}^2 $.
+For that purpose, \textcite{1983Edelsbrunnershape} introduced the concept of alpha-shapes, a mathematically well-defined extension to the convex hull of a finite set of points, closely related to the Delaunay triangulation and Voronoi diagrams of those points.
+
+This approach is used in areas such as the detection of features in images \parencite{2016VarytimidisAlpha}, reconstruction of surfaces from a cloud of points \parencite{2015WuAutomated} and Spectroscopy \parencite{2019XuModeling}, with the last work, which involves the estimation and removal of the Blaze function (a kind of envelope) of an echelle spectrograph, being particularly illustrative of the potential of cross-fertilization.
+
+Steps in the direction of translating geometric algorithms to the context of envelope detection were also made by \textcite{2015YangSkeleton} via an algorithm based on the construction of a skeleton underlying the wave, and also via the direct translation of computer vision methods to the task \parencite{2015YangRepresenting}.
+
+Following this path, we present a general approach to envelope detection, exploiting the intrinsic characteristics of a generic, spectrally complex wave, in order to avoid the need for manual intervention or parameter tuning.
+
+While the robustness of the proposed approach allows it to be used as a plug-in replacement for many methods encountered in the literature, we feel that it would be particularly useful for sound synthesis.
+
+The envelope is shown to add complexity to the spectral representation of a wave \parencite{2019TarjanoNeuro}, and an accurate description of the envelope, while describing the evolution of the instantaneous amplitude of a signal in time, would also greatly simplify further spectral analysis.
+Moreover, the algorithm developed in this work naturally divides a signal into its pseudo-cycles, pinpointing them in the time domain, providing the building blocks for the reconstruction of the fine structure of the wave.
+
+% The rest of this paper is divided as follows: After explaining and defining the various entities to be used subsequently, of which the concept of frontiers is arguably the most important, in the methodology section, we dedicate a section to present a new discrete curvature estimation approach, and how it was used to identify the frontiers of a discrete wave.
+
+We then proceed to illustrate the results of the application of the algorithm in a set of six diverse sound waves, ranging from voice to unpitched instruments comparing, afterwards, its performance with that of classic digital envelope estimation algorithms, both from a visual and numerical standpoint.
+
+We then discuss the implications of the work, emphasizing the focus on the construction of new interpretations of the relationship between time domain and frequency domain sound descriptions, and our vision of the impact of this methodology in the field of sound synthesis, providing directions for future developments.
+
+\subsection{Frequency content}
+
+\subsection{Representations}
+
+% one hot
+
+\section{Envelope Detection}
+
+Despite being an elusive concept, the temporal amplitude envelope of a signal is essential for its complete characterization, being the primary information-carrying medium in spoken voice and telecommunications, for example.
+Intuitively, the temporal envelope can be understood as a slowly varying function that multiplies the signal, being responsible for its outer shape.
+Envelope detection techniques have applications in areas like medicine, sound classification and synthesis, seismology and speech recognition.
+% Problem | Objectives
+Nevertheless, a general approach to digital envelope detection of signals with rich spectral content doesn't exist, as most methods involve manual intervention, in the form of filter design, smoothing, and other specific design choices, based on prior knowledge of the signals under investigation.
+To address this problem, we propose an algorithm that uses intrinsic characteristics of a signal to estimate its envelope, eliminating the necessity of parameter tuning.
+% Methods | Procedure | Approach
+The approach here described draws inspiration from geometric concepts to estimate the temporal envelope of an arbitrary signal; specifically, a new measure of discrete curvature is used to obtain the average radius of curvature of a discrete wave, that will serve as a threshold to identify the wave's samples that are part of the envelope.
+% Results | Findings | Product
+The algorithm compares favourably with classic envelope detection techniques based on smoothing, filtering and the Hilbert Transform, besides being physically plausible.
+We provide visualizations of the envelope extracted via the algorithm for various real-world signals, with very diverse characteristics, such as voice, spoken and sang, and pitched and non-pitched musical instruments, and discuss some approaches to assess the quality of the obtained envelopes.
+A Python module implementing the algorithm was made available via the Python Package Index (PyPI); interactive visualizations of envelopes for a diverse range of digital waves, as well as the source code for the Python implementation, are available online at \textcite{2020TarjanoEnvelope}.
+% Conclusions | Implications
+Besides the most direct applications of this work to audio classification and synthesis, we foresee impact in compression techniques and machine learning approaches to audio.
+We briefly discuss some potential paths in this direction.
+The discrete curvature definition presented could also be extended to three-dimensional settings, to improve shape detection algorithms based on alpha shapes.
+
+Envelope detection, also known as demodulation, is ubiquitous in both analogue and digital signal processing \parencite{2011CaetanoImproved}.
+Nevertheless, the literature in this area is fragmented \parencite{2017LyonsDigital}.
+Besides, most envelope detection techniques are designed to account for very specific settings, like pure sinusoids with moderate noise content, a limitation that excludes most physical signals, as is the case of recorded sound, for example; that limitation arises in part due to the lack of a strict mathematical definition of a temporal envelope \parencite{2013Mengempirical}.
+
+In many contexts, however, the temporal amplitude envelope of a signal plays a prominent role: according to \textcite{2017QiRelative}, for example, the envelope is at least as important as the fine structure of a sound wave in the context of the intelligibility of Mandarin tones.
+This is also the case for the English language \parencite{1995ShannonSpeech}, where even envelopes modulating mostly noise are still capable of conveying meaning.
+
+The envelope also helps to impart emotion and identity to the human voice \parencite{2018ZhuContributions} and music: Concert halls with envelope preserving characteristics are deemed more pleasant \parencite{2011LokkiEngaging}, for example.
+
+When dealing with broadband signals, approaches tailored to specific applications are prevalent, such as the one presented by \textcite{2014YangFast} for the distributed monitoring of fibre optic or the one formulated by \textcite{2018AssefModeling} in the context of medical ultrasound imaging.
+
+The problem of envelope detection, normally associated with the field of digital signal processing, can be made equivalent to the geometric problem of defining the (outer) shape of a set of points in $ \mathbb{R}^2 $, provided a method for addressing the difference of units in the horizontal, time related, and vertical, amplitude related, axes of a digital signal is available, rendering the original DSP problem readily approachable via geometric techniques.
+
+One direct way of defining the shape of a set of points in $ \mathbb{R}^2 $ is using the concept of a convex hull: as defined in \textcite{2008BergComputational}, the convex hull of a set of points $ S $ in two-dimensional space is the smallest convex subset of the plane that contains all the points in $ S $.
+Convex hulls are unique, in the sense that a set of points define one and only one convex hull.
+
+Generalizing that definition, \textcite{1983Edelsbrunnershape} introduced the concept of alpha shapes, that can be seen as a concave hull; a mathematically well-defined generalization of the convex hull of a finite set of points, closely related to the Delaunay triangulation and Voronoi diagrams of those points.
+
+The intuition behind the alpha shapes algorithm is that only points that can be touched by a circle of radius $ \alpha $ coming from an infinite distance towards the set of points of interest are part of the frontier, that is, of the concave hull of this set.
+
+Once a radius $ \alpha $ is defined, alpha shapes conserve the uniqueness property, in fact tending to the convex hull when $ \alpha \to \infty $.
+Consider, for example, the artificially generated discrete wave in figure \ref{fig:Hulls} below, where the links between consecutive samples were deemphasized to highlight the “set of points” nature of a signal: The convex hull and a concave hull, in the form of an alpha shape with a particular value of $ \alpha $ are illustrated.
+
+\begin{figure}[H]
+  \centering
+    \includegraphics{images/00ConvexHull.pdf}
+  \caption{A discrete signal interpreted as a set of points in the Cartesian plane, and the convex and concave hull at $ \alpha=300 $ defined by it.}
+  \label{fig:Hulls}
+\end{figure}
+
+The alpha shapes approach is used in areas such as detection of features in images \parencite{2016VarytimidisAlpha}, reconstruction of surfaces from a cloud of points \parencite{2015WuAutomated} and spectroscopy \parencite{2019XuModeling}.
+
+This last work, which involves the estimation and removal of the Blaze function (a kind of envelope) of an echelle spectrograph, being particularly illustrative of the potential synergy between geometric and DSP approaches.
+
+Other steps in the direction of translating geometric algorithms to the context of envelope detection were made by \textcite{2015YangSkeleton} via an algorithm based on the construction of a skeleton underlying the digital wave of interest, and also via the direct translation of computer vision methods to the task of envelope detection \parencite{2015YangRepresenting}.
+
+Following this path, we present a general approach to envelope detection, exploiting the intrinsic characteristics of a generic, spectrally complex wave, in order to completely avoid the need for manual intervention or parameter tuning.
+
+The rest of this paper is divided as follows: We start by defining as specifically as possible what a temporal envelope is, especially in the context of this work and providing a mathematical definition, in terms of the Hilbert transform, for the case of a narrowband signal.
+
+We proceed to illustrate the representation of a discrete wave used in this work, largely based on the concept of dividing the wave into its constituent pulses, and how it can simplify the algorithm, reducing the high dimensionality generally present in digital representations of waves.
+
+Next, we review methods for the estimation of the curvature of a discrete wave, after which we present our own; this method will then be used to estimate the equivalent circle, that will be used to define the envelope of the wave, via the procedure described in sequence.
+
+In the results section, we present some illustrations of the envelopes obtained with our method, comparing both accuracy and running time with the most common algorithms encountered in the literature.
+We also suggest an alternative use of the algorithm to identify the negative and positive frontiers of a wave, that can be seen as partial envelopes considering only the positive and negative pulses of a wave, respectively.
+
+We follow with suggestions about how the approach here presented can be useful beyond envelope detection, exemplifying with some applications, such as the extractions of the average waveform of a digital wave, and concluding with commentary on potential future developments of the method.
+
+\subsection{Methods}
+After defining what will be considered an envelope in the context of this work, in this section we explore some simplifications that can be applied to a discrete wave, defining the representation that will be used in the subsequent part of the work.
+
+We next explore the concept of discrete curvature, putting forth our own method, tailored to the definition of the equivalent circle of a wave; we end this section by showing how this circle can be used to identify the envelope of a discrete wave, in a similar approach to that of the alpha shapes theory, but taking advantage of some unique structure present in discrete waves.
+
+\subsubsection{Characterization of a temporal envelope}
+
+The interest in the theory of envelope detection arose in the analog domain, with the widespread adoption of radio communications, where the problem is typically restricted to simple sinusoids, often with well-defined frequencies \parencite{2011TurnerDemodulation}.
+
+In this context, the most mathematically sound definition of an envelope involves the representation of its underlying wave as an analytic signal, as introduced by Gabor in 1946 \parencite{2007HahnHistory}.
+In his work, \textcite{1946GaborTheory} applies the then relatively new mathematical machinery of the quantum mechanics to unify the time and frequency-domain representations of a wave, showing how the Hilbert transform could be applied to a real signal in order to obtain an equivalent complex signal, later known as the analytic signal.
+This analytic signal has the form $ A(t) = S(t) + \mathbb{H}(S(t)) \ \text{i} $ \parencite{2016HePraat} where $ S(t) $ is the original real signal.
+$ \mathbb{H}(S(t)) $, the Hilbert transform of the original signal, becomes the imaginary part of the analytic signal; the envelope of a signal thus represented can be straightforwardly obtained by the computation of the complex modulus of $ \mathbb{H}(S(t)) $.
+
+Despite its widespread use, and effectiveness in the context of narrowband signals, envelope detection techniques based on the Hilbert transform don't behave well for broadband signals \parencite{2012DauSpeech}, being even physically paradoxical in some cases \parencite{1996Loughlinamplitude}.
+
+Figure \ref{fig:AnalyticSignal} exemplifies the Hilbert envelope, for a pure sinusoid modulated by a polynomial, illustrating the difference in the shape of the envelope in the presence and absence of Gaussian white noise.
+It is readily noticeable from the figure that, as soon as noise is introduced in the original signal, it is reflected in the envelope.
+
+\begin{figure}[H]
+  \centering
+    \includegraphics{images/01AnalyticSignal.pdf}
+  \caption{Envelope of a pure sinusoid with a local frequency of 20 cycles modulated by a polynomial of degree 3, as obtained by the Hilbert Transform approach.
+In the first half, the sinusoid is free of noise, while in the second half Gaussian noise with a standard deviation of $ 1/10 $ of the maximum amplitude of the wave was added to the base signal.}
+  \label{fig:AnalyticSignal}
+\end{figure}
+
+In general, the problem of envelope detection, or demodulation, can be interpreted as the task of, given a continuous wave $ W(t) $, decomposing it in two components $ E(t) $ and $ C(t) $, such that $ W(t) = E(t) C(t) $ \parencite{2011TurnerDemodulation}.
+
+$ E(t) $ represents the slow varying part of the wave, also known as the (temporal) envelope, modulator component, or amplitude modulation (AM) of $ W $, while $ C(t) $ models its fast varying part, called throughout the literature as its (temporal) fine structure, carrier component, or frequency modulation (FM).
+
+In the case of broadband signals, the problem of envelope detection of an arbitrary wave is ill-posed, in the sense that an infinite number of pairs of $ E(t) $ and $ C(t) $ can result in the same specific $ W(t) $ \parencite{2011TurnerDemodulation, 2013Mengempirical, 1996Loughlinamplitude}.
+
+We address this problem by assuming $ C(t) $ normalized between $ \{-1, 1\}$.
+This assumption doesn't cause any loss of generality since, given an arbitrary $ C(t) $, we can obtain its normalization by dividing the function by its absolute global maximum, that is, $ \hat{C}(t) = C(t) / \max(|C(t)|) $, provided that $ \max(|C(t)|) \ne 0 $.
+
+In this work we are concerned with the discrete version of this problem: given a finite digital wave represented by the vector $ \textbf{w} \in \mathbb{R}^n $, instead of the continuous function $ W(t) $, obtaining the temporal envelope of $ \textbf{w} $.
+
+The preceding definitions can be translated to this discrete scenario assuming that the discrete quantities arise from observing the continuous ones at regular time intervals.
+In that case, the equality $ i = t \ \text{fps} $ can be used to link both settings, where $ i $ is the index of each observation, $ t $ stands for the time in seconds, and fps is the frame rate, or the number of observations made in the period of a second.
+We can thus define:
+
+\begin{subequations}\label{eq:Envelope}
+\begin{align}
+  & \mathbf{w}, \mathbf{e}, \mathbf{c} \in \mathbb{R}^n \\
+  & \mathbf{w} = (w_0, w_1, \cdots, w_{n-1}) \\
+  & \mathbf{e} = (e_0, e_1, \cdots, e_{n-1}) \\  
+  & \mathbf{c} = (c_0, c_1, \cdots, c_{n-1}) \\
+  & \mathbf{w} = \mathbf{e} \odot \mathbf{c} \ \therefore \ w_i = e_i c_i \ \forall \ i, \ 0 \le i \le n-1
+\end{align}
+\end{subequations}
+
+The $ \odot $ operator in \ref{eq:Envelope} stands for the Hadamard product, denoting element-wise multiplication of two vectors.
+Figure \ref{fig:Envelope} provides an example of the vectors just defined: \textbf{w} is obtained by modulating the normalized carrier wave \textbf{c} using the envelope \textbf{e}.
+
+Note from Figure \ref{fig:Envelope} that, at some local maxima, the envelope is well-defined, being in fact equal to the value of the wave $ w_i $, since $ c_i = 1 $ at those points, rendering $ e_i = 1 w_i $.
+
+The carrier wave \textbf{c} in the figure is periodic, and normalized between $ \{-1, 1\}$.
+Each cycle of the wave, then, reaches the value 1 only at its maximum making $ w_i $ equal to the value of the envelope at one point in each cycle.
+
+Image \ref{fig:Envelope}, along with the concave hull illustrated in figure \ref{fig:Hulls}, helps forming the intuition behind the method developed in this work: the idea is to identify the local extrema that touch the envelope, those being also the extrema of each (pseudo-)cycle.
+
+This is accomplished with the use of a circle with a defined radius, large enough to prevent it from reaching the smallest local extrema when "rolled" around the (geometric) representation of the wave.
+Only the extrema touched by the circle will be marked as part of the envelope.
+
+In the next section, a compatible geometric representation will be formulated, along with the necessary tools to estimate the average curvature of the wave and, consequently, the appropriate radius of the circle.
+
+\begin{figure}[H]
+  \centering
+    \includegraphics{images/02Envelope.pdf}
+  \caption{Example of a discrete wave \textbf{w} composed by the element-wise multiplication of an envelope \textbf{e} and a carrier \textbf{c}.
+The local extrema of \textbf{w} are highlighted with a circle.}
+  \label{fig:Envelope}
+\end{figure}
+
+\subsubsection{Simplified Representation of a Wave}
+
+Instrumental to the method here presented is the definition of a pulse: each time the value of \textbf{w} changes sign, that is, every time the discrete wave \textbf{w} crosses the horizontal axis, the beginning of a pulse is defined, with the next crossing defining its end.
+This definition is in line with the one presented in \textcite{national1996telecommunications}, where a pulse is defined as a rapid change in the amplitude of a signal, followed by a fast return to the baseline value; zero, in our case.
+
+From \ref{fig:Envelope} and the discussion in the preceding chapter, we can see that only the maximum point of a positive pulse or the minimum point of a negative one can potentially be equal to the envelope and are, thus, our only points of interest.
+We can then proceed, for the rest of the method, considering only those points, to great computational economy.
+
+For this reason we define P as the set of the points $ \{P_0, P_1, \cdots, P_{m-1}\} $ where $ P_j = (i_j, \lvert w_j \lvert) $, that is, $ \lvert w_j \lvert $, the absolute value of the extrema of each pulse, becomes the ordinate of each point and $ i_j $, its original index, becomes the abscissa.
+
+This relation is illustrated in figure \ref{fig:Pulses}. We call it P to emphasize the connection with the pulses of \textbf{w}.
+
+\begin{figure}[H]
+  \centering
+    \includegraphics{images/03Pulses.pdf}
+  \caption{Example of a discrete wave \textbf{w} divided into pulses, of which the extrema are highlighted with a circle.
+P is the set of the points $ P_j = (i_j, \lvert w_j \lvert) $ representing the absolute value of those extrema.}
+  \label{fig:Pulses}
+\end{figure}
+
+The points in P are not fit for geometric interpretation yet, since the abscissa and ordinate of their orthogonal coordinate system have different units: in the vertical axis, we have a unit related to the instantaneous intensity of the wave, while in the horizontal axis we have the index at which each extremum occurs, ultimately a time-related unit.
+
+Specifically, this coordinate system has a basis vector $ \mathbf{b} = \{ \mathbf{b}_x, \mathbf{b}_y \}$ with $ \mathbf{b}_x = (a, 0), \mathbf{b}_y = (0, i) $ where $ a $ is an amplitude unit such as decibel, a measure of the instantaneous sound pressure, volt or even a fraction of a fixed maximum value.
+$ i $ represents the index of the sample, and is linked to time by the aforementioned relation $ i = t \ \text{fps} $.
+
+We are interested in finding a coordinate system where a pulse's amplitude $ \lvert w_j \lvert $ and length would define, on average, a square.
+
+To achieve that, we divide the basis vector of the original vertical axis by the average of all ordinates, effectively cancelling the unit of the vertical components of the points.
+We could divide the basis vector of the horizontal axis by the average of the difference between the abscissa of a point and the abscissa of the immediately posterior, likewise eliminating the axis' unit.
+
+The drawback is that we would lose the direct link between P and \textbf{w} that arises from the fact that the abscissas of the points in P are indices $ i $ of \textbf{w}: we choose instead to leave the abscissa intact and multiply the now unitless vertical basis axis by the average of the difference between the abscissa of a point and the abscissa of the immediately posterior.
+In this way, both axes will have unit $ i $, and the relationship is preserved, making it easier to recover the envelope points later.
+
+In this new coordinate system we have a basis vector $ \mathbf{b}^\prime = \{ \mathbf{b}^\prime_x, \mathbf{b}^\prime_y \}$ where $ \mathbf{b}^\prime_x = (i, 0) $ and $\mathbf{b}^\prime_y = (0, i) $, related to the original basis vector by equation \ref{eq:Basis}:
+
+\begin{subequations}  \label{eq:Basis}
+\begin{align}
+\mathbf{b}^\prime_x &= \mathbf{b}_x  \\
+\mathbf{b}^\prime_y &= \frac{\left( \frac{i_{m-1} - i_0}{m-1} \right)}{\left( \frac{\sum_{j=0}^{m-1} \lvert w_j \lvert}{m} \right)} \mathbf{b}_y 
+\end{align}
+\end{subequations}
+
+The effect of this normalization can be seen in the points shown in figure \ref{fig:Vectors}, shown in scale in both coordinate systems.
+The set V of the vectors from each point of P to the next is also shown in the picture, and it will be used to estimate the average curvature of P.
+We define this set in equation \ref{eq:Vectors} below.
+
+\begin{equation} \label{eq:Vectors}
+\text{V} = \{ v_0, v_1, v_{m-2} \} \quad \text{where} \quad v_k = P_{j+1} - P_j \quad \forall \quad 0 \le k \le m-2
+\end{equation}
+
+\begin{figure}[H]
+  \centering
+    \includegraphics{images/04Vectors.pdf}
+  \caption{The set of points P and the set V of vectors between two adjacent points of P, in both the original and normalized coordinate systems, shown in scale.}
+  \label{fig:Vectors}
+\end{figure}
+
+\subsubsection{Discrete Curvature Estimation}
+
+Translating the explanation of alpha shapes in \textcite{1994EdelsbrunnerThree} to our context, we can say that the points that are part of the envelope are those touched by a circle of a given radius $ r $, coming from $ y \to \infty $ towards the signal, that is not allowed to contain any point of the signal; Intuitively, one can picture a circle being rolled above the signal, and marking the points it touches as envelope points.
+
+To infer the appropriate radius $ r $ of such circle a measure of discrete curvature is needed.
+Discrete curvature estimation is an important task in image processing \parencite{2010FleischmannNovel} for which no default definition exists.
+
+The two possible general approaches are the derivation of direct methods, that use characteristics of the discrete wave to calculate the curvature, or the use of the curvature of a smooth, continuous curve fitted to the discrete wave \parencite{2001CoeurjollyDiscrete}.
+
+Concerning direct methods \textcite{2014CarrollSurvey}, for example, derive three such definitions based on the approximation of a circle by an inscribed, centered and circumscribed polygon.
+In the context of three-dimensional meshes, \textcite{2016VasaMultivariate} evaluate a range of existing estimators from a multivariate point of view.
+
+Those approaches define the discrete curvature in relation to the vertices of a wave (or mesh, in the three-dimensional case).
+In our particular case, it wouldn't make sense to talk about the curvature of a single pulse, since it is what our points ultimately represent.
+Instead, we are interested in the change of direction, what the term curvature ultimately means, between two adjacent pulses, as the vectors in the set V suggests.
+
+We thus proceed to define a discrete curvature measure over the edges of a discrete wave.
+To that end we are going to apply the definition of smooth curvature as the rate of change of the unit tangent to a curve, noting that this is equivalent to that of the osculating circle \parencite{2016VasaMesh}.
+
+\subsubsection{The Equivalent Circle Approach}
+
+The rationale is to find, for each vector $ v_k \in \text{V} $, the radius $ r_k $ of the equivalent circle whose tangent has the same change in direction, in the same horizontal distance, as the vector of interest, as shown in figure \ref{fig:DiscreteCurvature}.
+The average curvature of P will then be obtained by the average of all radii $ r_k $.
+
+From figure \ref{fig:DiscreteCurvature} is easy to see that $ r_k = v_{k,x} \sin(\theta_k)$ where $ v_{k,x} $ is the component of $ v_k $ in the horizontal direction; $ \theta_k $ itself can be obtained using the slope $ m_0 = 0 $ of a line in the horizontal direction and the slope $ m_k = v_{k,y} / v_{k,x} $ defined by $ v_k $ via the equality $ \theta_k = \arctan \big( (m_k - m_0) / (1 + m_k m_0) \big) $.
+
+\begin{figure}[H]
+  \centering
+    \includegraphics{images/05DiscreteCurvature.pdf}
+  \caption{The tangent unit vector of the circle changes from the horizontal direction in $ \hat{u}_0 $ to an inclination of $ \theta_k $ in $ \hat{u}_1 $, $ \theta_k $ being the angle that the vector $ v_k $ makes with the horizontal direction.}
+  \label{fig:DiscreteCurvature}
+\end{figure}
+
+The radius of the circle that represents the average curvature of V (and P), and will be used to obtain the envelope of the discrete wave \textbf{w} can be obtained via equation \ref{eq:Radius}:
+
+\begin{equation} \label{eq:Radius}
+r = \frac{\sum\limits_{k=0}^{m-2} \left( \frac{v_{k,x} \sqrt{v_{k,x}^2 + v_{k,y}^2}}{ v_{k,y}} \right)}{m-2}
+\end{equation}
+
+We need now to construct an algorithm, using the radius just obtained, to identify the points that belong to the envelope.
+Generally, the algorithm for the alpha shapes approach resorts to the construction of the Delaunay triangulation of the set of points, that is latter filtered to contain only the outer edges.
+We can adopt a more straightforward method, as our points have a defined structure:
+
+Let $ x_o, y_o \in \mathbb{R}, \quad 0 \le y_o < +\infty $ be the coordinates of the center of a circle with radius $ r $, that can be placed anywhere in the upper plane of the coordinate system;
+
+Let $ P_a, P_b, P_c \in \text{P}, \quad P_a \ne P_b \ne P_c $ be three different points of the set P;
+
+For all circles of center $ x_o, y_o $ and radius $ r $ that can be constructed to pass through any two different points of P, that is, $ \forall \quad P_b, P_c $ such that $ (P_{b,x} - x_o)^2 + (P_{b,y} - y_o)^2 = (P_{c,x} - x_o)^2 + (P_{c,y} - y_o)^2 = r^2 $, $ P_a $ will be part of the set $ E $ of the points belonging to the envelope if and only if none of those circles contain $ P_a $, or $ P_a \in \text{E} \leftrightarrow (P_{b,x} - x_o)^2 + (P_{b,y} - y_o)^2 > r^2 $.
+Summarizing this, we have equation \ref{eq:InEnvelope}:
+
+\begin{subequations} \label{eq:InEnvelope}
+\begin{align}
+& P_a \in \text{E} \leftrightarrow (P_{b,x} - x_o)^2 + (P_{b,y} - y_o)^2 > r^2, \\
+& \forall \ P_b, P_c \mid (P_{b,x} - x_o)^2 + (P_{b,y} - y_o)^2 = (P_{c,x} - x_o)^2 + (P_{c,y} - y_o)^2 = r^2, \\
+& P_a, P_b, P_c \in \text{P}, \quad P_a \ne P_b \ne P_c \quad \text{and} \quad x_o, y_o \in \mathbb{R}, \quad 0 \le y_o < +\infty
+\end{align}
+\end{subequations}
+
+The algorithm follows directly from the definition in \ref{eq:InEnvelope} after noting that the first and last members of P will always be part of the envelope, since they are reachable by a circle of any radius $ r $.
+From $ P_0 $, our first pivot point, we construct circles passing through $ P_1, P_2, \cdots, P_d $ until a circle that doesn't contain any point in P is found; $ P_d $ then becomes the new pivot point and is included in E, and this procedure continues until $ P_{m-1} $ is reached.
+
+The procedure is formalized in the algorithm \ref{FindEnvelope}.
+We don't provide similar algorithmic descriptions of the preceding steps since the Python implementation can easily serve as pseudocode.
+
+\begin{algorithm}[H]
+\caption{Retrieve Envelope} \label{FindEnvelope}
+\begin{algorithmic}
+\STATE \textbf{Given} $ \text{P} = \{P_0, P_1, \cdots, P_{m-1}\}, r $, \textbf{circle}, a function that returns the center of a circle of a given radius passing through two points and \textbf{distance}, a function that returns the Euclidean distance between two points,
+\STATE \textbf{Let} $ \text{E} = \{ P_0 \}, \text{id1} = 0, \text{id2} = 1 $
+\STATE \textbf{Define} $ P_c = \text{point}, \text{empty} = \text{boolean} $
+\WHILE {$ \text{id2} < m $}
+  \STATE $ P_c \leftarrow \textbf{circle}(r, P_\text{id1}, P_\text{id2}) $
+  \STATE empty $ \leftarrow $ \TRUE 
+  \FOR{$ i = \text{id2} + 1 $ \TO $ m $}
+    \IF {$ \textbf{distance}(P_c, P_i) < r $}
+      \STATE empty $ \leftarrow $ \FALSE
+      \STATE id2 $ \leftarrow \text{id2} + 1 $
+      \STATE \textbf{break}
+    \ENDIF
+  \ENDFOR
+  \IF {empty}
+    \STATE E $ \leftarrow $ E $ \cup \ \{ P_\text{id2} \} $
+    \STATE id1 $ \leftarrow $ id2
+    \STATE id2 $ \leftarrow \text{id2} + 1 $
+  \ENDIF
+\ENDWHILE
+\end{algorithmic}
+\end{algorithm}
+
+The set E obtained via algorithm \ref{FindEnvelope} is a set of points.
+The most straightforward way to transform those points into the vector $ \textbf{e} \in \mathbb{R}^n $ is via linear interpolation, the approach used in this work.
+It is worth noting, however, that many other procedures are available, both for the interpolation and the smooth approximation of points.
+
+\subsection{Results and Discussion}
+Given the lack of a precise definition of what an envelope should be, we begin this section suggesting an empirical metric, based on the behaviour of the carrier wave \textbf{c}.
+We also comment on some metrics presented in the literature, and compare our algorithm with traditional demodulation approaches, both in relation to execution time and accuracy; to this last end, a numerical indicator is also introduced.
+
+We also illustrate how the method here presented can be used to isolate not only a single envelope of the wave, but the superior and inferior envelopes as well, that we call frontiers, ending the section with a suggestion on how the proposed algorithm can be useful beyond envelope detection, by identifying the approximate locations of the pseudo-cycles in a quasi-periodic wave.
+
+Due to an assortment of factors, such as the unconventional approach presented in this work and the intrinsic difficulty in visualizing discrete waves given their high dimensionality, the authors deemed helpful to provide a companion website with interactive visualizations \parencite{2020TarjanoEnvelope}.
+
+An implementation of the method in pure Python, fully functional, can also be found at the repository dedicated to this work as well as a Python module available at the Python Package Index (PyPI) and conveniently installable via the command `pip install signal-envelope`.
+Besides, those working under Windows 64-bit can benefit from a specialized dynamic library coded in C++ that is automatically used within this module.
+
+\subsubsection{Quality of an envelope}
+
+There are no agreed-upon objective metrics to assess the quality of an envelope.
+Recalling figure \ref{fig:Envelope}, however, and the accompanying discussion, an intuitive way is to consider the carrier wave \textbf{c}, and how well it seems to be bounded by the interval $ \{-1, 1\} $.
+
+To illustrate how the algorithm performs on such metric, we provide the following figures of original discrete waves \textbf{w}, their extracted envelope \textbf{e}, and the inferred carrier, obtained after dividing, element-wise, the original wave by its envelope.
+
+From figure \ref{fig:spoken_voice} it can be seen that the algorithm performs well under the carrier metric.
+This is especially remarkable for the case of spoken voice, a notorious complex wave with fast changes in pitch and high noise content.
+
+These characteristics are in stark contrast with those of the sang voice in figure \ref{fig:singer_voice}, where a recording of an alto singer is shown; there, the wave is almost periodic, with a stable waveform.
+
+\begin{figure}[H]
+  \centering
+    \includegraphics{images/06Graph_bend.pdf}
+  \caption{Signal, envelope and carrier for a record of a bend performed on an electric guitar.}
+  \label{fig:Bend}
+\end{figure}
+
+\begin{figure}[H]
+  \centering
+    \includegraphics{images/06Graph_spoken_voice.pdf}
+  \caption{Signal, envelope and carrier for a record of male voice uttering the word “amazing”.}
+  \label{fig:spoken_voice}
+\end{figure}
+
+\begin{figure}[H]
+  \centering
+    \includegraphics{images/06Graph_alto.pdf}
+  \caption{Signal, envelope and carrier for a record of an alto singer.}
+  \label{fig:singer_voice}
+\end{figure}
+
+In an effort to formalize the assessment of the quality of temporal envelopes, \textcite{1996Loughlinamplitude} proposed some conditions necessary, but not sufficient, to ensure the physical plausibility of an envelope.
+We comment briefly below that the algorithm here presented satisfies the four conditions presented.
+
+The first one states that, if a signal is bounded in magnitude, then its envelope should be, as well.
+Our algorithm complies with this requisite, as the envelope is composed of selected samples from the original signal itself.
+
+The second statement is that if the signal has a finite frequency range, that frequency range must not be exceeded by the envelope.
+This is only directly applicable in the continuous case, where this condition was proposed.
+
+Nevertheless, if we picture both the wave and the envelope as samples from continuous functions, is easy to see that this is the case for our algorithm; since the envelope is composed by some samples of the digital wave, in the limit where all the samples of the wave are also part of the envelope, we have that the maximum frequency for both is the same.
+
+In all other cases, the maximum frequency of the envelope will be smaller than that of the original wave.
+The range of frequencies of the envelope, then, will always lie between zero and the maximum frequency of the original wave.
+
+The third condition states what, for a periodic signal, the envelope should be a straight line with the intercept equal to the amplitude of this signal.
+Since all the local maxima have the same amplitude in the case of a periodic wave, our envelope will indeed be a straight line, with the same amplitude as the original wave.
+An example of such behaviour can be observed in figure \ref{fig:Comparison_sinusoid}.
+
+The fourth and last condition states that if a wave is multiplied by a constant, the envelope should also be multiplied by the same constant.
+That is also the case, again because the points that form the envelope are points of the original wave itself.
+
+\subsubsection{Comparison with traditional algorithms}
+
+Direct comparison with many of the more recent algorithms is made difficult by the unavailability of digital implementations of such works, many designed to process analog signals \parencite[e.g.,][]{2018AssefModeling}.
+
+Nevertheless, insight can be gained from a comparison of the results of the method here proposed with some of the most common envelope extraction algorithms.
+Specifically, we compare the present method with the following approaches:
+
+\begin{enumerate}
+\item Smoothing: This is a relatively simple procedure that, applied to the absolute values of a discrete signal, can provide an approximation of the variation of its amplitude in time.
+We use the Savitzky-Golay algorithm \parencite{1964SavitzkySmoothing} with a window width of 3001 samples and a cubic polynomial.
+\item Filtering: The absolute value of the wave is filtered using a digital implementation of the Butterworth \parencite{butterworth1930theory} low pass filter.
+We use an order 2 filter with a cut off frequency of 10 Hz.
+\item Hilbert Transform: We apply the Hilbert transform to the absolute values of the wave after filtering the original signal with a Butterworth filter with a cut off frequency of 100 Hz, using the absolute value of the analytic signal to obtain the envelope.
+\end{enumerate}
+
+In \ref{fig:Comparison_sinusoid} we illustrate the envelope extracted by the conventional algorithms, contrasting it with the result of the method presented in this work, for a simple sinusoid.
+Note that none of the conventional algorithms complies with the third condition proposed in \textcite{1996Loughlinamplitude} that the envelope of a periodic wave should be a straight line.
+
+\begin{figure}[H]
+  \centering
+    \includegraphics{images/07Comparison_sinusoid.pdf}
+  \caption{Comparison of envelope detection algorithms for a simple sinusoid}
+  \label{fig:Comparison_sinusoid}
+\end{figure}
+
+For a recording of the key 33 of a grand piano, where parts of the sound are highly percussive and noisy, the Hilbert approach presents considerable oscillations, as can be seen in figure \ref{fig:Comparison_piano}.
+
+\begin{figure}[H]
+  \centering
+    \includegraphics{images/07Comparison_piano.pdf}
+  \caption{Comparison of envelope detection algorithms for the recording of the key 33 of a grand piano}
+  \label{fig:Comparison_piano}
+\end{figure}
+
+In figure \ref{fig:Comparison_soprano} all envelopes exhibit a similar movement along the samples, but the ones generated by traditional algorithms are undershooting the original wave.
+
+\begin{figure}[H]
+  \centering
+    \includegraphics{images/07Comparison_soprano.pdf}
+  \caption{Comparison of envelope detection algorithms for the recording of the vocalization of a soprano singer}
+  \label{fig:Comparison_soprano}
+\end{figure}
+
+All the traditional methods demanded careful choice of parameters.
+In the case of the Hilbert transform, pre-processing in the form of filtering was applied.
+That is in contrast with our method, that organically tunes itself via the automatic choice of the radius of the circle representing the average discrete curvature of the signal.
+
+Table \ref{table:Error_comparison} presents a comparison of the average errors of the techniques.
+The average error was calculated by the average of the squared distance of the points in the envelope divided by two and the absolute value of the wave.
+The rationale is that the values of the envelope divided by two should pass near the average of the absolute values of the wave.
+This metric is only valid for reasonably smooth envelopes, however.
+
+We can see from the table \ref{table:Error_comparison} that the method presented here consistently outperforms the traditional approaches for all waves tested, yielding an error 33\% smaller, on average, than the other methods.
+It is also worth noting that the errors of the traditional methods are very similar between them.
+
+\begin{table}[H]
+\centering
+\begin{tabular}{lcccc}
+\hline
+\multicolumn{5}{c}{\textbf{Average error per frame}} \\
+              & Present Work & Smoothing & Low pass & Hilbert \\ 
+\hline
+alto          & 0.073        & 0.083     & 0.083   & 0.082   \\
+bend          & 0.009        & 0.015     & 0.015   & 0.014   \\
+piano         & 0.005        & 0.006     & 0.006   & 0.005   \\
+brass         & 0.024        & 0.035     & 0.035   & 0.033   \\
+soprano       & 0.042        & 0.065     & 0.065   & 0.065   \\
+spoken voice  & 0.024        & 0.037     & 0.037   & 0.034   \\
+tom           & 0.003        & 0.005     & 0.005   & 0.005   \\
+sinusoid      & 0.113        & 0.193     & 0.195   & 0.191   \\ 
+\hline
+Average       & 0.037        & 0.055     & 0.055   & 0.054  
+\end{tabular}
+\caption{Comparison of the average errors of the algorithm presented in this work with the most common methods of digital envelope identification.
+The discrete waves used were normalized between -1 and 1.
+The average is between the absolute value of those waves and the envelope divided by two.}
+\label{table:Error_comparison}
+\end{table}
+
+The times taken for the algorithms compared here to process each wave are shown in \ref{table:Time_comparison}.
+We used specialized implementations available in libraries of the Python programming language.
+The presented algorithm is in average the second-fastest.
+
+\begin{table}[H]
+\centering
+\begin{tabular}{lcccc}
+\hline
+\multicolumn{5}{c}{\textbf{Time in seconds}} \\
+              & Present Work & Smoothing & Low pass & Hilbert \\ 
+\hline
+alto          & 0.008        & 0.101     & 0.007   & 0.011   \\
+bend          & 0.030        & 0.226     & 0.012   & 0.131   \\
+piano         & 0.034        & 0.374     & 0.014   & 0.042   \\
+brass         & 0.032        & 0.297     & 0.011   & 0.045   \\
+soprano       & 0.177        & 0.735     & 0.027   & 0.566   \\
+spoken voice  & 0.014        & 0.105     & 0.003   & 0.038   \\
+tom           & 0.007        & 0.084     & 0.018   & 0.026   \\
+sinusoid      & 0.007        & 0.091     & 0.005   & 0.011   \\ 
+\hline
+Average       & 0.039        & 0.252     & 0.012   & 0.109  
+\end{tabular}
+\caption{Comparison of the processing time of the algorithm presented in this work with the most common methods of digital envelope identification.}
+\label{table:Time_comparison}
+\end{table}
+
+\subsubsection{Frontiers}
+
+In practice, is not uncommon for a discrete wave, especially in the case of sound, to present somewhat different positive and negative contours; in those cases, the algorithm here presented can be used independently in the positive and negative pulses of the wave to define two “envelopes”, that we are going to call superior and inferior frontiers.
+
+Figure \ref{fig:FullFrontiers} illustrates the frontiers of six diverse discrete sound waves, as well as an in-detail view of the highlighted segment for each signal.
+All waves are records of physical sounds, chosen to represent the applicability of the algorithm in real-world scenarios.
+
+\begin{figure}[H]
+  \centering
+    \includegraphics{images/08FullFrontiers.pdf}
+  \caption{Positive and negative frontiers of six digital waves, as extracted by the algorithm here presented.
+For each wave, the region highlighted in black is shown in detail besides the whole wave.
+For each wave, the horizontal axis is the sample number $ i $, while the vertical axis is the normalized amplitude.
+All 6 waves were recorded at 44100 fps.}
+  \label{fig:FullFrontiers}
+\end{figure}
+
+The frontiers are satisfactorily detected in the case of harmonic and inharmonic sounds, and are robust in relation to the number of samples and the frequencies of the waves.
+That is a peculiarity of our algorithm: the traditional algorithms are unable to generate two distinct envelopes for the same wave.
+
+\subsection{Approximated location of pseudo-cycles}
+
+The envelope is shown to add complexity to the spectral representation of a wave \parencite{2019TarjanoNeuro}, and an accurate description of the envelope, while describing the evolution of the instantaneous amplitude of a signal in time, would also simplify further spectral analysis.
+Figure \ref{fig:Fourier} shows the frequency-domain power spectrum for the wave and the carrier presented in \ref{fig:Envelope}.
+For the carrier, the power spectrum presents two nonzero values, while the power spectrum for the wave composed of the carrier modulated by the envelope is more complex.
+
+\begin{figure}[H]
+  \centering
+    \includegraphics{images/11Fourier.pdf}
+  \caption{Fourier power spectrum for the wave and carrier shown in \ref{fig:Envelope}.}
+  \label{fig:Fourier}
+\end{figure}
+
+Moreover, the algorithm developed in this work naturally divides a signal into its pseudo-cycles, pinpointing them in the time-domain, thus providing the building blocks for the reconstruction of the fine structure of the wave.
+
+\begin{figure}[H]
+  \centering
+    \includegraphics{images/09PCs.pdf}
+  \caption{Section of the lines formed by the positions of the positive and negative frontiers of the digital wave illustrated in \ref{fig:singer_voice}, as extracted by the algorithm presented, and their average.}
+  \label{fig:PCs}
+\end{figure}
+
+One can see in figure \ref{fig:PCs} a section of the plot of where the local maxima and minima occurs in the recording of an alto singer show in \ref{fig:singer_voice}.
+The average between the two is also shown.
+With this information, one can split a discrete wave into its pseudo-cycles, as is illustrated in Figure \ref{fig:AvgPc}: It shows the superposition of this extracted pseudo-cycles, after being normalized by length, and their average.
+This same average is shown shifted, to exemplify that approximate $C^0$ and $C^1$ continuity is achieved.
+
+\begin{figure}[H]
+  \centering
+    \includegraphics{images/10AvgPc.pdf}
+  \caption{Pseudo-cycles of the wave illustrated in \ref{fig:singer_voice}, segmented and superposed, and their average.
+The average is also shown shifted by one average period, to illustrate the average continuity.}
+  \label{fig:AvgPc}
+\end{figure}
+
+\subsection{Conclusion}
+This work fills a gap identified in the theory of digital signal processing, where the lack of general procedures for the accurate identification of the temporal envelope of arbitrary waves poses an obstacle to the complete description and eventual manipulation of signals.
+
+By presenting a general approach for envelope detection, and eliminating the need for parameter tuning, we expect to facilitate further research at the many areas that rely on envelope detection techniques, and improve already existing algorithm that have envelope detection as an intermediary step.
+
+Illustrating the efficiency of an algorithm inspired by geometric aspects of a discrete signal on a wide range of real-world signals, we also hope to encourage research in this direction.
+
+The present work also highlights some gaps in our current understanding and definition of temporal envelope and, in so doing, endeavours to motivate further research in this topic, that can ultimately give rise to a more mathematically sound envelope theory, especially concerning broadband waves; suggestions in this direction are also presented.
+
+While relevant in its own accord, the procedure here presented approximately isolates the individual pseudo-cycles of a wave, pinpointing them in the time-domain.
+
+Similarly, by decomposing a wave into its envelope and carrier, and simplifying the representation of such carrier in the frequency-domain, a rich set of advancements, especially in the areas of sound compression and synthesis in the time and frequency-domains, can be built upon this initial algorithm.
+
+By illustrating how this information can be used to obtain the average waveform of a discrete wave, we hope to encourage research in this area, ultimately leading to alternative representations of discrete waves.
 

### 2021-03-14 images/01AnalyticSignal.py

@@ -35,39 +35,42 @@ analytic_signal = hilbert(W)
 hilbert_envelope = np.abs(analytic_signal)
 
 
-'''Plotting'''
+'''============================================================================'''
+'''                              PLOT LINES                                    '''
+'''============================================================================'''
 FONT = dict(
     family="Latin Modern Roman",
     color="black",
     size=13.3333
   )
 
+  
+
+'''Plotting'''
 fig = go.Figure()
 fig.layout.template ="plotly_white" 
 fig.update_layout(
   xaxis_title="<b><i>i</i></b>",
   yaxis_title="<b>Amplitude</b>",
-  paper_bgcolor='rgba(0,0,0,0)',
-  plot_bgcolor='rgba(0,0,0,0)',
-  legend=dict(orientation='h', yanchor='top', y=1.1),
-  margin=dict(l=0, r=0, b=0, t=0, pad=0),
+  legend=dict(orientation='h', yanchor='top', xanchor='left', y=1.1),
+  margin=dict(l=0, r=0, b=0, t=0),
   font=FONT,
-  # titlefont=FONT
+  titlefont=FONT
 )
-# fig.layout.xaxis.title.font=FONT
-# fig.layout.yaxis.title.font=FONT
-fig.update_xaxes(title_font = FONT, showline=False, showgrid=False, zeroline=False)
-fig.update_yaxes(title_font = FONT, showline=False, showgrid=False, zerolinewidth=2, zerolinecolor='gray')
+fig.layout.xaxis.title.font=FONT
+fig.layout.yaxis.title.font=FONT
+fig.update_xaxes(showline=False, showgrid=False, zeroline=False)
+fig.update_yaxes(showline=False, showgrid=False, zerolinewidth=2, zerolinecolor="silver")
 
 '''Plotting Signal'''
 fig.add_trace(
   go.Scatter(
-    name="Signal",
+    name="Signal      ",
     # x=X,
     y=W,
     mode="lines",
     line=dict(
-        width=1,
+        width=.8,
         color="gray",
     )
   )
@@ -75,12 +78,12 @@ fig.add_trace(
 
 fig.add_trace(
   go.Scatter(
-    name="Hilbert Envelope",
+    name="Hilbert Envelope      ",
     # x=X,
     y=hilbert_envelope,
     mode="lines",
     line=dict(
-        width=1,
+        width=.8,
         color="black",
     )
   )
[…]

### 2021-03-14 images/01Pulses.py

@@ -1,119 +1,0 @@
-import sys
-# sys.path.append("c:/Users/tesse/Desktop/Files/Dropbox/0_Thesis/Python")
-import numpy as np
-import plotly.graph_objects as go
-from Helper import signal_to_pulses, get_pulses_area
-
-'''Generating Wave'''
-
-np.random.seed(1)
-# fps = 10
-n = 100+1
-
-X = np.arange(n)# / fps
-W = np.zeros(n)
-
-afp = [
-  [2, 3, np.pi],
-  # [1, 2, 1.2 * np.pi],
-  # [1, 2.3, 1.4 * np.pi]
-  ]
-for a, f, p in afp:
-  W += a * np.cos(p + 2 * np.pi * f * X / n) + np.random.normal(0, a / 5, n)
-
-W = W / np.max(np.abs(W))
-
-'''Wave to pulses'''
-pulses = signal_to_pulses(W)
-pulses_X, pulses_Y = get_pulses_area(pulses)
-
-
-FONT = dict(
-    family="Latin Modern Roman",
-    color="black",
-    size=13.3333
-  )
-
-'''Plotting'''
-fig = go.Figure()
-fig.layout.template ="plotly_white" 
-fig.update_layout(
-  xaxis_title="<b><i>i</i></b>",
-  yaxis_title="<b>Amplitude</b>",
-  legend=dict(orientation='h', yanchor='top', xanchor='left', y=1.1),
-  margin=dict(l=0, r=0, b=0, t=0),
-  font=FONT,
-  titlefont=FONT
-)
-fig.layout.xaxis.title.font=FONT
-fig.layout.yaxis.title.font=FONT
-fig.update_xaxes(showline=False, showgrid=False, zeroline=False)
-fig.update_yaxes(showline=False, showgrid=False, zerolinewidth=2, zerolinecolor='gray')
-
-
-'''Samples'''
-fig.add_trace(
-  go.Scatter(
-    name= "Samples Stem",
-    showlegend=False,
-    x=[i for x in X for i in (x, x, None)],
-    y=[i for y in W for i in (0, y, None)],
-    mode='lines',
-    line=go.scatter.Line(color="black", width=1)
-  )
-)
-fig.add_trace(
-  go.Scatter(
-    name="Samples",
-    x=X,
-    y=W,
-    mode='markers',
-    marker=dict(
-        size=4,
-        color="black",
-    )
-  )
-)
-
-''' Pulses '''
-fig.add_trace(
[…]

### 2021-03-14 images/02Envelope.py

@@ -1,0 +1,117 @@
+import sys
+print(sys.version)
+import plotly.graph_objects as go
+sys.path.append("c:/Users/tesse/Desktop/Files/Dropbox/0_Thesis/Python/01_Envelope")
+import numpy as np
+from scipy.interpolate import interp1d
+from Helper import signal_to_pulses, get_pulses_area
+
+'''Generating Wave'''
+
+np.random.seed(0)
+# fps = 10
+n = 100
+
+X = np.arange(n)# / fps
+C = np.zeros(n)
+
+f = interp1d([0, 0.25 * n, 0.5 * n, 0.75 * n, n], np.abs(1 + np.random.normal(0, .3, 5)), "cubic")
+E = f(X)
+
+afp = [
+  [1, 5, 1.4 * np.pi],
+  [1.5, 10, 1.4 * np.pi],
+  # [1, 2 * 6, 1.4 * np.pi]
+  ]
+for a, f, p in afp:
+  C += a * np.cos(p + 2 * np.pi * f * X / n) #+ np.random.normal(0, a / 100, n)
+
+C = -C / np.max(np.abs(C))
+
+W = E * C
+
+'''Wave to pulses'''
+pulses = signal_to_pulses(C)
+pulses_X, pulses_Y = get_pulses_area(pulses)
+
+'''============================================================================'''
+'''                              PLOT LINES                                    '''
+'''============================================================================'''
+FONT = dict(
+    family="Latin Modern Roman",
+    color="black",
+    size=13.3333
+  )
+
+'''Plotting'''
+fig = go.Figure()
+fig.layout.template ="plotly_white" 
+fig.update_layout(
+  xaxis_title="<b><i>i</i></b>",
+  yaxis_title="<b>Amplitude</b>",
+  legend=dict(orientation='h', yanchor='top', xanchor='left', y=1.1),
+  margin=dict(l=0, r=0, b=0, t=0),
+  font=FONT,
+  titlefont=FONT
+)
+fig.layout.xaxis.title.font=FONT
+fig.layout.yaxis.title.font=FONT
+
+fig.update_xaxes(showline=False, showgrid=False, zeroline=False)
+fig.update_yaxes(showline=False, zeroline=False, showgrid=True, gridwidth=1, gridcolor='silver', tickvals=[-1, 0, 1])
+
+fig.add_trace(
+  go.Scatter(
+    name="Envelope <b>e</b>      ",
+    x=X,#pulses_X,
[…]

### 2021-03-14 images/06AvgCircle.py

@@ -1,246 +1,0 @@
-import sys
-sys.path.append("c:/Users/tesse/Desktop/Files/Dropbox/0_Thesis/Python")
-import os
-from pathlib import Path
-import plotly.graph_objects as go
-import numpy as np
-from scipy.interpolate.interpolate import interp1d
-from Helper import read_wav, get_pulses_area, split_pulses, signal_to_pulses, get_frontier, draw_circle, get_curvature_function
-from scipy.signal import savgol_filter
-
-# os.path.dirname(os.path.dirname(path))
-
-'''==============='''
-''' Read wav file '''
-'''==============='''
-fig = go.Figure()
-
-parent_path = str(Path(os.path.abspath('./')).parents[0])
-print(parent_path)
-name = "piano33"
-W, fps = read_wav(f"{parent_path}/Python/Samples/{name}.wav")
-
-W = W[100103 : 100103 + 585] #* 1000 / np.max(np.abs(W[100103 : 100775]))
-
-W = W - np.average(W)
-# amplitude = np.max(np.abs(W))
-W = W * .5
-n = W.size
-print(n)
-X = np.arange(n)
-
-pulses = signal_to_pulses(W)
-
-areas_X, areas_Y = get_pulses_area(pulses)
-
-pos_pulses, neg_pulses, pos_noises, neg_noises = split_pulses(pulses)
-
-pos_X, neg_X = np.array([p.x for p in pos_pulses]), np.array([p.x for p in neg_pulses])
-pos_Y, neg_Y = np.array([p.y for p in pos_pulses]), np.array([p.y for p in neg_pulses])
-pos_L, neg_L = np.array([p.len for p in pos_pulses]) , np.array([p.len for p in neg_pulses])
-
-'''pos frontier'''
-scaling = np.average(pos_L) / np.average(pos_Y)
-
-for p in pos_pulses:
-  p.y = p.y * scaling
-
-pos_frontier_X, pos_frontier_Y = get_frontier(pos_pulses)
-
-pos_frontier_X = np.array(pos_frontier_X)
-pos_frontier_Y = np.array(pos_frontier_Y)
-
-pos_frontier_Y = pos_frontier_Y / scaling
-
-'''neg frontier'''
-scaling = np.average(neg_L) / np.average(neg_Y)
-
-for p in neg_pulses:
[…]

### 2021-03-14 images/08Fourier.py

@@ -1,266 +1,0 @@
-import plotly.graph_objects as go
-import numpy as np
-from scipy.interpolate import interp1d
-import numpy.polynomial.polynomial as poly
-import wave
-from scipy.signal import savgol_filter
-import os
-from pathlib import Path
-import sys
-# import argparse
-from plotly.subplots import make_subplots
-
-def read_wav(path): 
-  """returns signal & fps"""
-  wav = wave.open(path , 'r')
-  signal = np.frombuffer(wav.readframes(-1) , np.int16)
-  fps = wav.getframerate()
-  return signal, fps
-
-
-def get_circle(x0, y0, x1, y1, r):
-  '''returns center of circle that passes through two points'''
-  
-  radsq = r * r
-  q = np.sqrt((x1 - x0)**2 + (y1 - y0)**2)
-
-  # assert q <= 2 * r, f"shit"
-
-  x3 = (x0 + x1) / 2
-  y3 = (y0 + y1) / 2
-
-  if y0 + y1 >= 0:
-    xc = x3 + np.sqrt(radsq - (q / 2)**2) * ((y0 - y1) / q)
-    yc = y3 + np.sqrt(radsq - (q / 2)**2) * ((x1 - x0) / q)
-  else:
-    xc = x3 - np.sqrt(radsq - (q / 2)**2) * ((y0 - y1) / q)
-    yc = y3 - np.sqrt(radsq - (q / 2)**2) * ((x1 - x0) / q)
-
-  return xc, yc
-
-
-def get_radius_function(X, Y):
-  m0 = np.average(Y[1:] - Y[:-1]) / np.average(X[1:] - X[:-1])
-  curvatures_X = []
-  curvatures_Y = []
-  for i in range(len(X) - 1):
-    m1 = (Y[i + 1] - Y[i]) / (X[i + 1] - X[i])
-
-    theta = np.arctan( (m1 - m0) / (1 + m1 * m0) )
-    k = np.sin(theta) / (X[i + 1] - X[i])
-
-    # if k >= 0:
-    curvatures_X.append((X[i + 1] + X[i]) / 2)
-    curvatures_Y.append(k)
-    
-  curvatures_Y = np.array(curvatures_Y)
-  coefs = poly.polyfit(curvatures_X, curvatures_Y, 0)
-  smooth_curvatures_Y = poly.polyval(curvatures_X, coefs)
-
[…]

### 2021-03-14 images/09PCs.py

@@ -1,0 +1,96 @@
+'''To be used AFTER signal_envelope is installed via pip'''
+import sys
+# sys.path.append("signal_envelope/")
+import numpy as np
+import signal_envelope as se
+from scipy import interpolate
+import plotly.graph_objects as go
+
+name = "alto"
+W, _ = se.read_wav(f"C:/Users/tesse/Desktop/Files/Dropbox/0_Science/reps/envelope/test_samples/{name}.wav")
+amp = np.max(np.abs(W))
+W = 4 * W / amp
+X = np.arange(W.size)
+
+Xpos, Xneg = se.get_frontiers(W, 0)
+E = se.get_frontiers(W, 1)
+
+m = min(Xpos.size, Xneg.size)
+Xavg = (Xpos[0 : m] + Xneg[0 : m]) / 2
+
+'''============================================================================'''
+'''                                    PLOT                                    '''
+'''============================================================================'''
+
+FONT = dict(
+    family="Latin Modern Roman",
+    color="black",
+    size=13.3333
+  )
+
+'''Plotting'''
+fig = go.Figure()
+fig.layout.template ="plotly_white" 
+fig.update_layout(
+  xaxis_title="",
+  yaxis_title="<b><i>i</i></b>",
+  legend=dict(orientation='h', yanchor='top', xanchor='left', y=1.1),
+  margin=dict(l=0, r=0, b=0, t=0),
+  font=FONT,
+  titlefont=FONT
+)
+fig.layout.xaxis.title.font=FONT
+fig.layout.yaxis.title.font=FONT
+
+fig.update_xaxes(showline=False, showgrid=False, zeroline=False, range=[288-.5, 293+.5])
+fig.update_yaxes(showline=False, showgrid=False, zeroline=False, range=[23500, 24030])
+
+
+fig.add_trace(
+  go.Scatter(
+    name="Positive Frontier      ",
+    # x=np.arange(W.size),
+    y=Xpos,
+    mode="lines+markers",
+    # fill="tozeroy",
+    # fillcolor="rgba(0,0,0,0.6)",
+    line=dict(width=1, color="blue",),
+  )
+)
+
+fig.add_trace(
+  go.Scatter(
+    name="Negative Frontier      ",
+    # x=np.arange(W.size),
+    y=Xneg,
+    mode="lines+markers",
+    # fill="tozeroy",
+    # fillcolor="rgba(0,0,0,0.3)",
[…]

### 2021-03-14 images/09yRobustness.py

@@ -1,256 +1,0 @@
-import plotly.graph_objects as go
-import numpy as np
-from scipy.interpolate import interp1d
-import numpy.polynomial.polynomial as poly
-import wave
-from scipy.signal import savgol_filter
-import os
-from pathlib import Path
-import sys
-# import argparse
-from plotly.subplots import make_subplots
-
-def read_wav(path): 
-  """returns signal & fps"""
-  wav = wave.open(path , 'r')
-  signal = np.frombuffer(wav.readframes(-1) , np.int16)
-  fps = wav.getframerate()
-  return signal, fps
-
-
-def get_circle(x0, y0, x1, y1, r):
-  '''returns center of circle that passes through two points'''
-  
-  radsq = r * r
-  q = np.sqrt((x1 - x0)**2 + (y1 - y0)**2)
-
-  # assert q <= 2 * r, f"shit"
-
-  x3 = (x0 + x1) / 2
-  y3 = (y0 + y1) / 2
-
-  if y0 + y1 >= 0:
-    xc = x3 + np.sqrt(radsq - (q / 2)**2) * ((y0 - y1) / q)
-    yc = y3 + np.sqrt(radsq - (q / 2)**2) * ((x1 - x0) / q)
-  else:
-    xc = x3 - np.sqrt(radsq - (q / 2)**2) * ((y0 - y1) / q)
-    yc = y3 - np.sqrt(radsq - (q / 2)**2) * ((x1 - x0) / q)
-
-  return xc, yc
-
-
-def get_radius_function(X, Y):
-  m0 = np.average(Y[1:] - Y[:-1]) / np.average(X[1:] - X[:-1])
-  curvatures_X = []
-  curvatures_Y = []
-  for i in range(len(X) - 1):
-    m1 = (Y[i + 1] - Y[i]) / (X[i + 1] - X[i])
-
-    theta = np.arctan( (m1 - m0) / (1 + m1 * m0) )
-    k = np.sin(theta) / (X[i + 1] - X[i])
-
-    # if k >= 0:
-    curvatures_X.append((X[i + 1] + X[i]) / 2)
-    curvatures_Y.append(k)
-    
-  curvatures_Y = np.array(curvatures_Y)
-  coefs = poly.polyfit(curvatures_X, curvatures_Y, 0)
-  smooth_curvatures_Y = poly.polyval(curvatures_X, coefs)
-
[…]

### 2021-03-14 images/10xRobustness.py

@@ -1,258 +1,0 @@
-import plotly.graph_objects as go
-import numpy as np
-from scipy.interpolate import interp1d
-import numpy.polynomial.polynomial as poly
-import wave
-from scipy.signal import savgol_filter
-import os
-from pathlib import Path
-import sys
-# import argparse
-from plotly.subplots import make_subplots
-
-def read_wav(path): 
-  """returns signal & fps"""
-  wav = wave.open(path , 'r')
-  signal = np.frombuffer(wav.readframes(-1) , np.int16)
-  fps = wav.getframerate()
-  return signal, fps
-
-
-def get_circle(x0, y0, x1, y1, r):
-  '''returns center of circle that passes through two points'''
-  
-  radsq = r * r
-  q = np.sqrt((x1 - x0)**2 + (y1 - y0)**2)
-
-  # assert q <= 2 * r, f"shit"
-
-  x3 = (x0 + x1) / 2
-  y3 = (y0 + y1) / 2
-
-  if y0 + y1 >= 0:
-    xc = x3 + np.sqrt(radsq - (q / 2)**2) * ((y0 - y1) / q)
-    yc = y3 + np.sqrt(radsq - (q / 2)**2) * ((x1 - x0) / q)
-  else:
-    xc = x3 - np.sqrt(radsq - (q / 2)**2) * ((y0 - y1) / q)
-    yc = y3 - np.sqrt(radsq - (q / 2)**2) * ((x1 - x0) / q)
-
-  return xc, yc
-
-
-def get_radius_function(X, Y):
-  m0 = np.average(Y[1:] - Y[:-1]) / np.average(X[1:] - X[:-1])
-  curvatures_X = []
-  curvatures_Y = []
-  for i in range(len(X) - 1):
-    m1 = (Y[i + 1] - Y[i]) / (X[i + 1] - X[i])
-
-    theta = np.arctan( (m1 - m0) / (1 + m1 * m0) )
-    k = np.sin(theta) / (X[i + 1] - X[i])
-
-    # if k >= 0:
-    curvatures_X.append((X[i + 1] + X[i]) / 2)
-    curvatures_Y.append(k)
-    
-  curvatures_Y = np.array(curvatures_Y)
-  coefs = poly.polyfit(curvatures_X, curvatures_Y, 0)
-  smooth_curvatures_Y = poly.polyval(curvatures_X, coefs)
-
[…]

### 2021-03-14 images/10xRobustnessTest.py

@@ -1,267 +1,0 @@
-import plotly.graph_objects as go
-import numpy as np
-from scipy.interpolate import interp1d
-import numpy.polynomial.polynomial as poly
-import wave
-from scipy.signal import savgol_filter
-import os
-from pathlib import Path
-import sys
-import argparse
-from plotly.subplots import make_subplots
-
-def read_wav(path): 
-  """returns signal & fps"""
-  wav = wave.open(path , 'r')
-  signal = np.frombuffer(wav.readframes(-1) , np.int16)
-  fps = wav.getframerate()
-  return signal, fps
-
-
-def get_circle(x0, y0, x1, y1, r):
-  '''returns center of circle that passes through two points'''
-  
-  radsq = r * r
-  q = np.sqrt((x1 - x0)**2 + (y1 - y0)**2)
-
-  # assert q <= 2 * r, f"shit"
-
-  x3 = (x0 + x1) / 2
-  y3 = (y0 + y1) / 2
-
-  if y0 + y1 >= 0:
-    xc = x3 + np.sqrt(radsq - (q / 2)**2) * ((y0 - y1) / q)
-    yc = y3 + np.sqrt(radsq - (q / 2)**2) * ((x1 - x0) / q)
-  else:
-    xc = x3 - np.sqrt(radsq - (q / 2)**2) * ((y0 - y1) / q)
-    yc = y3 - np.sqrt(radsq - (q / 2)**2) * ((x1 - x0) / q)
-
-  return xc, yc
-
-
-def get_radius_function(X, Y):
-  m0 = np.average(Y[1:] - Y[:-1]) / np.average(X[1:] - X[:-1])
-  curvatures_X = []
-  curvatures_Y = []
-  for i in range(len(X) - 1):
-    m1 = (Y[i + 1] - Y[i]) / (X[i + 1] - X[i])
-
-    theta = np.arctan( (m1 - m0) / (1 + m1 * m0) )
-    k = np.sin(theta) / (X[i + 1] - X[i])
-
-    # if k >= 0:
-    curvatures_X.append((X[i + 1] + X[i]) / 2)
-    curvatures_Y.append(k)
-    
-  curvatures_Y = np.array(curvatures_Y)
-  coefs = poly.polyfit(curvatures_X, curvatures_Y, 0)
-  smooth_curvatures_Y = poly.polyval(curvatures_X, coefs)
-
[…]

### 2021-03-14 THESIS.tex
reorganization of the thesis

@@
 \subsection{Envelope of a digital signal}
-
-Despite being an elusive concept, the temporal amplitude envelope of a signal is essential for its complete characterization, being the primary information-carrying medium in spoken voice and telecommunications, for example.
-Intuitively, the temporal envelope can be understood as a smooth function that multiplies the signal, being responsible for its outer shape.
-Envelope detection techniques have applications in areas like health, sound classification and synthesis, seismology and speech recognition.
-Nevertheless, a general method to digital envelope detection of signals with rich spectral content doesn't exist, as most methods involve manual intervention, in the form of filter design, smoothing, as well as other specific design choices, based on a priori knowledge about the nature of the specific waves under investigation.
-To address this problem, we propose a framework that uses intrinsic characteristics of a signal to estimate its envelope, completely eliminating the necessity of parameter tuning.
-The approach here described draws inspiration from geometric concepts to isolate the frontiers and thus estimate the temporal envelope of an arbitrary signal; to that end, alpha-shapes, concave hulls, and discrete curvature are explored.
-We also define entities, such as a pulse and frontiers, in the context of an arbitrary digital signal, as a means to reduce dimensionality and the complexity of the proposed algorithm.
-Specifically, a new measure of discrete curvature is used to obtain the average radius of curvature of a discrete wave, serving as a threshold to identify the wave's frontier points.
-We find the algorithm accurate in the identification of the frontiers of a wide range of digital sound waves with very diverse characteristics, while localizing each pseudo-cycle of the wave in the time domain.
-The algorithm also compares favourably with classic envelope detection techniques based on filtering and the Hilbert Transform.
-Besides the most direct applications of this work to audio classification and synthesis, we foresee impact in compression techniques and machine learning approaches to audio.
-The discrete curvature definition presented could also be extended to three-dimensional settings, to improve shape detection algorithms based on alpha-shapes.
-
-\subsubsection{Envelope Detection}
-Envelope detection is ubiquitous in both analogue and digital signal processing \parencite{2011CaetanoImproved}.
-Nevertheless, the literature in this area is very fragmented \parencite{2017LyonsDigital}.
-Besides, most envelope detection techniques are designed to account for very specific settings, like pure sinusoids with moderate noise content, a limitation that excludes most physical signals, as is the case of recorded sound, for example; that limitation arises in part due to the lack of a strict mathematical definition of a temporal envelope \parencite{2013Mengempirical}.
-
-In many natural signals, however, the temporal amplitude envelope of a signal plays a prominent role in the characteristics exhibited: according to \textcite{2017QiRelative}, for example, the envelope is at least as important as the fine structure of a sound wave in the context of the intelligibility of mandarin tones.
-This is also the case for English \parencite{1995ShannonSpeech}, where even envelopes modulating mostly noise were still capable of conveying meaning.
-
-The envelope helps to impart emotion and identity to the human voice \parencite{2018ZhuContributions}, and envelope preserving characteristics in concert halls are associated with their pleasantness \parencite{2011LokkiEngaging}.
-
-When dealing with broadband signals, approaches tailored to specific applications are prevalent, such as the one presented by \textcite{2014YangFast} for the distributed monitoring of fibre optic or the one formulated by \textcite{2018AssefModeling} in the context of medical ultrasound imaging.
-
-Addressing the different units in the horizontal and vertical axes, one can transform the DSP problem of envelope detection into the geometric problem of defining the shape of a set of points in $ \mathbb{R}^2 $.
-For that purpose, \textcite{1983Edelsbrunnershape} introduced the concept of alpha-shapes, a mathematically well-defined extension to the convex hull of a finite set of points, closely related to the Delaunay triangulation and Voronoi diagrams of those points.
-
-This approach is used in areas such as the detection of features in images \parencite{2016VarytimidisAlpha}, reconstruction of surfaces from a cloud of points \parencite{2015WuAutomated} and Spectroscopy \parencite{2019XuModeling}, with the last work, which involves the estimation and removal of the Blaze function (a kind of envelope) of an echelle spectrograph, being particularly illustrative of the potential of cross-fertilization.
-
-Steps in the direction of translating geometric algorithms to the context of envelope detection were also made by \textcite{2015YangSkeleton} via an algorithm based on the construction of a skeleton underlying the wave, and also via the direct translation of computer vision methods to the task \parencite{2015YangRepresenting}.
-
-Following this path, we present a general approach to envelope detection, exploiting the intrinsic characteristics of a generic, spectrally complex wave, in order to avoid the need for manual intervention or parameter tuning.
-
-While the robustness of the proposed approach allows it to be used as a plug-in replacement for many methods encountered in the literature, we feel that it would be particularly useful for sound synthesis.
-
-The envelope is shown to add complexity to the spectral representation of a wave \parencite{2019TarjanoNeuro}, and an accurate description of the envelope, while describing the evolution of the instantaneous amplitude of a signal in time, would also greatly simplify further spectral analysis.
-Moreover, the algorithm developed in this work naturally divides a signal into its pseudo-cycles, pinpointing them in the time domain, providing the building blocks for the reconstruction of the fine structure of the wave.
-
-% The rest of this paper is divided as follows: After explaining and defining the various entities to be used subsequently, of which the concept of frontiers is arguably the most important, in the methodology section, we dedicate a section to present a new discrete curvature estimation approach, and how it was used to identify the frontiers of a discrete wave.
-
-We then proceed to illustrate the results of the application of the algorithm in a set of six diverse sound waves, ranging from voice to unpitched instruments comparing, afterwards, its performance with that of classic digital envelope estimation algorithms, both from a visual and numerical standpoint.
-
-We then discuss the implications of the work, emphasizing the focus on the construction of new interpretations of the relationship between time domain and frequency domain sound descriptions, and our vision of the impact of this methodology in the field of sound synthesis, providing directions for future developments.
-
-\subsection{Frequency content}
-
-\subsection{Representations}
-
-% one hot
-
-\section{Envelope Detection}
 
@@
 
+\subsection{Frequency content}
+
+\subsection{Representations} % sound compression paper
+
+% one hot
+
 % \section{References}

### 2021-03-19 THESIS.tex
added compression paper to thesis

@@
 
-\bibliography{00_BIBLI}
+\bibliography{00_BIBLI,Compression}
 
@@
   \centering
-    \includegraphics{images/00ConvexHull.pdf}
+    \includegraphics{images/envelope/00ConvexHull.pdf}
   \caption{A discrete signal interpreted as a set of points in the Cartesian plane, and the convex and concave hull at $ \alpha=300 $ defined by it.}
@@
   \centering
-    \includegraphics{images/01AnalyticSignal.pdf}
+    \includegraphics{images/envelope/01AnalyticSignal.pdf}
   \caption{Envelope of a pure sinusoid with a local frequency of 20 cycles modulated by a polynomial of degree 3, as obtained by the Hilbert Transform approach.
@@
   \centering
-    \includegraphics{images/02Envelope.pdf}
+    \includegraphics{images/envelope/02Envelope.pdf}
   \caption{Example of a discrete wave \textbf{w} composed by the element-wise multiplication of an envelope \textbf{e} and a carrier \textbf{c}.
@@
   \centering
-    \includegraphics{images/03Pulses.pdf}
+    \includegraphics{images/envelope/03Pulses.pdf}
   \caption{Example of a discrete wave \textbf{w} divided into pulses, of which the extrema are highlighted with a circle.
@@
   \centering
-    \includegraphics{images/04Vectors.pdf}
+    \includegraphics{images/envelope/04Vectors.pdf}
   \caption{The set of points P and the set V of vectors between two adjacent points of P, in both the original and normalized coordinate systems, shown in scale.}
@@
   \centering
-    \includegraphics{images/05DiscreteCurvature.pdf}
+    \includegraphics{images/envelope/05DiscreteCurvature.pdf}
   \caption{The tangent unit vector of the circle changes from the horizontal direction in $ \hat{u}_0 $ to an inclination of $ \theta_k $ in $ \hat{u}_1 $, $ \theta_k $ being the angle that the vector $ v_k $ makes with the horizontal direction.}
@@
   \centering
-    \includegraphics{images/06Graph_bend.pdf}
+    \includegraphics{images/envelope/06Graph_bend.pdf}
   \caption{Signal, envelope and carrier for a record of a bend performed on an electric guitar.}
@@
   \centering
-    \includegraphics{images/06Graph_spoken_voice.pdf}
+    \includegraphics{images/envelope/06Graph_spoken_voice.pdf}
   \caption{Signal, envelope and carrier for a record of male voice uttering the word “amazing”.}
@@
   \centering
-    \includegraphics{images/06Graph_alto.pdf}
+    \includegraphics{images/envelope/06Graph_alto.pdf}
   \caption{Signal, envelope and carrier for a record of an alto singer.}
@@
   \centering
-    \includegraphics{images/07Comparison_sinusoid.pdf}
+    \includegraphics{images/envelope/07Comparison_sinusoid.pdf}
   \caption{Comparison of envelope detection algorithms for a simple sinusoid}
@@
   \centering
-    \includegraphics{images/07Comparison_piano.pdf}
+    \includegraphics{images/envelope/07Comparison_piano.pdf}
   \caption{Comparison of envelope detection algorithms for the recording of the key 33 of a grand piano}
@@
   \centering
-    \includegraphics{images/07Comparison_soprano.pdf}
+    \includegraphics{images/envelope/07Comparison_soprano.pdf}
   \caption{Comparison of envelope detection algorithms for the recording of the vocalization of a soprano singer}
@@
   \centering
-    \includegraphics{images/08FullFrontiers.pdf}
+    \includegraphics{images/envelope/08FullFrontiers.pdf}
   \caption{Positive and negative frontiers of six digital waves, as extracted by the algorithm here presented.
@@
   \centering
-    \includegraphics{images/11Fourier.pdf}
+    \includegraphics{images/envelope/11Fourier.pdf}
   \caption{Fourier power spectrum for the wave and carrier shown in \ref{fig:Envelope}.}
@@
   \centering
-    \includegraphics{images/09PCs.pdf}
+    \includegraphics{images/envelope/09PCs.pdf}
   \caption{Section of the lines formed by the positions of the positive and negative frontiers of the digital wave illustrated in \ref{fig:singer_voice}, as extracted by the algorithm presented, and their average.}
@@
   \centering
-    \includegraphics{images/10AvgPc.pdf}
+    \includegraphics{images/envelope/10AvgPc.pdf}
   \caption{Pseudo-cycles of the wave illustrated in \ref{fig:singer_voice}, segmented and superposed, and their average.
@@
 
+\section{Pseudo cycles based representation}
+
+Sample-based virtual instruments are the de facto standard for the emulation of real-world acoustic instruments.
+Nevertheless, little attention is given to the encoding of the samples used in those instruments; lossless formats like WAV or proprietary codecs are common, pushing the on-disk size of those libraries to sizes exceeding the gigabytes.
+This renders the most realistic digital instruments inaccessible in many but the most high-end hardware; controllers, like electronic drum kits and keyboards and even less robust personal computers are often unable to take advantage of those libraries.
+We propose a lossy compression algorithm tailored specifically to compress pseudo periodic signals, as those comprising these instruments; due to the highly redundant nature of these sounds, improvements in relation to traditional codecs can be obtained.
+The algorithm operates entirely in the time domain, avoiding domain conversions common in lossy compression algorithms, thus exhibiting low latency in the decoding phase.
+It operates iteratively, dividing a signal into its pseudo cycles according to an estimate of the average waveform of those pseudo cycles, which is thus updated, and the process is repeated.
+We provide C++ source code and implementation of the algorithm, that was used to compare its performance against the MP3, AAC, and Opus codecs, finding the proposed method to be at last 13 times faster in decoding tasks, while maintaining comparable compression rate and quality, as measured by MSE and the ViSQOL library.
+A website and a GitHub repository with complementary material, where results can be heard, are also available.
+
+Domain specific compression algorithms and codecs exist in many areas.
+\cite{2016ChenCompressing}, for example, presents a scheme for compressing convolutional neural networks, while \cite{2014CanovasLossy}, besides revising existing methods, proposes two lossy schemes for the compression of quality score data of genomic sequencing.
+The use of lossy compression for wireless networks of sensors is studied in \cite{2014ZordanPerformance}.
+
+More recently, the work of \cite{2019CalhounExploring} explores the benefits of specific lossy compression algorithms in the context of checkpoints of computational simulations that are stored between simulation sessions; New technologies, such as 3D immersive audio, have propelled specific compression algorithms, such as the one presented in \cite{2021HuAudio}.
+
+Regarding sample-based digital musical instruments, however, no such specialized algorithms exist despite those signals, being highly uniform, offering a potential for compression that is not present in most general sounds and remains largely unexplored.
+
+This overlook arises in part from the marketing strategy of the majority of those products, where the sheer size of the library is advertised as an equivalent of quality and versatility.
+An overall prejudice against compressed formats in general from the part of the music production community supports this biased view.
+
+At last some of the criticism directed to the MP3 standard, and lossy codecs in general, is meant to establish a symbolic capital in the face of the crescent struggle between established recording labels and alternative channels of music diffusion, such as streaming services \cite{2020OGradyRethinking}.
+
+The adoption of special compressed formats would enable state-of-the-art digital instruments to be implemented in a wider range of hardware, where the storage costs now make the cost prohibitive, such as electronic drum kits, digital pianos, and even cell phones and tablets, democratizing those tools.
+
+A recent review of the advancements in data compression in general can be found in \cite{2021Jayasankarsurvey}.
+Accounts of the research in more specific areas, such as video \cite{2019LaudeComprehensive} and image \cite{2014RehmanImage} were also produced recently.
+
+Concerning lossless audio compression, research seems to be lagging when compared to lossy approaches; \cite{2001HansLossless} suggested that a limit was about to be reached at the beginning of the 2000s.
+
+The renaissance of neural networks led to interesting results, for example the use of Generative adversarial networks for audio compression presented in \cite{2018HuangBandwidth}.
+
+The ACER codec, as presented in \cite{2014CunninghamData}, offers an unconventional approach to compression, in that it identifies redundant pieces in a signal that are indexed in a dictionary.
+The results seem to be satisfactory, from a perceptual standpoint, when compared with MP3 and AAC codecs \cite{2019CunninghamSubjective}.
+
+None of those approaches, however, are designed to take advantage of the redundancies present in harmonic sounds, nor were designed with the low latency requirements imposed by digital musical instruments, especially when operating in real-time; to fill that gap, we propose the method presented in this letter.
+
+\subsection{Methodology}
+
+A continuous or discrete signal can be interpreted as a succession of consecutive waveforms, each one representing a (pseudo) cycle of the wave.
+In the case of perfectly periodic signals, those waveforms are identical, and we need only the definition of one waveform to characterize the whole signal.
+In pseudo periodic waves, on the other hand, there are variations between pseudo cycles, that must be accounted for.
+
+Figure \ref{fig:pseudocycles} illustrates a possible division in pseudo cycles of a segment of the discrete representation of the sound of a bowed cello.
+It can be seen that the waveform variation between pseudo cycles is very small, on account of the sound being harmonic.
+
+\begin{figure}%[H]
+\centerline{\includegraphics[width=\columnwidth]{images/compression/01_pseudocycles.pdf}}
+\caption{A discrete wave representing the sound of a bowed cello. The red lines represent the frontiers between successive pseudo cycles, as divided using the presented algorithm.}
+\label{fig:pseudocycles}
+\end{figure}
+
+This rationale of defining a signal as a series of a modified basis function is presented more formally in \cite{1993SinhaLow}; in that work, however, the bases are wavelets obtained via filter banks applied to the original signal using a pyramid algorithm.
+A similar approach can be found in \cite{1998SrinivasanHigh}, where a perceptual metric is used.
+In the presented work this basis is derived directly from the division in pseudo cycles of the underlying signal.
+
+If the wave is sufficiently well-behaved, as is the case with the harmonic sounds used in sample-based digital musical instruments, the bulk of those changes can be attributed to variations in both the amplitude and length of the base waveform; The present algorithm is based on identifying the best basis waveform and registering how it changes over time.
+
+We accomplish that encoding via an iterative algorithm, that divides the signal of interest using a waveform estimate.
+An updated basis waveform can then be obtained, and the process repeats while improvement in the mean squared error between the initial uncompressed and the compressed signal is being made.
+
+Figure \ref{fig:averagewaveformcello} shows how the base waveform, shown in red, for the cello sample; the base waveform is the average of each individual waveform, after being scaled to the same length.
+To ensure the continuity of the final wave, the average waveform is rotated between each iteration of the algorithm, so that the absolute vertical distance to zero of its first and last elements is the smallest possible.
+
+\begin{figure}%[H]
+\centerline{\includegraphics[width=\columnwidth]{images/compression/02_averagewaveformcello.pdf}}
+\caption{The various pseudo cycles in which the signal is divided are shown superimposed, after being scaled to the same length. The average of those individual waveforms is shown in red.}
+\label{fig:averagewaveformcello}
+\end{figure}
+
+For some signals, the average waveform resembles very closely a sinusoid, as can be seen in Fig.
+\ref{fig:averagewaveformpiano} that illustrates the average waveform in the case of the sound of a piano key.
+From a physical perspective, the strings in a piano are excited with an initial impulse left to vibrate undisturbed.
+After an initial transient period, it is natural that a harmonic mode becomes predominant.
+
+\begin{figure}%[H]
+\centerline{\includegraphics[width=\columnwidth]{images/compression/02_averagewaveformpiano.pdf}}
+\caption{The various pseudo cycles in which the signal is divided are shown superimposed, after being scaled to the same length. The average of those waveforms is shown in red.}
+\label{fig:averagewaveformpiano}
+\end{figure}
+
+In figure \ref{fig:periodsenvelope} the envelope and the normalized periods of the wave are shown.
+Using this data to modulate both the amplitude and the length of the average waveform presented in Fig.
+\ref{fig:averagewaveformcello} above, one can approximate the original signal, as will be shown in the rest of this chapter.
+
+\begin{figure}%[H]
+\centerline{\includegraphics[width=\columnwidth]{images/compression/03_periodsenvelope.pdf}}
+\caption{The envelope and the normalized length of each pseudo cycle.}
+\label{fig:periodsenvelope}
+\end{figure}
+
+While the raw representation of the original cello signal can be viewed as an array with 129867 individual entries, for the compact representation we have 396 pairs of numbers representing the amplitude and length of each pseudo cycle, and the average waveform represented by an array of size 326.
+The length of the array storing the encoded representation, in the particular case of the cello sample, would be 1118, or approximately 0.86\% of the original, uncompressed representation.
+
+Starting with an initial estimate of the waveform of the signal being processed, one can identify the pseudo cycles of a signal by gradually multiplying, element-wise, different lengths of this estimated waveform with the signal, and observing the highest average value.
+
+For the placement of the initial waveform, one has also to account for the potential phase shift in the signal.
+This is accomplished by convolving the different lengths of the waveform starting from different positions.
+For the subsequent placements, the initial position will be already fixed at the end of the preceding pseudo cycle, the length being the only quantity to be estimated.
+
+Figure \ref{fig:tdists} shows how many times a period with a specific length occurs, for 4 different sound samples illustrating different settings.
+The first two samples are from instruments, a cello, and a piano, and the last two are voices, from a female soprano and a male bass.
+
+The data in blue refers to the initial division of the waves, as obtained by the algorithm presented in \cite{2020TarjanoDigital}.
+Since the algorithm was primarily designed to identify the envelope of a wave, the variation in the lengths of the pseudo cycles is naturally greater than the red data, where the final result after the application of the current method is shown.
+
+The distribution of the periods' length tends to present a smaller variance, concentrating around the mode, as the algorithm progresses, and better basis waveforms are found.
+None of these distributions, as per the Shapiro test, are significantly close to a normal distribution, as can be observed comparing the distributions with the equivalent normal distributions plotted in the figure.
+
+The normal distributions, as constructed with mean equal to the mode of the periods distribution and standard deviation equal to the standard variation around the mode tend to have a greater spread.
+Investigating the range between 3 standard deviations around the mode is a very conservative assumption, then, that we can afford to make without compromising the efficiency of the algorithm.
+
+\begin{figure}%[H]
+\centerline{\includegraphics[width=\columnwidth]{images/compression/04_Tdist.pdf}}
+\caption{Distribution of the number of occurrences of periods lengths for 4 samples, both for the initial waveform in blue and the final waveform in black.}
+\label{fig:tdists}
+\end{figure}
+
+Fig.
+\ref{fig:convolution} illustrates, for the cello wave, the procedure for the identification of the first pseudo cycle of the wave, for a particular period.
+Although only one example is shown in the picture, the various lengths of the waveform are convolved with the beginning of the original signal, and the optimum chosen.
+
+\begin{figure}%[H]
+\centerline{\includegraphics[width=\columnwidth]{images/compression/05_convolution.pdf}}
+\caption{Identification of the first pseudo cycle.
+The intensity of the color is proportional to the average of the multiplication between the basis waveform and the underlying original signal.
+Only one waveform length is shown, translated in intervals of 10 frames.}
+\label{fig:convolution}
+\end{figure}
+
+For the subsequent periods, the procedure is as illustrated in Fig.
+\ref{fig:scaling}, as taking the original cello wave as an example: only the length of the basis waveform is changed, using cubic interpolation.
+At the first iteration, the steps illustrated in both Fig.
+\ref{fig:convolution} and Fig.
+\ref{fig:scaling} are performed with the use of a sinusoid, to ensure the best initial division and faster convergence of the algorithm.
+
+\begin{figure}%[H]
+\centerline{\includegraphics[width=\columnwidth]{images/compression/06_scaling.pdf}}
+\caption{Identification of the second pseudo cycle, by multiplying different lengths of the basis waveform with the underlying original signal.
+The intensity of the color is proportional to the average of the multiplication.}
+\label{fig:scaling}
+\end{figure}
+
+\subsection{Results}
+In this section, we compare the algorithm here presented with the most common lossy codecs in a variety of sound samples chosen to reflect the proposed immediate application of the algorithm, and thus consist of excerpts from notes from various instruments, as well as singing voice, as would be encountered in the libraries of sample-based instruments.
+Results can be heard on the website prepared for this letter \cite{2021TarjanoCompression}.
+
+The samples from soprano, mezzo-soprano, and bass singers were obtained by processing sounds from the Vocalset library\cite{2018WilkinsVocalset}, and are licensed under the Creative Commons Attribution 4.0 International.
+The male singer sample was based on a sample from the Human Voice Dataset \cite{2015Vocoboxvocobox}, available under the MIT License.
+
+The double bass, french horn, flute, and saxophone samples were based on samples freely available at the resources page of the Phillarmonia symphony orchestra website \cite{Philharmonia}, while the piano, cello, and violin samples were processed from the basis samples offered in the University of Iowa Musical Instrument Samples (MIS) \cite{FrittsUniversity}, and are also freely available.
+
+All original samples are uncompressed audio encoded in the WAV format, at a standard frame rate of 44100 fps and a bit rate of 705 kbps, mono that were normalized with long silences manually removed.
+The format was chosen for its popularity both in general \cite{2014LuoIdentifying} and in sample-based digital instruments.
+
+For the compressed formats, three lossy codecs were chosen.
+The first format we compare to is the popular MP3, the third layer of the MPEG-1/2 specification, proposed in 1992 \cite{2018BritanakAudio} by the Moving Picture Experts Group.
+
+The AAC format, in addition to being the formal mp3 successor, is used in services like Apple iTunes and YouTube \cite{2016SeichterAAC}.
+It was developed by a joint effort between industry representatives such as Dolby and Sony and academic representation through Bell Labs and the University of Hannover, for example \cite{2018BritanakAudio}.
+
+Outside the MPEG committee, the open-source Opus format, standardized by the Internet Engineering Task Force (IETF) in 2012, was chosen.
+Besides being considered superior to its predecessor Vorbis and the AAC format \cite{2018GunawanPerformance}, the codec supports both speech and music, has very low delay \cite{2013Valinhigh}, being used in real-time audio applications such as WhatsApp \cite{2019SrivastavaSmartphone}.
+
+The Opus codec, along with ACC and MP3, forms the set of the most advanced \cite{2019LaudeComprehensive} and popular \cite{2018KwanObjective} lossy compression formats; Readily available implementations of more experimental algorithms, such as the aforementioned ACER codec \cite{2014CunninghamData}, or the wavelet-based approaches presented in \cite{1993SinhaLow} and \cite{1998SrinivasanHigh} couldn't be found.
+
+All conversions, except for the MP3 encoding done with the Lame library, were made with the FFmpeg library.
+The PowerShell scripts used are available in the letter's repository \cite{2021TarjanoCompression}.
+The encoding configurations were chosen in order to generate files with sizes compatible with the present algorithm: for the MP3 format, a bit rate of 8 kb/s and a frame rate of 8000 frames/s was chosen.
+For the other two formats, no frame rate reduction was needed, and the bit rates were 4 kb/s for the ACC format and 6 kb/s for the Opus format.
+
+As can be seen in the accompanying website, the present work exhibits an average compression rate of 1.09\%, slightly larger than the 0.92\% obtained with the Opus codec, and better than the AAC (1.19\%) and MP3 (1.20\%) codecs.
+
+\subsection{Decoding times}
+Encoding times, as can be seen in the accompanying website, are around two times the duration of the original signal, and thus considerably higher than conventional codecs: While part of this discrepancy can be attributed to the immaturity of the implementation when compared to well-established ones, like lame for mp3, with decades of refinement, the majority of the difference can be attributed to the different approach used by the presented algorithm, engaging in a semi exhaustive search for the best basis purely in the time domain.
+
+This approach pays off during the decoding phase, however, since the representation doesn't need to be translated from the frequency domain, for example.
+In the context of sample-based digital musical instruments, the encoding occurs only once, and takes place in more robust hardware; it is more important to alleviate as much as possible the decoding effort, since, besides enabling the use of more affordable hardware, this task will be performed a far greater number of times.
+
+For digital musical instruments, it is important to maintain latency below 10 milliseconds, since that is the limit above which users start to experiment performance degradation \cite{2002WesselProblems,2016JackEffect}.
+Table \ref{tab:decoding} compares the absolute decoding times, in milliseconds; it can be seen that the proposed algorithm presents the lowest decoding times, with an average of 12 ms against an average of 163.5 ms of the Opus codec, the second-lowest.
+Besides, as the reconstruction of the sound is made linearly, pseudo cycle by pseudo cycle, entirely in the time domain, it would be possible to alternate between decoding and reproduction tasks, keeping the latency below the acceptable limit for samples of arbitrary length.
+
+\begin{table}
+\centering
+\caption{Original Duration and Decoding Times (milliseconds) for the Present Work and the MP3, ACC, and Opus lossy formats.}
+\small
+\setlength{\tabcolsep}{3pt}
+\begin{tabular}{lccccc}
+\hline
+Sample & Duration & P.W. & MP3 & ACC & Opus \\ 
+\hline
+cello & 2945 & 11 & 105 & 111 & 136 \\
+doublebass & 1388 & 4 & 103 & 201 & 108 \\
+flute & 11115 & 17 & 110 & 125 & 117 \\
+frenchhorn & 8713 & 21 & 211 & 226 & 223 \\
+mezzosoprano & 3259 & 9 & 163 & 147 & 116 \\
+piano & 4729 & 19 & 728 & 506 & 520 \\
+saxophone & 19825 & 33 & 109 & 115 & 119 \\
+singerbass & 2624 & 8 & 123 & 198 & 197 \\
+singermale & 1214 & 3 & 104 & 199 & 106 \\
+sopranoA & 2598 & 8 & 104 & 120 & 106 \\
+sopranoB & 2299 & 5 & 115 & 106 & 106 \\
+violin & 2271 & 6 & 111 & 105 & 108 \\ 
+% \hline
+% Average & 5248.35 & 12.00 & 173.83 & 179.92 & 163.50 \\
+% SD & 5486.06 & 8.88 & 177.45 & 111.73 & 118.70 \\
+\hline
+\end{tabular}
+\label{tab:decoding}
+\end{table}
+
+\subsection{Quality}
+
+To assess the overall quality of the proposed algorithm, in addition to the computation of the mean square error in relation to the original signal, we made use of the ViSQOL \cite{2020ChinenViSQOL} open-source library developed by Google, designed to substitute subjective listening tests when those are unpractical.
+According to \cite{2017SloanObjective}, ViSQOL is a more accurate alternative to well-known methods, like PEAQ and POLQA.
+
+Table \ref{tab:quality} shows the ranking of the compressed formats in those metrics: the MP3 codec performed best in all of them; to reach the compatible size with the other codecs, however, it was the only one that needed to be downsampled.
+
+\begin{table}
+\centering
+\caption{Ranking of the codecs compared under different quality metrics.}
+\small
+\setlength{\tabcolsep}{3pt}
+\begin{tabular}{lcccc}
+\hline
+Method & Present Work & MP3 & ACC & Opus \\
+\hline
+MSE & 3 & 1 & 4 & 2 \\
+ViSQOL audio & 4 & 1 & 2 & 3 \\
+ViSQOL speech & 2 & 1 & 4 & 3 \\
+\hline
+\end{tabular}
+\label{tab:quality}
+\end{table}
+
+Although the ViSQOL is designed as a cost-efficient alternative to subjective listening tests, it by any means claims to be a perfect replacement; The reader is strongly encouraged to hear the results in the samples section of the accompanying website  \cite{2021TarjanoCompression}.
+
+Despite performing better than the AAC codec in the objective measures, files compressed with the Opus codec present artifacts, in the majority of the samples, that renders them almost unintelligible.
+To a lesser extent, that is also the case of the AAC codec.
+
+While presenting an efficient algorithm to compress harmonic signals, with quality on par with the MP3, AAC, and Opus codecs and faster decoding times, immediately applicable to digital musical instruments, this work introduces a new framework for representing pseudo periodic waves, that can serve as a basis for future research.
+Fig.
+\ref{fig:waveforms} illustrates this representation, in the case of the cello sample.
+
+Starting from the representation in Fig.
+\ref{fig:waveforms}, the surface can be approximated by parametric surfaces, or machine learning approaches, providing more robustness in relation to the average waveform approach, for example.
+
+\begin{figure}%[H]
+\centerline{\includegraphics[width=1 \columnwidth]{images/compression/07_heatmap.pdf}}
+\caption{The evolution of the cello sample waveform over time.}
+\label{fig:waveforms}
+\end{figure}
+
 % \section{References}

### 2021-03-19 images/compression/01_pseudocycles.py

@@ -1,0 +1,136 @@
+# from os import linesep
+from statistics import mode
+import numpy as np
+import plotly.graph_objects as go
+import signal_envelope as se
+from scipy import interpolate
+import sys
+import os
+
+def get_pcs(Xpc, W):
+  Xpc = Xpc.astype(int)
+  # amp = np.max(np.abs(W))
+  # max_T = int(np.max(np.abs(Xpc[1:] - Xpc[:-1])))
+  Xlocal = np.linspace(0, 1, mode(Xpc[1:] - Xpc[:-1]))
+
+  orig_pcs = []
+  norm_pcs = []
+  for i in range(2, Xpc.size):
+    x0 = Xpc[i - 1]
+    x1 = Xpc[i] + 1
+    orig_pcs.append(W[x0 : x1])
+    if x1 - x0 >= 4:
+      yx = interpolate.interp1d(np.linspace(0, 1, x1-x0), W[x0 : x1], "cubic")
+      Ylocal = yx(Xlocal)
+      # Ylocal = Ylocal / np.max(np.abs(Ylocal)) * amp
+      norm_pcs.append(Ylocal)
+  return np.average(np.array(norm_pcs), 0), orig_pcs, norm_pcs
+
+def to_plot(Matrix):
+  X = []
+  Y = []
+  for line in Matrix:
+    for x, y in enumerate(line):
+      X.append(x)
+      Y.append(y)
+    X.append(None)
+    Y.append(None)
+  return X, Y
+
+name = "cello"
+
+'''###### Read wav file ######'''
+W, fps = se.read_wav(f"./original_samples/{name}.wav")
+W = W - np.average(W)
+amp = np.max(np.abs(W))
+W = W / amp
+n = W.size
+
+'''###### Read Pseudo-cycles info ######'''
+Xpc = np.genfromtxt(f"./csvs/{name}.csv", delimiter=",")
+
+X_xpcs, Y_xpcs = [], []
+for x in Xpc:
+  X_xpcs.append(x)
+  X_xpcs.append(x)
+  X_xpcs.append(None)
+  Y_xpcs.append(-1)
+  Y_xpcs.append(1)
+  Y_xpcs.append(None)
+
+
+# average_waveform, orig_waveforms, norm_waveforms = get_pcs(Xpc, W)
+
+# norm_waveforms = norm_waveforms - average_waveform
+
+
+
+'''============================================================================'''
+
+FONT = dict(
+    family="Latin Modern Roman",
+    color="black",
+    size=13.3333
+  )
+
[…]

### 2021-03-19 images/compression/02_averagewaveformpiano.py

@@ -1,0 +1,180 @@
+# from os import linesep
+from statistics import mode
+import numpy as np
+import plotly.graph_objects as go
+import signal_envelope as se
+from scipy import interpolate
+from plotly.subplots import make_subplots
+import sys
+
+
+def get_pcs(Xpc, W):
+  Xpc = Xpc.astype(int)
+  # amp = np.max(np.abs(W))
+  # max_T = int(np.max(np.abs(Xpc[1:] - Xpc[:-1])))
+  Xlocal = np.linspace(0, 1, mode(Xpc[1:] - Xpc[:-1]))
+
+  orig_pcs = []
+  norm_pcs = []
+  for i in range(2, Xpc.size):
+    x0 = Xpc[i - 1]
+    x1 = Xpc[i] + 1
+    orig_pcs.append(W[x0 : x1])
+    if x1 - x0 >= 4:
+      yx = interpolate.interp1d(np.linspace(0, 1, x1-x0), W[x0 : x1], "cubic")
+      Ylocal = yx(Xlocal)
+      # Ylocal = Ylocal / np.max(np.abs(Ylocal)) * amp
+      norm_pcs.append(Ylocal)
+  return np.average(np.array(norm_pcs), 0), orig_pcs, norm_pcs
+
+def to_plot(Matrix):
+  X = []
+  Y = []
+  for line in Matrix:
+    for x, y in enumerate(line):
+      X.append(x)
+      Y.append(y)
+    X.append(None)
+    Y.append(None)
+  return X, Y
+
+name = "piano"
+
+'''###### Read wav file ######'''
+W, fps = se.read_wav(f"./original_samples/{name}.wav")
+W = W - np.average(W)
+amp = np.max(np.abs(W))
+W = W / amp
+n = W.size
+
+'''###### Read Pseudo-cycles info ######'''
+Xpc = np.genfromtxt(f"./csvs/{name}.csv", delimiter=",")
+
+X_xpcs, Y_xpcs = [], []
+for x in Xpc:
+  X_xpcs.append(x)
+  X_xpcs.append(x)
+  X_xpcs.append(None)
+  Y_xpcs.append(-1)
+  Y_xpcs.append(1)
+  Y_xpcs.append(None)
+
+
+average_waveform, orig_waveforms, norm_waveforms = get_pcs(Xpc, W)
+
+# norm_waveforms = norm_waveforms - average_waveform
+
+
+
+'''============================================================================'''
+
+FONT = dict(
[…]

### 2021-03-19 images/compression/03_periodsenvelope.py

@@ -1,0 +1,142 @@
+# from os import linesep
+from statistics import mode
+import numpy as np
+import plotly.graph_objects as go
+import signal_envelope as se
+from scipy import interpolate
+import sys
+
+
+def get_pcs(Xpc, W):
+  Xpc = Xpc.astype(int)
+  # amp = np.max(np.abs(W))
+  # max_T = int(np.max(np.abs(Xpc[1:] - Xpc[:-1])))
+  Xlocal = np.linspace(0, 1, mode(Xpc[1:] - Xpc[:-1]))
+
+  orig_pcs = []
+  norm_pcs = []
+  for i in range(2, Xpc.size):
+    x0 = Xpc[i - 1]
+    x1 = Xpc[i] + 1
+    orig_pcs.append(W[x0 : x1])
+    if x1 - x0 >= 4:
+      yx = interpolate.interp1d(np.linspace(0, 1, x1-x0), W[x0 : x1], "cubic")
+      Ylocal = yx(Xlocal)
+      # Ylocal = Ylocal / np.max(np.abs(Ylocal)) * amp
+      norm_pcs.append(Ylocal)
+  return np.average(np.array(norm_pcs), 0), orig_pcs, norm_pcs
+
+def to_plot(Matrix):
+  X = []
+  Y = []
+  for line in Matrix:
+    for x, y in enumerate(line):
+      X.append(x)
+      Y.append(y)
+    X.append(None)
+    Y.append(None)
+  return X, Y
+
+name = "cello"
+
+'''###### Read wav file ######'''
+W, fps = se.read_wav(f"./original_samples/{name}.wav")
+W = W - np.average(W)
+amp = np.max(np.abs(W))
+W = W / amp
+n = W.size
+
+'''###### Read Pseudo-cycles info ######'''
+Xpc = np.genfromtxt(f"./csvs/{name}.csv", delimiter=",")
+
+X_xpcs, Y_xpcs = [], []
+for x in Xpc:
+  X_xpcs.append(x)
+  X_xpcs.append(x)
+  X_xpcs.append(None)
+  Y_xpcs.append(-1)
+  Y_xpcs.append(1)
+  Y_xpcs.append(None)
+
+
+average_waveform, orig_waveforms, norm_waveforms = get_pcs(Xpc, W)
+
+T = Xpc[1 :] - Xpc[: -1]
+T = T / np.average(T)
+E = [np.max(np.abs(waveform)) for waveform in orig_waveforms]
+
[…]

### 2021-03-19 images/compression/04_Tdist.py

@@ -1,0 +1,266 @@
+from plotly.subplots import make_subplots
+from statistics import mode
+import numpy as np
+import plotly.graph_objects as go
+import signal_envelope as se
+from scipy import interpolate
+import sys
+from scipy.stats import shapiro, norm
+
+def get_pcs(Xpc, W):
+  Xpc = Xpc.astype(int)
+  # amp = np.max(np.abs(W))
+  # max_T = int(np.max(np.abs(Xpc[1:] - Xpc[:-1])))
+  Xlocal = np.linspace(0, 1, mode(Xpc[1:] - Xpc[:-1]))
+
+  orig_pcs = []
+  norm_pcs = []
+  for i in range(2, Xpc.size):
+    x0 = Xpc[i - 1]
+    x1 = Xpc[i] + 1
+    orig_pcs.append(W[x0 : x1])
+    if x1 - x0 >= 4:
+      yx = interpolate.interp1d(np.linspace(0, 1, x1-x0), W[x0 : x1], "cubic")
+      Ylocal = yx(Xlocal)
+      # Ylocal = Ylocal / np.max(np.abs(Ylocal)) * amp
+      norm_pcs.append(Ylocal)
+  return np.average(np.array(norm_pcs), 0), orig_pcs, norm_pcs
+
+def to_plot(Matrix):
+  X = []
+  Y = []
+  for line in Matrix:
+    for x, y in enumerate(line):
+      X.append(x)
+      Y.append(y)
+    X.append(None)
+    Y.append(None)
+  return X, Y
+
+def get_periods(name):
+  '''###### Read wav file ######'''
+  W, fps = se.read_wav(f"./original_samples/{name}.wav")
+  W = W - np.average(W)
+  amp = np.max(np.abs(W))
+  W = W / amp
+  n = W.size
+
+  Xpc0 = se.get_frontiers(W, mode=1)
+
+  '''###### Read Pseudo-cycles info ######'''
+  Xpc = np.genfromtxt(f"./csvs/{name}.csv", delimiter=",")
+
+  T0 = (Xpc0[1 :] - Xpc0[: -1]).astype(np.int)
+  T = (Xpc[1 :] - Xpc[: -1]).astype(np.int)
+  l = max(np.max(np.abs(T0)), np.max(np.abs(T)))
+
+  avg = mode(T0)
+  sd = np.sqrt(np.average((avg - T0)**2))
[…]

### 2021-03-19 images/compression/05_convolution.py

@@ -1,0 +1,179 @@
+# from os import linesep
+from statistics import mode
+import numpy as np
+import plotly.graph_objects as go
+import signal_envelope as se
+from scipy import interpolate
+import sys
+
+
+def get_pcs(Xpc, W):
+  Xpc = Xpc.astype(int)
+  # amp = np.max(np.abs(W))
+  # max_T = int(np.max(np.abs(Xpc[1:] - Xpc[:-1])))
+  Xlocal = np.linspace(0, 1, mode(Xpc[1:] - Xpc[:-1]))
+  orig_pcs = []
+  norm_pcs = []
+  for i in range(2, Xpc.size):
+    x0 = Xpc[i - 1]
+    x1 = Xpc[i] + 1
+    orig_pcs.append(W[x0 : x1])
+    if x1 - x0 >= 4:
+      yx = interpolate.interp1d(np.linspace(0, 1, x1-x0), W[x0 : x1], "cubic")
+      Ylocal = yx(Xlocal)
+      # Ylocal = Ylocal / np.max(np.abs(Ylocal)) * amp
+      norm_pcs.append(Ylocal)
+  return np.average(np.array(norm_pcs), 0), orig_pcs, norm_pcs
+
+def to_plot(Matrix):
+  X = []
+  Y = []
+  for line in Matrix:
+    for x, y in enumerate(line):
+      X.append(x)
+      Y.append(y)
+    X.append(None)
+    Y.append(None)
+  return X, Y
+
+name = "cello"
+
+'''###### Read wav file ######'''
+W, fps = se.read_wav(f"./original_samples/{name}.wav")
+W = W - np.average(W)
+amp = np.max(np.abs(W))
+W = W / amp
+n = W.size
+
+Xpc0 = se.get_frontiers(W)
+
+'''###### Read Pseudo-cycles info ######'''
+Xpc = np.genfromtxt(f"./csvs/{name}.csv", delimiter=",")
+
+# X_xpcs, Y_xpcs = [], []
+# for x in Xpc:
+#   X_xpcs.append(x)
+#   X_xpcs.append(x)
+#   X_xpcs.append(None)
+#   Y_xpcs.append(-1)
+#   Y_xpcs.append(1)
+#   Y_xpcs.append(None)
+
+
+average_waveform, orig_waveforms, norm_waveforms = get_pcs(Xpc, W)
+
+T = (Xpc[1 :] - Xpc[: -1]).astype(np.int)
+
+# W = np.pad(W, mode(T))
+
+
[…]

### 2021-03-19 images/compression/07_heatmap.py

@@ -1,0 +1,117 @@
+# from os import linesep
+from statistics import mode
+import numpy as np
+import plotly.graph_objects as go
+import signal_envelope as se
+from scipy import interpolate
+import sys
+
+
+def get_pcs(Xpc, W):
+  Xpc = Xpc.astype(int)
+  # amp = np.max(np.abs(W))
+  # max_T = int(np.max(np.abs(Xpc[1:] - Xpc[:-1])))
+  Xlocal = np.linspace(0, 1, mode(Xpc[1:] - Xpc[:-1]))
+
+  orig_pcs = []
+  norm_pcs = []
+  for i in range(2, Xpc.size):
+    x0 = Xpc[i - 1]
+    x1 = Xpc[i] + 1
+    orig_pcs.append(W[x0 : x1])
+    if x1 - x0 >= 4:
+      yx = interpolate.interp1d(np.linspace(0, 1, x1-x0), W[x0 : x1], "cubic")
+      Ylocal = yx(Xlocal)
+      # Ylocal = Ylocal / np.max(np.abs(Ylocal)) * amp
+      norm_pcs.append(Ylocal)
+  return np.average(np.array(norm_pcs), 0), orig_pcs, norm_pcs
+
+def to_plot(Matrix):
+  X = []
+  Y = []
+  for line in Matrix:
+    for x, y in enumerate(line):
+      X.append(x)
+      Y.append(y)
+    X.append(None)
+    Y.append(None)
+  return X, Y
+
+name = "cello"
+
+'''###### Read wav file ######'''
+W, fps = se.read_wav(f"./original_samples/{name}.wav")
+W = W - np.average(W)
+amp = np.max(np.abs(W))
+W = W / amp
+n = W.size
+
+'''###### Read Pseudo-cycles info ######'''
+Xpc = np.genfromtxt(f"./csvs/{name}.csv", delimiter=",")
+
+X_xpcs, Y_xpcs = [], []
+for x in Xpc:
+  X_xpcs.append(x)
+  X_xpcs.append(x)
+  X_xpcs.append(None)
+  Y_xpcs.append(-1)
+  Y_xpcs.append(1)
+  Y_xpcs.append(None)
+
+
+average_waveform, orig_waveforms, norm_waveforms = get_pcs(Xpc, W)
+
+# norm_waveforms = norm_waveforms - average_waveform
+
+
+
+'''============================================================================'''
+
+FONT = dict(
+    family="Latin Modern Roman",
+    color="black",
+    size=13.3333
+  )
+
+
[…]

### 2021-03-23 THESIS.tex
about to change font

@@
 % Tell the reader the problem you are tackling in this project.
-
-Humans are a creative bunch.
+The area of artificial intelligence and, more specifically, neural networks, is interdisciplinar since its conception, being born from biological inspirations.
+Implementation wise, connections with fields like programming, computing and mathematics imposed themselves since the begining.
+
+Attention, however, seems to revolver around the use of advancements in related disciplines to the betterment of neural networks theory in general, be it in the form of convolutional networks, better hyperparameter search, parallel implementations, etc.
+Generally, neural networks are applyied to some problem, with little concern about transforming the problem in order to make it optimally suited to work in a neural networks approach, outside of necessary adaptations, such as normalization.
+
+This lack is in part justified by the recent progress exhibted in the field, where impressive results are abundant; a word of caution is needed here, however, as this is a field that experimented its fair share of hype cycles, followed by winters.
+neural networks faced many winters since its introduction.
+The pattern seems to repeat itself: hardware improvements enable attractive new results, the interest rises, and some technical advancements are made.
+Although, in its current renaissance, the field is still blooming, there are some that are already pointing to a decrease in the research output.
+
+We feel that this tool still offers a great potential in its current state of development; a potential thtat will only be fully explored once it is realised that workflows and auxiliary theories are needed, to be used in conjunctions with neural networks.
+
+In this sense, we introduce a novel representation for pseudo periodic time series, of which sound representation are our primary interest.
+Although much of the theory is drawn from conventional areas related to digital signal processing, image processing, etc, the representation was formulated from the beginning having in mind the strengths and limitations of neural networks in general.
+% State clearly how you aim to deal with this problem.
+
+In this context, the primary objective of this work is to illustrate how such integration can be done, by developing a new sound representation designed specifically to be used by neural networks in sound synthesis tasks.
+
 Since the beginning of computing, even when mainframes were the only available medium, restricted to organizations with considerable budgets, people found ways and motivation to overcome the technical limitations and develop artistic applications, like video games, in their (and the machine's) spare time.
@@
 
-% State clearly how you aim to deal with this problem. 
-% Limit the scope of your study. 
-% Sketch out how the thesis is structured to achieve your aim. 
-% objective
 Ambitious as it may seem, it is the goal of this monograph to motivate a change of approach in domains that, despite their modest intersections, share more attributes than is apparent at a superficial glance: Both machine learning and digital musical acoustics are areas that, despite their current relevance, are fresh out of their respective infancies, and a scrutiny of both promptly reveals the difficulties that arise from this state of affairs.
@@
 
-% Perhaps less daring, and owing more to the advanced age of the author than to his research prowess, is the presentation of this monograph in a format more akin to a book; In addition to turning this somewhat dense tome into a more palatable experience.
+% Limit the scope of your study.
+
+% Sketch out how the thesis is structured to achieve your aim.
+
+% objective
 
@@
 
-with the advance of technology, some classic problems of bibliographic research were mitigates, while new ones were introduced.
-a modern problem encontered by the author during the development of this work is the fragmentation of data.
-In the bygone days of physical research, a paper existed in primarily  physical location.
+With the advance of technology, some classic problems of bibliographic research were mitigated, while new ones were introduced. Fe
+w would argue that the ability to carry thousands of books in a handheld device wasn't an improvement, with accessibility improvements such as the possibility of hearing text, instead of reading it, distribution, etc revolutionizing the consumption, creation of knoledge.
+This revolution was not without its drawbacks, however.
+
+A modern problem encountered by the author during the development of this work is the fragmentation of data.
+In the bygone days of physical research, a paper existed, in the researcher files, in one physical location.
+All the metadata available, perhaps even notes, where in a single logic object, such as a book or a photocopy.
+
+When access to the original document was limited, such as a piece of literature residing in a library, nevertheless the approach could be the same, the logical object now being a card sumarizing the metadata and relevant content.
+This approach is, of course, translatable to the digital world, where one could maintain a folder where digital files, such as pdfs, reside, with notes and the relevant metadata embedded in the file, its filename of even in the metadata section.
+
+This approach quickly becomes problematic, however, for scientific writting, specially in the context of bigger productions such as books and, as is the present case, a thesis, for a number of reasons.
+Digital documents, especially the most common ones, are designed to be as generic as possible, serving variated purposes, as are their metadata systems, if present.
+
+The management of metadata is best acomplished with the used of dedicated software, of which zotero and mendeley are some examples, Being specifically designed to this task, facilities such as metadata retrieval and validation are often available.
+If on intends to make use of a typesetting system, such as Latex, avoiding bibliographic management software is virtually impossible.
+
+Another problem in bibliography management in the digital era is to maintain syncronized the original files, most commonly available in pdf or epub format, and generally trated as general files by most file brownsing interfaces, and its metadata, that must exhib a more sophisticated and uniform structure.
+pdf, despite its ubiquity, is a tricky format, being somewhat opaque.
+Besides, multiple ways exist to encode a similar document.
+Fortunately, we don't need to extract correct or even meaninful text from it, the only thing we need is a way tu uniquely identify each file, impevious to file changes such as the addition of comments, changes in location in the filesystem or renaming.
+
+bibtex uses keys. as tempting as it might sound to use those keys as identifiers fro the files, that approach presents problems.
+
 we adress the meta problem of organizing the bibliography introducing a piece of bibliographic software based on an novel paradigm.
-
-neural networks faced many winters since its introduction.
-The pattern seems to repeat itself: hardware improvements enable attractive new results, the interest rises, and some technical advancements are made.
-Although, in its current renaissance, the field is still blooming, there are some that are already pointing to a decrease in the research output.
-
-We feel that this tool still offers a great potential in its current state of development; a potential thtat will only be fully explored once it is realised that workflows and auxiliary theories are needed, to be used in conjunctions with neural networks.
-
-In this sense, we introduce a novel representation for pseudo periodic time series, of which sound representation are our primary interest.
-Although much of the theory is drawn from conventional areas related to digital signal processing, image processing, etc, the representation was formulated from the beginning having in mind the strengths and limitations of neural networks in general.
 
@@
 \section{Theory}
+
+In this sections we present some topics that will serve as basis for later chapters, from a purely theoretical point of view.
+This is specially important in this work, since many different, seemingly disconnected areas are brought togheter.
 

### 2021-03-23 images/envelope/00ConvexHull.py

@@ -114,5 +114,5 @@ fig.add_trace(
 
 # fig.show(config=dict({'scrollZoom': True}))
 save_name = "./images/" + sys.argv[0].split('/')[-1].replace(".py", ".svg")
-fig.write_image(save_name, width=650, height=300, engine="kaleido", format="svg")
+fig.write_image(save_name, width=605, height=300, engine="kaleido", format="svg")
 print("saved:", save_name)

### 2021-03-23 THESIS.tex
found correct image and font sizes

@@
 
-\usepackage[backend=biber]{biblatex}
+\usepackage[backend=biber,style=authoryear]{biblatex}
+\renewcommand*{\nameyeardelim}{\addcomma\space}
 %\usepackage{biblatex}
@@
 
-\usepackage[no-math]{fontspec}
-\defaultfontfeatures{Ligatures=TeX}
-\setmainfont{Tex Gyre Termes}
+% \usepackage[no-math]{fontspec}
+% \defaultfontfeatures{Ligatures=TeX}
+% \setmainfont{Tex Gyre Termes}
+
+\usepackage{lmodern} % Latin Modern Font
 
@@
 
-\bibliography{00_BIBLI,Compression}
+\bibliography{Envelope,Compression}
 
 \title{Sound Synthesis}
-\author{Carlos Tarjano}
+\author{Carlos Henrique Tarjano Santos}
 
@@
 Effort was made to comply with the most commom nomenclature when terms are encountered frequently in literature.
+
+This work will make extensive use of the term sample, that, even in the context of digital signal processing, can be ambiguous.
+It can refer to a, usually short, sound file, as when refering to sample based instruments, or it can sometimes mean a unique point of a discrete sound wave, as when used in sample rate.
+We will adopt the first meaning, and prefer frame for the second; altough the usage is more comum when referring to videos, it is used in terms such as frames per second.
 
@@
 Consider, for example, the artificially generated discrete wave in figure \ref{fig:Hulls} below, where the links between consecutive samples were deemphasized to highlight the “set of points” nature of a signal: The convex hull and a concave hull, in the form of an alpha shape with a particular value of $ \alpha $ are illustrated.
+The successive points can be interpreted as a wave, in which case the vertical axis represents its instantaneous amplitude at each frame, the meaning of the horizontal axis.
+If we disconsider the lines linking each consecute point it is easy to see the image as a cloud of points in a cartesian plane, however.
 

### 2021-03-23 images/envelope/01AnalyticSignal.py

@@ -41,7 +41,7 @@ hilbert_envelope = np.abs(analytic_signal)
 FONT = dict(
     family="Latin Modern Roman",
     color="black",
-    size=13.3333
+    size=16
   )
 
   
@@ -50,7 +50,7 @@ FONT = dict(
 fig = go.Figure()
 fig.layout.template ="plotly_white" 
 fig.update_layout(
-  xaxis_title="<b><i>i</i></b>",
+  xaxis_title="<b>Frame <i>i</i></b>",
   yaxis_title="<b>Amplitude</b>",
   legend=dict(orientation='h', yanchor='top', xanchor='left', y=1.1),
   margin=dict(l=0, r=0, b=0, t=0),
@@ -104,6 +104,6 @@ fig.add_trace(
 )
 
 # fig.show(config=dict({'scrollZoom': True}))
-save_name = "./images/" + sys.argv[0].split('/')[-1].replace(".py", ".svg")
-fig.write_image(save_name, width=650, height=200, engine="kaleido", format="svg")
+save_name = "./images/envelope/" + sys.argv[0].split('/')[-1].replace(".py", ".svg")
+fig.write_image(save_name, width=605, height=300, format="svg")
 print("saved:", save_name)

### 2021-03-31 images/envelope/02Envelope.py
some progress

@@ -1,7 +1,7 @@
 import sys
-print(sys.version)
+# print(sys.version)
 import plotly.graph_objects as go
-sys.path.append("c:/Users/tesse/Desktop/Files/Dropbox/0_Thesis/Python/01_Envelope")
+# sys.path.append("c:/Users/tesse/Desktop/Files/Dropbox/0_Thesis/Python/01_Envelope")
 import numpy as np
 from scipy.interpolate import interp1d
 from Helper import signal_to_pulses, get_pulses_area
@@ -40,14 +40,14 @@ pulses_X, pulses_Y = get_pulses_area(pulses)
 FONT = dict(
     family="Latin Modern Roman",
     color="black",
-    size=13.3333
+    size=16
   )
 
 '''Plotting'''
 fig = go.Figure()
 fig.layout.template ="plotly_white" 
 fig.update_layout(
-  xaxis_title="<b><i>i</i></b>",
+  xaxis_title="<b>Frame <i>i</i></b>",
   yaxis_title="<b>Amplitude</b>",
   legend=dict(orientation='h', yanchor='top', xanchor='left', y=1.1),
   margin=dict(l=0, r=0, b=0, t=0),
@@ -74,7 +74,7 @@ fig.add_trace(
 
 fig.add_trace(
   go.Scatter(
-    name="Carrier <b>c</b>      ",
+    name="Carrier <b>c</b>",
     # showlegend=False,
     x=X,
     y=C,
@@ -88,7 +88,7 @@ fig.add_trace(
 
 fig.add_trace(
   go.Scatter(
-    name="Wave <b>w = e</b> ⊙ <b>c</b>      ",
+    name="Wave <b>w = e</b> ⊙ <b>c</b>",
     x=X,#pulses_X,
     y=W,#pulses_Y,
     mode="lines+markers",
@@ -100,7 +100,7 @@ fig.add_trace(
 
 fig.add_trace(
   go.Scatter(
-    name="Local Extrema      ",
+    name="Local Extrema",
     x=[p.x for p in pulses][1:],
     y=W[[p.x for p in pulses]][1:],
     mode="markers",
@@ -111,7 +111,7 @@ fig.add_trace(
   )
 )
 
-fig.show(config=dict({'scrollZoom': True}))
-save_name = "./images/" + sys.argv[0].split('/')[-1].replace(".py", ".svg")
-fig.write_image(save_name, width=650, height=350, engine="kaleido", format="svg")
[…]

### 2021-03-31 images/envelope/03Pulses.py

@@ -1,7 +1,7 @@
 import sys
-print(sys.version)
+# print(sys.version)
 import plotly.graph_objects as go
-sys.path.append("c:/Users/tesse/Desktop/Files/Dropbox/0_Thesis/Python/01_Envelope")
+# sys.path.append("c:/Users/tesse/Desktop/Files/Dropbox/0_Thesis/Python/01_Envelope")
 import numpy as np
 from scipy.interpolate import interp1d
 from Helper import signal_to_pulses, get_pulses_area
@@ -50,14 +50,14 @@ for p in pulses[1:-1]:
 FONT = dict(
     family="Latin Modern Roman",
     color="black",
-    size=13.3333
+    size=16
   )
 
 '''Plotting'''
 fig = go.Figure()
 fig.layout.template ="plotly_white" 
 fig.update_layout(
-  xaxis_title="<b><i>i</i></b>",
+  xaxis_title="<b>Frame <i>i</i></b>",
   yaxis_title="<b>Amplitude</b>",
   legend=dict(orientation='h', yanchor='top', xanchor='left', y=1.1),
   margin=dict(l=0, r=0, b=0, t=0),
@@ -88,7 +88,7 @@ fig.add_trace(
 ''' Samples Legend'''
 fig.add_trace(
   go.Scatter(
-    name="Samples      ",
+    name="Samples",
     x=[None],#pulses_X,
     y=[None],#pulses_Y,
     mode="markers",
@@ -100,7 +100,7 @@ fig.add_trace(
 ''' Pulses Legend'''
 fig.add_trace(
   go.Scatter(
-    name="Pulses      ",
+    name="Pulses",
     x=[None],#pulses_X,
     y=[None],#pulses_Y,
     fill="tozeroy",
@@ -116,7 +116,7 @@ fig.add_trace(
 
 fig.add_trace(
   go.Scatter(
-    name="Local Extrema      ",
+    name="Local Extrema",
     # showlegend=False,
     x=Xp,
     y=Yp,
@@ -131,7 +131,7 @@ fig.add_trace(
 
 fig.add_trace(
   go.Scatter(
-    name="Absolute Local Extrema (<i>P</i>)      ",
+    name="Absolute Local Extrema (<i>P</i>)",
     # showlegend=False,
     x=Xp,
     y=np.abs(Yp),
@@ -165,7 +165,7 @@ fig.add_trace(
   )
 )
 
-fig.show(config=dict({'scrollZoom': True}))
[…]

### 2021-04-26 Papers/01_Pulses_Envelope/PulsesEnvelope.tex
Arxiv version
reviewing

@@
 \documentclass{article}
+\usepackage{float}
 \usepackage{lmodern} % Latin Modern Font
@@
 \usepackage{authblk}
+\usepackage{amsmath}
 \usepackage{amsfonts} % \mathbb
+\usepackage{amssymb} % \therefore
 \usepackage[margin=0.5in]{geometry}
@@
 \usepackage[backend=biber]{biblatex}
-\usepackage{amsmath}
 
@@
 
+% TODO only samples already present are part of the envelope; no overshooting
+% TODO Frontiers are also provided
+
 \section{Introduction}
-Envelope detection, also known as demodulation, is ubiquitous in both analogue and digital signal processing \parencite{2011CaetanoImproved}.
-Nevertheless, the literature in this area is fragmented \parencite{2017LyonsDigital}.
-Besides, most envelope detection techniques are designed to account for very specific settings, like pure sinusoids with moderate noise content, a limitation that excludes most physical signals, as is the case of recorded sound, for example; that limitation arises in part due to the lack of a strict mathematical definition of a temporal envelope \parencite{2013Mengempirical}.
+Envelope detection, also known as demodulation, is ubiquitous in both analogue and digital signal processing (DSP) \parencite{2011CaetanoImproved}.
+Nevertheless, the literature in this area is fragmented \parencite{2017LyonsDigital}, with most envelope detection techniques being designed to account for very specific settings, like pure sinusoids with moderate noise content, for example.
+
+Those constraints exclude most physical signals, as is the case of recorded sound, for example; this problem arises in part due to the lack of a strict mathematical definition of a temporal envelope \parencite{2013Mengempirical}.
 
@@
 
-The envelope also helps to impart emotion and identity to the human voice \parencite{2018ZhuContributions} and music: Concert halls with envelope preserving characteristics are deemed more pleasant \parencite{2011LokkiEngaging}, for example.
+The temporal envelope also helps to impart emotion and identity to the human voice \parencite{2018ZhuContributions} and music: Concert halls with envelope preserving characteristics are deemed more pleasant \parencite{2011LokkiEngaging}, for example.
 
-When dealing with broadband signals, approaches tailored to specific applications are prevalent, such as the one presented by \textcite{2014YangFast} for the distributed monitoring of fibre optic or the one formulated by \textcite{2018AssefModeling} in the context of medical ultrasound imaging.
+Specially when dealing with broadband signals, approaches tailored to specific applications, exploiting a priori knowledge of the waves, are prevalent, such as the one presented by \textcite{2014YangFast} for the distributed monitoring of fibre optic or the one formulated by \textcite{2018AssefModeling} in the context of medical ultrasound imaging.
 
-The problem of envelope detection, normally associated with the field of digital signal processing, can be made equivalent to the geometric problem of defining the (outer) shape of a set of points in $ \mathbb{R}^2 $, provided a method for addressing the difference of units in the horizontal, time related, and vertical, amplitude related, axes of a digital signal is available, rendering the original DSP problem readily approachable via geometric techniques.
+The problem of envelope detection of a discrete signal, normally associated with the field of digital signal processing, can be made equivalent to the geometric problem of defining the shape of a set of points in $ \mathbb{R}^2 $.
 
-One direct way of defining the shape of a set of points in $ \mathbb{R}^2 $ is using the concept of a convex hull: as defined in \textcite{2008BergComputational}, the convex hull of a set of points $ S $ in two-dimensional space is the smallest convex subset of the plane that contains all the points in $ S $.
-Convex hulls are unique, in the sense that a set of points define one and only one convex hull.
+Considering the standard 2d representation of a discrete wave, where the horizontal axis represents time and the vertical axis represents the amplitude, one can think of an individual sample as a point.
+Provided a coherent mapping from this space to the standard cartesian plane is available, the original DSP problem can be transformed into a geometric one.
 
-Generalizing that definition, \textcite{1983Edelsbrunnershape} introduced the concept of alpha shapes, that can be seen as a concave hull; a mathematically well-defined generalization of the convex hull of a finite set of points, closely related to the Delaunay triangulation and Voronoi diagrams of those points.
+% TODO
 
-The intuition behind the alpha shapes algorithm is that only points that can be touched by a circle of radius $ \alpha $ coming from an infinite distance towards the set of points of interest are part of the frontier, that is, of the concave hull of this set.
+One direct way of defining the shape of a set of points in $ \mathbb{R}^2 $ is using the concept of a convex hull, as defined in \textcite{2008BergComputational}.
+The convex hull of a set of points $ S $ in two-dimensional space is the smallest convex subset of the plane that contains all the points in $ S $.
+Convex hulls are unique, in the sense that a given set of points define one and only one convex hull.
 
-Once a radius $ \alpha $ is defined, alpha shapes conserve the uniqueness property, in fact tending to the convex hull when $ \alpha \to \infty $.
-Consider, for example, the artificially generated discrete wave in figure \ref{fig:Hulls} below, where the links between consecutive samples were deemphasized to highlight the “set of points” nature of a signal: The convex hull and a concave hull, in the form of an alpha shape with a particular value of $ \alpha $ are illustrated.
+When the general shape of a set of points present prominent concavities, the convex hull overestimates the boundaries.
+To address this limitation \textcite{1983Edelsbrunnershape} introduced the concept of alpha shapes, that became known as the concave hull: a mathematically well-defined generalization of the convex hull of a finite set of points, closely related to the Delaunay triangulation and Voronoi diagrams of those points.
 
-\begin{figure}[ht!]
+The intuition behind the alpha shapes algorithm is that only points that can be touched by a circle of radius $ \alpha $ coming from an infinite distance towards the baricenter of the set of points are part of the frontier.
+The polygon obtained linking those points forms the concave hull of this set.
+
+Once a radius $ \alpha $ is defined, alpha shapes exhibit the uniqueness property meaning that, for a given set of points and a given radius, only one concave hull exists.
+When the radius of the circle $ \alpha \to \infty $, concave hulls tend to the convex hulls.
+
+Consider, for example, the artificially generated discrete wave in Fig.
+\ref{fig:Hulls} below, where the links between consecutive samples were deemphasized to highlight the "set of points" nature of a signal: The convex hull and a concave hull, in the form of an alpha shape with a particular value of $ \alpha $ are illustrated.
+
+\begin{figure}[H]
   \centering
@@
 
-Following this path, we present a general approach to envelope detection, exploiting the intrinsic characteristics of a generic, spectrally complex wave, in order to completely avoid the need for manual intervention or parameter tuning.
+Following this path, we present a general approach to envelope detection, exploiting the intrinsic characteristics of a spectrally complex wave, in order to completely avoid the need for manual intervention or parameter tuning.
+
+% TODO %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+After defining what will be considered an envelope in the context of this work, in this section we explore some simplifications that can be applied to a discrete wave, defining the representation that will be used in the subsequent part of the work.
+
+We next explore the concept of discrete curvature, putting forth our own method, tailored to the definition of the equivalent circle of a wave; we end this section by showing how this circle can be used to identify the envelope of a discrete wave, in a similar approach to that of the alpha shapes theory, but taking advantage of some unique structure present in discrete waves.
 
@@
 We follow with suggestions about how the approach here presented can be useful beyond envelope detection, exemplifying with some applications, such as the extractions of the average waveform of a digital wave, and concluding with commentary on potential future developments of the method.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\section{Methods}
-After defining what will be considered an envelope in the context of this work, in this section we explore some simplifications that can be applied to a discrete wave, defining the representation that will be used in the subsequent part of the work.
+\section{Temporal envelope}\label{sec:Characterization of a temporal envelope}
+The interest in the theory of envelope detection arose in the analog domain, with the widespread adoption of radio communications, where the problem is typically restricted to artificial waves, often with well-defined frequencies \parencite{2011TurnerDemodulation}.
 
-We next explore the concept of discrete curvature, putting forth our own method, tailored to the definition of the equivalent circle of a wave; we end this section by showing how this circle can be used to identify the envelope of a discrete wave, in a similar approach to that of the alpha shapes theory, but taking advantage of some unique structure present in discrete waves.
-
-\subsection{Characterization of a temporal envelope}
-
-The interest in the theory of envelope detection arose in the analog domain, with the widespread adoption of radio communications, where the problem is typically restricted to simple sinusoids, often with well-defined frequencies \parencite{2011TurnerDemodulation}.
+% TODO vague definition of envelope - smoothness and slow varying
 
@@
 In his work, \textcite{1946GaborTheory} applies the then relatively new mathematical machinery of the quantum mechanics to unify the time and frequency-domain representations of a wave, showing how the Hilbert transform could be applied to a real signal in order to obtain an equivalent complex signal, later known as the analytic signal.
+
 This analytic signal has the form $ A(t) = S(t) + \mathbb{H}(S(t)) \ \text{i} $ \parencite{2016HePraat} where $ S(t) $ is the original real signal.
@@
 
-Figure \ref{fig:AnalyticSignal} exemplifies the Hilbert envelope, for a pure sinusoid modulated by a polynomial, illustrating the difference in the shape of the envelope in the presence and absence of Gaussian white noise.
-It is readily noticeable from the figure that, as soon as noise is introduced in the original signal, it is reflected in the envelope.
+Fig.
+\ref{fig:AnalyticSignal} exemplifies the Hilbert envelope, for a pure sinusoid modulated by a polynomial, illustrating the difference in the smoothness of the envelope in the presence and absence of Gaussian white noise.
+It is readily noticeable from the Fig.
+\ref{fig:AnalyticSignal} that, as soon as noise is introduced in the original signal, it is directly reflected in the envelope.
 
-\begin{figure}[ht!]
+\begin{figure}[H]
   \centering
@@
 The preceding definitions can be translated to this discrete scenario assuming that the discrete quantities arise from observing the continuous ones at regular time intervals.
-In that case, the equality $ i = t \ \text{fps} $ can be used to link both settings, where $ i $ is the index of each observation, $ t $ stands for the time in seconds, and fps is the frame rate, or the number of observations made in the period of a second.
+In that case, the equality $ i = t \ \text{fps} $ can be used to link both settings, where $ i $ is the index of each observation, $ t $ stands for the time in seconds, and fps is the frame rate, or the number of observations made in one second.
 We can thus define:
 
-\begin{align} \label{eq:Envelope}
+\begin{subequations}\label{eq:Envelope}
+\begin{align}
   & \mathbf{w}, \mathbf{e}, \mathbf{c} \in \mathbb{R}^n \\
@@
 \end{align}
+\end{subequations}
 
 The $ \odot $ operator in \ref{eq:Envelope} stands for the Hadamard product, denoting element-wise multiplication of two vectors.
-Figure \ref{fig:Envelope} provides an example of the vectors just defined: \textbf{w} is obtained by modulating the normalized carrier wave \textbf{c} using the envelope \textbf{e}.
+Fig.
+\ref{fig:Envelope} provides an example of the vectors just defined: \textbf{w} is obtained by modulating the normalized carrier wave \textbf{c} using the envelope \textbf{e}.
 
-Note from Figure \ref{fig:Envelope} that, at some local maxima, the envelope is well-defined, being in fact equal to the value of the wave $ w_i $, since $ c_i = 1 $ at those points, rendering $ e_i = 1 w_i $.
+Note from Fig.
+\ref{fig:Envelope} that, at some local maxima, the envelope is well-defined, being in fact equal to the value of the wave $ w_i $.
+$ c_i = 1 $ at those points, rendering $ e_i = 1 w_i $.
 
-The carrier wave \textbf{c} in the figure is periodic, and normalized between $ \{-1, 1\}$.
-Each cycle of the wave, then, reaches the value 1 only at its maximum making $ w_i $ equal to the value of the envelope at one point in each cycle.
-
-Image \ref{fig:Envelope}, along with the concave hull illustrated in figure \ref{fig:Hulls}, helps forming the intuition behind the method developed in this work: the idea is to identify the local extrema that touch the envelope, those being also the extrema of each (pseudo-)cycle.
-
-This is accomplished with the use of a circle with a defined radius, large enough to prevent it from reaching the smallest local extrema when "rolled" around the (geometric) representation of the wave.
-Only the extrema touched by the circle will be marked as part of the envelope.
-
-In the next section, a compatible geometric representation will be formulated, along with the necessary tools to estimate the average curvature of the wave and, consequently, the appropriate radius of the circle.
-
-\begin{figure}[ht!]
+\begin{figure}[H]
   \centering
@@
 
-\subsection{Simplified Representation}
+% TODO
+This realization that the envelope of a wave is defined only at some of the wave's local extrema is crucial, as we can use it to generate a simplified representation of a signal that will make the algorithm more efficient by considering only the local extrema of a given signal.
 
-Instrumental to the method here presented is the definition of a pulse: each time the value of \textbf{w} changes sign, that is, every time the discrete wave \textbf{w} crosses the horizontal axis, the beginning of a pulse is defined, with the next crossing defining its end.
+\section{Simplified Representation}
+
+We start with the definition of a pulse: each time the value of $ w_i$ changes sign, that is, every time the amplitude of the discrete wave \textbf{w} crosses the horizontal axis, the beginning of a pulse is defined, with the next crossing defining its end.
 This definition is in line with the one presented in \textcite{national1996telecommunications}, where a pulse is defined as a rapid change in the amplitude of a signal, followed by a fast return to the baseline value; zero, in our case.
 
-From \ref{fig:Envelope} and the discussion in the preceding chapter, we can see that only the maximum point of a positive pulse or the minimum point of a negative one can potentially be equal to the envelope and are, thus, our only points of interest.
+From \ref{fig:Envelope} and the discussion in Section \ref{sec:Characterization of a temporal envelope}, we can see that only the maximum point of a positive pulse or the minimum point of a negative pulse can potentially be equal to the envelope and are, thus, our only points of interest.
 We can then proceed, for the rest of the method, considering only those points, to great computational economy.
@@
 
-This relation is illustrated in figure \ref{fig:Pulses}. We call it P to emphasize the connection with the pulses of \textbf{w}.
+This relation is illustrated in Fig. \ref{fig:Pulses}. We call it P to emphasize the connection with the pulses of \textbf{w}.
+While the exact relation between $ n $, the size of the original signal, and $ m $, the size of the set of points, is dependent on factors such as the frequency and sampling rate of the original wave, the inequality $ m \le n $ holds for all discrete signals.
+In practice, $ m \ll n $ for most real world discrete signals.
 
-\begin{figure}[ht!]
+\begin{figure}[H]
   \centering
@@
 
-The points in P are not fit for geometric interpretation yet, since the abscissa and ordinate of their orthogonal coordinate system have different units: in the vertical axis, we have a unit related to the instantaneous intensity of the wave, while in the horizontal axis we have the index at which each extremum occurs, ultimately a time-related unit.
+\section{Mapping to Cartesian coordinates}
 
-Specifically, this coordinate system has a basis vector $ \mathbf{b} = \{ \mathbf{b}_x, \mathbf{b}_y \}$ with $ \mathbf{b}_x = (a, 0), \mathbf{b}_y = (0, i) $ where $ a $ is an amplitude unit such as decibel, a measure of the instantaneous sound pressure, volt or even a fraction of a fixed maximum value.
+The points in P are not fit for geometric interpretation yet, since the abscissa and ordinate of their orthogonal coordinate system have different units: in the vertical axis, we have a unit related to the instantaneous amplitude of the wave, while in the horizontal axis we have the index at which each extremum occurs, ultimately a time-related unit.
+
+Geometrically, this coordinate system has a basis vector $ \mathbf{b} = \{ \mathbf{b}_x, \mathbf{b}_y \}$ with $ \mathbf{b}_x = (a, 0), \mathbf{b}_y = (0, i) $ where $ a $ is an amplitude unit such as decibel, a measure of the instantaneous sound pressure, volt or even a fraction of a fixed maximum value.
 $ i $ represents the index of the sample, and is linked to time by the aforementioned relation $ i = t \ \text{fps} $.
 
-We are interested in finding a coordinate system where a pulse's amplitude $ \lvert w_j \lvert $ and length would define, on average, a square.
+We are interested in finding a new basis vector that would make the geometric representation consistent, in the sense that pulse's amplitude $ \lvert w_j \lvert $, now mapped to the ordinates of a point, and a pulse's length, approximately equivalent to the horizontal distance between two consecutive points in the geometrical interpretation would define, on average, a square.
 
@@
 
-In this new coordinate system we have a basis vector $ \mathbf{b}^\prime = \{ \mathbf{b}^\prime_x, \mathbf{b}^\prime_y \}$ where $ \mathbf{b}^\prime_x = (i, 0) $ and $\mathbf{b}^\prime_y = (0, i) $, related to the original basis vector by equation \ref{eq:Basis}:
+In this new coordinate system we have a basis vector $ \mathbf{b}^\prime = \{ \mathbf{b}^\prime_x, \mathbf{b}^\prime_y \}$ where $ \mathbf{b}^\prime_x = (i, 0) $ and $\mathbf{b}^\prime_y = (0, i) $, related to the original basis vector by Eq.
+\ref{eq:Basis}:
 
-\begin{align} \label{eq:Basis}
+\begin{subequations}  \label{eq:Basis} % TODO
+\begin{align}
 \mathbf{b}^\prime_x &= \mathbf{b}_x  \\
@@
 \end{align}
+\end{subequations}
 
-The effect of this normalization can be seen in the points shown in figure \ref{fig:Vectors}, shown in scale in both coordinate systems.
+The effect of this normalization can be seen in the points shown in Fig.\ref{fig:Vectors}, shown in scale in both coordinate systems.
+
+\section{Discrete Curvature Estimation}
+
+We have now a map from the original space of the two-dimensional representation of a discrete wave to the Cartesian coordinate system and thus a way to consistently transform a discrete signal into a set of points in the plane.
+
+Translating the explanation of the alpha shapes algorithm presented in \textcite{1994EdelsbrunnerThree} to our context, we can say that the points that are part of the envelope are those touched by a circle of a given radius $ r $, coming from $ y \to \infty $ towards the Cartesian representation of the signal, that is not allowed to contain any point of the signal; Intuitively, one can picture a circle being rolled above the signal, and marking the points it touches as envelope points.
+
+To employ the alpha shapes algorithm in this set of points, however, the radius $ r $ of such circle must be estimated, and for that a measure of discrete curvature is needed.
+
 The set V of the vectors from each point of P to the next is also shown in the picture, and it will be used to estimate the average curvature of P.
-We define this set in equation \ref{eq:Vectors} below.
+We define this set in Eq.
+\ref{eq:Vectors} below.
+% TODO
 
@@
 
-\begin{figure}[ht!]
+\begin{figure}[H]
   \centering
@@
 
-\subsection{Discrete Curvature Estimation}
-
-Translating the explanation of alpha shapes in \textcite{1994EdelsbrunnerThree} to our context, we can say that the points that are part of the envelope are those touched by a circle of a given radius $ r $, coming from $ y \to \infty $ towards the signal, that is not allowed to contain any point of the signal; Intuitively, one can picture a circle being rolled above the signal, and marking the points it touches as envelope points.
-
-To infer the appropriate radius $ r $ of such circle a measure of discrete curvature is needed.
 Discrete curvature estimation is an important task in image processing \parencite{2010FleischmannNovel} for which no default definition exists.
 
-The two possible general approaches are the derivation of direct methods, that use characteristics of the discrete wave to calculate the curvature, or the use of the curvature of a smooth, continuous curve fitted to the discrete wave \parencite{2001CoeurjollyDiscrete}.
+The two general approaches are the derivation of direct methods that use characteristics of the discrete curve to calculate the curvature, or the use of the curvature of a smooth, continuous curve fitted to the original discrete curve \parencite{2001CoeurjollyDiscrete}.
 
@@
 
-Those approaches define the discrete curvature in relation to the vertices of a wave (or mesh, in the three-dimensional case).
+Those approaches define the discrete curvature in relation to the vertices of a discrete curve (or mesh, in the three-dimensional case).
 In our particular case, it wouldn't make sense to talk about the curvature of a single pulse, since it is what our points ultimately represent.
@@
 
-We thus proceed to define a discrete curvature measure over the edges of a discrete wave.
+We thus proceed to define a discrete curvature measure over the edges of a discrete curve.
 To that end we are going to apply the definition of smooth curvature as the rate of change of the unit tangent to a curve, noting that this is equivalent to that of the osculating circle \parencite{2016VasaMesh}.
 
-\subsubsection{The Equivalent Circle Approach}
+\subsection{The Equivalent Circle Approach}
 
-The rationale is to find, for each vector $ v_k \in \text{V} $, the radius $ r_k $ of the equivalent circle whose tangent has the same change in direction, in the same horizontal distance, as the vector of interest, as shown in figure \ref{fig:DiscreteCurvature}.
+The rationale is to find, for each vector $ v_k \in \text{V} $, the radius $ r_k $ of the equivalent circle whose tangent has the same change in direction, in the same horizontal distance, as the vector of interest, as shown in Fig.
+\ref{fig:DiscreteCurvature}.
 The average curvature of P will then be obtained by the average of all radii $ r_k $.
 
-From figure \ref{fig:DiscreteCurvature} is easy to see that $ r_k = v_{k,x} \sin(\theta_k)$ where $ v_{k,x} $ is the component of $ v_k $ in the horizontal direction; $ \theta_k $ itself can be obtained using the slope $ m_0 = 0 $ of a line in the horizontal direction and the slope $ m_k = v_{k,y} / v_{k,x} $ defined by $ v_k $ via the equality $ \theta_k = \arctan \big( (m_k - m_0) / (1 + m_k m_0) \big) $.
+From Fig.
+\ref{fig:DiscreteCurvature} is easy to see that $ r_k = v_{k,x} / \sin(\theta_k) $ where $ v_{k,x} $ is the component of $ v_k $ in the horizontal direction; $ \theta_k $ itself can be obtained using the slope $ m_0 = 0 $ of a line in the horizontal direction and the slope $ m_k = v_{k,y} / v_{k,x} $ defined by $ v_k $ via the equality $ \theta_k = \arctan \big( (m_k - m_0) / (1 + m_k m_0) \big) $.
 
-\begin{figure}[ht!]
+\begin{figure}[H]
   \centering
@@
 
-The radius of the circle that represents the average curvature of V (and P), and will be used to obtain the envelope of the discrete wave \textbf{w} can be obtained via equation \ref{eq:Radius}:
+The radius of the circle that represents the average curvature of V (and, consequently, of P), and will be used to obtain the envelope of the discrete wave \textbf{w} can be obtained via Eq.
+\ref{eq:Radius}:
 
@@
 We need now to construct an algorithm, using the radius just obtained, to identify the points that belong to the envelope.
-Generally, the algorithm for the alpha shapes approach resorts to the construction of the Delaunay triangulation of the set of points, that is latter filtered to contain only the outer edges.
-We can adopt a more straightforward method, as our points have a defined structure:
+Generally, alpha shapes based algorithms resort to the construction of the Delaunay triangulation of the set of points, that is latter filtered to contain only the outer edges.
+We can adopt a more straightforward method, as our points have a sequential structure.
+Although that is the case throughout this work, the original signal doesn't need to be sampled in equal spaces.
 
@@
 For all circles of center $ x_o, y_o $ and radius $ r $ that can be constructed to pass through any two different points of P, that is, $ \forall \quad P_b, P_c $ such that $ (P_{b,x} - x_o)^2 + (P_{b,y} - y_o)^2 = (P_{c,x} - x_o)^2 + (P_{c,y} - y_o)^2 = r^2 $, $ P_a $ will be part of the set $ E $ of the points belonging to the envelope if and only if none of those circles contain $ P_a $, or $ P_a \in \text{E} \leftrightarrow (P_{b,x} - x_o)^2 + (P_{b,y} - y_o)^2 > r^2 $.
-Summarizing this, we have equation \ref{eq:InEnvelope}:
+Summarizing this, we have Eq.
+\ref{eq:InEnvelope}:
 
-\begin{align} \label{eq:InEnvelope}
+\begin{subequations} \label{eq:InEnvelope}
+\begin{align}
 & P_a \in \text{E} \leftrightarrow (P_{b,x} - x_o)^2 + (P_{b,y} - y_o)^2 > r^2, \\
@@
 \end{align}
+\end{subequations}
 
@@
 
-\section{Results and Discussion}
+\section{Quality of an envelope}
 Given the lack of a precise definition of what an envelope should be, we begin this section suggesting an empirical metric, based on the behaviour of the carrier wave \textbf{c}.
@@
 
-\subsection{Quality of an envelope}
-
 There are no agreed-upon objective metrics to assess the quality of an envelope.
-Recalling figure \ref{fig:Envelope}, however, and the accompanying discussion, an intuitive way is to consider the carrier wave \textbf{c}, and how well it seems to be bounded by the interval $ \{-1, 1\} $.
+Recalling Fig.
+\ref{fig:Envelope}, however, and the accompanying discussion, an intuitive way is to consider the carrier wave \textbf{c}, and how well it is bounded by the interval $ \{-1, 1\} $.
 
@@
 
-From figure \ref{fig:spoken_voice} it can be seen that the algorithm performs well under the carrier metric.
+From Fig.
+\ref{fig:spoken_voice} it can be seen that the algorithm performs well under the carrier metric.
 This is especially remarkable for the case of spoken voice, a notorious complex wave with fast changes in pitch and high noise content.
 
-These characteristics are in stark contrast with those of the sang voice in figure \ref{fig:singer_voice}, where a recording of an alto singer is shown; there, the wave is almost periodic, with a stable waveform.
+These characteristics are in stark contrast with those of the sang voice in Fig.
+\ref{fig:singer_voice}, where a recording of an alto singer is shown; there, the wave is almost periodic, with a stable waveform.
 
-\begin{figure}[ht!]
-  \centering
-    \includegraphics{images/06Graph_bend.pdf}
-  \caption{Signal, envelope and carrier for a record of a bend performed on an electric guitar.}
-  \label{fig:Bend}
-\end{figure}
-
-\begin{figure}[ht!]
+\begin{figure}[H]
   \centering
@@
   \label{fig:singer_voice}
+\end{figure}
+
+Fig. \ref{fig:Bend} shows the envelope in the case of a bend performed in an eletric guitar. The variations in frequency in the signal didn't affect the envelope extracted.
+
+\begin{figure}[H]
+  \centering
+    \includegraphics{images/06Graph_bend.pdf}
+  \caption{Signal, envelope and carrier for a record of a bend performed on an electric guitar.}
+  \label{fig:Bend}
 \end{figure}
@@
 
-Nevertheless, if we picture both the wave and the envelope as samples from continuous functions, is easy to see that this is the case for our algorithm; since the envelope is composed by some samples of the digital wave, in the limit where all the samples of the wave are also part of the envelope, we have that the maximum frequency for both is the same.
+Nevertheless, if we picture both the wave and the envelope as samples from continuous functions, is easy to see that this is the case for our algorithm; since the envelope is composed by a subset of the samples of the original digital signal, in the limit where all the samples of the wave are also part of the envelope, we have that the maximum frequency for both is the same.
 
@@
 Since all the local maxima have the same amplitude in the case of a periodic wave, our envelope will indeed be a straight line, with the same amplitude as the original wave.
-An example of such behaviour can be observed in figure \ref{fig:Comparison_sinusoid}.
+An example of such behaviour can be observed in Fig.
+\ref{fig:Comparison_sinusoid}.
 
@@
 
-\subsection{Comparison with traditional algorithms}
+\section{Comparison with traditional algorithms}
 
@@
 
-For a recording of the key 33 of a grand piano, where parts of the sound are highly percussive and noisy, the Hilbert approach presents considerable oscillations, as can be seen in figure \ref{fig:Comparison_piano}.
+For a recording of the key 33 of a grand piano, where parts of the sound are highly percussive and noisy, the Hilbert approach presents considerable oscillations, as can be seen in Fig.
+\ref{fig:Comparison_piano}.
 
-\begin{figure}[ht!]
+\begin{figure}[H]
   \centering
@@
 
-In figure \ref{fig:Comparison_soprano} all envelopes exhibit a similar movement along the samples, but the ones generated by traditional algorithms are undershooting the original wave.
+In Fig. \ref{fig:Comparison_soprano} all envelopes exhibit a similar movement along the samples, but the ones generated by traditional algorithms are undershooting the original wave.
 
-\begin{figure}[ht!]
+\begin{figure}[H]
   \centering
@@
 
-\begin{table}[ht!]
+\begin{table}[H]
 \centering
@@
 
-\subsection{Frontiers}
+\section{Superior and inferior envelopes - Frontiers}
 
-In practice, is not uncommon for a discrete wave, especially in the case of sound, to present somewhat different positive and negative contours; in those cases, the algorithm here presented can be used independently in the positive and negative pulses of the wave to define two “envelopes”, that we are going to call superior and inferior frontiers.
+In practice, is not uncommon for a discrete wave, especially in the case of sound, to present somewhat different superior and inferior contours; in those cases, the algorithm here presented can be used independently in the positive and negative pulses defined by the original wave to define two “envelopes”, that we are going to call superior and inferior frontiers.
 
-Figure \ref{fig:FullFrontiers} illustrates the frontiers of six diverse discrete sound waves, as well as an in-detail view of the highlighted segment for each signal.
+Fig.
+\ref{fig:FullFrontiers} illustrates the frontiers of six diverse discrete sound waves, as well as an in-detail view of the highlighted segment for each signal.
 All waves are records of physical sounds, chosen to represent the applicability of the algorithm in real-world scenarios.
 
-\begin{figure}[ht!]
+\begin{figure}[H]
   \centering
@@
 
-\subsection{Approximated location of pseudo-cycles}
+\section{Approximated location of pseudo-cycles}
 
-The envelope is shown to add complexity to the spectral representation of a wave \parencite{2019TarjanoNeuro}, and an accurate description of the envelope, while describing the evolution of the instantaneous amplitude of a signal in time, would also simplify further spectral analysis.
-Figure \ref{fig:Fourier} shows the frequency-domain power spectrum for the wave and the carrier presented in \ref{fig:Envelope}.
+% TODO
+
+The temporal envelope adds complexity to the spectral representation of a wave \parencite{2019TarjanoNeuro}, and an accurate description of the envelope, while describing the evolution of the instantaneous amplitude of a signal in time, would also simplify further spectral analysis.
+Fig.
+\ref{fig:Fourier} shows the frequency-domain power spectrum for the wave and the carrier presented in \ref{fig:Envelope}.
 For the carrier, the power spectrum presents two nonzero values, while the power spectrum for the wave composed of the carrier modulated by the envelope is more complex.
 
-\begin{figure}[ht!]
+\begin{figure}[H]
   \centering
@@
 
-One can see in figure \ref{fig:PCs} a section of the plot of where the local maxima and minima occurs in the recording of an alto singer show in \ref{fig:singer_voice}.
+One can see in Fig.
+\ref{fig:PCs} a section of the plot of where the local maxima and minima occurs in the recording of an alto singer show in \ref{fig:singer_voice}.
 The average between the two is also shown.
-With this information, one can split a discrete wave into its pseudo-cycles, as is illustrated in Figure \ref{fig:AvgPc}: It shows the superposition of this extracted pseudo-cycles, after being normalized by length, and their average.
+With this information, one can split a discrete wave into its pseudo-cycles, as is illustrated in Fig.
+\ref{fig:AvgPc}: It shows the superposition of this extracted pseudo-cycles, after being normalized by length, and their average.
 This same average is shown shifted, to exemplify that approximate $C^0$ and $C^1$ continuity is achieved.
 
-\begin{figure}[ht!]
+\begin{figure}[H]
   \centering

### 2021-04-26 Python/02_Representation_old_2/LMS_tests.py

@@ -1,170 +1,0 @@
-import numpy as np
-import plotly.graph_objects as go
-
-m = 5000
-k = 3
-X = np.arange(m)
-Y = np.cos(X/100) + np.random.randn(m)/5 # b
-
-
-'''Least Squares'''
-
-Q = np.zeros((m, k + 1))
-
-for i in range(m):
-  for j in range(k + 1):
-    Q[i, j] = i**j
-
-A = np.linalg.inv(Q.T @ Q) @ Q.T @ Y
-
-A_ = np.linalg.lstsq(Q, Y)[0]
-
-# print(A)
-# print(A_)
-
-print(np.allclose(A, A_))
-# exit()
-
-Y_lms = np.zeros((m))
-
-for i in range(m):
-  for j in range(k + 1):
-    Y_lms[i] += A[j] * i**j
-
-
-
-'''Constrained Least Squares (with q intervals and k-1 polynomial)'''
-s = 10
-l = m // s
-Q = np.zeros((m, k * s)) # A
-V = np.zeros((2 * (s-1), k * s)) # C
-
-'''populating Q'''
-for i1 in range(s):
-  for i2 in range(l * i1, l * (i1 + 1)):
-    for i3 in range(k):
-      Q[i2, i1 * k + i3] = X[i2]**i3
-
-'''populating V (constraints)'''
-for i in range(s - 1):
-  for j in range(k):
-    V[2 * i, i * k + j] = X[(i + 1) * l]**j
-    V[2 * i, i * k + k + j] = -X[(i + 1) * l]**j
-for i in range(s - 1):
-  for j in range(1, k):
-    V[2 * i + 1, i * k + j] = j * X[(i + 1) * l] ** (j - 1)
-    V[2 * i + 1, i * k + k + j] = -j * X[(i + 1) * l] ** (j - 1)
-
-
-'''solving systems'''
-a1 = np.hstack([                         # cols           lines
-      np.zeros(((k * s, k * s))),        # k * s          k*s
-      Q.T,                               # m
-      V.T                                # 2 * (s-1)
-    ])
-
-a2 = np.hstack([
-    Q,                                   # k * s          m
-    -1/2 * np.identity(m),               # m
-    np.zeros((m, 2 * (s-1)))             # 2 * (s-1)
-  ])
-
-a3 = np.hstack([
-    V,                                   # k * s          2 * (s-1)
-    np.zeros((2 * (s-1), m + 2 * (s-1))) # m + 2 * (s-1)
-  ])
-
-
[…]

### 2021-04-26 Papers/01_Pulses_Envelope/PulsesEnvelope.tex
reviewing

@@
 \usepackage{graphicx}
-\usepackage[backend=biber]{biblatex}
+\usepackage[backend=biber, sorting=none]{biblatex}
 
@@
 Despite being an elusive concept, the temporal amplitude envelope of a signal is essential for its complete characterization, being the primary information-carrying medium in spoken voice and telecommunications, for example.
-Intuitively, the temporal envelope can be understood as a slowly varying function that multiplies the signal, being responsible for its outer shape.
 Envelope detection techniques have applications in areas like medicine, sound classification and synthesis, seismology and speech recognition.
 % Problem | Objectives
-Nevertheless, a general approach to digital envelope detection of signals with rich spectral content doesn't exist, as most methods involve manual intervention, in the form of filter design, smoothing, and other specific design choices, based on prior knowledge of the signals under investigation.
-To address this problem, we propose an algorithm that uses intrinsic characteristics of a signal to estimate its envelope, eliminating the necessity of parameter tuning.
+Nevertheless, a general approach to envelope detection doesn't exist, as most methods involve manual intervention, in the form of filter design, smoothing, and other specific design choices based on prior knowledge of the signals under investigation.
+To address this problem, we propose an algorithm that uses characteristics of a signal to estimate its envelope, eliminating the necessity of parameter tuning.
 % Methods | Procedure | Approach
-The approach here described draws inspiration from geometric concepts to estimate the temporal envelope of an arbitrary signal; specifically, a new measure of discrete curvature is used to obtain the average radius of curvature of a discrete wave, that will serve as a threshold to identify the wave's samples that are part of the envelope.
+The method draws inspiration from geometric concepts; the discrete signal is mapped to a set of points in the Cartesian plane, were a new measure of discrete curvature is used to identify the samples that are part of the envelope.
+Due to its geometric nature, the method can optionally generate the superior and inferior envelopes of a given wave.
 % Results | Findings | Product
 The algorithm compares favourably with classic envelope detection techniques based on smoothing, filtering and the Hilbert Transform, besides being physically plausible.
-We provide visualizations of the envelope extracted via the algorithm for various real-world signals, with very diverse characteristics, such as voice, spoken and sang, and pitched and non-pitched musical instruments, and discuss some approaches to assess the quality of the obtained envelopes.
-A Python module implementing the algorithm was made available via the Python Package Index (PyPI); interactive visualizations of envelopes for a diverse range of digital waves, as well as the source code for the Python implementation, are available online at \textcite{2020TarjanoEnvelope}.
-% Conclusions | Implications
-Besides the most direct applications of this work to audio classification and synthesis, we foresee impact in compression techniques and machine learning approaches to audio.
-We briefly discuss some potential paths in this direction.
-The discrete curvature definition presented could also be extended to three-dimensional settings, to improve shape detection algorithms based on alpha shapes.
+We provide visualizations of the envelope extracted for various real-world signals, with very diverse characteristics, such as voice, spoken and sang, and pitched and non-pitched musical instruments, and discuss some approaches to assess the quality of the obtained envelopes.
+An optimized implementation is available via the Python Package Index (PyPI); interactive visualizations and source code are available at the accompanying repository.
 \end{abstract}
 
-{\bf Keywords:} DSP; alpha shapes; envelope detection; discrete curvature estimation; demodulation
-
-% TODO only samples already present are part of the envelope; no overshooting
-% TODO Frontiers are also provided
+{\bf Keywords:} alpha shapes; envelope detection; discrete curvature estimation; demodulation
 
@@
 
-The procedure is formalized in the algorithm \ref{FindEnvelope}.
-We don't provide similar algorithmic descriptions of the preceding steps since the Python implementation can easily serve as pseudocode.
+The procedure is formalized in the algorithm \ref{FindEnvelope}. We don't provide similar algorithmic descriptions of the preceding steps since the Python implementation available at % TODO
+ can easily serve as pseudocode.
 
@@
 
-The set E obtained via algorithm \ref{FindEnvelope} is a set of points.
+The set E obtained via algorithm \ref{FindEnvelope} is a subset of P, and thus also a set of points.
 The most straightforward way to transform those points into the vector $ \textbf{e} \in \mathbb{R}^n $ is via linear interpolation, the approach used in this work.
@@
 \section{Quality of an envelope}
-Given the lack of a precise definition of what an envelope should be, we begin this section suggesting an empirical metric, based on the behaviour of the carrier wave \textbf{c}.
+Given the lack of a precise definition of what an envelope is, we begin this section suggesting an empirical metric, based on the behaviour of the carrier wave \textbf{c}.
 We also comment on some metrics presented in the literature, and compare our algorithm with traditional demodulation approaches, both in relation to execution time and accuracy; to this last end, a numerical indicator is also introduced.

### 2021-04-26 Papers/01_Pulses_Envelope/PulsesEnvelope.tex
progress with the revision

@@
 \section{Temporal envelope}\label{sec:Characterization of a temporal envelope}
+
 The interest in the theory of envelope detection arose in the analog domain, with the widespread adoption of radio communications, where the problem is typically restricted to artificial waves, often with well-defined frequencies \parencite{2011TurnerDemodulation}.
 
-% TODO vague definition of envelope - smoothness and slow varying
+While there is no general formal definition of the envelope of broadband signals \parencite{2014Yangnovel,2015Yangtheoretical,2019Jiaempiricalj}, the general intuition is that an envelope is a smooth curve situated as close as possible to the signal \parencite{2019Jiaempiricalj}.
 
@@
 
-Despite its widespread use, and effectiveness in the context of narrowband signals, envelope detection techniques based on the Hilbert transform don't behave well for broadband signals \parencite{2012DauSpeech}, being even physically paradoxical in some cases \parencite{1996Loughlinamplitude}.
+Despite its widespread use, and effectiveness in the context of narrowband signals, envelope detection techniques based on the Hilbert transform don't behave well for broadband signals \parencite{2012DauSpeech, 2019Jiaempiricalj}, being even physically paradoxical in some cases \parencite{1996Loughlinamplitude, 2014Yangnovel}.
 
@@
 
+For a real world signal, such as an alto singer sustaining a steady note as shown in Fig.
+\ref{fig:AnalyticSignalAlto} the Hilbert transform alone provides a poor envelope, since it retains most part of the frequencies of the underlying wave.
+With further processing, however, such as filtering, an output closer to the intuitive definition of envelope given in the beginning of this section can be obtained.
+
+\begin{figure}[H]
+  \centering
+    \includegraphics{images/01AnalyticSignalAlto.pdf.pdf}
+  \caption{Representation of an alto singer sustaining a steady note and a comparison of the envelopes obtained by the Hilbert transform with and without further filtering.}
+  \label{fig:AnalyticSignalAlto}
+\end{figure}
+
 In general, the problem of envelope detection, or demodulation, can be interpreted as the task of, given a continuous wave $ W(t) $, decomposing it in two components $ E(t) $ and $ C(t) $, such that $ W(t) = E(t) C(t) $ \parencite{2011TurnerDemodulation}.
 
-$ E(t) $ represents the slow varying part of the wave, also known as the (temporal) envelope, modulator component, or amplitude modulation (AM) of $ W $, while $ C(t) $ models its fast varying part, called throughout the literature as its (temporal) fine structure, carrier component, or frequency modulation (FM).
+$ E(t) $ represents the slow varying part of the wave, also known as the (temporal) envelope, modulator component, or amplitude modulation (AM) of $ W(t) $, while $ C(t) $ models its fast varying part, called throughout the literature as its (temporal) fine structure, carrier component, or frequency modulation (FM).
 
-In the case of broadband signals, the problem of envelope detection of an arbitrary wave is ill-posed, in the sense that an infinite number of pairs of $ E(t) $ and $ C(t) $ can result in the same specific $ W(t) $ \parencite{2011TurnerDemodulation, 2013Mengempirical, 1996Loughlinamplitude}.
-
+In the case of broadband signals, the problem of envelope detection of an arbitrary wave is ill-posed, in the sense that an infinite number of pairs of $ E(t) $ and $ C(t) $ can result in the same $ W(t) $ \parencite{2011TurnerDemodulation, 2013Mengempirical, 1996Loughlinamplitude}.
+% TODO still ill posed
 We address this problem by assuming $ C(t) $ normalized between $ \{-1, 1\}$.
@@
 
-In this work we are concerned with the discrete version of this problem: given a finite digital wave represented by the vector $ \textbf{w} \in \mathbb{R}^n $, instead of the continuous function $ W(t) $, obtaining the temporal envelope of $ \textbf{w} $.
+In this work we are concerned with the discrete version of this problem: given a finite digital wave represented by the vector $ \mathbf{w} \in \mathbb{R}^n $, instead of the continuous function $ W(t) $, obtaining the temporal envelope $ \mathbf{e} $ of $ \mathbf{w} $.
 
 The preceding definitions can be translated to this discrete scenario assuming that the discrete quantities arise from observing the continuous ones at regular time intervals.
-In that case, the equality $ i = t \ \text{fps} $ can be used to link both settings, where $ i $ is the index of each observation, $ t $ stands for the time in seconds, and fps is the frame rate, or the number of observations made in one second.
+In that case, the equality $ i = t \ \text{fps} $ can be used to link both settings, where $ i $ is the index of each sample of the digital signal, $ t $ stands for the time in seconds, and fps is the frame rate, or the number of observations made in one second.
 We can thus define:
@@
 \begin{align}
-  & \mathbf{w}, \mathbf{e}, \mathbf{c} \in \mathbb{R}^n \\
-  & \mathbf{w} = (w_0, w_1, \cdots, w_{n-1}) \\
-  & \mathbf{e} = (e_0, e_1, \cdots, e_{n-1}) \\  
-  & \mathbf{c} = (c_0, c_1, \cdots, c_{n-1}) \\
-  & \mathbf{w} = \mathbf{e} \odot \mathbf{c} \ \therefore \ w_i = e_i c_i \ \forall \ i, \ 0 \le i \le n-1
+  \text{Given:}& \\
+  \mathbf{w} & \in \mathbb{R}^n, \\
+  \mathbf{w} & = (w_0, w_1, \cdots, w_{n-1}) \\
+  \text{We define:}& \\
+  \mathbf{e}, \mathbf{c} & \in \mathbb{R}^n \\
+  \mathbf{e} & = (e_0, e_1, \cdots, e_{n-1}), w_i \ge 0 \ \forall \ i, \ 0 \le i \le n-1 \\  
+  \mathbf{c} & = (c_0, c_1, \cdots, c_{n-1}), |c_i| \le 1 \ \forall \ i, \ 0 \le i \le n-1 \\
+  \mathbf{w} & = \mathbf{e} \odot \mathbf{c} \ \therefore \ w_i = e_i c_i \ \forall \ i, \ 0 \le i \le n-1
 \end{align}
@@
 Fig.
-\ref{fig:Envelope} provides an example of the vectors just defined: \textbf{w} is obtained by modulating the normalized carrier wave \textbf{c} using the envelope \textbf{e}.
-
-Note from Fig.
-\ref{fig:Envelope} that, at some local maxima, the envelope is well-defined, being in fact equal to the value of the wave $ w_i $.
-$ c_i = 1 $ at those points, rendering $ e_i = 1 w_i $.
+\ref{fig:Envelope} provides an example of the vectors just defined: \textbf{w} is obtained by multiplying the normalized carrier wave \textbf{c} using the envelope \textbf{e}.
 
@@
     \includegraphics{images/02Envelope.pdf}
-  \caption{Example of a discrete wave \textbf{w} composed by the element-wise multiplication of an envelope \textbf{e} and a carrier \textbf{c}.
+  \caption{Example of a discrete wave \textbf{w} composed by the element-wise multiplication of a previously known envelope \textbf{e} and carrier \textbf{c}.
 The local extrema of \textbf{w} are highlighted with a circle.}
@@
 
+In Fig.
+\ref{fig:Envelope} a signal arising from the multiplication of a known carrier wave \textbf{c} and a known envelope \textbf{e} is shown; this helps to illustrate a key concept: note from Fig.
+\ref{fig:Envelope} that, only at some local extrema, the envelope can be inferred without prior knowledge of the underlying carrier, being in fact equal to the value of the wave $ w_i $.
+$ c_i = 1 $ at those points, rendering $ e_i = 1 w_i $.
+
 % TODO
+
 This realization that the envelope of a wave is defined only at some of the wave's local extrema is crucial, as we can use it to generate a simplified representation of a signal that will make the algorithm more efficient by considering only the local extrema of a given signal.
@@
 
-In this new coordinate system we have a basis vector $ \mathbf{b}^\prime = \{ \mathbf{b}^\prime_x, \mathbf{b}^\prime_y \}$ where $ \mathbf{b}^\prime_x = (i, 0) $ and $\mathbf{b}^\prime_y = (0, i) $, related to the original basis vector by Eq.
+In this new coordinate system we have a basis vector $ \mathbf{b}^\prime = \{ \mathbf{b}^\prime_x, \mathbf{b}^\prime_y \}$ where $ \mathbf{b}^\prime_x = (i, 0) $ and $\mathbf{b}^\prime_y = (0, i) $, related to the original basis vector by the vertical scaling factor $ \beta $ as formalized in Eq.
 \ref{eq:Basis}:
@@
 \mathbf{b}^\prime_x &= \mathbf{b}_x  \\
-\mathbf{b}^\prime_y &= \frac{\left( \frac{i_{m-1} - i_0}{m-1} \right)}{\left( \frac{\sum_{j=0}^{m-1} \lvert w_j \lvert}{m} \right)} \mathbf{b}_y 
+\beta &= \frac{\left( \frac{i_{m-1} - i_0}{m-1} \right)}{\left( \frac{\sum_{j=0}^{m-1} \lvert w_j \lvert}{m} \right)}  \\
+\mathbf{b}^\prime_y &= \beta \ \mathbf{b}_y 
 \end{align}
@@
 
-The effect of this normalization can be seen in the points shown in Fig.\ref{fig:Vectors}, shown in scale in both coordinate systems.
+The effect of this normalization can be seen in the points shown in Fig.\ref{fig:Coordinates}, shown in scale for both coordinate systems.
+
+\begin{figure}[H]
+  \centering
+    \includegraphics{images/04Coordinates.pdf}
+  \caption{The set of points P and the set V of vectors between two adjacent points of P, in both the original and normalized coordinate systems, shown in scale.}
+  \label{fig:Coordinates}
+\end{figure}
 
@@
 
-To employ the alpha shapes algorithm in this set of points, however, the radius $ r $ of such circle must be estimated, and for that a measure of discrete curvature is needed.
+To employ an alpha shapes inspired algorithm in this set of points, however, the radius $ r $ of such circle must be estimated, and for that a measure of discrete curvature is needed.
 
-The set V of the vectors from each point of P to the next is also shown in the picture, and it will be used to estimate the average curvature of P.
+The set V of the vectors from each point of P to the next is shown in the Fig.
+\ref{fig:Vectors}, and will be used to estimate the average curvature of P.
 We define this set in Eq.
@@
     \includegraphics{images/04Vectors.pdf}
-  \caption{The set of points P and the set V of vectors between two adjacent points of P, in both the original and normalized coordinate systems, shown in scale.}
+  \caption{The set of points P and the set V of vectors between two adjacent points of P, in the normalized coordinate system.}
   \label{fig:Vectors}
@@
 \section{Quality of an envelope}
-Given the lack of a precise definition of what an envelope is, we begin this section suggesting an empirical metric, based on the behaviour of the carrier wave \textbf{c}.
+Given the lack of a precise definition of what an envelope is, we begin this section suggesting an empirical metric, based on the behaviour of the carrier wave \textbf{c}, obtained dividing, element-wise, the original signal and the envelope obtained.
 We also comment on some metrics presented in the literature, and compare our algorithm with traditional demodulation approaches, both in relation to execution time and accuracy; to this last end, a numerical indicator is also introduced.
 
-We also illustrate how the method here presented can be used to isolate not only a single envelope of the wave, but the superior and inferior envelopes as well, that we call frontiers, ending the section with a suggestion on how the proposed algorithm can be useful beyond envelope detection, by identifying the approximate locations of the pseudo-cycles in a quasi-periodic wave.
+We also illustrate how the method here presented can be adapted to isolate not only a single envelope of the wave, but the superior and inferior envelopes as well, that we call frontiers, 
+% TODO
+ending the section with a suggestion on how the proposed algorithm can be useful beyond envelope detection, by identifying the approximate locations of the pseudo-cycles in a quasi-periodic wave.
 
-Due to an assortment of factors, such as the unconventional approach presented in this work and the intrinsic difficulty in visualizing discrete waves given their high dimensionality, the authors deemed helpful to provide a companion website with interactive visualizations \parencite{2020TarjanoEnvelope}.
-
-An implementation of the method in pure Python, fully functional, can also be found at the repository dedicated to this work as well as a Python module available at the Python Package Index (PyPI) and conveniently installable via the command `pip install signal-envelope`.
-Besides, those working under Windows 64-bit can benefit from a specialized dynamic library coded in C++ that is automatically used within this module.
+Interactive visualizations are available from the repository prepared for this work \parencite{2020TarjanoEnvelope}, that also contains an implementation of the method in pure Python.
+A Python module available at the Python Package Index (PyPI) installable via the command `pip install signal-envelope` for convenience, is also available.
+Besides, those working under Windows 64-bit can benefit from a specialized dynamic library coded in C++ that is automatically used within this module, and can be separately obtained via the release section in the repository.
 
@@
 Recalling Fig.
-\ref{fig:Envelope}, however, and the accompanying discussion, an intuitive way is to consider the carrier wave \textbf{c}, and how well it is bounded by the interval $ \{-1, 1\} $.
+\ref{fig:Envelope}, however, and the accompanying discussion, an intuitive way is to observe how ell the carrier wave \textbf{c} conforms to the interval $ \{-1, 1\} $.
 
-To illustrate how the algorithm performs on such metric, we provide the following figures of original discrete waves \textbf{w}, their extracted envelope \textbf{e}, and the inferred carrier, obtained after dividing, element-wise, the original wave by its envelope.
+To illustrate how the algorithm performs on such metric, we provide the following figures of original discrete waves \textbf{w}, their extracted envelope \textbf{e}, and the inferred carrier, obtained after dividing, element-wise, the original signal by its extracted.
 
@@
 This is especially remarkable for the case of spoken voice, a notorious complex wave with fast changes in pitch and high noise content.
-
-These characteristics are in stark contrast with those of the sang voice in Fig.
-\ref{fig:singer_voice}, where a recording of an alto singer is shown; there, the wave is almost periodic, with a stable waveform.
 
@@
 
+These characteristics are in stark contrast with those of the sang voice in Fig.
+\ref{fig:singer_voice}, where a recording of an alto singer sustaining a steady note is shown; there, the wave is almost periodic, with a stable waveform and envelope.
+
 \begin{figure}[H]
@@
 
-Fig. \ref{fig:Bend} shows the envelope in the case of a bend performed in an eletric guitar. The variations in frequency in the signal didn't affect the envelope extracted.
+Fig. \ref{fig:Bend} shows the envelope in the case of a bend performed in an electric guitar. The variations in frequency in the signal didn't affect the envelope extracted.
 
@@
 
+% TODO %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 The first one states that, if a signal is bounded in magnitude, then its envelope should be, as well.
@@
 That is also the case, again because the points that form the envelope are points of the original wave itself.
+% TODO %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 We use an order 2 filter with a cut off frequency of 10 Hz.
-\item Hilbert Transform: We apply the Hilbert transform to the absolute values of the wave after filtering the original signal with a Butterworth filter with a cut off frequency of 100 Hz, using the absolute value of the analytic signal to obtain the envelope.
+\item Hilbert Transform: We apply the Hilbert transform to the original signal, filtering the absolute value of the complex result with a Butterworth filter with a cut off frequency of 100 Hz, using the absolute value of the analytic signal to obtain the envelope.
 \end{enumerate}
 
-In \ref{fig:Comparison_sinusoid} we illustrate the envelope extracted by the conventional algorithms, contrasting it with the result of the method presented in this work, for a simple sinusoid.
-Note that none of the conventional algorithms complies with the third condition proposed in \textcite{1996Loughlinamplitude} that the envelope of a periodic wave should be a straight line.
+In Fig.
+\ref{fig:Comparison_sinusoid} we illustrate the envelope extracted by the conventional algorithms, contrasting it with the result of the method presented in this work, for a simple sinusoid.
+Note that none of the conventional algorithms complies with the third condition proposed by \textcite{1996Loughlinamplitude}, that states that the envelope of a periodic wave should be a straight line.
 
@@
 
-In Fig. \ref{fig:Comparison_soprano} all envelopes exhibit a similar movement along the samples, but the ones generated by traditional algorithms are undershooting the original wave.
+In Fig. \ref{fig:Comparison_soprano} all envelopes exhibit a similar movement along the samples, but the ones generated by traditional algorithms are consistently undershooting the original wave.
 
@@
 
-All the traditional methods demanded careful choice of parameters.
-In the case of the Hilbert transform, pre-processing in the form of filtering was applied.
+Besides, the traditional methods demanded careful choice of parameters.
+In the case of the Hilbert transform, post-processing in the form of filtering was applied.
 That is in contrast with our method, that organically tunes itself via the automatic choice of the radius of the circle representing the average discrete curvature of the signal.
@@
 The average error was calculated by the average of the squared distance of the points in the envelope divided by two and the absolute value of the wave.
-The rationale is that the values of the envelope divided by two should pass near the average of the absolute values of the wave.
+The rationale is that the values of the envelope divided by two should lie near the average of the absolute values of the wave.
 This metric is only valid for reasonably smooth envelopes, however.
@@
 We can see from the table \ref{table:Error_comparison} that the method presented here consistently outperforms the traditional approaches for all waves tested, yielding an error 33\% smaller, on average, than the other methods.
-It is also worth noting that the errors of the traditional methods are very similar between them.
+It is also worth noting that the errors of the traditional methods are very similar between them, in agreement with the visualizations.
 
@@
 
-The times taken for the algorithms compared here to process each wave are shown in \ref{table:Time_comparison}.
-We used specialized implementations available in libraries of the Python programming language.
+The times taken for the algorithms to process each wave are shown in \ref{table:Time_comparison}. 
+The Python source for the tests, ass well as the samples used, are available at the repository dedicated to this work, in the file "Comparison of Envelope Algorithms.py".
+
+The methods were performed as described in the beginning of this section, using the implementations available at the Scipy library, under the Signal processing module.
 The presented algorithm is in average the second-fastest.
@@
 
-\section{Superior and inferior envelopes - Frontiers}
+\section{Superior and inferior envelopes}
 

### 2021-04-26 Papers/01_Pulses_Envelope/images/01AnalyticSignalAlto.py

@@ -1,38 +1,33 @@
 from scipy.signal import hilbert
 import numpy as np
 import plotly.graph_objects as go
-import numpy.polynomial.polynomial as poly
+import signal_envelope as se
 import sys
+from scipy.signal import butter, filtfilt
 
-'''Generating Wave'''
+def butter_lowpass_filter(data, fps, cutoff = 10, order = 2):
+  nyq = 0.5 * fps
+  normal_cutoff = cutoff / nyq
+  # Get the filter coefficients 
+  b, a = butter(order, normal_cutoff, btype='low', analog=False)
+  y = filtfilt(b, a, data)
+  return y
 
-np.random.seed(1)
-# fps = 10
-n = 1000+1
+name = "alto"
 
-X = np.arange(n)# / fps
-W = np.zeros(n)
-
-fp = [
-  [40, np.pi],
-  # [2, 1.2 * np.pi],
-  # [2.3, 1.4 * np.pi]
-  ]
-
-coefs = poly.polyfit([0, n//4, 3 * n//4, n], [0.6, 0.1, 0.9, .2], 3)
-A = poly.polyval(X, coefs)
-
-noise = np.random.normal(0, 0.1, n)
-noise[0 : n//2] = 0
-
-for f, p in fp:
-  W += np.cos(p + 2 * np.pi * f * X / n) + noise
-
-W = A * W / np.max(np.abs(W))
+'''==============='''
+''' Read wav file '''
+'''==============='''
+W, fps = se.read_wav(f"test_samples/{name}.wav")
+W = W - np.average(W)
+amplitude = np.max(np.abs(W))
+W = W / amplitude
+n = W.size
 
 '''Analytic_Signal'''
 analytic_signal = hilbert(W)
 hilbert_envelope = np.abs(analytic_signal)
+hilbert_envelope_filtered = butter_lowpass_filter(hilbert_envelope, fps, 100)
 
 
 '''============================================================================'''
@@ -52,7 +47,7 @@ fig.layout.template ="plotly_white"
 fig.update_layout(
   xaxis_title="<b>Sample <i>i</i></b>",
   yaxis_title="<b>Amplitude</b>",
-  legend=dict(orientation='h', yanchor='top', xanchor='left', y=1.1, itemsizing='trace'),
+  legend=dict(orientation='h', yanchor='top', xanchor='left', y=1.1, itemsizing='constant'),
[…]

### 2021-04-26 Papers/01_Pulses_Envelope/images/02Envelope.py

@@ -47,7 +47,7 @@ FONT = dict(
 fig = go.Figure()
 fig.layout.template ="plotly_white" 
 fig.update_layout(
-  xaxis_title="<b><i>i</i></b>",
+  xaxis_title="<b>Sample <i>i</i></b>",
   yaxis_title="<b>Amplitude</b>",
   legend=dict(orientation='h', yanchor='top', xanchor='left', y=1.1),
   margin=dict(l=0, r=0, b=0, t=0),
@@ -66,8 +66,8 @@ fig.add_trace(
     x=X,#pulses_X,
     y=E,#pulses_Y,
     mode="lines",
-    line=dict(width=2, color="gray", shape = 'spline'),
-    marker=dict(size=3, color="gray")
+    line=dict(width=2, color="red", shape = 'spline'),
+    marker=dict(size=3, color="red")
     # visible = "legendonly"
   )
 )
@@ -79,21 +79,21 @@ fig.add_trace(
     x=X,
     y=C,
     mode='lines+markers',
-    fill="tozeroy",
-    fillcolor="silver",
-    line=dict(width=1, color="gray"),
-    marker=dict(size=4, color="gray")
+    # fill="tozeroy",
+    # fillcolor="silver",
+    line=dict(width=1, color="blue"),
+    marker=dict(size=4, color="blue")
   )
 )
 
 fig.add_trace(
   go.Scatter(
-    name="Wave <b>w = e</b> ⊙ <b>c</b>      ",
+    name="Signal <b>w = e</b> ⊙ <b>c</b>      ",
     x=X,#pulses_X,
     y=W,#pulses_Y,
     mode="lines+markers",
-    line=dict(width=1, color="black"),
-    marker=dict(size=4, color="black")
+    line=dict(width=1, color="gray"),
+    marker=dict(size=4, color="gray")
     # visible = "legendonly"
   )
 )

### 2021-04-26 Papers/01_Pulses_Envelope/images/03Pulses.py

@@ -57,7 +57,7 @@ FONT = dict(
 fig = go.Figure()
 fig.layout.template ="plotly_white" 
 fig.update_layout(
-  xaxis_title="<b><i>i</i></b>",
+  xaxis_title="<b>Sample <i>i</i></b>",
   yaxis_title="<b>Amplitude</b>",
   legend=dict(orientation='h', yanchor='top', xanchor='left', y=1.1),
   margin=dict(l=0, r=0, b=0, t=0),
@@ -70,6 +70,8 @@ fig.layout.yaxis.title.font=FONT
 fig.update_xaxes(showline=False, showgrid=False, zeroline=False)
 fig.update_yaxes(showline=False, showgrid=False, zerolinewidth=2, zerolinecolor='black')
 
+pulses_color = "rgba(250, 140, 132, 0.5)"
+
 '''Samples'''
 fig.add_trace(
   go.Scatter(
@@ -79,7 +81,7 @@ fig.add_trace(
     y=W,
     mode='lines+markers',
     fill="tozeroy",
-    fillcolor="rgba(0,0,0,0.2)",
+    fillcolor=pulses_color,
     line=dict(color="gray", width=1),
     marker=dict(size=5, color="gray")
   )
@@ -88,10 +90,11 @@ fig.add_trace(
 ''' Samples Legend'''
 fig.add_trace(
   go.Scatter(
-    name="Samples      ",
+    name="Signal",
     x=[None],#pulses_X,
     y=[None],#pulses_Y,
-    mode="markers",
+    mode="lines+markers",
+    line=dict(color="gray", width=1),
     marker=dict(size=5, color="gray")
     # visible = "legendonly"
   )
@@ -100,12 +103,12 @@ fig.add_trace(
 ''' Pulses Legend'''
 fig.add_trace(
   go.Scatter(
-    name="Pulses      ",
+    name="Pulses",
     x=[None],#pulses_X,
     y=[None],#pulses_Y,
     fill="tozeroy",
     mode="none",
-    fillcolor="silver",
+    fillcolor=pulses_color,
     # marker=dict(
     #   size=5,
     #   color="black",
@@ -116,7 +119,7 @@ fig.add_trace(
 
 fig.add_trace(
   go.Scatter(
-    name="Local Extrema      ",
+    name="Local Extrema",
     # showlegend=False,
     x=Xp,
     y=Yp,
@@ -131,7 +134,7 @@ fig.add_trace(
 
 fig.add_trace(
[…]

### 2021-04-26 Papers/01_Pulses_Envelope/images/04Coordinates.py

@@ -1,0 +1,257 @@
+import sys
+print(sys.version)
+import plotly.graph_objects as go
+sys.path.append("c:/Users/tesse/Desktop/Files/Dropbox/0_Thesis/Python/01_Envelope")
+import numpy as np
+from scipy.interpolate import interp1d
+from Helper import signal_to_pulses, get_pulses_area
+from plotly.subplots import make_subplots
+
+'''Generating Wave'''
+
+np.random.seed(0)
+# fps = 10
+n = 100
+
+X = np.arange(n)# / fps
+C = np.zeros(n)
+
+f = interp1d([0, 0.25 * n, 0.5 * n, 0.75 * n, n], np.abs(np.random.normal(0, .3, 5)), "cubic")
+E = 1+f(X)
+
+afp = [
+  [1, 5, 1.4 * np.pi],
+  [1.5, 10, 1.4 * np.pi],
+  # [1, 2 * 6, 1.4 * np.pi]
+  ]
+for a, f, p in afp:
+  C += a * np.cos(p + 2 * np.pi * f * X / n) #+ np.random.normal(0, a / 100, n)
+
+C = -C / np.max(np.abs(C))
+
+W = E * C
+
+'''Wave to pulses'''
+pulses = signal_to_pulses(W)
+Xp = []
+Yp = []
+for p in pulses[1:-1]:
+  Xp.append(p.x)
+  Yp.append(np.abs(p.y))
+
+Xp = np.array(Xp)
+Yp = np.array(Yp)
+
+
+scaling = (np.sum(Xp[1:] - Xp[:-1])) / np.sum(Yp)
+Y = Yp * scaling
+'''============================================================================'''
+'''                              PLOT LINES                                    '''
+'''============================================================================'''
+fig = make_subplots(
+    rows=2, cols=1, shared_xaxes=False, vertical_spacing=0.01,
+    subplot_titles=("<b>Original coordinate system</b>", "<b>Normalized coordinate system</b>")
+    )
+
+FONT = dict(
+    family="Latin Modern Roman",
+    color="black",
+    size=13.3333
+  )
+
+for i in fig['layout']['annotations']:
+    i['font'] = FONT
+    
+'''Plotting'''
+
+fig.layout.template ="plotly_white" 
+fig.update_layout(
+  # xaxis_title="<b><i>x</i></b>", # ( <b><i>i</i></b> ) ",
[…]

### 2021-04-26 Papers/01_Pulses_Envelope/images/04Vectors.py

@@ -35,7 +35,7 @@ W = E * C
 pulses = signal_to_pulses(W)
 Xp = []
 Yp = []
-for p in pulses[1:-1]:
+for p in pulses[1:]:
   Xp.append(p.x)
   Yp.append(np.abs(p.y))
 
@@ -45,13 +45,12 @@ Yp = np.array(Yp)
 
 scaling = (np.sum(Xp[1:] - Xp[:-1])) / np.sum(Yp)
 Y = Yp * scaling
+
+
 '''============================================================================'''
 '''                              PLOT LINES                                    '''
 '''============================================================================'''
-fig = make_subplots(
-    rows=2, cols=1, shared_xaxes=False, vertical_spacing=0.01,
-    subplot_titles=("<b>Original coordinate system</b>", "<b>Normalized coordinate system</b>")
-    )
+fig = go.Figure()
 
 FONT = dict(
     family="Latin Modern Roman",
@@ -66,32 +65,20 @@ for i in fig['layout']['annotations']:
 
 fig.layout.template ="plotly_white" 
 fig.update_layout(
-  # xaxis_title="<b><i>x</i></b>", # ( <b><i>i</i></b> ) ",
-  # yaxis_title="<b><i>y</i></b>", # ( <b><i>i</i></b> ) ",
+  xaxis_title="<b><i>x</i></b>", # ( <b><i>i</i></b> ) ",
+  yaxis_title="<b><i>y</i></b>", # ( <b><i>i</i></b> ) ",
   yaxis = dict(scaleanchor = "x", scaleratio = 1),
   legend=dict(orientation='h', yanchor='top', xanchor='left', y=1.1),
   margin=dict(l=0, r=0, b=0, t=0),
   font=FONT,
   titlefont=FONT
 )
-# fig.layout.xaxis.title.font=FONT
-# fig.layout.yaxis.title.font=FONT
+fig.layout.xaxis.title.font=FONT
+fig.layout.yaxis.title.font=FONT
 
-fig.update_xaxes(showline=False, showgrid=False, zeroline=False, showticklabels=False, title=dict(font=FONT), row=1, col=1)
-fig.update_xaxes(showline=False, showgrid=False, zeroline=False, showticklabels=True, title=dict(text="<b><i>x</i></b>",font=FONT), row=2, col=1)
[…]

### 2021-04-26 Papers/01_Pulses_Envelope/images/06Graph.py

@@ -1,4 +1,3 @@
-'''To be used AFTER signal_envelope is installed via pip'''
 import sys
 # sys.path.append("signal_envelope/")
 import numpy as np
@@ -6,8 +5,8 @@ import signal_envelope as se
 from scipy import interpolate
 import plotly.graph_objects as go
 
-name = "spoken_voice"
-W, _ = se.read_wav(f"C:/Users/tesse/Desktop/Files/Dropbox/0_Science/reps/envelope/test_samples/{name}.wav")
+name = "alto"
+W, _ = se.read_wav(f"test_samples/{name}.wav")
 amp = np.max(np.abs(W))
 W = 4 * W / amp
 X = np.arange(W.size)
@@ -38,9 +37,9 @@ FONT = dict(
 fig = go.Figure()
 fig.layout.template ="plotly_white" 
 fig.update_layout(
-  xaxis_title="<b><i>i</i></b>",
+  xaxis_title="<b>Sample <i>i</i></b>",
   yaxis_title="<b>Amplitude</b>",
-  legend=dict(orientation='h', yanchor='top', xanchor='left', y=1.1),
+  legend=dict(orientation='h', yanchor='top', xanchor='left', y=1.1, itemsizing='constant'),
   margin=dict(l=0, r=0, b=0, t=0),
   font=FONT,
   titlefont=FONT
@@ -58,7 +57,7 @@ fig.add_trace(
     # x=np.arange(W.size),
     y=W,
     mode="lines",
-    # fill="tozeroy",
+    line_shape='spline',
     # fillcolor="gray",
     line=dict(width=1, color="silver",),
   )
@@ -70,7 +69,7 @@ fig.add_trace(
     # x=np.arange(W.size),
     y=C,
     mode="lines",
-    # fill="tozeroy",
+    line_shape='spline',
     # fillcolor="rgba(0,0,0,0.3)",
     line=dict(width=1, color="rgba(235, 64, 52,0.3)",),
   )
@@ -79,7 +78,7 @@ fig.add_trace(
 fig.add_trace(
   go.Scatter(
     name="Envelope (<b>e</b>)      ",
-    # x=E,
+    line_shape='spline',
     y=E,
     mode="lines",
     line=dict(width=2, color="black"),
@@ -108,7 +107,7 @@ fig.add_trace(
 #   )
 # )
 
-fig.show(config=dict({'scrollZoom': True}))
+# fig.show(config=dict({'scrollZoom': True}))
[…]

### 2021-04-29 Papers/01_Pulses_Envelope/PulsesEnvelope.tex
progress with the revision

@@
 We follow with suggestions about how the approach here presented can be useful beyond envelope detection, exemplifying with some applications, such as the extractions of the average waveform of a digital wave, and concluding with commentary on potential future developments of the method.
+
+Interactive visualizations are available from the repository prepared for this work \parencite{2020TarjanoEnvelope}, that also contains an implementation of the method in pure Python.
+A Python module available at the Python Package Index (PyPI) installable via the command `pip install signal-envelope` for convenience, is also available.
+Besides, those working under Windows 64-bit can benefit from a specialized dynamic library coded in C++ that is automatically used within this module, and can be separately obtained via the release section in the repository.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
   \centering
-    \includegraphics{images/01AnalyticSignalAlto.pdf.pdf}
+    \includegraphics{images/01AnalyticSignalAlto.pdf}
   \caption{Representation of an alto singer sustaining a steady note and a comparison of the envelopes obtained by the Hilbert transform with and without further filtering.}
@@
 
-\section{Quality of an envelope}
-Given the lack of a precise definition of what an envelope is, we begin this section suggesting an empirical metric, based on the behaviour of the carrier wave \textbf{c}, obtained dividing, element-wise, the original signal and the envelope obtained.
-We also comment on some metrics presented in the literature, and compare our algorithm with traditional demodulation approaches, both in relation to execution time and accuracy; to this last end, a numerical indicator is also introduced.
+\section{Quality of an envelope}\label{sec:quality}
 
-We also illustrate how the method here presented can be adapted to isolate not only a single envelope of the wave, but the superior and inferior envelopes as well, that we call frontiers, 
-% TODO
-ending the section with a suggestion on how the proposed algorithm can be useful beyond envelope detection, by identifying the approximate locations of the pseudo-cycles in a quasi-periodic wave.
+The lack of a precise definition of what an envelope is makes the assessment of the quality of an envelope a difficult task, for which there are no agreed-upon objective metrics.
+Recalling Fig.
+\ref{fig:Envelope}, however, and the accompanying discussion, if one divides, element-wise, the original signal by the envelope, the carrier wave thus obtained is expected to be bounded between -1 and 1.
 
-Interactive visualizations are available from the repository prepared for this work \parencite{2020TarjanoEnvelope}, that also contains an implementation of the method in pure Python.
-A Python module available at the Python Package Index (PyPI) installable via the command `pip install signal-envelope` for convenience, is also available.
-Besides, those working under Windows 64-bit can benefit from a specialized dynamic library coded in C++ that is automatically used within this module, and can be separately obtained via the release section in the repository.
-
-There are no agreed-upon objective metrics to assess the quality of an envelope.
-Recalling Fig.
-\ref{fig:Envelope}, however, and the accompanying discussion, an intuitive way is to observe how ell the carrier wave \textbf{c} conforms to the interval $ \{-1, 1\} $.
-
-To illustrate how the algorithm performs on such metric, we provide the following figures of original discrete waves \textbf{w}, their extracted envelope \textbf{e}, and the inferred carrier, obtained after dividing, element-wise, the original signal by its extracted.
+Figs.
+\ref{fig:spoken_voice}, \ref{fig:singer_voice} and \ref{fig:Bend} present discrete waves \textbf{w}, their extracted envelope \textbf{e}, and the inferred carrier \textbf{c}, obtained after dividing, element-wise, the original signal by its extracted.
 
@@
 
-These characteristics are in stark contrast with those of the sang voice in Fig.
+Those characteristics are in stark contrast with those of the sang voice in Fig.
 \ref{fig:singer_voice}, where a recording of an alto singer sustaining a steady note is shown; there, the wave is almost periodic, with a stable waveform and envelope.
@@
 
-In an effort to formalize the assessment of the quality of temporal envelopes, \textcite{1996Loughlinamplitude} proposed some conditions necessary, but not sufficient, to ensure the physical plausibility of an envelope.
+In an effort to formalize the assessment of the quality of temporal envelopes, \textcite{1996Loughlinamplitude} proposed some theoretical conditions necessary, but not sufficient, to ensure the physical plausibility of an envelope.
 We comment briefly below that the algorithm here presented satisfies the four conditions presented.
@@
 That is also the case, again because the points that form the envelope are points of the original wave itself.
-% TODO %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+In order to obtain a numeric indicator of the quality of an envelope, we construct a reference envelope, that trades off pratical applicability in favour of mathematical formality.
+In \textcite{1992BruceCurves} the envelope is defined as the boundary of the region filled by a family of curves.
+Fig.
+\ref{fig:RefEnvelopeTheory} illustrates this concept, for a family of sinusoids that share the same frequency and amplitude, and whose phases vary from 0 to $ 2 \pi $.
+
+\begin{figure}[H]
+  \centering
+    \includegraphics{images/09RefEnvelopeTheory.pdf}
+  \caption{Envelope as the boundary of the region filled by a family of curves, for a family of sinusoids with the same frequency and amplitude, modulated by a polynomial.}
+  \label{fig:RefEnvelopeTheory}
+\end{figure}
+
+% TODO
+In practical applications, one seldom has access to a family of curve, instead being interested in obtaining the envelope of a single signal.
+Nevertheless, an approximation can be obtained, by shifting the original signal along its initial position in the horizontal axis, in order to cover the time of one period of the signal.
+Fig. \ref{fig:RefEnvelope} shows the reference envelope for the sound of a tom.
+
+\begin{figure}[H]
+  \centering
+    \includegraphics{images/10RefEnvelope.pdf}
+  \caption{Reference envelope obtained by shifting the original signal, for part of a sample of a tom.}
+  \label{fig:RefEnvelope}
+\end{figure}
+
+Since envelope is not smooth, being composed of a series of horizontal segments, as can be seen in \ref{fig:RefEnvelope}, besides being computationally expensive to calculate.
+However, arising from a mathematically sound base, it can serve as a reference to infer the quality of more empirical methods.
 
@@
 \item Filtering: The absolute value of the wave is filtered using a digital implementation of the Butterworth \parencite{butterworth1930theory} low pass filter.
-We use an order 2 filter with a cut off frequency of 10 Hz.
-\item Hilbert Transform: We apply the Hilbert transform to the original signal, filtering the absolute value of the complex result with a Butterworth filter with a cut off frequency of 100 Hz, using the absolute value of the analytic signal to obtain the envelope.
+We use an order 2 filter with a cut-off frequency of 10 Hz.
+\item Hilbert Transform: We apply the Hilbert transform to the original signal, filtering the absolute value of the complex result with a Butterworth filter with a cut-off frequency of 1/10 of the fundamental frequency of the original signal.
 \end{enumerate}
 
-In Fig.
-\ref{fig:Comparison_sinusoid} we illustrate the envelope extracted by the conventional algorithms, contrasting it with the result of the method presented in this work, for a simple sinusoid.
+In Fig. \ref{fig:Comparison_sinusoid} we illustrate the envelope extracted by the conventional algorithms, contrasting it with the result of the method presented in this work, for a simple sinusoid. 
+% TODO
 Note that none of the conventional algorithms complies with the third condition proposed by \textcite{1996Loughlinamplitude}, that states that the envelope of a periodic wave should be a straight line.
@@
 Besides, the traditional methods demanded careful choice of parameters.
-In the case of the Hilbert transform, post-processing in the form of filtering was applied.
+In the case of the Hilbert transform, post-processing in the form of filtering was applied, and an appropriate cut-off frequency had to be chosen.
 That is in contrast with our method, that organically tunes itself via the automatic choice of the radius of the circle representing the average discrete curvature of the signal.
 
-Table \ref{table:Error_comparison} presents a comparison of the average errors of the techniques.
-The average error was calculated by the average of the squared distance of the points in the envelope divided by two and the absolute value of the wave.
-The rationale is that the values of the envelope divided by two should lie near the average of the absolute values of the wave.
-This metric is only valid for reasonably smooth envelopes, however.
-
-We can see from the table \ref{table:Error_comparison} that the method presented here consistently outperforms the traditional approaches for all waves tested, yielding an error 33\% smaller, on average, than the other methods.
-It is also worth noting that the errors of the traditional methods are very similar between them, in agreement with the visualizations.
+Table \ref{table:Error_comparison} presents the average of the absolute errors of the techniques, when compared with the reference envelope introduced in the end of section \ref{sec:quality}
 
@@
 
+We can see from the table \ref{table:Error_comparison} that the method presented here consistently outperforms the traditional approaches for all waves tested, yielding an error 33\% smaller, on average, than the other methods.
+It is also worth noting that the errors of the traditional methods are very similar between them, in agreement with the visualizations.
+
 The times taken for the algorithms to process each wave are shown in \ref{table:Time_comparison}. 
@@
 The methods were performed as described in the beginning of this section, using the implementations available at the Scipy library, under the Signal processing module.
-The presented algorithm is in average the second-fastest.
+The presented algorithm, is on average, the second-fastest.
 

### 2021-04-29 Papers/01_Pulses_Envelope/code/phases test.py

@@ -1,0 +1,106 @@
+import plotly.graph_objects as go
+import numpy as np
+from scipy.interpolate import interp1d
+from scipy.signal import savgol_filter, hilbert, butter, filtfilt
+import sys
+from timeit import default_timer as timer
+import signal_envelope as se
+
+def butter_lowpass_filter(data, fps, cutoff = 10, order = 2):
+  nyq = 0.5 * fps
+  normal_cutoff = cutoff / nyq
+  # Get the filter coefficients 
+  b, a = butter(order, normal_cutoff, btype='low', analog=False)
+  y = filtfilt(b, a, data)
+  return y
+
+def get_reference_envelope(W):
+  Wp = np.pad(W, (0, n//2))
+  Wp_roll = np.copy(Wp)
+  conv = np.zeros(n//2)
+
+  t = int(n / np.argmax(np.abs(np.fft.rfft(W))))
+
+  for i in range(t//2, 10 * t):
+    Wp_roll = np.roll(Wp_roll, 1)
+    conv[i] = np.sum(Wp_roll * Wp)
+
+  t = int(np.argmax(conv) * 2)
+  rolled_W = np.roll(np.pad(W, (t//2, t//2) ), -t//2)
+
+  # print(t)
+  for i in range(t):
+    rolled_W = np.roll(rolled_W, 1)
+    for j in range(W.size):
+      W[j] = max(abs(W[j]), abs(rolled_W[t//2 + j]))
+  return W
+
+# name = "piano" # change the name here for any of the files in the "test_samples" folder
+# name = "sinusoid"
+name = "tom"
+
+'''==============='''
+''' Read wav file '''
+'''==============='''
+W, fps = se.read_wav(f"test_samples/{name}.wav")
+W = W - np.average(W)
+amplitude = np.max(np.abs(W))
+W = W / amplitude
+n = W.size
+print(f"n={n}")
+X = np.arange(n)
+
+
+'''============================================================================'''
+'''                                    PLOT                                    '''
+'''============================================================================'''
+FONT = dict(
+    family="Latin Modern Roman",
+    color="black",
+    size=13.3333
+  )
+
+'''Plotting'''
+fig = go.Figure()
+fig.layout.template ="plotly_white" 
+fig.update_layout(
+  xaxis_title="<b>Sample <i>i</i></b>",
+  yaxis_title="<b>Amplitude</b>",
[…]

### 2021-04-29 Papers/01_Pulses_Envelope/images/01AnalyticSignalAlto.py

@@ -60,7 +60,7 @@ fig.update_yaxes(showline=False, showgrid=False, zerolinewidth=.8, zerolinecolor
 '''Plotting Signal'''
 fig.add_trace(
   go.Scatter(
-    name="Signal",
+    name="Signal   ",
     # x=X,
     y=W,
     mode="lines",
@@ -74,7 +74,7 @@ fig.add_trace(
 
 fig.add_trace(
   go.Scatter(
-    name="Hilbert Transform",
+    name="Hilbert Transform   ",
     # x=X,
     y=hilbert_envelope,
     mode="lines",
@@ -90,7 +90,7 @@ fig.add_trace(
 
 fig.add_trace(
   go.Scatter(
-    name="Filtered Hilbert Transform",
+    name="Filtered Hilbert Transform   ",
     # x=X,
     y=hilbert_envelope_filtered,
     mode="lines",

### 2021-04-29 Papers/01_Pulses_Envelope/images/06Graph.py

@@ -5,7 +5,9 @@ import signal_envelope as se
 from scipy import interpolate
 import plotly.graph_objects as go
 
-name = "alto"
+# name = "alto"
+# name="bend"
+name="spoken_voice"
 W, _ = se.read_wav(f"test_samples/{name}.wav")
 amp = np.max(np.abs(W))
 W = 4 * W / amp
@@ -109,7 +111,7 @@ fig.add_trace(
 
 # fig.show(config=dict({'scrollZoom': True}))
 save_name = "./images/" + sys.argv[0].split('/')[-1].replace(".py", "") + "_" + name + ".svg"
-fig.write_image(save_name, width=650, height=270, engine="kaleido", format="svg")
+fig.write_image(save_name, width=650, height=300, engine="kaleido", format="svg")
 print("saved:", save_name)
 
 

### 2021-04-29 Papers/01_Pulses_Envelope/images/09RefEnvelopeTheory.py

@@ -1,0 +1,96 @@
+import sys
+print(sys.version)
+import plotly.graph_objects as go
+sys.path.append("c:/Users/tesse/Desktop/Files/Dropbox/0_Thesis/Python/01_Envelope")
+import numpy as np
+from scipy.interpolate import interp1d
+from Helper import signal_to_pulses, get_pulses_area
+
+'''Generating Wave'''
+
+def get_W(X, E, p=0):
+  C = np.zeros(E.size)
+  afp = [
+    [1, 5, p],
+    # [1.5, 10, p],
+    ]
+  for a, f, p in afp:
+    C += a * np.cos(p + 2 * np.pi * f * X / n) #+ np.random.normal(0, a / 100, n)
+  C = C / np.max(np.abs(C))
+  return E * C
+
+np.random.seed(0)
+# fps = 10
+n = 100
+X = np.arange(n)# / fps
+f = interp1d(np.linspace(0, n, 7, endpoint=True), np.abs(1 + np.random.normal(0, .3, 7)), "cubic")
+E = f(X)
+
+
+
+
+
+
+
+'''============================================================================'''
+'''                              PLOT LINES                                    '''
+'''============================================================================'''
+FONT = dict(
+    family="Latin Modern Roman",
+    color="black",
+    size=13.3333
+  )
+
+'''Plotting'''
+fig = go.Figure()
+fig.layout.template ="plotly_white" 
+fig.update_layout(
+  xaxis_title="<b>Sample <i>i</i></b>",
+  yaxis_title="<b>Amplitude</b>",
+  legend=dict(orientation='h', yanchor='top', xanchor='left', y=1.1),
+  margin=dict(l=0, r=0, b=0, t=0),
+  font=FONT,
+  titlefont=FONT
+)
+fig.layout.xaxis.title.font=FONT
+fig.layout.yaxis.title.font=FONT
+
+fig.update_xaxes(showline=False, showgrid=False, zeroline=False)
+fig.update_yaxes(showline=False, zeroline=False, showgrid=True)
+
+fig.add_trace(
+  go.Scatter(
+    name="Envelope      ",
+    x=X,#pulses_X,
+    y=E,#pulses_Y,
+    mode="lines",
+    line=dict(width=2, color="red", shape = 'spline'),
+    marker=dict(size=3, color="red")
+    # visible = "legendonly"
+  )
+)
+
+Xf = []
+Yf = []
[…]

### 2021-04-30 Papers/01_Pulses_Envelope/PulsesEnvelope.tex
progress

@@
 
+% TODO proof curvature %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+Our problem is to derive a measure of the average curvature of a discrete curve from its oscillations in the vertical axis.
+Recalling that we derived the points in P from the local maxima, there is no local maxima between p0 and p1, and the envelope between both of them doesnt go below the minimum of them.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 The set V of the vectors from each point of P to the next is shown in the Fig.
@@
 \multicolumn{5}{c}{\textbf{Average error per frame}} \\
-              & Present Work & Smoothing & Low pass & Hilbert \\ 
+Signal & Present Work & Smoothing & Low pass & Hilbert \\ 
 \hline
-alto          & 0.073        & 0.083     & 0.083   & 0.082   \\
-bend          & 0.009        & 0.015     & 0.015   & 0.014   \\
-piano         & 0.005        & 0.006     & 0.006   & 0.005   \\
-brass         & 0.024        & 0.035     & 0.035   & 0.033   \\
-soprano       & 0.042        & 0.065     & 0.065   & 0.065   \\
-spoken voice  & 0.024        & 0.037     & 0.037   & 0.034   \\
-tom           & 0.003        & 0.005     & 0.005   & 0.005   \\
-sinusoid      & 0.113        & 0.193     & 0.195   & 0.191   \\ 
+alto  & 0.001 & 0.337 & 0.342 & 0.163 \\
+bend  & 0.000 & 0.029 & 0.030 & 0.008 \\
+brass & 0.001 & 0.107 & 0.107 & 0.038 \\
+nonperiodic & 0.000 & 0.211 & 0.158 & 0.000 \\
+piano & 0.003 & 0.016 & 0.016 & 0.007 \\
+sinusoid & 0.000 & 0.132 & 0.131 & 0.000 \\
+soprano & 0.000 & 0.155 & 0.155 & 0.044 \\
+spoken\_voice & 0.006 & 0.109 & 0.110 & 0.045 \\
+tom   & 0.001 & 0.011 & 0.015 & 0.005 \\
 \hline
-Average       & 0.037        & 0.055     & 0.055   & 0.054  
+Average & 0.001 & 0.123 & 0.118 & 0.035 \\
+frac. of max. & 0.010 & 1.000 & 0.961 & 0.281 \\
 \end{tabular}
@@
 \multicolumn{5}{c}{\textbf{Time in seconds}} \\
-              & Present Work & Smoothing & Low pass & Hilbert \\ 
+Signal  & Present Work & Smoothing & Low pass & Hilbert \\ 
 \hline
-alto          & 0.008        & 0.101     & 0.007   & 0.011   \\
-bend          & 0.030        & 0.226     & 0.012   & 0.131   \\
-piano         & 0.034        & 0.374     & 0.014   & 0.042   \\
-brass         & 0.032        & 0.297     & 0.011   & 0.045   \\
-soprano       & 0.177        & 0.735     & 0.027   & 0.566   \\
-spoken voice  & 0.014        & 0.105     & 0.003   & 0.038   \\
-tom           & 0.007        & 0.084     & 0.018   & 0.026   \\
-sinusoid      & 0.007        & 0.091     & 0.005   & 0.011   \\ 
+alto  & 0.010 & 0.096 & 0.007 & 0.006 \\
+bend  & 0.009 & 0.146 & 0.005 & 0.067 \\
+brass & 0.015 & 0.253 & 0.005 & 0.035 \\
+nonperiodic & 0.002 & 0.064 & 0.003 & 0.007 \\
+piano & 0.022 & 0.316 & 0.009 & 0.030 \\
+sinusoid & 0.003 & 0.069 & 0.003 & 0.006 \\
+soprano & 0.111 & 0.574 & 0.013 & 0.293 \\
+spoken\_voice & 0.003 & 0.062 & 0.003 & 0.017 \\
+tom   & 0.002 & 0.070 & 0.001 & 0.004 \\
 \hline
-Average       & 0.039        & 0.252     & 0.012   & 0.109  
+Average & 0.020 & 0.183 & 0.005 & 0.052 \\
+frac. of max. & 0.108 & 1.000 & 0.030 & 0.283 \\
 \end{tabular}
-\caption{Comparison of the processing time of the algorithm presented in this work with the most common methods of digital envelope identification.}
+\caption{Comparison of the processing time of the algorithm presented in this work with the most common methods of digital envelope detection.}
 \label{table:Time_comparison}

### 2021-05-03 Papers/01_Pulses_Envelope/PulsesEnvelope.tex
progress with the revision

@@
 
-We have now a map from the original space of the two-dimensional representation of a discrete wave to the Cartesian coordinate system and thus a way to consistently transform a discrete signal into a set of points in the plane.
+Having established a map from the original space of the two-dimensional representation of a discrete wave to the Cartesian coordinate system and, thus, a way to consistently transform our simplified description of a discrete signal into a set of points in the plane, the shape of which can be estimated with the use of geometric algorithms.
 
-Translating the explanation of the alpha shapes algorithm presented in \textcite{1994EdelsbrunnerThree} to our context, we can say that the points that are part of the envelope are those touched by a circle of a given radius $ r $, coming from $ y \to \infty $ towards the Cartesian representation of the signal, that is not allowed to contain any point of the signal; Intuitively, one can picture a circle being rolled above the signal, and marking the points it touches as envelope points.
+The goal of those algorithms, such as the convex hull and the alpha-shapes, is to discriminate which points of a set are part of its frontier.
+
+Translating the explanation of the alpha-shapes algorithm presented in \textcite{1994EdelsbrunnerThree} to our context, we can say that the points that are part of the envelope are those that can be touched by a circle of a given radius $ r $, that must not contain any point of the signal, coming from $ y \to \infty $ towards the Cartesian representation of the signal.
+
+% TODO formal definition of alpha shapes, maybe image
+
+Intuitively, one can picture a circle of fixed radius being rolled above the signal, and marking the points it touches as envelope points.
 
 To employ an alpha shapes inspired algorithm in this set of points, however, the radius $ r $ of such circle must be estimated, and for that a measure of discrete curvature is needed.
-
-% TODO proof curvature %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-Our problem is to derive a measure of the average curvature of a discrete curve from its oscillations in the vertical axis.
-Recalling that we derived the points in P from the local maxima, there is no local maxima between p0 and p1, and the envelope between both of them doesnt go below the minimum of them.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-The set V of the vectors from each point of P to the next is shown in the Fig.
-\ref{fig:Vectors}, and will be used to estimate the average curvature of P.
-We define this set in Eq.
-\ref{eq:Vectors} below.
-% TODO
-
-\begin{equation} \label{eq:Vectors}
-\text{V} = \{ v_0, v_1, v_{m-2} \} \quad \text{where} \quad v_k = P_{j+1} - P_j \quad \forall \quad 0 \le k \le m-2
-\end{equation}
-
-\begin{figure}[H]
-  \centering
-    \includegraphics{images/04Vectors.pdf}
-  \caption{The set of points P and the set V of vectors between two adjacent points of P, in the normalized coordinate system.}
-  \label{fig:Vectors}
-\end{figure}
 
@@
 Those approaches define the discrete curvature in relation to the vertices of a discrete curve (or mesh, in the three-dimensional case).
-In our particular case, it wouldn't make sense to talk about the curvature of a single pulse, since it is what our points ultimately represent.
-Instead, we are interested in the change of direction, what the term curvature ultimately means, between two adjacent pulses, as the vectors in the set V suggests.
+Instead, we are interested in the change of direction, what the term curvature ultimately means, between two adjacent points.
 
-We thus proceed to define a discrete curvature measure over the edges of a discrete curve.
-To that end we are going to apply the definition of smooth curvature as the rate of change of the unit tangent to a curve, noting that this is equivalent to that of the osculating circle \parencite{2016VasaMesh}.
+Since we are dealing with a potentially large set of points, it is important for this measure to act locally, to assure $ O(n) $ complexity.
+
+Thus, the set V of the vectors from each point of P to the next, shown in the Fig.
+\ref{fig:Vectors}, will be used to estimate the average curvature of P: We will calculate the equivalent circle for each vector, and average the values thus obtained to infer the average curvature of the whole discrete curve.
+
+\begin{figure}[H]
+  \centering
+    \includegraphics{images/04Vectors.pdf}
+  \caption{The set of points P and the set V of vectors between two adjacent points of P, in the normalized coordinate system.}
+  \label{fig:Vectors}
+\end{figure}
+
+We formalize the set of vectors V in Eq. \ref{eq:Vectors} below.
+
+\begin{equation} \label{eq:Vectors}
+\text{V} = \{ v_0, v_1, v_{m-2} \} \quad \text{where} \quad v_k = P_{j+1} - P_j \quad \forall \quad 0 \le k \le m-2
+\end{equation}
 
@@
 
-The rationale is to find, for each vector $ v_k \in \text{V} $, the radius $ r_k $ of the equivalent circle whose tangent has the same change in direction, in the same horizontal distance, as the vector of interest, as shown in Fig.
+We thus proceed to define a discrete curvature measure over the edges of a discrete curve.
+To that end we are going to apply the definition of smooth curvature as the rate of change of the unit tangent to a curve.
+This definition is convenient, since it is equivalent to that of the osculating circle \parencite{2016VasaMesh}.
+
+One of those edges, a vector $ v_k \in \text{V} $, is singled out in Fig.
+\ref{fig:DiscreteCurvature}.
+$ v_k $ represents the net displacement of the envelope from the extrema $ P_{j} $ to $ P_{j+1} $.
+Since we dont yet know which points of P are part of the envelope, instead of assuming the trajectory before $ P_{j} $ passes through
+$ P_{j-1} $, it is more prudent to assume that it can have passed anywhere in the line segment from 
+% TODO
+That is because all we can assume from $ v_k $ is that, in this current region of the envelope, it is able to a vertical displacement of $ v_{k,y} $ in the same time as it performs a horizontal displacement of $ v_{k,x} $.
+In average, we can consider the envelope entering from the horizontal direction.
+
+The rationale then is to find, for each vector $ v_k $, the radius $ r_k $ of the equivalent circle whose tangent has the same change in direction, in the same horizontal distance, as the vector of interest, as shown in Fig.
 \ref{fig:DiscreteCurvature}.
@@
 From Fig.
-\ref{fig:DiscreteCurvature} is easy to see that $ r_k = v_{k,x} / \sin(\theta_k) $ where $ v_{k,x} $ is the component of $ v_k $ in the horizontal direction; $ \theta_k $ itself can be obtained using the slope $ m_0 = 0 $ of a line in the horizontal direction and the slope $ m_k = v_{k,y} / v_{k,x} $ defined by $ v_k $ via the equality $ \theta_k = \arctan \big( (m_k - m_0) / (1 + m_k m_0) \big) $.
+\ref{fig:DiscreteCurvature} one can see that $ r_k = v_{k,x} / \sin(\theta_k) $ where $ v_{k,x} $ is the component of $ v_k $ in the horizontal direction; $ \theta_k $ itself can be obtained using the slope $ m_0 = 0 $ of the horizontal line and the slope $ m_k = v_{k,y} / v_{k,x} $ defined by $ v_k $ using the equality $ \theta_k = \arctan \big( (m_k - m_0) / (1 + m_k m_0) \big) $.
 
@@
 
-The radius of the circle that represents the average curvature of V (and, consequently, of P), and will be used to obtain the envelope of the discrete wave \textbf{w} can be obtained via Eq.
-\ref{eq:Radius}:
+The radius of the circle equivalent to each $ v_k $ is defined in Eq.
+\ref{eq:Radius}.
+This equation has the desirable property that $ v_{k,y} \to 0 \implies v_{k,y} \to \infty $, that is, the radius tends to infinity when no vertical oscillation occurs, effectively turning the circle into a line.
+% It is also directly proportional to $ v_{k,x} $, making the radius larger the more the points are horizontally far from each other.
 
 \begin{equation} \label{eq:Radius}
-r = \frac{\sum\limits_{k=0}^{m-2} \left( \frac{v_{k,x} \sqrt{v_{k,x}^2 + v_{k,y}^2}}{ v_{k,y}} \right)}{m-2}
+  r_k = \frac{v_{k,x} \sqrt{v_{k,x}^2 + v_{k,y}^2}}{ v_{k,y}}
+\end{equation}
+
+For calculation purposes, however, this can cause a problem, as a zero could make its way into the denominator of a fraction; in the actual algorithm, the quantity calculated for each vector is the curvature itself, that is, the inverse of the radius.
+In this way, the problem is avoided, since $ v_{k,y} $ is guaranteed to be greater than zero.
+The average of curvatures is then inverted in order to obtain the average radius, as illustrated in Eq.
+\ref{eq:Curvature}.
+
+\begin{equation} \label{eq:Curvature}
+r = \left( \frac{\sum\limits_{k=0}^{m-2} \left( \frac{ v_{k,y}}{v_{k,x} \sqrt{v_{k,x}^2 + v_{k,y}^2}} \right)}{m-2} \right)^{-1}
 \end{equation}
@@
 We need now to construct an algorithm, using the radius just obtained, to identify the points that belong to the envelope.
-Generally, alpha shapes based algorithms resort to the construction of the Delaunay triangulation of the set of points, that is latter filtered to contain only the outer edges.
+Generally, alpha-shapes based algorithms resort to the construction of the Delaunay triangulation of the set of points, that is latter filtered to contain only the outer edges.
 We can adopt a more straightforward method, as our points have a sequential structure.
-Although that is the case throughout this work, the original signal doesn't need to be sampled in equal spaces.
+
+It is worth noting that, although that is the case throughout this work, the original signal doesn't need to be sampled at equal intervals.
 
@@
 The algorithm follows directly from the definition in \ref{eq:InEnvelope} after noting that the first and last members of P will always be part of the envelope, since they are reachable by a circle of any radius $ r $.
-From $ P_0 $, our first pivot point, we construct circles passing through $ P_1, P_2, \cdots, P_d $ until a circle that doesn't contain any point in P is found; $ P_d $ then becomes the new pivot point and is included in E, and this procedure continues until $ P_{m-1} $ is reached.
+From $ P_0 $, our first pivot point, we construct circles passing through $ P_1, P_2, \cdots, P_d $ until a circle that doesn't contain any point in P is found; $ P_d $ then becomes the new pivot point and is included in E, and this process continues until $ P_{m-1} $ is reached.
 
-The procedure is formalized in the algorithm \ref{FindEnvelope}. We don't provide similar algorithmic descriptions of the preceding steps since the Python implementation available at % TODO
- can easily serve as pseudocode.
+The procedure is formalized in the algorithm \ref{FindEnvelope}.
+We don't provide similar algorithmic descriptions of the preceding steps since the Python implementation available at \textcite{2020TarjanoEnvelope} can easily serve as pseudocode.
 
@@
 The most straightforward way to transform those points into the vector $ \textbf{e} \in \mathbb{R}^n $ is via linear interpolation, the approach used in this work.
+% TODO
 It is worth noting, however, that many other procedures are available, both for the interpolation and the smooth approximation of points.

### 2021-05-04 Papers/01_Pulses_Envelope/PulsesEnvelope.tex
some progress with the review

@@
 Considering the standard 2d representation of a discrete wave, where the horizontal axis represents time and the vertical axis represents the amplitude, one can think of an individual sample as a point.
-Provided a coherent mapping from this space to the standard cartesian plane is available, the original DSP problem can be transformed into a geometric one.
-
-% TODO
+Provided a coherent mapping from this space to the standard Cartesian plane is available, the original DSP problem can be transformed into the geometric problem of finding the shape of a set of points.
 
@@
 In the case of broadband signals, the problem of envelope detection of an arbitrary wave is ill-posed, in the sense that an infinite number of pairs of $ E(t) $ and $ C(t) $ can result in the same $ W(t) $ \parencite{2011TurnerDemodulation, 2013Mengempirical, 1996Loughlinamplitude}.
-% TODO still ill posed
-We address this problem by assuming $ C(t) $ normalized between $ \{-1, 1\}$.
+
+We address this problem by assuming $ C(t) $ normalized between $ \{-1, 1\}$ or, in other words $ |\max(C(t))| = 1 $.
 This assumption doesn't cause any loss of generality since, given an arbitrary $ C(t) $, we can obtain its normalization by dividing the function by its absolute global maximum, that is, $ \hat{C}(t) = C(t) / \max(|C(t)|) $, provided that $ \max(|C(t)|) \ne 0 $.
@@
 
-% TODO
-
 This realization that the envelope of a wave is defined only at some of the wave's local extrema is crucial, as we can use it to generate a simplified representation of a signal that will make the algorithm more efficient by considering only the local extrema of a given signal.
 
-\section{Simplified Representation}
+\section{Simplified Representation} \label{sec:representation}
 
@@
 
-For this reason we define P as the set of the points $ \{P_0, P_1, \cdots, P_{m-1}\} $ where $ P_j = (i_j, \lvert w_j \lvert) $, that is, $ \lvert w_j \lvert $, the absolute value of the extrema of each pulse, becomes the ordinate of each point and $ i_j $, its original index, becomes the abscissa.
+For this reason we define P as the set of the points $ \{P_0, P_1, \cdots, P_{m-1}\} $ where $ P_j = (i_j, \lvert w_{i_j} \lvert) $, that is, $ \lvert w_{i_j} \lvert $, the absolute value of the extrema of each pulse, becomes the ordinate of each point and $ i_j $, its original index, becomes the abscissa.
 
@@
 While the exact relation between $ n $, the size of the original signal, and $ m $, the size of the set of points, is dependent on factors such as the frequency and sampling rate of the original wave, the inequality $ m \le n $ holds for all discrete signals.
-In practice, $ m \ll n $ for most real world discrete signals.
+In practice, $ m \ll n $ for most real world signals.
 
@@
   \caption{Example of a discrete wave \textbf{w} divided into pulses, of which the extrema are highlighted with a circle.
-P is the set of the points $ P_j = (i_j, \lvert w_j \lvert) $ representing the absolute value of those extrema.}
+P is the set of the points $ P_j = (i_j, \lvert w_{i_j} \lvert) $ representing the absolute value of those extrema.}
   \label{fig:Pulses}
@@
 
-\section{Mapping to Cartesian coordinates}
+\section{Mapping to the Cartesian coordinate system} \label{sec:mapping}
 
@@
 
-Geometrically, this coordinate system has a basis vector $ \mathbf{b} = \{ \mathbf{b}_x, \mathbf{b}_y \}$ with $ \mathbf{b}_x = (a, 0), \mathbf{b}_y = (0, i) $ where $ a $ is an amplitude unit such as decibel, a measure of the instantaneous sound pressure, volt or even a fraction of a fixed maximum value.
+Geometrically, we can think of these points as points in a vector space described by a non-standard basis vector $ \mathbf{b} = \{ \mathbf{b}_x, \mathbf{b}_y \}$ with $ \mathbf{b}_x = (a, 0), \mathbf{b}_y = (0, i) $ where $ a $ is an amplitude unit such as decibel, a measure of the instantaneous sound pressure, volt or even a fraction of a fixed maximum value.
 $ i $ represents the index of the sample, and is linked to time by the aforementioned relation $ i = t \ \text{fps} $.
 
-We are interested in finding a new basis vector that would make the geometric representation consistent, in the sense that pulse's amplitude $ \lvert w_j \lvert $, now mapped to the ordinates of a point, and a pulse's length, approximately equivalent to the horizontal distance between two consecutive points in the geometrical interpretation would define, on average, a square.
+We are interested in finding a change of basis formula from the original basis to the standard basis $ \mathbf{e} = \{ \mathbf{e}_x, \mathbf{e}_y \} = \{ (1,0),(0,1) \}$ of this vector space, making the geometric representation consistent.
 
-To achieve that, we divide the basis vector of the original vertical axis by the average of all ordinates, effectively cancelling the unit of the vertical components of the points.
-We could divide the basis vector of the horizontal axis by the average of the difference between the abscissa of a point and the abscissa of the immediately posterior, likewise eliminating the axis' unit.
+To achieve that, we divide the ordinate of each point by the average of all ordinates, effectively cancelling the unit of the vertical components of the points.
+If we proceeded similarly in relation to the horizontal components,
+the direct link between P and \textbf{w} that arises from the fact that the abscissas of the points in P are indices $ i $ of \textbf{w} would be lost.
 
-The drawback is that we would lose the direct link between P and \textbf{w} that arises from the fact that the abscissas of the points in P are indices $ i $ of \textbf{w}: we choose instead to leave the abscissa intact and multiply the now unitless vertical basis axis by the average of the difference between the abscissa of a point and the abscissa of the immediately posterior.
-In this way, both axes will have unit $ i $, and the relationship is preserved, making it easier to recover the envelope points later.
+It is best to leave the abscissas intact and multiply the now unitless vertical coordinates of each point by the average of the difference between the abscissa of a point and the abscissa of the immediately posterior.
+In this way, we can consider both axes as having unit $ i $, and the relationship is preserved, making it easier to recover the envelope points later.
 
-In this new coordinate system we have a basis vector $ \mathbf{b}^\prime = \{ \mathbf{b}^\prime_x, \mathbf{b}^\prime_y \}$ where $ \mathbf{b}^\prime_x = (i, 0) $ and $\mathbf{b}^\prime_y = (0, i) $, related to the original basis vector by the vertical scaling factor $ \beta $ as formalized in Eq.
-\ref{eq:Basis}:
+Considering $P_{\mathbf{b}}$ as a point of the set P represented by the original basis and $P_{\mathbf{e}}$ the same point as represented with the standard basis, we can formalize this change of basis as in \ref{eq:Basis}:
 
-\begin{subequations}  \label{eq:Basis} % TODO
+\begin{subequations}  \label{eq:Basis}
 \begin{align}
-\mathbf{b}^\prime_x &= \mathbf{b}_x  \\
-\beta &= \frac{\left( \frac{i_{m-1} - i_0}{m-1} \right)}{\left( \frac{\sum_{j=0}^{m-1} \lvert w_j \lvert}{m} \right)}  \\
-\mathbf{b}^\prime_y &= \beta \ \mathbf{b}_y 
+P_{\mathbf{b}} &= (x_1, x_2) \\
+\beta &= \frac{\left( \frac{i_{m-1} - i_m}{m-1} \right)}{\left( \frac{\sum_{j=0}^{m-1} \lvert w_j \lvert}{m} \right)}  \\
+P_{\mathbf{e}} &= (x_1, \beta x_2) \quad \forall \  P \in \text{P} \\ 
 \end{align}
@@
 
-The effect of this normalization can be seen in the points shown in Fig.\ref{fig:Coordinates}, shown in scale for both coordinate systems.
+The effect of this normalization can be seen in the points shown in Fig.\ref{fig:Coordinates}, shown in scale for both basis.
+We can consider our points now residing in the Cartesian coordinate system.
 
@@
     \includegraphics{images/04Coordinates.pdf}
-  \caption{The set of points P and the set V of vectors between two adjacent points of P, in both the original and normalized coordinate systems, shown in scale.}
+  \caption{The set of points P, in both the original (top) and standard (bottom) coordinate basis, shown in scale.}
   \label{fig:Coordinates}
@@
 
-Having established a map from the original space of the two-dimensional representation of a discrete wave to the Cartesian coordinate system and, thus, a way to consistently transform our simplified description of a discrete signal into a set of points in the plane, the shape of which can be estimated with the use of geometric algorithms.
+Having established a transformation from the original space of the two-dimensional representation of a discrete wave to the Cartesian coordinate system and, thus, a way to consistently transform our simplified description of a discrete signal into a set of points in the plane, the shape of which can be estimated with the use of geometric algorithms.
 
@@
 \begin{equation} \label{eq:Vectors}
-\text{V} = \{ v_0, v_1, v_{m-2} \} \quad \text{where} \quad v_k = P_{j+1} - P_j \quad \forall \quad 0 \le k \le m-2
+\text{V} = \{ v_0, v_1, \cdots, v_{m-2} \} \quad \text{where} \quad v_k = P_{j+1} - P_j \quad \forall \quad 0 \le k \le m-2
 \end{equation}
@@
 \ref{fig:DiscreteCurvature}.
-$ v_k $ represents the net displacement of the envelope from the extrema $ P_{j} $ to $ P_{j+1} $.
-Since we dont yet know which points of P are part of the envelope, instead of assuming the trajectory before $ P_{j} $ passes through
-$ P_{j-1} $, it is more prudent to assume that it can have passed anywhere in the line segment from 
-% TODO
-That is because all we can assume from $ v_k $ is that, in this current region of the envelope, it is able to a vertical displacement of $ v_{k,y} $ in the same time as it performs a horizontal displacement of $ v_{k,x} $.
-In average, we can consider the envelope entering from the horizontal direction.
+$ v_k $ represents the net displacement of the envelope from the extrema $ P_{j} = (x_j, y_j) $ to $ P_{j+1} = (x_{j+1}, y_{j+1}) $.
+Since we don't yet know which points of P are part of the envelope, instead of assuming that the trajectory before $ P_{j} $ passes through $ P_{j-1} $, it is more prudent to assume that it can have passed anywhere in the line segment from points $ (x_j - (x_{j+1} - x_j), y_j - (y_{j+1} - y_j)) $ to $ (x_j - (x_{j+1} - x_j), y_j + (y_{j+1} - y_j)) $, the bold line segment before  $ v_k $ in Fig.
+\ref{fig:DiscreteCurvature}.
 
-The rationale then is to find, for each vector $ v_k $, the radius $ r_k $ of the equivalent circle whose tangent has the same change in direction, in the same horizontal distance, as the vector of interest, as shown in Fig.
+That is because all we can assume from $ v_k $ is that, in this current region of the envelope, it is able to a vertical displacement of $ v_{k,y} $ in the same time as it performs a horizontal displacement of $ v_{k,x} $;
+
+Since this is true for all the vectors $ v_k \in \text{V} $, we can consider that, on average, the envelope insides horizontally at $ P_{j} $.
+
+The rationale then is to find, for each vector $ v_k $, the radius $ r_k $ of the equivalent circle whose tangent has the same change in direction, in the same horizontal distance, as the vector, as shown in Fig.
 \ref{fig:DiscreteCurvature}.
@@
 \end{equation}
+
+\section{Extracting the envelope}
 
@@
 
-The set E obtained via algorithm \ref{FindEnvelope} is a subset of P, and thus also a set of points.
+The set E obtained via algorithm \ref{FindEnvelope} is a subset of P, and thus also a set of points in the Euclidean space.
+
+Recall from section \ref{sec:mapping} that the vertical coordinates of these points are currently scaled by the change of basis described in Eq.
+\ref{eq:Basis}, in relation to the original signal as represented in his original basis.
+We could invert this transformation, but since we were careful to maintain the direct link between the horizontal coordinates in both basis, we can simply consider the horizontal coordinates of P as the indices of samples of the original discrete signal, whose absolute value are part of the envelope.
+
 The most straightforward way to transform those points into the vector $ \textbf{e} \in \mathbb{R}^n $ is via linear interpolation, the approach used in this work.
-% TODO
+
 It is worth noting, however, that many other procedures are available, both for the interpolation and the smooth approximation of points.
+Depending on the intended use of the envelope, k-curves interpolation \parencite{2017Yank}is a specially interesting approach, since it guarantees that the maximum curvature occurs at the interpolated points.
 
@@
 
+More formally, we have the set P of the absolute values of the local extrema of the original signal $ \mathbf{w} = (w_0, w_1, \cdots, w_{n-1}) $ that contains, by definition, the global extrema of $ \mathbf{w} $.
+The set E of the points that belong to the envelope is a subset of P. We can thus write Eq. \ref{eq:Req1}:
+
+\begin{subequations} \label{eq:Req1}
+\begin{align}
+\text{P} &= \{ P_0, P_1, \cdots, P_{m-1} \} \quad \text{where} \quad P_j = (i_j, \lvert w_{i_j} \lvert) \quad \forall \quad 0 \le j \le m-1 \\
+\text{E} &\subset \text{P} \implies \max (\text{E}) \le \max (\text{P}) = \max (|w_0|, |w_1|, \cdots, |w_{n-1}|)
+\end{align}
+\end{subequations}
+
+If the original signal is bounded, that is $ \max (|w_0|, |w_1|, \cdots, |w_{n-1}|) = b \in \mathbb{R} $, not only our envelope is bounded, as required, but is guaranteed to have the same bound $ b $.
+
 The second statement is that if the signal has a finite frequency range, that frequency range must not be exceeded by the envelope.
@@
 
-The fourth and last condition states that if a wave is multiplied by a constant, the envelope should also be multiplied by the same constant.
-That is also the case, again because the points that form the envelope are points of the original wave itself.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-In order to obtain a numeric indicator of the quality of an envelope, we construct a reference envelope, that trades off pratical applicability in favour of mathematical formality.
+The fourth and last condition states that if a wave is multiplied by a constant, the envelope should also be multiplied by the same constant.
+
+Recall from section \ref{mapping} that, before the extraction of the envelope, the original points undergo a change of basis to lie in the Cartesian coordinate system.
+After the extraction, the inverse operation is performed on the points of the Envelope.
+
+Consider a discrete signal $ \mathbf{w} = (w_0, w_1, \cdots, w_{n-1}) $ and the set P of its absolute extrema, represented in the standard basis, $ \text{P} = \{ P_0, P_1, \cdots, P_{m-1} \}, \quad P_j = (i_j, \beta \lvert w_{i_j} \lvert) $, where $ \beta $ is obtained from Eq.
+\ref{eq:Basis}.
+
+Given a constant $ \alpha > 0, \alpha \in \mathbb{R} $, we can define the scaled signal as $ \mathbf{w}^\prime = (\alpha w_0, \alpha w_1, \cdots, \alpha w_{n-1}) $.
+Observing, from Eq.
+\ref{eq:Basis}, that $ \alpha $ can be factored out of denominator of the summation, we have $ \beta^\prime = \beta / \alpha $ and the new set $ \text{P}^\prime = \{ P^\prime_0, P^\prime_1, \cdots, P^\prime_{m-1} \}, \quad P^\prime_j = (i_j, \frac{\beta}{\alpha} \lvert \alpha w_{i_j} \lvert) $ is equal to the set P for the signal before scaling.
+
+This guarantees that the same subset of points E will be identified as part of the envelope in both cases.
+In the respective original basis system, applying the inverse change of basis, those points are $ P_j = (i_j, \frac{\beta}{\beta} \lvert w_{i_j} \lvert)  \quad \forall \  P \in \text{E} $, for the original signal, and $ P^\prime_j = (i_j, \frac{\alpha}{\beta}\frac{\beta}{\alpha} \lvert \alpha w_{i_j} \lvert) $ for the scaled signal.
+
+In order to obtain a numeric indicator of the quality of an envelope, we construct a reference envelope, that trades off practical applicability in favour of mathematical formality.
 In \textcite{1992BruceCurves} the envelope is defined as the boundary of the region filled by a family of curves.
@@
 
-% TODO
-In practical applications, one seldom has access to a family of curve, instead being interested in obtaining the envelope of a single signal.
+In practical applications, one seldom has access to a family of curves, instead being interested in obtaining the envelope of a single signal.
 Nevertheless, an approximation can be obtained, by shifting the original signal along its initial position in the horizontal axis, in order to cover the time of one period of the signal.
@@
 In Fig. \ref{fig:Comparison_sinusoid} we illustrate the envelope extracted by the conventional algorithms, contrasting it with the result of the method presented in this work, for a simple sinusoid. 
-% TODO
-Note that none of the conventional algorithms complies with the third condition proposed by \textcite{1996Loughlinamplitude}, that states that the envelope of a periodic wave should be a straight line.
+Only the present algorithm and the one based on the Hilbert transform comply with the third condition proposed by \textcite{1996Loughlinamplitude}, that states that the envelope of a periodic wave should be a straight line.
 
@@
     \includegraphics{images/07Comparison_sinusoid.pdf}
-  \caption{Comparison of envelope detection algorithms for a simple sinusoid}
+  \caption{Comparison of envelope detection algorithms for a simple sinusoid. The envelope for the Hilbert transform and for the present work are coincident in the image.}
   \label{fig:Comparison_sinusoid}
@@
 
-For a recording of the key 33 of a grand piano, where parts of the sound are highly percussive and noisy, the Hilbert approach presents considerable oscillations, as can be seen in Fig.
+For a recording of the key 33 of a grand piano, where parts of the sound, specially at the start, are highly percussive and noisy, the Hilbert approach presents a considerable deviation, as can be seen in Fig.
 \ref{fig:Comparison_piano}.
@@
 
-We can see from the table \ref{table:Error_comparison} that the method presented here consistently outperforms the traditional approaches for all waves tested, yielding an error 33\% smaller, on average, than the other methods.
-It is also worth noting that the errors of the traditional methods are very similar between them, in agreement with the visualizations.
+We can see from the table \ref{table:Error_comparison} that the method presented here consistently outperforms the traditional approaches for all waves tested, yielding an error 
+33\% smaller, % TODO
+on average, than the other methods. It is also worth noting that the errors of the traditional methods are very similar between them, in agreement with the visualizations.
 
-The times taken for the algorithms to process each wave are shown in \ref{table:Time_comparison}. 
-The Python source for the tests, ass well as the samples used, are available at the repository dedicated to this work, in the file "Comparison of Envelope Algorithms.py".
+The times taken for the algorithms to process each signal are shown in \ref{table:Time_comparison}. 
+The Python source for the tests, ass well as the samples used, are available at the repository dedicated to this work, in the file “Comparison of Envelope Algorithms.py”.
 
@@
 
-\section{Superior and inferior envelopes}
+\section{Extensions}
 
-In practice, is not uncommon for a discrete wave, especially in the case of sound, to present somewhat different superior and inferior contours; in those cases, the algorithm here presented can be used independently in the positive and negative pulses defined by the original wave to define two “envelopes”, that we are going to call superior and inferior frontiers.
+In this section we comment briefly on some characteristics and extensions of the presented algorithm, in order to suggest possible applications beyond envelope extraction.
+One of the less obvious is, perhaps, its use in the segmentation of a signal into pseudo cycles.
+
+\subsection{Superior and inferior envelopes}
+
+In practice, is not uncommon for a discrete wave, especially in the case of sound, to present somewhat different superior and inferior contours; in those cases, the algorithm can be easily adapted to infer from the positive and negative pulses defined by the original wave to define two independent “envelopes”, that we are going to call superior and inferior frontiers.
+
+If we split P, as defined in section \ref{sec:representation}, into $\text{P}^+$, the set of the local maxima of the original signal and $\text{P}^-$, the set of its local minima, we can apply the remaining steps of the algorithm in those two sets independently, arriving at a positive and a negative frontier.
 
 Fig.
-\ref{fig:FullFrontiers} illustrates the frontiers of six diverse discrete sound waves, as well as an in-detail view of the highlighted segment for each signal.
+\ref{fig:FullFrontiers} illustrates these frontiers for six diverse discrete sound waves, as well as an in-detail view of the highlighted segment for each signal.
 All waves are records of physical sounds, chosen to represent the applicability of the algorithm in real-world scenarios.
+The implementation available at the PyPI repository provides an option to extract the frontiers.
 
@@
 
-The frontiers are satisfactorily detected in the case of harmonic and inharmonic sounds, and are robust in relation to the number of samples and the frequencies of the waves.
-That is a peculiarity of our algorithm: the traditional algorithms are unable to generate two distinct envelopes for the same wave.
+The frontiers were are satisfactorily detected in the case of harmonic and inharmonic sounds, and are robust in relation to the number of samples and the frequencies of the waves.
+It is worth noting that the positive and negative frontiers are generally smoother than the unified envelope and seems to conform better to the countours of the underlying digital signal.
 
-\section{Approximated location of pseudo-cycles}
-
-% TODO
+\subsection{Simplified spectral representation}
 
 The temporal envelope adds complexity to the spectral representation of a wave \parencite{2019TarjanoNeuro}, and an accurate description of the envelope, while describing the evolution of the instantaneous amplitude of a signal in time, would also simplify further spectral analysis.
+
 Fig.
@@
 
-Moreover, the algorithm developed in this work naturally divides a signal into its pseudo-cycles, pinpointing them in the time-domain, thus providing the building blocks for the reconstruction of the fine structure of the wave.
+Fig.
+\ref{fig:FourierBend} shows part the power spectra for both the original wave and the carrier obtained by the present algorithm, for the guitar bend sound in Fig.
+\ref{fig:Bend}, around its fundamental frequency.
+Two prominent frequencies can be seen, the initial and final notes of the bend.
+In the carrier, the other frequencies appear more smooth.
+
+\begin{figure}[H]
+  \centering
+    \includegraphics{images/12Fourier.pdf}
+  \caption{Fourier power spectrum for the wave and carrier of the sound of a guitar bend shown in Fig. \ref{fig:Bend}.}
+  \label{fig:FourierBend}
+\end{figure}
+
+\subsection{Approximated location of pseudo-cycles}
+
+The algorithm developed in this work naturally divides a signal into its pseudo-cycles, pinpointing them in the time-domain, as illustrated in Fig.
+\ref{fig:singer_voice}, for the sound of an alto singer singing a sustained note, as in Fig.
+\ref{fig:singer_voice}.
+
+The vertical lines mark the indices of the samples that belong to the positive frontier of the original signal, and the section between each vertical line isolates a pseudo cycle of the wave.
+Note that, although two local maxima exist in each pseudo cycle, only the largest was marked.
 
@@
     \includegraphics{images/09PCs.pdf}
-  \caption{Section of the lines formed by the positions of the positive and negative frontiers of the digital wave illustrated in \ref{fig:singer_voice}, as extracted by the algorithm presented, and their average.}
+  \caption{Section of the digital wave illustrated in \ref{fig:singer_voice}, with vertical lines at the indices of the samples that belong to the positive frontier.}
   \label{fig:PCs}
@@
 
-One can see in Fig.
-\ref{fig:PCs} a section of the plot of where the local maxima and minima occurs in the recording of an alto singer show in \ref{fig:singer_voice}.
-The average between the two is also shown.
-With this information, one can split a discrete wave into its pseudo-cycles, as is illustrated in Fig.
-\ref{fig:AvgPc}: It shows the superposition of this extracted pseudo-cycles, after being normalized by length, and their average.
+Therefore, using those indices, one can split a discrete wave into its pseudo cycles.
+One application of that is illustrated in Fig.
+\ref{fig:AvgPc}: It shows the superposition of the extracted pseudo cycles, after being normalized by length, and their average.
+They were also rotated to make their starting point more close to zero.
+
 This same average is shown shifted, to exemplify that approximate $C^0$ and $C^1$ continuity is achieved.
+This can be seen as the average waveform of the original discrete signal, and can be used, along with the information of its envelope, to generate an alternative representation of the wave, useful, for example, for compression and classification purposes.
 
@@
 
-By presenting a general approach for envelope detection, and eliminating the need for parameter tuning, we expect to facilitate further research at the many areas that rely on envelope detection techniques, and improve already existing algorithm that have envelope detection as an intermediary step.
+By presenting a general approach for envelope detection, and eliminating the need for parameter tuning, we expect to facilitate further research in the many areas that rely on envelope detection techniques, and improve already existing algorithms that have envelope detection as an intermediary step.
 
@@
 
-The present work also highlights some gaps in our current understanding and definition of temporal envelope and, in so doing, endeavours to motivate further research in this topic, that can ultimately give rise to a more mathematically sound envelope theory, especially concerning broadband waves; suggestions in this direction are also presented.
+The present work also highlights some gaps in our current understanding and definition of temporal envelope and, in so doing, endeavours to motivate further research in this topic, that can ultimately give rise to a more mathematically rigorous envelope theory, especially concerning broadband waves; suggestions in this direction were also presented.
 

### 2021-05-04 Papers/01_Pulses_Envelope/images/04Coordinates.py

@@ -50,7 +50,7 @@ Y = Yp * scaling
 '''============================================================================'''
 fig = make_subplots(
     rows=2, cols=1, shared_xaxes=False, vertical_spacing=0.01,
-    subplot_titles=("<b>Original coordinate system</b>", "<b>Normalized coordinate system</b>")
+    subplot_titles=("<b>Original basis</b>", "<b>Standard basis</b>")
     )
 
 FONT = dict(
@@ -79,10 +79,10 @@ fig.update_layout(
 
 fig.update_xaxes(showline=False, showgrid=False, zeroline=False, showticklabels=False, title=dict(font=FONT), row=1, col=1)
 fig.update_xaxes(showline=False, showgrid=False, zeroline=False, showticklabels=True, 
-title=dict(text="$ i \mapsto x $",font=FONT), row=2, col=1)
-fig.update_yaxes(showline=False, showgrid=False, zeroline=False, title=dict(text="Amplitude",font=FONT), row=1, col=1, tickvals=[i for i in range(0, 3, 2)])
+title=dict(text="$ x $",font=FONT), row=2, col=1)
+fig.update_yaxes(showline=False, showgrid=False, zeroline=False, title=dict(text="$ y $",font=FONT), row=1, col=1, tickvals=[i for i in range(0, 3, 2)])
 fig.update_yaxes(showline=False, showgrid=False, zeroline=False, 
-title=dict(text="$\\text{Amplitude} . \\beta \mapsto y $",font=FONT), row=2, col=1, tickvals=[i for i in range(0, 11, 2)])
+title=dict(text="$ y $",font=FONT), row=2, col=1, tickvals=[i for i in range(0, 11, 2)])
 
 fig.add_trace(
   go.Scatter(

### 2021-05-04 Papers/01_Pulses_Envelope/images/07Comparison.py

@@ -17,8 +17,8 @@ def butter_lowpass_filter(data, fps, cutoff = 10, order = 2):
 
 
 # name = "piano" # change the name here for any of the files in the "test_samples" folder
-# name = "sinusoid"
-name = "soprano"
+name = "sinusoid"
+# name = "soprano"
 
 '''==============='''
 ''' Read wav file '''
@@ -115,7 +115,7 @@ fig.add_trace(
     x=X,
     y=envY_hilbert,
     mode="lines",
-    line_shape='spline',
+    # line_shape='spline',
     line=dict(width=1.3, color="red"),
   )
 )
@@ -127,7 +127,7 @@ fig.add_trace(
     x=X,
     y=envY_smooth,
     mode="lines",
-    line_shape='spline',
+    # line_shape='spline',
     line=dict(width=1.2, color="blue"),
   )
 )
@@ -138,7 +138,7 @@ fig.add_trace(
     x=X,
     y=envY_lowpass,
     mode="lines",
-    line_shape='spline',
+    # line_shape='spline',
     line=dict(width=1.1, color="green"),
   )
 )
@@ -149,7 +149,7 @@ fig.add_trace(
     x=X,
     y=envY,
     mode="lines",
-    line_shape='spline',
+    # line_shape='spline',
     line=dict(width=1, color="black"),
   )
 )

### 2021-05-05 Papers/01_Pulses_Envelope/PulsesEnvelope.tex
progress

@@
 When the general shape of a set of points present prominent concavities, the convex hull overestimates the boundaries.
-To address this limitation \textcite{1983Edelsbrunnershape} introduced the concept of alpha shapes, that became known as the concave hull: a mathematically well-defined generalization of the convex hull of a finite set of points, closely related to the Delaunay triangulation and Voronoi diagrams of those points.
+To address this limitation \textcite{1983Edelsbrunnershape} introduced the concept of alpha shapes, also known as concave hull: a mathematically well-defined generalization of the convex hull of a finite set of points, closely related to the Delaunay triangulation and Voronoi diagrams of those points.
 
@@
 
-We are interested in finding a change of basis formula from the original basis to the standard basis $ \mathbf{e} = \{ \mathbf{e}_x, \mathbf{e}_y \} = \{ (1,0),(0,1) \}$ of this vector space, making the geometric representation consistent.
+We are interested in finding a change of basis formula from the original basis to the standard basis $ \mathbf{s} = \{ \mathbf{s}_x, \mathbf{s}_y \} = \{ (1,0),(0,1) \}$ of this vector space, making the geometric representation consistent.
 
@@
 
-Considering $P_{\mathbf{b}}$ as a point of the set P represented by the original basis and $P_{\mathbf{e}}$ the same point as represented with the standard basis, we can formalize this change of basis as in \ref{eq:Basis}:
+Considering $P_{\mathbf{b}}$ as a point of the set P represented by the original basis and $P_{\mathbf{s}}$ the same point as represented with the standard basis, we can formalize this change of basis as in \ref{eq:Basis}:
 
@@
 \beta &= \frac{\left( \frac{i_{m-1} - i_m}{m-1} \right)}{\left( \frac{\sum_{j=0}^{m-1} \lvert w_j \lvert}{m} \right)}  \\
-P_{\mathbf{e}} &= (x_1, \beta x_2) \quad \forall \  P \in \text{P} \\ 
+P_{\mathbf{s}} &= (x_1, \beta x_2) \quad \forall \  P \in \text{P} \\ 
 \end{align}
@@
 
-Translating the explanation of the alpha-shapes algorithm presented in \textcite{1994EdelsbrunnerThree} to our context, we can say that the points that are part of the envelope are those that can be touched by a circle of a given radius $ r $, that must not contain any point of the signal, coming from $ y \to \infty $ towards the Cartesian representation of the signal.
+A formal definition of alpha-shapes is given at \textcite{1983Edelsbrunnershape}.
+In the context of this work, gaining the intuition of the algorithm is more important, and thus the explanation of the alpha-shapes algorithm presented in \textcite{1994EdelsbrunnerThree} can be more insightful.
 
-% TODO formal definition of alpha shapes, maybe image
+The explanation, translated to our context, invites one to think of the plane were the samples are represented as a soft material, such as Styrofoam.
+The samples themselves are a harder material, such as rock, and occupy a fixed position in this plane.
+The alpha shapes algorithm consists of carving this plane from the outside of those points, with a circular eraser of fixed radius $ \alpha $.
+In the end of the process, the remaining Styrofoam that couldn't be erased due to the impeding rocks is the alpha-shape of this set of points.
 
-Intuitively, one can picture a circle of fixed radius being rolled above the signal, and marking the points it touches as envelope points.
+Alternatively, one can picture a circle of fixed radius being rolled above the signal, and marking the points it touches as envelope points.
 
@@
 \begin{equation} \label{eq:Curvature}
-r = \left( \frac{\sum\limits_{k=0}^{m-2} \left( \frac{ v_{k,y}}{v_{k,x} \sqrt{v_{k,x}^2 + v_{k,y}^2}} \right)}{m-2} \right)^{-1}
+\overline{r} = \left( \frac{\sum\limits_{k=0}^{m-2} \left( \frac{ v_{k,y}}{v_{k,x} \sqrt{v_{k,x}^2 + v_{k,y}^2}} \right)}{m-2} \right)^{-1}
 \end{equation}
@@
 
-In an effort to formalize the assessment of the quality of temporal envelopes, \textcite{1996Loughlinamplitude} proposed some theoretical conditions necessary, but not sufficient, to ensure the physical plausibility of an envelope.
-We comment briefly below that the algorithm here presented satisfies the four conditions presented.
+\subsection{Theoretical guarantees}
 
-% TODO %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+In an effort to formalize the assessment of the quality of amplitude and frequency modulations of a signal, \textcite{1996Loughlinamplitude} proposed some theoretical conditions necessary, but not sufficient, to ensure the physical plausibility of an envelope.
+
+We comment below on how those conditions translate to the discrete setting, and show that the algorithm here presented satisfies the four conditions presented.
+
 The first one states that, if a signal is bounded in magnitude, then its envelope should be, as well.
-Our algorithm complies with this requisite, as the envelope is composed of selected samples from the original signal itself.
+Our algorithm complies with this requirement, as the envelope is composed of selected samples from the original signal itself.
 
@@
 
-If the original signal is bounded, that is $ \max (|w_0|, |w_1|, \cdots, |w_{n-1}|) = b \in \mathbb{R} $, not only our envelope is bounded, as required, but is guaranteed to have the same bound $ b $.
+If the amplitude of the original signal is bounded, that is $ \max (|w_0|, |w_1|, \cdots, |w_{n-1}|) = b \in \mathbb{R} $, not only the amplitude of the envelope is bounded, as required, but is guaranteed to have the same bound $ b $.
 
-The second statement is that if the signal has a finite frequency range, that frequency range must not be exceeded by the envelope.
-This is only directly applicable in the continuous case, where this condition was proposed.
+The second statement dictates that if the signal has a finite frequency range, that frequency range must not be exceeded by the carrier wave.
 
-Nevertheless, if we picture both the wave and the envelope as samples from continuous functions, is easy to see that this is the case for our algorithm; since the envelope is composed by a subset of the samples of the original digital signal, in the limit where all the samples of the wave are also part of the envelope, we have that the maximum frequency for both is the same.
+To see that this condition holds for the proposed method, consider a discrete signal $ \mathbf{w} = (w_0, w_1, \cdots, w_{n-1}) $, where all the absolute values $ |P_j| $ of its local extrema are greater than one.
 
-In all other cases, the maximum frequency of the envelope will be smaller than that of the original wave.
-The range of frequencies of the envelope, then, will always lie between zero and the maximum frequency of the original wave.
+This requirement doesn't imply loss of generality since, for an arbitrary signal, we can assure this condition by dividing all samples by the value of the minimum absolute local extrema.
+
+The relation $ \min(\text{E}) \ge 1 $ thus holds, since $ \text{E} \subset \text{P}$.
+
+Provided that one uses a method that preserves the original bounds when interpolating the points in E, such as linear interpolation, we can be sure that all the points in the envelope vector $ \mathbf{e} = (e_0, e_1, \cdots, e_{n-1}) $ are also greater than one.
+
+Recalling Eq.
+\ref{eq:Envelope}, where the relation between the original signal, the carrier and the envelope was defined as $ \mathbf{w} = \mathbf{e} \odot \mathbf{c} $, we can write the magnitude of an arbitrary frequency of the original signal ($ m_\mathbf{w}(f) $) and, likewise, the magnitude of an arbitrary frequency of the carrier ($ m_\mathbf{c}(f) $) in terms of the Discrete Fourier Transform (DFT) as in Eq.
+\ref{eq:DFT}:
+
+\begin{subequations}\label{eq:DFT}
+\begin{align}
+m_\mathbf{w}(f) &= \sum_{k=0}^{n-1} \sqrt{\left( w_k \cos(2 \pi k f / n) \right)^2 + \left( w_k \sin(2 \pi k f / n) \right)^2} \\
+m_\mathbf{c}(f) &= \sum_{k=0}^{n-1} \sqrt{\left( \frac{w_k}{e_k} \cos(2 \pi k f / n) \right)^2 + \left( \frac{w_k}{e_k} \sin(2 \pi k f / n) \right)^2} 
+\end{align}
+\end{subequations}
+
+For any given frequency $ f $, both $ m_\mathbf{w}(f) $ and $ m_\mathbf{c}(f) $ are nonnegative.
+Also, since $ e_k \ge 1 \ \forall \ k, 0 \le k \le n - 1 $ we establish that $ 0 \le m_\mathbf{c}(f) \le m_\mathbf{w}(f) \ \forall \ f $ which is a more restrictive condition than the original: If $ \mathbf{w} $ has a finite frequency range, the magnitude of frequencies outside that range are zero, as are those of the same frequencies of the carrier $ \mathbf{c} $.
+
+% TODO %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 In the respective original basis system, applying the inverse change of basis, those points are $ P_j = (i_j, \frac{\beta}{\beta} \lvert w_{i_j} \lvert)  \quad \forall \  P \in \text{E} $, for the original signal, and $ P^\prime_j = (i_j, \frac{\alpha}{\beta}\frac{\beta}{\alpha} \lvert \alpha w_{i_j} \lvert) $ for the scaled signal.
+
+\subsection{Reference envelope}
 
@@
 
-\section{Comparison with traditional algorithms}
+\subsection{Comparison with traditional algorithms}
 
@@
 
-We can see from the table \ref{table:Error_comparison} that the method presented here consistently outperforms the traditional approaches for all waves tested, yielding an error 
-33\% smaller, % TODO
-on average, than the other methods. It is also worth noting that the errors of the traditional methods are very similar between them, in agreement with the visualizations.
+We can see from the table \ref{table:Error_comparison} that the method presented here consistently outperforms the traditional approaches for all waves tested, yielding an error of less than 2\% on average, of the error of the traditional methods.
 
 The times taken for the algorithms to process each signal are shown in \ref{table:Time_comparison}. 
-The Python source for the tests, ass well as the samples used, are available at the repository dedicated to this work, in the file “Comparison of Envelope Algorithms.py”.
+The Python source for the tests, as well as the samples used, are available at the repository dedicated to this work, in the file “Comparison of Envelope Algorithms.py”.
 

### 2021-05-06 Papers/01_Pulses_Envelope/PulsesEnvelope.tex
finished adding suggestions

@@
 
-To see that this condition holds for the proposed method, consider a discrete signal $ \mathbf{w} = (w_0, w_1, \cdots, w_{n-1}) $, where all the absolute values $ |P_j| $ of its local extrema are greater than one.
+To see that this condition holds for the proposed method, consider a discrete signal $ \mathbf{w} = (w_0, w_1, \cdots, w_{n-1}) $, where all the absolute values of its local extrema are greater than one.
 
@@
 
-% TODO %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+The third condition states that a physically plausible algorithm should, for a signal $ \mathbf{w} $ of constant amplitude $ a > 0, \ a \in \mathbb{R} $ and constant frequency, of the form $ a \cos(\omega i + \phi) $, result in a constant envelope $ \mathbf{e} $ of magnitude $ a $ and o a carrier wave $ \mathbf{c} $ of the form $ \cos(\omega i + \phi) $.
 
-The third condition states what, for a periodic signal, the envelope should be a straight line with the intercept equal to the amplitude of this signal.
-Since all the local maxima have the same amplitude in the case of a periodic wave, our envelope will indeed be a straight line, with the same amplitude as the original wave.
 An example of such behaviour can be observed in Fig.
-\ref{fig:Comparison_sinusoid}.
+\ref{fig:Comparison_sinusoid}, where the envelope of a simple sinusoid is shown.
+To see that this is always the case, note that all the extrema of a sinusoid are of the form $ a (\pm 1) $ and $ P_j = (i_j, \lvert a (\pm 1) \lvert) = (i_j, a) \ \forall P_j \in \text{P} $.
+As $ \text{E} \subset \text{P} $, we have that the envelope $ \mathbf{e} = (a, a, \cdots, a) $ and is, thus constant.
+Likewise, since $ \mathbf{w} = \mathbf{e} \odot \mathbf{c} $, we have $ c_i = a \cos(\omega i + \phi) / a = \cos(\omega i + \phi) \ \forall \ i, 0 \le i \le n - 1 $.
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-The fourth and last condition states that if a wave is multiplied by a constant, the envelope should also be multiplied by the same constant.
+The fourth and last condition states that if a wave is multiplied by a constant, its envelope should also be multiplied by the same constant.
 

### 2021-05-07 Papers/01_Pulses_Envelope/00ElsevierPulsesEnvelope.tex
moving to elsevier template

@@
 \documentclass{elsarticle}
-% \usepackage{lmodern} % Latin Modern Font
-% \usepackage[T1]{fontenc}
-% \usepackage{authblk}
-\usepackage{amsfonts} % \mathbb
+% \usepackage[natbib=true]{biblatex}
 \usepackage[margin=0.5in]{geometry}
@@
 % \usepackage{graphicx}
-\usepackage[backend=biber]{biblatex}
+% \usepackage[backend=biber]{biblatex}
 \usepackage{amsmath}
-
+\usepackage{amsfonts} % \mathbb
+\usepackage{amssymb}  % \therefore
 \usepackage{algorithm}
@@
 
-% \usepackage{hyperref}
-% \usepackage{hypcap}
-% \hypersetup{
-%   colorlinks=true, 
-%   allcolors=black,
-%   pdfauthor={Carlos Henrique Tarjano Santos},
-%   pdftitle={Robust Digital Envelope Estimation Via Geometric Properties of an Arbitrary Real Signal}
-% }
-
-\DeclareMathOperator{\sign}{sign}
-\bibliography{bibli}
+\bibliographystyle{elsarticle-num}
+\begin{document}
+\begin{frontmatter}
 
@@
 \begin{abstract} 
-% Background
 Despite being an elusive concept, the temporal amplitude envelope of a signal is essential for its complete characterization, being the primary information-carrying medium in spoken voice and telecommunications, for example.
-Intuitively, the temporal envelope can be understood as a slowly varying function that multiplies the signal, being responsible for its outer shape.
 Envelope detection techniques have applications in areas like medicine, sound classification and synthesis, seismology and speech recognition.
-% Problem | Objectives
-Nevertheless, a general approach to digital envelope detection of signals with rich spectral content doesn't exist, as most methods involve manual intervention, in the form of filter design, smoothing, and other specific design choices, based on prior knowledge of the signals under investigation.
-To address this problem, we propose an algorithm that uses intrinsic characteristics of a signal to estimate its envelope, eliminating the necessity of parameter tuning.
-% Methods | Procedure | Approach
-The approach here described draws inspiration from geometric concepts to estimate the temporal envelope of an arbitrary signal; specifically, a new measure of discrete curvature is used to obtain the average radius of curvature of a discrete wave, that will serve as a threshold to identify the wave's samples that are part of the envelope.
-% Results | Findings | Product
+Nevertheless, a general approach to envelope detection doesn't exist, as most methods involve manual intervention, in the form of filter design, smoothing, and other specific design choices based on prior knowledge of the signals under investigation.
+To address this problem, we propose an algorithm that uses characteristics of a signal to estimate its envelope, eliminating the necessity of parameter tuning.
+The method draws inspiration from geometric concepts; the discrete signal is mapped to a set of points in the Cartesian plane, were a new measure of discrete curvature is used to identify the samples that are part of the envelope.
+Due to its geometric nature, the method can optionally generate the superior and inferior envelopes of a given wave.
 The algorithm compares favourably with classic envelope detection techniques based on smoothing, filtering and the Hilbert Transform, besides being physically plausible.
-We provide visualizations of the envelope extracted via the algorithm for various real-world signals, with very diverse characteristics, such as voice, spoken and sang, and pitched and non-pitched musical instruments, and discuss some approaches to assess the quality of the obtained envelopes.
-A Python module implementing the algorithm was made available via the Python Package Index (PyPI); interactive visualizations of envelopes for a diverse range of digital waves, as well as the source code for the Python implementation, are available online.
-% Conclusions | Implications
-Besides the most direct applications of this work to audio classification and synthesis, we foresee impact in compression techniques and machine learning approaches to audio.
-We briefly discuss some potential paths in this direction.
-The discrete curvature definition presented could also be extended to three-dimensional settings, to improve shape detection algorithms based on alpha shapes.
+We provide visualizations of the envelope extracted for various real-world signals, with very diverse characteristics, such as voice, spoken and sang, and pitched and non-pitched musical instruments, and discuss some approaches to assess the quality of the obtained envelopes.
+An optimized implementation is available via the Python Package Index (PyPI); interactive visualizations and source code are available at the accompanying repository.
 \end{abstract}
@@
 
-\begin{document}
+\end{frontmatter}
 
@@
 \section{Introduction}
-Envelope detection, also known as demodulation, is ubiquitous in both analogue and digital signal processing \parencite{2011CaetanoImproved}.
-Nevertheless, the literature in this area is fragmented \parencite{2017LyonsDigital}.
-Besides, most envelope detection techniques are designed to account for very specific settings, like pure sinusoids with moderate noise content, a limitation that excludes most physical signals, as is the case of recorded sound, for example; that limitation arises in part due to the lack of a strict mathematical definition of a temporal envelope \parencite{2013Mengempirical}.
+Envelope detection, also known as demodulation, is ubiquitous in both analogue and digital signal processing (DSP) \cite{2011CaetanoImproved}.
+Nevertheless, the literature in this area is fragmented \cite{2017LyonsDigital}, with most envelope detection techniques being designed to account for very specific settings, like pure sinusoids with moderate noise content, for example.
 
-In many contexts, however, the temporal amplitude envelope of a signal plays a prominent role: according to \textcite{2017QiRelative}, for example, the envelope is at least as important as the fine structure of a sound wave in the context of the intelligibility of Mandarin tones.
-This is also the case for the English language \parencite{1995ShannonSpeech}, where even envelopes modulating mostly noise are still capable of conveying meaning.
+Those constraints exclude most physical signals, as is the case of recorded sound, for example; this problem arises in part due to the lack of a strict mathematical definition of a temporal envelope \cite{2013Mengempirical}.
 
-The envelope also helps to impart emotion and identity to the human voice \parencite{2018ZhuContributions} and music: Concert halls with envelope preserving characteristics are deemed more pleasant \parencite{2011LokkiEngaging}, for example.
+In many contexts, however, the temporal amplitude envelope of a signal plays a prominent role: according to \cite{2017QiRelative}, for example, the envelope is at least as important as the fine structure of a sound wave in the context of the intelligibility of Mandarin tones.
+This is also the case for the English language \cite{1995ShannonSpeech}, where even envelopes modulating mostly noise are still capable of conveying meaning.
 
-When dealing with broadband signals, approaches tailored to specific applications are prevalent, such as the one presented by \textcite{2014YangFast} for the distributed monitoring of fibre optic or the one formulated by \textcite{2018AssefModeling} in the context of medical ultrasound imaging.
+The temporal envelope also helps to impart emotion and identity to the human voice \cite{2018ZhuContributions} and music: Concert halls with envelope preserving characteristics are deemed more pleasant \cite{2011LokkiEngaging}, for example.
 
-The problem of envelope detection, normally associated with the field of digital signal processing, can be made equivalent to the geometric problem of defining the (outer) shape of a set of points in $ \mathbb{R}^2 $, provided a method for addressing the difference of units in the horizontal, time related, and vertical, amplitude related, axes of a digital signal is available, rendering the original DSP problem readily approachable via geometric techniques.
+Specially when dealing with broadband signals, approaches tailored to specific applications, exploiting a priori knowledge of the waves, are prevalent, such as the one presented by \cite{2014YangFast} for the distributed monitoring of fibre optic or the one formulated by \cite{2018AssefModeling} in the context of medical ultrasound imaging.
 
-One direct way of defining the shape of a set of points in $ \mathbb{R}^2 $ is using the concept of a convex hull: as defined in \textcite{2008BergComputational}, the convex hull of a set of points $ S $ in two-dimensional space is the smallest convex subset of the plane that contains all the points in $ S $.
-Convex hulls are unique, in the sense that a set of points define one and only one convex hull.
+The problem of envelope detection of a discrete signal, normally associated with the field of digital signal processing, can be made equivalent to the geometric problem of defining the shape of a set of points in $ \mathbb{R}^2 $.
 
-Generalizing that definition, \textcite{1983Edelsbrunnershape} introduced the concept of alpha shapes, that can be seen as a concave hull; a mathematically well-defined generalization of the convex hull of a finite set of points, closely related to the Delaunay triangulation and Voronoi diagrams of those points.
+Considering the standard 2d representation of a discrete wave, where the horizontal axis represents time and the vertical axis represents the amplitude, one can think of an individual sample as a point.
+Provided a coherent mapping from this space to the standard Cartesian plane is available, the original DSP problem can be transformed into the geometric problem of finding the shape of a set of points.
 
-The intuition behind the alpha shapes algorithm is that only points that can be touched by a circle of radius $ \alpha $ coming from an infinite distance towards the set of points of interest are part of the frontier, that is, of the concave hull of this set.
+One direct way of defining the shape of a set of points in $ \mathbb{R}^2 $ is using the concept of a convex hull, as defined in \cite{2008BergComputational}.
+The convex hull of a set of points $ S $ in two-dimensional space is the smallest convex subset of the plane that contains all the points in $ S $.
+Convex hulls are unique, in the sense that a given set of points define one and only one convex hull.
 
-Once a radius $ \alpha $ is defined, alpha shapes conserve the uniqueness property, in fact tending to the convex hull when $ \alpha \to \infty $.
-Consider, for example, the artificially generated discrete wave in figure \ref{fig:Hulls} below, where the links between consecutive samples were deemphasized to highlight the “set of points” nature of a signal: The convex hull and a concave hull, in the form of an alpha shape with a particular value of $ \alpha $ are illustrated.
+When the general shape of a set of points present prominent concavities, the convex hull overestimates the boundaries.
+To address this limitation \cite{1983Edelsbrunnershape} introduced the concept of alpha shapes, also known as concave hull: a mathematically well-defined generalization of the convex hull of a finite set of points, closely related to the Delaunay triangulation and Voronoi diagrams of those points.
 
-\begin{figure}[ht!]
+The intuition behind the alpha shapes algorithm is that only points that can be touched by a circle of radius $ \alpha $ coming from an infinite distance towards the baricenter of the set of points are part of the frontier.
+The polygon obtained linking those points forms the concave hull of this set.
+
+Once a radius $ \alpha $ is defined, alpha shapes exhibit the uniqueness property meaning that, for a given set of points and a given radius, only one concave hull exists.
+When the radius of the circle $ \alpha \to \infty $, concave hulls tend to the convex hulls.
+
+Consider, for example, the artificially generated discrete wave in Fig.
+\ref{fig:Hulls} below, where the links between consecutive samples were deemphasized to highlight the "set of points" nature of a signal: The convex hull and a concave hull, in the form of an alpha shape with a particular value of $ \alpha $ are illustrated.
+
+\begin{figure}[H]
   \centering
@@
 
-The alpha shapes approach is used in areas such as detection of features in images \parencite{2016VarytimidisAlpha}, reconstruction of surfaces from a cloud of points \parencite{2015WuAutomated} and spectroscopy \parencite{2019XuModeling}.
+The alpha shapes approach is used in areas such as detection of features in images \cite{2016VarytimidisAlpha}, reconstruction of surfaces from a cloud of points \cite{2015WuAutomated} and spectroscopy \cite{2019XuModeling}.
 
@@
 
-Other steps in the direction of translating geometric algorithms to the context of envelope detection were made by \textcite{2015YangSkeleton} via an algorithm based on the construction of a skeleton underlying the digital wave of interest, and also via the direct translation of computer vision methods to the task of envelope detection \parencite{2015YangRepresenting}.
+Other steps in the direction of translating geometric algorithms to the context of envelope detection were made by \cite{2015YangSkeleton} via an algorithm based on the construction of a skeleton underlying the digital wave of interest, and also via the direct translation of computer vision methods to the task of envelope detection \cite{2015YangRepresenting}.
 
-Following this path, we present a general approach to envelope detection, exploiting the intrinsic characteristics of a generic, spectrally complex wave, in order to completely avoid the need for manual intervention or parameter tuning.
+Following this path, we present a general approach to envelope detection, exploiting the intrinsic characteristics of a spectrally complex wave, in order to completely avoid the need for manual intervention or parameter tuning.
+
+% TODO %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+After defining what will be considered an envelope in the context of this work, in this section we explore some simplifications that can be applied to a discrete wave, defining the representation that will be used in the subsequent part of the work.
+
+We next explore the concept of discrete curvature, putting forth our own method, tailored to the definition of the equivalent circle of a wave; we end this section by showing how this circle can be used to identify the envelope of a discrete wave, in a similar approach to that of the alpha shapes theory, but taking advantage of some unique structure present in discrete waves.
 
@@
 
-\section{Methods}
-After defining what will be considered an envelope in the context of this work, in this section we explore some simplifications that can be applied to a discrete wave, defining the representation that will be used in the subsequent part of the work.
+Interactive visualizations are available from the repository prepared for this work \cite{2020TarjanoEnvelope}, that also contains an implementation of the method in pure Python.
+A Python module available at the Python Package Index (PyPI) installable via the command `pip install signal-envelope` for convenience, is also available.
+Besides, those working under Windows 64-bit can benefit from a specialized dynamic library coded in C++ that is automatically used within this module, and can be separately obtained via the release section in the repository.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-We next explore the concept of discrete curvature, putting forth our own method, tailored to the definition of the equivalent circle of a wave; we end this section by showing how this circle can be used to identify the envelope of a discrete wave, in a similar approach to that of the alpha shapes theory, but taking advantage of some unique structure present in discrete waves.
+\section{Temporal envelope}\label{sec:Characterization of a temporal envelope}
 
-\subsection{Characterization of a temporal envelope}
+The interest in the theory of envelope detection arose in the analog domain, with the widespread adoption of radio communications, where the problem is typically restricted to artificial waves, often with well-defined frequencies \cite{2011TurnerDemodulation}.
 
-The interest in the theory of envelope detection arose in the analog domain, with the widespread adoption of radio communications, where the problem is typically restricted to simple sinusoids, often with well-defined frequencies \parencite{2011TurnerDemodulation}.
+While there is no general formal definition of the envelope of broadband signals \cite{2014Yangnovel,2015Yangtheoretical,2019Jiaempiricalj}, the general intuition is that an envelope is a smooth curve situated as close as possible to the signal \cite{2019Jiaempiricalj}.
 
-In this context, the most mathematically sound definition of an envelope involves the representation of its underlying wave as an analytic signal, as introduced by Gabor in 1946 \parencite{2007HahnHistory}.
-In his work, \textcite{1946GaborTheory} applies the then relatively new mathematical machinery of the quantum mechanics to unify the time and frequency-domain representations of a wave, showing how the Hilbert transform could be applied to a real signal in order to obtain an equivalent complex signal, later known as the analytic signal.
-This analytic signal has the form $ A(t) = S(t) + \mathbb{H}(S(t)) \ \text{i} $ \parencite{2016HePraat} where $ S(t) $ is the original real signal.
+In this context, the most mathematically sound definition of an envelope involves the representation of its underlying wave as an analytic signal, as introduced by Gabor in 1946 \cite{2007HahnHistory}.
+In his work, \cite{1946GaborTheory} applies the then relatively new mathematical machinery of the quantum mechanics to unify the time and frequency-domain representations of a wave, showing how the Hilbert transform could be applied to a real signal in order to obtain an equivalent complex signal, later known as the analytic signal.
+
+This analytic signal has the form $ A(t) = S(t) + \mathbb{H}(S(t)) \ \text{i} $ \cite{2016HePraat} where $ S(t) $ is the original real signal.
 $ \mathbb{H}(S(t)) $, the Hilbert transform of the original signal, becomes the imaginary part of the analytic signal; the envelope of a signal thus represented can be straightforwardly obtained by the computation of the complex modulus of $ \mathbb{H}(S(t)) $.
 
-Despite its widespread use, and effectiveness in the context of narrowband signals, envelope detection techniques based on the Hilbert transform don't behave well for broadband signals \parencite{2012DauSpeech}, being even physically paradoxical in some cases \parencite{1996Loughlinamplitude}.
+Despite its widespread use, and effectiveness in the context of narrowband signals, envelope detection techniques based on the Hilbert transform don't behave well for broadband signals \cite{2012DauSpeech, 2019Jiaempiricalj}, being even physically paradoxical in some cases \cite{1996Loughlinamplitude, 2014Yangnovel}.
 
-Figure \ref{fig:AnalyticSignal} exemplifies the Hilbert envelope, for a pure sinusoid modulated by a polynomial, illustrating the difference in the shape of the envelope in the presence and absence of Gaussian white noise.
-It is readily noticeable from the figure that, as soon as noise is introduced in the original signal, it is reflected in the envelope.
+Fig.
+\ref{fig:AnalyticSignal} exemplifies the Hilbert envelope, for a pure sinusoid modulated by a polynomial, illustrating the difference in the smoothness of the envelope in the presence and absence of Gaussian white noise.
+It is readily noticeable from the Fig.
+\ref{fig:AnalyticSignal} that, as soon as noise is introduced in the original signal, it is directly reflected in the envelope.
 
-\begin{figure}[ht!]
+\begin{figure}[H]
   \centering
@@
 
-In general, the problem of envelope detection, or demodulation, can be interpreted as the task of, given a continuous wave $ W(t) $, decomposing it in two components $ E(t) $ and $ C(t) $, such that $ W(t) = E(t) C(t) $ \parencite{2011TurnerDemodulation}.
+For a real world signal, such as an alto singer sustaining a steady note as shown in Fig.
+\ref{fig:AnalyticSignalAlto} the Hilbert transform alone provides a poor envelope, since it retains most part of the frequencies of the underlying wave.
+With further processing, however, such as filtering, an output closer to the intuitive definition of envelope given in the beginning of this section can be obtained.
 
-$ E(t) $ represents the slow varying part of the wave, also known as the (temporal) envelope, modulator component, or amplitude modulation (AM) of $ W $, while $ C(t) $ models its fast varying part, called throughout the literature as its (temporal) fine structure, carrier component, or frequency modulation (FM).
+\begin{figure}[H]
+  \centering
+    \includegraphics{images/01AnalyticSignalAlto.pdf}
+  \caption{Representation of an alto singer sustaining a steady note and a comparison of the envelopes obtained by the Hilbert transform with and without further filtering.}
+  \label{fig:AnalyticSignalAlto}
+\end{figure}
 
-In the case of broadband signals, the problem of envelope detection of an arbitrary wave is ill-posed, in the sense that an infinite number of pairs of $ E(t) $ and $ C(t) $ can result in the same specific $ W(t) $ \parencite{2011TurnerDemodulation, 2013Mengempirical, 1996Loughlinamplitude}.
+In general, the problem of envelope detection, or demodulation, can be interpreted as the task of, given a continuous wave $ W(t) $, decomposing it in two components $ E(t) $ and $ C(t) $, such that $ W(t) = E(t) C(t) $ \cite{2011TurnerDemodulation}.
 
-We address this problem by assuming $ C(t) $ normalized between $ \{-1, 1\}$.
+$ E(t) $ represents the slow varying part of the wave, also known as the (temporal) envelope, modulator component, or amplitude modulation (AM) of $ W(t) $, while $ C(t) $ models its fast varying part, called throughout the literature as its (temporal) fine structure, carrier wave, or frequency modulation (FM).
+
+In the case of broadband signals, the problem of envelope detection of an arbitrary wave is ill-posed, in the sense that an infinite number of pairs of $ E(t) $ and $ C(t) $ can result in the same $ W(t) $ \cite{2011TurnerDemodulation, 2013Mengempirical, 1996Loughlinamplitude}.
+
+We address this problem by assuming $ C(t) $ normalized between $ \{-1, 1\}$ or, in other words $ |\max(C(t))| = 1 $.
 This assumption doesn't cause any loss of generality since, given an arbitrary $ C(t) $, we can obtain its normalization by dividing the function by its absolute global maximum, that is, $ \hat{C}(t) = C(t) / \max(|C(t)|) $, provided that $ \max(|C(t)|) \ne 0 $.
 
-In this work we are concerned with the discrete version of this problem: given a finite digital wave represented by the vector $ \textbf{w} \in \mathbb{R}^n $, instead of the continuous function $ W(t) $, obtaining the temporal envelope of $ \textbf{w} $.
+In this work we are concerned with the discrete version of this problem: given a finite digital wave represented by the vector $ \mathbf{w} \in \mathbb{R}^n $, instead of the continuous function $ W(t) $, obtaining the temporal envelope $ \mathbf{e} $ of $ \mathbf{w} $.
 
 The preceding definitions can be translated to this discrete scenario assuming that the discrete quantities arise from observing the continuous ones at regular time intervals.
-In that case, the equality $ i = t \ \text{fps} $ can be used to link both settings, where $ i $ is the index of each observation, $ t $ stands for the time in seconds, and fps is the frame rate, or the number of observations made in the period of a second.
+In that case, the equality $ i = t \ \text{fps} $ can be used to link both settings, where $ i $ is the index of each sample of the digital signal, $ t $ stands for the time in seconds, and fps is the frame rate, or the number of observations made in one second.
 We can thus define:
 
-\begin{align} \label{eq:Envelope}
-\begin{split}
-  & \mathbf{w}, \mathbf{e}, \mathbf{c} \in \mathbb{R}^n \\
-  & \mathbf{w} = (w_0, w_1, \cdots, w_{n-1}) \\
-  & \mathbf{e} = (e_0, e_1, \cdots, e_{n-1}) \\
-  & \mathbf{c} = (c_0, c_1, \cdots, c_{n-1}) \\
-  & \mathbf{w} = \mathbf{e} \odot \mathbf{c} \ \therefore \ w_i = e_i c_i \ \forall \ i, \ 0 \le i \le n-1
-\end{split}
+\begin{subequations}\label{eq:Envelope}
+\begin{align}
+  \text{Given:}& \\
+  \mathbf{w} & \in \mathbb{R}^n, \\
+  \mathbf{w} & = (w_0, w_1, \cdots, w_{n-1}) \\
+  \text{We define:}& \\
+  \mathbf{e}, \mathbf{c} & \in \mathbb{R}^n \\
+  \mathbf{e} & = (e_0, e_1, \cdots, e_{n-1}), w_i \ge 0 \ \forall \ i, \ 0 \le i \le n-1 \\  
+  \mathbf{c} & = (c_0, c_1, \cdots, c_{n-1}), |c_i| \le 1 \ \forall \ i, \ 0 \le i \le n-1 \\
+  \mathbf{w} & = \mathbf{e} \odot \mathbf{c} \ \therefore \ w_i = e_i c_i \ \forall \ i, \ 0 \le i \le n-1
 \end{align}
+\end{subequations}
 
 The $ \odot $ operator in \ref{eq:Envelope} stands for the Hadamard product, denoting element-wise multiplication of two vectors.
-Figure \ref{fig:Envelope} provides an example of the vectors just defined: \textbf{w} is obtained by modulating the normalized carrier wave \textbf{c} using the envelope \textbf{e}.
+Fig.
+\ref{fig:Envelope} provides an example of the vectors just defined: \textbf{w} is obtained by multiplying the normalized carrier wave \textbf{c} using the envelope \textbf{e}.
 
-Note from Figure \ref{fig:Envelope} that, at some local maxima, the envelope is well-defined, being in fact equal to the value of the wave $ w_i $, since $ c_i = 1 $ at those points, rendering $ e_i = 1 w_i $.
-
-The carrier wave \textbf{c} in the figure is periodic, and normalized between $ \{-1, 1\}$.
-Each cycle of the wave, then, reaches the value 1 only at its maximum making $ w_i $ equal to the value of the envelope at one point in each cycle.
-
-Image \ref{fig:Envelope}, along with the concave hull illustrated in figure \ref{fig:Hulls}, helps forming the intuition behind the method developed in this work: the idea is to identify the local extrema that touch the envelope, those being also the extrema of each (pseudo-)cycle.
-
-This is accomplished with the use of a circle with a defined radius, large enough to prevent it from reaching the smallest local extrema when "rolled" around the (geometric) representation of the wave.
-Only the extrema touched by the circle will be marked as part of the envelope.
-
-In the next section, a compatible geometric representation will be formulated, along with the necessary tools to estimate the average curvature of the wave and, consequently, the appropriate radius of the circle.
-
-\begin{figure}[ht!]
+\begin{figure}[H]
   \centering
     \includegraphics{images/02Envelope.pdf}
-  \caption{Example of a discrete wave \textbf{w} composed by the element-wise multiplication of an envelope \textbf{e} and a carrier \textbf{c}.
+  \caption{Example of a discrete wave \textbf{w} composed by the element-wise multiplication of a previously known envelope \textbf{e} and carrier \textbf{c}.
 The local extrema of \textbf{w} are highlighted with a circle.}
@@
 
-\subsection{Simplified Representation}
+In Fig.
+\ref{fig:Envelope} a signal arising from the multiplication of a known carrier wave \textbf{c} and a known envelope \textbf{e} is shown; this helps to illustrate a key concept: note from Fig.
+\ref{fig:Envelope} that, only at some local extrema, the envelope can be inferred without prior knowledge of the underlying carrier, being in fact equal to the value of the wave $ w_i $.
+$ c_i = 1 $ at those points, rendering $ e_i = 1 w_i $.
 
-Instrumental to the method here presented is the definition of a pulse: each time the value of \textbf{w} changes sign, that is, every time the discrete wave \textbf{w} crosses the horizontal axis, the beginning of a pulse is defined, with the next crossing defining its end.
-This definition is in line with the one presented in \textcite{national1996telecommunications}, where a pulse is defined as a rapid change in the amplitude of a signal, followed by a fast return to the baseline value; zero, in our case.
+This realization that the envelope of a wave is defined only at some of the wave's local extrema is crucial, as we can use it to generate a simplified representation of a signal that will make the algorithm more efficient by considering only the local extrema of a given signal.
 
-From \ref{fig:Envelope} and the discussion in the preceding chapter, we can see that only the maximum point of a positive pulse or the minimum point of a negative one can potentially be equal to the envelope and are, thus, our only points of interest.
+\section{Simplified Representation} \label{sec:representation}
+
+We start with the definition of a pulse: each time the value of $ w_i$ changes sign, that is, every time the amplitude of the discrete wave \textbf{w} crosses the horizontal axis, the beginning of a pulse is defined, with the next crossing defining its end.
+This definition is in line with the one presented in \cite{national1996telecommunications}, where a pulse is defined as a rapid change in the amplitude of a signal, followed by a fast return to the baseline value; zero, in our case.
+
+From \ref{fig:Envelope} and the discussion in Section \ref{sec:Characterization of a temporal envelope}, we can see that only the maximum point of a positive pulse or the minimum point of a negative pulse can potentially be equal to the envelope and are, thus, our only points of interest.
 We can then proceed, for the rest of the method, considering only those points, to great computational economy.
 
-For this reason we define P as the set of the points $ \{P_0, P_1, \cdots, P_{m-1}\} $ where $ P_j = (i_j, \lvert w_j \lvert) $, that is, $ \lvert w_j \lvert $, the absolute value of the extrema of each pulse, becomes the ordinate of each point and $ i_j $, its original index, becomes the abscissa.
+For this reason we define P as the set of the points $ \{P_0, P_1, \cdots, P_{m-1}\} $ where $ P_j = (i_j, \lvert w_{i_j} \lvert) $, that is, $ \lvert w_{i_j} \lvert $, the absolute value of the extrema of each pulse, becomes the ordinate of each point and $ i_j $, its original index, becomes the abscissa.
 
-This relation is illustrated in figure \ref{fig:Pulses}. We call it P to emphasize the connection with the pulses of \textbf{w}.
+This relation is illustrated in Fig. \ref{fig:Pulses}. We call it P to emphasize the connection with the pulses of \textbf{w}.
+While the exact relation between $ n $, the size of the original signal, and $ m $, the size of the set of points, is dependent on factors such as the frequency and sampling rate of the original wave, the inequality $ m \le n $ holds for all discrete signals.
+In practice, $ m \ll n $ for most real world signals.
 
-\begin{figure}[ht!]
+\begin{figure}[H]
   \centering
@@
 
-The points in P are not fit for geometric interpretation yet, since the abscissa and ordinate of their orthogonal coordinate system have different units: in the vertical axis, we have a unit related to the instantaneous intensity of the wave, while in the horizontal axis we have the index at which each extremum occurs, ultimately a time-related unit.
+\section{Mapping to the Cartesian coordinate system} \label{sec:mapping}
 
-Specifically, this coordinate system has a basis vector $ \mathbf{b} = \{ \mathbf{b}_x, \mathbf{b}_y \}$ with $ \mathbf{b}_x = (a, 0), \mathbf{b}_y = (0, i) $ where $ a $ is an amplitude unit such as decibel, a measure of the instantaneous sound pressure, volt or even a fraction of a fixed maximum value.
+The points in P are not fit for geometric interpretation yet, since the abscissa and ordinate of their orthogonal coordinate system have different units: in the vertical axis, we have a unit related to the instantaneous amplitude of the wave, while in the horizontal axis we have the index at which each extremum occurs, ultimately a time-related unit.
+
+Geometrically, we can think of these points as points in a vector space described by a non-standard basis vector $ \mathbf{b} = \{ \mathbf{b}_x, \mathbf{b}_y \}$ with $ \mathbf{b}_x = (a, 0), \mathbf{b}_y = (0, i) $ where $ a $ is an amplitude unit such as decibel, a measure of the instantaneous sound pressure, volt or even a fraction of a fixed maximum value.
 $ i $ represents the index of the sample, and is linked to time by the aforementioned relation $ i = t \ \text{fps} $.
 
-We are interested in finding a coordinate system where a pulse's amplitude $ \lvert w_j \lvert $ and length would define, on average, a square.
+We are interested in finding a change of basis formula from the original basis to the standard basis $ \mathbf{s} = \{ \mathbf{s}_x, \mathbf{s}_y \} = \{ (1,0),(0,1) \}$ of this vector space, making the geometric representation consistent.
 
-To achieve that, we divide the basis vector of the original vertical axis by the average of all ordinates, effectively cancelling the unit of the vertical components of the points.
-We could divide the basis vector of the horizontal axis by the average of the difference between the abscissa of a point and the abscissa of the immediately posterior, likewise eliminating the axis' unit.
+To achieve that, we divide the ordinate of each point by the average of all ordinates, effectively cancelling the unit of the vertical components of the points.
+If we proceeded similarly in relation to the horizontal components,
+the direct link between P and \textbf{w} that arises from the fact that the abscissas of the points in P are indices $ i $ of \textbf{w} would be lost.
 
-The drawback is that we would lose the direct link between P and \textbf{w} that arises from the fact that the abscissas of the points in P are indices $ i $ of \textbf{w}: we choose instead to leave the abscissa intact and multiply the now unitless vertical basis axis by the average of the difference between the abscissa of a point and the abscissa of the immediately posterior.
-In this way, both axes will have unit $ i $, and the relationship is preserved, making it easier to recover the envelope points later.
+It is best to leave the abscissas intact and multiply the now unitless vertical coordinates of each point by the average of the difference between the abscissa of a point and the abscissa of the immediately posterior.
+In this way, we can consider both axes as having unit $ i $, and the relationship is preserved, making it easier to recover the envelope points later.
 
-In this new coordinate system we have a basis vector $ \mathbf{b}^\prime = \{ \mathbf{b}^\prime_x, \mathbf{b}^\prime_y \}$ where $ \mathbf{b}^\prime_x = (i, 0) $ and $\mathbf{b}^\prime_y = (0, i) $, related to the original basis vector by equation \ref{eq:Basis}:
+Considering $P_{\mathbf{b}}$ as a point of the set P represented by the original basis and $P_{\mathbf{s}}$ the same point as represented with the standard basis, we can formalize this change of basis as in \ref{eq:Basis}:
 
-\begin{align} \label{eq:Basis}
-\begin{split}
-\mathbf{b}^\prime_x &= \mathbf{b}_x  \\
-\mathbf{b}^\prime_y &= \frac{\left( \frac{i_{m-1} - i_0}{m-1} \right)}{\left( \frac{\sum_{j=0}^{m-1} \lvert w_j \lvert}{m} \right)} \mathbf{b}_y 
-\end{split}
+\begin{subequations}  \label{eq:Basis}
+\begin{align}
+P_{\mathbf{b}} &= (x_1, x_2) \\
+\beta &= \frac{\left( \frac{i_{m-1} - i_m}{m-1} \right)}{\left( \frac{\sum_{j=0}^{m-1} \lvert w_j \lvert}{m} \right)}  \\
+P_{\mathbf{s}} &= (x_1, \beta x_2) \quad \forall \  P \in \text{P} \\ 
 \end{align}
+\end{subequations}
 
-The effect of this normalization can be seen in the points shown in figure \ref{fig:Vectors}, shown in scale in both coordinate systems.
-The set V of the vectors from each point of P to the next is also shown in the picture, and it will be used to estimate the average curvature of P.
-We define this set in equation \ref{eq:Vectors} below.
+The effect of this normalization can be seen in the points shown in Fig.\ref{fig:Coordinates}, shown in scale for both basis.
+We can consider our points now residing in the Cartesian coordinate system.
 
-\begin{equation} \label{eq:Vectors}
-\text{V} = \{ v_0, v_1, v_{m-2} \} \quad \text{where} \quad v_k = P_{j+1} - P_j \quad \forall \quad 0 \le k \le m-2
-\end{equation}
+\begin{figure}[H]
+  \centering
+    \includegraphics{images/04Coordinates.pdf}
+  \caption{The set of points P, in both the original (top) and standard (bottom) coordinate basis, shown in scale.}
+  \label{fig:Coordinates}
+\end{figure}
 
-\begin{figure}[ht!]
+\section{Discrete Curvature Estimation}
+
+Having established a transformation from the original space of the two-dimensional representation of a discrete wave to the Cartesian coordinate system and, thus, a way to consistently transform our simplified description of a discrete signal into a set of points in the plane, the shape of which can be estimated with the use of geometric algorithms.
+
+The goal of those algorithms, such as the convex hull and the alpha-shapes, is to discriminate which points of a set are part of its frontier.
+
+A formal definition of alpha-shapes is given at \cite{1983Edelsbrunnershape}.
+In the context of this work, gaining the intuition of the algorithm is more important, and thus the explanation of the alpha-shapes algorithm presented in \cite{1994EdelsbrunnerThree} can be more insightful.
+
+The explanation, translated to our context, invites one to think of the plane were the samples are represented as a soft material, such as Styrofoam.
+The samples themselves are a harder material, such as rock, and occupy a fixed position in this plane.
+The alpha shapes algorithm consists of carving this plane from the outside of those points, with a circular eraser of fixed radius $ \alpha $.
+In the end of the process, the remaining Styrofoam that couldn't be erased due to the impeding rocks is the alpha-shape of this set of points.
+
+Alternatively, one can picture a circle of fixed radius being rolled above the signal, and marking the points it touches as envelope points.
+
+To employ an alpha shapes inspired algorithm in this set of points, however, the radius $ r $ of such circle must be estimated, and for that a measure of discrete curvature is needed.
+
+Discrete curvature estimation is an important task in image processing \cite{2010FleischmannNovel} for which no default definition exists.
+
+The two general approaches are the derivation of direct methods that use characteristics of the discrete curve to calculate the curvature, or the use of the curvature of a smooth, continuous curve fitted to the original discrete curve \cite{2001CoeurjollyDiscrete}.
+
+Concerning direct methods \cite{2014CarrollSurvey}, for example, derive three such definitions based on the approximation of a circle by an inscribed, centered and circumscribed polygon.
+In the context of three-dimensional meshes, \cite{2016VasaMultivariate} evaluate a range of existing estimators from a multivariate point of view.
+
+Those approaches define the discrete curvature in relation to the vertices of a discrete curve (or mesh, in the three-dimensional case).
+Instead, we are interested in the change of direction, what the term curvature ultimately means, between two adjacent points.
+
+Since we are dealing with a potentially large set of points, it is important for this measure to act locally, to assure $ O(n) $ complexity.
+
+Thus, the set V of the vectors from each point of P to the next, shown in the Fig.
+\ref{fig:Vectors}, will be used to estimate the average curvature of P: We will calculate the equivalent circle for each vector, and average the values thus obtained to infer the average curvature of the whole discrete curve.
+
+\begin{figure}[H]
   \centering
     \includegraphics{images/04Vectors.pdf}
-  \caption{The set of points P and the set V of vectors between two adjacent points of P, in both the original and normalized coordinate systems, shown in scale.}
+  \caption{The set of points P and the set V of vectors between two adjacent points of P, in the normalized coordinate system.}
   \label{fig:Vectors}
@@
 
-\subsection{Discrete Curvature Estimation}
+We formalize the set of vectors V in Eq. \ref{eq:Vectors} below.
 
-Translating the explanation of alpha shapes in \textcite{1994EdelsbrunnerThree} to our context, we can say that the points that are part of the envelope are those touched by a circle of a given radius $ r $, coming from $ y \to \infty $ towards the signal, that is not allowed to contain any point of the signal; Intuitively, one can picture a circle being rolled above the signal, and marking the points it touches as envelope points.
+\begin{equation} \label{eq:Vectors}
+\text{V} = \{ v_0, v_1, \cdots, v_{m-2} \} \quad \text{where} \quad v_k = P_{j+1} - P_j \quad \forall \quad 0 \le k \le m-2
+\end{equation}
 
-To infer the appropriate radius $ r $ of such circle a measure of discrete curvature is needed.
-Discrete curvature estimation is an important task in image processing \parencite{2010FleischmannNovel} for which no default definition exists.
+\subsection{The Equivalent Circle Approach}
 
-The two possible general approaches are the derivation of direct methods, that use characteristics of the discrete wave to calculate the curvature, or the use of the curvature of a smooth, continuous curve fitted to the discrete wave \parencite{2001CoeurjollyDiscrete}.
+We thus proceed to define a discrete curvature measure over the edges of a discrete curve.
+To that end we are going to apply the definition of smooth curvature as the rate of change of the unit tangent to a curve.
+This definition is convenient, since it is equivalent to that of the osculating circle \cite{2016VasaMesh}, the value we are interested in finding.
 
-Concerning direct methods \textcite{2014CarrollSurvey}, for example, derive three such definitions based on the approximation of a circle by an inscribed, centered and circumscribed polygon.
-In the context of three-dimensional meshes, \textcite{2016VasaMultivariate} evaluate a range of existing estimators from a multivariate point of view.
+One of those edges, in the form of the vector $ v_k \in \text{V} $, is singled out in Fig.
+\ref{fig:DiscreteCurvature}.
+$ v_k $ represents the net displacement of the envelope from the extrema $ P_{j} = (x_j, y_j) $ to $ P_{j+1} = (x_{j+1}, y_{j+1}) $.
+Since we don't yet know which points of P are part of the envelope, instead of assuming that the trajectory before $ P_{j} $ passes through $ P_{j-1} $, it is more prudent to assume that it can have passed anywhere in the line segment from points $ (x_j - (x_{j+1} - x_j), y_j - (y_{j+1} - y_j)) $ to $ (x_j - (x_{j+1} - x_j), y_j + (y_{j+1} - y_j)) $, the bold line segment before $ v_k $ in Fig.
+\ref{fig:DiscreteCurvature}.
 
-Those approaches define the discrete curvature in relation to the vertices of a wave (or mesh, in the three-dimensional case).
-In our particular case, it wouldn't make sense to talk about the curvature of a single pulse, since it is what our points ultimately represent.
-Instead, we are interested in the change of direction, what the term curvature ultimately means, between two adjacent pulses, as the vectors in the set V suggests.
+That is because all we can safely assume from $ v_k $ is that, in the immediate vicinity of the vector, the envelope is capable of a vertical displacement of $ v_{k,y} $ in the same time as it performs a horizontal displacement of $ v_{k,x} $;
 
-We thus proceed to define a discrete curvature measure over the edges of a discrete wave.
-To that end we are going to apply the definition of smooth curvature as the rate of change of the unit tangent to a curve, noting that this is equivalent to that of the osculating circle \parencite{2016VasaMesh}.
+Since this is true for all the vectors $ v_k \in \text{V} $, we can consider that, on average, the envelope insides horizontally at $ P_{j} $.
 
-\subsubsection{The Equivalent Circle Approach}
-
-The rationale is to find, for each vector $ v_k \in \text{V} $, the radius $ r_k $ of the equivalent circle whose tangent has the same change in direction, in the same horizontal distance, as the vector of interest, as shown in figure \ref{fig:DiscreteCurvature}.
+The rationale then is to find, for each vector $ v_k $, the radius $ r_k $ of the equivalent circle whose tangent has the same change in direction, in the same horizontal distance, as the vector, as shown in Fig.
+\ref{fig:DiscreteCurvature}.
 The average curvature of P will then be obtained by the average of all radii $ r_k $.
 
-From figure \ref{fig:DiscreteCurvature} is easy to see that $ r_k = v_{k,x} \sin(\theta_k)$ where $ v_{k,x} $ is the component of $ v_k $ in the horizontal direction; $ \theta_k $ itself can be obtained using the slope $ m_0 = 0 $ of a line in the horizontal direction and the slope $ m_k = v_{k,y} / v_{k,x} $ defined by $ v_k $ via the equality $ \theta_k = \arctan \big( (m_k - m_0) / (1 + m_k m_0) \big) $.
+From Fig.
+\ref{fig:DiscreteCurvature} one can see that $ r_k = v_{k,x} / \sin(\theta_k) $ where $ v_{k,x} $ is the component of $ v_k $ in the horizontal direction; $ \theta_k $ itself can be obtained using the slope $ m_0 = 0 $ of the horizontal line and the slope $ m_k = v_{k,y} / v_{k,x} $ defined by $ v_k $ using the equality $ \theta_k = \arctan \big( (m_k - m_0) / (1 + m_k m_0) \big) $.
 
-\begin{figure}[ht!]
+\begin{figure}[H]
   \centering
@@
 
-The radius of the circle that represents the average curvature of V (and P), and will be used to obtain the envelope of the discrete wave \textbf{w} can be obtained via equation \ref{eq:Radius}:
+The radius $ r_k $ of the circle equivalent to each $ v_k $ is defined in Eq.
+\ref{eq:Radius}.
+This equation has the desirable property that $ v_{k,y} \to 0 \implies v_{k,y} \to \infty $, that is, the radius tends to infinity when no vertical oscillation occurs, effectively turning the circle into a line.
+
+The radius is also directly proportional to $ v_{k,x} $, making the radius larger the more the points are horizontally far from each other.
 
 \begin{equation} \label{eq:Radius}
-\begin{split}
-r = \frac{\sum\limits_{k=0}^{m-2} \left( \frac{v_{k,x} \sqrt{v_{k,x}^2 + v_{k,y}^2}}{ v_{k,y}} \right)}{m-2}
-\end{split}
+  r_k = \frac{v_{k,x} \sqrt{v_{k,x}^2 + v_{k,y}^2}}{ v_{k,y}}
 \end{equation}
 
+For calculation purposes, however, this can pose a problem, as a zero could make its way into the denominator of a fraction; in the actual algorithm, the quantity calculated for each vector is the curvature itself, that is, the inverse of each radius $ r_k $.
+In this way, the problem is avoided, since $ v_{k,x} $ is guaranteed to be greater than zero.
+The average of curvatures is then inverted in order to obtain the average radius $ \overline{r} $, as illustrated in Eq.
+\ref{eq:Curvature}.
+
+\begin{equation} \label{eq:Curvature}
+\overline{r} = \left( \frac{\sum\limits_{k=0}^{m-2} \left( \frac{ v_{k,y}}{v_{k,x} \sqrt{v_{k,x}^2 + v_{k,y}^2}} \right)}{m-2} \right)^{-1}
+\end{equation}
+
+\section{Extracting the envelope}
+
 We need now to construct an algorithm, using the radius just obtained, to identify the points that belong to the envelope.
-Generally, the algorithm for the alpha shapes approach resorts to the construction of the Delaunay triangulation of the set of points, that is latter filtered to contain only the outer edges.
-We can adopt a more straightforward method, as our points have a defined structure:
+Generally, alpha-shapes based algorithms resort to the construction of the Delaunay triangulation of the set of points, that is latter filtered to contain only the outer edges.
+We can adopt a more straightforward method, as our points have a sequential structure.
+
+It is worth noting that, although that is the case throughout this work, the original signal doesn't need to be sampled at equal intervals.
 
@@
 For all circles of center $ x_o, y_o $ and radius $ r $ that can be constructed to pass through any two different points of P, that is, $ \forall \quad P_b, P_c $ such that $ (P_{b,x} - x_o)^2 + (P_{b,y} - y_o)^2 = (P_{c,x} - x_o)^2 + (P_{c,y} - y_o)^2 = r^2 $, $ P_a $ will be part of the set $ E $ of the points belonging to the envelope if and only if none of those circles contain $ P_a $, or $ P_a \in \text{E} \leftrightarrow (P_{b,x} - x_o)^2 + (P_{b,y} - y_o)^2 > r^2 $.
-Summarizing this, we have equation \ref{eq:InEnvelope}:
+Summarizing this, we have Eq.
+\ref{eq:InEnvelope}:
 
-\begin{align} \label{eq:InEnvelope}
-\begin{split}
+\begin{subequations} \label{eq:InEnvelope}
+\begin{align}
 & P_a \in \text{E} \leftrightarrow (P_{b,x} - x_o)^2 + (P_{b,y} - y_o)^2 > r^2, \\
@@
 & P_a, P_b, P_c \in \text{P}, \quad P_a \ne P_b \ne P_c \quad \text{and} \quad x_o, y_o \in \mathbb{R}, \quad 0 \le y_o < +\infty
-\end{split}
 \end{align}
+\end{subequations}
 
 The algorithm follows directly from the definition in \ref{eq:InEnvelope} after noting that the first and last members of P will always be part of the envelope, since they are reachable by a circle of any radius $ r $.
-From $ P_0 $, our first pivot point, we construct circles passing through $ P_1, P_2, \cdots, P_d $ until a circle that doesn't contain any point in P is found; $ P_d $ then becomes the new pivot point and is included in E, and this procedure continues until $ P_{m-1} $ is reached.
+From $ P_0 $, our first pivot point, we construct circles passing through $ P_1, P_2, \cdots, P_d $ until a circle that doesn't contain any point in P is found; $ P_d $ then becomes the new pivot point and is included in E, and this process continues until $ P_{m-1} $ is reached.
 
 The procedure is formalized in the algorithm \ref{FindEnvelope}.
-We don't provide similar algorithmic descriptions of the preceding steps since the Python implementation can easily serve as pseudocode.
+We don't provide similar algorithmic descriptions of the preceding steps since the Python implementation available at \cite{2020TarjanoEnvelope} can easily serve as pseudocode.
 
@@
 
-The set E obtained via algorithm \ref{FindEnvelope} is a set of points.
+The set E obtained via algorithm \ref{FindEnvelope} is a subset of P, and thus also a set of points in the Euclidean space.
+
+Recall from section \ref{sec:mapping} that the vertical coordinates of these points are currently scaled by the change of basis described in Eq.
+\ref{eq:Basis}, in relation to the original signal as represented in his original basis.
+We could invert this transformation, but since we were careful to maintain the direct link between the horizontal coordinates in both basis, we can simply consider the horizontal coordinates of P as the indices of samples of the original discrete signal, whose absolute value are part of the envelope.
+
 The most straightforward way to transform those points into the vector $ \textbf{e} \in \mathbb{R}^n $ is via linear interpolation, the approach used in this work.
+
 It is worth noting, however, that many other procedures are available, both for the interpolation and the smooth approximation of points.
+Depending on the intended use of the envelope, k-curves interpolation \cite{2017Yank}is a specially interesting approach, since it guarantees that the maximum curvature occurs at the interpolated points.
 
-\section{Results and Discussion}
-Given the lack of a precise definition of what an envelope should be, we begin this section suggesting an empirical metric, based on the behaviour of the carrier wave \textbf{c}.
-We also comment on some metrics presented in the literature, and compare our algorithm with traditional demodulation approaches, both in relation to execution time and accuracy; to this last end, a numerical indicator is also introduced.
+\section{Quality of an envelope}\label{sec:quality}
 
-We also illustrate how the method here presented can be used to isolate not only a single envelope of the wave, but the superior and inferior envelopes as well, that we call frontiers, ending the section with a suggestion on how the proposed algorithm can be useful beyond envelope detection, by identifying the approximate locations of the pseudo-cycles in a quasi-periodic wave.
+The lack of a precise definition of what an envelope is makes the assessment of the quality of an envelope a difficult task, for which there are no agreed-upon objective metrics.
+Recalling Fig.
+\ref{fig:Envelope}, however, and the accompanying discussion, if one divides, element-wise, the original signal by the envelope, the carrier wave thus obtained is expected to be bounded between -1 and 1.
 
-Due to an assortment of factors, such as the unconventional approach presented in this work and the intrinsic difficulty in visualizing discrete waves given their high dimensionality, the authors deemed helpful to provide a companion website with interactive visualizations \parencite{2020TarjanoEnvelope}.
+Figs.
+\ref{fig:spoken_voice}, \ref{fig:singer_voice} and \ref{fig:Bend} present discrete waves \textbf{w}, their extracted envelope \textbf{e}, and the inferred carrier \textbf{c}, obtained after dividing, element-wise, the original signal by its extracted.
 
-An implementation of the method in pure Python, fully functional, can also be found at the repository dedicated to this work as well as a Python module available at the Python Package Index (PyPI) and conveniently installable via the command `pip install signal-envelope`.
-Besides, those working under Windows 64-bit can benefit from a specialized dynamic library coded in C++ that is automatically used within this module.
-
-\subsection{Quality of an envelope}
-
-There are no agreed-upon objective metrics to assess the quality of an envelope.
-Recalling figure \ref{fig:Envelope}, however, and the accompanying discussion, an intuitive way is to consider the carrier wave \textbf{c}, and how well it seems to be bounded by the interval $ \{-1, 1\} $.
-
-To illustrate how the algorithm performs on such metric, we provide the following figures of original discrete waves \textbf{w}, their extracted envelope \textbf{e}, and the inferred carrier, obtained after dividing, element-wise, the original wave by its envelope.
-
-From figure \ref{fig:spoken_voice} it can be seen that the algorithm performs well under the carrier metric.
+From Fig.
+\ref{fig:spoken_voice} it can be seen that the algorithm performs well under the carrier metric.
 This is especially remarkable for the case of spoken voice, a notorious complex wave with fast changes in pitch and high noise content.
 
-These characteristics are in stark contrast with those of the sang voice in figure \ref{fig:singer_voice}, where a recording of an alto singer is shown; there, the wave is almost periodic, with a stable waveform.
+\begin{figure}[H]
+  \centering
+    \includegraphics{images/06Graph_spoken_voice.pdf}
+  \caption{Signal, envelope and carrier for a record of male voice uttering the word “amazing”.}
+  \label{fig:spoken_voice}
+\end{figure}
 
-\begin{figure}[ht!]
+Those characteristics are in stark contrast with those of the sang voice in Fig.
+\ref{fig:singer_voice}, where a recording of an alto singer sustaining a steady note is shown; there, the wave is almost periodic, with a stable waveform and envelope.
+
+\begin{figure}[H]
+  \centering
+    \includegraphics{images/06Graph_alto.pdf}
+  \caption{Signal, envelope and carrier for a record of an alto singer.}
+  \label{fig:singer_voice}
+\end{figure}
+
+Fig. \ref{fig:Bend} shows the envelope in the case of a bend performed in an electric guitar. The variations in frequency in the signal didn't affect the envelope extracted.
+
+\begin{figure}[H]
   \centering
@@
 
-\begin{figure}[ht!]
+\subsection{Theoretical guarantees}
+
+In an effort to formalize the assessment of the quality of amplitude and frequency modulations of a signal, \cite{1996Loughlinamplitude} proposed some theoretical conditions necessary, but not sufficient, to ensure the physical plausibility of an envelope.
+
+We comment below on how those conditions translate to the discrete setting, and show that the algorithm here presented satisfies the four conditions presented.
+
+The first one states that, if a signal is bounded in magnitude, then its envelope should be, as well.
+Our algorithm complies with this requirement, as the envelope is composed of selected samples from the original signal itself.
+
+More formally, we have the set P of the absolute values of the local extrema of the original signal $ \mathbf{w} = (w_0, w_1, \cdots, w_{n-1}) $ that contains, by definition, the global extrema of $ \mathbf{w} $.
+The set E of the points that belong to the envelope is a subset of P. We can thus write Eq. \ref{eq:Req1}:
+
+\begin{subequations} \label{eq:Req1}
+\begin{align}
+\text{P} &= \{ P_0, P_1, \cdots, P_{m-1} \} \quad \text{where} \quad P_j = (i_j, \lvert w_{i_j} \lvert) \quad \forall \quad 0 \le j \le m-1 \\
+\text{E} &\subset \text{P} \implies \max (\text{E}) \le \max (\text{P}) = \max (|w_0|, |w_1|, \cdots, |w_{n-1}|)
+\end{align}
+\end{subequations}
+
+If the amplitude of the original signal is bounded, that is $ \max (|w_0|, |w_1|, \cdots, |w_{n-1}|) = b \in \mathbb{R} $, not only the amplitude of the envelope is bounded, as required, but is guaranteed to have the same bound $ b $.
+
+The second statement dictates that if the signal has a finite frequency range, that frequency range must not be exceeded by the carrier wave.
+
+To see that this condition holds for the proposed method, consider a discrete signal $ \mathbf{w} = (w_0, w_1, \cdots, w_{n-1}) $, where all the absolute values of its local extrema are greater than one.
+
+This requirement doesn't imply loss of generality since, for an arbitrary signal, we can assure this condition by dividing all samples by the value of the minimum absolute local extrema.
+
+The relation $ \min(\text{E}) \ge 1 $ thus holds, since $ \text{E} \subset \text{P}$.
+
+Provided that one uses a method that preserves the original bounds when interpolating the points in E, such as linear interpolation, we can be sure that all the points in the envelope vector $ \mathbf{e} = (e_0, e_1, \cdots, e_{n-1}) $ are also greater than one.
+
+Recalling Eq.
+\ref{eq:Envelope}, where the relation between the original signal, the carrier and the envelope was defined as $ \mathbf{w} = \mathbf{e} \odot \mathbf{c} $, we can write the magnitude of an arbitrary frequency of the original signal ($ m_\mathbf{w}(f) $) and, likewise, the magnitude of an arbitrary frequency of the carrier ($ m_\mathbf{c}(f) $) in terms of the Discrete Fourier Transform (DFT) as in Eq.
+\ref{eq:DFT}:
+
+\begin{subequations}\label{eq:DFT}
+\begin{align}
+m_\mathbf{w}(f) &= \sum_{k=0}^{n-1} \sqrt{\left( w_k \cos(2 \pi k f / n) \right)^2 + \left( w_k \sin(2 \pi k f / n) \right)^2} \\
+m_\mathbf{c}(f) &= \sum_{k=0}^{n-1} \sqrt{\left( \frac{w_k}{e_k} \cos(2 \pi k f / n) \right)^2 + \left( \frac{w_k}{e_k} \sin(2 \pi k f / n) \right)^2} 
+\end{align}
+\end{subequations}
+
+For any given frequency $ f $, both $ m_\mathbf{w}(f) $ and $ m_\mathbf{c}(f) $ are nonnegative.
+Also, since $ e_k \ge 1 \ \forall \ k, 0 \le k \le n - 1 $ we establish that $ 0 \le m_\mathbf{c}(f) \le m_\mathbf{w}(f) \ \forall \ f $ which is a more restrictive condition than the original: If $ \mathbf{w} $ has a finite frequency range, the magnitude of frequencies outside that range are zero, as are those of the same frequencies of the carrier $ \mathbf{c} $.
+
+The third condition states that a physically plausible algorithm should, for a signal $ \mathbf{w} $ of constant amplitude $ a > 0, \ a \in \mathbb{R} $ and constant frequency, of the form $ a \cos(\omega i + \phi) $, result in a constant envelope $ \mathbf{e} $ of magnitude $ a $ and o a carrier wave $ \mathbf{c} $ of the form $ \cos(\omega i + \phi) $.
+
+An example of such behaviour can be observed in Fig.
+\ref{fig:Comparison_sinusoid}, where the envelope of a simple sinusoid is shown.
+To see that this is always the case, note that all the extrema of a sinusoid are of the form $ a (\pm 1) $ and $ P_j = (i_j, \lvert a (\pm 1) \lvert) = (i_j, a) \ \forall P_j \in \text{P} $.
+As $ \text{E} \subset \text{P} $, we have that the envelope $ \mathbf{e} = (a, a, \cdots, a) $ and is, thus constant.
+Likewise, since $ \mathbf{w} = \mathbf{e} \odot \mathbf{c} $, we have $ c_i = a \cos(\omega i + \phi) / a = \cos(\omega i + \phi) \ \forall \ i, 0 \le i \le n - 1 $.
+
+The fourth and last condition states that if a wave is multiplied by a constant, its envelope should also be multiplied by the same constant.
+
+Recall from section \ref{mapping} that, before the extraction of the envelope, the original points undergo a change of basis to lie in the Cartesian coordinate system.
+After the extraction, the inverse operation is performed on the points of the Envelope.
+
+Consider a discrete signal $ \mathbf{w} = (w_0, w_1, \cdots, w_{n-1}) $ and the set P of its absolute extrema, represented in the standard basis, $ \text{P} = \{ P_0, P_1, \cdots, P_{m-1} \}, \quad P_j = (i_j, \beta \lvert w_{i_j} \lvert) $, where $ \beta $ is obtained from Eq.
+\ref{eq:Basis}.
+
+Given a constant $ \alpha > 0, \alpha \in \mathbb{R} $, we can define the scaled signal as $ \mathbf{w}^\prime = (\alpha w_0, \alpha w_1, \cdots, \alpha w_{n-1}) $.
+Observing, from Eq.
+\ref{eq:Basis}, that $ \alpha $ can be factored out of denominator of the summation, we have $ \beta^\prime = \beta / \alpha $ and the new set $ \text{P}^\prime = \{ P^\prime_0, P^\prime_1, \cdots, P^\prime_{m-1} \}, \quad P^\prime_j = (i_j, \frac{\beta}{\alpha} \lvert \alpha w_{i_j} \lvert) $ is equal to the set P for the signal before scaling.
+
+This guarantees that the same subset of points E will be identified as part of the envelope in both cases.
+In the respective original basis system, applying the inverse change of basis, those points are $ P_j = (i_j, \frac{\beta}{\beta} \lvert w_{i_j} \lvert)  \quad \forall \  P \in \text{E} $, for the original signal, and $ P^\prime_j = (i_j, \frac{\alpha}{\beta}\frac{\beta}{\alpha} \lvert \alpha w_{i_j} \lvert) $ for the scaled signal.
+
+\subsection{Reference envelope}
+
+In order to obtain a numeric indicator of the quality of an envelope, we construct a reference envelope, that trades off practical applicability in favour of mathematical formality.
+In \cite{1992BruceCurves} the envelope is defined as the boundary of the region filled by a family of curves.
+Fig.
+\ref{fig:RefEnvelopeTheory} illustrates this concept, for a family of sinusoids that share the same frequency and amplitude, and whose phases vary from 0 to $ 2 \pi $.
+
+\begin{figure}[H]
   \centering
-    \includegraphics{images/06Graph_spoken_voice.pdf}
-  \caption{Signal, envelope and carrier for a record of male voice uttering the word “amazing”.}
-  \label{fig:spoken_voice}
+    \includegraphics{images/09RefEnvelopeTheory.pdf}
+  \caption{Envelope as the boundary of the region filled by a family of curves, for a family of sinusoids with the same frequency and amplitude, modulated by a polynomial.}
+  \label{fig:RefEnvelopeTheory}
 \end{figure}
 
-\begin{figure}[ht!]
+In practical applications, one seldom has access to a family of curves, instead being interested in obtaining the envelope of a single signal.
+Nevertheless, an approximation can be obtained, by shifting the original signal along its initial position in the horizontal axis, in order to cover the time of one period of the signal.
+Fig. \ref{fig:RefEnvelope} shows the reference envelope for the sound of a tom.
+
+\begin{figure}[H]
   \centering
-    \includegraphics{images/06Graph_alto.pdf}
-  \caption{Signal, envelope and carrier for a record of an alto singer.}
-  \label{fig:singer_voice}
+    \includegraphics{images/10RefEnvelope.pdf}
+  \caption{Reference envelope obtained by shifting the original signal, for part of a sample of a tom.}
+  \label{fig:RefEnvelope}
 \end{figure}
 
-In an effort to formalize the assessment of the quality of temporal envelopes, \textcite{1996Loughlinamplitude} proposed some conditions necessary, but not sufficient, to ensure the physical plausibility of an envelope.
-We comment briefly below that the algorithm here presented satisfies the four conditions presented.
-
-The first one states that, if a signal is bounded in magnitude, then its envelope should be, as well.
-Our algorithm complies with this requisite, as the envelope is composed of selected samples from the original signal itself.
-
-The second statement is that if the signal has a finite frequency range, that frequency range must not be exceeded by the envelope.
-This is only directly applicable in the continuous case, where this condition was proposed.
-
-Nevertheless, if we picture both the wave and the envelope as samples from continuous functions, is easy to see that this is the case for our algorithm; since the envelope is composed by some samples of the digital wave, in the limit where all the samples of the wave are also part of the envelope, we have that the maximum frequency for both is the same.
-
-In all other cases, the maximum frequency of the envelope will be smaller than that of the original wave.
-The range of frequencies of the envelope, then, will always lie between zero and the maximum frequency of the original wave.
-
-The third condition states what, for a periodic signal, the envelope should be a straight line with the intercept equal to the amplitude of this signal.
-Since all the local maxima have the same amplitude in the case of a periodic wave, our envelope will indeed be a straight line, with the same amplitude as the original wave.
-An example of such behaviour can be observed in figure \ref{fig:Comparison_sinusoid}.
-
-The fourth and last condition states that if a wave is multiplied by a constant, the envelope should also be multiplied by the same constant.
-That is also the case, again because the points that form the envelope are points of the original wave itself.
+Since envelope is not smooth, being composed of a series of horizontal segments, as can be seen in \ref{fig:RefEnvelope}, besides being computationally expensive to calculate.
+However, arising from a mathematically sound base, it can serve as a reference to infer the quality of more empirical methods.
 
@@
 
-Direct comparison with many of the more recent algorithms is made difficult by the unavailability of digital implementations of such works, many designed to process analog signals \parencite[e.g.,][]{2018AssefModeling}.
+Direct comparison with many of the more recent algorithms is made difficult by the unavailability of digital implementations of such works, many designed to process analog signals \cite[e.g.,][]{2018AssefModeling}.
 
@@
 \item Smoothing: This is a relatively simple procedure that, applied to the absolute values of a discrete signal, can provide an approximation of the variation of its amplitude in time.
-We use the Savitzky-Golay algorithm \parencite{1964SavitzkySmoothing} with a window width of 3001 samples and a cubic polynomial.
-\item Filtering: The absolute value of the wave is filtered using a digital implementation of the Butterworth \parencite{butterworth1930theory} low pass filter.
-We use an order 2 filter with a cut off frequency of 10 Hz.
-\item Hilbert Transform: We apply the Hilbert transform to the absolute values of the wave after filtering the original signal with a Butterworth filter with a cut off frequency of 100 Hz, using the absolute value of the analytic signal to obtain the envelope.
+We use the Savitzky-Golay algorithm \cite{1964SavitzkySmoothing} with a window width of 3001 samples and a cubic polynomial.
+\item Filtering: The absolute value of the wave is filtered using a digital implementation of the Butterworth \cite{butterworth1930theory} low pass filter.
+We use an order 2 filter with a cut-off frequency of 10 Hz.
+\item Hilbert Transform: We apply the Hilbert transform to the original signal, filtering the absolute value of the complex result with a Butterworth filter with a cut-off frequency of 1/10 of the fundamental frequency of the original signal.
 \end{enumerate}
 
-In \ref{fig:Comparison_sinusoid} we illustrate the envelope extracted by the conventional algorithms, contrasting it with the result of the method presented in this work, for a simple sinusoid.
-Note that none of the conventional algorithms complies with the third condition proposed in \textcite{1996Loughlinamplitude} that the envelope of a periodic wave should be a straight line.
+In Fig. \ref{fig:Comparison_sinusoid} we illustrate the envelope extracted by the conventional algorithms, contrasting it with the result of the method presented in this work, for a simple sinusoid. 
+Only the present algorithm and the one based on the Hilbert transform comply with the third condition proposed by \cite{1996Loughlinamplitude}, that states that the envelope of a periodic wave should be a straight line.
 
-\begin{figure}[ht!]
+\begin{figure}[H]
   \centering
     \includegraphics{images/07Comparison_sinusoid.pdf}
-  \caption{Comparison of envelope detection algorithms for a simple sinusoid}
+  \caption{Comparison of envelope detection algorithms for a simple sinusoid. The envelope for the Hilbert transform and for the present work are coincident in the image.}
   \label{fig:Comparison_sinusoid}
@@
 
-For a recording of the key 33 of a grand piano, where parts of the sound are highly percussive and noisy, the Hilbert approach presents considerable oscillations, as can be seen in figure \ref{fig:Comparison_piano}.
+For a recording of the key 33 of a grand piano, where parts of the sound, specially at the start, are highly percussive and noisy, the Hilbert approach presents a considerable deviation, as can be seen in Fig.
+\ref{fig:Comparison_piano}.
 
-\begin{figure}[ht!]
+\begin{figure}[H]
   \centering
@@
 
-In figure \ref{fig:Comparison_soprano} all envelopes exhibit a similar movement along the samples, but the ones generated by traditional algorithms are undershooting the original wave.
+In Fig. \ref{fig:Comparison_soprano} all envelopes exhibit a similar movement along the samples, but the ones generated by traditional algorithms are consistently undershooting the original wave.
 
-\begin{figure}[ht!]
+\begin{figure}[H]
   \centering
@@
 
-All the traditional methods demanded careful choice of parameters.
-In the case of the Hilbert transform, pre-processing in the form of filtering was applied.
+Besides, the traditional methods demanded careful choice of parameters.
+In the case of the Hilbert transform, post-processing in the form of filtering was applied, and an appropriate cut-off frequency had to be chosen.
 That is in contrast with our method, that organically tunes itself via the automatic choice of the radius of the circle representing the average discrete curvature of the signal.
 
-Table \ref{table:Error_comparison} presents a comparison of the average errors of the techniques.
-The average error was calculated by the average of the squared distance of the points in the envelope divided by two and the absolute value of the wave.
-The rationale is that the values of the envelope divided by two should pass near the average of the absolute values of the wave.
-This metric is only valid for reasonably smooth envelopes, however.
+Table \ref{table:Error_comparison} presents the average of the absolute errors of the techniques, when compared with the reference envelope introduced in the end of section \ref{sec:quality}
 
-We can see from the table \ref{table:Error_comparison} that the method presented here consistently outperforms the traditional approaches for all waves tested, yielding an error 33\% smaller, on average, than the other methods.
-It is also worth noting that the errors of the traditional methods are very similar between them.
-
-\begin{table}[ht!]
+\begin{table}[H]
 \centering
@@
 
-The times taken for the algorithms compared here to process each wave are shown in \ref{table:Time_comparison}.
-We used specialized implementations available in libraries of the Python programming language.
-The presented algorithm is in average the second-fastest.
+We can see from the table \ref{table:Error_comparison} that the method presented here consistently outperforms the traditional approaches for all waves tested, yielding an error of less than 2\% on average, of the error of the traditional methods.
 
-\begin{table}[ht!]
+The times taken for the algorithms to process each signal are shown in \ref{table:Time_comparison}. 
+The Python source for the tests, as well as the samples used, are available at the repository dedicated to this work, in the file “Comparison of Envelope Algorithms.py”.
+
+The methods were performed as described in the beginning of this section, using the implementations available at the Scipy library, under the Signal processing module.
+The presented algorithm, is on average, the second-fastest.
+
+\begin{table}[H]
 \centering
@@
 
-\subsection{Frontiers}
+\section{Extensions}
 
-In practice, is not uncommon for a discrete wave, especially in the case of sound, to present somewhat different positive and negative contours; in those cases, the algorithm here presented can be used independently in the positive and negative pulses of the wave to define two “envelopes”, that we are going to call superior and inferior frontiers.
+In this section we comment briefly on some characteristics and extensions of the presented algorithm, in order to suggest possible applications beyond envelope extraction.
+One of the less obvious is, perhaps, its use in the segmentation of a signal into pseudo cycles.
 
-Figure \ref{fig:FullFrontiers} illustrates the frontiers of six diverse discrete sound waves, as well as an in-detail view of the highlighted segment for each signal.
+\subsection{Superior and inferior envelopes}
+
+In practice, is not uncommon for a discrete wave, especially in the case of sound, to present somewhat different superior and inferior contours; in those cases, the algorithm can be easily adapted to infer from the positive and negative pulses defined by the original wave to define two independent “envelopes”, that we are going to call superior and inferior frontiers.
+
+If we split P, as defined in section \ref{sec:representation}, into $\text{P}^+$, the set of the local maxima of the original signal and $\text{P}^-$, the set of its local minima, we can apply the remaining steps of the algorithm in those two sets independently, arriving at a positive and a negative frontier.
+
+Fig.
+\ref{fig:FullFrontiers} illustrates these frontiers for six diverse discrete sound waves, as well as an in-detail view of the highlighted segment for each signal.
 All waves are records of physical sounds, chosen to represent the applicability of the algorithm in real-world scenarios.
+The implementation available at the PyPI repository provides an option to extract the frontiers.
 
-\begin{figure}[ht!]
+\begin{figure}[H]
   \centering
@@
 
-The frontiers are satisfactorily detected in the case of harmonic and inharmonic sounds, and are robust in relation to the number of samples and the frequencies of the waves.
-That is a peculiarity of our algorithm: the traditional algorithms are unable to generate two distinct envelopes for the same wave.
+The frontiers were are satisfactorily detected in the case of harmonic and inharmonic sounds, and are robust in relation to the number of samples and the frequencies of the waves.
+It is worth noting that the positive and negative frontiers are generally smoother than the unified envelope and seems to conform better to the countours of the underlying digital signal.
 
-\subsection{Approximated location of pseudo-cycles}
+\subsection{Simplified spectral representation}
 
-The envelope is shown to add complexity to the spectral representation of a wave \parencite{2019TarjanoNeuro}, and an accurate description of the envelope, while describing the evolution of the instantaneous amplitude of a signal in time, would also simplify further spectral analysis.
-Figure \ref{fig:Fourier} shows the frequency-domain power spectrum for the wave and the carrier presented in \ref{fig:Envelope}.
+The temporal envelope adds complexity to the spectral representation of a wave \cite{2019TarjanoNeuro}, and an accurate description of the envelope, while describing the evolution of the instantaneous amplitude of a signal in time, would also simplify further spectral analysis.
+
+Fig.
+\ref{fig:Fourier} shows the frequency-domain power spectrum for the wave and the carrier presented in \ref{fig:Envelope}.
 For the carrier, the power spectrum presents two nonzero values, while the power spectrum for the wave composed of the carrier modulated by the envelope is more complex.
 
-\begin{figure}[ht!]
+\begin{figure}[H]
   \centering
@@
 
-Moreover, the algorithm developed in this work naturally divides a signal into its pseudo-cycles, pinpointing them in the time-domain, thus providing the building blocks for the reconstruction of the fine structure of the wave.
+Fig.
+\ref{fig:FourierBend} shows part the power spectra for both the original wave and the carrier obtained by the present algorithm, for the guitar bend sound in Fig.
+\ref{fig:Bend}, around its fundamental frequency.
+Two prominent frequencies can be seen, the initial and final notes of the bend.
+In the carrier, the other frequencies appear more smooth.
 
-\begin{figure}[ht!]
+\begin{figure}[H]
+  \centering
+    \includegraphics{images/12Fourier.pdf}
+  \caption{Fourier power spectrum for the wave and carrier of the sound of a guitar bend shown in Fig. \ref{fig:Bend}.}
+  \label{fig:FourierBend}
+\end{figure}
+
+\subsection{Approximated location of pseudo-cycles}
+
+The algorithm developed in this work naturally divides a signal into its pseudo-cycles, pinpointing them in the time-domain, as illustrated in Fig.
+\ref{fig:singer_voice}, for the sound of an alto singer singing a sustained note, as in Fig.
+\ref{fig:singer_voice}.
+
+The vertical lines mark the indices of the samples that belong to the positive frontier of the original signal, and the section between each vertical line isolates a pseudo cycle of the wave.
+Note that, although two local maxima exist in each pseudo cycle, only the largest was marked.
+
+\begin{figure}[H]
   \centering
     \includegraphics{images/09PCs.pdf}
-  \caption{Section of the lines formed by the positions of the positive and negative frontiers of the digital wave illustrated in \ref{fig:singer_voice}, as extracted by the algorithm presented, and their average.}
+  \caption{Section of the digital wave illustrated in \ref{fig:singer_voice}, with vertical lines at the indices of the samples that belong to the positive frontier.}
   \label{fig:PCs}
@@
 
-One can see in figure \ref{fig:PCs} a section of the plot of where the local maxima and minima occurs in the recording of an alto singer show in \ref{fig:singer_voice}.
-The average between the two is also shown.
-With this information, one can split a discrete wave into its pseudo-cycles, as is illustrated in Figure \ref{fig:AvgPc}: It shows the superposition of this extracted pseudo-cycles, after being normalized by length, and their average.
+Therefore, using those indices, one can split a discrete wave into its pseudo cycles.
+One application of that is illustrated in Fig.
+\ref{fig:AvgPc}: It shows the superposition of the extracted pseudo cycles, after being normalized by length, and their average.
+They were also rotated to make their starting point more close to zero.
+
 This same average is shown shifted, to exemplify that approximate $C^0$ and $C^1$ continuity is achieved.
+This can be seen as the average waveform of the original discrete signal, and can be used, along with the information of its envelope, to generate an alternative representation of the wave, useful, for example, for compression and classification purposes.
 
-\begin{figure}[ht!]
+\begin{figure}[H]
   \centering
@@
 
-\section{References}
-
-\printbibliography[heading=none]
+\bibliography{bibli}
 

### 2021-05-07 Papers/01_Pulses_Envelope/PulsesEnvelope.tex
moving to elsevier template

@@
 
-$ E(t) $ represents the slow varying part of the wave, also known as the (temporal) envelope, modulator component, or amplitude modulation (AM) of $ W(t) $, while $ C(t) $ models its fast varying part, called throughout the literature as its (temporal) fine structure, carrier component, or frequency modulation (FM).
+$ E(t) $ represents the slow varying part of the wave, also known as the (temporal) envelope, modulator component, or amplitude modulation (AM) of $ W(t) $, while $ C(t) $ models its fast varying part, called throughout the literature as its (temporal) fine structure, carrier wave, or frequency modulation (FM).
 
@@
 To that end we are going to apply the definition of smooth curvature as the rate of change of the unit tangent to a curve.
-This definition is convenient, since it is equivalent to that of the osculating circle \parencite{2016VasaMesh}.
+This definition is convenient, since it is equivalent to that of the osculating circle \parencite{2016VasaMesh}, the value we are interested in finding.
 
-One of those edges, a vector $ v_k \in \text{V} $, is singled out in Fig.
+One of those edges, in the form of the vector $ v_k \in \text{V} $, is singled out in Fig.
 \ref{fig:DiscreteCurvature}.
@@
 
-That is because all we can assume from $ v_k $ is that, in this current region of the envelope, it is able to a vertical displacement of $ v_{k,y} $ in the same time as it performs a horizontal displacement of $ v_{k,x} $;
+That is because all we can safely assume from $ v_k $ is that, in the immediate vicinity of the vector, the envelope is capable of a vertical displacement of $ v_{k,y} $ in the same time as it performs a horizontal displacement of $ v_{k,x} $;
 
@@
 
-The radius of the circle equivalent to each $ v_k $ is defined in Eq.
+The radius $ r_k $ of the circle equivalent to each $ v_k $ is defined in Eq.
 \ref{eq:Radius}.
 This equation has the desirable property that $ v_{k,y} \to 0 \implies v_{k,y} \to \infty $, that is, the radius tends to infinity when no vertical oscillation occurs, effectively turning the circle into a line.
-% It is also directly proportional to $ v_{k,x} $, making the radius larger the more the points are horizontally far from each other.
+
+The radius is also directly proportional to $ v_{k,x} $, making the radius larger the more the points are horizontally far from each other.
 
@@
 
-For calculation purposes, however, this can cause a problem, as a zero could make its way into the denominator of a fraction; in the actual algorithm, the quantity calculated for each vector is the curvature itself, that is, the inverse of the radius.
-In this way, the problem is avoided, since $ v_{k,y} $ is guaranteed to be greater than zero.
-The average of curvatures is then inverted in order to obtain the average radius, as illustrated in Eq.
+For calculation purposes, however, this can pose a problem, as a zero could make its way into the denominator of a fraction; in the actual algorithm, the quantity calculated for each vector is the curvature itself, that is, the inverse of each radius $ r_k $.
+In this way, the problem is avoided, since $ v_{k,x} $ is guaranteed to be greater than zero.
+The average of curvatures is then inverted in order to obtain the average radius $ \overline{r} $, as illustrated in Eq.
 \ref{eq:Curvature}.

### 2021-05-07 Papers/01_Pulses_Envelope/00ElsevierPulsesEnvelope.tex
will try to clean bib

@@
 Once a radius $ \alpha $ is defined, alpha shapes exhibit the uniqueness property meaning that, for a given set of points and a given radius, only one concave hull exists.
-When the radius of the circle $ \alpha \to \infty $, concave hulls tend to the convex hulls.
+When the radius $ \alpha $ of the circle tends to infinity, concave hulls tend to the convex hulls.
 
@@
 
-While there is no general formal definition of the envelope of broadband signals \cite{2014Yangnovel,2015Yangtheoretical,2019Jiaempiricalj}, the general intuition is that an envelope is a smooth curve situated as close as possible to the signal \cite{2019Jiaempiricalj}.
+While there is no general formal definition of the envelope of broadband signals \cite{2014Yangnovel,2015Yangtheoretical,2019Jiaempirical}, the general intuition is that an envelope is a smooth curve situated as close as possible to the signal \cite{2019Jiaempirical}.
 
@@
 
-Despite its widespread use, and effectiveness in the context of narrowband signals, envelope detection techniques based on the Hilbert transform don't behave well for broadband signals \cite{2012DauSpeech, 2019Jiaempiricalj}, being even physically paradoxical in some cases \cite{1996Loughlinamplitude, 2014Yangnovel}.
+Despite its widespread use, and effectiveness in the context of narrowband signals, envelope detection techniques based on the Hilbert transform don't behave well for broadband signals \cite{2012DauSpeech, 2019Jiaempirical}, being even physically paradoxical in some cases \cite{1996Loughlinamplitude, 2014Yangnovel}.
 
@@
 
-Recall from section \ref{mapping} that, before the extraction of the envelope, the original points undergo a change of basis to lie in the Cartesian coordinate system.
+Recall from section \ref{sec:mapping} that, before the extraction of the envelope, the original points undergo a change of basis to lie in the Cartesian coordinate system.
 After the extraction, the inverse operation is performed on the points of the Envelope.

### 2021-05-07 Papers/01_Pulses_Envelope/00ElsevierPulsesEnvelope.tex
formatted according to elsevier template

@@
-\documentclass{elsarticle}
+\documentclass[preprint]{elsarticle}
 % \usepackage[natbib=true]{biblatex}
@@
 \usepackage{algorithmic}
+\usepackage{lineno,hyperref}
+% \modulolinenumbers[5]
 
@@
 
+\linenumbers
+
 \section{Introduction}
@@
 
-\begin{figure}[H]
+\begin{figure}[ht!]
   \centering
@@
 
-\begin{subequations}\label{eq:Envelope}
-\begin{align}
+\begin{align} \label{eq:Envelope}
+\begin{split}
   \text{Given:}& \\
@@
   \mathbf{w} & = \mathbf{e} \odot \mathbf{c} \ \therefore \ w_i = e_i c_i \ \forall \ i, \ 0 \le i \le n-1
+\end{split}
 \end{align}
-\end{subequations}
 
@@
 
-\begin{subequations}  \label{eq:Basis}
-\begin{align}
+\begin{align}\label{eq:Basis}
+\begin{split}
 P_{\mathbf{b}} &= (x_1, x_2) \\
@@
 P_{\mathbf{s}} &= (x_1, \beta x_2) \quad \forall \  P \in \text{P} \\ 
+\end{split}
 \end{align}
-\end{subequations}
 
@@
 
-\begin{equation} \label{eq:Vectors}
+\begin{align} \label{eq:Vectors}
 \text{V} = \{ v_0, v_1, \cdots, v_{m-2} \} \quad \text{where} \quad v_k = P_{j+1} - P_j \quad \forall \quad 0 \le k \le m-2
-\end{equation}
+\end{align}
 
@@
 
-\begin{equation} \label{eq:Radius}
+\begin{align} \label{eq:Radius}
   r_k = \frac{v_{k,x} \sqrt{v_{k,x}^2 + v_{k,y}^2}}{ v_{k,y}}
-\end{equation}
+\end{align}
 
@@
 
-\begin{equation} \label{eq:Curvature}
+\begin{align} \label{eq:Curvature}
 \overline{r} = \left( \frac{\sum\limits_{k=0}^{m-2} \left( \frac{ v_{k,y}}{v_{k,x} \sqrt{v_{k,x}^2 + v_{k,y}^2}} \right)}{m-2} \right)^{-1}
-\end{equation}
+\end{align}
 
@@
 
-\begin{subequations} \label{eq:InEnvelope}
-\begin{align}
+\begin{align} \label{eq:InEnvelope}
+\begin{split}
 & P_a \in \text{E} \leftrightarrow (P_{b,x} - x_o)^2 + (P_{b,y} - y_o)^2 > r^2, \\
@@
 & P_a, P_b, P_c \in \text{P}, \quad P_a \ne P_b \ne P_c \quad \text{and} \quad x_o, y_o \in \mathbb{R}, \quad 0 \le y_o < +\infty
+\end{split}
 \end{align}
-\end{subequations}
 
@@
 
-\begin{algorithm}[H]
+\begin{algorithm}[ht!]
 \caption{Retrieve Envelope} \label{FindEnvelope}
@@
 Figs.
-\ref{fig:spoken_voice}, \ref{fig:singer_voice} and \ref{fig:Bend} present discrete waves \textbf{w}, their extracted envelope \textbf{e}, and the inferred carrier \textbf{c}, obtained after dividing, element-wise, the original signal by its extracted.
+\ref{fig:spokenvoice}, \ref{fig:singer_voice} and \ref{fig:Bend} present discrete waves \textbf{w}, their extracted envelope \textbf{e}, and the inferred carrier \textbf{c}, obtained after dividing, element-wise, the original signal by its extracted.
 
 From Fig.
-\ref{fig:spoken_voice} it can be seen that the algorithm performs well under the carrier metric.
+\ref{fig:spokenvoice} it can be seen that the algorithm performs well under the carrier metric.
 This is especially remarkable for the case of spoken voice, a notorious complex wave with fast changes in pitch and high noise content.
 
-\begin{figure}[H]
+\begin{figure}[ht!]
   \centering
@@
   \caption{Signal, envelope and carrier for a record of male voice uttering the word “amazing”.}
-  \label{fig:spoken_voice}
+  \label{fig:spokenvoice}
 \end{figure}
@@
 
-\begin{subequations} \label{eq:Req1}
-\begin{align}
+\begin{align} \label{eq:Req1}
+\begin{split}
 \text{P} &= \{ P_0, P_1, \cdots, P_{m-1} \} \quad \text{where} \quad P_j = (i_j, \lvert w_{i_j} \lvert) \quad \forall \quad 0 \le j \le m-1 \\
 \text{E} &\subset \text{P} \implies \max (\text{E}) \le \max (\text{P}) = \max (|w_0|, |w_1|, \cdots, |w_{n-1}|)
+\end{split}
 \end{align}
-\end{subequations}
 
@@
 
-\begin{subequations}\label{eq:DFT}
-\begin{align}
+\begin{align} \label{eq:DFT}
+\begin{split}
 m_\mathbf{w}(f) &= \sum_{k=0}^{n-1} \sqrt{\left( w_k \cos(2 \pi k f / n) \right)^2 + \left( w_k \sin(2 \pi k f / n) \right)^2} \\
 m_\mathbf{c}(f) &= \sum_{k=0}^{n-1} \sqrt{\left( \frac{w_k}{e_k} \cos(2 \pi k f / n) \right)^2 + \left( \frac{w_k}{e_k} \sin(2 \pi k f / n) \right)^2} 
+\end{split}
 \end{align}
-\end{subequations}
 
@@
 
-\begin{table}[H]
+\begin{table}[ht!]
 \centering

### 2021-05-08 Papers/01_Pulses_Envelope/00ElsevierPulsesEnvelope.tex
preparing for final revisions

@@
 
+% \author{Carlos Henrique Tarjano Santos\corref{cor1}}
+% \author{Valdecy Pereira}
+% \address{Department of Production Engineering, Universidade Federal Fluminense, Rua Passo da Pátria, 156, Campus Praia Vermelha, Bloco D - sala 309, São Domingos, Niterói, RJ, Brasil, CEP: 24.210-240}
+% \cortext[cor1]{Corresponding author}
+% \ead{carlostarjano@id.uff.br}
+
+\cortext[cor1]{Corresponding author}
+\address{Department of Production Engineering, Universidade Federal Fluminense, Rua Passo da Pátria, 156, Campus Praia Vermelha, Bloco D - sala 309, São Domingos, Niterói, RJ, Brasil, CEP: 24.210-240}
+
 \author{Carlos Henrique Tarjano Santos\corref{cor1}}
 \ead{carlostarjano@id.uff.br}
-\address{Department of Production Engineering, Universidade Federal Fluminense, Rua Passo da Pátria, 156, Campus Praia Vermelha, Bloco D - sala 309, São Domingos, Niterói, RJ, Brasil, CEP: 24.210-240}
-\cortext[cor1]{Corresponding author, carlostarjano@id.uff.br}
 
 \author{Valdecy Pereira}
-\address{Department of Production Engineering, Universidade Federal Fluminense, Rua Passo da Pátria, 156, Campus Praia Vermelha, Bloco D - sala 309, São Domingos, Niterói, RJ, Brasil, CEP: 24.210-240}
+% \ead{support@elsevier.com}
+
+% \author{Valdecy Pereira}
+% \address{Department of Production Engineering, Universidade Federal Fluminense, Rua Passo da Pátria, 156, Campus Praia Vermelha, Bloco D - sala 309, São Domingos, Niterói, RJ, Brasil, CEP: 24.210-240}
 
@@
 
+We start with a brief background of envelope detection in Section \ref{sec:envelope} and define, as specifically as possible, what constitutes a temporal envelope.
+
+In Section \ref{sec:representation} we explore some simplifications to the task of detecting the envelope of a signal that arise from the envelope definition adopted in this work and simplify the task of envelope detection.
+
+Section \ref{sec:mapping} explores the geometric characteristics of envelope detection and formalizes how this classic DSP problem can be translated to the geometric problem of finding the shape of a set of points.
+
+We next explore, in Section \ref{sec:curvature}, the concept of discrete curvature, that will be used in Section \ref{sec:envelopedetection}, where the envelope detection algorithm is presented.
+
+Section \ref{sec:quality} discusses the difficulties in the assessment of the quality of an envelope, proposing qualitative and quantitative approaches, and demonstrates that the method propose conforms to four requirements for a physically plausible envelope proposed in \cite{1996Loughlinamplitude}.
+
+In Section \ref{sec:comparison} we compare the presented algorithm with traditional envelope detection methods, both in terms of accuracy and time and present, in Section \ref{sec:extensions},  potential applications of the algorithm, some of which might not be immediately obvious.
+
 % TODO %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-After defining what will be considered an envelope in the context of this work, in this section we explore some simplifications that can be applied to a discrete wave, defining the representation that will be used in the subsequent part of the work.
-
-We next explore the concept of discrete curvature, putting forth our own method, tailored to the definition of the equivalent circle of a wave; we end this section by showing how this circle can be used to identify the envelope of a discrete wave, in a similar approach to that of the alpha shapes theory, but taking advantage of some unique structure present in discrete waves.
-
-The rest of this paper is divided as follows: We start by defining as specifically as possible what a temporal envelope is, especially in the context of this work and providing a mathematical definition, in terms of the Hilbert transform, for the case of a narrowband signal.
-
-We proceed to illustrate the representation of a discrete wave used in this work, largely based on the concept of dividing the wave into its constituent pulses, and how it can simplify the algorithm, reducing the high dimensionality generally present in digital representations of waves.
-
-Next, we review methods for the estimation of the curvature of a discrete wave, after which we present our own; this method will then be used to estimate the equivalent circle, that will be used to define the envelope of the wave, via the procedure described in sequence.
-
-In the results section, we present some illustrations of the envelopes obtained with our method, comparing both accuracy and running time with the most common algorithms encountered in the literature.
-We also suggest an alternative use of the algorithm to identify the negative and positive frontiers of a wave, that can be seen as partial envelopes considering only the positive and negative pulses of a wave, respectively.
-
-We follow with suggestions about how the approach here presented can be useful beyond envelope detection, exemplifying with some applications, such as the extractions of the average waveform of a digital wave, and concluding with commentary on potential future developments of the method.
-
 Interactive visualizations are available from the repository prepared for this work \cite{2020TarjanoEnvelope}, that also contains an implementation of the method in pure Python.
@@
 
-\section{Temporal envelope}\label{sec:Characterization of a temporal envelope}
+\section{Temporal envelope}\label{sec:envelope}
 
@@
 
-From \ref{fig:Envelope} and the discussion in Section \ref{sec:Characterization of a temporal envelope}, we can see that only the maximum point of a positive pulse or the minimum point of a negative pulse can potentially be equal to the envelope and are, thus, our only points of interest.
+From \ref{fig:Envelope} and the discussion in Section \ref{sec:envelope}, we can see that only the maximum point of a positive pulse or the minimum point of a negative pulse can potentially be equal to the envelope and are, thus, our only points of interest.
 We can then proceed, for the rest of the method, considering only those points, to great computational economy.
@@
 
-The effect of this normalization can be seen in the points shown in Fig.\ref{fig:Coordinates}, shown in scale for both basis.
+The effect of this normalization can be seen in the points shown in Fig.
+\ref{fig:Coordinates}, shown in scale for both basis.
 We can consider our points now residing in the Cartesian coordinate system.
@@
 
-\section{Discrete Curvature Estimation}
+\section{Discrete Curvature Estimation} \label{sec:curvature}
 
@@
 
-\section{Extracting the envelope}
+\section{Identifying the envelope} \label{sec:envelopedetection}
 
@@
 
-\section{Quality of an envelope}\label{sec:quality}
+\section{Quality of an envelope} \label{sec:quality}
 
@@
 
-\subsection{Comparison with traditional algorithms}
+\subsection{Comparison with traditional algorithms}\label{sec:comparison}
 
@@
 
-\section{Extensions}
+\section{Extensions} \label{sec:extensions}
 

### 2021-05-09 Papers/01_Pulses_Envelope/00ElsevierPulsesEnvelope.tex
about to start grammar check
revising

@@
 \documentclass[preprint]{elsarticle}
-% \usepackage[natbib=true]{biblatex}
 \usepackage[margin=0.5in]{geometry}
-% \usepackage[utf8]{inputenc}
-% \usepackage{graphicx}
-% \usepackage[backend=biber]{biblatex}
 \usepackage{amsmath}
@@
 \usepackage{lineno,hyperref}
-% \modulolinenumbers[5]
 
@@
 \title{Envelope Estimation using Geometric Properties of a Discrete Real Signal}
-
-% \author{Carlos Henrique Tarjano Santos\corref{cor1}}
-% \author{Valdecy Pereira}
-% \address{Department of Production Engineering, Universidade Federal Fluminense, Rua Passo da Pátria, 156, Campus Praia Vermelha, Bloco D - sala 309, São Domingos, Niterói, RJ, Brasil, CEP: 24.210-240}
-% \cortext[cor1]{Corresponding author}
-% \ead{carlostarjano@id.uff.br}
 
@@
 \author{Valdecy Pereira}
-% \ead{support@elsevier.com}
-
-% \author{Valdecy Pereira}
-% \address{Department of Production Engineering, Universidade Federal Fluminense, Rua Passo da Pátria, 156, Campus Praia Vermelha, Bloco D - sala 309, São Domingos, Niterói, RJ, Brasil, CEP: 24.210-240}
 
@@
 Despite being an elusive concept, the temporal amplitude envelope of a signal is essential for its complete characterization, being the primary information-carrying medium in spoken voice and telecommunications, for example.
-Envelope detection techniques have applications in areas like medicine, sound classification and synthesis, seismology and speech recognition.
+Envelope detection techniques have applications in areas like medicine, sound classification and synthesis, seismology, and speech recognition.
 Nevertheless, a general approach to envelope detection doesn't exist, as most methods involve manual intervention, in the form of filter design, smoothing, and other specific design choices based on prior knowledge of the signals under investigation.
 To address this problem, we propose an algorithm that uses characteristics of a signal to estimate its envelope, eliminating the necessity of parameter tuning.
-The method draws inspiration from geometric concepts; the discrete signal is mapped to a set of points in the Cartesian plane, were a new measure of discrete curvature is used to identify the samples that are part of the envelope.
+The method draws inspiration from geometric concepts; the discrete signal is mapped to a set of points in the Cartesian plane, where a new measure of discrete curvature is used to identify the samples that are part of the envelope.
 Due to its geometric nature, the method can optionally generate the superior and inferior envelopes of a given wave.
-The algorithm compares favourably with classic envelope detection techniques based on smoothing, filtering and the Hilbert Transform, besides being physically plausible.
-We provide visualizations of the envelope extracted for various real-world signals, with very diverse characteristics, such as voice, spoken and sang, and pitched and non-pitched musical instruments, and discuss some approaches to assess the quality of the obtained envelopes.
+The algorithm compares favorably with classic envelope detection techniques based on smoothing, filtering, and the Hilbert Transform, besides being physically plausible.
+We provide visualizations of the envelope extracted for various real-world signals with very diverse characteristics, such as voice, spoken and sang, and pitched and unpitched musical instruments, and discuss some approaches to objectively assess the quality of the obtained envelopes.
 An optimized implementation is available via the Python Package Index (PyPI); interactive visualizations and source code are available at the accompanying repository.
@@
 \maketitle
-
 \linenumbers
-
 \section{Introduction}
@@
 
-The intuition behind the alpha shapes algorithm is that only points that can be touched by a circle of radius $ \alpha $ coming from an infinite distance towards the baricenter of the set of points are part of the frontier.
+% TODO
+The intuition behind the alpha shapes algorithm is that only points that can be touched by a disc of radius $ \alpha $ coming from an infinite distance towards the baricenter of the set of points are part of the frontier.
 The polygon obtained linking those points forms the concave hull of this set.
 
-Once a radius $ \alpha $ is defined, alpha shapes exhibit the uniqueness property meaning that, for a given set of points and a given radius, only one concave hull exists.
+Once a radius $ \alpha $ is defined, alpha shapes exhibit the uniqueness property: for a given set of points and a given radius, only one concave hull exists.
 When the radius $ \alpha $ of the circle tends to infinity, concave hulls tend to the convex hulls.
+This
 
 Consider, for example, the artificially generated discrete wave in Fig.
-\ref{fig:Hulls} below, where the links between consecutive samples were deemphasized to highlight the "set of points" nature of a signal: The convex hull and a concave hull, in the form of an alpha shape with a particular value of $ \alpha $ are illustrated.
+\ref{fig:Hulls} below, where the links between consecutive samples were deemphasized to highlight the “set of points” nature of a signal: two alpha shapes with different values for $ \alpha $ and the convex hull are illustrated.
+As $ \alpha $ increases, the contour of the set of points becomes smoother, and closer to the convex hull.
 
@@
 For a real world signal, such as an alto singer sustaining a steady note as shown in Fig.
-\ref{fig:AnalyticSignalAlto} the Hilbert transform alone provides a poor envelope, since it retains most part of the frequencies of the underlying wave.
-With further processing, however, such as filtering, an output closer to the intuitive definition of envelope given in the beginning of this section can be obtained.
+\ref{fig:AnalyticSignalAlto} the Hilbert transform alone provides a poor envelope, since it retains most of the frequencies of the underlying wave.
+With further processing, however, such as filtering, a smoother output can be obtained.
 
@@
 
-Geometrically, we can think of these points as points in a vector space described by a non-standard basis vector $ \mathbf{b} = \{ \mathbf{b}_x, \mathbf{b}_y \}$ with $ \mathbf{b}_x = (a, 0), \mathbf{b}_y = (0, i) $ where $ a $ is an amplitude unit such as decibel, a measure of the instantaneous sound pressure, volt or even a fraction of a fixed maximum value.
+Geometrically, we can think of these points as points in a vector space on a non-standard, orthogonal basis vector $ \mathbf{b} = \{ \mathbf{b}_x, \mathbf{b}_y \}$ with $ \mathbf{b}_x = (a, 0), \mathbf{b}_y = (0, i) $ where $ a $ is an amplitude unit such as decibel, a measure of the instantaneous sound pressure, volt or even a fraction of a fixed maximum value.
 $ i $ represents the index of the sample, and is linked to time by the aforementioned relation $ i = t \ \text{fps} $.
 
-We are interested in finding a change of basis formula from the original basis to the standard basis $ \mathbf{s} = \{ \mathbf{s}_x, \mathbf{s}_y \} = \{ (1,0),(0,1) \}$ of this vector space, making the geometric representation consistent.
+We are interested in representing the point in the standard basis $ \mathbf{s} = \{ \mathbf{s}_x, \mathbf{s}_y \} = \{ (1,0),(0,1) \}$ of this vector space, making the geometric representation consistent.
 
@@
 If we proceeded similarly in relation to the horizontal components,
-the direct link between P and \textbf{w} that arises from the fact that the abscissas of the points in P are indices $ i $ of \textbf{w} would be lost.
+the direct link between P and \textbf{w} that arises from the fact that the abscissas of the points in P are numerically equivalent to the indices $ i $ of \textbf{w} would be lost.
 
 It is best to leave the abscissas intact and multiply the now unitless vertical coordinates of each point by the average of the difference between the abscissa of a point and the abscissa of the immediately posterior.
-In this way, we can consider both axes as having unit $ i $, and the relationship is preserved, making it easier to recover the envelope points later.
+In this way, we can consider both axes as having unit $ i $, and the relationship is preserved, making it easier to recover the envelope points in the original basis later.
 
-Considering $P_{\mathbf{b}}$ as a point of the set P represented by the original basis and $P_{\mathbf{s}}$ the same point as represented with the standard basis, we can formalize this change of basis as in \ref{eq:Basis}:
+Considering $P_{\mathbf{b}}$ as a point of the set P represented on the original basis and $P_{\mathbf{s}}$ the same point as represented on the standard basis, we can formalize this transformation as in \ref{eq:Basis}:
 

### 2021-05-09 Papers/01_Pulses_Envelope/images/00ConvexHull.py

@@ -29,18 +29,38 @@ W = W / np.max(np.abs(W))
 
 points = np.array([[x, w] for x, w in zip(X, W)])
 hull = ConvexHull(points)
-alpha = 300
-alpha_edges = alpha_shape(points, alpha)
+
+'''============================================================================'''
+'''                              Alpha 1                                    '''
+'''============================================================================'''
+
+alpha1 = 50
+alpha_edges = alpha_shape(points, alpha1)
 alpha_edges = sorted(alpha_edges, key=lambda tup: tup[0])
 
-XX = []
-YY = []
+XX1 = []
+YY1 = []
 for e in alpha_edges:
   idx0 = e[0]
   idx1 = e[1]
+  XX1.append(points[idx0][0]), XX1.append(points[idx1][0]), XX1.append(None)
+  YY1.append(points[idx0][1]), YY1.append(points[idx1][1]), YY1.append(None)
 
-  XX.append(points[idx0][0]), XX.append(points[idx1][0]), XX.append(None)
-  YY.append(points[idx0][1]), YY.append(points[idx1][1]), YY.append(None)
+'''============================================================================'''
+'''                              Alpha 2                                    '''
+'''============================================================================'''
+
+alpha2 = 300
+alpha_edges = alpha_shape(points, alpha2)
+alpha_edges = sorted(alpha_edges, key=lambda tup: tup[0])
+
+XX2 = []
+YY2 = []
+for e in alpha_edges:
+  idx0 = e[0]
+  idx1 = e[1]
+  XX2.append(points[idx0][0]), XX2.append(points[idx1][0]), XX2.append(None)
+  YY2.append(points[idx0][1]), YY2.append(points[idx1][1]), YY2.append(None)
 
 
 '''============================================================================'''
@@ -51,6 +71,11 @@ FONT = dict(
     color="black",
     size=13.3333
   )
+
+al = .5
+red = f"rgba(191, 25, 13, {al})"
+blue = f"rgba(13, 92, 161, {al})"
+green = f"rgba(0, 0, 0, {al})"
 
 '''Plotting'''
 fig = go.Figure()
@@ -70,6 +95,37 @@ fig.update_xaxes(showline=False, showgrid=False, zerolinewidth=1, zerolinecolor=
[…]

### 2021-05-09 Papers/01_Pulses_Envelope/images/01AnalyticSignal.py

@@ -65,7 +65,7 @@ fig.update_yaxes(showline=False, showgrid=False, zerolinewidth=.8, zerolinecolor
 '''Plotting Signal'''
 fig.add_trace(
   go.Scatter(
-    name="Signal",
+    name="Signal    ",
     # x=X,
     y=W,
     mode="lines+markers",
@@ -79,7 +79,7 @@ fig.add_trace(
 
 fig.add_trace(
   go.Scatter(
-    name="Hilbert Envelope",
+    name="Hilbert Envelope    ",
     # x=X,
     y=hilbert_envelope,
     mode="lines",

### 2021-05-09 Papers/01_Pulses_Envelope/images/01AnalyticSignalAlto.py

@@ -47,14 +47,14 @@ fig.layout.template ="plotly_white"
 fig.update_layout(
   xaxis_title="<b>Sample <i>i</i></b>",
   yaxis_title="<b>Amplitude</b>",
-  legend=dict(orientation='h', yanchor='top', xanchor='left', y=1.1, itemsizing='constant'),
+  legend=dict(orientation='h', yanchor='top', xanchor='left', y=1.1, itemsizing='trace'),
   margin=dict(l=0, r=0, b=0, t=0),
   font=FONT,
   titlefont=FONT
 )
 fig.layout.xaxis.title.font=FONT
 fig.layout.yaxis.title.font=FONT
-fig.update_xaxes(showline=False, showgrid=False, zeroline=False)
+fig.update_xaxes(showline=False, showgrid=False, zeroline=False, range=[10744, 13683])
 fig.update_yaxes(showline=False, showgrid=False, zerolinewidth=.8, zerolinecolor="silver")
 
 '''Plotting Signal'''
@@ -74,29 +74,29 @@ fig.add_trace(
 
 fig.add_trace(
   go.Scatter(
-    name="Hilbert Transform   ",
+    name="Hilbert Envelope   ",
     # x=X,
     y=hilbert_envelope,
     mode="lines",
-    line_shape='spline',
+    # line_shape='spline',
     # fillcolor="rgba(250, 140, 132, 0.5)",
     # fill="toself",
     line=dict(
         width=1,
-        color="rgba(250, 0, 0, 0.5)",
+        color="red",
     )
   )
 )
 
 fig.add_trace(
   go.Scatter(
-    name="Filtered Hilbert Transform   ",
+    name="Filtered Hilbert Envelope   ",
     # x=X,
     y=hilbert_envelope_filtered,
     mode="lines",
-    line_shape='spline',
+    # line_shape='spline',
     line=dict(
-        width=1,
+        width=2,
         color="black",
     )
   )

### 2021-05-09 Papers/01_Pulses_Envelope/images/04Coordinates.py

@@ -49,8 +49,8 @@ Y = Yp * scaling
 '''                              PLOT LINES                                    '''
 '''============================================================================'''
 fig = make_subplots(
-    rows=2, cols=1, shared_xaxes=False, vertical_spacing=0.01,
-    subplot_titles=("<b>Original basis</b>", "<b>Standard basis</b>")
+    rows=2, cols=1, shared_xaxes=False, #vertical_spacing=0.01,
+    subplot_titles=("<b>Original coordinates</b>", "<b>Cartesian coordinates</b>")
     )
 
 FONT = dict(
@@ -77,10 +77,10 @@ fig.update_layout(
 # fig.layout.xaxis.title.font=FONT
 # fig.layout.yaxis.title.font=FONT
 
-fig.update_xaxes(showline=False, showgrid=False, zeroline=False, showticklabels=False, title=dict(font=FONT), row=1, col=1)
+fig.update_xaxes(showline=False, showgrid=False, zeroline=False, showticklabels=False, title=dict(text="$ i $",font=FONT), row=1, col=1, title_standoff=.0)
 fig.update_xaxes(showline=False, showgrid=False, zeroline=False, showticklabels=True, 
-title=dict(text="$ x $",font=FONT), row=2, col=1)
-fig.update_yaxes(showline=False, showgrid=False, zeroline=False, title=dict(text="$ y $",font=FONT), row=1, col=1, tickvals=[i for i in range(0, 3, 2)])
+title=dict(text="$ x \equiv i $",font=FONT), row=2, col=1)
+fig.update_yaxes(showline=False, showgrid=False, zeroline=False, title=dict(text="Amplitude",font=FONT), row=1, col=1, tickvals=[i for i in range(0, 3, 2)])
 fig.update_yaxes(showline=False, showgrid=False, zeroline=False, 
 title=dict(text="$ y $",font=FONT), row=2, col=1, tickvals=[i for i in range(0, 11, 2)])
 
@@ -106,41 +106,6 @@ fig.add_trace(
   ), row=2, col=1
 )
 
-# for i in range(1, len(Xp)):  
-#   fig.add_annotation(
-#     x=Xp[i],  # arrows' head
-#     y=Yp[i],  # arrows' head
-#     ax=Xp[i - 1],  # arrows' tail
-#     ay=Yp[i - 1],  # arrows' tail
-#     xref='x',
-#     yref='y',
[…]

### 2021-05-10 Papers/01_Pulses_Envelope/00ElsevierPulsesEnvelope.tex
revising

@@
 \linenumbers
+
 \section{Introduction}
-Envelope detection, also known as demodulation, is ubiquitous in both analogue and digital signal processing (DSP) \cite{2011CaetanoImproved}.
-Nevertheless, the literature in this area is fragmented \cite{2017LyonsDigital}, with most envelope detection techniques being designed to account for very specific settings, like pure sinusoids with moderate noise content, for example.
+Envelope detection, also known as demodulation, is ubiquitous in both analog and digital signal processing (DSP) \cite{2011CaetanoImproved}.
+Nevertheless, the literature in this area is fragmented \cite{2017LyonsDigital}, with most envelope detection techniques being designed to account for very specific settings, in general involving narrowband signals with moderate noise content.
 
@@
 
-In many contexts, however, the temporal amplitude envelope of a signal plays a prominent role: according to \cite{2017QiRelative}, for example, the envelope is at least as important as the fine structure of a sound wave in the context of the intelligibility of Mandarin tones.
+In many contexts, however, the temporal amplitude envelope of a signal plays a prominent role: according to \cite{2017QiRelative}, for example, the envelope is at least as important as the fine structure of a sound wave in the context of intelligibility of Mandarin tones.
 This is also the case for the English language \cite{1995ShannonSpeech}, where even envelopes modulating mostly noise are still capable of conveying meaning.
 
-The temporal envelope also helps to impart emotion and identity to the human voice \cite{2018ZhuContributions} and music: Concert halls with envelope preserving characteristics are deemed more pleasant \cite{2011LokkiEngaging}, for example.
+The temporal envelope also helps to impart emotion and identity to the human voice \cite{2018ZhuContributions} and music: Concert halls with envelope preserving characteristics are deemed more pleasant \cite{2011LokkiEngaging}.
 
-Specially when dealing with broadband signals, approaches tailored to specific applications, exploiting a priori knowledge of the waves, are prevalent, such as the one presented by \cite{2014YangFast} for the distributed monitoring of fibre optic or the one formulated by \cite{2018AssefModeling} in the context of medical ultrasound imaging.
+Specially when dealing with broadband signals, approaches tailored to specific applications, exploiting prior knowledge of the waves, are prevalent, such as the one presented by \cite{2014YangFast} for the distributed monitoring of fiber optic or the one formulated by \cite{2018AssefModeling} in the context of medical ultrasound imaging.
 
@@
 
-Considering the standard 2d representation of a discrete wave, where the horizontal axis represents time and the vertical axis represents the amplitude, one can think of an individual sample as a point.
-Provided a coherent mapping from this space to the standard Cartesian plane is available, the original DSP problem can be transformed into the geometric problem of finding the shape of a set of points.
+Considering the standard graphic representation of a discrete wave, where the horizontal axis represents time and the vertical axis represents the amplitude, one can think of each sample as a point in a plane.
+Provided a coherent mapping from this space to the Cartesian plane is available, the original DSP problem can be transformed into the geometric problem of finding the shape of this set of points.
 
-One direct way of defining the shape of a set of points in $ \mathbb{R}^2 $ is using the concept of a convex hull, as defined in \cite{2008BergComputational}.
+One direct way of defining the shape of a set of points in $ \mathbb{R}^2 $ is by using the concept of a convex hull, as defined in \cite{2008BergComputational}.
 The convex hull of a set of points $ S $ in two-dimensional space is the smallest convex subset of the plane that contains all the points in $ S $.
@@
 
-When the general shape of a set of points present prominent concavities, the convex hull overestimates the boundaries.
-To address this limitation \cite{1983Edelsbrunnershape} introduced the concept of alpha shapes, also known as concave hull: a mathematically well-defined generalization of the convex hull of a finite set of points, closely related to the Delaunay triangulation and Voronoi diagrams of those points.
+When the general shape of a set of points presents prominent concavities, the convex hull overestimates the boundaries.
+To address this limitation, \cite{1983Edelsbrunnershape} introduced the concept of alpha shapes, also known as concave hulls: a mathematically well-defined generalization of the convex hull of a finite set of points, closely related to the Delaunay triangulation and Voronoi diagrams of those points.
 
-% TODO
-The intuition behind the alpha shapes algorithm is that only points that can be touched by a disc of radius $ \alpha $ coming from an infinite distance towards the baricenter of the set of points are part of the frontier.
-The polygon obtained linking those points forms the concave hull of this set.
+The intuition behind the alpha shapes algorithm is that, given a set of points and a fixed radius $ \alpha $, if it is possible to construct a circle containing one or more points of the set in such a way that the disc defined by the circle does not contain any other point of the set, then these points are part of the frontier.
+The polygon obtained linking all such points forms the alpha shape, or concave hull, of this set.
 
-Once a radius $ \alpha $ is defined, alpha shapes exhibit the uniqueness property: for a given set of points and a given radius, only one concave hull exists.
-When the radius $ \alpha $ of the circle tends to infinity, concave hulls tend to the convex hulls.
-This
+Fig.
+\ref{fig:AlphaShape} illustrates an alpha shape, for $ \alpha=10 $, of a set of points.
+Note that two points of the set lie at the circle illustrated, but the open disc defined by it contains no points of the set.
+An analogous construction is possible for every point at the frontier.
+For all the interior points, it is impossible to construct a circle of radius 10 that contains one or more of them without defining a disc containing at least one point of the set.
+
+\begin{figure}[ht!]
+  \centering
+    \includegraphics{images/00AlphaShape.pdf}
+  \caption{The alpha shape of a set of points for $ \alpha=10 $. A circle of radius 10, containing two points of the set, but defining a disc that doesn't contain any point of the set, is also shown.}
+  \label{fig:AlphaShape}
+\end{figure}
+
+Once a radius $ \alpha $ is defined, alpha shapes exhibit the uniqueness property: for a given set of points and a given radius, only one alpha shape exists.
+When the radius $ \alpha $ of the circle tends to infinity, alpha shapes, also known as concave hulls, tend to the convex hulls.
 
 Consider, for example, the artificially generated discrete wave in Fig.
-\ref{fig:Hulls} below, where the links between consecutive samples were deemphasized to highlight the “set of points” nature of a signal: two alpha shapes with different values for $ \alpha $ and the convex hull are illustrated.
+\ref{fig:Hulls} below, where the links between consecutive samples were de-emphasized to highlight the “set of points” nature of a signal: two alpha shapes, with different values for $ \alpha $, and the convex hull are illustrated.
 As $ \alpha $ increases, the contour of the set of points becomes smoother, and closer to the convex hull.
@@
     \includegraphics{images/00ConvexHull.pdf}
-  \caption{A discrete signal interpreted as a set of points in the Cartesian plane, and the convex and concave hull at $ \alpha=300 $ defined by it.}
+  \caption{A discrete signal interpreted as a set of points in the Cartesian plane, two alpha shapes, and the convex hull defined by it.}
   \label{fig:Hulls}
@@
 
-The alpha shapes approach is used in areas such as detection of features in images \cite{2016VarytimidisAlpha}, reconstruction of surfaces from a cloud of points \cite{2015WuAutomated} and spectroscopy \cite{2019XuModeling}.
+The alpha shapes theory is used in areas such as detection of features in images \cite{2016VarytimidisAlpha}, reconstruction of surfaces from a cloud of points \cite{2015WuAutomated} and spectroscopy \cite{2019XuModeling}.
 
-This last work, which involves the estimation and removal of the Blaze function (a kind of envelope) of an echelle spectrograph, being particularly illustrative of the potential synergy between geometric and DSP approaches.
+This last work, which involves the estimation and removal of the Blaze function (a kind of envelope) of an echelle spectrograph, is particularly illustrative of the potential synergy between geometric and DSP approaches.
 
@@
 
-In Section \ref{sec:representation} we explore some simplifications to the task of detecting the envelope of a signal that arise from the envelope definition adopted in this work and simplify the task of envelope detection.
+In Section \ref{sec:representation} we explore some simplifications to the task of detecting the envelope of a signal that arise from the envelope definition adopted in this work.
 
-Section \ref{sec:mapping} explores the geometric characteristics of envelope detection and formalizes how this classic DSP problem can be translated to the geometric problem of finding the shape of a set of points.
+Section \ref{sec:mapping} explores the geometric characteristics of envelope detection and formalizes how this classic DSP problem can be transformed into the geometric problem of finding the shape of a set of points.
 
@@
 
-Section \ref{sec:quality} discusses the difficulties in the assessment of the quality of an envelope, proposing qualitative and quantitative approaches, and demonstrates that the method propose conforms to four requirements for a physically plausible envelope proposed in \cite{1996Loughlinamplitude}.
+Section \ref{sec:quality} discusses the difficulties in the assessment of the quality of an envelope, proposing qualitative and quantitative approaches, and demonstrates that the method presented conforms to four requirements for a physically plausible envelope proposed in \cite{1996Loughlinamplitude}.
 
-In Section \ref{sec:comparison} we compare the presented algorithm with traditional envelope detection methods, both in terms of accuracy and time and present, in Section \ref{sec:extensions},  potential applications of the algorithm, some of which might not be immediately obvious.
+In Section \ref{sec:comparison} we compare the presented algorithm with traditional envelope detection methods in terms of accuracy and time and present, in Section \ref{sec:extensions},  potential applications of the algorithm, some of which might not be immediately obvious.
 
-% TODO %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-Interactive visualizations are available from the repository prepared for this work \cite{2020TarjanoEnvelope}, that also contains an implementation of the method in pure Python.
-A Python module available at the Python Package Index (PyPI) installable via the command `pip install signal-envelope` for convenience, is also available.
+Links to interactive visualizations are available from the repository prepared for this work \cite{2020TarjanoEnvelope}, which also contains an implementation of the method in pure Python.
+A Python module can be installed from the Python Package Index (PyPI) with the command `pip install signal-envelope`.
 Besides, those working under Windows 64-bit can benefit from a specialized dynamic library coded in C++ that is automatically used within this module, and can be separately obtained via the release section in the repository.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\section{Temporal envelope}\label{sec:envelope}
+\section{Temporal envelopes}\label{sec:envelope}
 
@@
 
-While there is no general formal definition of the envelope of broadband signals \cite{2014Yangnovel,2015Yangtheoretical,2019Jiaempirical}, the general intuition is that an envelope is a smooth curve situated as close as possible to the signal \cite{2019Jiaempirical}.
+While there is no general formal definition for the envelope of broadband signals \cite{2014Yangnovel,2015Yangtheoretical,2019Jiaempirical}, the consensus is that an envelope is a smooth curve situated as close as possible to the signal \cite{2019Jiaempirical}.
 
@@
 \ref{fig:AnalyticSignal} exemplifies the Hilbert envelope, for a pure sinusoid modulated by a polynomial, illustrating the difference in the smoothness of the envelope in the presence and absence of Gaussian white noise.
-It is readily noticeable from the Fig.
+It is readily noticeable from Fig.
 \ref{fig:AnalyticSignal} that, as soon as noise is introduced in the original signal, it is directly reflected in the envelope.
@@
     \includegraphics{images/01AnalyticSignal.pdf}
-  \caption{Envelope of a pure sinusoid with a local frequency of 20 cycles modulated by a polynomial of degree 3, as obtained by the Hilbert Transform approach.
-In the first half, the sinusoid is free of noise, while in the second half Gaussian noise with a standard deviation of $ 1/10 $ of the maximum amplitude of the wave was added to the base signal.}
+  \caption{Envelope of a pure sinusoid with a local frequency of 20 cycles, modulated by a polynomial of degree 3, as obtained by the Hilbert Transform approach.
+In the first half of the samples, the sinusoid is free of noise, while in the second half Gaussian noise, with a standard deviation of $ 1/10 $ of the maximum amplitude of the wave, was added to the base signal.}
   \label{fig:AnalyticSignal}
@@
 For a real world signal, such as an alto singer sustaining a steady note as shown in Fig.
-\ref{fig:AnalyticSignalAlto} the Hilbert transform alone provides a poor envelope, since it retains most of the frequencies of the underlying wave.
+\ref{fig:AnalyticSignalAlto}, the Hilbert transform alone provides a poor envelope, since it retains most of the frequencies of the underlying wave.
 With further processing, however, such as filtering, a smoother output can be obtained.
@@
     \includegraphics{images/01AnalyticSignalAlto.pdf}
-  \caption{Representation of an alto singer sustaining a steady note and a comparison of the envelopes obtained by the Hilbert transform with and without further filtering.}
+  \caption{A section of the representation of an alto singer sustaining a steady note and a comparison of the envelopes obtained by the Hilbert transform with and without further filtering.}
   \label{fig:AnalyticSignalAlto}
@@
 
-$ E(t) $ represents the slow varying part of the wave, also known as the (temporal) envelope, modulator component, or amplitude modulation (AM) of $ W(t) $, while $ C(t) $ models its fast varying part, called throughout the literature as its (temporal) fine structure, carrier wave, or frequency modulation (FM).
+$ E(t) $ represents the slow varying part of the wave, also known as the (temporal) envelope, modulator component, or amplitude modulation (AM) of $ W(t) $, while $ C(t) $ models its fast-varying part, called throughout the literature as its (temporal) fine structure, carrier wave, or frequency modulation (FM).
 
@@
 
-We address this problem by assuming $ C(t) $ normalized between $ \{-1, 1\}$ or, in other words $ |\max(C(t))| = 1 $.
-This assumption doesn't cause any loss of generality since, given an arbitrary $ C(t) $, we can obtain its normalization by dividing the function by its absolute global maximum, that is, $ \hat{C}(t) = C(t) / \max(|C(t)|) $, provided that $ \max(|C(t)|) \ne 0 $.
+We address this problem by assuming $ C(t) $ to be normalized between $ \{-1, 1\}$ or, in other words $ |\max(C(t))| = 1 $.
+This assumption does not cause any loss of generality since, given an arbitrary $ C(t) $, we can obtain its normalization by dividing the function by its absolute global maximum, that is, $ \hat{C}(t) = C(t) / \max(|C(t)|) $, provided that $ \max(|C(t)|) \ne 0 $.
 
-In this work we are concerned with the discrete version of this problem: given a finite digital wave represented by the vector $ \mathbf{w} \in \mathbb{R}^n $, instead of the continuous function $ W(t) $, obtaining the temporal envelope $ \mathbf{e} $ of $ \mathbf{w} $.
+In this work we are concerned with the discrete version of the envelope detection problem: given a finite digital wave represented by the vector $ \mathbf{w} \in \mathbb{R}^n $, obtaining the temporal envelope $ \mathbf{e} $ of $ \mathbf{w} $.
 
@@
   \mathbf{e}, \mathbf{c} & \in \mathbb{R}^n \\
-  \mathbf{e} & = (e_0, e_1, \cdots, e_{n-1}), w_i \ge 0 \ \forall \ i, \ 0 \le i \le n-1 \\  
-  \mathbf{c} & = (c_0, c_1, \cdots, c_{n-1}), |c_i| \le 1 \ \forall \ i, \ 0 \le i \le n-1 \\
+  \mathbf{e} & = (e_0, e_1, \cdots, e_{n-1}), e_i \ge 0 \ \forall \ i, \ 0 \le i \le n-1 \\  
+  \mathbf{c} & = (c_0, c_1, \cdots, c_{n-1}), |c_i| \le 1 \ \forall \ i, \ 0 \le i \le n-1, |\max(\mathbf{c})| =1 \\
   \mathbf{w} & = \mathbf{e} \odot \mathbf{c} \ \therefore \ w_i = e_i c_i \ \forall \ i, \ 0 \le i \le n-1
@@
 \ref{fig:Envelope} a signal arising from the multiplication of a known carrier wave \textbf{c} and a known envelope \textbf{e} is shown; this helps to illustrate a key concept: note from Fig.
-\ref{fig:Envelope} that, only at some local extrema, the envelope can be inferred without prior knowledge of the underlying carrier, being in fact equal to the value of the wave $ w_i $.
-$ c_i = 1 $ at those points, rendering $ e_i = 1 w_i $.
+\ref{fig:Envelope} that, only at some local extrema, the envelope can be inferred without prior knowledge of the underlying carrier, being, in fact, equal to the value of the instantaneous amplitude of the wave $ w_i $.
+That is, $ c_i = 1 $ at those points, rendering $ e_i = 1 w_i $.
 
-This realization that the envelope of a wave is defined only at some of the wave's local extrema is crucial, as we can use it to generate a simplified representation of a signal that will make the algorithm more efficient by considering only the local extrema of a given signal.
+This realization that the envelope of a wave is defined only at some of the wave's local extrema is important, as we can use it to generate a simplified representation of a signal that will make the algorithm more efficient.
 
-\section{Simplified Representation} \label{sec:representation}
+\section{Simplified Representation of a signal} \label{sec:representation}
 
-We start with the definition of a pulse: each time the value of $ w_i$ changes sign, that is, every time the amplitude of the discrete wave \textbf{w} crosses the horizontal axis, the beginning of a pulse is defined, with the next crossing defining its end.
-This definition is in line with the one presented in \cite{national1996telecommunications}, where a pulse is defined as a rapid change in the amplitude of a signal, followed by a fast return to the baseline value; zero, in our case.
+We start with the definition of a pulse: each time the value of $ w_i $ changes sign, that is, every time the instantaneous amplitude of the discrete wave \textbf{w} crosses the horizontal axis, the beginning of a pulse is defined, with the next crossing defining its end.
+This interpretation of a pulse is in line with the one presented in \cite{national1996telecommunications}, where a pulse is defined as a rapid change in the amplitude of a signal, followed by a fast return to the baseline value; zero, in our case.
 
-From \ref{fig:Envelope} and the discussion in Section \ref{sec:envelope}, we can see that only the maximum point of a positive pulse or the minimum point of a negative pulse can potentially be equal to the envelope and are, thus, our only points of interest.
-We can then proceed, for the rest of the method, considering only those points, to great computational economy.
+From Fig.
+\ref{fig:Envelope} and the discussion in Section \ref{sec:envelope}, we can see that only the maximum point of a positive pulse or the minimum point of a negative pulse can potentially be equal to the envelope and are, thus, our only points of interest.
+We can then proceed, for the rest of the method, considering only those points, to a great computational economy.
 
@@
 This relation is illustrated in Fig. \ref{fig:Pulses}. We call it P to emphasize the connection with the pulses of \textbf{w}.
-While the exact relation between $ n $, the size of the original signal, and $ m $, the size of the set of points, is dependent on factors such as the frequency and sampling rate of the original wave, the inequality $ m \le n $ holds for all discrete signals.
-In practice, $ m \ll n $ for most real world signals.
+While the exact relation between $ n $, the number of samples of the original signal, and $ m $, the cardinality of the set of points, is dependent on factors such as the frequency and sampling rate of the original wave, the inequality $ m \le n $ holds for all discrete signals.
+In practice, $ m \ll n $ for most real-world signals.
 
@@
 
-To achieve that, we divide the ordinate of each point by the average of all ordinates, effectively cancelling the unit of the vertical components of the points.
+To achieve that, we divide the ordinate of each point by the average of all ordinates, effectively canceling the unit of the vertical components of the points.
 If we proceeded similarly in relation to the horizontal components,
@@
 It is best to leave the abscissas intact and multiply the now unitless vertical coordinates of each point by the average of the difference between the abscissa of a point and the abscissa of the immediately posterior.
-In this way, we can consider both axes as having unit $ i $, and the relationship is preserved, making it easier to recover the envelope points in the original basis later.
+In this way, we can consider both axes as having unit $ i $, and the relationship is preserved, making it easier to recover the envelope points in the original coordinates later.
 
@@
 P_{\mathbf{b}} &= (x_1, x_2) \\
-\beta &= \frac{\left( \frac{i_{m-1} - i_m}{m-1} \right)}{\left( \frac{\sum_{j=0}^{m-1} \lvert w_j \lvert}{m} \right)}  \\
+\beta &= \frac{\left( \frac{i_{j-1} - i_j}{m-1} \right)}{\left( \frac{\sum_{j=0}^{m-1} \lvert w_j \lvert}{m} \right)}  \\
 P_{\mathbf{s}} &= (x_1, \beta x_2) \quad \forall \  P \in \text{P} \\ 
@@
 
-The effect of this normalization can be seen in the points shown in Fig.
-\ref{fig:Coordinates}, shown in scale for both basis.
-We can consider our points now residing in the Cartesian coordinate system.
+The effect of this transformation can be seen in the points represented in Fig.
+\ref{fig:Coordinates}, shown in scale for both coordinate systems.
+We can consider our points residing in the Cartesian coordinate system after the transformation.
 
@@
     \includegraphics{images/04Coordinates.pdf}
-  \caption{The set of points P, in both the original (top) and standard (bottom) coordinate basis, shown in scale.}
+  \caption{The set P of points, in both the original (top) and standard (bottom) coordinate systems, shown in scale.}
   \label{fig:Coordinates}
@@
 
-\section{Discrete Curvature Estimation} \label{sec:curvature}
+\section{Discrete curvature estimation} \label{sec:curvature}
 
-Having established a transformation from the original space of the two-dimensional representation of a discrete wave to the Cartesian coordinate system and, thus, a way to consistently transform our simplified description of a discrete signal into a set of points in the plane, the shape of which can be estimated with the use of geometric algorithms.
+Having established a transformation from the original space of the two-dimensional representation of a discrete wave to the Cartesian coordinate system, we are now in a position to apply the geometric concept of the shape of a set of points.
 
-The goal of those algorithms, such as the convex hull and the alpha-shapes, is to discriminate which points of a set are part of its frontier.
+Algorithms such as the convex hull and the alpha shapes solve that problem by discriminating which points of a set are part of its frontier.
 
-A formal definition of alpha-shapes is given at \cite{1983Edelsbrunnershape}.
-In the context of this work, gaining the intuition of the algorithm is more important, and thus the explanation of the alpha-shapes algorithm presented in \cite{1994EdelsbrunnerThree} can be more insightful.
+A formal definition of alpha shapes is given at \cite{1983Edelsbrunnershape}.
+In the context of this work, however, gaining the intuition of the algorithm is more important, and thus the explanation presented in \cite{1994EdelsbrunnerThree} can be more insightful.
 
-The explanation, translated to our context, invites one to think of the plane were the samples are represented as a soft material, such as Styrofoam.
-The samples themselves are a harder material, such as rock, and occupy a fixed position in this plane.
+The explanation, translated to our context, invites one to think of the plane where the points of P are represented as a soft material, such as Styrofoam.
+The points themselves are made of a harder material, such as rock, and occupy a fixed position in this plane.
 The alpha shapes algorithm consists of carving this plane from the outside of those points, with a circular eraser of fixed radius $ \alpha $.
-In the end of the process, the remaining Styrofoam that couldn't be erased due to the impeding rocks is the alpha-shape of this set of points.
+At the end of the process, the remaining Styrofoam that couldn't be erased due to the impeding rocks is the alpha-shape of this set of points.
 
-Alternatively, one can picture a circle of fixed radius being rolled above the signal, and marking the points it touches as envelope points.
+Alternatively, one can picture a circle of fixed radius being rolled above the set P, and marking the points it touches as part of the envelope.
 
-To employ an alpha shapes inspired algorithm in this set of points, however, the radius $ r $ of such circle must be estimated, and for that a measure of discrete curvature is needed.
+To employ an alpha shapes inspired algorithm, however, the radius $ r $ of such a circle must be estimated, and for that, a measure of discrete curvature is needed.
 
@@
 The two general approaches are the derivation of direct methods that use characteristics of the discrete curve to calculate the curvature, or the use of the curvature of a smooth, continuous curve fitted to the original discrete curve \cite{2001CoeurjollyDiscrete}.
+Since we are dealing with a potentially large set of points, it is important for this measure to act locally, to assure $ O(n) $ complexity, as is generally not the case for approaches that involve curve fitting.
 
-Concerning direct methods \cite{2014CarrollSurvey}, for example, derive three such definitions based on the approximation of a circle by an inscribed, centered and circumscribed polygon.
+Concerning direct methods \cite{2014CarrollSurvey}, for example, derive three such definitions based on the approximation of a circle by an inscribed, centered, and circumscribed polygon.
 In the context of three-dimensional meshes, \cite{2016VasaMultivariate} evaluate a range of existing estimators from a multivariate point of view.
-
 Those approaches define the discrete curvature in relation to the vertices of a discrete curve (or mesh, in the three-dimensional case).
-Instead, we are interested in the change of direction, what the term curvature ultimately means, between two adjacent points.
-
-Since we are dealing with a potentially large set of points, it is important for this measure to act locally, to assure $ O(n) $ complexity.
+Since we are interested in the change of direction, what the term curvature ultimately means, between two adjacent points, a definition of discrete curvature over the edges of a discrete curve would be preferable.
 
@@
     \includegraphics{images/04Vectors.pdf}
-  \caption{The set of points P and the set V of vectors between two adjacent points of P, in the normalized coordinate system.}
+  \caption{The set P of points and the set V of vectors between two adjacent points of P, in the Cartesian coordinate system.}
   \label{fig:Vectors}
@@
 
-We formalize the set of vectors V in Eq. \ref{eq:Vectors} below.
+We formalize the set V of vectors in Eq. \ref{eq:Vectors} below.
 

### 2021-05-10 Papers/01_Pulses_Envelope/images/00AlphaShape.py

@@ -1,0 +1,157 @@
+import sys
+print(sys.version)
+import plotly.graph_objects as go
+sys.path.append("c:/Users/tesse/Desktop/Files/Dropbox/0_Thesis/Python/01_Envelope")
+import numpy as np
+from scipy.spatial import ConvexHull
+from Helper import signal_to_pulses, get_pulses_area, alpha_shape, get_circle
+
+
+def draw_circle(xc, yc, r, fig, n=100, color="silver"):
+  '''draws circle as plotly scatter from center and radius'''
+  X = []
+  Y = []
+  pulses_color = "rgba(250, 140, 132, 0.5)"
+
+  for t in np.linspace(0, 2 * np.pi, n):
+    X.append(xc + r * np.cos(t))
+    Y.append(yc + r * np.sin(t))
+
+
+  fig.add_trace(
+    go.Scatter(
+      name=f"Circle of radius \u03B1={radius}",
+      # legendgroup="Circles",
+      x=X,
+      y=Y,
+      # showlegend=False,
+      # visible = "legendonly",
+      mode="lines",
+      line=dict(
+          width=1,
+          color="rgba(3, 19, 252, 0.8)"
+      )
+    )
+  )
+
+  fig.add_trace(
+    go.Scatter(
+      name="Open disc",
+      # legendgroup="Circles",
+      x=X,
+      y=Y,
+      fill="tozeroy",
+      fillcolor="rgba(3, 19, 252, 0.1)",
+      mode="none",
+      # line=dict(
+      #     width=1,
+      #     color=color
+      # )
+    )
+  )
+  return X, Y
+
+
+
+np.random.seed(1)
+# fps = 10
+n = 50+1
+
+X = np.arange(n)# / fps
+
+W = np.random.normal(0, np.sqrt(n), n)
+
+W -= W.min()
+
+radius = 10
+
+
+
+points = np.array([[x, w] for x, w in zip(X, W)])
+hull = ConvexHull(points)
+
+x0, y0 = get_circle(points[10][0], points[10][1], points[14][0], points[14][1], radius)
+
+'''============================================================================'''
+'''                              Alpha 1                                    '''
+'''============================================================================'''
+
+
+alpha_edges = alpha_shape(points, radius)
+alpha_edges = sorted(alpha_edges, key=lambda tup: tup[0])
+
+XX1 = []
+YY1 = []
+for e in alpha_edges:
+  idx0 = e[0]
[…]

### 2021-05-10 Papers/01_Pulses_Envelope/images/00ConvexHull.py

@@ -98,7 +98,7 @@ fig.update_yaxes(showline=False, showgrid=False, zerolinewidth=1, zerolinecolor=
 
 fig.add_trace(
   go.Scatter(
-    name=f"Alpha-Shape (\u03B1={alpha1})    ",
+    name=f"Alpha shape (\u03B1={alpha1})    ",
     # showlegend=False,
     x=XX1,
     y=YY1,
@@ -113,7 +113,7 @@ fig.add_trace(
 
 fig.add_trace(
   go.Scatter(
-    name=f"Alpha-Shape (\u03B1={alpha2})    ",
+    name=f"Alpha shape (\u03B1={alpha2})    ",
     # showlegend=False,
     x=XX2,
     y=YY2,

### 2021-05-10 Papers/01_Pulses_Envelope/00ElsevierPulsesEnvelope.tex
reworking algorithm

@@
 We thus proceed to define a discrete curvature measure over the edges of a discrete curve.
-To that end we are going to apply the definition of smooth curvature as the rate of change of the unit tangent to a curve.
+To that end, we are going to apply the definition of smooth curvature as the rate of change of the unit tangent to a curve.
 This definition is convenient, since it is equivalent to that of the osculating circle \cite{2016VasaMesh}, the value we are interested in finding.
@@
 
-The rationale then is to find, for each vector $ v_k $, the radius $ r_k $ of the equivalent circle whose tangent has the same change in direction, in the same horizontal distance, as the vector, as shown in Fig.
+The rationale then is to find, for each vector $ v_k $, the radius $ r_k $ of the equivalent circle whose tangent has the same change in direction, in the same horizontal distance, as the vector and the horizontal direction, as shown in Fig.
 \ref{fig:DiscreteCurvature}.
@@
     \includegraphics{images/05DiscreteCurvature.pdf}
-  \caption{The tangent unit vector of the circle changes from the horizontal direction in $ \hat{u}_0 $ to an inclination of $ \theta_k $ in $ \hat{u}_1 $, $ \theta_k $ being the angle that the vector $ v_k $ makes with the horizontal direction.}
+  \caption{The unit vector tangent to the circle changes from the horizontal direction in $ \hat{u}_0 $ to an inclination of $ \theta_k $ in $ \hat{u}_1 $, $ \theta_k $ being the angle that the vector $ v_k $ makes with the horizontal direction.}
   \label{fig:DiscreteCurvature}
@@
 
-For calculation purposes, however, this can pose a problem, as a zero could make its way into the denominator of a fraction; in the actual algorithm, the quantity calculated for each vector is the curvature itself, that is, the inverse of each radius $ r_k $.
+For calculation purposes, however, the formulation in Eq.
+\ref{eq:Radius} poses a problem, as a zero could make its way into the denominator of a fraction; in the actual algorithm, the quantity calculated for each vector is the curvature itself, that is, the inverse of each radius $ r_k $.
 In this way, the problem is avoided, since $ v_{k,x} $ is guaranteed to be greater than zero.
@@
 We need now to construct an algorithm, using the radius just obtained, to identify the points that belong to the envelope.
-Generally, alpha-shapes based algorithms resort to the construction of the Delaunay triangulation of the set of points, that is latter filtered to contain only the outer edges.
+Generally, alpha shapes based algorithms resort to the construction of the Delaunay triangulation of the set of points, that is latter filtered to contain only the outer edges.
 We can adopt a more straightforward method, as our points have a sequential structure.
 
-It is worth noting that, although that is the case throughout this work, the original signal doesn't need to be sampled at equal intervals.
-
-Let $ x_o, y_o \in \mathbb{R}, \quad 0 \le y_o < +\infty $ be the coordinates of the center of a circle with radius $ r $, that can be placed anywhere in the upper plane of the coordinate system;
+Let $ x_o, y_o \in \mathbb{R}, \quad 0 \le y_o $ be the coordinates of the center of a circle with radius $ r $, that can be placed anywhere in the upper plane of the coordinate system;
 
@@
 & \forall \ P_b, P_c \mid (P_{b,x} - x_o)^2 + (P_{b,y} - y_o)^2 = (P_{c,x} - x_o)^2 + (P_{c,y} - y_o)^2 = r^2, \\
-& P_a, P_b, P_c \in \text{P}, \quad P_a \ne P_b \ne P_c \quad \text{and} \quad x_o, y_o \in \mathbb{R}, \quad 0 \le y_o < +\infty
+& P_a, P_b, P_c \in \text{P}, \quad P_a \ne P_b \ne P_c \quad \text{and} \quad x_o, y_o \in \mathbb{R}, \quad 0 \le y_o
 \end{split}
@@
 
-The algorithm follows directly from the definition in \ref{eq:InEnvelope} after noting that the first and last members of P will always be part of the envelope, since they are reachable by a circle of any radius $ r $.
+The algorithm follows directly from the definition in \ref{eq:InEnvelope} after noting that the first and last members of P will always be part of the envelope, since they are reachable by any circle of any radius $ r $.
 From $ P_0 $, our first pivot point, we construct circles passing through $ P_1, P_2, \cdots, P_d $ until a circle that doesn't contain any point in P is found; $ P_d $ then becomes the new pivot point and is included in E, and this process continues until $ P_{m-1} $ is reached.

### 2021-05-10 Papers/01_Pulses_Envelope/00ElsevierPulsesEnvelope.tex
revised

@@
 We need now to construct an algorithm, using the radius just obtained, to identify the points that belong to the envelope.
-Generally, alpha shapes based algorithms resort to the construction of the Delaunay triangulation of the set of points, that is latter filtered to contain only the outer edges.
+Generally, alpha shapes based algorithms resort to the construction of the Delaunay triangulation of the set of points, which is later filtered to contain only the outer edges.
 We can adopt a more straightforward method, as our points have a sequential structure.
@@
 
-Let $ P_a, P_b, P_c \in \text{P}, \quad P_a \ne P_b \ne P_c $ be three different points of the set P;
+Let $ P_a, P_b \in \text{P}, \quad P_a \ne P_b $ be two different points of the set P;
 
-For all circles of center $ x_o, y_o $ and radius $ r $ that can be constructed to pass through any two different points of P, that is, $ \forall \quad P_b, P_c $ such that $ (P_{b,x} - x_o)^2 + (P_{b,y} - y_o)^2 = (P_{c,x} - x_o)^2 + (P_{c,y} - y_o)^2 = r^2 $, $ P_a $ will be part of the set $ E $ of the points belonging to the envelope if and only if none of those circles contain $ P_a $, or $ P_a \in \text{E} \leftrightarrow (P_{b,x} - x_o)^2 + (P_{b,y} - y_o)^2 > r^2 $.
+For all circles of center $ x_o, y_o $ and radius $ r $ that can be constructed to pass through any two different points of P, that is, all circles such that $ (P_{a,x} - x_o)^2 + (P_{a,y} - y_o)^2 = (P_{b,x} - x_o)^2 + (P_{b,y} - y_o)^2 = r^2 $, $ P_a $ and $ P_b $ will be part of the set $ E $ of the points belonging to the envelope if and only if exist one such circle that defines a disc that does not contain any other point of P.
 Summarizing this, we have Eq.
@@
 \begin{split}
-& P_a \in \text{E} \leftrightarrow (P_{b,x} - x_o)^2 + (P_{b,y} - y_o)^2 > r^2, \\
-& \forall \ P_b, P_c \mid (P_{b,x} - x_o)^2 + (P_{b,y} - y_o)^2 = (P_{c,x} - x_o)^2 + (P_{c,y} - y_o)^2 = r^2, \\
-& P_a, P_b, P_c \in \text{P}, \quad P_a \ne P_b \ne P_c \quad \text{and} \quad x_o, y_o \in \mathbb{R}, \quad 0 \le y_o
+& (P_{a,x} - x_o)^2 + (P_{a,y} - y_o)^2 = (P_{b,x} - x_o)^2 + (P_{b,y} - y_o)^2 = r^2 \\
+& P_a, P_b \in \text{E} \leftrightarrow (P_{c,x} - x_o)^2 + (P_{c,y} - y_o)^2 > r^2 \ \forall \ P_c \in \text{P} - \{P_a, P_b\} \\
+& P_a, P_b, P_c \in \text{P}, \quad P_a \ne P_c, P_b \ne P_c \quad \text{and} \quad x_o, y_o \in \mathbb{R}, \ 0 \le y_o
 \end{split}
@@
 
-The algorithm follows directly from the definition in \ref{eq:InEnvelope} after noting that the first and last members of P will always be part of the envelope, since they are reachable by any circle of any radius $ r $.
+The algorithm follows directly from the definition in Eq.
+\ref{eq:InEnvelope} after noting that the first and last points of P will always be part of the envelope.
+Since $ P_a $ can be equal to $ P_b $, it is always possible to construct a circle containing those points, by placing the center of the circle to the left of or right of the first or last point, respectively.
+The discs defined by those circles can't contain any other point.
+
 From $ P_0 $, our first pivot point, we construct circles passing through $ P_1, P_2, \cdots, P_d $ until a circle that doesn't contain any point in P is found; $ P_d $ then becomes the new pivot point and is included in E, and this process continues until $ P_{m-1} $ is reached.
@@
 
-The set E obtained via algorithm \ref{FindEnvelope} is a subset of P, and thus also a set of points in the Euclidean space.
+The set E obtained via the algorithm \ref{FindEnvelope} is a subset of P, and thus also a set of points in the Euclidean space.
 
-Recall from section \ref{sec:mapping} that the vertical coordinates of these points are currently scaled by the change of basis described in Eq.
-\ref{eq:Basis}, in relation to the original signal as represented in his original basis.
-We could invert this transformation, but since we were careful to maintain the direct link between the horizontal coordinates in both basis, we can simply consider the horizontal coordinates of P as the indices of samples of the original discrete signal, whose absolute value are part of the envelope.
+Recall from section \ref{sec:mapping} that the vertical coordinates of these points are currently scaled, in relation to the original signal, by the transformation described in Eq.
+\ref{eq:Basis}.
+We could invert this transformation, but since we were careful to maintain the direct link between the horizontal coordinates in both coordinate systems, we can simply consider the horizontal coordinates of P as the indices of the samples of the original discrete signal, and retrieve the envelope at each point directly.
 
-The most straightforward way to transform those points into the vector $ \textbf{e} \in \mathbb{R}^n $ is via linear interpolation, the approach used in this work.
+After converting E to the original coordinate system, the most straightforward way to transform those points into the vector $ \textbf{e} \in \mathbb{R}^n $ is via linear interpolation, the approach used in this work.
 
 It is worth noting, however, that many other procedures are available, both for the interpolation and the smooth approximation of points.
-Depending on the intended use of the envelope, k-curves interpolation \cite{2017Yank}is a specially interesting approach, since it guarantees that the maximum curvature occurs at the interpolated points.
+Depending on the intended use of the envelope, k-curves interpolation \cite{2017Yank} is a specially interesting approach, since it guarantees that the maximum curvature occurs at the interpolated points.
 
@@
 Figs.
-\ref{fig:spokenvoice}, \ref{fig:singer_voice} and \ref{fig:Bend} present discrete waves \textbf{w}, their extracted envelope \textbf{e}, and the inferred carrier \textbf{c}, obtained after dividing, element-wise, the original signal by its extracted.
+\ref{fig:spokenvoice}, \ref{fig:singer_voice} and \ref{fig:Bend} present discrete waves \textbf{w}, their extracted envelope \textbf{e}, and the inferred carrier \textbf{c}, obtained after dividing, element-wise, the original signal by its extracted envelope.
 
 From Fig.
-\ref{fig:spokenvoice} it can be seen that the algorithm performs well under the carrier metric.
+\ref{fig:spokenvoice} it can be seen that the algorithm performs well under the carrier criteria.
 This is especially remarkable for the case of spoken voice, a notorious complex wave with fast changes in pitch and high noise content.
@@
 
-Those characteristics are in stark contrast with those of the sang voice in Fig.
-\ref{fig:singer_voice}, where a recording of an alto singer sustaining a steady note is shown; there, the wave is almost periodic, with a stable waveform and envelope.
+Those characteristics are in stark contrast with those of the singing voice in Fig.
+\ref{fig:singer_voice}, where a recording of an alto singer sustaining a steady note is shown.
+There, the wave is almost periodic, with a stable waveform and envelope.
 
@@
 
-Fig. \ref{fig:Bend} shows the envelope in the case of a bend performed in an electric guitar. The variations in frequency in the signal didn't affect the envelope extracted.
+Fig. \ref{fig:Bend} shows the envelope in the case of a bend performed in an electric guitar. The fast frequency variations in the signal didn't affect the envelope extracted.
 
@@
 
-We comment below on how those conditions translate to the discrete setting, and show that the algorithm here presented satisfies the four conditions presented.
+We comment below on how those conditions translate to the discrete setting, and show that the algorithm proposed satisfies all the four conditions presented.
 
@@
 Recalling Eq.
-\ref{eq:Envelope}, where the relation between the original signal, the carrier and the envelope was defined as $ \mathbf{w} = \mathbf{e} \odot \mathbf{c} $, we can write the magnitude of an arbitrary frequency of the original signal ($ m_\mathbf{w}(f) $) and, likewise, the magnitude of an arbitrary frequency of the carrier ($ m_\mathbf{c}(f) $) in terms of the Discrete Fourier Transform (DFT) as in Eq.
+\ref{eq:Envelope}, where the relation between the original signal, the carrier and the envelope was defined as $ \mathbf{w} = \mathbf{e} \odot \mathbf{c} $, we can write the magnitude of an arbitrary frequency of the original signal ($ m_\mathbf{w}(f) $) and, likewise, the magnitude of an arbitrary frequency of the carrier ($ m_\mathbf{c}(f) $), in terms of the Discrete Fourier Transform (DFT) as in Eq.
 \ref{eq:DFT}:
@@
 
-The third condition states that a physically plausible algorithm should, for a signal $ \mathbf{w} $ of constant amplitude $ a > 0, \ a \in \mathbb{R} $ and constant frequency, of the form $ a \cos(\omega i + \phi) $, result in a constant envelope $ \mathbf{e} $ of magnitude $ a $ and o a carrier wave $ \mathbf{c} $ of the form $ \cos(\omega i + \phi) $.
+The third condition states that a physically plausible algorithm should, for a signal $ \mathbf{w} $ of constant amplitude $ a > 0, \ a \in \mathbb{R} $ and constant frequency, of the form $ a \cos(\omega i + \phi) $, generate a constant envelope $ \mathbf{e} $ of magnitude $ a $ and a carrier wave $ \mathbf{c} $ of the form $ \cos(\omega i + \phi) $.
 
@@
 To see that this is always the case, note that all the extrema of a sinusoid are of the form $ a (\pm 1) $ and $ P_j = (i_j, \lvert a (\pm 1) \lvert) = (i_j, a) \ \forall P_j \in \text{P} $.
-As $ \text{E} \subset \text{P} $, we have that the envelope $ \mathbf{e} = (a, a, \cdots, a) $ and is, thus constant.
+As $ \text{E} \subset \text{P} $, we have that the envelope $ \mathbf{e} = (a, a, \cdots, a) $ and is, thus, constant.
 Likewise, since $ \mathbf{w} = \mathbf{e} \odot \mathbf{c} $, we have $ c_i = a \cos(\omega i + \phi) / a = \cos(\omega i + \phi) \ \forall \ i, 0 \le i \le n - 1 $.
@@
 
-Recall from section \ref{sec:mapping} that, before the extraction of the envelope, the original points undergo a change of basis to lie in the Cartesian coordinate system.
+Recall from section \ref{sec:mapping} that, before the extraction of the envelope, the original points undergo a transformation to lie in the Cartesian coordinate system.
 After the extraction, the inverse operation is performed on the points of the Envelope.
@@
 
-In order to obtain a numeric indicator of the quality of an envelope, we construct a reference envelope, that trades off practical applicability in favour of mathematical formality.
+In order to obtain a numeric indicator of the quality of an envelope, we construct a reference envelope, that trades off applicability in favour of mathematical formality.
 In \cite{1992BruceCurves} the envelope is defined as the boundary of the region filled by a family of curves.
@@
 In practical applications, one seldom has access to a family of curves, instead being interested in obtaining the envelope of a single signal.
-Nevertheless, an approximation can be obtained, by shifting the original signal along its initial position in the horizontal axis, in order to cover the time of one period of the signal.
+Nevertheless, an approximation can be obtained, by shifting the original signal along its initial position in the horizontal axis by half of its period.
 Fig. \ref{fig:RefEnvelope} shows the reference envelope for the sound of a tom.
@@
 
-Since envelope is not smooth, being composed of a series of horizontal segments, as can be seen in \ref{fig:RefEnvelope}, besides being computationally expensive to calculate.
-However, arising from a mathematically sound base, it can serve as a reference to infer the quality of more empirical methods.
+The envelope obtained is not smooth, being composed of a series of horizontal segments, as can be seen in \ref{fig:RefEnvelope}, besides being computationally expensive to calculate.
+However, arising from a mathematically sound basis, it can serve as a reference to infer the quality of more empirical methods.
 
@@
 \item Smoothing: This is a relatively simple procedure that, applied to the absolute values of a discrete signal, can provide an approximation of the variation of its amplitude in time.
-We use the Savitzky-Golay algorithm \cite{1964SavitzkySmoothing} with a window width of 3001 samples and a cubic polynomial.
+We use the Savitzky-Golay algorithm \cite{1964SavitzkySmoothing} with a window width of 3001 samples and a cubic polynomial fitting.
 \item Filtering: The absolute value of the wave is filtered using a digital implementation of the Butterworth \cite{butterworth1930theory} low pass filter.
@@
 
-For a recording of the key 33 of a grand piano, where parts of the sound, specially at the start, are highly percussive and noisy, the Hilbert approach presents a considerable deviation, as can be seen in Fig.
+For a recording of the key 33 of a grand piano, where parts of the sound, specially at the start, are highly percussive and noisy, the Hilbert approach presents a considerable undershooting, as can be seen in Fig.
 \ref{fig:Comparison_piano}.
@@
 \caption{Comparison of the average errors of the algorithm presented in this work with the most common methods of digital envelope identification.
-The discrete waves used were normalized between -1 and 1.
-The average is between the absolute value of those waves and the envelope divided by two.}
+The discrete waves used were normalized between -1 and 1.}
 \label{table:Error_comparison}
@@
 
-The frontiers were are satisfactorily detected in the case of harmonic and inharmonic sounds, and are robust in relation to the number of samples and the frequencies of the waves.
-It is worth noting that the positive and negative frontiers are generally smoother than the unified envelope and seems to conform better to the countours of the underlying digital signal.
+The frontiers were satisfactorily detected in the case of harmonic and inharmonic sounds, and are robust in relation to the number of samples and the frequencies of the waves.
+It is worth noting that the positive and negative frontiers are generally smoother than the unified envelope and seems to conform better to the contours of the underlying digital signal.
 
@@
 
-The temporal envelope adds complexity to the spectral representation of a wave \cite{2019TarjanoNeuro}, and an accurate description of the envelope, while describing the evolution of the instantaneous amplitude of a signal in time, would also simplify further spectral analysis.
+The temporal envelope adds complexity to the spectral representation of a wave \cite{2019TarjanoNeuro}, and an accurate representation of the envelope, while describing the evolution of the instantaneous amplitude of a signal in time, would also simplify further spectral analysis.
 
@@
 Fig.
-\ref{fig:FourierBend} shows part the power spectra for both the original wave and the carrier obtained by the present algorithm, for the guitar bend sound in Fig.
+\ref{fig:FourierBend} shows part of the power spectra for both the original wave and the carrier obtained by the present algorithm, for the guitar bend sound in Fig.
 \ref{fig:Bend}, around its fundamental frequency.
 Two prominent frequencies can be seen, the initial and final notes of the bend.
-In the carrier, the other frequencies appear more smooth.
+In the carrier, the other frequencies appear smoother.
 
@@
 The vertical lines mark the indices of the samples that belong to the positive frontier of the original signal, and the section between each vertical line isolates a pseudo cycle of the wave.
-Note that, although two local maxima exist in each pseudo cycle, only the largest was marked.
+Note that, although two local maxima exist in each pseudo cycle, only the largest was marked by the algorithm.
 
@@
 \ref{fig:AvgPc}: It shows the superposition of the extracted pseudo cycles, after being normalized by length, and their average.
-They were also rotated to make their starting point more close to zero.
+They were also rotated to make their starting point closer to zero.
 

### 2021-05-10 Papers/01_Pulses_Envelope/00ElsevierPulsesEnvelope.tex
refining

@@
 Links to interactive visualizations are available from the repository prepared for this work \cite{2020TarjanoEnvelope}, which also contains an implementation of the method in pure Python.
-A Python module can be installed from the Python Package Index (PyPI) with the command `pip install signal-envelope`.
+A Python module can be installed from the Python Package Index (PyPI) with the command “pip install signal-envelope”.
 Besides, those working under Windows 64-bit can benefit from a specialized dynamic library coded in C++ that is automatically used within this module, and can be separately obtained via the release section in the repository.
@@
 The preceding definitions can be translated to this discrete scenario assuming that the discrete quantities arise from observing the continuous ones at regular time intervals.
-In that case, the equality $ i = t \ \text{fps} $ can be used to link both settings, where $ i $ is the index of each sample of the digital signal, $ t $ stands for the time in seconds, and fps is the frame rate, or the number of observations made in one second.
+In that case, the equality $ i = t \ \text{fps} $ can be used to link both settings, where $ i $ is the index of each sample of the discrete signal, $ t $ stands for the time in seconds, and fps is the frame rate, or the number of observations made in one second.
 We can thus define:
@@
 This realization that the envelope of a wave is defined only at some of the wave's local extrema is important, as we can use it to generate a simplified representation of a signal that will make the algorithm more efficient.
+
+Note that, in Fig.
+\ref{fig:Envelope}, the carrier, the original signal and even the envelope are representations of similar entities: signals in the same coordinate system.
+That is in contrast with some representations found in the literature of amplitude and frequency modulation of narrowband signals \cite[e.g.,][]{1996Loughlinamplitude} where the FM and AM of a signal are represented in different coordinate systems, commonly frequency versus time and amplitude versus time, respectively.
 
@@
 The frontiers were satisfactorily detected in the case of harmonic and inharmonic sounds, and are robust in relation to the number of samples and the frequencies of the waves.
-It is worth noting that the positive and negative frontiers are generally smoother than the unified envelope and seems to conform better to the contours of the underlying digital signal.
+It is worth noting that the positive and negative frontiers are generally smoother than the unified envelope and seems to conform better to the contours of the underlying discrete signal.
 

### 2021-05-10 Papers/01_Pulses_Envelope/images/09PCs.py

@@ -38,8 +38,8 @@ FONT = dict(
 fig = go.Figure()
 fig.layout.template ="plotly_white" 
 fig.update_layout(
-  xaxis_title="<b>Amplitude</b>",
-  yaxis_title="<b>Sample <i>i</i></b>",
+  yaxis_title="<b>Amplitude</b>",
+  xaxis_title="<b>Sample <i>i</i></b>",
   legend=dict(orientation='h', yanchor='top', xanchor='left', y=1.1),
   margin=dict(l=0, r=0, b=0, t=0),
   font=FONT,

### 2021-05-10 Papers/01_Pulses_Envelope/images/10AvgPc.py

@@ -65,7 +65,7 @@ FONT = dict(
 fig = go.Figure()
 fig.layout.template ="plotly_white" 
 fig.update_layout(
-  xaxis_title="<b><i>i</i></b>",
+  xaxis_title="<b>Sample <i>i</i></b>",
   yaxis_title="<b>Amplitude</b>",
   legend=dict(orientation='h', yanchor='top', xanchor='left', y=1.1),
   margin=dict(l=0, r=0, b=0, t=0),

### 2021-05-10 Papers/01_Pulses_Envelope/00ElsevierPulsesEnvelope.tex
revising

@@
 Despite being an elusive concept, the temporal amplitude envelope of a signal is essential for its complete characterization, being the primary information-carrying medium in spoken voice and telecommunications, for example.
-Envelope detection techniques have applications in areas like medicine, sound classification and synthesis, seismology, and speech recognition.
-Nevertheless, a general approach to envelope detection doesn't exist, as most methods involve manual intervention, in the form of filter design, smoothing, and other specific design choices based on prior knowledge of the signals under investigation.
+Envelope detection techniques have applications in medicine, sound classification, sound synthesis, seismology, and speech recognition.
+Nevertheless, a general approach to envelope detection doesn't exist.
+Most methods involve manual intervention, in the form of filter design, smoothing, and other specific design choices based on prior knowledge of the signals under investigation.
 To address this problem, we propose an algorithm that uses characteristics of a signal to estimate its envelope, eliminating the necessity of parameter tuning.
@@
 \section{Introduction}
-Envelope detection, also known as demodulation, is ubiquitous in both analog and digital signal processing (DSP) \cite{2011CaetanoImproved}.
-Nevertheless, the literature in this area is fragmented \cite{2017LyonsDigital}, with most envelope detection techniques being designed to account for very specific settings, in general involving narrowband signals with moderate noise content.
+Envelope detection, also known as demodulation, is ubiquitous in analog and digital signal processing (DSP) \cite{2011CaetanoImproved}.
+Nevertheless, the literature in this area is fragmented \cite{2017LyonsDigital}, with most envelope detection techniques being designed to account for particular settings, in general involving narrowband signals with moderate noise content.
 
-Those constraints exclude most physical signals, as is the case of recorded sound, for example; this problem arises in part due to the lack of a strict mathematical definition of a temporal envelope \cite{2013Mengempirical}.
+Those constraints exclude most physical signals, as is the case of recorded sound, for example; this problem arises in part due to the lack of a strict mathematical definition of a temporal envelope \cite{2014Yangnovel,2015Yangtheoretical,2019Jiaempirical}.
 
@@
 An analogous construction is possible for every point at the frontier.
-For all the interior points, it is impossible to construct a circle of radius 10 that contains one or more of them without defining a disc containing at least one point of the set.
+
+Conversely, any circle of radius $ \alpha=10 $ that contains at least one of the interior defines an open disc that also contains at least another point of the set P.
 
@@
 
-Following this path, we present a general approach to envelope detection, exploiting the intrinsic characteristics of a spectrally complex wave, in order to completely avoid the need for manual intervention or parameter tuning.
+Following this path, we present a general approach to envelope detection, exploiting geometric characteristics of a spectrally complex wave to completely avoid the need for parameter tuning.
 
-We start with a brief background of envelope detection in Section \ref{sec:envelope} and define, as specifically as possible, what constitutes a temporal envelope.
+We start with a brief background of envelope detection in Section \ref{sec:envelope} and define, as precisely as possible, what constitutes a temporal envelope.
 
-In Section \ref{sec:representation} we explore some simplifications to the task of detecting the envelope of a signal that arise from the envelope definition adopted in this work.
+In Section \ref{sec:representation} we explore some simplifications to the task of envelope detection that arise from the envelope definition adopted in this work.
 
@@
 
-Section \ref{sec:quality} discusses the difficulties in the assessment of the quality of an envelope, proposing qualitative and quantitative approaches, and demonstrates that the method presented conforms to four requirements for a physically plausible envelope proposed in \cite{1996Loughlinamplitude}.
+Section \ref{sec:quality} discusses the difficulties in assessing the quality of an envelope, proposing qualitative and quantitative approaches, and demonstrates that the method presented conforms to four requirements for a physically plausible envelope proposed in \cite{1996Loughlinamplitude}.
 
@@
 
-In this context, the most mathematically sound definition of an envelope involves the representation of its underlying wave as an analytic signal, as introduced by Gabor in 1946 \cite{2007HahnHistory}.
-In his work, \cite{1946GaborTheory} applies the then relatively new mathematical machinery of the quantum mechanics to unify the time and frequency-domain representations of a wave, showing how the Hilbert transform could be applied to a real signal in order to obtain an equivalent complex signal, later known as the analytic signal.
+In this context, the most mathematically sound definition of an envelope, as introduced by Gabor in 1946 \cite{2007HahnHistory}, involves representing the underlying wave as an analytic signal.
+In his work, \cite{1946GaborTheory} applies the then relatively new mathematical machinery of quantum mechanics to unify the time and frequency-domain representations of a wave, showing how the Hilbert transform could be applied to a real signal in order to obtain an equivalent complex signal, later known as the analytic signal.
 
@@
 $ \mathbb{H}(S(t)) $, the Hilbert transform of the original signal, becomes the imaginary part of the analytic signal; the envelope of a signal thus represented can be straightforwardly obtained by the computation of the complex modulus of $ \mathbb{H}(S(t)) $.
+This representation is common in classical AM-FM literature.
 
@@
 
-For a real world signal, such as an alto singer sustaining a steady note as shown in Fig.
+For real-world signals, such as an alto singer sustaining a steady note as shown in Fig.
 \ref{fig:AnalyticSignalAlto}, the Hilbert transform alone provides a poor envelope, since it retains most of the frequencies of the underlying wave.
@@
     \includegraphics{images/01AnalyticSignalAlto.pdf}
-  \caption{A section of the representation of an alto singer sustaining a steady note and a comparison of the envelopes obtained by the Hilbert transform with and without further filtering.}
+  \caption{A section of the representation of an alto singer sustaining a steady note comparing the envelopes obtained by the Hilbert transform with and without further filtering.}
   \label{fig:AnalyticSignalAlto}
@@
 
-In the case of broadband signals, the problem of envelope detection of an arbitrary wave is ill-posed, in the sense that an infinite number of pairs of $ E(t) $ and $ C(t) $ can result in the same $ W(t) $ \cite{2011TurnerDemodulation, 2013Mengempirical, 1996Loughlinamplitude}.
+In the case of broadband signals, the problem of envelope detection of an arbitrary wave is ill-posed, in the sense that an infinite number of pairs of $ E(t) $ and $ C(t) $ can result in the same $ W(t) $ \cite{2011TurnerDemodulation, 1996Loughlinamplitude}.
 
@@
   \mathbf{e}, \mathbf{c} & \in \mathbb{R}^n \\
-  \mathbf{e} & = (e_0, e_1, \cdots, e_{n-1}), e_i \ge 0 \ \forall \ i, \ 0 \le i \le n-1 \\  
-  \mathbf{c} & = (c_0, c_1, \cdots, c_{n-1}), |c_i| \le 1 \ \forall \ i, \ 0 \le i \le n-1, |\max(\mathbf{c})| =1 \\
+  \mathbf{e} & \triangleq (e_0, e_1, \cdots, e_{n-1}), e_i \ge 0 \ \forall \ i, \ 0 \le i \le n-1 \\  
+  \mathbf{c} & \triangleq (c_0, c_1, \cdots, c_{n-1}), |c_i| \le 1 \ \forall \ i, \ 0 \le i \le n-1, |\max(\mathbf{c})| =1 \\
+  \text{Such as:}& \\
   \mathbf{w} & = \mathbf{e} \odot \mathbf{c} \ \therefore \ w_i = e_i c_i \ \forall \ i, \ 0 \le i \le n-1
@@
 
-This realization that the envelope of a wave is defined only at some of the wave's local extrema is important, as we can use it to generate a simplified representation of a signal that will make the algorithm more efficient.
+This realization that the envelope of a wave is defined only at some of the wave's local extrema is essential, as it can be used to generate a simplified representation of a signal, making the algorithm more efficient.
 
@@
 \ref{fig:Envelope}, the carrier, the original signal and even the envelope are representations of similar entities: signals in the same coordinate system.
-That is in contrast with some representations found in the literature of amplitude and frequency modulation of narrowband signals \cite[e.g.,][]{1996Loughlinamplitude} where the FM and AM of a signal are represented in different coordinate systems, commonly frequency versus time and amplitude versus time, respectively.
+This is in line with \cite{2011CaetanoImproved,2011TurnerDemodulation,2017LyonsDigital} but in contrast with some representations found in the literature of amplitude and frequency modulation of narrowband signals \cite[e.g.,][]{1996Loughlinamplitude} where the FM and AM of a signal are represented in different coordinate systems, commonly frequency versus time and amplitude versus time, respectively.
 
@@
 \ref{fig:Envelope} and the discussion in Section \ref{sec:envelope}, we can see that only the maximum point of a positive pulse or the minimum point of a negative pulse can potentially be equal to the envelope and are, thus, our only points of interest.
-We can then proceed, for the rest of the method, considering only those points, to a great computational economy.
+We can then proceed, for the rest of the method, considering only those points, to a tremendous computational economy.
 
-For this reason we define P as the set of the points $ \{P_0, P_1, \cdots, P_{m-1}\} $ where $ P_j = (i_j, \lvert w_{i_j} \lvert) $, that is, $ \lvert w_{i_j} \lvert $, the absolute value of the extrema of each pulse, becomes the ordinate of each point and $ i_j $, its original index, becomes the abscissa.
+For this reason, we define P as the set of the points $ \{P_0, P_1, \cdots, P_{m-1}\} $ where $ P_j = (i_j, \lvert w_{i_j} \lvert) $, that is, $ \lvert w_{i_j} \lvert $, the absolute value of the extrema of each pulse, becomes the ordinate of each point and $ i_j $, its original index, becomes the abscissa.
 
@@
 To achieve that, we divide the ordinate of each point by the average of all ordinates, effectively canceling the unit of the vertical components of the points.
-If we proceeded similarly in relation to the horizontal components,
-the direct link between P and \textbf{w} that arises from the fact that the abscissas of the points in P are numerically equivalent to the indices $ i $ of \textbf{w} would be lost.
+If we proceeded similarly concerning the horizontal components,
+the direct link between P and \textbf{w}, that arises from the fact that the abscissas of the points in P are numerically equivalent to the indices $ i $ of \textbf{w}, would be lost.
 
@@
 
-Having established a transformation from the original space of the two-dimensional representation of a discrete wave to the Cartesian coordinate system, we are now in a position to apply the geometric concept of the shape of a set of points.
+Having established a transformation from the original space of the two-dimensional representation of a discrete wave to the Cartesian coordinate system, we are now able to apply the geometric concept of the shape of a set of points.
 
@@
 
-Concerning direct methods \cite{2014CarrollSurvey}, for example, derive three such definitions based on the approximation of a circle by an inscribed, centered, and circumscribed polygon.
+Concerning direct methods \cite{2014CarrollSurvey}, derive three such definitions based on the approximation of a circle by an inscribed, centered, and circumscribed polygon.
 In the context of three-dimensional meshes, \cite{2016VasaMultivariate} evaluate a range of existing estimators from a multivariate point of view.
@@
 $ v_k $ represents the net displacement of the envelope from the extrema $ P_{j} = (x_j, y_j) $ to $ P_{j+1} = (x_{j+1}, y_{j+1}) $.
-Since we don't yet know which points of P are part of the envelope, instead of assuming that the trajectory before $ P_{j} $ passes through $ P_{j-1} $, it is more prudent to assume that it can have passed anywhere in the line segment from points $ (x_j - (x_{j+1} - x_j), y_j - (y_{j+1} - y_j)) $ to $ (x_j - (x_{j+1} - x_j), y_j + (y_{j+1} - y_j)) $, the bold line segment before $ v_k $ in Fig.
+
+We assume from $ v_k $ that, near the vector, the envelope defines a vertical displacement of $ v_{k,y} $ in the horizontal distance of $ v_{k,x} $; since we don't yet know which points of P are part of the envelope, we can't be certain that the envelope passes through $ P_{j-1} $ before reaching $ P_{j} $, or even if it reaches $ P_{j} $.
+
+Nevertheless, assuming it passes through $ P_{j} $, and is limited to the diplacement observed, the envelope might have passed anywhere in the line segment from points $ (x_j - (x_{j+1} - x_j), y_j - (y_{j+1} - y_j)) $ to $ (x_j - (x_{j+1} - x_j), y_j + (y_{j+1} - y_j)) $, the bold line segment before $ v_k $ in Fig.
 \ref{fig:DiscreteCurvature}.
 
-That is because all we can safely assume from $ v_k $ is that, in the immediate vicinity of the vector, the envelope is capable of a vertical displacement of $ v_{k,y} $ in the same time as it performs a horizontal displacement of $ v_{k,x} $;
+Generalizing those observations for all vectors $ v_k \in \text{V} $, we can consider that, on average, the envelope insides horizontally at $ P_{j} $ and smoothly changes from the horizontal direction to the direction of the vector $ v_k $.
 
-Since this is true for all the vectors $ v_k \in \text{V} $, we can consider that, on average, the envelope insides horizontally at $ P_{j} $.
-
-The rationale then is to find, for each vector $ v_k $, the radius $ r_k $ of the equivalent circle whose tangent has the same change in direction, in the same horizontal distance, as the vector and the horizontal direction, as shown in Fig.
+The rationale then is to find, for each vector $ v_k $, the radius $ r_k $ of the equivalent circle whose unit tangent performs the same change in direction, in the same horizontal distance, as illustrated in Fig.
 \ref{fig:DiscreteCurvature}.
-The average curvature of P will then be obtained by the average of all radii $ r_k $.
+The average curvature of P will be obtained by the average of all radii $ r_k $.
 
@@
 
-We need now to construct an algorithm, using the radius just obtained, to identify the points that belong to the envelope.
+We need to construct an algorithm, using the radius just obtained, to identify the points that belong to the envelope.
 Generally, alpha shapes based algorithms resort to the construction of the Delaunay triangulation of the set of points, which is later filtered to contain only the outer edges.
-We can adopt a more straightforward method, as our points have a sequential structure.
+We can adopt a more straightforward approach, since our points have a sequential structure.
 
@@
 
-The procedure is formalized in the algorithm \ref{FindEnvelope}.
+The procedure is formalized in algorithm \ref{FindEnvelope}.
 We don't provide similar algorithmic descriptions of the preceding steps since the Python implementation available at \cite{2020TarjanoEnvelope} can easily serve as pseudocode.
@@
 
-The first one states that, if a signal is bounded in magnitude, then its envelope should be, as well.
+The first one states that, if a signal is bounded in magnitude, its envelope should also be bounded.
 Our algorithm complies with this requirement, as the envelope is composed of selected samples from the original signal itself.
@@
 
-In order to obtain a numeric indicator of the quality of an envelope, we construct a reference envelope, that trades off applicability in favour of mathematical formality.
+In order to obtain a numeric indicator of the quality of an envelope, we construct a reference envelope, that trades off applicability in favor of mathematical formality.
 In \cite{1992BruceCurves} the envelope is defined as the boundary of the region filled by a family of curves.
@@
 In practical applications, one seldom has access to a family of curves, instead being interested in obtaining the envelope of a single signal.
-Nevertheless, an approximation can be obtained, by shifting the original signal along its initial position in the horizontal axis by half of its period.
+Nevertheless, an approximation can be obtained, by shifting the original signal along its initial position in the horizontal axis by half of its period, and taking the absolute maximum value at each index $i$.
 Fig. \ref{fig:RefEnvelope} shows the reference envelope for the sound of a tom.
@@
 The envelope obtained is not smooth, being composed of a series of horizontal segments, as can be seen in \ref{fig:RefEnvelope}, besides being computationally expensive to calculate.
-However, arising from a mathematically sound basis, it can serve as a reference to infer the quality of more empirical methods.
+However, arising from a mathematically sound background, it can serve as a reference to infer the quality of empirical methods.
 
@@
 In Fig. \ref{fig:Comparison_sinusoid} we illustrate the envelope extracted by the conventional algorithms, contrasting it with the result of the method presented in this work, for a simple sinusoid. 
-Only the present algorithm and the one based on the Hilbert transform comply with the third condition proposed by \cite{1996Loughlinamplitude}, that states that the envelope of a periodic wave should be a straight line.
+Only the present algorithm and the one based on the Hilbert transform comply with the third condition proposed by \cite{1996Loughlinamplitude}, which states that the envelope of a periodic wave should be a straight line.
 
@@
 In the case of the Hilbert transform, post-processing in the form of filtering was applied, and an appropriate cut-off frequency had to be chosen.
-That is in contrast with our method, that organically tunes itself via the automatic choice of the radius of the circle representing the average discrete curvature of the signal.
+That is in contrast with our method, which organically tunes itself via the automatic choice of the circle's radius representing the average discrete curvature of the signal.
 
@@
 
-The frontiers were satisfactorily detected in the case of harmonic and inharmonic sounds, and are robust in relation to the number of samples and the frequencies of the waves.
+The frontiers were satisfactorily detected in the case of harmonic and inharmonic sounds, and are robust about the number of samples and the frequencies of the waves.
 It is worth noting that the positive and negative frontiers are generally smoother than the unified envelope and seems to conform better to the contours of the underlying discrete signal.
@@
 
-The temporal envelope adds complexity to the spectral representation of a wave \cite{2019TarjanoNeuro}, and an accurate representation of the envelope, while describing the evolution of the instantaneous amplitude of a signal in time, would also simplify further spectral analysis.
+The temporal envelope adds complexity to the spectral representation of a wave \cite{2019TarjanoNeuro}.
+An accurate representation of the envelope, while describing the instantaneous amplitude of a signal, would also simplify further spectral analysis.
 
 Fig.
-\ref{fig:Fourier} shows the frequency-domain power spectrum for the wave and the carrier presented in \ref{fig:Envelope}.
-For the carrier, the power spectrum presents two nonzero values, while the power spectrum for the wave composed of the carrier modulated by the envelope is more complex.
+\ref{fig:Fourier} shows the frequency-domain power spectrum for the wave and the carrier presented in Fig.
+\ref{fig:Envelope}.
+For the carrier, the power spectrum presents two nonzero values, while the power spectrum for the wave, composed of the carrier modulated by the envelope, is more complex.
 
@@
 
-The algorithm developed in this work naturally divides a signal into its pseudo-cycles, pinpointing them in the time-domain, as illustrated in Fig.
+The algorithm developed in this work naturally divides a signal into its pseudo cycles, pinpointing them in the time-domain, as illustrated in Fig.
 \ref{fig:singer_voice}, for the sound of an alto singer singing a sustained note, as in Fig.
@@
 The vertical lines mark the indices of the samples that belong to the positive frontier of the original signal, and the section between each vertical line isolates a pseudo cycle of the wave.
-Note that, although two local maxima exist in each pseudo cycle, only the largest was marked by the algorithm.
+Although two local maxima exist in each pseudo cycle, only the largest was marked by the algorithm.
 
@@
 
-The present work also highlights some gaps in our current understanding and definition of temporal envelope and, in so doing, endeavours to motivate further research in this topic, that can ultimately give rise to a more mathematically rigorous envelope theory, especially concerning broadband waves; suggestions in this direction were also presented.
+The present work also highlights some gaps in our current understanding and definition of temporal envelope and, in so doing, endeavors to motivate further research in this topic, that can ultimately give rise to a more mathematically rigorous envelope theory, especially concerning broadband waves; suggestions in this direction were also presented.
 
@@
 
-Similarly, by decomposing a wave into its envelope and carrier, and simplifying the representation of such carrier in the frequency-domain, a rich set of advancements, especially in the areas of sound compression and synthesis in the time and frequency-domains, can be built upon this initial algorithm.
+Similarly, by decomposing a wave into its envelope and carrier, and simplifying the representation of such carrier in the frequency-domain, a rich set of advancements, especially in sound compression and synthesis, both in the time and frequency-domains, can be built upon this initial algorithm.
 

### 2021-05-10 Papers/01_Pulses_Envelope/images/10RefEnvelope.py

@@ -67,7 +67,7 @@ fig.update_yaxes(showline=False, zeroline=False, showgrid=True)
 
 fig.add_trace(
   go.Scatter(
-    name="Signal       ",
+    name="Signal",
     # x=np.arange(rolled_W.size).tolist() + [None],
     y=W,
     mode="lines",
@@ -81,12 +81,12 @@ fig.add_trace(
 
 fig.add_trace(
   go.Scatter(
-    name="Signal, shifted from -T/2 to T/2       ",
+    name="$ \\text{Family of signals, shifted from } i-T/2 \\text{ to } i + T/2 $",
     x=Xf[0:-1:20].flatten()[t : ],#pulses_X,
     y=Yf[0:-1:20].flatten()[t : ],#pulses_Y,
     mode="lines",
     line=dict(width=1, color="rgba(120, 120, 120, .5)"
-    , shape = 'spline'
+    # , shape = 'spline'
     ),
     marker=dict(size=4, color="gray"),
     # showlegend=False
@@ -96,19 +96,19 @@ fig.add_trace(
 
 fig.add_trace(
   go.Scatter(
-    name="Reference envelope       ",
+    name="Reference envelope",
     x=(np.arange(rolled_W.size) - t//2)[t//2 : ],
     y=(np.max(np.array(Yf)[:, 0:-1], 0)[t//2 : ]),
     mode="lines",
     line=dict(width=2, color="red"
-    , shape = 'spline'
+    # , shape = 'spline'
     ),
     marker=dict(size=3, color="red")
     # visible = "legendonly"
   )
 )
 
-fig.show(config=dict({'scrollZoom': True}))
+# fig.show(config=dict({'scrollZoom': True}))
 save_name = "./images/" + sys.argv[0].split('/')[-1].replace(".py", ".svg")
 fig.write_image(save_name, width=650, height=350, engine="kaleido", format="svg")
 print("saved:", save_name)

### 2021-05-10 Papers/01_Pulses_Envelope/00ElsevierPulsesEnvelope.tex
revising

@@
 
-The $ \odot $ operator in \ref{eq:Envelope} stands for the Hadamard product, denoting element-wise multiplication of two vectors.
+The $ \odot $ operator in Equation \ref{eq:Envelope} stands for the Hadamard product, denoting element-wise multiplication of two vectors.
 Fig.
@@
 
-Considering $P_{\mathbf{b}}$ as a point of the set P represented on the original basis and $P_{\mathbf{s}}$ the same point as represented on the standard basis, we can formalize this transformation as in \ref{eq:Basis}:
+Considering $P_{\mathbf{b}}$ as a point of the set P represented on the original basis and $P_{\mathbf{s}}$ the same point as represented on the standard basis, we can formalize this transformation as in Equation \ref{eq:Basis}:
 
@@
 
-We formalize the set V of vectors in Eq. \ref{eq:Vectors} below.
+We formalize the set V of vectors in Equation \ref{eq:Vectors} below.
 
@@
 
-The radius $ r_k $ of the circle equivalent to each $ v_k $ is defined in Eq.
-\ref{eq:Radius}.
+The radius $ r_k $ of the circle equivalent to each $ v_k $ is defined in Equation \ref{eq:Radius}.
 This equation has the desirable property that $ v_{k,y} \to 0 \implies v_{k,y} \to \infty $, that is, the radius tends to infinity when no vertical oscillation occurs, effectively turning the circle into a line.
@@
 
-For calculation purposes, however, the formulation in Eq.
-\ref{eq:Radius} poses a problem, as a zero could make its way into the denominator of a fraction; in the actual algorithm, the quantity calculated for each vector is the curvature itself, that is, the inverse of each radius $ r_k $.
+For calculation purposes, however, the formulation in Equation \ref{eq:Radius} poses a problem, as a zero could make its way into the denominator of a fraction; in the actual algorithm, the quantity calculated for each vector is the curvature itself, that is, the inverse of each radius $ r_k $.
 In this way, the problem is avoided, since $ v_{k,x} $ is guaranteed to be greater than zero.
-The average of curvatures is then inverted in order to obtain the average radius $ \overline{r} $, as illustrated in Eq.
-\ref{eq:Curvature}.
+The average of curvatures is then inverted in order to obtain the average radius $ \overline{r} $, as illustrated in Equation \ref{eq:Curvature}.
 
@@
 
-The algorithm follows directly from the definition in Eq.
-\ref{eq:InEnvelope} after noting that the first and last points of P will always be part of the envelope.
+The algorithm follows directly from the definition in Equation \ref{eq:InEnvelope} after noting that the first and last points of P will always be part of the envelope.
 Since $ P_a $ can be equal to $ P_b $, it is always possible to construct a circle containing those points, by placing the center of the circle to the left of or right of the first or last point, respectively.
@@
 
-Recall from section \ref{sec:mapping} that the vertical coordinates of these points are currently scaled, in relation to the original signal, by the transformation described in Eq.
-\ref{eq:Basis}.
+Recall from Section \ref{sec:mapping} that the vertical coordinates of these points are currently scaled, in relation to the original signal, by the transformation described in Equation \ref{eq:Basis}.
 We could invert this transformation, but since we were careful to maintain the direct link between the horizontal coordinates in both coordinate systems, we can simply consider the horizontal coordinates of P as the indices of the samples of the original discrete signal, and retrieve the envelope at each point directly.
@@
 More formally, we have the set P of the absolute values of the local extrema of the original signal $ \mathbf{w} = (w_0, w_1, \cdots, w_{n-1}) $ that contains, by definition, the global extrema of $ \mathbf{w} $.
-The set E of the points that belong to the envelope is a subset of P. We can thus write Eq. \ref{eq:Req1}:
+The set E of the points that belong to the envelope is a subset of P. We can thus write Equation \ref{eq:Req1}:
 
@@
 
-Recalling Eq.
-\ref{eq:Envelope}, where the relation between the original signal, the carrier and the envelope was defined as $ \mathbf{w} = \mathbf{e} \odot \mathbf{c} $, we can write the magnitude of an arbitrary frequency of the original signal ($ m_\mathbf{w}(f) $) and, likewise, the magnitude of an arbitrary frequency of the carrier ($ m_\mathbf{c}(f) $), in terms of the Discrete Fourier Transform (DFT) as in Eq.
-\ref{eq:DFT}:
+Recalling Equation \ref{eq:Envelope}, where the relation between the original signal, the carrier and the envelope was defined as $ \mathbf{w} = \mathbf{e} \odot \mathbf{c} $, we can write the magnitude of an arbitrary frequency of the original signal ($ m_\mathbf{w}(f) $) and, likewise, the magnitude of an arbitrary frequency of the carrier ($ m_\mathbf{c}(f) $), in terms of the Discrete Fourier Transform (DFT) as in Equation \ref{eq:DFT}:
 
@@
 
-Recall from section \ref{sec:mapping} that, before the extraction of the envelope, the original points undergo a transformation to lie in the Cartesian coordinate system.
+Recall from Section \ref{sec:mapping} that, before the extraction of the envelope, the original points undergo a transformation to lie in the Cartesian coordinate system.
 After the extraction, the inverse operation is performed on the points of the Envelope.
 
-Consider a discrete signal $ \mathbf{w} = (w_0, w_1, \cdots, w_{n-1}) $ and the set P of its absolute extrema, represented in the standard basis, $ \text{P} = \{ P_0, P_1, \cdots, P_{m-1} \}, \quad P_j = (i_j, \beta \lvert w_{i_j} \lvert) $, where $ \beta $ is obtained from Eq.
-\ref{eq:Basis}.
+Consider a discrete signal $ \mathbf{w} = (w_0, w_1, \cdots, w_{n-1}) $ and the set P of its absolute extrema, represented in the standard basis, $ \text{P} = \{ P_0, P_1, \cdots, P_{m-1} \}, \quad P_j = (i_j, \beta \lvert w_{i_j} \lvert) $, where $ \beta $ is obtained from Equation \ref{eq:Basis}.
 
 Given a constant $ \alpha > 0, \alpha \in \mathbb{R} $, we can define the scaled signal as $ \mathbf{w}^\prime = (\alpha w_0, \alpha w_1, \cdots, \alpha w_{n-1}) $.
-Observing, from Eq.
-\ref{eq:Basis}, that $ \alpha $ can be factored out of denominator of the summation, we have $ \beta^\prime = \beta / \alpha $ and the new set $ \text{P}^\prime = \{ P^\prime_0, P^\prime_1, \cdots, P^\prime_{m-1} \}, \quad P^\prime_j = (i_j, \frac{\beta}{\alpha} \lvert \alpha w_{i_j} \lvert) $ is equal to the set P for the signal before scaling.
+Observing, from Equation \ref{eq:Basis}, that $ \alpha $ can be factored out of denominator of the summation, we have $ \beta^\prime = \beta / \alpha $ and the new set $ \text{P}^\prime = \{ P^\prime_0, P^\prime_1, \cdots, P^\prime_{m-1} \}, \quad P^\prime_j = (i_j, \frac{\beta}{\alpha} \lvert \alpha w_{i_j} \lvert) $ is equal to the set P for the signal before scaling.
 
@@
 
-The envelope obtained is not smooth, being composed of a series of horizontal segments, as can be seen in \ref{fig:RefEnvelope}, besides being computationally expensive to calculate.
+The envelope obtained is not smooth, being composed of a series of horizontal segments, as can be seen in Fig.
+\ref{fig:RefEnvelope}, besides being computationally expensive to calculate.
 However, arising from a mathematically sound background, it can serve as a reference to infer the quality of empirical methods.
@@
 
-Table \ref{table:Error_comparison} presents the average of the absolute errors of the techniques, when compared with the reference envelope introduced in the end of section \ref{sec:quality}
+Table \ref{table:Error_comparison} presents the average of the absolute errors of the techniques, when compared with the reference envelope introduced in the end of Section \ref{sec:quality}
 
@@
 
-We can see from the table \ref{table:Error_comparison} that the method presented here consistently outperforms the traditional approaches for all waves tested, yielding an error of less than 2\% on average, of the error of the traditional methods.
+We can see from Table \ref{table:Error_comparison} that the method presented here consistently outperforms the traditional approaches for all waves tested, yielding an error of less than 2\% on average, of the error of the traditional methods.
 
-The times taken for the algorithms to process each signal are shown in \ref{table:Time_comparison}. 
+The times taken for the algorithms to process each signal are shown in Table \ref{table:Time_comparison}. 
 The Python source for the tests, as well as the samples used, are available at the repository dedicated to this work, in the file “Comparison of Envelope Algorithms.py”.
@@
 
-If we split P, as defined in section \ref{sec:representation}, into $\text{P}^+$, the set of the local maxima of the original signal and $\text{P}^-$, the set of its local minima, we can apply the remaining steps of the algorithm in those two sets independently, arriving at a positive and a negative frontier.
+If we split P, as defined in Section \ref{sec:representation}, into $\text{P}^+$, the set of the local maxima of the original signal and $\text{P}^-$, the set of its local minima, we can apply the remaining steps of the algorithm in those two sets independently, arriving at a positive and a negative frontier.
 
@@
     \includegraphics{images/11Fourier.pdf}
-  \caption{Fourier power spectrum for the wave and carrier shown in \ref{fig:Envelope}.}
+  \caption{Fourier power spectrum for the wave and carrier shown in Fig. \ref{fig:Envelope}.}
   \label{fig:Fourier}
@@
 \ref{fig:Bend}, around its fundamental frequency.
-Two prominent frequencies can be seen, the initial and final notes of the bend.
+Two prominent frequencies can be seen, corresponding to the initial and final notes of the bend.
 In the carrier, the other frequencies appear smoother.
@@
     \includegraphics{images/09PCs.pdf}
-  \caption{Section of the digital wave illustrated in \ref{fig:singer_voice}, with vertical lines at the indices of the samples that belong to the positive frontier.}
+  \caption{Section of the digital wave illustrated in Fig. \ref{fig:singer_voice}, with vertical lines at the indices of the samples that belong to the positive frontier.}
   \label{fig:PCs}
@@
     \includegraphics{images/10AvgPc.pdf}
-  \caption{Pseudo-cycles of the wave illustrated in \ref{fig:singer_voice}, segmented and superposed, and their average.
+  \caption{Pseudo-cycles of the wave illustrated in Fig.
+\ref{fig:singer_voice}, segmented and superposed, and their average.
 The average is also shown shifted by one average period, to illustrate the average continuity.}

### 2021-05-11 Papers/01_Pulses_Envelope/00ElsevierPulsesEnvelope.tex
revising

@@
 To address this problem, we propose an algorithm that uses characteristics of a signal to estimate its envelope, eliminating the necessity of parameter tuning.
-The method draws inspiration from geometric concepts; the discrete signal is mapped to a set of points in the Cartesian plane, where a new measure of discrete curvature is used to identify the samples that are part of the envelope.
+The method draws inspiration from geometric concepts: the discrete signal is mapped to a set of points in the Cartesian plane, where a new measure of discrete curvature is used to identify the samples that are part of the envelope.
 Due to its geometric nature, the method can optionally generate the superior and inferior envelopes of a given wave.
 The algorithm compares favorably with classic envelope detection techniques based on smoothing, filtering, and the Hilbert Transform, besides being physically plausible.
-We provide visualizations of the envelope extracted for various real-world signals with very diverse characteristics, such as voice, spoken and sang, and pitched and unpitched musical instruments, and discuss some approaches to objectively assess the quality of the obtained envelopes.
-An optimized implementation is available via the Python Package Index (PyPI); interactive visualizations and source code are available at the accompanying repository.
+We provide visualizations of the envelope extracted for various real-world signals with diverse characteristics, such as voice, spoken and sang, and pitched and unpitched musical instruments, and discuss some approaches to objectively assess the quality of the obtained envelopes.
+An optimized implementation is available via the Python Package Index (PyPI), while interactive visualizations and source code are available from the accompanying GitHub repository.
 \end{abstract}
@@
 Envelope detection, also known as demodulation, is ubiquitous in analog and digital signal processing (DSP) \cite{2011CaetanoImproved}.
-Nevertheless, the literature in this area is fragmented \cite{2017LyonsDigital}, with most envelope detection techniques being designed to account for particular settings, in general involving narrowband signals with moderate noise content.
+Nevertheless, the literature in this area is fragmented \cite{2017LyonsDigital}, with most envelope detection techniques being designed to account for particular settings involving narrowband signals with moderate noise content.
 
-Those constraints exclude most physical signals, as is the case of recorded sound, for example; this problem arises in part due to the lack of a strict mathematical definition of a temporal envelope \cite{2014Yangnovel,2015Yangtheoretical,2019Jiaempirical}.
+Those constraints exclude most physical signals, such recorded sounds, for example; this problem arises in part due to the lack of a strict mathematical definition of a temporal envelope \cite{2014Yangnovel,2015Yangtheoretical,2019Jiaempirical}.
 
@@
 
-The temporal envelope also helps to impart emotion and identity to the human voice \cite{2018ZhuContributions} and music: Concert halls with envelope preserving characteristics are deemed more pleasant \cite{2011LokkiEngaging}.
+The temporal envelope also helps to impart emotion and identity to the human voice \cite{2018ZhuContributions} and music in general: Concert halls with envelope preserving characteristics are deemed more pleasant \cite{2011LokkiEngaging}.
 
-Specially when dealing with broadband signals, approaches tailored to specific applications, exploiting prior knowledge of the waves, are prevalent, such as the one presented by \cite{2014YangFast} for the distributed monitoring of fiber optic or the one formulated by \cite{2018AssefModeling} in the context of medical ultrasound imaging.
+Specially when dealing with broadband signals, approaches tailored to specific applications, exploiting prior knowledge of the waves, are prevalent, such as the one presented by \cite{2014YangFast} for the distributed monitoring of fiber optic, or the one formulated by \cite{2018AssefModeling} in the context of medical ultrasound imaging.
 
-The problem of envelope detection of a discrete signal, normally associated with the field of digital signal processing, can be made equivalent to the geometric problem of defining the shape of a set of points in $ \mathbb{R}^2 $.
+The problem of envelope detection of a discrete signal, traditionally associated with the field of digital signal processing, can be made equivalent to the geometric problem of defining the shape of a set of points in $ \mathbb{R}^2 $.
 
-Considering the standard graphic representation of a discrete wave, where the horizontal axis represents time and the vertical axis represents the amplitude, one can think of each sample as a point in a plane.
-Provided a coherent mapping from this space to the Cartesian plane is available, the original DSP problem can be transformed into the geometric problem of finding the shape of this set of points.
+Considering the standard graphic representation of a discrete wave, where the horizontal axis represents time and the vertical axis represents amplitude, one can think of each sample as a point in a plane.
+Provided a coherent mapping from this space to the Cartesian coordinate system is available, the original DSP problem can be transformed into the geometric problem of finding the shape of this set of points.
 
@@
 The convex hull of a set of points $ S $ in two-dimensional space is the smallest convex subset of the plane that contains all the points in $ S $.
+
+To get an intuition, one can think of the points in a set as nails in a wooden board, around which a rubber band is stretched. The polygon defined by the rubber band is the convex hull of this set.
 Convex hulls are unique, in the sense that a given set of points define one and only one convex hull.
@@
 \ref{fig:AlphaShape} illustrates an alpha shape, for $ \alpha=10 $, of a set of points.
-Note that two points of the set lie at the circle illustrated, but the open disc defined by it contains no points of the set.
+Note that two points of the set lie on the circle illustrated, but the open disc it defines contains no points of the set.
 An analogous construction is possible for every point at the frontier.
 
-Conversely, any circle of radius $ \alpha=10 $ that contains at least one of the interior defines an open disc that also contains at least another point of the set P.
+Conversely, any circle of radius $ \alpha=10 $ containing at least one of the interior points defines an open disc that also contains at least another point of the set P.
 
@@
 Once a radius $ \alpha $ is defined, alpha shapes exhibit the uniqueness property: for a given set of points and a given radius, only one alpha shape exists.
-When the radius $ \alpha $ of the circle tends to infinity, alpha shapes, also known as concave hulls, tend to the convex hulls.
+When the radius $ \alpha $ of the circle tends to infinity, the alpha shape of the set tends to its convex hull.
 
 Consider, for example, the artificially generated discrete wave in Fig.
-\ref{fig:Hulls} below, where the links between consecutive samples were de-emphasized to highlight the “set of points” nature of a signal: two alpha shapes, with different values for $ \alpha $, and the convex hull are illustrated.
-As $ \alpha $ increases, the contour of the set of points becomes smoother, and closer to the convex hull.
+\ref{fig:Hulls}, where the links between consecutive samples were de-emphasized to highlight the “set of points” nature of a signal: two alpha shapes, with different values of $ \alpha $, and the convex hull are illustrated.
+As $ \alpha $ increases, the contour of the set of points becomes coarser, and closer to the convex hull.
 
@@
 
-The alpha shapes theory is used in areas such as detection of features in images \cite{2016VarytimidisAlpha}, reconstruction of surfaces from a cloud of points \cite{2015WuAutomated} and spectroscopy \cite{2019XuModeling}.
+Alpha shapes are used in areas such as detection of features in images \cite{2016VarytimidisAlpha}, reconstruction of surfaces from clouds of points \cite{2015WuAutomated} and spectroscopy \cite{2019XuModeling}.
 
@@
 
-Other steps in the direction of translating geometric algorithms to the context of envelope detection were made by \cite{2015YangSkeleton} via an algorithm based on the construction of a skeleton underlying the digital wave of interest, and also via the direct translation of computer vision methods to the task of envelope detection \cite{2015YangRepresenting}.
+Other steps in the direction of applying geometric algorithms to envelope detection were made by \cite{2015YangSkeleton}, with an algorithm based on the construction of a skeleton underlying the digital wave of interest, and by \cite{2015YangRepresenting}, via the direct translation of computer vision methods to the task of envelope detection.
 
@@
 
-We start with a brief background of envelope detection in Section \ref{sec:envelope} and define, as precisely as possible, what constitutes a temporal envelope.
+We start with a brief background of envelope detection in Section \ref{sec:envelope} where we define, as precisely as possible, what constitutes a temporal envelope.
 
@@
 
-We next explore, in Section \ref{sec:curvature}, the concept of discrete curvature, that will be used in Section \ref{sec:envelopedetection}, where the envelope detection algorithm is presented.
+We next explore, in Section \ref{sec:curvature}, the concept of discrete curvature, used in Section \ref{sec:envelopedetection} to discriminate which samples of a discrete signal belong to the envelope.
 
-Section \ref{sec:quality} discusses the difficulties in assessing the quality of an envelope, proposing qualitative and quantitative approaches, and demonstrates that the method presented conforms to four requirements for a physically plausible envelope proposed in \cite{1996Loughlinamplitude}.
+Section \ref{sec:quality} discusses the difficulties in assessing the quality of an envelope, suggesting qualitative and quantitative approaches.
+It is also demonstrated that the method presented conforms to the four requirements for a physically plausible envelope proposed in \cite{1996Loughlinamplitude}.
+We also compare the presented algorithm with traditional envelope detection methods in terms of accuracy and processing time.
 
-In Section \ref{sec:comparison} we compare the presented algorithm with traditional envelope detection methods in terms of accuracy and time and present, in Section \ref{sec:extensions}, potential applications of the algorithm, some of which might not be immediately obvious.
+Section \ref{sec:extensions} discusses potential applications of the algorithm, some of which might not be immediately obvious.
 
 Links to interactive visualizations are available from the repository prepared for this work \cite{2020TarjanoEnvelope}, which also contains an implementation of the method in pure Python.
+
 A Python module can be installed from the Python Package Index (PyPI) with the command “pip install signal-envelope”.
-Besides, those working under Windows 64-bit can benefit from a specialized dynamic library coded in C++ that is automatically used within this module, and can be separately obtained via the release section in the repository.
+Those working under Windows 64-bit can benefit from a specialized dynamic library implemented in C++, automatically used within this module.
+This library can be obtained separately via the repository's release section.
 
@@
 
-The interest in the theory of envelope detection arose in the analog domain, with the widespread adoption of radio communications, where the problem is typically restricted to artificial waves, often with well-defined frequencies \cite{2011TurnerDemodulation}.
+The interest in the theory of envelope detection arose in the analog domain, with the widespread adoption of radio communications, where the problem is typically restricted to artificial waves with well-defined frequencies \cite{2011TurnerDemodulation}.
 
-While there is no general formal definition for the envelope of broadband signals \cite{2014Yangnovel,2015Yangtheoretical,2019Jiaempirical}, the consensus is that an envelope is a smooth curve situated as close as possible to the signal \cite{2019Jiaempirical}.
+While there is no general formal definition for the envelope of broadband signals \cite{2014Yangnovel,2015Yangtheoretical,2019Jiaempirical}, the intuition is that the ideal envelope is a smooth curve situated as close as possible to the signal \cite{2019Jiaempirical}, without intersecting with it.
 
-In this context, the most mathematically sound definition of an envelope, as introduced by Gabor in 1946 \cite{2007HahnHistory}, involves representing the underlying wave as an analytic signal.
-In his work, \cite{1946GaborTheory} applies the then relatively new mathematical machinery of quantum mechanics to unify the time and frequency-domain representations of a wave, showing how the Hilbert transform could be applied to a real signal in order to obtain an equivalent complex signal, later known as the analytic signal.
+The most mathematically sound definition of an envelope, introduced by Gabor in 1946 \cite{2007HahnHistory}, involves representing the underlying wave as an analytic signal.
+In his work \cite{1946GaborTheory}, Gabor employs the then relatively new mathematical machinery of quantum mechanics to unify the time and frequency-domain representations of a wave, showing how the Hilbert transform can be applied to a real signal in order to create an equivalent complex signal.
 
-This analytic signal has the form $ A(t) = S(t) + \mathbb{H}(S(t)) \ \text{i} $ \cite{2016HePraat} where $ S(t) $ is the original real signal.
+This complex signal became known as the analytic signal, and has the form $ A(t) = S(t) + \mathbb{H}(S(t)) \ \text{i} $ \cite{2016HePraat}, where $ S(t) $ is the original real signal.
 $ \mathbb{H}(S(t)) $, the Hilbert transform of the original signal, becomes the imaginary part of the analytic signal; the envelope of a signal thus represented can be straightforwardly obtained by the computation of the complex modulus of $ \mathbb{H}(S(t)) $.
-This representation is common in classical AM-FM literature.
+The analytic signal representation is common in traditional AM-FM literature.
 
@@
 It is readily noticeable from Fig.
-\ref{fig:AnalyticSignal} that, as soon as noise is introduced in the original signal, it is directly reflected in the envelope.
+\ref{fig:AnalyticSignal} that, as soon as noise is introduced in the original signal, it is directly reflected on the envelope.
 
@@
     \includegraphics{images/01AnalyticSignal.pdf}
-  \caption{Envelope of a pure sinusoid with a local frequency of 20 cycles, modulated by a polynomial of degree 3, as obtained by the Hilbert Transform approach.
-In the first half of the samples, the sinusoid is free of noise, while in the second half Gaussian noise, with a standard deviation of $ 1/10 $ of the maximum amplitude of the wave, was added to the base signal.}
+  \caption{Envelope of a pure sinusoid with a local frequency of 20 cycles, modulated by a polynomial of degree 3, obtained by the Hilbert Transform approach.
+In the first half of the samples, the sinusoid is free of noise.
+In the second half Gaussian noise with a standard deviation of $ 1/10 $ of the maximum amplitude of the wave was added to the signal.}
   \label{fig:AnalyticSignal}
@@
 
-For real-world signals, such as an alto singer sustaining a steady note as shown in Fig.
+For real-world signals, such the one of an alto singer sustaining a steady note partially shown in Fig.
 \ref{fig:AnalyticSignalAlto}, the Hilbert transform alone provides a poor envelope, since it retains most of the frequencies of the underlying wave.
-With further processing, however, such as filtering, a smoother output can be obtained.
+With further processing, such as filtering, a smoother output can be obtained.
 
@@
     \includegraphics{images/01AnalyticSignalAlto.pdf}
-  \caption{A section of the representation of an alto singer sustaining a steady note comparing the envelopes obtained by the Hilbert transform with and without further filtering.}
+  \caption{Comparison of the envelopes obtained by the Hilbert transform with and without further filtering, for a section of the representation of an alto singer sustaining a steady note.}
   \label{fig:AnalyticSignalAlto}
@@
 
-In general, the problem of envelope detection, or demodulation, can be interpreted as the task of, given a continuous wave $ W(t) $, decomposing it in two components $ E(t) $ and $ C(t) $, such that $ W(t) = E(t) C(t) $ \cite{2011TurnerDemodulation}.
+In general, the problem of envelope detection can be interpreted as the task of, given a continuous wave $ W(t) $, decomposing it in two waves $ E(t) $ and $ C(t) $, such that $ W(t) = E(t) C(t) $ \cite{2011TurnerDemodulation}.
 
@@
 
-We address this problem by assuming $ C(t) $ to be normalized between $ \{-1, 1\}$ or, in other words $ |\max(C(t))| = 1 $.
-This assumption does not cause any loss of generality since, given an arbitrary $ C(t) $, we can obtain its normalization by dividing the function by its absolute global maximum, that is, $ \hat{C}(t) = C(t) / \max(|C(t)|) $, provided that $ \max(|C(t)|) \ne 0 $.
+We address this problem by assuming $ C(t) $ to be normalized between $ \{-1, 1\}$ such that $ |\max(C(t))| = 1 $.
+This assumption does not cause any loss of generality, since, given an arbitrary $ C(t) $, we can obtain its normalization by dividing the function by its absolute global maximum, that is, $ \hat{C}(t) = C(t) / \max(|C(t)|) $, provided that $ \max(|C(t)|) \ne 0 $.
 
-In this work we are concerned with the discrete version of the envelope detection problem: given a finite digital wave represented by the vector $ \mathbf{w} \in \mathbb{R}^n $, obtaining the temporal envelope $ \mathbf{e} $ of $ \mathbf{w} $.
+In this work we are concerned with the discrete version of the envelope detection problem: given a finite discrete wave represented by the vector $ \mathbf{w} \in \mathbb{R}^n $, obtaining the temporal envelope $ \mathbf{e} \in \mathbb{R}^n $ of $ \mathbf{w} $.
 
-The preceding definitions can be translated to this discrete scenario assuming that the discrete quantities arise from observing the continuous ones at regular time intervals.
+The continuous definitions can be translated to this discrete scenario assuming that the discrete quantities arise from observing the continuous ones at regular time intervals.
 In that case, the equality $ i = t \ \text{fps} $ can be used to link both settings, where $ i $ is the index of each sample of the discrete signal, $ t $ stands for the time in seconds, and fps is the frame rate, or the number of observations made in one second.
@@
 
-The $ \odot $ operator in Equation \ref{eq:Envelope} stands for the Hadamard product, denoting element-wise multiplication of two vectors.
+The $ \odot $ operator in Equation \ref{eq:Envelope} stands for the Hadamard product and denotes element-wise multiplication of two vectors.
 Fig.
-\ref{fig:Envelope} provides an example of the vectors just defined: \textbf{w} is obtained by multiplying the normalized carrier wave \textbf{c} using the envelope \textbf{e}.
+\ref{fig:Envelope} provides an example of the vectors just defined: \textbf{w} is obtained by element-wise multiplication between the normalized carrier wave \textbf{c} and the envelope \textbf{e}.
 
@@
     \includegraphics{images/02Envelope.pdf}
-  \caption{Example of a discrete wave \textbf{w} composed by the element-wise multiplication of a previously known envelope \textbf{e} and carrier \textbf{c}.
+  \caption{Example of a discrete wave \textbf{w} arising from the element-wise multiplication of an envelope \textbf{e} and an carrier \textbf{c}, both previously known.
 The local extrema of \textbf{w} are highlighted with a circle.}
@@
 
-In Fig.
-\ref{fig:Envelope} a signal arising from the multiplication of a known carrier wave \textbf{c} and a known envelope \textbf{e} is shown; this helps to illustrate a key concept: note from Fig.
-\ref{fig:Envelope} that, only at some local extrema, the envelope can be inferred without prior knowledge of the underlying carrier, being, in fact, equal to the value of the instantaneous amplitude of the wave $ w_i $.
-That is, $ c_i = 1 $ at those points, rendering $ e_i = 1 w_i $.
+Fig. \ref{fig:Envelope} helps to illustrate a key concept: only at some local extrema the envelope can be inferred without prior knowledge of the underlying carrier.
+
+TODO
+% being, in fact, equal to the value of the instantaneous amplitude of the wave $ w_i $. That is, $ c_i = 1 $ at those points, rendering $ e_i = 1 w_i $.
 
@@
 \ref{fig:Envelope}, the carrier, the original signal and even the envelope are representations of similar entities: signals in the same coordinate system.
-This is in line with \cite{2011CaetanoImproved,2011TurnerDemodulation,2017LyonsDigital} but in contrast with some representations found in the literature of amplitude and frequency modulation of narrowband signals \cite[e.g.,][]{1996Loughlinamplitude} where the FM and AM of a signal are represented in different coordinate systems, commonly frequency versus time and amplitude versus time, respectively.
+This is in line with \cite{2011CaetanoImproved,2011TurnerDemodulation,2017LyonsDigital}, but in contrast with some representations found in the literature of amplitude modulation of narrowband signals \cite[e.g.][]{1996Loughlinamplitude}, where the FM and AM of a signal are represented in different coordinate systems, commonly frequency versus time and amplitude versus time, respectively.
 
@@
 
-For this reason, we define P as the set of the points $ \{P_0, P_1, \cdots, P_{m-1}\} $ where $ P_j = (i_j, \lvert w_{i_j} \lvert) $, that is, $ \lvert w_{i_j} \lvert $, the absolute value of the extrema of each pulse, becomes the ordinate of each point and $ i_j $, its original index, becomes the abscissa.
+For this reason, we define P as the set of the points $ \{P_0, P_1, \cdots, P_{m-1}\} $ where $ P_j = (i_j, \lvert w_{i_j} \lvert) $, that is, $ \lvert w_{i_j} \lvert $, the absolute value of the extremum of each pulse, becomes the ordinate of each point and $ i_j $, its original index, becomes the abscissa.
+This is illustrated in Fig.
+\ref{fig:Pulses}.
 
-This relation is illustrated in Fig. \ref{fig:Pulses}. We call it P to emphasize the connection with the pulses of \textbf{w}.
-While the exact relation between $ n $, the number of samples of the original signal, and $ m $, the cardinality of the set of points, is dependent on factors such as the frequency and sampling rate of the original wave, the inequality $ m \le n $ holds for all discrete signals.
-In practice, $ m \ll n $ for most real-world signals.
+While the exact relationship between $ n $, the number of samples of the original signal, and $ m $, the cardinality of the set P, is dependent on factors such as the frequency and sampling rate of the original wave, the inequality $ m \le n $ holds for all discrete signals.
+In practice, $ m \ll n $ is true for most real-world signals.
 
@@
 
-Geometrically, we can think of these points as points in a vector space on a non-standard, orthogonal basis vector $ \mathbf{b} = \{ \mathbf{b}_x, \mathbf{b}_y \}$ with $ \mathbf{b}_x = (a, 0), \mathbf{b}_y = (0, i) $ where $ a $ is an amplitude unit such as decibel, a measure of the instantaneous sound pressure, volt or even a fraction of a fixed maximum value.
+Geometrically, we can think of these points as points in a vector space on a non-standard, orthogonal basis vector $ \mathbf{b} = \{ \mathbf{b}_x, \mathbf{b}_y \}$ with $ \mathbf{b}_x = (i, 0), \mathbf{b}_y = (0, a) $, where $ a $ is an amplitude unit such as decibel, a measure of the instantaneous sound pressure, volt or even a fraction of a fixed maximum value.
 $ i $ represents the index of the sample, and is linked to time by the aforementioned relation $ i = t \ \text{fps} $.
 
-We are interested in representing the point in the standard basis $ \mathbf{s} = \{ \mathbf{s}_x, \mathbf{s}_y \} = \{ (1,0),(0,1) \}$ of this vector space, making the geometric representation consistent.
+We are interested in representing those points on the Cartesian coordinate system, and for that we need a transformation that places those points on the standard basis $ \mathbf{s} = \{ \mathbf{s}_x, \mathbf{s}_y \} = \{ (1,0),(0,1) \}$ of this vector space.
 
 To achieve that, we divide the ordinate of each point by the average of all ordinates, effectively canceling the unit of the vertical components of the points.
-If we proceeded similarly concerning the horizontal components,
-the direct link between P and \textbf{w}, that arises from the fact that the abscissas of the points in P are numerically equivalent to the indices $ i $ of \textbf{w}, would be lost.
+If we were to proceed similarly concerning the horizontal components,
+the direct link between P and \textbf{w} would be lost.
 
-It is best to leave the abscissas intact and multiply the now unitless vertical coordinates of each point by the average of the difference between the abscissa of a point and the abscissa of the immediately posterior.
-In this way, we can consider both axes as having unit $ i $, and the relationship is preserved, making it easier to recover the envelope points in the original coordinates later.
+It is best to leave the abscissas intact and multiply the now unitless vertical coordinates of each point by the average of the difference between the abscissa of a point and the abscissa of the immediately posterior point.
 
-Considering $P_{\mathbf{b}}$ as a point of the set P represented on the original basis and $P_{\mathbf{s}}$ the same point as represented on the standard basis, we can formalize this transformation as in Equation \ref{eq:Basis}:
+In this way, we can consider both axes as having unit $ i $.
+Dividing them by $ i $, we obtain the effect of representing the points on the standard basis.
+Numerically, the relationship is preserved, since the abscissas of the points $ P_{\mathbf{s}} $ are numerically equivalent to the indices $ i $ of \textbf{w}, making it straightforward to recover the envelope points on the original coordinates later.
+
+Considering $ P_{\mathbf{b}} $, a point of the set P represented on the original coordinate system and $ P_{\mathbf{s}} $ the same point represented on the Cartesian coordinate system, we formalize this transformation in Equation \ref{eq:Basis}:
 
@@
 \begin{split}
-P_{\mathbf{b}} &= (x_1, x_2) \\
-\beta &= \frac{\left( \frac{i_{j-1} - i_j}{m-1} \right)}{\left( \frac{\sum_{j=0}^{m-1} \lvert w_j \lvert}{m} \right)}  \\
-P_{\mathbf{s}} &= (x_1, \beta x_2) \quad \forall \  P \in \text{P} \\ 
+P & = (x, y) \\
+\beta &= \frac{\left( \frac{i_{j+1} - i_j}{m-1} \right)}{\left( \frac{\sum_{j=0}^{m-1} \lvert w_j \lvert}{m} \right)}  \\
+P_{\mathbf{b}} &= x \mathbf{b}_x + y \mathbf{b}_y = x (i, 0) + y (0, a) = (x i, y a) \\
+P_{\mathbf{s}} &= x \mathbf{s}_x + \beta y \mathbf{s}_y = x (1, 0) + \beta y (0, 1) = (x, \beta y) \\
+&\forall \  P \in \text{P} \\ 
 \end{split}
@@
 
-Algorithms such as the convex hull and the alpha shapes solve that problem by discriminating which points of a set are part of its frontier.
+Algorithms, such as the convex hull and the alpha shapes, solve this problem by discriminating which points of a set are part of its frontier.
 
@@
 
-Alternatively, one can picture a circle of fixed radius being rolled above the set P, and marking the points it touches as part of the envelope.
+Alternatively, one can picture a circle of fixed radius being rolled above the points in the set P, marking the points it touches as part of the envelope.
 
@@
 The two general approaches are the derivation of direct methods that use characteristics of the discrete curve to calculate the curvature, or the use of the curvature of a smooth, continuous curve fitted to the original discrete curve \cite{2001CoeurjollyDiscrete}.
-Since we are dealing with a potentially large set of points, it is important for this measure to act locally, to assure $ O(n) $ complexity, as is generally not the case for approaches that involve curve fitting.
+Since we are dealing with a potentially large set of points, it is important for this measure to act locally, to assure $ O(n) $ complexity; this is generally not the case for approaches involving curve fitting.
 
-Concerning direct methods \cite{2014CarrollSurvey}, derive three such definitions based on the approximation of a circle by an inscribed, centered, and circumscribed polygon.
-In the context of three-dimensional meshes, \cite{2016VasaMultivariate} evaluate a range of existing estimators from a multivariate point of view.
+Concerning direct methods, \cite{2014CarrollSurvey} presents three such definitions based on the approximation of a circle by an inscribed, centered, and circumscribed polygon.
+In the context of three-dimensional meshes, \cite{2016VasaMultivariate} evaluate a range of existing estimators, from a multivariate point of view.
+
 Those approaches define the discrete curvature in relation to the vertices of a discrete curve (or mesh, in the three-dimensional case).
@@
 Thus, the set V of the vectors from each point of P to the next, shown in the Fig.
-\ref{fig:Vectors}, will be used to estimate the average curvature of P: We will calculate the equivalent circle for each vector, and average the values thus obtained to infer the average curvature of the whole discrete curve.
+\ref{fig:Vectors}, will be used to estimate the average curvature of P: We will calculate the equivalent circle for each vector, and average the values obtained to infer the average curvature of the discrete curve.
 
@@
 
-We formalize the set V of vectors in Equation \ref{eq:Vectors} below.
+We formalize the set V of vectors in Equation \ref{eq:Vectors}.
 
@@
 \ref{fig:DiscreteCurvature}.
-$ v_k $ represents the net displacement of the envelope from the extrema $ P_{j} = (x_j, y_j) $ to $ P_{j+1} = (x_{j+1}, y_{j+1}) $.
+The vector $ v_k $ represents the envelope's displacement from the extrema $ P_{j} = (x_j, y_j) $ to $ P_{j+1} = (x_{j+1}, y_{j+1}) $.
 
-We assume from $ v_k $ that, near the vector, the envelope defines a vertical displacement of $ v_{k,y} $ in the horizontal distance of $ v_{k,x} $; since we don't yet know which points of P are part of the envelope, we can't be certain that the envelope passes through $ P_{j-1} $ before reaching $ P_{j} $, or even if it reaches $ P_{j} $.
+We assume that, near the vector $ v_k $, the envelope defines a vertical displacement $ v_{k,y} = y_{j+1} - y_j $ and a horizontal displacement $ v_{k,x} = x_{j+1} - x_j $.
+Since we don't yet know which points of P are part of the envelope, we can't be certain that the envelope passes through $ P_{j-1} $ before reaching $ P_{j} $, or even if it reaches $ P_{j} $.
 
-Nevertheless, assuming it passes through $ P_{j} $, and is limited to the diplacement observed, the envelope might have passed anywhere in the line segment from points $ (x_j - (x_{j+1} - x_j), y_j - (y_{j+1} - y_j)) $ to $ (x_j - (x_{j+1} - x_j), y_j + (y_{j+1} - y_j)) $, the bold line segment before $ v_k $ in Fig.
-\ref{fig:DiscreteCurvature}.
+Nevertheless, assuming it passes through $ P_{j} $, and is limited to the displacements just observed, the envelope might have passed anywhere in the line segment from points $ (x_j - (x_{j+1} - x_j), y_j - (y_{j+1} - y_j)) $ to $ (x_j - (x_{j+1} - x_j), y_j + (y_{j+1} - y_j)) $, the bold line segment before $ v_k $ in Fig.
+\ref{fig:DiscreteCurvature}, before reaching $ P_{j} $.
 
@@
 
-The rationale then is to find, for each vector $ v_k $, the radius $ r_k $ of the equivalent circle whose unit tangent performs the same change in direction, in the same horizontal distance, as illustrated in Fig.
-\ref{fig:DiscreteCurvature}.
+The rationale then is to find, for each vector $ v_k $, the radius $ r_k $ of the equivalent circle whose unit tangent performs the same change in direction, in the same horizontal distance.
 The average curvature of P will be obtained by the average of all radii $ r_k $.
@@
 From Fig.
-\ref{fig:DiscreteCurvature} one can see that $ r_k = v_{k,x} / \sin(\theta_k) $ where $ v_{k,x} $ is the component of $ v_k $ in the horizontal direction; $ \theta_k $ itself can be obtained using the slope $ m_0 = 0 $ of the horizontal line and the slope $ m_k = v_{k,y} / v_{k,x} $ defined by $ v_k $ using the equality $ \theta_k = \arctan \big( (m_k - m_0) / (1 + m_k m_0) \big) $.
+\ref{fig:DiscreteCurvature} one can see that $ r_k = v_{k,x} / \sin(\theta_k) $, where $ v_{k,x} $ is the component of $ v_k $ in the horizontal direction; $ \theta_k $ itself can be obtained using the slope $ m_0 = 0 $ of the horizontal line and the slope $ m_k = v_{k,y} / v_{k,x} $, defined by $ v_k $, using the equality $ \theta_k = \arctan \big( (m_k - m_0) / (1 + m_k m_0) \big) $.
 
@@
 The radius $ r_k $ of the circle equivalent to each $ v_k $ is defined in Equation \ref{eq:Radius}.
-This equation has the desirable property that $ v_{k,y} \to 0 \implies v_{k,y} \to \infty $, that is, the radius tends to infinity when no vertical oscillation occurs, effectively turning the circle into a line.
+This equation has the desirable property $ v_{k,y} \to 0 \implies v_{k,y} \to \infty $, that is, the radius tends to infinity when no vertical oscillation occurs, effectively turning the circle into a line.
 
-The radius is also directly proportional to $ v_{k,x} $, making the radius larger the more the points are horizontally far from each other.
+The radius is also directly proportional to $ v_{k,x} $, making the radius larger the more the points are horizontally distant from each other.
 
@@
 We need to construct an algorithm, using the radius just obtained, to identify the points that belong to the envelope.
-Generally, alpha shapes based algorithms resort to the construction of the Delaunay triangulation of the set of points, which is later filtered to contain only the outer edges.
+Generally, alpha shapes algorithms resort to the construction of the Delaunay triangulation of the set of points, which is later filtered to contain only the outer edges.
 We can adopt a more straightforward approach, since our points have a sequential structure.
 
-Let $ x_o, y_o \in \mathbb{R}, \quad 0 \le y_o $ be the coordinates of the center of a circle with radius $ r $, that can be placed anywhere in the upper plane of the coordinate system;
+Let $ x_o, y_o \in \mathbb{R}, \quad 0 \le y_o $ be the coordinates of the center of a circle with radius $ r $, that can be placed anywhere in the upper plane of the coordinate system, and $ P_a, P_b \in \text{P} $ be two points of the set P;
 
-Let $ P_a, P_b \in \text{P}, \quad P_a \ne P_b $ be two different points of the set P;
-
-For all circles of center $ x_o, y_o $ and radius $ r $ that can be constructed to pass through any two different points of P, that is, all circles such that $ (P_{a,x} - x_o)^2 + (P_{a,y} - y_o)^2 = (P_{b,x} - x_o)^2 + (P_{b,y} - y_o)^2 = r^2 $, $ P_a $ and $ P_b $ will be part of the set $ E $ of the points belonging to the envelope if and only if exist one such circle that defines a disc that does not contain any other point of P.
-Summarizing this, we have Eq.
-\ref{eq:InEnvelope}:
+For all circles of center $ x_o, y_o $ and radius $ r $ that can be constructed to pass through any two points of P, that is, all circles such that $ (P_{a,x} - x_o)^2 + (P_{a,y} - y_o)^2 = (P_{b,x} - x_o)^2 + (P_{b,y} - y_o)^2 = r^2 $, $ P_a $ and $ P_b $ will be part of the set $ E $ of the points belonging to the envelope if and only if exists one such circle defining a disc that does not contain any other point of P.
+Summarizing this, we have Equation \ref{eq:InEnvelope}:
 
@@
 The algorithm follows directly from the definition in Equation \ref{eq:InEnvelope} after noting that the first and last points of P will always be part of the envelope.
-Since $ P_a $ can be equal to $ P_b $, it is always possible to construct a circle containing those points, by placing the center of the circle to the left of or right of the first or last point, respectively.
+Since $ P_a $ can be equal to $ P_b $, it is always possible to construct a circle containing those points, by placing the center of the circle as far as possible to the left or right of the first or last point, respectively.
 The discs defined by those circles can't contain any other point.
@@
 
-The set E obtained via the algorithm \ref{FindEnvelope} is a subset of P, and thus also a set of points in the Euclidean space.
+The set E obtained by the algorithm \ref{FindEnvelope} is a subset of P, and thus also a set of points in the Euclidean space.
 

### 2021-05-11 Papers/01_Pulses_Envelope/00ElsevierPulsesEnvelope.tex
revising

@@
 Envelope detection techniques have applications in medicine, sound classification, sound synthesis, seismology, and speech recognition.
-Nevertheless, a general approach to envelope detection doesn't exist.
+Nevertheless, a general approach to envelope detection does not exist.
 Most methods involve manual intervention, in the form of filter design, smoothing, and other specific design choices based on prior knowledge of the signals under investigation.
@@
     \includegraphics{images/00AlphaShape.pdf}
-  \caption{The alpha shape of a set of points for $ \alpha=10 $. A circle of radius 10, containing two points of the set, but defining a disc that doesn't contain any point of the set, is also shown.}
+  \caption{Alpha shape of a set of points for $ \alpha=10 $. A circle of radius 10, containing two points of the set, but defining a disc that does not contain any point of the set, is also shown.}
   \label{fig:AlphaShape}
@@
 
-While there is no general formal definition for the envelope of broadband signals \cite{2014Yangnovel,2015Yangtheoretical,2019Jiaempirical}, the intuition is that the ideal envelope is a smooth curve situated as close as possible to the signal \cite{2019Jiaempirical}, without intersecting with it.
+While there is no general formal definition for the envelope of broadband signals \cite{2008HlawatschTime,2012HuEMD,2014Yangnovel,2015Yangtheoretical,2019Jiaempirical}, the intuition is that the ideal envelope is a smooth curve situated as close as possible to the signal \cite{2014Yangnovel,2015Yangtheoretical,2019Jiaempirical}, without intersecting with it.
 
@@
 
-Despite its widespread use, and effectiveness in the context of narrowband signals, envelope detection techniques based on the Hilbert transform don't behave well for broadband signals \cite{2012DauSpeech, 2019Jiaempirical}, being even physically paradoxical in some cases \cite{1996Loughlinamplitude, 2014Yangnovel}.
+Despite its widespread use, and effectiveness in the context of narrowband signals, envelope detection techniques based on the Hilbert transform do not behave well for broadband signals \cite{2012DauSpeech, 2019Jiaempirical}, being even physically paradoxical in some cases \cite{1996Loughlinamplitude, 2014Yangnovel}.
 
@@
 
-Geometrically, we can think of these points as points in a vector space on a non-standard, orthogonal basis vector $ \mathbf{b} = \{ \mathbf{b}_x, \mathbf{b}_y \}$ with $ \mathbf{b}_x = (i, 0), \mathbf{b}_y = (0, a) $, where $ a $ is an amplitude unit such as decibel, a measure of the instantaneous sound pressure, volt or even a fraction of a fixed maximum value.
+Geometrically, we can think of these points as lying in a vector space with a non-standard, orthogonal basis vector $ \mathbf{b} = \{ \mathbf{b}_x, \mathbf{b}_y \}$ with $ \mathbf{b}_x = (i, 0), \mathbf{b}_y = (0, a) $, where $ a $ is an amplitude unit such as decibel, a measure of the instantaneous sound pressure, volt or even a fraction of a fixed maximum value.
 $ i $ represents the index of the sample, and is linked to time by the aforementioned relation $ i = t \ \text{fps} $.
 
-We are interested in representing those points on the Cartesian coordinate system, and for that we need a transformation that places those points on the standard basis $ \mathbf{s} = \{ \mathbf{s}_x, \mathbf{s}_y \} = \{ (1,0),(0,1) \}$ of this vector space.
+We are interested in representing those points in the Cartesian coordinate system, and for that we need a transformation that allows those points to be described in the standard basis $ \mathbf{s} = \{ \mathbf{s}_x, \mathbf{s}_y \} = \{ (1,0),(0,1) \}$ of this vector space.
 
@@
 In this way, we can consider both axes as having unit $ i $.
-Dividing them by $ i $, we obtain the effect of representing the points on the standard basis.
-Numerically, the relationship is preserved, since the abscissas of the points $ P_{\mathbf{s}} $ are numerically equivalent to the indices $ i $ of \textbf{w}, making it straightforward to recover the envelope points on the original coordinates later.
+Dividing them by $ i $, we obtain the effect of representing the points in the standard basis.
+The relationship is preserved, since the abscissas of the points $ P_{\mathbf{s}} $ are numerically equivalent to the indices $ i $ of \textbf{w}, making it straightforward to recover the envelope points in the original coordinate system.
 
-Considering $ P_{\mathbf{b}} $, a point of the set P represented on the original coordinate system and $ P_{\mathbf{s}} $ the same point represented on the Cartesian coordinate system, we formalize this transformation in Equation \ref{eq:Basis}:
+Considering $ P_{\mathbf{b}} $, a point of the set P represented in the original coordinate system and $ P_{\mathbf{s}} $ the same point represented in the Cartesian coordinate system, we formalize this transformation in Equation \ref{eq:Basis}:
 
@@
 \beta &= \frac{\left( \frac{i_{j+1} - i_j}{m-1} \right)}{\left( \frac{\sum_{j=0}^{m-1} \lvert w_j \lvert}{m} \right)}  \\
-P_{\mathbf{b}} &= x \mathbf{b}_x + y \mathbf{b}_y = x (i, 0) + y (0, a) = (x i, y a) \\
-P_{\mathbf{s}} &= x \mathbf{s}_x + \beta y \mathbf{s}_y = x (1, 0) + \beta y (0, 1) = (x, \beta y) \\
+P_{\mathbf{b}} &= x \ \mathbf{b}_x + y \ \mathbf{b}_y = x (i, 0) + y (0, a) = (x i, y a) \\
+P_{\mathbf{s}} &= x \ \mathbf{s}_x + \beta \ y \ \mathbf{s}_y = x (1, 0) + \beta y (0, 1) = (x, \beta y) \\
 &\forall \  P \in \text{P} \\ 
@@
 The effect of this transformation can be seen in the points represented in Fig.
-\ref{fig:Coordinates}, shown in scale for both coordinate systems.
-We can consider our points residing in the Cartesian coordinate system after the transformation.
+\ref{fig:Coordinates}.
+Despite the two bases being different, the horizontal relationship is preserved, since one unit along the $ x $ axis in Cartesian coordinates corresponds to one unit of the sample index $ i $ in the original coordinate system.
 
@@
     \includegraphics{images/04Coordinates.pdf}
-  \caption{The set P of points, in both the original (top) and standard (bottom) coordinate systems, shown in scale.}
+  \caption{The set P of points, in both the original (top) and Cartesian (bottom) coordinate systems, shown in scale.}
   \label{fig:Coordinates}
@@
 
-Algorithms, such as the convex hull and the alpha shapes, solve this problem by discriminating which points of a set are part of its frontier.
+Algorithms such as the convex hull and the alpha shapes solve this problem by discriminating which points of a set are part of its frontier.
 
-A formal definition of alpha shapes is given at \cite{1983Edelsbrunnershape}.
-In the context of this work, however, gaining the intuition of the algorithm is more important, and thus the explanation presented in \cite{1994EdelsbrunnerThree} can be more insightful.
+A formal definition of alpha shapes is given in \cite{1983Edelsbrunnershape}.
+In the context of this work, however, gaining intuition about the algorithm is more important, and for that the explanation presented in \cite{1994EdelsbrunnerThree} can be more insightful.
 
-The explanation, translated to our context, invites one to think of the plane where the points of P are represented as a soft material, such as Styrofoam.
+The explanation, translated to our context, invites one to think of the plane where the points in the set P are represented as a soft material, such as Styrofoam.
 The points themselves are made of a harder material, such as rock, and occupy a fixed position in this plane.
-The alpha shapes algorithm consists of carving this plane from the outside of those points, with a circular eraser of fixed radius $ \alpha $.
-At the end of the process, the remaining Styrofoam that couldn't be erased due to the impeding rocks is the alpha-shape of this set of points.
+The alpha shapes algorithm can be seen as carving this plane from the outside of those points with a circular eraser of fixed radius $ \alpha $.
+At the end of the process, the remaining Styrofoam that could not be erased due to the impeding rocks is the alpha-shape of this set of points.
 
@@
 
-Discrete curvature estimation is an important task in image processing \cite{2010FleischmannNovel} for which no default definition exists.
+Discrete curvature estimation is an important task in image processing \cite{2010FleischmannNovel}, for which no default definition exists.
 
@@
 Thus, the set V of the vectors from each point of P to the next, shown in the Fig.
-\ref{fig:Vectors}, will be used to estimate the average curvature of P: We will calculate the equivalent circle for each vector, and average the values obtained to infer the average curvature of the discrete curve.
+\ref{fig:Vectors}, will be used to estimate the average curvature of P: We will calculate the equivalent circle for each vector, and average the values obtained to infer the curvature of the set.
 
@@
 We assume that, near the vector $ v_k $, the envelope defines a vertical displacement $ v_{k,y} = y_{j+1} - y_j $ and a horizontal displacement $ v_{k,x} = x_{j+1} - x_j $.
-Since we don't yet know which points of P are part of the envelope, we can't be certain that the envelope passes through $ P_{j-1} $ before reaching $ P_{j} $, or even if it reaches $ P_{j} $.
+Since we do not yet know which points of P are part of the envelope, we cannot be certain that the envelope passes through $ P_{j-1} $ before reaching $ P_{j} $, or even if it reaches $ P_{j} $.
 
@@
 The radius $ r_k $ of the circle equivalent to each $ v_k $ is defined in Equation \ref{eq:Radius}.
-This equation has the desirable property $ v_{k,y} \to 0 \implies v_{k,y} \to \infty $, that is, the radius tends to infinity when no vertical oscillation occurs, effectively turning the circle into a line.
+This equation has the desirable property $ v_{k,y} \to 0 \implies v_{k,y} \to \infty $, that is, the radius tends to infinity when there is no vertical displacement, effectively turning the circle into a line.
 
-The radius is also directly proportional to $ v_{k,x} $, making the radius larger the more the points are horizontally distant from each other.
+% The radius is also directly proportional to $ v_{k,x} $, becoming larger in proportion to the horizontal distance between the adjacent points.
 
@@
 In this way, the problem is avoided, since $ v_{k,x} $ is guaranteed to be greater than zero.
-The average of curvatures is then inverted in order to obtain the average radius $ \overline{r} $, as illustrated in Equation \ref{eq:Curvature}.
+The average of the curvatures is then inverted in order to obtain the average radius $ \overline{r} $, as illustrated in Equation \ref{eq:Curvature}.
 
@@
 
-For all circles of center $ x_o, y_o $ and radius $ r $ that can be constructed to pass through any two points of P, that is, all circles such that $ (P_{a,x} - x_o)^2 + (P_{a,y} - y_o)^2 = (P_{b,x} - x_o)^2 + (P_{b,y} - y_o)^2 = r^2 $, $ P_a $ and $ P_b $ will be part of the set $ E $ of the points belonging to the envelope if and only if exists one such circle defining a disc that does not contain any other point of P.
-Summarizing this, we have Equation \ref{eq:InEnvelope}:
+For all circles of center $ (x_o, y_o) $ and radius $ r $ that can be constructed to pass through any two points of P, that is, all circles such that $ (P_{a,x} - x_o)^2 + (P_{a,y} - y_o)^2 = (P_{b,x} - x_o)^2 + (P_{b,y} - y_o)^2 = r^2 $, $ P_a $ and $ P_b $ will be part of the set E of the points belonging to the envelope if and only if exists one such circle defining a disc that does not contain any other point of P.
+Equation \ref{eq:InEnvelope} summarizes this.
 
@@
 
-The algorithm follows directly from the definition in Equation \ref{eq:InEnvelope} after noting that the first and last points of P will always be part of the envelope.
-Since $ P_a $ can be equal to $ P_b $, it is always possible to construct a circle containing those points, by placing the center of the circle as far as possible to the left or right of the first or last point, respectively.
-The discs defined by those circles can't contain any other point.
+The algorithm follows directly from the definition in Equation \ref{eq:InEnvelope}, after noting that the first and last points of P will always be part of the envelope.
+Since $ P_a $ can be equal to $ P_b $, it is always possible to construct a circle containing the first, leftmost point of P, by placing the center of the circle at $ (P_{0,x} - r, P_{0,y}) $.
+The disc defined by this circle cannot contain any other point of P, since all other points have abscissas greater than $ P_{0,x} $.
+The same is true for $ P_{m-1} $, the last and rightmost point of P.
 
-From $ P_0 $, our first pivot point, we construct circles passing through $ P_1, P_2, \cdots, P_d $ until a circle that doesn't contain any point in P is found; $ P_d $ then becomes the new pivot point and is included in E, and this process continues until $ P_{m-1} $ is reached.
+From $ P_0 $, our first pivot point, we construct circles passing through $ P_1, P_2, \cdots, P_d $ until a circle that does not contain any point is found; $ P_d $ then becomes the new pivot point and is included in E, and this process continues until $ P_{m-1} $ is reached.
 
 The procedure is formalized in algorithm \ref{FindEnvelope}.
-We don't provide similar algorithmic descriptions of the preceding steps since the Python implementation available at \cite{2020TarjanoEnvelope} can easily serve as pseudocode.
+We do not provide similar algorithmic descriptions of the preceding steps since the Python implementation available at \cite{2020TarjanoEnvelope} can easily serve as pseudocode.
 
@@
 
-Recall from Section \ref{sec:mapping} that the vertical coordinates of these points are currently scaled, in relation to the original signal, by the transformation described in Equation \ref{eq:Basis}.
-We could invert this transformation, but since we were careful to maintain the direct link between the horizontal coordinates in both coordinate systems, we can simply consider the horizontal coordinates of P as the indices of the samples of the original discrete signal, and retrieve the envelope at each point directly.
+Recall from Section \ref{sec:mapping} that the coordinates of these points are currently scaled, in relation to the original signal, by the transformation described in Equation \ref{eq:Basis}.
+We could invert the transformation, but since we were careful to maintain a link between the horizontal coordinates in both coordinate systems, we can simply consider the horizontal coordinates of P as the indices of the samples of the original discrete signal, and retrieve the envelope at each point directly.
 
-After converting E to the original coordinate system, the most straightforward way to transform those points into the vector $ \textbf{e} \in \mathbb{R}^n $ is via linear interpolation, the approach used in this work.
+After converting E back to the original coordinate system, the most straightforward way to transform those points into the vector $ \textbf{e} \in \mathbb{R}^n $ is via linear interpolation, the approach used in this work.
 
@@
 
-The lack of a precise definition of what an envelope is makes the assessment of the quality of an envelope a difficult task, for which there are no agreed-upon objective metrics.
+The lack of a precise definition of what an envelope is makes the assessment of the quality of an envelope a difficult task \cite{2012HuEMD}, for which there are no agreed-upon objective metrics.
 Recalling Fig.
-\ref{fig:Envelope}, however, and the accompanying discussion, if one divides, element-wise, the original signal by the envelope, the carrier wave thus obtained is expected to be bounded between -1 and 1.
+\ref{fig:Envelope}, however, and the accompanying discussion, if one divides, element-wise, the original signal by the extracted envelope, the carrier wave obtained is expected to be bounded between -1 and 1.
 
-Figs.
-\ref{fig:spokenvoice}, \ref{fig:singer_voice} and \ref{fig:Bend} present discrete waves \textbf{w}, their extracted envelope \textbf{e}, and the inferred carrier \textbf{c}, obtained after dividing, element-wise, the original signal by its extracted envelope.
+Figs. \ref{fig:spokenvoice}, \ref{fig:singer_voice} and \ref{fig:Bend} present real-world discrete waves \textbf{w}, their extracted envelope \textbf{e}, and the inferred carrier \textbf{c}.
 
@@
     \includegraphics{images/06Graph_spoken_voice.pdf}
-  \caption{Signal, envelope and carrier for a record of male voice uttering the word “amazing”.}
+  \caption{Signal, envelope and carrier a male voice uttering the word “amazing”.}
   \label{fig:spokenvoice}
@@
 There, the wave is almost periodic, with a stable waveform and envelope.
+The superior frontier of the carrier wave in this case is approximately horizontal.
 
@@
     \includegraphics{images/06Graph_alto.pdf}
-  \caption{Signal, envelope and carrier for a record of an alto singer.}
+  \caption{Signal, envelope and carrier for an alto singer sustaining a steady note.}
   \label{fig:singer_voice}
@@
 
-Fig. \ref{fig:Bend} shows the envelope in the case of a bend performed in an electric guitar. The fast frequency variations in the signal didn't affect the envelope extracted.
+Fig. \ref{fig:Bend} shows the representation of a bend performed in an electric guitar. The fast frequency variations in the signal did not affect the envelope extracted, nor the carrier wave.
 
@@
     \includegraphics{images/06Graph_bend.pdf}
-  \caption{Signal, envelope and carrier for a record of a bend performed on an electric guitar.}
+  \caption{Signal, envelope and carrier for a bend performed on an electric guitar.}
   \label{fig:Bend}
@@
 
-In an effort to formalize the assessment of the quality of amplitude and frequency modulations of a signal, \cite{1996Loughlinamplitude} proposed some theoretical conditions necessary, but not sufficient, to ensure the physical plausibility of an envelope.
+In an effort to formalize the assessment of the quality of the amplitude and frequency modulations of a signal, \cite{1996Loughlinamplitude} proposed some theoretical conditions necessary, but not sufficient, to ensure the physical plausibility of an envelope.
 
@@
 More formally, we have the set P of the absolute values of the local extrema of the original signal $ \mathbf{w} = (w_0, w_1, \cdots, w_{n-1}) $ that contains, by definition, the global extrema of $ \mathbf{w} $.
-The set E of the points that belong to the envelope is a subset of P. We can thus write Equation \ref{eq:Req1}:
+The set E of the points that belong to the envelope is a subset of P. We can thus write Equation \ref{eq:Req1}, with P represented in the original coordinate system:
 
@@
 
-This requirement doesn't imply loss of generality since, for an arbitrary signal, we can assure this condition by dividing all samples by the value of the minimum absolute local extrema.
+This requirement does not imply loss of generality since, for an arbitrary signal, we can assure this condition by dividing all samples by the value of the minimum absolute local extrema.
 
@@
 For any given frequency $ f $, both $ m_\mathbf{w}(f) $ and $ m_\mathbf{c}(f) $ are nonnegative.
-Also, since $ e_k \ge 1 \ \forall \ k, 0 \le k \le n - 1 $ we establish that $ 0 \le m_\mathbf{c}(f) \le m_\mathbf{w}(f) \ \forall \ f $ which is a more restrictive condition than the original: If $ \mathbf{w} $ has a finite frequency range, the magnitude of frequencies outside that range are zero, as are those of the same frequencies of the carrier $ \mathbf{c} $.
+Also, since $ e_k \ge 1 \ \forall \ k, 0 \le k \le n - 1 $, we establish that $ 0 \le m_\mathbf{c}(f) \le m_\mathbf{w}(f) \ \forall \ f $, which is a more restrictive condition than the original.
+If $ \mathbf{w} $ has a finite frequency range, the magnitude of the frequencies outside that range are zero, as are those of the carrier $ \mathbf{c} $.
 
-The third condition states that a physically plausible algorithm should, for a signal $ \mathbf{w} $ of constant amplitude $ a > 0, \ a \in \mathbb{R} $ and constant frequency, of the form $ a \cos(\omega i + \phi) $, generate a constant envelope $ \mathbf{e} $ of magnitude $ a $ and a carrier wave $ \mathbf{c} $ of the form $ \cos(\omega i + \phi) $.
+The third condition states that a physically plausible algorithm should, for a signal $ \mathbf{w} $ of constant amplitude $ a \in \mathbb{R}, \ a > 0  $ and constant frequency, of the form $ a \cos(\omega i + \phi) $, generate a constant envelope $ \mathbf{e} $ of magnitude $ a $ and a carrier wave $ \mathbf{c} $ of the form $ \cos(\omega i + \phi) $.
 
@@
 \ref{fig:Comparison_sinusoid}, where the envelope of a simple sinusoid is shown.
-To see that this is always the case, note that all the extrema of a sinusoid are of the form $ a (\pm 1) $ and $ P_j = (i_j, \lvert a (\pm 1) \lvert) = (i_j, a) \ \forall P_j \in \text{P} $.
+To see that this is always the case, note that all the extrema of a sinusoid are of the form $ a (\pm 1) $ and $ P_j = (i_j, \lvert a (\pm 1) \lvert) = (i_j, a) \ \forall P_j \in \text{P} $ in the original coordinate system.
 As $ \text{E} \subset \text{P} $, we have that the envelope $ \mathbf{e} = (a, a, \cdots, a) $ and is, thus, constant.
@@
 Recall from Section \ref{sec:mapping} that, before the extraction of the envelope, the original points undergo a transformation to lie in the Cartesian coordinate system.
-After the extraction, the inverse operation is performed on the points of the Envelope.
+After the extraction, the inverse operation is performed on the points of the envelope.
 
-Consider a discrete signal $ \mathbf{w} = (w_0, w_1, \cdots, w_{n-1}) $ and the set P of its absolute extrema, represented in the standard basis, $ \text{P} = \{ P_0, P_1, \cdots, P_{m-1} \}, \quad P_j = (i_j, \beta \lvert w_{i_j} \lvert) $, where $ \beta $ is obtained from Equation \ref{eq:Basis}.
+Consider a discrete signal $ \mathbf{w} = (w_0, w_1, \cdots, w_{n-1}) $ and the set P of its absolute extrema.
+Represented in the Cartesian coordinate system, $ P_j = (i_j, \beta \lvert w_{i_j} \lvert) $, where $ \beta $ is obtained from Equation \ref{eq:Basis}.
 
-Given a constant $ \alpha > 0, \alpha \in \mathbb{R} $, we can define the scaled signal as $ \mathbf{w}^\prime = (\alpha w_0, \alpha w_1, \cdots, \alpha w_{n-1}) $.
-Observing, from Equation \ref{eq:Basis}, that $ \alpha $ can be factored out of denominator of the summation, we have $ \beta^\prime = \beta / \alpha $ and the new set $ \text{P}^\prime = \{ P^\prime_0, P^\prime_1, \cdots, P^\prime_{m-1} \}, \quad P^\prime_j = (i_j, \frac{\beta}{\alpha} \lvert \alpha w_{i_j} \lvert) $ is equal to the set P for the signal before scaling.
+Given a constant $ \alpha \in \mathbb{R}, \ \alpha > 0 $, we can define the scaled signal as $ \mathbf{w}^\prime = (\alpha w_0, \alpha w_1, \cdots, \alpha w_{n-1}) $.
+Observing, from Equation \ref{eq:Basis}, that $ \alpha $ can be factored out of denominator of the summation, we have $ \beta^\prime = \beta / \alpha $ and the new set $ \text{P}^\prime = \{ P^\prime_0, P^\prime_1, \cdots, P^\prime_{m-1} \} $, with $ P^\prime_j = (i_j, \frac{\beta}{\alpha} \lvert \alpha w_{i_j} \lvert) $, is equal to the original set P.
 
 This guarantees that the same subset of points E will be identified as part of the envelope in both cases.
-In the respective original basis system, applying the inverse change of basis, those points are $ P_j = (i_j, \frac{\beta}{\beta} \lvert w_{i_j} \lvert)  \quad \forall \  P \in \text{E} $, for the original signal, and $ P^\prime_j = (i_j, \frac{\alpha}{\beta}\frac{\beta}{\alpha} \lvert \alpha w_{i_j} \lvert) $ for the scaled signal.
+In their respective original basis system, applying the inverse transformation, those points are $ P_j = (i_j, \frac{\beta}{\beta} \lvert w_{i_j} \lvert) = (i_j, |w_{i_j}|) \quad \forall \ P \in \text{E} $, for the original signal, and $ P^\prime_j = (i_j, \frac{\alpha}{\beta}\frac{\beta}{\alpha} \lvert \alpha w_{i_j} \lvert) = (i_j, |\alpha w_{i_j}|)  \quad \forall \ P^\prime \in \text{E}^\prime $ for the scaled signal.
 
@@
 
-In order to obtain a numeric indicator of the quality of an envelope, we construct a reference envelope, that trades off applicability in favor of mathematical formality.
-In \cite{1992BruceCurves} the envelope is defined as the boundary of the region filled by a family of curves.
+% In order to obtain a numeric indicator of the quality of an envelope, we construct a reference envelope, 
+In \cite{1992BruceCurves,2019Jiaempirical}, the envelope is defined as the boundary of the region filled by a family of curves.
 Fig.
@@
 Nevertheless, an approximation can be obtained, by shifting the original signal along its initial position in the horizontal axis by half of its period, and taking the absolute maximum value at each index $i$.
-Fig. \ref{fig:RefEnvelope} shows the reference envelope for the sound of a tom.
+Fig. \ref{fig:RefEnvelope} shows part of the reference envelope for the sound of a tom drum.
 
@@
     \includegraphics{images/10RefEnvelope.pdf}
-  \caption{Reference envelope obtained by shifting the original signal, for part of a sample of a tom.}
+  \caption{Part of the reference envelope obtained by shifting the original signal horizontally, for the sound of a tom drum.}
   \label{fig:RefEnvelope}
@@
 The envelope obtained is not smooth, being composed of a series of horizontal segments, as can be seen in Fig.
-\ref{fig:RefEnvelope}, besides being computationally expensive to calculate.
-However, arising from a mathematically sound background, it can serve as a reference to infer the quality of empirical methods.
+\ref{fig:RefEnvelope}.
+However, those segments agree with the general contour of the wave, and overestimate and underestimate the envelope in equal measures.
+
+Although lacking in applicability, both given its discontinuous shape and the high computational cost, an envelope constructed in this way arises from a mathematically sound background, and can serve as a reference to infer the quality of more practical methods.
 
@@
 
-Direct comparison with many of the more recent algorithms is made difficult by the unavailability of digital implementations of such works, many designed to process analog signals \cite[e.g.,][]{2018AssefModeling}.
+Direct comparison with many recent algorithms is made difficult by the unavailability of digital implementations of such works, some of whice where designed to process analog signals \cite[e.g.][]{2018AssefModeling}.
 
-Nevertheless, insight can be gained from a comparison of the results of the method here proposed with some of the most common envelope extraction algorithms.
+Nevertheless, insight can be gained from comparing the method here proposed with some of the most common envelope extraction algorithms.
 Specifically, we compare the present method with the following approaches:
@@
 
-In Fig. \ref{fig:Comparison_sinusoid} we illustrate the envelope extracted by the conventional algorithms, contrasting it with the result of the method presented in this work, for a simple sinusoid. 
-Only the present algorithm and the one based on the Hilbert transform comply with the third condition proposed by \cite{1996Loughlinamplitude}, which states that the envelope of a periodic wave should be a straight line.
+Fig. \ref{fig:Comparison_sinusoid} presents the comparison in the case of a simple sinusoid. 
+The present algorithm and the Hilbert transform approach comply with the third condition proposed by \cite{1996Loughlinamplitude}, which states that the envelope of a periodic wave should be a straight line.
 
@@
 
-For a recording of the key 33 of a grand piano, where parts of the sound, specially at the start, are highly percussive and noisy, the Hilbert approach presents a considerable undershooting, as can be seen in Fig.
+For a recording of the key 33 of a grand piano, where parts of the sound, specially at the start, are highly percussive and noisy, the Hilbert approach undershoots the original signal, as can be seen in Fig.
 \ref{fig:Comparison_piano}.
@@
     \includegraphics{images/07Comparison_piano.pdf}
-  \caption{Comparison of envelope detection algorithms for the recording of the key 33 of a grand piano}
+  \caption{Comparison of envelope detection algorithms for the sound of key 33 of a grand piano}
   \label{fig:Comparison_piano}
@@
 
-In Fig. \ref{fig:Comparison_soprano} all envelopes exhibit a similar movement along the samples, but the ones generated by traditional algorithms are consistently undershooting the original wave.
+In Fig.
+\ref{fig:Comparison_soprano} a soprano singer sustaining a steady note is shown.
+All envelopes exhibit a similar movement along the samples, but the ones generated by traditional algorithms are consistently undershooting the original wave.
 
@@
     \includegraphics{images/07Comparison_soprano.pdf}
-  \caption{Comparison of envelope detection algorithms for the recording of the vocalization of a soprano singer}
+  \caption{Comparison of envelope detection algorithms for the vocalization of a soprano singer}
   \label{fig:Comparison_soprano}
@@
 Besides, the traditional methods demanded careful choice of parameters.
-In the case of the Hilbert transform, post-processing in the form of filtering was applied, and an appropriate cut-off frequency had to be chosen.
+In the case of the Hilbert transform, post-processing in the form of filtering was needed, and an appropriate cut-off frequency had to be chosen.
 That is in contrast with our method, which organically tunes itself via the automatic choice of the circle's radius representing the average discrete curvature of the signal.
 
-Table \ref{table:Error_comparison} presents the average of the absolute errors of the techniques, when compared with the reference envelope introduced in the end of Section \ref{sec:quality}
+Table \ref{table:Error_comparison} presents the average of the absolute errors of the techniques, when compared with the reference envelope introduced in the end of Section \ref{sec:quality}.
 
@@
 \hline
-\multicolumn{5}{c}{\textbf{Average error per frame}} \\
+\multicolumn{5}{c}{\textbf{Average absolute error}} \\
 Signal & Present Work & Smoothing & Low pass & Hilbert \\ 
 \hline
-alto  & 0.001 & 0.337 & 0.342 & 0.163 \\
-bend  & 0.000 & 0.029 & 0.030 & 0.008 \\
-brass & 0.001 & 0.107 & 0.107 & 0.038 \\
-nonperiodic & 0.000 & 0.211 & 0.158 & 0.000 \\
-piano & 0.003 & 0.016 & 0.016 & 0.007 \\
-sinusoid & 0.000 & 0.132 & 0.131 & 0.000 \\
-soprano & 0.000 & 0.155 & 0.155 & 0.044 \\
-spoken\_voice & 0.006 & 0.109 & 0.110 & 0.045 \\
-tom   & 0.001 & 0.011 & 0.015 & 0.005 \\
+alto  & 0.0033 & 0.5800 & 0.5836 & 0.4021 \\
+bend  & 0.0038 & 0.1268 & 0.1278 & 0.0435 \\
+brass & 0.0109 & 0.3154 & 0.3154 & 0.1866 \\
+nonperiodic & 0.0000 & 0.3623 & 0.3680 & 0.0002 \\
+piano & 0.0195 & 0.0678 & 0.0680 & 0.0388 \\
+sinusoid & 0.0000 & 0.3636 & 0.3614 & 0.0001 \\
+soprano & 0.0027 & 0.3911 & 0.3911 & 0.2091 \\
+spoken\_voice & 0.0496 & 0.2793 & 0.2794 & 0.1612 \\
+tom   & 0.0073 & 0.0506 & 0.0532 & 0.0228 \\
 \hline
-Average & 0.001 & 0.123 & 0.118 & 0.035 \\
-frac. of max. & 0.010 & 1.000 & 0.961 & 0.281 \\
+Average & 0.0108 & 0.2819 & 0.2831 & 0.1182 \\
 \end{tabular}
-\caption{Comparison of the average errors of the algorithm presented in this work with the most common methods of digital envelope identification.
-The discrete waves used were normalized between -1 and 1.}
+\caption{Average absolute errors of each algorithm in relation to the reference envelope. All signals were previously normalized.}
 \label{table:Error_comparison}
@@
 
-We can see from Table \ref{table:Error_comparison} that the method presented here consistently outperforms the traditional approaches for all waves tested, yielding an error of less than 2\% on average, of the error of the traditional methods.
+We can see from Table \ref{table:Error_comparison} that the proposed method consistently outperforms the traditional approaches for all waves tested, yielding an error of less than 2\%, on average, of the error of the traditional methods.
 
@@
 The methods were performed as described in the beginning of this section, using the implementations available at the Scipy library, under the Signal processing module.
-The presented algorithm, is on average, the second-fastest.
+The proposed algorithm, is on average, the second-fastest.
 
@@
 Average & 0.020 & 0.183 & 0.005 & 0.052 \\
-frac. of max. & 0.108 & 1.000 & 0.030 & 0.283 \\
 \end{tabular}
-\caption{Comparison of the processing time of the algorithm presented in this work with the most common methods of digital envelope detection.}
+\caption{Processing time of the algorithms, in seconds.}
 \label{table:Time_comparison}
@@
 
-In practice, is not uncommon for a discrete wave, especially in the case of sound, to present somewhat different superior and inferior contours; in those cases, the algorithm can be easily adapted to infer from the positive and negative pulses defined by the original wave to define two independent “envelopes”, that we are going to call superior and inferior frontiers.
+In practice, is not uncommon for a discrete wave, especially in the case of sound, to present somewhat different superior and inferior contours.
+In those cases, the algorithm can be easily modified to generate two independent envelopes, that we are going to call superior and inferior frontiers of the signal.
 
-If we split P, as defined in Section \ref{sec:representation}, into $\text{P}^+$, the set of the local maxima of the original signal and $\text{P}^-$, the set of its local minima, we can apply the remaining steps of the algorithm in those two sets independently, arriving at a positive and a negative frontier.
+If we split P, defined in Section \ref{sec:representation}, into $\text{P}^+$, the set of the local maxima of the original signal and $\text{P}^-$, the set of its local minima, we can apply the remaining steps of the algorithm in those two sets independently, arriving at a superior and an inferior frontier, respectively.
 
@@
 \ref{fig:FullFrontiers} illustrates these frontiers for six diverse discrete sound waves, as well as an in-detail view of the highlighted segment for each signal.
-All waves are records of physical sounds, chosen to represent the applicability of the algorithm in real-world scenarios.
+All waves are records of physical sounds, chosen to represent the applicability of the algorithm in a real-world scenario.
 The implementation available at the PyPI repository provides an option to extract the frontiers.
@@
     \includegraphics{images/08FullFrontiers.pdf}
-  \caption{Positive and negative frontiers of six digital waves, as extracted by the algorithm here presented.
-For each wave, the region highlighted in black is shown in detail besides the whole wave.
-For each wave, the horizontal axis is the sample number $ i $, while the vertical axis is the normalized amplitude.
-All 6 waves were recorded at 44100 fps.}
+  \caption{Superior and inferior frontiers of six discrete signals, extracted by the proposed algorithm.
+For each wave, the region highlighted in black is shown in detail.
+The horizontal axis of each subplot represents the sample index $ i $, while the vertical axis represents the normalized amplitude.}
   \label{fig:FullFrontiers}
@@
 The frontiers were satisfactorily detected in the case of harmonic and inharmonic sounds, and are robust about the number of samples and the frequencies of the waves.
-It is worth noting that the positive and negative frontiers are generally smoother than the unified envelope and seems to conform better to the contours of the underlying discrete signal.
+It is worth noting that the positive and negative frontiers are generally smoother than the unified envelope presented in Figs.
+\ref{fig:spokenvoice}, \ref{fig:singer_voice} and \ref{fig:Bend}, and conforms better to the contours of the underlying discrete signal.
 
@@
 
-The temporal envelope adds complexity to the spectral representation of a wave \cite{2019TarjanoNeuro}.
-An accurate representation of the envelope, while describing the instantaneous amplitude of a signal, would also simplify further spectral analysis.
+In general, the temporal envelope adds complexity to the spectral representation of a wave \cite{2019TarjanoNeuro}.
+As demonstrated in Section \ref{sec:quality}, the carrier wave obtained with the proposed method is guaranteed to have a narrower frequency range than the original signal.
+
+Using the carrier wave in place of the original signal would simplify algorithms based on spectral analysis, without considerable drawbacks, since the envelope is fully known and can be applied back to the signal after eventual modifications.
 
@@
 Two prominent frequencies can be seen, corresponding to the initial and final notes of the bend.
-In the carrier, the other frequencies appear smoother.
+In the carrier, the other frequencies, introduced by the envelope, are less pronounced.
 
@@
     \includegraphics{images/12Fourier.pdf}
-  \caption{Fourier power spectrum for the wave and carrier of the sound of a guitar bend shown in Fig. \ref{fig:Bend}.}
+  \caption{Fourier power spectrum for wave and carrier of the sound of a guitar bend shown in Fig. \ref{fig:Bend}.}
   \label{fig:FourierBend}
@@
 
-The algorithm developed in this work naturally divides a signal into its pseudo cycles, pinpointing them in the time-domain, as illustrated in Fig.
-\ref{fig:singer_voice}, for the sound of an alto singer singing a sustained note, as in Fig.
+The algorithm developed in this work naturally divides a signal into its pseudo cycles, pinpointing them in the time-domain.
+Fig.
+\ref{fig:singer_voice} exemplifies this, for the sound of an alto singer singing a sustained note seen in Fig.
 \ref{fig:singer_voice}.
 
-The vertical lines mark the indices of the samples that belong to the positive frontier of the original signal, and the section between each vertical line isolates a pseudo cycle of the wave.
+The vertical lines mark the indices of the samples that belong to the positive frontier of the original signal.
+The section between each vertical line can be seen as a pseudo cycle of the wave.
 Although two local maxima exist in each pseudo cycle, only the largest was marked by the algorithm.
@@
 One application of that is illustrated in Fig.
-\ref{fig:AvgPc}: It shows the superposition of the extracted pseudo cycles, after being normalized by length, and their average.
-They were also rotated to make their starting point closer to zero.
+\ref{fig:AvgPc}: It shows the superposition of the extracted waveforms, after being normalized by length, and their average.
+They were horizontally shifted to make their starting point close to zero.
 

### 2021-05-11 Papers/01_Pulses_Envelope/00ElsevierPulsesEnvelope.tex
revising

@@
 
-\subsection{Approximated location of pseudo-cycles}
+\subsection{Approximated location of pseudo cycles}
 
@@
 \ref{fig:AvgPc}: It shows the superposition of the extracted waveforms, after being normalized by length, and their average.
-They were horizontally shifted to make their starting point close to zero.
 
 This same average is shown shifted, to exemplify that approximate $C^0$ and $C^1$ continuity is achieved.
-This can be seen as the average waveform of the original discrete signal, and can be used, along with the information of its envelope, to generate an alternative representation of the wave, useful, for example, for compression and classification purposes.
+This can be seen as the average waveform of the original discrete signal, and can be used, along with the description of the signal's envelope, to generate an alternative representation of the wave, useful, for example, for compression purposes.
 
@@
     \includegraphics{images/10AvgPc.pdf}
-  \caption{Pseudo-cycles of the wave illustrated in Fig.
-\ref{fig:singer_voice}, segmented and superposed, and their average.
-The average is also shown shifted by one average period, to illustrate the average continuity.}
+  \caption{Waveform of each pseudo cycles of the wave illustrated in Fig.
+\ref{fig:singer_voice}, interpolated to the same length and superposed.
+The average in its original position, and shifted, is also shown.}
   \label{fig:AvgPc}
@@
 \section{Conclusion}
-This work fills a gap identified in the theory of digital signal processing, where the lack of general procedures for the accurate identification of the temporal envelope of arbitrary waves poses an obstacle to the complete description and eventual manipulation of signals.
+
+This work fills a gap in the theory of digital signal processing, especially in the context of broadband signals, where the lack of general procedures for envelope detection is an obstacle to the complete description and manipulation of signals.
 
@@
 
-Illustrating the efficiency of an algorithm inspired by geometric aspects of a discrete signal on a wide range of real-world signals, we also hope to encourage research in this direction.
+Illustrating the efficiency of a geometric envelope detection algorithm on a wide range of real-world signals, we also hope to encourage research in this direction.
 
-The present work also highlights some gaps in our current understanding and definition of temporal envelope and, in so doing, endeavors to motivate further research in this topic, that can ultimately give rise to a more mathematically rigorous envelope theory, especially concerning broadband waves; suggestions in this direction were also presented.
+Showing that some important guarantees can be derived for envelope detection approaches based on selecting samples from the local extrema of the original signal, this works a direction to future research, that can give rise to a more mathematically rigorous envelope theory.
 
-While relevant in its own accord, the procedure here presented approximately isolates the individual pseudo-cycles of a wave, pinpointing them in the time-domain.
+The present work also highlights some gaps in our current understanding and definition of temporal envelope and, in so doing, endeavors to motivate further research in this topic.
 
-Similarly, by decomposing a wave into its envelope and carrier, and simplifying the representation of such carrier in the frequency-domain, a rich set of advancements, especially in sound compression and synthesis, both in the time and frequency-domains, can be built upon this initial algorithm.
+While relevant on its own, the procedure here presented approximately isolates the individual pseudo cycles of a wave, pinpointing them in the time-domain.
+This approach can serve as a basis for sound compression techniques.
 
-By illustrating how this information can be used to obtain the average waveform of a discrete wave, we hope to encourage research in this area, ultimately leading to alternative representations of discrete waves.
+Similarly, decomposing a wave into its envelope and carrier, thus simplifying the wave's spectral representation, can improve frequency-domain approaches to sound classification, compression, and synthesis.
+
+By illustrating how this algorithm can be used to segment a discrete signal into its pseudo cycles, and generate a description of the evolution of the signal's waveform over time, we hope to encourage research in this direction, potentially leading to alternative representations of discrete waves.
 

### 2021-05-11 Papers/01_Pulses_Envelope/images/10AvgPc.py

@@ -40,7 +40,7 @@ transp_black = "rgba(38, 12, 12, 0.2)"
 name = "alto"
 W, _ = se.read_wav(f"C:/Users/tesse/Desktop/Files/Dropbox/0_Science/reps/envelope/test_samples/{name}.wav")
 amp = np.max(np.abs(W))
-W = 4 * W / amp
+W = W / amp
 X = np.arange(W.size)
 
 Xpos, Xneg = se.get_frontiers(W, 0)
@@ -49,7 +49,7 @@ E = se.get_frontiers(W, 1)
 m = min(Xpos.size, Xneg.size)
 Xavg = np.round((Xpos[0 : m] + Xneg[0 : m]) / 2).astype(np.int)
 
-avg_pcs, orig_pcs, norm_pcs = average_pc_waveform(Xavg, W)
+avg_pcs, orig_pcs, norm_pcs = average_pc_waveform(Xpos.astype(np.int), W)
 
 '''============================================================================'''
 '''                                    PLOT                                    '''
@@ -81,37 +81,54 @@ fig.update_yaxes(showline=False, showgrid=False, zeroline=False)
 XX, YY = to_plot(norm_pcs)
 fig.add_trace(
   go.Scatter(
-    name="Normalized Pseudo Cycles    ",
+    name="Normalized waveforms  ",
     x=XX,
     y=YY,
     mode="lines",
     # fill="tozeroy",
     # fillcolor="rgba(0,0,0,0.3)",
-    line=dict(width=1, color="rgba(235, 64, 52,0.2)",),
+    line=dict(width=1, color=transp_black,),
   )
 )
 
 fig.add_trace(
   go.Scatter(
-    name="Average Waveform    ",
+    name="Split    ",
+    showlegend=False,
+    x=[avg_pcs.size - 1, avg_pcs.size - 1],
+    y=[-0.7, 1.1],
+    mode="lines",
+    # fill="tozeroy",
+    # fillcolor="rgba(0,0,0,0.3)",
+    line=dict(width=1, color="blue",),
+  )
+)
+
+fig.add_trace(
+  go.Scatter(
+    name="Average waveform  ",
     # x=np.arange(W.size),
     y=avg_pcs,
-    mode="lines",
+    mode="lines+markers",
+    line_shape='spline',
     # fill="tozeroy",
     # fillcolor="rgba(0,0,0,0.6)",
-    line=dict(width=3, color="black",),
+    marker=dict(size=3),
[…]

### 2021-05-12 Papers/01_Pulses_Envelope/00ElsevierPulsesEnvelope.tex
revising

@@
     \includegraphics{images/00AlphaShape.pdf}
-  \caption{Alpha shape of a set of points for $ \alpha=10 $. A circle of radius 10, containing two points of the set, but defining a disc that does not contain any point of the set, is also shown.}
+  \caption{Alpha shape of a set of points, for $ \alpha=10 $. A circle of radius 10, containing two points of the set, but defining a disc that does not contain any point of the set, is also shown.}
   \label{fig:AlphaShape}
@@
 Fig.
-\ref{fig:Envelope} provides an example of the vectors just defined: \textbf{w} is obtained by element-wise multiplication between the normalized carrier wave \textbf{c} and the envelope \textbf{e}.
+\ref{fig:Envelope} provides an example of the vectors just defined: \textbf{w} is obtained by element-wise multiplication between the carrier wave \textbf{c} and the envelope \textbf{e}.
 

### 2021-05-12 Papers/01_Pulses_Envelope/00ElsevierPulsesEnvelope.tex
revising

@@
 
-TODO
-% being, in fact, equal to the value of the instantaneous amplitude of the wave $ w_i $. That is, $ c_i = 1 $ at those points, rendering $ e_i = 1 w_i $.
+Consider the point $ w_i = c_i e_i $, with $ c_i $ as a local maximum of \textbf{c}, and assume $ c_i, e_i > 0 $.
+A point close to $ w_i $ can be written as $ w_i + \Delta w = (c_i - \Delta c) (e_i + \Delta e) $.
+We can assume $ \Delta e \approx 0 $, since the rate of change of the envelope is considerably smaller than that of the carrier.
+Making $ \Delta c > 0 $ guarantees the assumption that $ c_i $ is a local maximum.
 
-This realization that the envelope of a wave is defined only at some of the wave's local extrema is essential, as it can be used to generate a simplified representation of a signal, making the algorithm more efficient.
+Expanding this equation gives $ w_i + \Delta w = - \Delta{c} e_{i} + c_{i} e_{i} $, where $ c_i e_i = w_i $ and can be cancelled.
+Since $ \Delta c, e_i > 0 $, we conclude that $ \Delta w < 0 $ and $ w_i $ must  be a local maximum of \textbf{w}.
+
+This realization that the envelope is observable only at some of the wave's local extrema is important, as it can be used to generate a simplified representation of a signal, making the algorithm more efficient.
 
@@
 \ref{fig:Envelope}, the carrier, the original signal and even the envelope are representations of similar entities: signals in the same coordinate system.
-This is in line with \cite{2011CaetanoImproved,2011TurnerDemodulation,2017LyonsDigital}, but in contrast with some representations found in the literature of amplitude modulation of narrowband signals \cite[e.g.][]{1996Loughlinamplitude}, where the FM and AM of a signal are represented in different coordinate systems, commonly frequency versus time and amplitude versus time, respectively.
+This is in line with \cite{2011CaetanoImproved,2011TurnerDemodulation,2017LyonsDigital}, but in contrast with some representations found in the literature of amplitude modulation \cite[e.g.][]{1996Loughlinamplitude}, where the FM and AM of a signal are represented in different coordinate systems, commonly frequency versus time and amplitude versus time, respectively.
 

### 2021-05-12 Papers/01_Pulses_Envelope/Response to Reviewers.tex
response to reviewers

\documentclass{article}
\usepackage[margin=0.5in]{geometry}
\usepackage{amsmath}
\usepackage{amsfonts} % \mathbb
% \usepackage{amssymb}  % \therefore
\usepackage{algorithm}
\usepackage{algorithmic}
% \usepackage{lineno,hyperref}

% \bibliographystyle{elsarticle-num}
\begin{document}

\title{Response to Reviewers}

\maketitle

\section{Reviewer 2}
This is an interesting research. It seems that the authors have already put this article publicly accessible at arXiv.org. The authors are suggested to consider and address the following issues:

\subsection{Writing needs to be refined.}
We extensively reviewed the text, rewriting parts of it, besides restructuring the sections to improve flow and coherence.

\subsection{Issues with the Highlights: maximum 85 characters, including spaces, per bullet point}
Highlights were rewritten to conform with the requirement of less than 85 characters.

\subsection{Abstract looks too long. Should be presented in a more concise way.}
We revised the abstract, removing superfluous information and verbose phrasing in order to make it more succinct, resulting in a reduction from the original 349 words to 248 words in the revised version.

\subsection{Citation of references should be in order and no jump of reference numbers.}
This issue was caused by the set of commands used to compile the TeX source, and we are sorry for the overlook. It was fixed in the revised version.

\subsection{What's the Abscissa in the highlights (50k, 100k, etc.) and some figures?}
The abscissas refer to the index $i$ of the discrete signal's samples. We revised the legends of figures 3 to 6, 10 to 17, 21 and 22 to include this information, and also made this more explicit in the paragraph where the discrete version of the problem is introduced (page 5, line 109).

\subsection{Legends in some figures are confusing and hard to distinguish, such as in Fig 2: signal and Hilbert Envelope.}
We reworked all figures, colour coding different entities when deemed helpful, with careful attention to the impact in the readability of the legends.

In the case of Fig. 2 in the original manuscript, now Fig. 3 in the revised version (page 4, line 90), we changed the colour of the Hilbert envelope to red, and the line indicating the start of the noisy section to black, making both stand apart from the signal.

\subsection{Figure 3: the illustration of the carrier c and wave w does not comply with the common sense of envelope and modulation?}
We changed the legend of the figure (page 6, Fig. 5, line 116, top of the page) and Equation 1 (page 5, line 110) to make more explicit that the example was constructed from a known carrier and envelope.

Besides, we added a paragraph in the revised version (page 6, line 123) calling attention to the different representations found in the literature.

% TODO
\subsection{Definition of the envelope is very vague and controversial. What's the relationship with the conventional amplitude modulation?}
We made more explicit that a general definition of envelope, especially in the case of broadband signals, is still an open question in the literature, adding recent citations to corroborate this.(page 4, line 76; page 5, line 100)

In an effort to better define an envelope, we added a general definition for the envelope of a family of curves (page 13, line 311), that motivated a rewrite of the metric used to assess the general quality of envelopes. This metric is now based on the construction of a reference envelope from an approximation of this definition to a discrete acenario (page 13, line 310).

Moreover, we greatly extended the discussion on how the proposed method satisfies the 4 conditions necessary for the plausibility of the amplitude and frequency modulation of a signal, proposed by \cite{1996Loughlinamplitude}, reinforcing the relationship with conventional amplitude modulation (page 12, line 261).

\subsection{For the proposed "Equivalent Circle Approach", an acceptable mathematical background and justification is missing.}
We expanded the explanation for the Equivalent circle, justifying the assumption of the horizontal initial direction, and making other assumptions explicit (page 9, line 190).

\subsection{It's quite doubtful about the assumption of $ w = e \odot c $?}
We revised the explanation about this equation to make clear that, besides being assumed in \cite{2011TurnerDemodulation}, the relation holds by definition in our work (page 5, equation 1, line 108).

% \cite{2019Jiaempirical}

We also noted that the definition based on the analytic signal is common in classic AM-FM literature that deals with narrowband signals.

\subsection{P8: the authors claimed that "the algorithm here presented satisfies the four conditions presented", however, no convincing evidence is provided. Please prove it mathematically.}
We added mathematical proofs for each one of the four conditions.

\subsection{How was the benchmarking carried out? Details on the implementation should be provided.}
We explained the benchmarking more extensively, citing that the implementations used were from the signal processing module of the Scipy Python library, and also made the source code for the tests, as well as the used samples, available at the repository dedicated for the work.

\section{Reviewer 4}
This paper proposes a new method for envelope estimation by using the geometric properties of a discrete real signal. Using discrete curvatures to determine the envelope of a discrete signal is an interesting idea and results obtained by this proposed method seem to be acceptable. However, the reviewer cannot recommend the publication of the paper in its current form. The authors need to address the reviewers below and revise the paper accordingly:

\subsection{The description of the key concepts is not clear. The paragraphs of the current paper were poorly written and readers cannot follow them easily. The authors are suggested to revise the paper extensively to improve its readability.}
We revised the text extensively, with the aim of improving the overall readability and making the flow of ideas more linear. To that end, the overall structure of sections and subsections was substantially changed.

We also extended the definition of the basic concepts, adding Figures to better illustrate

\subsection{The authors describe basic concepts of convex and concave hulls in an intuitive manner and show that “the idea is to identify the local extrema that tough the envelope”. But the reviewer cannot fully understand the logical connection between them. The authors are suggested to explain their idea in a more explicit manner.}

\subsection{The reviewer cannot understand why filtering is needed in the pre-processing of Hilbert transform. Clarify the reasoning and necessity.}
We explained in the revised text that the results of the pure Hilbert transform retain great part of the frequency content of the underlying wave, specially in the case of broadband signals, and added a figure to exemplify this effect in the case of a real world signal of an alto singer uttering a sustained note. We moved the filtering to the post-processing, however, and results for the Hilbert transform were improved substantially.

\subsection{Envelopes obtained by the proposed method seem to be good, but those obtained by other comparable methods, such as Hilbert transform, are not reasonable. The reviewer believes that results obtained by Hilbert transform should not be bad as shown by the authors.}
Besides moving the filtering to the post-processing phase, we changed the cut-off frequency of the filter, that was fixed at 100 Hz, to 1/10 of the fundamental frequency of the original signal, improving the envelopes obtained by the Hilbert transform substantially.

\subsubsection{Besides, its end effects are serious. Is it caused by the filtering, which is mentioned by the reviewer in comment 3?}
The effects were caused by filtering the wave prior to the transform. We changed the order of the operations, eliminating the effects in the revised version.

\subsubsection{Provide an explanation for the results in Figures 10, 11, and 12.}

\subsection{5. There are a lot of mistakes in the texts, please proofread the paper carefully. Some of them listed in the following:}

\subsubsection{a. Full name at the first time for the abbreviate: being particularly illustrative of the potential synergy between geometric and DSP approaches.}
We inserted the full name and abbreviation for DSP in its first use in the first paragraph of the revised version, and did the same with other abbreviations.

\subsubsection{b. Figure or table or equation should be illustrated:}

\begin{itemize}
  \item From 3 and the discussion in the preceding chapter;
  \item The algorithm follows directly from the definition in 5 after noting;
  \item In 10 we illustrate the envelope extracted by the conventional algorithms;
  \item The times taken for the algorithms compared here to process each wave are shown in 2;
  \item Figure 14 shows the frequency-domain power spectrum for the wave and the carrier presented in 3.
\end{itemize}

This issue arose from a misconception that those terms would be added during the latex compilation, discovered late in the process of elaborating the manuscript, that caused many of the faulty references to remained unnoticed. We took steps to ensure proper indication of figures, tables and algorithms in the revised version.

\subsubsection{c. Wrong equation, should be $ r_k = v_{k,x} / \sin(\theta_k) $ : From figure 6 is easy to see that $ r_k = v_{k,x}  \sin(\theta_k) $}
We thank the reviewer for pointing out this error. This mistake was fixed in the revised version. It turned out to be a typo, in the sense that the following equation, derived in part from this one, was correct. We seized the opportunity to also double check all other equations in the revised version.

\subsection{6. Figure 7 is shown in the manuscript, but there is no explanation on the main body for it. Please add an explanation in the revised paper.}
We added an explanation to the figure 7 in the revised paper, that illustrates the envelope of a guitar bend, as suggested.

\section{Reviewer 5}

\subsection{Once the carrier frequency is known the technique presented in this paper is a special kind of limited curvature interpolation technique.}

We apologize if we misunderstood the commentary, but we think we addressed it when making clear that the wave in Fig. XX 
was constructed from a previously known carrier wave and envelope; that is the only instance when the carrier frequency, being a sinusoid, was known.

\subsection{The author should improve the citation by referencing corresponding techniques, such as K-curve, and bounded curvature interpolation methods. It is interesting to investigate the application of such technique to signal envelope extraction.}
The K-curves algorithm involves an iterative optimization problem that would add a considerable amount of time to the interpolation step, since the number of interpolation points is usually in the hundreds. Our main concern, in the interpolation step, is not so much about curvature, but rather about boundedness between the interpolated values.

As for the curvature estimation step, we tested interpolation and fitting methods during the development of an estimate of the curvature of a discrete curve. Some computational problems arose during piecewise polynomial fitting for example, due to the general high dimensionality of the problem, and some artifacts with B-splines approach motivated us to pursue alternative, local methods, that culminated in the development of the equivalent circle approach. We plan an extension to the algorithm in the future, where the curvature will be allowed to vary along the time, and for that k-curves seems ideal, because of its guarantee of maximum curvature at the control points.

Moreover, we indicated in the revised version that k-curves might be an interesting option, depending on the requisites for the extracted envelope.

\bibliographystyle{unsrt}
\bibliography{bibli}

\end{document}

### 2021-05-13 Papers/01_Pulses_Envelope/Response to Reviewers.tex
revised

@@
 \subsection{Writing needs to be refined.}
-We extensively reviewed the text, rewriting parts of it, besides restructuring the sections to improve flow and coherence.
+We extensively reviewed the text, rewriting parts of it, besides restructuring the structure of the sections to improve flow and coherence.
 
 \subsection{Issues with the Highlights: maximum 85 characters, including spaces, per bullet point}
-Highlights were rewritten to conform with the requirement of less than 85 characters.
+Highlights were rewritten to conform to the requirement of fewer than 85 characters.
 
 \subsection{Abstract looks too long. Should be presented in a more concise way.}
-We revised the abstract, removing superfluous information and verbose phrasing in order to make it more succinct, resulting in a reduction from the original 349 words to 248 words in the revised version.
+We revised the abstract, removing superfluous information and verbose phrasing to make it more succinct, resulting in a reduction from the original 349 words to 248 words in the revised version.
 
@@
 The abscissas refer to the index $i$ of the discrete signal's samples.
-We revised the legends of figures 3 to 6, 10 to 17, 21 and 22 to include this information, and also made this more explicit in the paragraph where the discrete version of the problem is introduced (page 5, line 109).
+We revised the highlights and the legends of Figs.
+3 to 6, 10 to 17, 21 and 22 to include this information, and also made this more explicit in the paragraph where the discrete version of the problem is introduced (page 5, line 109).
 
 \subsection{Legends in some figures are confusing and hard to distinguish, such as in Fig 2: signal and Hilbert Envelope.}
-We reworked all figures, colour coding different entities when deemed helpful, with careful attention to the impact in the readability of the legends.
-
+We reworked all figures, color coding different entities when deemed helpful, with careful attention to the impact on readability.  All legends were also revised to be clearer.
 In the case of Fig.
 2 in the original manuscript, now Fig.
-3 in the revised version (page 4, line 90), we changed the colour of the Hilbert envelope to red, and the line indicating the start of the noisy section to black, making both stand apart from the signal.
+3 in the revised version (page 4, line 90), we changed the color of the Hilbert envelope to red, and the line indicating the start of the noisy section to black, making both stand apart from the signal.
+The legend was completely reformulated in this case.
 
 \subsection{Figure 3: the illustration of the carrier c and wave w does not comply with the common sense of envelope and modulation?}
-We changed the legend of the figure (page 6, Fig.
-5, line 116, top of the page) and Equation 1 (page 5, line 110) to make more explicit that the example was constructed from a known carrier and envelope.
+We changed the legend of Fig. 5 (page 6, line 116, top of the page) and Equation 1 (page 5, line 110) to make it more explicit that the example was constructed from a known carrier and envelope.
 
@@
 
-% TODO
-\subsection{Definition of the envelope is very vague and controversial. What's the relationship with the conventional amplitude modulation?}
-We made more explicit that a general definition of envelope, especially in the case of broadband signals, is still an open question in the literature, adding recent citations to corroborate this.(page 4, line 76; page 5, line 100)
+\subsection{Definition of the envelope is very vague and controversial.}
+
+We made it more explicit that a general definition of envelope, especially in the case of broadband signals, is still an open question in the literature, adding recent citations to corroborate this.
+(page 4, line 76; page 5, line 100)
+
+A common intuitive definition is also given on page 4, line 76.
 
 In an effort to better define an envelope, we added a general definition for the envelope of a family of curves (page 13, line 311), that motivated a rewrite of the metric used to assess the general quality of envelopes.
-This metric is now based on the construction of a reference envelope from an approximation of this definition to a discrete acenario (page 13, line 310).
+This metric is now based on the construction of a reference envelope from an approximation of this definition to the discrete scenario (page 13, line 310).
 
-Moreover, we greatly extended the discussion on how the proposed method satisfies the 4 conditions necessary for the plausibility of the amplitude and frequency modulation of a signal, proposed by \cite{1996Loughlinamplitude}, reinforcing the relationship with conventional amplitude modulation (page 12, line 261).
+In the context of the work, we improved the definition adopted, as seen in Equation 1 (page 5, line 110). We also formalized the characteristics of a carrier wave on page 5, line 102.
+
+\subsection{What's the relationship with the conventional amplitude modulation?}
+We greatly extended the discussion on how the proposed method satisfies the 4 conditions necessary for the physical plausibility of the amplitude and frequency modulation of a signal, proposed by \cite{1996Loughlinamplitude} (page 11, line 264, Theoretical guarantees).
+
+As those conditions were postulated in the context of conventional amplitude modulation, this reinforces the relationship between the proposed algorithm and more traditional approaches.
 
 \subsection{For the proposed "Equivalent Circle Approach", an acceptable mathematical background and justification is missing.}
-We expanded the explanation for the Equivalent circle, justifying the assumption of the horizontal initial direction, and making other assumptions explicit (page 9, line 190).
+
+We expanded the explanation for the Equivalent circle, justifying the assumption of the horizontal initial direction, and making the other assumptions explicit (page 9, line 201).
+We further comment on some properties of this approach on page 9, line 215
 
 \subsection{It's quite doubtful about the assumption of $ w = e \odot c $?}
-We revised the explanation about this equation to make clear that, besides being assumed in \cite{2011TurnerDemodulation}, the relation holds by definition in our work (page 5, equation 1, line 108).
+We revised the explanation about this equation to make clear that, besides being assumed in \cite{2011TurnerDemodulation} (page 5, line 95), the relation holds by definition in our work (equation 1, page 5, line 110).
 
-% \cite{2019Jiaempirical}
+We also noted that the definition based on the analytic signal is common in classic AM-FM literature. (page 4, line 85)
 
-We also noted that the definition based on the analytic signal is common in classic AM-FM literature that deals with narrowband signals.
-
-\subsection{P8: the authors claimed that "the algorithm here presented satisfies the four conditions presented", however, no convincing evidence is provided. Please prove it mathematically.}
-We added mathematical proofs for each one of the four conditions.
+\subsection{P8: the authors claimed that “the algorithm here presented satisfies the four conditions presented”, however, no convincing evidence is provided. Please prove it mathematically.}
+We added mathematical proofs for each one of the four conditions. (page 11, line 264, Theoretical guarantees)
 
@@
 We explained the benchmarking more extensively, citing that the implementations used were from the signal processing module of the Scipy Python library, and also made the source code for the tests, as well as the used samples, available at the repository dedicated for the work.
+(page 15, line 350)
 
@@
 The authors are suggested to revise the paper extensively to improve its readability.}
-We revised the text extensively, with the aim of improving the overall readability and making the flow of ideas more linear.
-To that end, the overall structure of sections and subsections was substantially changed.
 
-We also extended the definition of the basic concepts, adding Figures to better illustrate
+We revised the text extensively, to improve the overall readability and make the flow of ideas more linear. To that end, the overall structure of sections and subsections was substantially changed.
+
+We also extended the definition of the basic concepts, adding Figures to better illustrate them.
+
+We inserted a more intuitive explanation of the alpha shapes algorithm in page 8, line 171, complemented by Fig. 1 (page 3, top of the page)
+
+The concept of convex hull received a similar treatment, with an additional explanation on page 2, line 25.
+
+Fig. 2, where both concepts are illustrated, was improved, and two alpha shapes, with different alphas, are shown.
 
@@
 
+The general overview of the paper at the end of the introduction was improved (page 4, line 55).
+Besides, the flow of ideas was made more linear by the reorganization of the sections and subsections; the main steps were divided into sections, generally beginning with an explanation of what is to be developed ahead.
+
+A section dedicated to how the signal is mapped to Cartesian coordinates(Section 4, page 6) and a section explaining how the envelope is identified (Section 6, page 10) were introduced.
+
 \subsection{The reviewer cannot understand why filtering is needed in the pre-processing of Hilbert transform. Clarify the reasoning and necessity.}
-We explained in the revised text that the results of the pure Hilbert transform retain great part of the frequency content of the underlying wave, specially in the case of broadband signals, and added a figure to exemplify this effect in the case of a real world signal of an alto singer uttering a sustained note.
+We explained in the revised text that the results of the pure Hilbert transform retain part of the frequency content of the underlying wave, especially in the case of broadband signals, and added Fig.
+4 (page 5, line 94) to exemplify this effect in the case of a real-world signal.
 We moved the filtering to the post-processing, however, and results for the Hilbert transform were improved substantially.
@@
 The reviewer believes that results obtained by Hilbert transform should not be bad as shown by the authors.}
-Besides moving the filtering to the post-processing phase, we changed the cut-off frequency of the filter, that was fixed at 100 Hz, to 1/10 of the fundamental frequency of the original signal, improving the envelopes obtained by the Hilbert transform substantially.
+Besides moving the filtering to the post-processing phase, we changed the cut-off frequency of the filter, which was fixed at 100 Hz, to 1/10 of the fundamental frequency of the original signal, improving the envelopes obtained by the Hilbert transform.
 
 \subsubsection{Besides, its end effects are serious. Is it caused by the filtering, which is mentioned by the reviewer in comment 3?}
-The effects were caused by filtering the wave prior to the transform. We changed the order of the operations, eliminating the effects in the revised version.
+The effects were worsened by filtering the wave before the transform.
+We changed the order of the operations, attenuating the effects in the revised version.
+Note that the effects are present even in unfiltered signals, such as the one in Fig.
+3 (page 4, bottom of the page), being a characteristic of the Hilbert transform.
 
 \subsubsection{Provide an explanation for the results in Figures 10, 11, and 12.}
+An explanation was added for each Figure:
+\begin{itemize}
+  \item Former Fig. 10, now Fig. 15 - page 14, line 335;
+  \item Former Fig. 11, now Fig. 16 - page 14, line 338;
+  \item Former Fig. 12, now Fig. 17 - page 15, line 340.
+\end{itemize}
 
@@
 \subsubsection{a. Full name at the first time for the abbreviate: being particularly illustrative of the potential synergy between geometric and DSP approaches.}
-We inserted the full name and abbreviation for DSP in its first use in the first paragraph of the revised version, and did the same with other abbreviations.
+We inserted the full name and abbreviation for DSP in its first use in the first paragraph of the revised version (page 2, line 2), and did the same with other abbreviations, such as the terms amplitude modulation (AM), first appearing on page 5, line 98; frequency modulation (FM), in page 5, line 100 and Python Package Index (PyPI), at page 4, line 70
 
@@
 
-This issue arose from a misconception that those terms would be added during the latex compilation, discovered late in the process of elaborating the manuscript, that caused many of the faulty references to remained unnoticed.
-We took steps to ensure proper indication of figures, tables and algorithms in the revised version.
+This issue arose from a misconception that those terms would be added during the latex compilation, discovered late in the process of elaborating the manuscript, which caused many of the faulty references to remain unnoticed.
+We took steps to ensure proper indication of figures, tables, equations, and algorithms in the revised version.
 
@@
 This mistake was fixed in the revised version.
-It turned out to be a typo, in the sense that the following equation, derived in part from this one, was correct.
-We seized the opportunity to also double check all other equations in the revised version.
+It turned out to be a typo, in the sense that the following equation (equation 4, page 10, top of the page), derived in part from this one, was correct.
+We seized the opportunity to double-check all equations in the revised version.
 
 \subsection{6. Figure 7 is shown in the manuscript, but there is no explanation on the main body for it. Please add an explanation in the revised paper.}
-We added an explanation to the figure 7 in the revised paper, that illustrates the envelope of a guitar bend, as suggested.
+We added an explanation to the former figure 7, now figure 12 (page 12) in the revised paper, on page 11, line 262.
 
@@
 
-We apologize if we misunderstood the commentary, but we think we addressed it when making clear that the wave in Fig. XX 
-was constructed from a previously known carrier wave and envelope; that is the only instance when the carrier frequency, being a sinusoid, was known.
+We apologize if we misunderstood the suggestion, but we think we addressed it when making clear that the wave in Fig.
+5, page 6, was constructed from a previously known carrier wave and envelope; that is the only instance when the carrier frequency, being a sinusoid, is known.
 
@@
 As for the curvature estimation step, we tested interpolation and fitting methods during the development of an estimate of the curvature of a discrete curve.
-Some computational problems arose during piecewise polynomial fitting for example, due to the general high dimensionality of the problem, and some artifacts with B-splines approach motivated us to pursue alternative, local methods, that culminated in the development of the equivalent circle approach.
-We plan an extension to the algorithm in the future, where the curvature will be allowed to vary along the time, and for that k-curves seems ideal, because of its guarantee of maximum curvature at the control points.
+Some computational problems arose and motivated us to pursue local methods, culminating in the development of the equivalent circle approach.
 
-Moreover, we indicated in the revised version that k-curves might be an interesting option, depending on the requisites for the extracted envelope.
+We indicated in the revised version that k-curves might be an interesting option to use in the interpolation step, depending on the requisites for the extracted envelope.
 

### 2021-05-13 Papers/01_Pulses_Envelope/00ElsevierPulsesEnvelope.tex
revised

@@
 Due to its geometric nature, the method can optionally generate the superior and inferior envelopes of a given wave.
-The algorithm compares favorably with classic envelope detection techniques based on smoothing, filtering, and the Hilbert Transform, besides being physically plausible.
+The algorithm compares favorably with classic envelope detection techniques based on smoothing, filtering, and the Hilbert transform, besides being physically plausible.
 We provide visualizations of the envelope extracted for various real-world signals with diverse characteristics, such as voice, spoken and sang, and pitched and unpitched musical instruments, and discuss some approaches to objectively assess the quality of the obtained envelopes.
@@
 
-We next explore, in Section \ref{sec:curvature}, the concept of discrete curvature, used in Section \ref{sec:envelopedetection} to discriminate which samples of a discrete signal belong to the envelope.
+We next investigate, in Section \ref{sec:curvature}, the concept of discrete curvature, used in Section \ref{sec:envelopedetection} to discriminate which samples of a discrete signal belong to the envelope.
 
@@
 Fig.
-\ref{fig:AnalyticSignal} exemplifies the Hilbert envelope, for a pure sinusoid modulated by a polynomial, illustrating the difference in the smoothness of the envelope in the presence and absence of Gaussian white noise.
+\ref{fig:AnalyticSignal} exemplifies the Hilbert envelope, for a pure sinusoid modulated by a polynomial, illustrating the difference in the smoothness of the envelope in the presence and absence of Gaussian noise.
 It is readily noticeable from Fig.
@@
     \includegraphics{images/01AnalyticSignal.pdf}
-  \caption{Envelope of a pure sinusoid with a local frequency of 20 cycles, modulated by a polynomial of degree 3, obtained by the Hilbert Transform approach.
+  \caption{Envelope of a pure sinusoid with a local frequency of 20 cycles, modulated by a polynomial of degree 3, obtained by the Hilbert transform approach.
 In the first half of the samples, the sinusoid is free of noise.
-In the second half Gaussian noise with a standard deviation of $ 1/10 $ of the maximum amplitude of the wave was added to the signal.}
+In the second half, Gaussian noise was added to the signal.}
   \label{fig:AnalyticSignal}
@@
 
-The times taken for the algorithms to process each signal are shown in Table \ref{table:Time_comparison}. 
+The processing times of the algorithms, for each signal, are shown in Table \ref{table:Time_comparison}. 
 The Python source for the tests, as well as the samples used, are available at the repository dedicated to this work, in the file “Comparison of Envelope Algorithms.py”.

### 2021-08-25 Papers/01_Pulses_Envelope/Response to Reviewers.tex
incorporated suggestions review 2

--- 2021-05-12 Papers/01_Pulses_Envelope/Response to Reviewers.tex
@@
 
-\section{Reviewer 2}
-This is an interesting research. It seems that the authors have already put this article publicly accessible at arXiv.org. The authors are suggested to consider and address the following issues:
+\section{Reviewer #4}
+The comments have been well address by the authors. The manuscript has also been well revised based on comments.
 
-\subsection{Writing needs to be refined.}
-We extensively reviewed the text, rewriting parts of it, besides restructuring the sections to improve flow and coherence.
+\section{Reviewer #8} The revised manuscript was much better than the original one.
+According to previous comments, the writing style of this revised version was largely improved, and some key concepts were much clearer compared with the original one.
+Most of the previous comments were well addressed.
+This reviewer just has several minor comments.
 
-\subsection{Issues with the Highlights: maximum 85 characters, including spaces, per bullet point}
-Highlights were rewritten to conform with the requirement of less than 85 characters.
+\subsection{The term 'benchmarking' always refers to the 'true value', or some 'gold standard' method that best estimates the true value.
+In the revised manuscript, section 7.2, the authors described two methods, which one was used as the 'benchmarking' for calculating the results listed in Table 1?}
 
-\subsection{Abstract looks too long. Should be presented in a more concise way.}
-We revised the abstract, removing superfluous information and verbose phrasing in order to make it more succinct, resulting in a reduction from the original 349 words to 248 words in the revised version.
+Only one method is described in section 7.2, the reference envelope.
+We changed the text to make this clear (line 320), and better defined this method, that was formalized in an equation (line 325).
+We also made clear that this is the method used as reference in Table 1 (line 331).
 
-\subsection{Citation of references should be in order and no jump of reference numbers.}
-This issue was caused by the set of commands used to compile the TeX source, and we are sorry for the overlook. It was fixed in the revised version.
+\subsection{What's the main drawback of POST-processing(filtering) of Hilbert transform?
+Although the Table 1 has already shown the error of this method.
+The reviewer suggests the author give an intuitive explanation.
+For example, around Figure.4, can we say the POST- filtering of Hilbert transform mis-estimated the envelope in magnitude?}
 
-\subsection{What's the Abscissa in the highlights (50k, 100k, etc.) and some figures?}
-The abscissas refer to the index $i$ of the discrete signal's samples.
-We revised the legends of figures 3 to 6, 10 to 17, 21 and 22 to include this information, and also made this more explicit in the paragraph where the discrete version of the problem is introduced (page 5, line 109).
+We can affirm that post filtering the result of the Hilbert transform will generally underestimate the envelope.
+This can be better understood in the frequency domain: By low pass filtering in the time domain, we are essentially zeroing frequency bins in the frequency domain above a certain frequency threshold and lowering the total energy of the transform result.
 
-\subsection{Legends in some figures are confusing and hard to distinguish, such as in Fig 2: signal and Hilbert Envelope.}
-We reworked all figures, colour coding different entities when deemed helpful, with careful attention to the impact in the readability of the legends.
+This procedure isn't intrinsically detrimental, however, in the sense that if the original result of the transform is smooth, those frequencies would already be absent, and the corresponding bins would be zero from the beginning.
+Filtering wouldn't change the original output of the transform.
 
-In the case of Fig.
-2 in the original manuscript, now Fig.
-3 in the revised version (page 4, line 90), we changed the colour of the Hilbert envelope to red, and the line indicating the start of the noisy section to black, making both stand apart from the signal.
+We changed the text to make this more intuitive (line 92).
 
-\subsection{Figure 3: the illustration of the carrier c and wave w does not comply with the common sense of envelope and modulation?}
-We changed the legend of the figure (page 6, Fig.
-5, line 116, top of the page) and Equation 1 (page 5, line 110) to make more explicit that the example was constructed from a known carrier and envelope.
+\subsection{The Figure.15 should be moved before Figure.13, since this Figure was mentioned earlier (line 294) than Figure.13 and Figure.14.}
+Despite being mentioned earlier (line 301 of the revised manuscript), Fig.
+15 is better understood after the description of the envelope extraction algorithms it exemplifies, presented after Figs 13 and 14, in Section "7.3.
+Comparison with traditional algorithms".
+To avoid a major reorganization of the text, we therefore indicated in its first mention (line 301) that the figure is to be found in section "7.3.
+Comparison with traditional algorithms".
 
-Besides, we added a paragraph in the revised version (page 6, line 123) calling attention to the different representations found in the literature.
+\subsection{Did the proposed algorithm require the known carrier c as a priori knowledge? This algorithm may not need this information. However, the author should clarify that in the manuscript.}
 
-% TODO
-\subsection{Definition of the envelope is very vague and controversial. What's the relationship with the conventional amplitude modulation?}
-We made more explicit that a general definition of envelope, especially in the case of broadband signals, is still an open question in the literature, adding recent citations to corroborate this.(page 4, line 76; page 5, line 100)
+A priori knowledge of the carrier c is not required.
+We feel that the misunderstanding in this regard is introduced shortly after the definition of the vectors c, e and w in equation 1, in the subsequent paragraph.
+We therefore reworked this paragraph to make this more apparent (line 115).
 
-In an effort to better define an envelope, we added a general definition for the envelope of a family of curves (page 13, line 311), that motivated a rewrite of the metric used to assess the general quality of envelopes.
-This metric is now based on the construction of a reference envelope from an approximation of this definition to a discrete acenario (page 13, line 310).
+\subsection{Line 216, did the author mean 'vk,y->0 => rk -> infinite'?}
 
-Moreover, we greatly extended the discussion on how the proposed method satisfies the 4 conditions necessary for the plausibility of the amplitude and frequency modulation of a signal, proposed by \cite{1996Loughlinamplitude}, reinforcing the relationship with conventional amplitude modulation (page 12, line 261).
+That was indeed what we intended, and we are sorry for the overlook. We corrected this error in line 223 of the revised version.
 
-\subsection{For the proposed "Equivalent Circle Approach", an acceptable mathematical background and justification is missing.}
-We expanded the explanation for the Equivalent circle, justifying the assumption of the horizontal initial direction, and making other assumptions explicit (page 9, line 190).
-
-\subsection{It's quite doubtful about the assumption of $ w = e \odot c $?}
-We revised the explanation about this equation to make clear that, besides being assumed in \cite{2011TurnerDemodulation}, the relation holds by definition in our work (page 5, equation 1, line 108).
-
-% \cite{2019Jiaempirical}
-
-We also noted that the definition based on the analytic signal is common in classic AM-FM literature that deals with narrowband signals.
-
-\subsection{P8: the authors claimed that "the algorithm here presented satisfies the four conditions presented", however, no convincing evidence is provided. Please prove it mathematically.}
-We added mathematical proofs for each one of the four conditions.
-
-\subsection{How was the benchmarking carried out? Details on the implementation should be provided.}
-We explained the benchmarking more extensively, citing that the implementations used were from the signal processing module of the Scipy Python library, and also made the source code for the tests, as well as the used samples, available at the repository dedicated for the work.
-
-\section{Reviewer 4}
-This paper proposes a new method for envelope estimation by using the geometric properties of a discrete real signal.
-Using discrete curvatures to determine the envelope of a discrete signal is an interesting idea and results obtained by this proposed method seem to be acceptable.
-However, the reviewer cannot recommend the publication of the paper in its current form.
-The authors need to address the reviewers below and revise the paper accordingly:
-
-\subsection{The description of the key concepts is not clear.
-The paragraphs of the current paper were poorly written and readers cannot follow them easily.
-The authors are suggested to revise the paper extensively to improve its readability.}
-We revised the text extensively, with the aim of improving the overall readability and making the flow of ideas more linear.
-To that end, the overall structure of sections and subsections was substantially changed.
-
-We also extended the definition of the basic concepts, adding Figures to better illustrate
-
-\subsection{The authors describe basic concepts of convex and concave hulls in an intuitive manner and show that “the idea is to identify the local extrema that tough the envelope”.
-But the reviewer cannot fully understand the logical connection between them.
-The authors are suggested to explain their idea in a more explicit manner.}
-
-\subsection{The reviewer cannot understand why filtering is needed in the pre-processing of Hilbert transform. Clarify the reasoning and necessity.}
-We explained in the revised text that the results of the pure Hilbert transform retain great part of the frequency content of the underlying wave, specially in the case of broadband signals, and added a figure to exemplify this effect in the case of a real world signal of an alto singer uttering a sustained note.
-We moved the filtering to the post-processing, however, and results for the Hilbert transform were improved substantially.
-
-\subsection{Envelopes obtained by the proposed method seem to be good, but those obtained by other comparable methods, such as Hilbert transform, are not reasonable.
-The reviewer believes that results obtained by Hilbert transform should not be bad as shown by the authors.}
-Besides moving the filtering to the post-processing phase, we changed the cut-off frequency of the filter, that was fixed at 100 Hz, to 1/10 of the fundamental frequency of the original signal, improving the envelopes obtained by the Hilbert transform substantially.
-
-\subsubsection{Besides, its end effects are serious. Is it caused by the filtering, which is mentioned by the reviewer in comment 3?}
-The effects were caused by filtering the wave prior to the transform. We changed the order of the operations, eliminating the effects in the revised version.
-
-\subsubsection{Provide an explanation for the results in Figures 10, 11, and 12.}
-
-\subsection{5. There are a lot of mistakes in the texts, please proofread the paper carefully. Some of them listed in the following:}
-
-\subsubsection{a. Full name at the first time for the abbreviate: being particularly illustrative of the potential synergy between geometric and DSP approaches.}
-We inserted the full name and abbreviation for DSP in its first use in the first paragraph of the revised version, and did the same with other abbreviations.
-
-\subsubsection{b. Figure or table or equation should be illustrated:}
-
-\begin{itemize}
-  \item From 3 and the discussion in the preceding chapter;
-  \item The algorithm follows directly from the definition in 5 after noting;
-  \item In 10 we illustrate the envelope extracted by the conventional algorithms;
-  \item The times taken for the algorithms compared here to process each wave are shown in 2;
-  \item Figure 14 shows the frequency-domain power spectrum for the wave and the carrier presented in 3.
-\end{itemize}
-
-This issue arose from a misconception that those terms would be added during the latex compilation, discovered late in the process of elaborating the manuscript, that caused many of the faulty references to remained unnoticed.
-We took steps to ensure proper indication of figures, tables and algorithms in the revised version.
-
-\subsubsection{c. Wrong equation, should be $ r_k = v_{k,x} / \sin(\theta_k) $ : From figure 6 is easy to see that $ r_k = v_{k,x}  \sin(\theta_k) $}
-We thank the reviewer for pointing out this error.
-This mistake was fixed in the revised version.
-It turned out to be a typo, in the sense that the following equation, derived in part from this one, was correct.
-We seized the opportunity to also double check all other equations in the revised version.
-
-\subsection{6. Figure 7 is shown in the manuscript, but there is no explanation on the main body for it. Please add an explanation in the revised paper.}
-We added an explanation to the figure 7 in the revised paper, that illustrates the envelope of a guitar bend, as suggested.
-
-\section{Reviewer 5}
-
-\subsection{Once the carrier frequency is known the technique presented in this paper is a special kind of limited curvature interpolation technique.}
-
-We apologize if we misunderstood the commentary, but we think we addressed it when making clear that the wave in Fig. XX 
-was constructed from a previously known carrier wave and envelope; that is the only instance when the carrier frequency, being a sinusoid, was known.
-
-\subsection{The author should improve the citation by referencing corresponding techniques, such as K-curve, and bounded curvature interpolation methods.
-It is interesting to investigate the application of such technique to signal envelope extraction.}
-The K-curves algorithm involves an iterative optimization problem that would add a considerable amount of time to the interpolation step, since the number of interpolation points is usually in the hundreds.
-Our main concern, in the interpolation step, is not so much about curvature, but rather about boundedness between the interpolated values.
-
-As for the curvature estimation step, we tested interpolation and fitting methods during the development of an estimate of the curvature of a discrete curve.
-Some computational problems arose during piecewise polynomial fitting for example, due to the general high dimensionality of the problem, and some artifacts with B-splines approach motivated us to pursue alternative, local methods, that culminated in the development of the equivalent circle approach.
-We plan an extension to the algorithm in the future, where the curvature will be allowed to vary along the time, and for that k-curves seems ideal, because of its guarantee of maximum curvature at the control points.
-
-Moreover, we indicated in the revised version that k-curves might be an interesting option, depending on the requisites for the extracted envelope.
+\section{Reviewer #9} The authors have put great effort to revise the manuscript and addressed all comments from other reviewers. I don't have other comments to add.
 

### 2021-08-25 Papers/01_Pulses_Envelope/00ElsevierPulsesEnvelope.tex
incorporated suggestions review 2

@@
 
-For real-world signals, such the one of an alto singer sustaining a steady note partially shown in Fig.
+%%%%%%%%%%%%%%%%%%%%% TODO
+As seen, the intuitive definition of an envelope describes it as a smooth curve; this smoothness requirement implies that the envelope should ideally be comprised of lower frequencies than the original signal.
+For real-world signals, such as the one of an alto singer sustaining a steady note, partially shown in Fig.
 \ref{fig:AnalyticSignalAlto}, the Hilbert transform alone provides a poor envelope, since it retains most of the frequencies of the underlying wave.
-With further processing, such as filtering, a smoother output can be obtained.
+
+Applying a low pass filter, we can coerce the result of the Hilbert transform to comply with the smoothness requirement, by removing high frequencies from the original result.
+This will potentially lead to undershoots, proportional to the initial high frequency content of the transform removed without compensation.
+%%%%%%%%%%%%%%%%%%%%%
 
@@
 
+%%%%%%%%%%%%%%%%%%%%% TODO
 The $ \odot $ operator in Equation \ref{eq:Envelope} stands for the Hadamard product and denotes element-wise multiplication of two vectors.
@@
 \ref{fig:Envelope} provides an example of the vectors just defined: \textbf{w} is obtained by element-wise multiplication between the carrier wave \textbf{c} and the envelope \textbf{e}.
+
+Equation \ref{eq:Envelope} connects the three vectors \textbf{e}, \textbf{w} and \textbf{c} in the sense that each one can be derived from the other two.
+In most real world scenarios, however, only the original signal \textbf{w} is known, and the envelope \textbf{e} is to be retrieved without a priori knowledge of the carrier wave \textbf{c}.
+%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-Fig. \ref{fig:Envelope} helps to illustrate a key concept: only at some local extrema the envelope can be inferred without prior knowledge of the underlying carrier.
+In this context, Fig. \ref{fig:Envelope} helps to illustrate a key concept: only at some local extrema the envelope can be inferred without prior knowledge of the underlying carrier.
 
@@
 The radius $ r_k $ of the circle equivalent to each $ v_k $ is defined in Equation \ref{eq:Radius}.
-This equation has the desirable property $ v_{k,y} \to 0 \implies v_{k,y} \to \infty $, that is, the radius tends to infinity when there is no vertical displacement, effectively turning the circle into a line.
+This equation has the desirable property $ v_{k,y} \to 0 \implies r_k \to \infty $, that is, the radius tends to infinity when there is no vertical displacement, effectively turning the circle into a line.
 
@@
 
-An example of such behaviour can be observed in Fig.
+An example of such behaviour can be observed in Section \ref{sec:comparison}, Fig.
 \ref{fig:Comparison_sinusoid}, where the envelope of a simple sinusoid is shown.
@@
 \subsection{Reference envelope}
-
 % In order to obtain a numeric indicator of the quality of an envelope, we construct a reference envelope, 
@@
 
-In practical applications, one seldom has access to a family of curves, instead being interested in obtaining the envelope of a single signal.
-Nevertheless, an approximation can be obtained, by shifting the original signal along its initial position in the horizontal axis by half of its period, and taking the absolute maximum value at each index $i$.
-Fig. \ref{fig:RefEnvelope} shows part of the reference envelope for the sound of a tom drum.
+In practical applications, one is generally interested in obtaining the envelope of a single known signal, seldom having access to a family of curves.
+By shifting this signal along its initial position, an approximation to a family of curves can be obtained, however, that can be used to construct a reference envelope.
+
+Given a discrete signal $ \mathbf{w} $, the reference envelope $ \mathbf{r} = (r_0, r_1, \cdots, r_{n-1}) $ is formalized in equation \ref{eq:referenceenvelope}, where $ f $ is the predominant frequency of the original signal, as obtained by the maximum argument of the DFT of the original signal, and $ t $ its predominant period.
+Moreover, the value of the original signal $ \mathbf{w} $ outside the interval $ 0 \le i < n $ is assumed to be zero.
+
+\begin{align} \label{eq:referenceenvelope}
+  \begin{split}
+    f &= \text{argmax} \big( \mathcal{F} (\mathbf{w}) \big) \\
+    t &= n / f \\
+    r_i &= \max(w_{i - t/2}, w_{i - t/2 + 1}, w_{i - t/2 + 2}, \cdots, w_{i + t/2})
+  \end{split}
+\end{align}
+
+% Given an arbitrary signal, one can use the FFT to obtain its predominant period. The reference envelope is obtained by gradually shifting the original signal from a lag of
+
+% Nevertheless, ,  in the horizontal axis by half of its period, and taking the absolute maximum value at each index $i$.
+% Fig. \ref{fig:RefEnvelope} shows part of the reference envelope for the sound of a tom drum.
 
@@
 
-Although lacking in applicability, both given its discontinuous shape and the high computational cost, an envelope constructed in this way arises from a mathematically sound background, and can serve as a reference to infer the quality of more practical methods.
+Although lacking in applicability, both given its discontinuous shape and the high computational cost of its estimation, an envelope constructed in this way arises from a mathematically sound background, and can serve as a reference to infer the quality of more practical methods.
+This envelope is used in the comparisons presented in Table \ref{table:Error_comparison}
 
@@
 
-Direct comparison with many recent algorithms is made difficult by the unavailability of digital implementations of such works, some of whice where designed to process analog signals \cite[e.g.][]{2018AssefModeling}.
+Direct comparison with many recent algorithms is made difficult by the unavailability of digital implementations of such works, some of which where designed to process analog signals \cite[e.g.][]{2018AssefModeling}.
 

### 2021-08-25 Papers/01_Pulses_Envelope/Response to Reviewers.tex
second revision submitted

@@
 \usepackage{amsfonts} % \mathbb
-% \usepackage{amssymb}  % \therefore
+\usepackage[utf8]{inputenc}
+\usepackage [autostyle, english = american]{csquotes}
+\MakeOuterQuote{"}
 \usepackage{algorithm}
@@
 
-\section{Reviewer #4}
+\section{Reviewer 4}
 The comments have been well address by the authors. The manuscript has also been well revised based on comments.
 
-\section{Reviewer #8} The revised manuscript was much better than the original one.
+\section{Reviewer 8} The revised manuscript was much better than the original one.
 According to previous comments, the writing style of this revised version was largely improved, and some key concepts were much clearer compared with the original one.
@@
 Only one method is described in section 7.2, the reference envelope.
-We changed the text to make this clear (line 320), and better defined this method, that was formalized in an equation (line 325).
+We changed the text to make this clear (line 320), and better defined this method, that was formalized with the addition of equation 9 (line 325).
 We also made clear that this is the method used as reference in Table 1 (line 331).
@@
 
-This procedure isn't intrinsically detrimental, however, in the sense that if the original result of the transform is smooth, those frequencies would already be absent, and the corresponding bins would be zero from the beginning.
+This procedure isn't intrinsically detrimental to the Hilbert envelope, however, in the sense that if the original result of the transform is smooth, those frequencies would already be absent, and the corresponding bins would be zero from the beginning.
 Filtering wouldn't change the original output of the transform.
 
-We changed the text to make this more intuitive (line 92).
+We changed the text to add a more intuitive explanation, and mentioned this potential effect of low pass filtering (line 92).
 
@@
 Despite being mentioned earlier (line 301 of the revised manuscript), Fig.
-15 is better understood after the description of the envelope extraction algorithms it exemplifies, presented after Figs 13 and 14, in Section "7.3.
-Comparison with traditional algorithms".
-To avoid a major reorganization of the text, we therefore indicated in its first mention (line 301) that the figure is to be found in section "7.3.
-Comparison with traditional algorithms".
+15 is better understood after the description of the envelope extraction algorithms it exemplifies, presented after Figs 13 and 14, in Section 7.3.
+"Comparison with traditional algorithms".
+To avoid a major reorganization of the text, we therefore indicated in its first mention (line 301) that the figure is to be found in section 7.3.
+"Comparison with traditional algorithms".
 
@@
 
-\subsection{Line 216, did the author mean 'vk,y->0 => rk -> infinite'?}
+\subsection{Line 216, did the author mean 'vk,y$\to$0 $\implies$ rk $\to$ infinite'?}
 
@@
 
-\section{Reviewer #9} The authors have put great effort to revise the manuscript and addressed all comments from other reviewers. I don't have other comments to add.
+\section{Reviewer 9} The authors have put great effort to revise the manuscript and addressed all comments from other reviewers. I don't have other comments to add.
 
-\bibliographystyle{unsrt}
-\bibliography{bibli}
+% \bibliographystyle{unsrt}
+% \bibliography{bibli}
 

### 2021-08-25 Papers/01_Pulses_Envelope/00ElsevierPulsesEnvelope.tex
second revision submitted

@@
 Applying a low pass filter, we can coerce the result of the Hilbert transform to comply with the smoothness requirement, by removing high frequencies from the original result.
-This will potentially lead to undershoots, proportional to the initial high frequency content of the transform removed without compensation.
+This will potentially lead to undershoots, proportional to the initial high-frequency content of the transform removed without compensation.
 %%%%%%%%%%%%%%%%%%%%%
@@
 
-%%%%%%%%%%%%%%%%%%%%% TODO
 The $ \odot $ operator in Equation \ref{eq:Envelope} stands for the Hadamard product and denotes element-wise multiplication of two vectors.
@@
 
+%%%%%%%%%%%%%%%%%%%%% TODO
 Equation \ref{eq:Envelope} connects the three vectors \textbf{e}, \textbf{w} and \textbf{c} in the sense that each one can be derived from the other two.
@@
 \end{figure}
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%% TODO
 In practical applications, one is generally interested in obtaining the envelope of a single known signal, seldom having access to a family of curves.
-By shifting this signal along its initial position, an approximation to a family of curves can be obtained, however, that can be used to construct a reference envelope.
+By shifting this signal around its initial position, an approximation to a family of curves can be obtained, however, that can be used to construct a reference envelope.
 
@@
 This envelope is used in the comparisons presented in Table \ref{table:Error_comparison}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
